### 2018-07-25 sections/content.tex
Initial commit of template files

%--------------------------------------------------------%
%	ABSTRACT
%--------------------------------------------------------%

% NOTE: AJS Wants about 100 words.

\begin{abstract}
\noindent Suspendisse potenti. Suspendisse quis sem elit, et mattis nisl. Phasellus consequat erat eu velit rhoncus non pharetra neque auctor. Phasellus eu lacus quam. Ut ipsum dolor, euismod aliquam congue sed, lobortis et orci. Mauris eget velit id arcu ultricies auctor in eget dolor. Pellentesque suscipit adipiscing sem, imperdiet laoreet dolor elementum ut. Mauris condimentum est sed velit lacinia placerat. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Nullam diam metus, pharetra vitae euismod sed, placerat ultrices eros. Aliquam tincidunt dapibus venenatis. In interdum tellus nec justo accumsan aliquam. Nulla sit amet massa augue.
\end{abstract}

% Insert keywords here
\setlength\parindent{.45in} \keywords{Sociology, Publication, Manuscript}

%--------------------------------------------------------%
%	BODY TEXT
%--------------------------------------------------------%

% Start double spacing here if you want
\doublespacing

% Main Text
\section*{Section One: Citations, Lists \& A Long Table}

Maecenas \cite{anderson1999} fermentum \cite{bourdieu1984} urna ac sapien tincidunt lobortis. Nunc feugiat faucibus varius. Ut sed purus nunc. Ut eget eros quis lectus mollis pharetra ut in tellus. Pellentesque ultricies velit sed orci pharetra et fermentum lacus imperdiet. Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos. Suspendisse commodo ultrices mauris, condimentum hendrerit lorem condimentum et. Pellentesque urna augue, semper et rutrum ac, consequat id quam. Proin lacinia aliquet justo, ut suscipit massa commodo sit amet. Proin vehicula nibh nec mauris tempor interdum. Donec orci ante, tempor a viverra vel, volutpat sed orci. \cite{bader2007}

Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. \cite{coleman1990} Pellentesque quis interdum velit. Nulla tincidunt sem quis nisi molestie nec hendrerit nulla interdum. Nunc at lectus at neque dapibus dapibus sit amet in massa. Nam ut nisl in diam consectetur dignissim. Sed lacinia diam id nunc suscipit vitae semper lorem semper. In vehicula velit at tortor fringilla elementum aliquam erat blandit. \cite{dimaggio1983} Donec pretium libero et neque vehicula blandit. Curabitur consequat interdum sem at ultrices. Sed at tincidunt metus. Etiam vulputate, lacus eget fermentum posuere, ante mi dignissim augue, et ultrices felis tortor sed nisl. \cite{giddens1991}

Here are two short lists, one bullet-point, and the other numbered.

\begin{itemize}
\item Bullet point one
\item Bullet point two
\end{itemize}

\begin{enumerate}
\item Numbered list item one
\item Numbered list item two
\end{enumerate}

\vspace{.25in}
\begin{center}
{\LARGE \textit{TABLE 1 ABOUT HERE.}}
\end{center}

\subsection*{Subsection One: A Sample Small Table}

Quisque elit ipsum, porttitor et imperdiet in, facilisis ac diam. Nunc facilisis interdum felis eget tincidunt. \cite{glaser2009} In condimentum fermentum leo, non consequat leo imperdiet pharetra. Fusce ac massa ipsum, vel convallis diam. Quisque eget turpis felis. Curabitur posuere, risus eu placerat porttitor, magna metus mollis ipsum, eu volutpat nisl erat ac justo. Nullam semper, mi at iaculis viverra, nunc velit iaculis nunc, eu tempor ligula eros in nulla. Aenean dapibus eleifend convallis. \cite{goffman1959} Cras ut libero tellus. Integer mollis eros eget risus malesuada fringilla mattis leo facilisis. Etiam interdum turpis eget odio ultricies sed convallis magna accumsan. Morbi in leo a mauris sollicitudin molestie at non nisl. \cite{granovetter1992}

\vspace{.25in}
\begin{center}
{\LARGE \textit{TABLE 2 ABOUT HERE.}}
\end{center}

\subsection*{Subsection Two: A Sample Figure \& A Sample Equation}

Donec eget ligula venenatis est posuere eleifend in sit amet diam. Vestibulum sollicitudin mauris ac augue blandit ultricies. Nulla facilisi. Etiam ut turpis nunc. \cite{granovetter1973} Praesent leo orci, tincidunt vitae feugiat eu, feugiat a massa. Duis mauris ipsum, tempor vel condimentum nec, suscipit non mi. Fusce quis urna dictum felis posuere sagittis ac sit amet erat. In in ultrices lectus. Nulla vitae ipsum lectus, a gravida erat. Etiam quam nisl, blandit ut porta in, accumsan a nibh. \cite{hochschild1983} Phasellus sodales euismod dolor sit amet elementum. Phasellus varius placerat erat, nec gravida libero pellentesque id. Fusce nisi ante, euismod nec cursus at, suscipit a enim. Nulla facilisi.

\vspace{.25in}
\begin{center}
{\LARGE \textit{FIGURE 1 ABOUT HERE.}}
\end{center}

Integer risus dui, condimentum et gravida vitae, adipiscing et enim. Aliquam erat volutpat. \cite{lareau2003} Pellentesque diam sapien, egestas eget gravida ut, tempor eu nulla. Vestibulum mollis pretium lacus eget venenatis. Fusce gravida nisl quis est molestie eu luctus ipsum pretium. Maecenas non eros lorem, vel adipiscing odio. \cite{massey1993} Etiam dolor risus, mattis in pellentesque id, pellentesque eu nibh. Mauris nec ante at orci ultricies placerat ac non massa. Aenean imperdiet, ante eu sollicitudin vestibulum, dolor felis dapibus arcu, sit amet fermentum urna nibh sit amet mauris. \cite{mcpherson2001} Suspendisse adipiscing mollis dolor quis lobortis. \cite{meyer1977}

\begin{equation}
\label{eq:emc}
e = mc^2
\end{equation}

\section*{The Second Section: Highlights, Markup, \& Comments}

\hl{Etiam congue sollicitudin diam non porttitor.} Etiam turpis nulla, auctor a pretium non, luctus quis ipsum. Fusce pretium gravida libero non accumsan. \cite{putnam1995} \ul{Donec eget augue ut nulla placerat hendrerit ac ut mi. Phasellus euismod ornare mollis.} Proin tempus fringilla ultricies. Donec pretium feugiat libero quis convallis. Nam interdum ante sed magna congue eu semper tellus sagittis. Curabitur eu augue elit. \cite{raudenbush2004} \todo[size=\footnotesize]{Sample margin comment. Footnote size.}

\st{Aenean eleifend purus et massa consequat facilisis. Etiam volutpat placerat dignissim.} \cite{sampson1997} Ut nec nibh nulla. Aliquam erat volutpat. Nam at massa velit, eu malesuada augue. Maecenas sit amet nunc mauris. Maecenas eu ligula quis turpis molestie elementum nec at est. Sed adipiscing neque ac sapien viverra sit amet vestibulum arcu rhoncus. \todo[inline,color=blue!40]{Sample inline comment, with blue color. Good for bigger comments. }

Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas. Quisque lacus diam, congue vel laoreet id, iaculis eu sapien. \cite{swidler1986} In id risus ac leo pellentesque pellentesque et in dui. Aliquam erat volutpat. Nullam ac viverra augue. In tempus venenatis nibh quis semper. Maecenas ac nisl eu ligula dictum lobortis. Sed lacus ante, tempor eu dictum eu, accumsan in velit. Integer accumsan convallis porttitor. Maecenas pretium tincidunt metus sit amet gravida. Maecenas pretium blandit felis, ac interdum ante semper sed. \cite{west1987}

In auctor ultrices elit, vel feugiat ligula aliquam sed. Curabitur aliquam elit sed dui rhoncus consectetur. Cras elit ipsum, lobortis a tempor at, viverra vitae mi. Cras sed urna sed eros bibendum faucibus. Morbi vel leo orci, vel faucibus orci. Vivamus urna nisl, sodales vitae posuere in, tempus vel tellus. Donec magna est, luctus non commodo sit amet, placerat et enim. \cite{wilson2012}

### 2018-07-25 sections/coverpage.tex
Initial commit of template files

%--------------------------------------------------------%
%	COVER PAGE
%--------------------------------------------------------%

\begin{titlepage}

  \newcommand{\HRule}{\rule{\linewidth}{0.5mm}} % Defines a new command for the horizontal lines, change thickness here

  \center % Center everything on the page

%	HEADING SECTION

  %\textsc{\LARGE University Name}\\[1.5cm] % Name of your university/college
  \textsc{\Large Manuscript Submission}\\[0.5cm] % Major heading such as course name
  \textsc{\large The Journal of Blah Blah}\\[0.5cm] % Minor heading such as course title

%	TITLE SECTION

  \vspace{1.5 cm}
  \HRule \\[0.4cm]
  { \huge \bfseries Wordy Manuscript Title: Even Wordier Manuscript Subtitle}\\[0.4cm] % Title of your document
  \HRule \\[1.5cm]

%	AUTHOR SECTION

  {\em\Large\textbf Authors:}\\
  \vspace{.5 cm}
  First \textsc{Author}\\
  Department of XYZ,\\
  The University of ABC\\
  \vspace{.5 cm}
  Second \textsc{Author}\\
  Department of XYZ,\\
  The University of ABC\\
  \vspace{.5 cm}
  Third \textsc{Author}\\
  Department of ABC,\\
  XYZ University\\

%	DATE SECTION

  \vspace{1.5 cm}
  {\large Submitted: \today}\\[3cm] % Date, change the \today to a set date if you want to be precise

\vfill % Fill the rest of the page with whitespace

\end{titlepage}

%-----------------------------------------------------------------

\newpage

### 2018-07-25 sections/title_authors.tex
Initial commit of template files

%--------------------------------------------------------%
%	TITLE SECTION with authors
%--------------------------------------------------------%

% Article title
  \title{\vspace{-15mm}\fontsize{21pt}{10pt}\selectfont\textbf{Wordy Manuscript Title: \\Even Wordier Manuscript Subtitle\thanks{Draft manuscript. Please do not cite without the authors' permission.}}}

% Authors and Affiliations
  \author[1]{\large First Author}
  \author[1]{\large Second Author}
  \author[2]{\large Third Author}
  \affil[1]{\normalsize Department of XYZ, The University of ABC}
  \affil[2]{\normalsize Department of ABC, The University of XYZ}
  \renewcommand\Authands{ and }

% Today's date
	\date{Submitted: \usvardate\today}

### 2018-07-25 sections/coverpage.tex
Update coverpage to simpler style

@@
-%--------------------------------------------------------%
-%	COVER PAGE
-%--------------------------------------------------------%
+%!TEX root = ../main.tex
 
@@
   %\textsc{\LARGE University Name}\\[1.5cm] % Name of your university/college
-  \textsc{\Large Manuscript Submission}\\[0.5cm] % Major heading such as course name
-  \textsc{\large The Journal of Blah Blah}\\[0.5cm] % Minor heading such as course title
+  % \textsc{\Large Manuscript Submission}\\[0.5cm] % Major heading such as course name
+  % \textsc{\large The Journal of Blah Blah}\\[0.5cm] % Minor heading such as course title
 
@@
 
-  \vspace{1.5 cm}
-  \HRule \\[0.4cm]
-  { \huge \bfseries Wordy Manuscript Title: Even Wordier Manuscript Subtitle}\\[0.4cm] % Title of your document
-  \HRule \\[1.5cm]
+  \vspace{3.5 cm}
+  { \huge \bfseries Correlating Microbial Abundances for the Microbial Interaction Network Database}\\[0.4cm] % Title of your document
 
@@
 
-  {\em\Large\textbf Authors:}\\
-  \vspace{.5 cm}
-  First \textsc{Author}\\
+  \vspace{1.5 cm}
+  Dileep Kishore, Gabriel Birzu, Zhenjun Hu, Charles DeLisi, Kirill Korolev, Daniel Segre\\
+  \vspace{1cm}
   Department of XYZ,\\
   The University of ABC\\
-  \vspace{.5 cm}
-  Second \textsc{Author}\\
-  Department of XYZ,\\
-  The University of ABC\\
-  \vspace{.5 cm}
-  Third \textsc{Author}\\
   Department of ABC,\\
@@
 
-  \vspace{1.5 cm}
-  {\large Submitted: \today}\\[3cm] % Date, change the \today to a set date if you want to be precise
+  % \vspace{1.5 cm}
+  % {\large Submitted: \today}\\[3cm] % Date, change the \today to a set date if you want to be precise
 

### 2018-07-27 sections/abstract.tex
Add abstract

%!TEX root = ../main.tex

% QUESTION: How many words is this limited to?

\begin{abstract}
  {
    \noindent
    Microorganisms naturally occur in communities which can consist of tens of thousands of species involved in complex interactions with each other.
    16S ribosomal RNA (rRNA) gene profiling provides snapshots that reveal the abundances and phylogenies of microorganisms in microbial communities.
    These snapshots can then be used to unveil the complex interactions governing these microbial communities.
    However, the nature of 16S data-sets makes the data processing difficult and error prone.
    The development of appropriate and well-validated analysis methods that do not produce statistical artifacts is still ongoing and is receiving much attention.
    Here, we perform a thorough analysis of each step in this process to determine the best tools and parameter choices that generate the most accurate and robust networks.
    Ultimately, we create a sequence analysis pipeline to standardize the processing of high-throughput community 16S rRNA sequencing data into networks of microbial interactions derived from their co-occurrences.
    \hl{The development of a standard pipeline for 16S data analysis would enable accurate comparison of data from different data-sets.} \todo[size=\footnotesize]{TODO: Talk briefly about the main result}
    The pipeline is availabe as a Python package and a docker container.
  }
\end{abstract}

% Insert keywords here
\setlength\parindent{.45in} \keywords{Microbiome, 16S rRNA, Pipeline, Interaction, Network, Correlations, Qiime}

### 2018-07-27 sections/introduction.tex
Add introduction.tex

\section*{Introduction}

    Microorganisms naturally occur in communities which can consist of tens of thousands of species involved in complex interactions with millions of metabolic and signaling relationships.
    Each interaction within this complex ecological network can have a positive effect, a negative effect or no effect on the species involved.
    For example, bacteria that compete for the same nutrient will have a negative impact on each other when the media is deficient in that nutrient \cite{Ghoul2016}.
    There can also be cases of cross-feeding where two species produce metabolic products that are beneficial to each other.

    The microbial interaction networks are highly dynamic, and they constantly reorganize in a varying environment.
    Even under static conditions, these interactions are often difficult to predict due to their non-linear nature \cite{Konopka2015}.
    Compounding on this complexity is the fact that these interaction networks usually involve thousands of species.
    While the species diversity is now measurable through metagenomic sequencing and other technologies, the detailed characterization of inter-species interactions still escapes our understanding.
    Knowledge of the community composition at a particular instance in time would enable us to derive partial insights into the underlying dynamics.
    The study of microbial communities using high throughput genomic surveys (such as 16S rRNA gene profiling) has become routine in the resolution of microbial diversity, abundances and associations within microbial populations \cite{Jovel2016}.
    There have recently been many large-scale data collection efforts involving soil (Earth Microbiome Project) \cite{Thompson2017}, marine (Tara Oceans Project) \cite{Zhang2015} and human-associated microbiota (Human Microbiome Project) \cite{HumanMicrobiomeProjectConsortium2012}.
    With data being acquired at exponential rates, it becomes increasingly difficult and essential to be able to extract useful knowledge from them.
    The first goal of this study is to alleviate the data-to-knowledge bottleneck by creating a comprehensive database of microbial interactions comprising of several kinds of heterogeneous data from different data sources.
    The first stage of the project will involve populating the database with associations derived from experimental predictions, literature review and raw 16S sequence data.

    Due to the compositional nature of 16S data, standard data analysis tools are not very effective.
    The development of appropriate and well-validated analysis methods that do not produce statistical artifacts is still ongoing and is receiving much attention. There have been many approaches taken to try to solve this issue, but, there is no single method that works perfectly in all situations \cite{Golob2017,Weiss2016}.
    The methodology for the analysis of 16S rRNA sequence data is not very straightforward and there are a myriad of tools for each step of the process.
    This makes the analysis complicated and not very reproducible.
    Hence, we propose to build a standardized 16S data analysis pipeline that will ensure accuracy and uniformity across all the data stored in the database.
    There exist many platforms like MG-RAST \cite{Keegan2016}, Qiita \cite{qiita} and others that provide pipelines for 16S data analysis, but the main motivations behind the construction of our own pipeline are two-fold.
    Firstly, these tools do not analyze co-occurrence networks and the effect of upstream statistical methods on the inferred networks.
    Secondly, a comprehensive study of existing methods is needed in order to suggest optimal methods and parameters for users of the pipeline and database.
    The pipeline would also allow a user to process their sequence data using the same standards applied to the database and compare their results to the networks in the database.
    This would also ensure reproducible comparisons.

    We will further collect metagenomic data (whole genome shotgun sequence data), metatranscriptomic data (mRNA sequence data) and metabolomic data (metabolite profile from mass spectroscopy) and incorporate these into the database in the next phase of the project.
    This addition would enable us to obtain a mechanistic perspective of the dynamics in a microbial community.
    The heterogeneous multi-omic data-sets will open new avenues and exciting approaches for generating new knowledge.
    This data can be leveraged to provide insights into previously unknown microbial ecological interaction network dynamics upon integration into the database.
    We will also make use of \ac{comets} simulations to predict interactions in microbial communities and to design synthetic communities.

### 2018-07-27 sections/acronyms.tex
Add acronyms to project

%!TEX root = ../main.tex

\section*{Acronyms}

\vspace{1em}

\begin{acronym}[XXXXXXXX]
    \acro{ngs}[NGS]{next-generation sequencing}
    \acro{mind}[MIND]{Microbial Interaction Network Database}
    \acro{otu}[OTU]{Operational Taxanomic Unit}
    \acro{esv}[ESV]{Exact Sequence Variant}
    \acro{pcr}[PCR]{Polymerase Chain Reaction}
    \acro{gg}[GG]{Green Genes}
    \acro{ncbi}[NCBI]{National Center for Biotechnology Information}
    \acro{comets}[COMETS]{Computation of Microbial Ecosystems in Time and Space}
    \acro{dada2}[DADA2]{Divisive Amplicon Denoising Algorithm 2}
    \acro{qiime1}[QIIME1]{Quantitative Insights Into Microbial Ecology 1}
    \acro{qiime2}[QIIME2]{Quantitative Insights Into Microbial Ecology 2}
    \acro{sparcc}[SparCC]{Sparse Correlations for Compositional data}
    \acro{spieceasi}[SpiecEasi]{Sparse InversE Covariance estimation for Ecological Association and Statistical Inference}
    \acro{mldm}[mLDM]{metagenomic Lognormal-Dirichlet-Multinomial}
    \acro{daa}[DAA]{Direct Association Analysis}
    \acro{kegg}[KEGG]{Kyoto Encyclopedia of Genes and Genomes}
    \acro{fba}[FBA]{Flux Balance Analysis}
\end{acronym}

### 2018-07-27 sections/introduction.tex
Update introduction
1. Delete paragraph about metagenomics and metatranscriptomics
2. Add tex root
3. Fix indentation

@@
+%!TEX root = ../main.tex
+
 \section*{Introduction}
@@
   This would also ensure reproducible comparisons.
-
-    We will further collect metagenomic data (whole genome shotgun sequence data), metatranscriptomic data (mRNA sequence data) and metabolomic data (metabolite profile from mass spectroscopy) and incorporate these into the database in the next phase of the project.
-    This addition would enable us to obtain a mechanistic perspective of the dynamics in a microbial community.
-    The heterogeneous multi-omic data-sets will open new avenues and exciting approaches for generating new knowledge.
-    This data can be leveraged to provide insights into previously unknown microbial ecological interaction network dynamics upon integration into the database.
-    We will also make use of \ac{comets} simulations to predict interactions in microbial communities and to design synthetic communities.

### 2018-07-27 sections/methods.tex
Add methods.tex and pipeline figure

%!TEX root = ../main.tex

\section*{Methods}

  The flowchart describing the workflow of the \ac{mind} pipeline is as shown in Figure \ref{fig:mind_pipeline}.
  The pipeline seamlessly integrates many publicly available tools as well as custom Python modules and scripts to extract co-occurrence associations from 16S sequence data.
  The input to the pipeline by default is the raw 16S rRNA sequence data, but it can be configured to use processed sequences, \ac{otu} tables or other types of intermediate data as well.
  The final output of the pipeline is the inferred network of co-occurrence associations among the microbes present in the samples.
  In order to determine the best combination of tools for the pipeline, we examine the effects of each tool on the final co-occurrence network.
  We can then estimate the variance contributed by each step in the pipeline to the total variance present across the networks.
  We propose to use mock synthetic data simulated using an Illumina read-simulator \cite{Escalona2016} and a mock community data-set \cite{Bokulich2016} as standards for comparing the counts matrices and the assigned taxonomies.
  We will then use a voting method to obtain a co-occurrence network that has links predicted by multiple network inference methods.

  Some key features of the pipeline include modularity, flexibility and ease-of-use.
  The entire pipeline consists of modules which can be modified, reordered or replaced appropriately using a simple configuration script.
  The only thing needed to process data through the pipeline is the package itself and a configuration script that lists the steps to be performed, along with the parameters to be used (optional).
  The pipeline allows for automatic parallelization of all possible processes, both within and across samples.
  The entire pipeline has been containerized into a Docker \cite{Merkel1994} image for easy deployment and setup.
  The main components of the pipeline are detailed in the following section.

  \begin{figure}[h]
    \centering
    \includegraphics[width=0.7\textwidth]{mind_pipeline.pdf}
    \caption{
      \textbf{The workflow of the microbial co-occurrence analysis pipeline}.
        The processes can be grouped into four major components - denoising and clustering (A), taxonomy assignment (B), \ac{otu}/\ac{esv} processing (C), and network inference (D).
        Each of these incorporate several processes, each of which in turn have several alternate algorithms for the same task (as indicated by the text to the right of the blue boxes).
    }
    \label{fig:mind_pipeline}
  \end{figure}

  \subsubsection*{Denoising and Clustering}
    \vspace{-5mm}
    This module deals with processing the raw 16S sequence data into \ac{otu} or \ac{esv} count tables.
    It incorporates three processes: quality control, denoise/cluster and chimera checking.
    \begin{itemize}
    \item The quality control process handles the demultiplexing and quality control steps such as trimming adapters and trimming low-quality nucleotide stretches from the sequences.
    \item The denoise/cluster process handles the conversion of the demultiplexed, trimmed sequences into \ac{otu} or \ac{esv} count tables \footnote{Some methods perform clustering and taxonomy assignment in the same step}.
    \item The chimera checking process handles the removal of chimeric sequences created during the \ac{pcr} step.
    \end{itemize}
    The output of this module is a counts matrix that describes the number of reads of a particular \ac{otu} or \ac{esv} (rows of the matrix) present in each sample (columns of the matrix). The pipeline incorporates \ac{qiime1} \cite{Caporaso2010}, \ac{qiime2}, \ac{dada2} \cite{Callahan2016} and Deblur \cite{Amir2017} tools in this module.

  \subsubsection*{Taxonomy Assignment}
    \vspace{-5mm}
    This module deals with assigning taxonomies to either the representative sequences of the \ac{otu}s or directly to the \ac{esv}s.
    In order to assign taxonomies to a particular sequence we need two things: a query tool and a taxonomy database.
    The taxonomy database is a collection of 16S sequences of known organisms.
    The query tool allows one to compare a sequence of interest to all the sequences in the database to find the best match.
    The pipeline already incorporates \ac{gg} \cite{DeSantis2006}, SILVA \cite{Quast2012} and will soon include the \ac{ncbi} \cite{Sayers2009} as well.

  \subsubsection*{OTU and ESV Processing}
    \vspace{-5mm}
    This module deals with normalization, filtering and applying transformations on the \ac{otu} or \ac{esv} counts matrix.
    \begin{itemize}
    \item Rarefaction is a normalization technique used to overcome the bias that might arise due to variable sampling depth \footnote{The total number of reads obtained from a particular sample} in different samples. This is commonly performed either by sub-sampling or by normalization of the matrix to the lowest sampling depth.
    \item Filtering is usually performed to remove samples or features from the count matrix that are sparse.
    \item Transformations are performed in order to correct for and overcome the compositional bias that is inherent in a counts matrix.
    \end{itemize}

  \subsubsection*{Network Inference}
    \vspace{-5mm}
    This module deals with the inference of co-occurrence associations from the \ac{otu} or \ac{otu} counts matrix. These associations can be represented as a network.
    \begin{itemize}
    \item Associations can either be correlations or direct associations.
    \item Significance of these associations is determined by re-sampling, bootstrapping and calculating pvalues.
    \end{itemize}
    The pipeline incorporates \ac{sparcc} \cite{Friedman2012} and \ac{spieceasi} \cite{Kurtz2015} and will soon include \ac{mldm} \cite{Yang2017} and \ac{daa} \cite{Menon2018}.

### 2018-07-27 sections/methods.tex
Add data visualizer section to methods

@@
     The pipeline incorporates \ac{sparcc} \cite{Friedman2012} and \ac{spieceasi} \cite{Kurtz2015} and will soon include \ac{mldm} \cite{Yang2017} and \ac{daa} \cite{Menon2018}.
+
+  \subsubsection*{MIND Data Visualizer}
+
+    In addition to the web application and pipeline, we have also developed a tool that facilitates easy visualization and exploration of the data generated by the \ac{mind} pipeline.
+    This tool is a Python package and is built using dash \cite{dash}.
+    It aids in the analysis of the effects of different processing methods on the generated counts matrix or network.
+    It takes in as input the results of a pipeline run and allows one to visualize various properties of intermediate \ac{otu} or \ac{esv} tables and final association networks.
+    This tool makes it easy and straightforward to compare different methods and combinations of methods in the 16S data analysis workflow.
+
+    The visualization tool currently supports:
+    \begin{itemize}
+      \item Filtering and aggregating data from multiple sources (data-sets)
+      \item Comparing properties of the generated counts matrices and networks across different methods
+      \item Visualization of these properties
+    \end{itemize}

### 2018-07-27 sections/results.tex
Add results.tex and draft result figures

%!TEX root = ../main.tex

\section*{Results}

  \begin{figure}[h]
    \begin{center}
      \includegraphics[width=17.8cm]{main_figure.png}
      \caption{\textbf{Percentage of variance in the final co-occurrence network due to each processing step.} }
      \label{fig:variance}
    \end{center}
  \end{figure}

  \begin{figure}[h]
    \begin{center}
      \includegraphics[width=15.8cm]{otu_corr_full.pdf}
      \caption{\textbf{Correlation between the abundances of taxa in OTU tables generated using various clustering and denoising methods}. }
      \label{fig:otu_correlations}
    \end{center}
  \end{figure}

  \begin{figure}[h]
    \begin{center}
      \includegraphics[width=17.8cm]{tax_comp_full.pdf}
      \caption{\textbf{Taxonomy composition of the 20 most abundant genera predicted using different taxonomy references databases}. NOTE: I will replace the legend with the taxonomy tree}
      \label{fig:tax_comp}
    \end{center}
  \end{figure}

  \begin{figure}[h]
    \begin{center}
      \includegraphics[width=15.8cm]{tax_distance_deblur.pdf}
      \caption{\textbf{Average percentage of mismatches in taxonomy assignment at various taxonomy levels}. NOTE: I will include a boxplot of the weighted distance in the taxonomy tree}
      \label{fig:otu_correlations}
    \end{center}
  \end{figure}

  \begin{figure}[h]
    \begin{center}
      \includegraphics[width=17.8cm]{net_inference_comparison.pdf}
      \caption{\textbf{Variance of the co-occurrence networks for different processing pipelines}.  Node (\textbf{a})  and edge (\textbf{b}) overlap between all different networks generated from the data.}
      \label{fig:network_heatmap}
    \end{center}
  \end{figure}

  \begin{figure}[h]
    \begin{center}
      %\includegraphics[width=17.8cm]{../figures/network_comparison.pdf}
      \caption{\textbf{Core network captures known interactions in vaginal microbiota.} \textbf{a} Core network, obtained by taking the intersection between all the different pipelines. \textbf{b} Vaginal microbiota network as determined by XXX.}
      \label{fig:network_comparison}
    \end{center}
  \end{figure}

### 2018-07-27 sections/discussion.tex
Add root and section heading to discussion.tex

--- 2018-07-27 sections/results.tex
@@
 
-\section*{Results}
+\section*{Discussion}
 
-  \begin{figure}[h]
-    \begin{center}
-      \includegraphics[width=17.8cm]{main_figure.png}
-      \caption{\textbf{Percentage of variance in the final co-occurrence network due to each processing step.} }
-      \label{fig:variance}
-    \end{center}
-  \end{figure}
-
-  \begin{figure}[h]
-    \begin{center}
-      \includegraphics[width=15.8cm]{otu_corr_full.pdf}
-      \caption{\textbf{Correlation between the abundances of taxa in OTU tables generated using various clustering and denoising methods}. }
-      \label{fig:otu_correlations}
-    \end{center}
-  \end{figure}
-
-  \begin{figure}[h]
-    \begin{center}
-      \includegraphics[width=17.8cm]{tax_comp_full.pdf}
-      \caption{\textbf{Taxonomy composition of the 20 most abundant genera predicted using different taxonomy references databases}. NOTE: I will replace the legend with the taxonomy tree}
-      \label{fig:tax_comp}
-    \end{center}
-  \end{figure}
-
-  \begin{figure}[h]
-    \begin{center}
-      \includegraphics[width=15.8cm]{tax_distance_deblur.pdf}
-      \caption{\textbf{Average percentage of mismatches in taxonomy assignment at various taxonomy levels}. NOTE: I will include a boxplot of the weighted distance in the taxonomy tree}
-      \label{fig:otu_correlations}
-    \end{center}
-  \end{figure}
-
-  \begin{figure}[h]
-    \begin{center}
-      \includegraphics[width=17.8cm]{net_inference_comparison.pdf}
-      \caption{\textbf{Variance of the co-occurrence networks for different processing pipelines}.
-Node (\textbf{a})  and edge (\textbf{b}) overlap between all different networks generated from the data.}
-      \label{fig:network_heatmap}
-    \end{center}
-  \end{figure}
-
-  \begin{figure}[h]
-    \begin{center}
-      %\includegraphics[width=17.8cm]{../figures/network_comparison.pdf}
-      \caption{\textbf{Core network captures known interactions in vaginal microbiota.} \textbf{a} Core network, obtained by taking the intersection between all the different pipelines.
-\textbf{b} Vaginal microbiota network as determined by XXX.}
-      \label{fig:network_comparison}
-    \end{center}
-  \end{figure}
+  Discussion

### 2018-07-27 sections/acknowledgments.tex
Add acknowledgments

%!TEX root = ../main.tex

\section*{Acknowledgments}

  We are grateful for ...

### 2018-07-27 sections/coverpage.tex
Add affiliations to the coverpage

@@
   \vspace{1.5 cm}
-  Dileep Kishore, Gabriel Birzu, Zhenjun Hu, Charles DeLisi, Kirill Korolev, Daniel Segre\\
+  Dileep Kishore\textsuperscript{1},
+  Gabriel Birzu\textsuperscript{2},
+  Zhenjun Hu\textsuperscript{1},
+  Charles DeLisi\textsuperscript{1,3},
+  Kirill Korolev\textsuperscript{1,2},
+  Daniel Segre\textsuperscript{1,3,4}\\
   \vspace{1cm}
-  Department of XYZ,\\
-  The University of ABC\\
-  Department of ABC,\\
-  XYZ University\\
+  \textsuperscript{1}Bioinformatics Program, Boston University, Boston, Massachusetts, USA\\
+  \textsuperscript{2}Department of Physics, Boston University, Boston, Massachusetts, USA\\
+  \textsuperscript{3}Department of Biomedical Engineering, Boston University, Boston, Massachusetts, USA\\
+  \textsuperscript{4}Department of Biology, Boston University, Boston, Massachusetts, USA\\
 

### 2018-07-27 sections/results.tex
Figures are now confined to their own sections
Update caption for 'otu_correlations' and 'network_comparison' figures

@@
       \includegraphics[width=15.8cm]{otu_corr_full.pdf}
-      \caption{\textbf{Correlation between the abundances of taxa in OTU tables generated using various clustering and denoising methods}. }
+      \caption{
+        \textbf{Comparison of various denoising and clustering algorithms used in the pipeline}.
+        (A, B) Correlation of the abundances of the taxa that are common between the count matrices created by two different methods.
+        (A) The best correlation (most similar methods) is between open-reference and denovo.
+        (B) The worst correlation (least similar methods) is between open-reference and dada2.
+        (C) A heatmap showing the $\mathrm{R}^2$ of all pairwise comparisons of the methods.
+      }
       \label{fig:otu_correlations}
@@
       \includegraphics[width=17.8cm]{net_inference_comparison.pdf}
-      \caption{\textbf{Variance of the co-occurrence networks for different processing pipelines}.
-Node (\textbf{a})  and edge (\textbf{b}) overlap between all different networks generated from the data.}
-      \label{fig:network_heatmap}
+      \caption{
+        \textbf{Networks generated using different network inference methods}.
+        The four different networks generated by different network inference methods are very dissimilar (C).
+        (A) The overlaps between the edges among the network5 generated is shown. The number of edges that are common to all networks are very low (5).
+        (B) The hamming distance between the networks is shown. The similarity between various methods was found to vary with the data-source used.
+      }
+      \label{fig:network_comparison}
     \end{center}

### 2018-07-27 sections/results.tex
Update figure labels
1. 'network_comparison' -> 'real_network'
2. 'tax_comp' -> 'tax_comparison'
Add some preliminary text to the results

@@
 \section*{Results}
+
+  There is significant variation both in the networks as well as the \ac{otu}/\ac{esv} tables when different combinations of methods are used to generate them.
+  Figure \ref{fig:otu_correlations} shows the comparison of several denoising and clustering methods.
+  The data shows that all the methods that perform either clustering or denoising are highly correlated with each other.
+  The correlation is low when comparing a method that performs clustering on the sequences with a method that performs denoising (Figure \ref{fig:otu_correlations}).
+  The common core network (network made up of edges that are predicted by all methods) is very small compared to the total number of edges predicted by all the methods (Figure \ref{fig:denoise_network}).
+  In this result, only the denoising method was changed (database used: \ac{gg} and network inference: \ac{sparcc}).
+  But, nonetheless this approach of identifying a common network will help identify links with a good degree of confidence.
+
+  SILVA and \ac{gg} are two popular 16S databases used for taxonomy identification.
+  The two databases vastly differ in terms of their last update status - \ac{gg} was last updated on May 2013 and SILVA was last updated on December 2017.
+  Since updates to taxonomic classifications are frequent, these two databases vary significantly \cite{Balvociute2017}.
+  As a consequence of this, when different taxonomy databases are used on the same data-set we see large numbers of misclassifications \ref{fig:tax_mismatches}.
+
+  Different network inference methods infer significantly different association networks.
+  We can divide the network inference methods into two sets, the first set of methods (Pearson, Spearman, \ac{sparcc}) infer pairwise correlations while the second set infer direct associations (\ac{spieceasi}, \ac{mldm}, \ac{daa}).
+  Our preliminary analysis (Figure \ref{fig:network_comparison}) shows that networks generated using 4 different methods (Pearson, Spearman, \ac{sparcc}, \ac{spieceasi}) are vastly different and the degree of similarity (number of edges that overlapped) between the methods was dependent on the data-set used.
+  The abundance profiles of the different data-sets used are show in Figure \ref{fig:abundance_profile}.
+
+  The pipeline and the data visualization tool help the user explore the effect of the various methods on their data.
+  Exploring these differences are difficult and inconvenient without the tool since different tools differ in their input and output formats and require inter-converting between the various formats.
+  In the case of our pipeline, this is handled automatically.
 
@@
       \caption{\textbf{Taxonomy composition of the 20 most abundant genera predicted using different taxonomy references databases}. NOTE: I will replace the legend with the taxonomy tree}
-      \label{fig:tax_comp}
+      \label{fig:tax_comparison}
     \end{center}
@@
       \caption{\textbf{Average percentage of mismatches in taxonomy assignment at various taxonomy levels}. NOTE: I will include a boxplot of the weighted distance in the taxonomy tree}
-      \label{fig:otu_correlations}
+      \label{fig:tax_mismatches}
     \end{center}
@@
 \textbf{b} Vaginal microbiota network as determined by XXX.}
-      \label{fig:network_comparison}
+      \label{fig:real_network}
     \end{center}

### 2018-07-31 sections/abstract.tex
Remove indent for keywords
Fixed typo

@@
     \hl{The development of a standard pipeline for 16S data analysis would enable accurate comparison of data from different data-sets.} \todo[size=\footnotesize]{TODO: Talk briefly about the main result}
-    The pipeline is availabe as a Python package and a docker container.
+    The pipeline is available as a Python package and a docker container.
   }
@@
 % Insert keywords here
-\setlength\parindent{.45in} \keywords{Microbiome, 16S rRNA, Pipeline, Interaction, Network, Correlations, Qiime}
+\keywords{Microbiome, 16S rRNA, Pipeline, Interaction, Network, Correlations, Qiime}

### 2018-07-31 sections/abstract.tex
Update abstract - minor word deletions and substitutions

@@
     Microorganisms naturally occur in communities which can consist of tens of thousands of species involved in complex interactions with each other.
-    16S ribosomal RNA (rRNA) gene profiling provides snapshots that reveal the abundances and phylogenies of microorganisms in microbial communities.
+    16S ribosomal RNA (rRNA) gene profiling provides snapshots that reveal the abundances and phylogenies of microorganisms.
     These snapshots can then be used to unveil the complex interactions governing these microbial communities.
-    However, the nature of 16S data-sets makes the data processing difficult and error prone.
+    However, the nature of 16S data-sets makes the data-analysis process, difficult and error-prone.
     The development of appropriate and well-validated analysis methods that do not produce statistical artifacts is still ongoing and is receiving much attention.
-    Here, we perform a thorough analysis of each step in this process to determine the best tools and parameter choices that generate the most accurate and robust networks.
-    Ultimately, we create a sequence analysis pipeline to standardize the processing of high-throughput community 16S rRNA sequencing data into networks of microbial interactions derived from their co-occurrences.
+    In this study, we perform a thorough analysis of each step in the 16S data-analysis process to determine the best tools and parameter choices that generate the most accurate and robust networks.
+    Ultimately, we create a 16S sequence analysis pipeline to standardize the processing of high-throughput community 16S rRNA sequencing data into networks of microbial interactions derived from their co-occurrences.
     \hl{The development of a standard pipeline for 16S data analysis would enable accurate comparison of data from different data-sets.} \todo[size=\footnotesize]{TODO: Talk briefly about the main result}

### 2018-07-31 sections/introduction.tex
Split sentence into two lines
Update introduction
1. Added a better start
2. Split it into better logical sections
3. Added todos

@@
 
-  Microorganisms naturally occur in communities which can consist of tens of thousands of species involved in complex interactions with millions of metabolic and signaling relationships.
+  Microbiomes are ubiquitous and play an important role in biogeochemical transformations in both natural as well as man-made ecosystems.
+  These microbiomes comprise several thousand different microbes interacting with each other as well as their environment often through intricate metabolic and signaling relationships.
+
   Each interaction within this complex ecological network can have a positive effect, a negative effect or no effect on the species involved.
+\todo[size=\footnotesize]{Combine with an explanation of co-occurrences}
   For example, bacteria that compete for the same nutrient will have a negative impact on each other when the media is deficient in that nutrient \cite{Ghoul2016}.
   There can also be cases of cross-feeding where two species produce metabolic products that are beneficial to each other.
+
+  High-throughput metagenomic sequencing techniques help detect, identify and quantify a large part of the constitutive microorganisms of a microbiome.
+  Even though the species diversity is now measurable through metagenomics, the detailed characterization of inter-species interactions still escapes our understanding.
 
@@
   Compounding on this complexity is the fact that these interaction networks usually involve thousands of species.
-  While the species diversity is now measurable through metagenomic sequencing and other technologies, the detailed characterization of inter-species interactions still escapes our understanding.
+
   Knowledge of the community composition at a particular instance in time would enable us to derive partial insights into the underlying dynamics.
   The study of microbial communities using high throughput genomic surveys (such as 16S rRNA gene profiling) has become routine in the resolution of microbial diversity, abundances and associations within microbial populations \cite{Jovel2016}.
-  There have recently been many large-scale data collection efforts involving soil (Earth Microbiome Project) \cite{Thompson2017}, marine (Tara Oceans Project) \cite{Zhang2015} and human-associated microbiota (Human Microbiome Project) \cite{HumanMicrobiomeProjectConsortium2012}.
+
+  Advances in next-generation sequencing (NGS) technologies had led to the generation of a wealth of data about these microbial populations.
+  There have recently been many large-scale data collection efforts involving soil (Earth Microbiome Project) \todo[size=\footnotesize]{Don't need to mention the names, just cite} \cite{Thompson2017}, marine (Tara Oceans Project) \cite{Zhang2015} and human-associated microbiota (Human Microbiome Project) \cite{HumanMicrobiomeProjectConsortium2012}.
   With data being acquired at exponential rates, it becomes increasingly difficult and essential to be able to extract useful knowledge from them.
-  The first goal of this study is to alleviate the data-to-knowledge bottleneck by creating a comprehensive database of microbial interactions comprising of several kinds of heterogeneous data from different data sources.
-  The first stage of the project will involve populating the database with associations derived from experimental predictions, literature review and raw 16S sequence data.
 

### 2018-07-31 sections/coverpage.tex
Add grave to Daniel's name

@@
   Kirill Korolev\textsuperscript{1,2},
-  Daniel Segre\textsuperscript{1,3,4}\\
+  Daniel Segr\`{e}\textsuperscript{1,3,4}\\
   \vspace{1cm}

### 2018-07-31 sections/methods.tex
Fix indentation of items in methods
Change section heading to "Materials and Methods"
Add datasets subsection and reformat methods

@@
 
-\section*{Methods}
+\section*{Materials and Methods}
+
+  \subsection*{Datasets}
+
+    \vspace{-5mm}
+
+    The mock community 16S datasets was obtained from \cite{Bokulich2016}.
+
+    The real datasets were control samples (49samples) \todo[size=\footnotesize]{Check the exact number} obtained from \cite{Kang2017}.
+
+    Simulated data was generated using Grinder \cite{Angly2012}.
+
+  \subsection*{16S processing pipeline}
+
+    \vspace{-5mm}
 
     The flowchart describing the workflow of the \ac{mind} pipeline is as shown in Figure \ref{fig:mind_pipeline}.
-  The pipeline seamlessly integrates many publicly available tools as well as custom Python modules and scripts to extract co-occurrence associations from 16S sequence data.
-  The input to the pipeline by default is the raw 16S rRNA sequence data, but it can be configured to use processed sequences, \ac{otu} tables or other types of intermediate data as well.
+    The pipeline integrates many publicly available tools as well as custom Python modules and scripts to extract co-occurrence associations from 16S sequence data.
+    The input to the pipeline by default is the raw community 16S rRNA sequence data, but it can be configured to use processed sequences, \ac{otu} tables or other types of intermediate data as well.
     The final output of the pipeline is the inferred network of co-occurrence associations among the microbes present in the samples.
@@
 
+    % TODO: Talk about what is happening in the pipeline NOT what that steps does in theory
     \subsubsection*{Denoising and Clustering}
@@
       \begin{itemize}
-    \item Rarefaction is a normalization technique used to overcome the bias that might arise due to variable sampling depth \footnote{The total number of reads obtained from a particular sample} in different samples.
+        \item Rarefaction is a normalization technique used to overcome the bias that might arise due to variable sampling depth in different samples.
 This is commonly performed either by sub-sampling or by normalization of the matrix to the lowest sampling depth.
@@
 
-  \subsubsection*{MIND Data Visualizer}
+  \subsection*{Statistical Analysis}
+
+    Tests
+
+  \subsection*{MIND Data Visualizer}
 

### 2018-08-06 sections/introduction.tex
Update Introduction

@@
   Microbiomes are ubiquitous and play an important role in biogeochemical transformations in both natural as well as man-made ecosystems.
-  These microbiomes comprise several thousand different microbes interacting with each other as well as their environment often through intricate metabolic and signaling relationships.
+  These microbiomes comprise several thousand different microbes interacting with each other and their environment, often through intricate metabolic and signaling relationships.
+  The changes in the microbiome can impact the health of the environmental niche in which they reside \cite{HumanMicrobiomeProjectConsortium2012,Lloyd-Price2016} and this can have major implications in health of the human body.
+  A mechanistic understanding of the key interactions driving the microbe-microbe and microbe-environment dynamics is necessary in order to analyze the dysbiosis in a healthy microbiome and in the design of modifications to these correct these microbiomes.
 
+  Studies have shown the importance of specific microbial interactions in the healthy human microbiome \cite{Lloyd-Price2016} and others have shown how changes in these interactions can lead to dysbiosis \cite{Wang2017,Gilbert2016,Belizario2015}.
+  Co-occurrence relationships in microbial communities are be driven by microbial interactions.
   Each interaction within this complex ecological network can have a positive effect, a negative effect or no effect on the species involved.
-\todo[size=\footnotesize]{Combine with an explanation of co-occurrences}
-  For example, bacteria that compete for the same nutrient will have a negative impact on each other when the media is deficient in that nutrient \cite{Ghoul2016}.
-  There can also be cases of cross-feeding where two species produce metabolic products that are beneficial to each other.
-
-  High-throughput metagenomic sequencing techniques help detect, identify and quantify a large part of the constitutive microorganisms of a microbiome.
-  Even though the species diversity is now measurable through metagenomics, the detailed characterization of inter-species interactions still escapes our understanding.
-
+  Bacteria that compete for the same nutrient will have a negative co-occurrence relationship \cite{Ghoul2016} whereas cross-feeding between species leads to a positive co-occurrence relationship \cite{DSouza2018}.
   The microbial interaction networks are highly dynamic, and they constantly reorganize in a varying environment.
@@
 
+  High-throughput metagenomic sequencing techniques (such as 16S rRNA gene profiling) help detect, identify and quantify a large part of the constitutive microorganisms of a microbiome \cite{Jovel2016}.
   Knowledge of the community composition at a particular instance in time would enable us to derive partial insights into the underlying dynamics.
-  The study of microbial communities using high throughput genomic surveys (such as 16S rRNA gene profiling) has become routine in the resolution of microbial diversity, abundances and associations within microbial populations \cite{Jovel2016}.
+  With the advancement in DNA sequencing technologies \cite{Narihiro2017} and data processing methods much more information can be extracted from these microbial community samples than ever before.
+  These advances have led to large-scale data collection efforts involving soil (Earth Microbiome Project) \cite{Thompson2017}, marine (Tara Oceans Project) \cite{Zhang2015} and human-associated microbiota (Human Microbiome Project) \cite{HumanMicrobiomeProjectConsortium2012} and thus led the generation of a wealth of data about these microbial communities.
+  These metagenomic data-sets can be used to infer networks of co-occurrence relationships between microorganisms in the samples.
+  The co-occurrence networks can in turn be used to identify important interactions driving the dynamics of the microbial community, as well as, analyze emergent properties such as robustness and modularity.
+  However, 16S sequencing data is limited by factors such as resolution, sequencing depth, compositional nature, sequencing errors and copy number variations.
+  To counter these limitations and remove statistical artifacts, many different analysis methods have been developed \cite{Callahan2016,Amir2017,Friedman2012,Kurtz2015}.
+  But, the different methods and tools developed to solve these issues infer vastly different community compositions and co-occurrence networks \cite{Golob2017,Weiss2016}.
 
-  Advances in next-generation sequencing (NGS) technologies had led to the generation of a wealth of data about these microbial populations.
-  There have recently been many large-scale data collection efforts involving soil (Earth Microbiome Project) \todo[size=\footnotesize]{Don't need to mention the names, just cite} \cite{Thompson2017}, marine (Tara Oceans Project) \cite{Zhang2015} and human-associated microbiota (Human Microbiome Project) \cite{HumanMicrobiomeProjectConsortium2012}.
-  With data being acquired at exponential rates, it becomes increasingly difficult and essential to be able to extract useful knowledge from them.
-
-  Due to the compositional nature of 16S data, standard data analysis tools are not very effective.
-  The development of appropriate and well-validated analysis methods that do not produce statistical artifacts is still ongoing and is receiving much attention.
-  There have been many approaches taken to try to solve this issue, but, there is no single method that works perfectly in all situations \cite{Golob2017,Weiss2016}.
-  The methodology for the analysis of 16S rRNA sequence data is not very straightforward and there are a myriad of tools for each step of the process.
-  This makes the analysis complicated and not very reproducible.
-  Hence, we propose to build a standardized 16S data analysis pipeline that will ensure accuracy and uniformity across all the data stored in the database.
-  There exist many platforms like MG-RAST \cite{Keegan2016}, Qiita \cite{qiita} and others that provide pipelines for 16S data analysis, but the main motivations behind the construction of our own pipeline are two-fold.
-  Firstly, these tools do not analyze co-occurrence networks and the effect of upstream statistical methods on the inferred networks.
-  Secondly, a comprehensive study of existing methods is needed in order to suggest optimal methods and parameters for users of the pipeline and database.
-  The pipeline would also allow a user to process their sequence data using the same standards applied to the database and compare their results to the networks in the database.
-  This would also ensure reproducible comparisons.
+  The methodology for the analysis of 16S rRNA sequence data is not very straightforward and there are a myriad of tools for each step of the process, making the analysis complicated and not very reproducible.
+  In this study, we propose to build a standardized 16S data-analysis pipeline that will produce robust and reproducible co-occurrence networks from community 16S sequence data.
+\todo[size=\footnotesize]{Mention DB?}
+  There exist many platforms like MG-RAST \cite{Keegan2016}, Qiita \cite{qiita} and tools such as QIIME \cite{Caporaso2010} that provide pipelines for 16S data analysis, but the main motivations behind the construction of our own pipeline are two-fold.
+  Firstly, the existing tools do not focus their analysis on co-occurrence networks and the effect of upstream statistical methods on the inferred networks.
+  Secondly, we undertake a comprehensive study of existing methods in order to suggest optimal methods, tools and parameters for users of the pipeline.
+  We believe that these steps will ensure better reproducibility and easier comparison of networks across data-sets.

### 2018-08-07 sections/methods.tex
Update Methods

@@
     \vspace{-5mm}
-
-    The mock community 16S datasets was obtained from \cite{Bokulich2016}.
-
-    The real datasets were control samples (49samples) \todo[size=\footnotesize]{Check the exact number} obtained from \cite{Kang2017}.
-
-    Simulated data was generated using Grinder \cite{Angly2012}.
+    \begin{itemize}
+      \item The mock community 16S datasets were obtained from mockrobiota \cite{Bokulich2016}. We used samples (\hl{enter id number}) for this study.
+      \item The real datasets were control samples (49samples) \todo[size=\footnotesize]{Check the exact number} of a fecal microbiome transplant study obtained from \cite{Kang2017}.
+      \item Simulated data was generated using Grinder \cite{Angly2012}.
+    \end{itemize}
 
@@
 
-    The flowchart describing the workflow of the \ac{mind} pipeline is as shown in Figure \ref{fig:mind_pipeline}.
-    The pipeline integrates many publicly available tools as well as custom Python modules and scripts to extract co-occurrence associations from 16S sequence data.
-    The input to the pipeline by default is the raw community 16S rRNA sequence data, but it can be configured to use processed sequences, \ac{otu} tables or other types of intermediate data as well.
-    The final output of the pipeline is the inferred network of co-occurrence associations among the microbes present in the samples.
-    In order to determine the best combination of tools for the pipeline, we examine the effects of each tool on the final co-occurrence network.
-    We can then estimate the variance contributed by each step in the pipeline to the total variance present across the networks.
-    We propose to use mock synthetic data simulated using an Illumina read-simulator \cite{Escalona2016} and a mock community data-set \cite{Bokulich2016} as standards for comparing the counts matrices and the assigned taxonomies.
-    We will then use a voting method to obtain a co-occurrence network that has links predicted by multiple network inference methods.
+    The flowchart describing the workflow of the 16S data-analysis pipeline is as shown in Figure \ref{fig:mind_pipeline}.
+    The pipeline (\hl{github link here}) integrates many publicly available tools as well as custom Python modules and scripts to extract co-occurrence associations from 16S sequence data.
+    The input to the pipeline by default is the raw community 16S rRNA sequence data, but it can additionally be configured to use processed sequences, \ac{otu} tables or other types of intermediate data.
+    The final output of the pipeline is the inferred network of co-occurrence relationships among the microbes present in the samples.
 
@@
     The entire pipeline consists of modules which can be modified, reordered or replaced appropriately using a simple configuration script.
-    The only thing needed to process data through the pipeline is the package itself and a configuration script that lists the steps to be performed, along with the parameters to be used (optional).
+    The configuration script lists the steps to be performed during runtime, along with the parameters to be used (optional) for the various steps.
+    Since the entire pipeline run-through is stored in the form of a text file (the configuration file), subsequent runs are highly reproducible and changes can be easily tracked using version control.
     The pipeline allows for automatic parallelization of all possible processes, both within and across samples.
-    The entire pipeline has been containerized into a Docker \cite{Merkel1994} image for easy deployment and setup.
-    The main components of the pipeline are detailed in the following section.
+    In addition to the Python package, the entire pipeline has been containerized into a Docker \cite{Merkel1994} image (\hl{dockerhub link}) for easy deployment and setup.
+    The main components of the pipeline are detailed in the following sections.
 
-    \begin{figure}[h]
-      \centering
-      \includegraphics[width=0.7\textwidth]{mind_pipeline.pdf}
-      \caption{
-        \textbf{The workflow of the microbial co-occurrence analysis pipeline}.
-          The processes can be grouped into four major components - denoising and clustering (A), taxonomy assignment (B), \ac{otu}/\ac{esv} processing (C), and network inference (D).
-          Each of these incorporate several processes, each of which in turn have several alternate algorithms for the same task (as indicated by the text to the right of the blue boxes).
-      }
-      \label{fig:mind_pipeline}
-    \end{figure}
-
-    % TODO: Talk about what is happening in the pipeline NOT what that steps does in theory
+    \todo[size=\footnotesize]{Should be in the supplementary section}
     \subsubsection*{Denoising and Clustering}
@@
       This module deals with processing the raw 16S sequence data into \ac{otu} or \ac{esv} count tables.
-      It incorporates three processes: quality control, denoise/cluster and chimera checking.
+      It consists of three processes:
       \begin{itemize}
         \item The quality control process handles the demultiplexing and quality control steps such as trimming adapters and trimming low-quality nucleotide stretches from the sequences.
-        \item The denoise/cluster process handles the conversion of the demultiplexed, trimmed sequences into \ac{otu} or \ac{esv} count tables \footnote{Some methods perform clustering and taxonomy assignment in the same step}.
+        \item The denoise/cluster process handles the conversion of the demultiplexed, trimmed sequences into \ac{otu} or \ac{esv} count tables (some methods perform clustering and taxonomy assignment in the same step).
         \item The chimera checking process handles the removal of chimeric sequences created during the \ac{pcr} step.
@@
       This module deals with assigning taxonomies to either the representative sequences of the \ac{otu}s or directly to the \ac{esv}s.
-      In order to assign taxonomies to a particular sequence we need two things: a query tool and a taxonomy database.
-      The taxonomy database is a collection of 16S sequences of known organisms.
-      The query tool allows one to compare a sequence of interest to all the sequences in the database to find the best match.
-      The pipeline already incorporates \ac{gg} \cite{DeSantis2006}, SILVA \cite{Quast2012} and will soon include the \ac{ncbi} \cite{Sayers2009} as well.
+      In order to assign taxonomies to a particular sequence we use:
+      \begin{itemize}
+        \item A taxonomy database, which is a collection of 16S sequences of known organisms.
+        \item A query tool, which allows one to compare a sequence of interest to all the sequences in the database to find the best match.
+      \end{itemize}
+      The pipeline incorporates \ac{gg} \cite{DeSantis2006}, SILVA \cite{Quast2012} and the \ac{ncbi} \cite{Sayers2009} databases for taxonomy assignment
 
+    % TODO: Add Gabriel's new stuff here
     \subsubsection*{OTU and ESV Processing}
@@
         \item Rarefaction is a normalization technique used to overcome the bias that might arise due to variable sampling depth in different samples.
-This is commonly performed either by sub-sampling or by normalization of the matrix to the lowest sampling depth.
-        \item Filtering is usually performed to remove samples or features from the count matrix that are sparse.
-        \item Transformations are performed in order to correct for and overcome the compositional bias that is inherent in a counts matrix.
+This is performed either by sub-sampling or by normalization of the matrix to the lowest sampling depth.
+        \item Filtering is performed to remove samples or features (\ac{otu}s or \ac{esv}s) from the count matrix that are sparse.
+In order to determine the filtering threshold we fix the number of samples and correlation detection power needed  and determine the number of features to be used.
+        \item Transformations are performed in order to correct for and overcome the compositional bias that is inherent in a counts matrix (in most cases this is handled by the network inference algorithm).
       \end{itemize}
@@
       \vspace{-5mm}
-      This module deals with the inference of co-occurrence associations from the \ac{otu} or \ac{otu} counts matrix. These associations can be represented as a network.
+      This module deals with the inference of co-occurrence associations from the \ac{otu} or \ac{otu} counts matrix.
+These associations can be represented as a network, with nodes representing taxonomies and edges the association between them.
       \begin{itemize}
-        \item Associations can either be correlations or direct associations.
-        \item Significance of these associations is determined by re-sampling, bootstrapping and calculating pvalues.
+        \item A null model is created by re-sampling and bootstrapping the correlation/interaction matrix
+        \item Significance of these associations is determined by calculating pvalues against this null model.
       \end{itemize}
-      The pipeline incorporates \ac{sparcc} \cite{Friedman2012} and \ac{spieceasi} \cite{Kurtz2015} and will soon include \ac{mldm} \cite{Yang2017} and \ac{daa} \cite{Menon2018}.
+      The pipeline includes Pearson, Spearman and \ac{sparcc} \cite{Friedman2012} as the pairwise correlation metrics, and \ac{spieceasi} \cite{Kurtz2015}, \ac{mldm} \cite{Yang2017} and \ac{daa} \cite{Menon2018} as the direct interaction metrics.
 
-  \subsection*{Statistical Analysis}
+  \subsection*{Statistical Analysis of Count Matrices}
 
-    Tests
+   \textbf{Taxonomy comparisons}
+
+    \textbf{Power calculations}
+
+  \subsection*{Statistical Analysis of Co-occurrence Networks}
+
+    \textbf{Network Variability}:
+    In order to determine the best combination of tools for the pipeline, we examine the effects of each tool on the final co-occurrence network.
+    We can then estimate the variance contributed by each step in the pipeline to the total variance present across the networks.
+    \begin{itemize}
+      \item The distance metric used
+      \item How we determine the mean network
+      \item ANOVA?
+    \end{itemize}
+
+    \textbf{P-value merging}
+    We then use a p-value merging strategy (refer to the CoNet paper)
 
@@
 
-    In addition to the web application and pipeline, we have also developed a tool that facilitates easy visualization and exploration of the data generated by the \ac{mind} pipeline.
-    This tool is a Python package and is built using dash \cite{dash}.
+    The data visualizer facilitates easy visualization and exploration of the data generated by the pipeline.
+    This tool is a Python web-based interface built using dash \cite{dash}.
     It aids in the analysis of the effects of different processing methods on the generated counts matrix or network.
-    It takes in as input the results of a pipeline run and allows one to visualize various properties of intermediate \ac{otu} or \ac{esv} tables and final association networks.
-    This tool makes it easy and straightforward to compare different methods and combinations of methods in the 16S data analysis workflow.
-
-    The visualization tool currently supports:
+    The data from a pipeline run can be directly fed into the visualizer, allowing one to visualize various properties of intermediate \ac{otu} or \ac{esv} tables and final association networks.
+    This tool makes it easy and straightforward to compare different methods and combinations of methods in the 16S data analysis workflow. The visualization tool supports:
     \begin{itemize}
@@
     \end{itemize}
+
+  \subsection*{Code and Data Availability}
+
+    Links here

### 2018-08-07 sections/supplementary_figures.tex
Add 'supplementary_figures' section

%!TEX root = ../main.tex

\section*{Supplementary figures}

  \renewcommand{\thefigure}{S\arabic{figure}}
  \setcounter{figure}{0}

  \begin{figure}[h!]
    \centering
    \includegraphics[width=0.93\textwidth]{all_denoise_reg.pdf}
    \caption{
      \textbf{Pairwise regression plots for all the denoising methods}.
    }
    \label{fig:all_denoise_reg}
  \end{figure}

  \begin{figure}[h!]
    \centering
    \includegraphics[width=.9\textwidth]{denoise_network.pdf}
    \caption{
      \textbf{Networks generated using various denoising methods.}
      The denoising methods used in this result are closed-reference clustering, open-reference clustering, denovo clustering and \ac{dada2}.
      The database (\ac{gg}) and network inference methods (\ac{sparcc}) were kept constant.
      (A) The set of interactions predicted by using all the methods (union)
      (B) The set of interactions that were common to all the methods (intersection)
    }
    \label{fig:denoise_network}
  \end{figure}

### 2018-08-07 sections/results.tex
Update Results

@@
 
+We introduce (\hl{insert name here}) a flexible and modular pipeline for 16S data analysis. It incorporates various popular, publicly available tools as well as custom
+
+  \begin{figure}[h]
+    \centering
+    \includegraphics[width=0.9\textwidth]{mind_pipeline.pdf}
+    \caption{
+      \textbf{The workflow of the microbial co-occurrence analysis pipeline}.
+      The processes can be grouped into four major steps: \textbf{(A)} denoising and clustering, \textbf{(B)} taxonomy assignment, \textbf{(C)} \ac{otu}/\ac{esv} processing, and \textbf{(D)} network inference.
+      Each step incorporates several processes, each of which in turn have several alternate algorithms for the same task (indicated by the text to the right of the blue boxes).
+    }
+    \label{fig:mind_pipeline}
+  \end{figure}
+
+  \FloatBarrier
+
+  \subsection*{Variance among co-occurrence networks}
+
+  Figure \ref{fig:variance} shows the fraction of total variation in the networks as contributed by four different steps in the pipeline.
+  \todo[size=\footnotesize]{Results being processed}
+
+  \begin{figure}[h]
+    \centering
+    \includegraphics[width=0.9\textwidth]{main_figure.png}
+    \caption{\textbf{Percentage of variance in the final co-occurrence network due to each processing step.}}
+    \label{fig:variance}
+  \end{figure}
+
+  \FloatBarrier
+
+  \subsection*{Denoising and clustering methods are very dissimilar}
+
   There is significant variation both in the networks as well as the \ac{otu}/\ac{esv} tables when different combinations of methods are used to generate them.
-  Figure \ref{fig:otu_correlations} shows the comparison of several denoising and clustering methods.
-  The data shows that all the methods that perform either clustering or denoising are highly correlated with each other.
-  The correlation is low when comparing a method that performs clustering on the sequences with a method that performs denoising (Figure \ref{fig:otu_correlations}).
+  The \ac{esv} tables generated by methods that perform denoising are very similar to each other and the \ac{otu} tables generated by the clustering methods are very similar to each other, but results of denoising and clustering are highly uncorrelated as shown in Figure \ref{fig:otu_correlations}
   The common core network (network made up of edges that are predicted by all methods) is very small compared to the total number of edges predicted by all the methods (Figure \ref{fig:denoise_network}).
   In this result, only the denoising method was changed (database used: \ac{gg} and network inference: \ac{sparcc}).
-  But, nonetheless this approach of identifying a common network will help identify links with a good degree of confidence.
-
-  SILVA and \ac{gg} are two popular 16S databases used for taxonomy identification.
-  The two databases vastly differ in terms of their last update status - \ac{gg} was last updated on May 2013 and SILVA was last updated on December 2017.
-  Since updates to taxonomic classifications are frequent, these two databases vary significantly \cite{Balvociute2017}.
-  As a consequence of this, when different taxonomy databases are used on the same data-set we see large numbers of misclassifications \ref{fig:tax_mismatches}.
-
-  Different network inference methods infer significantly different association networks.
-  We can divide the network inference methods into two sets, the first set of methods (Pearson, Spearman, \ac{sparcc}) infer pairwise correlations while the second set infer direct associations (\ac{spieceasi}, \ac{mldm}, \ac{daa}).
-  Our preliminary analysis (Figure \ref{fig:network_comparison}) shows that networks generated using 4 different methods (Pearson, Spearman, \ac{sparcc}, \ac{spieceasi}) are vastly different and the degree of similarity (number of edges that overlapped) between the methods was dependent on the data-set used.
-  The abundance profiles of the different data-sets used are show in Figure \ref{fig:abundance_profile}.
-
-  The pipeline and the data visualization tool help the user explore the effect of the various methods on their data.
-  Exploring these differences are difficult and inconvenient without the tool since different tools differ in their input and output formats and require inter-converting between the various formats.
-  In the case of our pipeline, this is handled automatically.
 
   \begin{figure}[h]
-    \begin{center}
-      \includegraphics[width=17.8cm]{main_figure.png}
-      \caption{\textbf{Percentage of variance in the final co-occurrence network due to each processing step.} }
-      \label{fig:variance}
-    \end{center}
-  \end{figure}
-
-  \begin{figure}[h]
-    \begin{center}
-      \includegraphics[width=15.8cm]{otu_corr_full.pdf}
+    \centering
+    \includegraphics[width=0.9\textwidth]{otu_corr_full.pdf}
     \caption{
@@
     \label{fig:otu_correlations}
-    \end{center}
+  \end{figure}
+
+  \FloatBarrier
+
+  \subsection*{Taxonomy databases vary widely in taxonomy hierarchy and update frequency}
+
+  SILVA and \ac{gg} are two popular 16S databases used for taxonomy identification.
+  The two databases vastly differ in terms of their last update status - \ac{gg} was last updated on May 2013 and SILVA was last updated on December 2017.
+  Since updates to taxonomic classifications are frequent, these two databases vary significantly \cite{Balvociute2017}.
+  As a consequence of this, when different taxonomy databases are used on the same data-set we see large numbers of misclassifications \ref{fig:tax_mismatches}.
+
+  \begin{figure}[h]
+    \centering
+    \includegraphics[width=0.9\textwidth]{tax_comp_full.pdf}
+    \caption{\textbf{Taxonomy composition of the 20 most abundant genera predicted using different taxonomy references databases}. NOTE: I will replace the legend with the taxonomy tree}
+    \label{fig:tax_comparison}
   \end{figure}
@@
   \begin{figure}[h]
-    \begin{center}
-      \includegraphics[width=17.8cm]{tax_comp_full.pdf}
-      \caption{\textbf{Taxonomy composition of the 20 most abundant genera predicted using different taxonomy references databases}. NOTE: I will replace the legend with the taxonomy tree}
-      \label{fig:tax_comparison}
-    \end{center}
+    \centering
+    \includegraphics[width=0.95\textwidth]{tax_distance_deblur.pdf}
+    \caption{\textbf{Average percentage of mismatches in taxonomy assignment at various taxonomy levels}. NOTE: I will include a boxplot of the weighted distance in the taxonomy tree}
+    \label{fig:tax_mismatches}
   \end{figure}
 
-  \begin{figure}[h]
-    \begin{center}
-      \includegraphics[width=15.8cm]{tax_distance_deblur.pdf}
-      \caption{\textbf{Average percentage of mismatches in taxonomy assignment at various taxonomy levels}. NOTE: I will include a boxplot of the weighted distance in the taxonomy tree}
-      \label{fig:tax_mismatches}
-    \end{center}
-  \end{figure}
+  \subsection*{Network inference contributes to the most variance}
+
+  Different network inference methods infer significantly different association networks.
+  We can divide the network inference methods into two sets, the first set of methods (Pearson, Spearman, \ac{sparcc}) infer pairwise correlations while the second set infer direct associations (\ac{spieceasi}, \ac{mldm}, \ac{daa}).
+  Our preliminary analysis (Figure \ref{fig:network_comparison}) shows that networks generated using 4 different methods (Pearson, Spearman, \ac{sparcc}, \ac{spieceasi}) are vastly different and the degree of similarity (number of edges that overlapped) between the methods was dependent on the data-set used.
+  The abundance profiles of the different data-sets used are show in Figure \ref{fig:abundance_profile}.
 
   \begin{figure}[h]
-    \begin{center}
-      \includegraphics[width=17.8cm]{net_inference_comparison.pdf}
+    \centering
+    \includegraphics[width=0.9\textwidth]{net_inference_comparison.pdf}
     \caption{
@@
     \label{fig:network_comparison}
-    \end{center}
   \end{figure}
@@
   \begin{figure}[h]
-    \begin{center}
+    \centering
     %\includegraphics[width=17.8cm]{../figures/network_comparison.pdf}
-      \caption{\textbf{Core network captures known interactions in vaginal microbiota.} \textbf{a} Core network, obtained by taking the intersection between all the different pipelines.
-\textbf{b} Vaginal microbiota network as determined by XXX.}
+    \caption{\textbf{Core network captures known interactions in vaginal microbiota.} \textbf{(A)} Core network, obtained by taking the intersection between all the different pipelines.
+\textbf{(B)} Vaginal microbiota network as determined by XXX.}
     \label{fig:real_network}
-    \end{center}
   \end{figure}
+
+  % TODO: Based on these results: What is the best pipeline we recommend?
+  The pipeline and the data visualization tool help the user explore the effect of the various methods on their data.
+  Exploring these differences are difficult and inconvenient without the tool since different tools differ in their input and output formats and require inter-converting between the various formats.
+  In the case of our pipeline, this is handled automatically.

### 2018-08-08 sections/coverpage.tex
Update Coverpage

--- 2018-07-25 sections/coverpage.tex
@@
 
-  \vspace{3.5 cm}
+  \vspace*{\fill}
   { \huge \bfseries Correlating Microbial Abundances for the Microbial Interaction Network Database}\\[0.4cm] % Title of your document
@@
 
-  \vspace{1.5 cm}
-  Dileep Kishore, Gabriel Birzu, Zhenjun Hu, Charles DeLisi, Kirill Korolev, Daniel Segre\\
-  \vspace{1cm}
-  Department of XYZ,\\
-  The University of ABC\\
-  Department of ABC,\\
-  XYZ University\\
+  % \vspace{1.5 cm}
+  % Dileep Kishore\textsuperscript{1},
+  % Gabriel Birzu\textsuperscript{2},
+  % Zhenjun Hu\textsuperscript{1},
+  % Charles DeLisi\textsuperscript{1,3},
+  % Kirill Korolev\textsuperscript{1,2},
+  % Daniel Segr\`{e}\textsuperscript{1,3,4}\\
+  % \vspace{1cm}
+  % \textsuperscript{1}Bioinformatics Program, Boston University, Boston, Massachusetts, USA\\
+  % \textsuperscript{2}Department of Physics, Boston University, Boston, Massachusetts, USA\\
+  % \textsuperscript{3}Department of Biomedical Engineering, Boston University, Boston, Massachusetts, USA\\
+  % \textsuperscript{4}Department of Biology, Boston University, Boston, Massachusetts, USA\\
 
-% DATE SECTION
+% % DATE SECTION
 
@@
 
-\vfill % Fill the rest of the page with whitespace
+  \vspace*{\fill} % Fill the rest of the page with whitespace
 

### 2018-08-08 sections/introduction.tex
Update Introduction
1. No gold standard
2. Motivation for why we compare different methods with each other
3. How we conduct the study

@@
   The methodology for the analysis of 16S rRNA sequence data is not very straightforward and there are a myriad of tools for each step of the process, making the analysis complicated and not very reproducible.
+  But because there does not exist a dataset wherein the microbial interactions have been fully characterized, there is no straightforward way to determine which set of tools or methods generate the most accurate networks.
   In this study, we propose to build a standardized 16S data-analysis pipeline that will produce robust and reproducible co-occurrence networks from community 16S sequence data.
 \todo[size=\footnotesize]{Mention DB?}
+  We do this by conducting a systematic study of the various methods currently available.
+  We compare different methods to each other and additionally use mock microbial community datasets and synthetically generated reads to benchmark them.
+
   There exist many platforms like MG-RAST \cite{Keegan2016}, Qiita \cite{qiita} and tools such as QIIME \cite{Caporaso2010} that provide pipelines for 16S data analysis, but the main motivations behind the construction of our own pipeline are two-fold.

### 2018-08-16 sections/coverpage.tex
Update title options

@@
   \vspace*{\fill}
-  { \huge \bfseries Correlating Microbial Abundances for the Microbial Interaction Network Database}\\[0.4cm] % Title of your document
+  { \huge  1. Investigating the best practices for inference of microbial co-occurrence networks from 16S data}\\[0.4cm] % Title of your document
+  { \huge  2. Attempting to find the best practice pipeline for inferring co-occurrence networks from 16S data}\\[0.4cm] % Title of your document
+  { \huge  3. Deciphering the complexities in co-occurrence network inference from 16S data}\\[0.4cm] % Title of your document
 

### 2018-08-16 sections/methods.tex
Do not include method descriptions in the supplementary section
Minor style updates
Update methods

@@
     \begin{itemize}
-      \item The mock community 16S datasets were obtained from mockrobiota \cite{Bokulich2016}. We used samples (\hl{enter id number}) for this study.
+      \item The mock community 16S datasets were obtained from mockrobiota \cite{Bokulich2016}. We used mock numbers (\hl{enter id number}) for this study.
       \item The real datasets were control samples (49samples) \todo[size=\footnotesize]{Check the exact number} of a fecal microbiome transplant study obtained from \cite{Kang2017}.
@@
     \end{itemize}
+
+    \subsubsection*{Synthetic data}
+
+    \todo[size=\footnotesize]{Detailed options used to generate the data}
 
@@
     In addition to the Python package, the entire pipeline has been containerized into a Docker \cite{Merkel1994} image (\hl{dockerhub link}) for easy deployment and setup.
+
+    \hl{Package usage details}
+
     The main components of the pipeline are detailed in the following sections.
 
-    \todo[size=\footnotesize]{Should be in the supplementary section}
     \subsubsection*{Denoising and Clustering}
@@
       \end{itemize}
-      The pipeline incorporates \ac{gg} \cite{DeSantis2006}, SILVA \cite{Quast2012} and the \ac{ncbi} \cite{Sayers2009} databases for taxonomy assignment
+      The pipeline incorporates \ac{gg} \cite{DeSantis2006}, SILVA \cite{Quast2012} and the \ac{ncbi}~\cite{Sayers2009} databases for taxonomy assignment
 
@@
     The data from a pipeline run can be directly fed into the visualizer, allowing one to visualize various properties of intermediate \ac{otu} or \ac{esv} tables and final association networks.
-    This tool makes it easy and straightforward to compare different methods and combinations of methods in the 16S data analysis workflow. The visualization tool supports:
+    This tool makes it easy and straightforward to compare different methods and combinations of methods in the 16S data analysis workflow.
+    The visualization tool supports:
     \begin{itemize}

### 2018-08-16 sections/results.tex
Update results

@@
 
-We introduce (\hl{insert name here}) a flexible and modular pipeline for 16S data analysis. It incorporates various popular, publicly available tools as well as custom
+  We have developed \hl{insert package name here}, a flexible and modular pipeline for 16S data analysis.
+  It incorporates various popular, publicly available tools as well as custom Python modules and scripts to facilitate 16S data processing.
+  The pipeline workflow is as shown in Figure \ref{fig:mind_pipeline}, it supports the conversion of raw 16S sequence data or counts matrices into co-occurrence networks through multiple methods.
+  Each process in the pipeline supports alternate tools for performing the same task, users can use the configuration file to change these values.
+  The set of default options for the pipeline can be found in Table \ref{table:default_options}.
+  We also provide the users with a set of guidelines to facilitate tool selection as appropriate for their data.
+
+  \begin{table}[h]
+    \centering
+    \small
+    \begin{tabular}{|c|c|c|}
+      %\begin{tabularx}{\textwidth}{XXXX}
+      \hline
+      \textbf{Process} & \textbf{Tool} & \textbf{Parameters} \\
+      \hline
+      Denoising and Clustering & Deblur & ?? \\
+      Taxonomy Assignment & \ac{ncbi} + Blast & ?? \\
+      OTU Processing & Dynamic cutoff & ?? \\
+      Network Inference & User choice & ?? \\
+      \hline
+    \end{tabular}
+    %\end{tabularx}
+    \caption{Default tools and parameters for the pipeline}
+    \label{table:default_options}
+  \end{table}
 
@@
       Each step incorporates several processes, each of which in turn have several alternate algorithms for the same task (indicated by the text to the right of the blue boxes).
-    }
+    }%
     \label{fig:mind_pipeline}
@@
 
-  Figure \ref{fig:variance} shows the fraction of total variation in the networks as contributed by four different steps in the pipeline.
-  \todo[size=\footnotesize]{Results being processed}
+  In order to analyze the effect of different statistical methods on the inferred co-occurrence networks, we generate networks using all possible combinations of methods and estimate the variability in the networks.
+  Figure \ref{fig:variance} shows the fraction of total variation among the co-occurrence networks as contributed by each of the four main steps of the pipeline.
+  \todo[size=\footnotesize]{Results soon!}
 
@@
     \includegraphics[width=0.9\textwidth]{main_figure.png}
-    \caption{\textbf{Percentage of variance in the final co-occurrence network due to each processing step.}}
+    \caption{\textbf{Percentage of variance in the final co-occurrence network due to each processing step.}}%
     \label{fig:variance}
@@
 
-  There is significant variation both in the networks as well as the \ac{otu}/\ac{esv} tables when different combinations of methods are used to generate them.
-  The \ac{esv} tables generated by methods that perform denoising are very similar to each other and the \ac{otu} tables generated by the clustering methods are very similar to each other, but results of denoising and clustering are highly uncorrelated as shown in Figure \ref{fig:otu_correlations}
+  There is significant variation both in the \ac{otu}/\ac{esv} tables when different combinations of methods are used to generate them.
+  We process the same 16S dataset using 5 different methods - open-reference clustering, closed-reference clustering and denovo clustering from \ac{qiime1}~\cite{Caporaso2010}, \ac{dada2}~\cite{Callahan2016} and Deblur~\cite{Amir2017}.
+  For each of these methods we use the same taxonomy database (\ac{gg}) to assign taxonomies to the representative sequences.
+  We then correlate the abundances of matching taxonomies between every combination of a pair of methods (Figure \ref{fig:all_denoise_reg}).
+  The \ac{esv} tables generated by methods that perform denoising are very similar to each other $\sim0.9$ and the \ac{otu} tables generated by the clustering methods are very similar to each other $\sim0.9$, but results of denoising and clustering are highly uncorrelated with each other $\sim0.4$ (Figure \ref{fig:otu_correlations}).
+
   The common core network (network made up of edges that are predicted by all methods) is very small compared to the total number of edges predicted by all the methods (Figure \ref{fig:denoise_network}).
-  In this result, only the denoising method was changed (database used: \ac{gg} and network inference: \ac{sparcc}).
+  In these results, only the denoising method was changed (database used: \ac{gg} and network inference: \ac{sparcc}).
+
+  These comparisons only elucidate the pairwise similarity or dissimilarity of a pair of methods.
+In order to determine the best method we use synthetically generated reads from an Illumina read simulator called Grinder~\cite{Angly2012}.
+  The reads were simulated using the taxonomy distribution profile generated by \ac{qiime1} open-reference clustering using the \hl{FMT} dataset.
+  Reference sequences were obtained from \ac{gg} and other metrics such as sampling depths, chimera percentage were estimated from the original sequences.
+  Detailed information on all the other settings are provided in the methods section.
+  We find that the \ac{qiime1} methods (clustering) are not as robust when compared to the denoising methods in re-generating the same abundance distribution.
+\todo[size=\footnotesize]{Not yet confirmed}
 
@@
     \centering
-    \includegraphics[width=0.9\textwidth]{otu_corr_full.pdf}
+    \includegraphics[width=0.9\textwidth]{edited/otu_corr_full.pdf}
     \caption{
@@
   SILVA and \ac{gg} are two popular 16S databases used for taxonomy identification.
-  The two databases vastly differ in terms of their last update status - \ac{gg} was last updated on May 2013 and SILVA was last updated on December 2017.
-  Since updates to taxonomic classifications are frequent, these two databases vary significantly \cite{Balvociute2017}.
-  As a consequence of this, when different taxonomy databases are used on the same data-set we see large numbers of misclassifications \ref{fig:tax_mismatches}.
+  The two databases vastly differ in terms of their last update status - \ac{gg} was last updated on May 2013 and SILVA was last updated on December 2017 at the time of writing.
+  Since updates to taxonomic classifications are frequent, these two databases vary significantly in terms of taxonomy hierarchies~\cite{Balvociute2017}.
+
+  The deblur denoising pipeline was first used to obtain \ac{esv} counts table by processing the FMT 16S sequence data.
+  These \ac{esv}s are then assigned a particular taxonomic label using two different reference 16S taxonomy databases - SILVA and \ac{gg}.
+  Figure~\ref{fig:tax_comparison} shows that the top 20 most abundant genera in the two resulting taxonomy composition tables are \hl{different}.
+  The most abundant genus in the \ac{gg} taxonomy table was \textit{Escherichia} whereas in the SILVA taxonomy table it was \textit{Escherichia-Shigella}.
+  These might seem like minor differences, but when comparing a large number of composition tables these problems are hard to diagnose.
+  Figure~\ref{fig:tax_comparison_all} shows that the result is very similar when you look at all genera instead of just the top 20 most abundant genera.
+  The percentage of this misclassification was found to change with the taxonomy resolution needed.
+  The mismatches were $\sim10\%$ at the Phylum level but were seen to go up to $\sim60\%$ at the Species level, as shown in Figure~\ref{fig:tax_mismatches}.
+  Moreover, Figure~\ref{fig:venn_tax} shows that there does not seem to exist a correlation between abundance and mismatch.
+  This implies that the most abundant sequences are not necessarily the ones that are consistently matched to the same taxonomies in the different reference databases.
+  \todo[size=\footnotesize]{Incorporate NCBI}
+
+  These results prompted us to turn to \ac{ncbi} for taxonomy referencing.
+  Our reasons for choosing \ac{ncbi} are that it is regularly maintained and updated and accuracy of the assignment is comparable to the SILVA and \ac{gg} reference databases.
+  \hl{If we use NCBI we need to show how accurate that is. We can show this using the synthetic data or mock communities}
 
@@
     \includegraphics[width=0.9\textwidth]{tax_comp_full.pdf}
-    \caption{\textbf{Taxonomy composition of the 20 most abundant genera predicted using different taxonomy references databases}. NOTE: I will replace the legend with the taxonomy tree}
+    \caption{\textbf{Taxonomy composition of the 20 most abundant genera predicted using different taxonomy references databases}}
     \label{fig:tax_comparison}
@@
     \includegraphics[width=0.95\textwidth]{tax_distance_deblur.pdf}
-    \caption{\textbf{Average percentage of mismatches in taxonomy assignment at various taxonomy levels}. NOTE: I will include a boxplot of the weighted distance in the taxonomy tree}
+    \caption{\textbf{Average percentage of mismatches in taxonomy assignment at various taxonomy levels.}}
     \label{fig:tax_mismatches}
@@
 
+  \FloatBarrier
+
   \subsection*{Network inference contributes to the most variance}
+
+  \hl{How did we get these results? What do the figures tell us? What is the conclusion?}
+
+  The taxonomy composition table obtained using the \ac{gg} reference database is used to infer co-occurrence associations between the microbes.
 

### 2018-12-07 figures.tex
Add figures draft #2

%--------------------------------------------------------%
% Figures
%--------------------------------------------------------%

\input{sections/preamble}

% \titleformat{\subsection}[runin]{}{}{}{}[]

\title{ \huge Microbial Interaction Database pipeline - Figures draft \#2 }
\author{DK, GB, ZH, CD, KK, DS}
\date{\today}

\begin{document}

\maketitle

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{figure1.pdf}
  \caption{
    \textbf{The workflow of the microbial co-occurrence analysis pipeline}.
    The processes can be grouped into four major steps: \textbf{(A)} denoising and clustering, \textbf{(B)} taxonomy assignment, \textbf{(C)} OTU/ESV processing, and \textbf{(D)} network inference.
    Each step incorporates several processes, each of which in turn have several alternate algorithms for the same task (indicated by the text to the right of the blue boxes).
    \textbf{(E)} shows the average variances in the final networks contributed by each of the four groups over all the test datasets
  }
  \label{fig:main_figure}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{figure2.pdf}
  \caption{
    \textbf{Comparison of various denoising and clustering algorithms used in the workflow.}
    \textbf{(A)} The average weighted UniFrac distance between the representative sequences generated by each denosing/clustering method
    \textbf{(B)} The average unweighted UniFrac distance between the representative sequences generated by each denosing/clustering method
    \textbf{(C)} The distributions of the average unweighted UniFrac distance between original sequence profile and calculated profile in mock datasets
  }
  \label{fig:figure2}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{figure3.pdf}
  \caption{
    \textbf{Comparison of taxonomy assignments using different taxonomy reference databases}
    \textbf{(A)} Taxonomy composition of the 20 most abundant genera predicted using different taxonomy references databases: Greengenes, SILVA and NCBI.
    The legend shows the common and the unique genera among the taxonomy assignments.
    \textbf{(B)}
  }
  \label{fig:figure3}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{figure4.pdf}
  \caption{Effect of OTU processsing measures on the OTU table. This is just a placeholder}
  \label{fig:figure4}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{figure5.pdf}
  \caption{
    \textbf{Networks generated using different network inference methods}.
    The four different networks generated by different network inference methods are very dissimilar (C).
    (A) The overlaps between the edges among the network5 generated is shown. The number of edges that are common to all networks are very low (5).
    (B) The hamming distance between the networks is shown. The similarity between various methods was found to vary with the data-source used.
}
  \label{fig:figure5}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{figure6.pdf}
  \caption{Example networks generated using our pipeline: Final consensus networks for some datasets}
  \label{fig:figure6}
\end{figure}

\FloatBarrier

\subsection*{Supplementary}%

\begin{figure}[h]
  \centering
  \includegraphics[width=0.67\linewidth]{figures1.pdf}
  \caption{
    \textbf{Comparison of various denoising and clustering algorithms used in the pipeline}.
    (A, B) Correlation of the abundances of the taxa that are common between the count matrices created by two different methods.
    (A) The best correlation (most similar methods) is between open-reference and denovo.
    (B) The worst correlation (least similar methods) is between open-reference and dada2.
    (C) A heatmap showing the $\mathrm{R}^2$ of all pairwise comparisons of the methods.
  }
  \label{fig:figures1}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=0.8\linewidth]{pdf/all_denoise_reg.pdf}
  \caption{All pairwise correlations comparing the similarity between different denoising and clustering methods}
  \label{fig:figures2}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=0.8\linewidth]{pdf/denoise_network.pdf}
  \caption{A network showing union and intersection of networks generated using certain combination of methods}
  \label{fig:pdf/denoise_network}
\end{figure}

\end{document}

### 2018-12-19 figures.tex
Update Figure2 caption

@@
   \caption{
-    \textbf{Comparison of various denoising and clustering algorithms used in the workflow.}
-    \textbf{(A)} The average weighted UniFrac distance between the representative sequences generated by each denosing/clustering method
-    \textbf{(B)} The average unweighted UniFrac distance between the representative sequences generated by each denosing/clustering method
-    \textbf{(C)} The distributions of the average unweighted UniFrac distance between original sequence profile and calculated profile in mock datasets
+    \textbf{Comparison of representative sequences generated by the various denoising and clustering algorithms used in the workflow.}
+    \textbf{(A)} The average weighted UniFrac distance between the representative sequences,
+    \textbf{(B)} The average unweighted UniFrac distance,
+    \textbf{(C)} The distributions of the average unweighted UniFrac distance between original sequence profile and calculated profile in mock datasets (this will be replaced with a violoin plot)
   }

### 2018-12-19 figures.tex
Update figure2

@@
     \textbf{(B)} The average unweighted UniFrac distance,
-    \textbf{(C)} The distributions of the average unweighted UniFrac distance between original sequence profile and calculated profile in mock datasets (this will be replaced with a violoin plot)
+    \textbf{(C)} The distributions of the average unweighted UniFrac distance between original sequence profile and calculated sequence profile in mock datasets (this will be replaced with a violoin plot)
   }

### 2018-12-19 figures.tex
Update figure3

@@
     The legend shows the common and the unique genera among the taxonomy assignments.
-    \textbf{(B)}
+    \textbf{(B)} The percentage of OTUs assigned to the same taxonomic label when using different reference databases.
+    The percentage of matches are shown at various taxonomic levels.
+    \textbf{(C)} The bray-curtis dissimilarity between the original taxonomy profile and calculated taxonomy profile in the mock datasets.
   }

### 2018-12-19 figures.tex
Move figure4 to figures2
Update figure4

@@
   \includegraphics[width=\linewidth]{figure4.pdf}
-  \caption{Effect of OTU processsing measures on the OTU table. This is just a placeholder}
-  \label{fig:figure4}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=\linewidth]{figure5.pdf}
   \caption{
     \textbf{Networks generated using different network inference methods}.
-    The four different networks generated by different network inference methods are very dissimilar (C).
-    (A) The overlaps between the edges among the network5 generated is shown. The number of edges that are common to all networks are very low (5).
-    (B) The hamming distance between the networks is shown. The similarity between various methods was found to vary with the data-source used.
+    \textbf{(A)} The four different networks generated by different network inference methods are very dissimilar
+    \textbf{(B)} The edge overlap upset plot
+    \textbf{(C)} The node overlap upset plot.
   }
-  \label{fig:figure5}
+  \label{fig:figure4}
 \end{figure}
@@
   \centering
+  \includegraphics[width=0.7\linewidth]{figures2.pdf}
+  \caption{Effect of OTU processsing measures on the OTU table. This is just a placeholder}
+  \label{fig:figures2}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
   \includegraphics[width=0.8\linewidth]{pdf/all_denoise_reg.pdf}

### 2019-03-25 figures.tex
Update supplementary figures
Update: Figure 1
Add figure 6

@@
   \centering
-  \includegraphics[width=\linewidth]{figure1.pdf}
+  \includegraphics[width=0.7\linewidth]{figure1.pdf}
   \caption{
@@
 
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.8\linewidth]{figure6.pdf}
+  \caption{
+    \textbf{The choice of database contributes to the most variance in the networks}.
+    \textbf{(A)} The relative variance in the networks contributed by the denoising, taxonomy assignment and filtering steps.
+    \textbf{(B)} The PCA plot clearly shows that the networks can be separated based on database
+  }
+  \label{fig:figure1}
+\end{figure}
+
 \FloatBarrier
@@
   \caption{All pairwise correlations comparing the similarity between different denoising and clustering methods}
-  \label{fig:figures2}
+  \label{fig:figures3}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.8\linewidth]{png/hamming_distance.png}
+  \caption{The hamming distance between the networks is shown. The similarity between various methods was found to vary with the data-source used.}
+  \label{fig:figures4}
 \end{figure}
@@
   \caption{A network showing union and intersection of networks generated using certain combination of methods}
-  \label{fig:pdf/denoise_network}
+  \label{fig:figures5}
 \end{figure}

### 2019-04-16 sections/coverpage.tex
Add: New title suggestion for the article

--- 2018-08-08 sections/coverpage.tex
@@
   \vspace*{\fill}
-  { \huge \bfseries Correlating Microbial Abundances for the Microbial Interaction Network Database}\\[0.4cm] % Title of your document
+  {\huge 1. Inferring microbial co-occurrence networks from 16S data: A systematic evaluations}\\[0.4cm]
+  {\huge  2. Investigating the best practices for inference of microbial co-occurrence networks from 16S data}\\[0.4cm] % Title of your document
+  {\huge  3. Attempting to find the best practice pipeline for inferring co-occurrence networks from 16S data}\\[0.4cm] % Title of your document
+  {\huge  4. Deciphering the complexities in co-occurrence network inference from 16S data}\\[0.4cm] % Title of your document
 

### 2019-04-16 sections/abstract.tex
Minor rephrasing
Change: Original abstract replaced with ISMB 2019 abstract

--- 2018-07-31 sections/abstract.tex
@@
 %!TEX root = ../main.tex
-
-% QUESTION: How many words is this limited to?
 
@@
     \noindent
-    Microorganisms naturally occur in communities which can consist of tens of thousands of species involved in complex interactions with each other.
-    16S ribosomal RNA (rRNA) gene profiling provides snapshots that reveal the abundances and phylogenies of microorganisms in microbial communities.
-    These snapshots can then be used to unveil the complex interactions governing these microbial communities.
-    However, the nature of 16S data-sets makes the data processing difficult and error prone.
-    The development of appropriate and well-validated analysis methods that do not produce statistical artifacts is still ongoing and is receiving much attention.
-    Here, we perform a thorough analysis of each step in this process to determine the best tools and parameter choices that generate the most accurate and robust networks.
-    Ultimately, we create a sequence analysis pipeline to standardize the processing of high-throughput community 16S rRNA sequencing data into networks of microbial interactions derived from their co-occurrences.
-    \hl{The development of a standard pipeline for 16S data analysis would enable accurate comparison of data from different data-sets.} \todo[size=\footnotesize]{TODO: Talk briefly about the main result}
+    Microbes tend to organize into communities consisting of hundreds of species involved in complex interactions with each other.
+    16S ribosomal RNA (16S rRNA) gene profiling provides snapshots that reveal the phylogenies and abundance distributions of these microbial communities.
+    These snapshots, when collected from multiple samples, have the potential to reveal which microbes co-occur, providing a glimpse into the network of inter-dependencies underlying these communities.
+    The inference of networks from 16S data is prone to statistical artifacts, but the extent to which the different steps of the workflow affect the resultant network is still unclear.
+    In this study, we perform a meticulous analysis of each step of a pipeline that processes 16S sequencing data into a network of microbial associations.
+    Through this process, we determine the tools and parameters that generate the most accurate and robust co-occurrence.
+    Ultimately, we develop a standardized pipeline that follows these default tools and parameters, but that can also help explore the outcome of any other combination of choices.
+    We envisage that this standard pipeline for processing 16S sequencing data into networks of microbial co-occurrences could be used for integrating multiple data-sets, and for generating comparative analyses and consensus networks useful for detecting disease-related patterns.
+\todo[size=\footnotesize]{NOTE: Abstract constrained to 200 words} \\
     The pipeline is available as a Python package and a docker container.
@@
 % Insert keywords here
-\keywords{Microbiome, 16S rRNA, Pipeline, Interaction, Network, Correlations, Qiime}
+\keywords{Microbiome, 16S rRNA, Pipeline, Interaction, Network, Correlations, Qiime, Co-occurrence}

### 2019-04-23 scripts/figure4/figure4_data.py
Add: All figures after 4th draft

@@ -13,8 +13,6 @@ def write_tables(file_paths, notus, output_path):
         otu_df = otu.to_dataframe(dense=True)
         obs_metadata = otu.metadata_to_dataframe(axis="observation")
         obs_metadata["Abundance"] = otu_df.sum(axis=1)
-        if notus < 0:
-            notus = obs_metadata.shape[0]
         final_df = obs_metadata.sort_values(by="Abundance", ascending=False).iloc[
             :notus
         ]

### 2019-04-23 figures.tex
Update figure 6(b) legend
Change: Update figures.tex for 4th draft and presentation

@@
 
-\title{ \huge Microbial Interaction Database pipeline - Figures draft \#2 }
+\title{ \huge Microbial Interaction Database pipeline - Figures draft \#4 }
 \author{DK, GB, ZH, CD, KK, DS}
@@
   \centering
-  \includegraphics[width=0.7\linewidth]{figure1.pdf}
+  \includegraphics[width=0.67\linewidth]{figure1.pdf}
   \caption{
     \textbf{The workflow of the microbial co-occurrence analysis pipeline}.
-    The processes can be grouped into four major steps: \textbf{(A)} denoising and clustering, \textbf{(B)} taxonomy assignment, \textbf{(C)} OTU/ESV processing, and \textbf{(D)} network inference.
+    The processes can be grouped into four major steps: \textbf{(DC)} denoising and clustering, \textbf{(TA)} taxonomy assignment, \textbf{(OP)} OTU/ESV processing, and \textbf{(NI)} network inference.
     Each step incorporates several processes, each of which in turn have several alternate algorithms for the same task (indicated by the text to the right of the blue boxes).
-    \textbf{(E)} shows the average variances in the final networks contributed by each of the four groups over all the test datasets
   }
-  \label{fig:main_figure}
+  \label{fig:figure1}
 \end{figure}
@@
   \centering
-  \includegraphics[width=\linewidth]{figure2.pdf}
+  \includegraphics[width=0.75\linewidth]{figure2.pdf}
   \caption{
-    \textbf{Comparison of representative sequences generated by the various denoising and clustering algorithms used in the workflow.}
-    \textbf{(A)} The average weighted UniFrac distance between the representative sequences,
-    \textbf{(B)} The average unweighted UniFrac distance,
-    \textbf{(C)} The distributions of the average unweighted UniFrac distance between original sequence profile and calculated sequence profile in mock datasets (this will be replaced with a violoin plot)
+    \textbf{The choice of taxonomic reference database contributes to the most variance in the networks}.
+    \textbf{(A)} The relative variance in the networks contributed by the denoising, taxonomy assignment and filtering steps.
+    \textbf{(B)} The PCA plot shows that the networks can be clearly separated based on the database used for the taxonomy assignment
   }
@@
   \caption{
-    \textbf{Comparison of taxonomy assignments using different taxonomy reference databases}
-    \textbf{(A)} Taxonomy composition of the 20 most abundant genera predicted using different taxonomy references databases: Greengenes, SILVA and NCBI.
-    The legend shows the common and the unique genera among the taxonomy assignments.
-    \textbf{(B)} The percentage of OTUs assigned to the same taxonomic label when using different reference databases.
-    The percentage of matches are shown at various taxonomic levels.
-    \textbf{(C)} The bray-curtis dissimilarity between the original taxonomy profile and calculated taxonomy profile in the mock datasets.
+    \textbf{The representative sequences generated by the different denoising/clustering methods mostly differ in the sequences that are in low abundance.}
+    \textbf{(A)} The average weighted UniFrac distance between the representative sequences shows that the representative sequences and their compositions are fairly identical between the methods,
+    \textbf{(B)} The largers average unweighted UniFrac distance indicates that methods differ in their identifaction of sequences of low abundance,
+    \textbf{(C)} The distributions of the average unweighted UniFrac distance between the expected sequence profile and the calculated sequence profile in mock datasets show that dada2 and deblur were the best performing methods in most of the datasets.
   }
@@
   \caption{
-    \textbf{Networks generated using different network inference methods}.
-    \textbf{(A)} The four different networks generated by different network inference methods are very dissimilar
-    \textbf{(B)} The edge overlap upset plot
-    \textbf{(C)} The node overlap upset plot.
+    \textbf{Taxonomic reference databases vary widely in terms of their taxonomies.}
+    \textbf{(A)} The assignment of the top 50 representative sequences to their respective taxonomies using the three different reference databases shows how the same sequences are assigned to different genus.
+    \textbf{(B)} The percentage of OTUs assigned to the same taxonomic label when using different reference databases.
+    The percentage of mis-matches decrease at higher taxonomic levels but even at the Phylum level there exists around 10\% of mis-matches.
+    \textbf{(C)} The bray-curtis dissimilarity between the expected taxonomy profile and calculated taxonomy profile in the mock datasets shows that there is no best choice of database for every dataset.
   }
@@
   \centering
-  \includegraphics[width=\linewidth]{figure6.pdf}
-  \caption{Example networks generated using our pipeline: Final consensus networks for some datasets}
+  \includegraphics[width=0.85\linewidth]{figure5.pdf}
+  \caption{
+    \textbf{Networks generated using different network inference methods show marked differences both in terms of edge-density and connectivity}.
+    \textbf{(A)} The six different networks generated by the different network inference methods are very dissimilar
+    \textbf{(B)} The node overlap upset plot indicates that all the networks have a large number of common nodes involved in connections.
+    Whereas, \textbf{(C)} The edge overlap upset plot shows that a very small fraction of these connections are actually shared.
+  }
   \label{fig:figure6}
@@
   \centering
-  \includegraphics[width=0.8\linewidth]{figure6.pdf}
+  \includegraphics[width=0.9\linewidth]{figure6.pdf}
   \caption{
-    \textbf{The choice of database contributes to the most variance in the networks}.
-    \textbf{(A)} The relative variance in the networks contributed by the denoising, taxonomy assignment and filtering steps.
-    \textbf{(B)} The PCA plot clearly shows that the networks can be separated based on database
+    \textbf{Network inference and taxonomic assignment have the highest influence on the inferred network structures.}
+    \textbf{(A)} The network constructed using the default pipeline parameters (denosing=dada2, database=ncbi, otu\_filtering=on, network\_inference=sparcc) is compared with networks when one of the steps use a different tool.
+    The common connections are in black, connections unique to the network are colored and connections in the default network and but not present in the current network are green.
+    \textbf{(B)} The L1 distance of the networks generated by changing one step of the default pipeline against the network generated using the default parameters.
   }
@@
 \subsection*{Supplementary}%
+
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{S\arabic{figure}}
 
@@
   \centering
-  \includegraphics[width=0.7\linewidth]{figures2.pdf}
-  \caption{Effect of OTU processsing measures on the OTU table. This is just a placeholder}
+  \includegraphics[width=0.8\linewidth]{figures2.pdf}
+  \caption{Results of the mock community analysis. (A) Weighted unifrac (B) Unweighted unifrac}
   \label{fig:figures2}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.8\linewidth]{figures3.pdf}
+  \caption{Results of the synthetic community analysis. (A) Weighted unifrac (B) Unweighted unifrac}
+  \label{fig:figures3}
 \end{figure}
@@
   \caption{All pairwise correlations comparing the similarity between different denoising and clustering methods}
-  \label{fig:figures3}
+  \label{fig:figures4}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=\linewidth]{figures4.pdf}
+  \caption{
+    \textbf{(A)} Taxonomy composition of the 20 most abundant genera predicted using different taxonomy references databases: Greengenes, SILVA and NCBI.
+    The legend shows the common and the unique genera among the taxonomy assignments.
+}
+  \label{fig:figures4}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=\linewidth]{figures5.pdf}
+  \caption{
+    The bray-curtis dissmilarity between the expected taxonomic composition and generated taxonomic composiion for the synthetic datasets.
+}
+  \label{fig:figures4}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=\linewidth]{figures6.pdf}
+  \caption{
+    The effect of OTU processing measurs on the OTU table
+}
+  \label{fig:figures4}
 \end{figure}

### 2019-05-01 sections/abstract.tex
Add: GitHub link to mindpipe

@@
 \todo[size=\footnotesize]{NOTE: Abstract constrained to 200 words} \\
-    The pipeline is available as a Python package and a docker container.
+    The pipeline is available as a Python package and a docker container at \href{https://github.com/dileep-kishore/mindpipe}{https://github.com/dileep-kishore/mindpipe}
   }

### 2019-05-01 sections/methods.tex
Style: Change indentation

@@
 
-  \subsection*{MIND Data Visualizer}
-
-    The data visualizer facilitates easy visualization and exploration of the data generated by the pipeline.
-    This tool is a Python web-based interface built using dash \cite{dash}.
-    It aids in the analysis of the effects of different processing methods on the generated counts matrix or network.
-    The data from a pipeline run can be directly fed into the visualizer, allowing one to visualize various properties of intermediate \ac{otu} or \ac{esv} tables and final association networks.
-    This tool makes it easy and straightforward to compare different methods and combinations of methods in the 16S data analysis workflow.
-    The visualization tool supports:
-    \begin{itemize}
-      \item Filtering and aggregating data from multiple sources (data-sets)
-      \item Comparing properties of the generated counts matrices and networks across different methods
-      \item Visualization of these properties
-    \end{itemize}
-
   \subsection*{Code and Data Availability}

### 2019-05-09 sections/methods.tex
Change: Update methods section

@@
   \vspace{-5mm}
-  \begin{itemize}
-    \item The mock community 16S datasets were obtained from mockrobiota \cite{Bokulich2016}. We used mock numbers (\hl{enter id number}) for this study.
-    \item The real datasets were control samples (49samples) \todo[size=\footnotesize]{Check the exact number} of a fecal microbiome transplant study obtained from \cite{Kang2017}.
-    \item Simulated data was generated using Grinder \cite{Angly2012}.
-  \end{itemize}
+  The study uses three kinds of 16S rRNA sequencing datasets \- real datasets, mock datasets and synthetic datasets.
+  Real datasets are those sequencing reads obtained from naturally occurring microbial community samples.
+  The current study used healthy stool samples from a fecal microbiome transplant study \cite{Kang2017} and healhty saliva samples from a periodontal disease study~\cite{Chen2018} as real datasets for analysis.
+  The mock community 16S datasets were obtained from mockrobiota~\cite{Bokulich2016}.
+  The mock datasets used for this study were labelled mock4, mock12 and mock16 datasets.
+  The synthetic datasets were generated using an artificial read simulator called ART~\cite{Huang2012}.
+  Three different microbial composition profiles were used as input
+  Synthetic reads were generated using a soil and water microbiome composition profiles from the Earth Microbiome Project~\cite{Thompson2017} and healthy gut microbiome project from the Human Microbiome Project~\cite{HumanMicrobiomeProjectConsortium2012}
+  The scripts used to generate the synthetic data are in the \hl{scripts folder of the respository}
 
-  \subsubsection*{Synthetic data}
-
-  \todo[size=\footnotesize]{Detailed options used to generate the data}
-
-  \subsection*{16S processing pipeline}
+  \subsection*{mindpipe}
 
   \vspace{-5mm}
-
-  The flowchart describing the workflow of the 16S data-analysis pipeline is as shown in Figure \ref{fig:mind_pipeline}.
-  The pipeline (\hl{github link here}) integrates many publicly available tools as well as custom Python modules and scripts to extract co-occurrence associations from 16S sequence data.
-  The input to the pipeline by default is the raw community 16S rRNA sequence data, but it can additionally be configured to use processed sequences, \ac{otu} tables or other types of intermediate data.
+  The flowchart describing the workflow of the 16S data-analysis pipeline is as shown in Figure \ref{fig:figure1}.
+  The pipeline integrates many publicly available tools as well as custom Python modules and scripts to extract co-occurrence associations from 16S sequence data.
+  The inputs to the pipeline by default are the raw community 16S rRNA sequence reads, but it can also additionally be configured to use trimmed sequences, \ac{otu} tables and other types of intermediate data.
   The final output of the pipeline is the inferred network of co-occurrence relationships among the microbes present in the samples.
 
-  Some key features of the pipeline include modularity, flexibility and ease-of-use.
-  The entire pipeline consists of modules which can be modified, reordered or replaced appropriately using a simple configuration script.
-  The configuration script lists the steps to be performed during runtime, along with the parameters to be used (optional) for the various steps.
+  The mindpipe pipeline provides both a Python API as well as a command-line interface and only requires a single configuration file.
+  The configuration file lists the inputs, output and the steps to be performed during runtime, along with the parameters to be used (if different from defaults) for the various steps.
   Since the entire pipeline run-through is stored in the form of a text file (the configuration file), subsequent runs are highly reproducible and changes can be easily tracked using version control.
-  The pipeline allows for automatic parallelization of all possible processes, both within and across samples.
-  In addition to the Python package, the entire pipeline has been containerized into a Docker \cite{Merkel1994} image (\hl{dockerhub link}) for easy deployment and setup.
+  It uses the nextflow workflow manager~\cite{Tommaso2015} under the hood and facilitates the usage of the pipeline on the local machine, cluster or cloud with minimal configuration change.
+  It also allows for automatic parallelization of all possible processes, both within and across samples.
+  The pipeline is designed to be modular, where each tool or method is organized into modules which can be easily modified or replaced.
+  This also simplifies the process of adding new tools (refer to modules section in the mindpipe documentation).
+  In addition to the Python package, the entire pipeline has been containerized into a Docker~\cite{Merkel1994} image (\hl{dockerhub link}) for easy deployment and setup.
+  The main components of the pipeline are detailed in the subsequent sections.
 
-  \hl{Package usage details}
-
-  The main components of the pipeline are detailed in the following sections.
-
-  \subsubsection*{Denoising and Clustering}
+  \subsection*{Denoising and Clustering}
   \vspace{-5mm}
@@
   The output of this module is a counts matrix that describes the number of reads of a particular \ac{otu} or \ac{esv} (rows of the matrix) present in each sample (columns of the matrix).
-    The pipeline incorporates \ac{qiime1} \cite{Caporaso2010}, \ac{qiime2}, \ac{dada2} \cite{Callahan2016} and Deblur \cite{Amir2017} tools in this module.
+  The pipeline incorporates open reference clustering, closed reference clustering and de novo clustering methods from \ac{qiime1}~\cite{Caporaso2010} and denoising methods from \ac{dada2}~\cite{Callahan2016} and Deblur~\cite{Amir2017}.
 
-  \subsubsection*{Taxonomy Assignment}
+  \subsection*{Taxonomy Assignment}
   \vspace{-5mm}
@@
   \end{itemize}
-    The pipeline incorporates \ac{gg} \cite{DeSantis2006}, SILVA \cite{Quast2012} and the \ac{ncbi}~\cite{Sayers2009} databases for taxonomy assignment
+  The pipeline incorporates \ac{gg}~\cite{DeSantis2006}, SILVA~\cite{Quast2012} and the \ac{ncbi}~\cite{Sayers2009} databases for taxonomy assignment and the Naive Bayes classifier from \ac{qiime2} and \ac{ncbi} blast as the query tools.
 
   % TODO: Add Gabriel's new stuff here
-  \subsubsection*{OTU and ESV Processing}
+  \subsection*{OTU and ESV Processing}
   \vspace{-5mm}
@@
 
-  \subsubsection*{Network Inference}
+  \subsection*{Network Inference}
   \vspace{-5mm}
@@
   \end{itemize}
-    The pipeline includes Pearson, Spearman and \ac{sparcc} \cite{Friedman2012} as the pairwise correlation metrics, and \ac{spieceasi} \cite{Kurtz2015}, \ac{mldm} \cite{Yang2017} and \ac{daa} \cite{Menon2018} as the direct interaction metrics.
+  The pipeline includes Pearson, Spearman and \ac{sparcc}~\cite{Friedman2012} as the pairwise correlation metrics, and \ac{spieceasi}~\cite{Kurtz2015}, \ac{mldm}~\cite{Yang2017} and \ac{magma}~\cite{} as the direct association metrics.
 
-  \subsection*{Statistical Analysis of Count Matrices}
-
- \textbf{Taxonomy comparisons}
-
-  \textbf{Power calculations}
-
-  \subsection*{Statistical Analysis of Co-occurrence Networks}
+  \subsection*{Statistical Analysis}
 
   \textbf{Network Variability}:
-  In order to determine the best combination of tools for the pipeline, we examine the effects of each tool on the final co-occurrence network.
-  We can then estimate the variance contributed by each step in the pipeline to the total variance present across the networks.
-  \begin{itemize}
-    \item The distance metric used
-    \item How we determine the mean network
-    \item ANOVA?
-  \end{itemize}
+  The adjacency matrices of the networks generated from all possible combinations of tools are merged.
+  The variance contributed by each step of the pipeline is calculated for every connection in the merged network using ANOVA.
+  The total variance for the network is calculated by adding the variances for each connection.
+  The PCA analysis is also performed on the merged network.
+
+  \textbf{Power calculations:}
 
   \textbf{P-value merging}
-  We then use a p-value merging strategy (refer to the CoNet paper)
+  The empirical Browns method is used for combining p-values from the various methods to obtain a consensus pvalue, which is used to create the consensus network.
 

### 2019-05-09 sections/results.tex
Change: Update results section

--- 2018-07-27 sections/results.tex
@@
 
-  There is significant variation both in the networks as well as the \ac{otu}/\ac{esv} tables when different combinations of methods are used to generate them.
-  Figure \ref{fig:otu_correlations} shows the comparison of several denoising and clustering methods.
-  The data shows that all the methods that perform either clustering or denoising are highly correlated with each other.
-  The correlation is low when comparing a method that performs clustering on the sequences with a method that performs denoising (Figure \ref{fig:otu_correlations}).
-  The common core network (network made up of edges that are predicted by all methods) is very small compared to the total number of edges predicted by all the methods (Figure \ref{fig:denoise_network}).
-  In this result, only the denoising method was changed (database used: \ac{gg} and network inference: \ac{sparcc}).
-  But, nonetheless this approach of identifying a common network will help identify links with a good degree of confidence.
+  \subsection*{mindpipe}
 
-  SILVA and \ac{gg} are two popular 16S databases used for taxonomy identification.
-  The two databases vastly differ in terms of their last update status - \ac{gg} was last updated on May 2013 and SILVA was last updated on December 2017.
-  Since updates to taxonomic classifications are frequent, these two databases vary significantly \cite{Balvociute2017}.
-  As a consequence of this, when different taxonomy databases are used on the same data-set we see large numbers of misclassifications \ref{fig:tax_mismatches}.
+  We have developed \hl{mindpipe}, a flexible and modular pipeline for 16S data analysis.
+  It incorporates various popular, publicly available tools as well as custom Python modules and scripts to facilitate inference of co-occurrence networks from 16S data.
+  The pipeline workflow is as shown in Figure~\ref{fig:figure1}, it supports the conversion of raw 16S sequence data or counts matrices into co-occurrence networks through multiple methods.
+  Each process in the pipeline supports alternate tools for performing the same task, users can use the configuration file to change these values.
+  We perform a systematic analysis of each step of the pipeline to determine the set of default options for the pipeline (Table~\ref{tab:default_options}).
+  We also provide the users with a set of guidelines to facilitate tool selection as appropriate for their data.
 
-  Different network inference methods infer significantly different association networks.
-  We can divide the network inference methods into two sets, the first set of methods (Pearson, Spearman, \ac{sparcc}) infer pairwise correlations while the second set infer direct associations (\ac{spieceasi}, \ac{mldm}, \ac{daa}).
-  Our preliminary analysis (Figure \ref{fig:network_comparison}) shows that networks generated using 4 different methods (Pearson, Spearman, \ac{sparcc}, \ac{spieceasi}) are vastly different and the degree of similarity (number of edges that overlapped) between the methods was dependent on the data-set used.
-  The abundance profiles of the different data-sets used are show in Figure \ref{fig:abundance_profile}.
-
-  The pipeline and the data visualization tool help the user explore the effect of the various methods on their data.
-  Exploring these differences are difficult and inconvenient without the tool since different tools differ in their input and output formats and require inter-converting between the various formats.
-  In the case of our pipeline, this is handled automatically.
+  \begin{table}[h]
+    \centering
+    \small
+    \begin{tabular}{|c|c|c|}
+      %\begin{tabularx}{\textwidth}{XXXX}
+      \hline
+      \textbf{Process} & \textbf{Tool} & \textbf{Parameters} \\
+      \hline
+      Denoising and Clustering & Dada2 & ?? \\
+      Taxonomy Assignment & \ac{ncbi} + Blast & ?? \\
+      OTU Processing & Dynamic cutoff & ?? \\
+      Network Inference & Consensus & ?? \\
+      \hline
+    \end{tabular}
+    %\end{tabularx}
+    \caption{Default tools and parameters for the pipeline}
+    \label{tab:default_options}
+  \end{table}
 
   \begin{figure}[h]
-    \begin{center}
-      \includegraphics[width=17.8cm]{main_figure.png}
-      \caption{\textbf{Percentage of variance in the final co-occurrence network due to each processing step.} }
-      \label{fig:variance}
-    \end{center}
+    \centering
+    \includegraphics[width=0.7\linewidth]{figure1.pdf}
+    \caption{
+      \textbf{The workflow of the microbial co-occurrence analysis pipeline}.
+      The processes can be grouped into four major steps: \textbf{(DC)} denoising and clustering, \textbf{(TA)} taxonomy assignment, \textbf{(OP)} OTU/ESV processing, and \textbf{(NI)} network inference.
+      Each step incorporates several processes, each of which in turn have several alternate algorithms for the same task (indicated by the text to the right of the blue boxes).
+      The text along the arrows describes the data that is being passed from one step to another.
+    }
+    \label{fig:figure1}
+  \end{figure}
+
+  % TODO: Talk about the datasets: real, mock and synthetic (with references)
+  In order to objectively determine the best method we use both mock data from mockrobiota~\cite{Bokulich2016} as well as, synthetically generated reads from an Illumina read simulator called ART~\cite{Huang2012}.
+  The mock datasets selected were mock4, mock12 and mock16 since they had expected sequences.
+  The synthetic reads were simulated using the three different taxonomy distribution profiles generated from soil, water and hl{risk} datasets.
+  Reference sequences were generated using \ac{ncbi} and the \hl{decard} tool \cite{Golob2017} for the required taxonomy profiles.
+  Detailed information on the mock communities and all the other settings used to generate the sequence data are provided in the methods section.
+
+  \FloatBarrier
+
+  \subsection*{The choice of reference database and network inference algorithm have the biggest impact on inferred networks}
+
+  In order to analyze the effect of different statistical methods on the inferred co-occurrence networks, we generate co-occurrence networks using all possible combinations of methods and estimate the variability in the networks due to each choice.
+  Figure \ref{fig:figure2} shows the fraction of total variation among the co-occurrence networks as contributed by each of the four main steps of the pipeline.
+  The 16S reference database contributes to the most variation ($\sim25\%$) in the networks and the networks can be clearly separated based on the database used.
+  \todo[size=\footnotesize]{Wait for results from network inference methods}
+  This indicates that the taxonomy assigned to the reference sequences drastically alters the co-occurrence network and this change is much more significant than how the reference sequences themselves are identified (in the denoise/cluster step).
+  The residual variation can be seen as an artifact that arises due to the effect on the resultant network when multiple steps are changed at the same time.
+  Another interesting observation has been that the dissimilarity between the networks decreases when the low abundance OTUs are removed from the network.
+
+  \begin{figure}
+    \centering
+    \includegraphics[width=0.85\linewidth]{figure2.pdf}
+    \caption{
+      \textbf{The choice of database contributes to the most variance in the networks}.
+      \textbf{(A)} The relative variance in the networks contributed by the DC, TA and OP steps of the pipeline.
+      \textbf{(B)} All combinations of inferred networks are shown as points on a PCA plot.
+The color of the points corresponds to the taxonomy database, the shape corresponds to the denoising/clustering method and the size corresponds to whether low abundance OTUs were removed or not.
+      The plot clearly shows that the points can be separated based on the TA step and that the differences due to the DC and OP steps are not as significant.
+    }
+    \label{fig:figure2}
+  \end{figure}
+
+  \FloatBarrier
+
+  \subsection*{Denoising and clustering methods are differ in identifying less common reference sequences}
+
+  There is good agreement in the \ac{otu}/\ac{esv} tables when different combinations of methods are used to generate them.
+  We process the same 16S sequencing reads (\hl{FMT} dataset) using 5 different methods \- open-reference clustering, closed-reference clustering, denovo clustering, \ac{dada2}~\cite{Callahan2016} and Deblur~\cite{Amir2017}.
+  The first three methods are from the \ac{qiime1}~\cite{Caporaso2010} package.
+
+  To compare the representative sequences generated by these methods we employ both the weighted~\cite{Lozupone2007} (Figure~\ref{fig:figure3}A) and unweighted UniFrac method~\cite{Lozupone2005} (Figure~\ref{fig:figure3}B).
+  The weighted UniFrac distance metric takes into account the counts of the representative sequences, whereas the unweighted UniFrac distance metric does not and hence gives equal weights to each sequence.
+  From Figure~\ref{fig:figure3}A we can observe that the representative sequences generated by the different methods are similar to each other when weighted by their abundance.
+  Figure~\ref{fig:figure3}B on the other hand shows an increase in dissimilarity between each pair of methods suggesting that the methods might differ in the sequences of low abundance.
+  In order to verify this claim, we changed the threshold for the sequence count and recalculated the weighted and unweighted UniFrac distance metrics and also looked at the number of representative sequences identified in each case.
+\todo[size=\footnotesize]{Haven't done this.
+This should be in supplementary}
+
+  %% TODO: Check if this is necessary
+  \textcolor{gray}{
+    For each of these methods we use the same taxonomy database (\ac{gg}) to assign taxonomies to the representative sequences.
+    We then correlate the abundances of matching taxonomies between every combination of a pair of methods (Figure \ref{fig:all_denoise_reg}).
+    The \ac{esv} tables generated by methods that perform denoising are very similar to each other $\sim0.9$ and the \ac{otu} tables generated by the clustering methods are very similar to each other $\sim0.9$, but results of denoising and clustering are highly uncorrelated with each other $\sim0.4$ (Figure \ref{fig:figureS2}).
+    The common core network (network made up of edges that are predicted by all methods) is very small compared to the total number of edges predicted by all the methods (Figure \ref{fig:denoise_network}).
+    In these results, only the denoising method was changed (database used: \ac{gg} and network inference: \ac{sparcc}).
+  }
+
+  These comparisons only elucidate the pairwise similarity or dissimilarity of a pair of methods.
+  In order to determine the tool that most accurately recapitulates the reference sequences in the samples we the mock datasets.
+  The mock community datasets we processed by the various methods and the predicted representative sequences were compared with expected representative sequences and their distribution.
+  The results (Figure~\ref{fig:figure3}C and D) show that the performance of the methods are similar to each other as in the case of the real dataset, but, are very different from the expected sequence distribution.
+  This result is more exaggerated in the case of the unweighted UniFrac metric where some of the datasets show a very high deviation from the expected sequences.
+  These high deviations are primarily in two of the three datasets that were analyzed and show that the datasets themselves play a big role in the performance of these methods.
+  This can be clearly seen in the performance (weighted UniFrac distance) of \ac{dada2} and Deblur on mock12 and mock16 datasets, where, Deblur outperforms \ac{dada2} on mock12 but the under-performs on mock16.
+
+  There is no method that clearly outperforms the rest in all datasets.
+  Based on their slightly better performance on the mock datasets, their \textit{de novo} error correcting nature and other previous studies (\hl{refs}) we recommend either \ac{dada2} or Deblur for users.
+  However, as a result of the unexpected bad performance of Deblur on the synthetic data, the default algorithm in the pipeline was chosen to be \ac{dada2}.
+
+  \begin{figure}
+    \centering
+    \includegraphics[width=0.75\textwidth]{figure3.pdf}
+    \caption{
+      \textbf{The representative sequences generated by the different denoising/clustering methods are very similar but differ slightly in the sequences that are in low abundance.}
+      \textbf{(A)} The average weighted UniFrac distance between the representative sequences shows that the representative sequences and their compositions are fairly identical between the methods,
+      \textbf{(B)} The relatively larger average unweighted UniFrac distance indicates that methods differ in their identification of sequences of low abundance,
+      \textbf{(C)} The distributions of the average unweighted UniFrac distance between the expected sequence profile and the calculated sequence profile in mock datasets show that dada2 and Deblur were the best performing methods in most of the datasets.
+    }
+    \label{fig:figure3}
+  \end{figure}
+
+  \FloatBarrier
+
+  \subsection*{Taxonomy databases vary widely in taxonomy hierarchy and update frequency}
+
+  The three 16S taxonomic reference databases used in this study are SILVA~\cite{Quast2012}, \ac{gg}~\cite{DeSantis2006} and \ac{ncbi} RefSeq.
+  SILVA and \ac{gg} are two popular 16S databases used for taxonomy identification.
+  The \ac{ncbi} RefSeq nucleotide database contains 16S rRNA sequences as a part of two BioProjects - 33175 and 33317.
+  The three databases vastly differ in terms of their last update status - \ac{gg} was last updated on May 2013, SILVA was last updated on December 2017 at the time of writing and \ac{ncbi} is updated as new sequences are curated.
+  Since updates to taxonomic classifications are frequent, these databases vary significantly in terms of taxonomy hierarchies~\cite{Balvociute2017}.
+
+  The representative sequences obtained from the \ac{dada2} method in DC step were used for taxonomic assignment using the three reference databases.
+  Figure~\ref{fig:figure4}A depicts a flow diagram that shows how the top 50 representative sequences (sorted by abundance) are assigned a Genus according to the three different databases.
+  We observe that not only does the assigned Genus composition vary significantly, but, the percentage of unassigned representative sequences (gray) also differ.
+  Even the most abundant representative sequence is assigned to an "unknown" Genus in two of the three databases.
+  A representative sequence might be get assigned an "unknown" Genus if either the taxonomy identifier associated with the sequence in the database did not contain a Genus or the more likely case where the database contains multiple sequences that are very similar to the query (representative) sequence and the consensus algorithm (from \ac{qiime2}) is unable to assign one particular Genus at the required confidence.
+  After assigning all the representative sequences to taxonomies we perform a pairwise comparison of the similarity between assignments from different databases at every taxonomic level (Figure~\ref{fig:figure4}B).
+  The assignments beyond Family level (Family, Genus and Species) are very dissimilar with $<70\%$ similarity between any pair of databases.
+  There are no two reference databases that are more similar than the other, with \ac{gg} and SILVA producing only marginally similar assignments compared to \ac{ncbi}.
+
+  As in the previous section, these comparisons only indicate similarity or dissimilarity between methods.
+  In order to obtain an absolute measure of accuracy of the taxonomic assignments we use the expected sequences from the mock datasets as the query sequences for the databases and the expected taxonomic composition as the standard to compare against (Figure~\ref{fig:figure4}C).
+  Again, we observe that none of the databases perform better than the others. \todo[size=\footnotesize]{Has to be verified}
+
+  % TODO: Check if this is necessary
+  \textcolor{gray}{
+    Figure~\ref{fig:figureS3} shows that the top 20 most abundant genera in the three resulting taxonomy composition tables are different.
+    The most abundant genus in the \ac{gg} taxonomy table was \textit{Escherichia} whereas in the SILVA taxonomy table it was \textit{Escherichia-Shigella}.
+    These might seem like minor differences, but when comparing a large number of composition tables these problems are hard to diagnose.
+    Figure~\ref{fig:tax_comparison_all} shows that the result is very similar when you look at all genera instead of just the top 20 most abundant genera.
+    Moreover, Figure~\ref{fig:venn_tax} shows that there does not seem to exist a correlation between abundance and mismatch.
+    This implies that the most abundant sequences are not necessarily the ones that are consistently matched to the same taxonomies in the different reference databases.
+  }
+
+  These results prompted us to turn to \ac{ncbi} for taxonomy assignment.
+  Our reasons for choosing \ac{ncbi} are that it is regularly maintained and updated and accuracy of the assignment is still comparable to the SILVA and \ac{gg} reference databases that are routinely used as reference databases.
+  \hl{Or combine results from multiple databases?}
+
+  \begin{figure}[h]
+    \centering
+    \includegraphics[width=0.72\textwidth]{figure4.pdf}
+    \caption{
+      \textbf{Taxonomic reference databases vary widely in terms of their taxonomies.}
+      \textbf{(A)} The assignment of the top 50 representative sequences to their respective taxonomies using the three different reference databases shows how the same sequences are assigned to different genus.
+      \textbf{(B)} The percentage of \ac{OTU}s assigned to the same taxonomic label when using different reference databases.
+      The percentage of mismatches decrease at higher taxonomic levels but even at the Phylum level there exists around 10\% of mismatches.
+      \textbf{(C)} The Bray-Curtis dissimilarity between the expected taxonomy profile and calculated taxonomy profile in the mock datasets shows that there is no best choice of database for every dataset.
+    }
+    \label{fig:figure4}
+  \end{figure}
+
+  \FloatBarrier
+
+  \subsection*{Networks generated using different network inference methods show marked difference in edge-density and connectivity}
+
+  % TODO: Add reference here
+  The six different network inference methods used in this study are \ac{magma}, \ac{mldm}, \ac{spieceasi}, \ac{sparcc}, Spearman and Pearson.
+  These network inference methods into two groups, the first set of methods (Pearson, Spearman, \ac{sparcc}) infer pairwise correlations while the second set infer direct associations (\ac{spieceasi}, \ac{mldm}, \ac{magma}).
+
+  The taxonomy composition table obtained using the \ac{ncbi} reference database is used as the input to these algorithms to infer co-occurrence associations between the microbes.
+  Figure~\ref{fig:figure5}A shows the networks inferred from this dataset using the different inference algorithms.
+  An edge weight threshold of 0.3 was applied to the networks generated from the methods that infer pairwise correlations and no threshold was applied in case of the other networks.
+  We can clearly see that the different networks differ vastly in their edge-density and connectivity and even some of the edges in common to these networks have their signs inverted.
+  To get a more quantitative idea, we can take a look at the nodes and edges (Figure~\ref{fig:figure5}B) in common between the networks using UpSet plots (only \ac{magma}, \ac{mldm}, \ac{spieceasi}, \ac{sparcc} are used for comparison).
+  The results for the node intersections show that the networks have a large number of edges in common ($63$ out of $67$ nodes in the smallest network - \ac{magma}) and no network possesses any unique node.
+  The edge intersections in contrast show that only $19$ edges (out of $98$ edges in the smallest network - \ac{magma}) are in common between all the methods and each network has a large number of unique edges.
+  These results indicate that there is a substantial rewiring of connections in the inferred networks.
+
+  Unlike in the previous sections, where were we evaluated the performance of methods on mock datasets, there are no such datasets that contain a set of known interactions for the evaluation of the network inference algorithms.
+  Therefore, we propose the construction of a \hl{consensus network} (Figure~\ref{fig:figure5}C) involving \ac{magma}, \ac{mldm}, \ac{spieceasi} and \ac{sparcc} by merging the p-values generated from bootstraps of the original taxonomy composition table.
+  The resulting network will have weights reported by all four methods.
+
+  \begin{figure}[h]
+    \centering
+    \includegraphics[width=0.85\textwidth]{figure5.pdf}
+    \caption{
+      \textbf{Networks generated using different network inference methods show marked differences both in terms of edge-density and connectivity}.
+      \textbf{(A)} The six different networks generated by the different network inference methods are very dissimilar.
+      The green links are positive associations and the orange links are negative associations.
+      A threshold of 0.3 was set for the methods that infer pairwise correlations (\ac{sparcc}, Spearman, Pearson) and no threshold was set for the other methods.
+      \textbf{(B)} The node overlap upset plot indicates that all the networks have a large number of common nodes involved in connections.
+      Whereas, \textbf{(C)} The edge overlap upset plot shows that a very small fraction of these connections are actually shared.
+    }
+    \label{fig:figure5}
+  \end{figure}
+
+  \FloatBarrier
+
+  \subsection*{The default network}
+
+  The default tools and parameters chosen for the pipeline are given in Table~\ref{tab:default_options}.
+  The default
+
+  Figure~\ref{fig:figure6}A shows the default network compared against networks generated by altering one of the steps of the pipeline from the default.
+  These results indicate that the biggest differences in networks occur when the reference database or the network inference algorithm are changed.
+  Furthermore, the L1 distance of networks generated by altering one of the steps of the pipeline from the default against the default network (Figure~\ref{fig:figure6}B) show that the biggest deviations from the default network occur when the \ac{ta} and \ac{ni} steps are changed, reinforcing the same results observed in Figure~\ref{fig:figure2}.
+
+  \begin{figure}[h]
+    \centering
+    \includegraphics[width=0.9\linewidth]{figure6.pdf}
+    \caption{
+      \textbf{Network inference and taxonomic assignment have the highest influence on the inferred network structures.}
+      \textbf{(A)} The network constructed using the default pipeline parameters (DC=\ac{dada2}, TA=\ac{ncbi}, OP=on, NI=\ac{sparcc}) is compared with networks when one of the steps use a different tool.
+      The common connections are in black, connections unique to the network are colored purple and connections in the default network and but not present in the current network are gray.
+      \textbf{(B)} The L1 distance of the networks generated by changing one step of the default pipeline against the network generated using the default parameters.
+    }
+    \label{fig:figure6}
   \end{figure}
@@
   \begin{figure}[h]
-    \begin{center}
-      \includegraphics[width=15.8cm]{otu_corr_full.pdf}
+    \centering
+    \includegraphics[width=17cm]{figure7.pdf}
     \caption{
-        \textbf{Comparison of various denoising and clustering algorithms used in the pipeline}.
-        (A, B) Correlation of the abundances of the taxa that are common between the count matrices created by two different methods.
-        (A) The best correlation (most similar methods) is between open-reference and denovo.
-        (B) The worst correlation (least similar methods) is between open-reference and dada2.
-        (C) A heatmap showing the $\mathrm{R}^2$ of all pairwise comparisons of the methods.
+      \textbf{The networks generated using the default pipeline settings.}
   }
-      \label{fig:otu_correlations}
-    \end{center}
+    \label{fig:figure7}
   \end{figure}
-
-  \begin{figure}[h]
-    \begin{center}
-      \includegraphics[width=17.8cm]{tax_comp_full.pdf}
-      \caption{\textbf{Taxonomy composition of the 20 most abundant genera predicted using different taxonomy references databases}. NOTE: I will replace the legend with the taxonomy tree}
-      \label{fig:tax_comparison}
-    \end{center}
-  \end{figure}
-
-  \begin{figure}[h]
-    \begin{center}
-      \includegraphics[width=15.8cm]{tax_distance_deblur.pdf}
-      \caption{\textbf{Average percentage of mismatches in taxonomy assignment at various taxonomy levels}. NOTE: I will include a boxplot of the weighted distance in the taxonomy tree}
-      \label{fig:tax_mismatches}
-    \end{center}
-  \end{figure}
-
-  \begin{figure}[h]
-    \begin{center}
-      \includegraphics[width=17.8cm]{net_inference_comparison.pdf}
-      \caption{
-        \textbf{Networks generated using different network inference methods}.
-        The four different networks generated by different network inference methods are very dissimilar (C).
-        (A) The overlaps between the edges among the network5 generated is shown. The number of edges that are common to all networks are very low (5).
-        (B) The hamming distance between the networks is shown. The similarity between various methods was found to vary with the data-source used.
-      }
-      \label{fig:network_comparison}
-    \end{center}
-  \end{figure}
-
-  \begin{figure}[h]
-    \begin{center}
-      %\includegraphics[width=17.8cm]{../figures/network_comparison.pdf}
-      \caption{\textbf{Core network captures known interactions in vaginal microbiota.} \textbf{a} Core network, obtained by taking the intersection between all the different pipelines.
-\textbf{b} Vaginal microbiota network as determined by XXX.}
-      \label{fig:real_network}
-    \end{center}
-  \end{figure}

### 2019-05-30 sections/abstract.tex
Change: Remove todo

--- 2019-04-16 sections/abstract.tex
@@
     We envisage that this standard pipeline for processing 16S sequencing data into networks of microbial co-occurrences could be used for integrating multiple data-sets, and for generating comparative analyses and consensus networks useful for detecting disease-related patterns.
-\todo[size=\footnotesize]{NOTE: Abstract constrained to 200 words} \\
-    The pipeline is available as a Python package and a docker container.
+\\
+    The pipeline is available as a Python package and a docker container at \href{https://github.com/dileep-kishore/mindpipe}{https://github.com/dileep-kishore/mindpipe}
   }

### 2019-05-30 sections/results.tex
Change: Remove todo and fix typos

@@
   The 16S reference database contributes to the most variation ($\sim25\%$) in the networks and the networks can be clearly separated based on the database used.
-  \todo[size=\footnotesize]{Wait for results from network inference methods}
+  \todo[size=\footnotesize]{Update with full results}
   This indicates that the taxonomy assigned to the reference sequences drastically alters the co-occurrence network and this change is much more significant than how the reference sequences themselves are identified (in the denoise/cluster step).
@@
 
-  \subsection*{Denoising and clustering methods are differ in identifying less common reference sequences}
+  \subsection*{Denoising and clustering methods differ in their identificaiton of less common reference sequences}
 
@@
   In order to verify this claim, we changed the threshold for the sequence count and recalculated the weighted and unweighted UniFrac distance metrics and also looked at the number of representative sequences identified in each case.
-\todo[size=\footnotesize]{Haven't done this.
-This should be in supplementary}
+\todo[size=\footnotesize]{In supplementary}
 
@@
   In order to obtain an absolute measure of accuracy of the taxonomic assignments we use the expected sequences from the mock datasets as the query sequences for the databases and the expected taxonomic composition as the standard to compare against (Figure~\ref{fig:figure4}C).
-  Again, we observe that none of the databases perform better than the others. \todo[size=\footnotesize]{Has to be verified}
+  Again, we observe that none of the databases perform better than the others.
 

### 2019-06-05 sections/methods.tex
Add: Add item
Add: Line about db consensus algorithm and magma ref

@@
     \item A query tool, which allows one to compare a sequence of interest to all the sequences in the database to find the best match.
+    \item Consensus method
   \end{itemize}
-  The pipeline incorporates \ac{gg}~\cite{DeSantis2006}, SILVA~\cite{Quast2012} and the \ac{ncbi}~\cite{Sayers2009} databases for taxonomy assignment and the Naive Bayes classifier from \ac{qiime2} and \ac{ncbi} blast as the query tools.
+  The pipeline incorporates \ac{gg}~\cite{DeSantis2006}, SILVA~\cite{Quast2012} and the \ac{ncbi}~\cite{Sayers2009} databases for taxonomy assignment and the Naive Bayes classifier from \ac{qiime2} and \ac{ncbi} blast as the query tools (from \ac{qiime2}).
+  The consensus algorithm used is also defaults used by the classifiers in \ac{qiime2}.
 
@@
   \end{itemize}
-  The pipeline includes Pearson, Spearman and \ac{sparcc}~\cite{Friedman2012} as the pairwise correlation metrics, and \ac{spieceasi}~\cite{Kurtz2015}, \ac{mldm}~\cite{Yang2017} and \ac{magma}~\cite{} as the direct association metrics.
+  The pipeline includes Pearson, Spearman and \ac{sparcc}~\cite{Friedman2012} as the pairwise correlation metrics, and \ac{spieceasi}~\cite{Kurtz2015}, \ac{mldm}~\cite{Yang2017} and \ac{magma} (\hl{in biorxiv}) as the direct association metrics.
 

### 2019-06-05 sections/results.tex
Change: Minor grammatical changes

@@
 
-  \subsection*{Networks generated using different network inference methods show marked difference in edge-density and connectivity}
+  \subsection*{Networks generated using different network inference methods show notable difference in edge-density and connectivity}
 
@@
   We can clearly see that the different networks differ vastly in their edge-density and connectivity and even some of the edges in common to these networks have their signs inverted.
-  To get a more quantitative idea, we can take a look at the nodes and edges (Figure~\ref{fig:figure5}B) in common between the networks using UpSet plots (only \ac{magma}, \ac{mldm}, \ac{spieceasi}, \ac{sparcc} are used for comparison).
+  To get a more quantitative idea, we can take a look at the nodes and edges (Figure~\ref{fig:figure5}B) in common between the networks using UpSet plots (only \ac{magma}, \ac{mldm}, \ac{spieceasi}, \ac{sparcc} are used in the comparison).
   The results for the node intersections show that the networks have a large number of edges in common ($63$ out of $67$ nodes in the smallest network - \ac{magma}) and no network possesses any unique node.
@@
     \caption{
-      \textbf{Networks generated using different network inference methods show marked differences both in terms of edge-density and connectivity}.
+      \textbf{Networks generated using different network inference methods show notable differences both in terms of edge-density and connectivity}.
       \textbf{(A)} The six different networks generated by the different network inference methods are very dissimilar.

### 2019-06-05 sections/discussion.tex
Update discussion
Add: Write first draft of discussion section

@@
 
-  Discussion
+% TODO:
+% * Emphasize that the tool makes it really easy to analyze your data using multiple methods and then compare
+% * It is also easy to add new tools
+% * Write about each step and potential things people could do to better analyze their data
+% * Future steps
+
+% General statements
+Co-occurrence associations in microbial communities will identify important interactions that drive microbial community structure and organization.
+There is still much disagreement between the networks generated using various methods.
+This study makes a significant effort to identify those tools and parameters that enable more accurate inference of these co-occurrence networks.
+The mindpipe pipeline also enables the user to explore the effect of various tools and their combinations on their dataset of interest, thereby enabling one to identify those tools that perform their best on that particular dataset.
+
+The different denoising/clustering methods differ mostly in their identification of sequences that are in low abundances.
+Hence, they do not have much of an impact on the inferred co-occurrence networks when the sequences of abundance are removed.
+However, comparison of inferred and expected reference sequences and their abundances in mock community datasets has allowed us to identify \ac{dada2} as the method which best recapitulates expected the sequence composition.
+There are newer recently published methods like dbOTU3~\cite{Olesen2017} as well as older popular methods like mothur~\cite{Schloss2009} that haven't been included in the study, and are possible avenues for future investigation.
+
+The choice of taxonomy database was found to be the most important factor in the inference of a microbial co-occurrence network, contributing $\sim20\%$ of the total variance.
+The frequent changes in the taxonomy nomenclature coupled with the frequency of updates to the various 16S reference databases create inherent differences \cite{Balvociute2017} in taxonomy hierarchies in these databases.
+Our analysis did not reveal one particular reference database to perform better in all scenarios and we believe that that choice of the database should be made based on the organisms expected to be present in the dataset.
+The default reference database in the pipeline is the NCBI 16S RefSeq database as it is more frequently updated.
+We also enable users to use custom databases \cite{Griffen2011,Ritari2015} with the blast and naive bayes classifiers that are incorporated into the pipeline (from \ac{qiime2})
+
+% From "Accuracy of taxonomy prediction for 16S rRNA and fungal ITS sequences":
+% (It is clear from the results reported here that taxonomy prediction methods cannot be definitively ranked according to their accuracy using TAXXI or previous benchmark tests.
+However, some trends are apparent which can guide a biologist toward an appropriate method for addressing a given research question.
+In most cases, the highest true-positive rates are achieved by the simple top-hit classifiers (see e.g., Table 4).
+This shows, as might be expected, that the true name of a known rank is usually present in the top hit annotation, and there is no indication that more complicated methods achieve any improvement in this respect by successfully identifying anomalous cases where a taxon is known but has a lower identity than the top hit.
+This shows that the biggest challenge in algorithm design can be framed as predicting the LCR, or, equivalently, deciding how many ranks of the taxonomy in the top hit should be deleted in the prediction.
+Most methods incorporate thresholds which determine the lowest rank to predict using a measure such as identity (MEGAN, Metaxa2), bootstrap (RDP, SINTAX, Q2SK), posterior probability (BLCA), and/or frequency in the top hits (MEGAN, Q1, Q2BLAST and Q2VS).
+Adjusting a threshold necessarily makes a trade-off between true positives and false positives, because more true positives can be obtained only at the expense of including more false positives, reflecting the fact that taxonomy correlates only approximately with any measure of sequence similarity.
+Thus, different thresholds may be appropriate in different studies depending on whether false negatives or false positives are more important.)
+
+Filtering out taxa that are present in low abundances in all samples did not increase (in most datasets tested) the proportion of taxa in common between taxonomy tables generated using different reference databases.
+However, we do observe that the reduction in the number of taxa leads better agreement in the networks inferred through different methods.
+Moreover, filtering is necessary in order to increase the power in tests of significance when the number of taxa is much greater than the number of samples.
+
+The networks generated by different network inference methods show considerable differences in edge-density and connectivity.
+One reason for this is because of the underlying assumptions regarding sparsity, distribution and compositionality that the algorithms make.
+The consensus network created by merging the networks inferred using the different network inference methods enables the creation of a network whose links have evidence based on multiple inference algorithms.
+
+\hl{Other factors that play an important role in network inference} \\
+Differences in environments of the samples could lead to the inference of spurious interactions.
+
+Exploring the effects of these combinations of methods on the resultant networks is difficult and inconvenient since different tools differ in their input and output formats and require inter-converting between the various formats.
+The pipeline enables these by providing a variety of modules for inter-conversion between various formats, additionally it also allows easy incorporation of new tools as modules.
+The mindpipe pipeline not only enables users to generate co-occurrence networks using the default recommended options, but also allows users to identify tools or parameter options that tailored to their datasets.
+Thus, in this study we have identified a set of tools and parameters that help infer more accurately and consistently co-occurrence networks from microbial 16S sequencing data.

### 2019-06-05 sections/supplementary_figures.tex
Add: Supplementary figures
Add: Add all supplementary figures

@@
 
+\newpage
 \section*{Supplementary figures}
@@
 
-  \begin{figure}[h!]
+  \begin{figure}[h]
   \centering
-    \includegraphics[width=0.93\textwidth]{all_denoise_reg.pdf}
+  \includegraphics[width=0.67\linewidth]{figureS1.pdf}
   \caption{
-      \textbf{Pairwise regression plots for all the denoising methods}.
+    \textbf{Comparison of various denoising and clustering algorithms used in the pipeline}.
+    (A, B) Correlation of the abundances of the taxa that are common between the count matrices created by two different methods.
+    (A) The best correlation (most similar methods) is between open-reference and denovo.
+    (B) The worst correlation (least similar methods) is between open-reference and dada2.
+    (C) A heatmap showing the $\mathrm{R}^2$ of all pairwise comparisons of the methods.
   }
-    \label{fig:all_denoise_reg}
+  \label{fig:figureS1}
 \end{figure}
 
-  \begin{figure}[h!]
+  \begin{figure}[h]
     \centering
-    \includegraphics[width=.9\textwidth]{denoise_network.pdf}
+    \includegraphics[width=0.8\linewidth]{figureS2.pdf}
+    \caption{Results of the mock community analysis. (A) Weighted unifrac (B) Unweighted unifrac}
+    \label{fig:figureS2}
+  \end{figure}
+
+  \begin{figure}[h]
+    \centering
+    \includegraphics[width=0.8\linewidth]{figureS3.pdf}
+    \caption{Results of the synthetic community analysis. (A) Weighted unifrac (B) Unweighted unifrac}
+    \label{fig:figureS3}
+  \end{figure}
+
+  \begin{figure}[h]
+    \centering
+    \includegraphics[width=0.8\linewidth]{pdf/all_denoise_reg.pdf}
+    \caption{All pairwise correlations comparing the similarity between different denoising and clustering methods}
+    \label{fig:figureS4}
+  \end{figure}
+
+  \begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{figureS4.pdf}
     \caption{
-      \textbf{Networks generated using various denoising methods.}
-      The denoising methods used in this result are closed-reference clustering, open-reference clustering, denovo clustering and \ac{dada2}.
-      The database (\ac{gg}) and network inference methods (\ac{sparcc}) were kept constant.
-      (A) The set of interactions predicted by using all the methods (union)
-      (B) The set of interactions that were common to all the methods (intersection)
+      \textbf{(A)} Taxonomy composition of the 20 most abundant genera predicted using different taxonomy references databases: Greengenes, SILVA and NCBI.
+      The legend shows the common and the unique genera among the taxonomy assignments.
   }
-    \label{fig:denoise_network}
+    \label{fig:figureS4}
   \end{figure}
+
+  \begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{figureS5.pdf}
+    \caption{
+      The bray-curtis dissmilarity between the expected taxonomic composition and generated taxonomic composiion for the synthetic datasets.
+  }
+    \label{fig:figureS4}
+  \end{figure}
+
+  \begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{figureS6.pdf}
+    \caption{
+      The effect of OTU processing measurs on the OTU table
+  }
+    \label{fig:figureS4}
+  \end{figure}
+
+  \begin{figure}[h]
+    \centering
+    \includegraphics[width=0.8\linewidth]{png/hamming_distance.png}
+    \caption{The hamming distance between the networks is shown. The similarity between various methods was found to vary with the data-source used.}
+    \label{fig:figureS4}
+  \end{figure}
+
+  \begin{figure}[h]
+    \centering
+    \includegraphics[width=0.8\linewidth]{pdf/denoise_network.pdf}
+    \caption{A network showing union and intersection of networks generated using certain combination of methods}
+    \label{fig:figureS5}
+  \end{figure}

### 2019-07-10 sections/introduction.tex
Change: Remove todo
Update intro (Daniel's corrections)

@@
   Studies have shown the importance of specific microbial interactions in the healthy human microbiome \cite{Lloyd-Price2016} and others have shown how changes in these interactions can lead to dysbiosis \cite{Wang2017,Gilbert2016,Belizario2015}.
-  Co-occurrence relationships in microbial communities are be driven by microbial interactions.
+  While direct measurement of interactions is challenging, a lot of effort has gone into measuring correlations.
+  Co-occurrence relationships in microbial communities may be driven by microbial interactions, although the extent and nature of their relationship is highly debated.
   Each interaction within this complex ecological network can have a positive effect, a negative effect or no effect on the species involved.
@@
   In this study, we propose to build a standardized 16S data-analysis pipeline that will produce robust and reproducible co-occurrence networks from community 16S sequence data.
-\todo[size=\footnotesize]{Mention DB?}
   We do this by conducting a systematic study of the various methods currently available.

### 2019-08-16 sections/results.tex
Add: Update table 1

@@
 
-  We have developed \hl{mindpipe}, a flexible and modular pipeline for 16S data analysis.
+  We have developed \hl{mindpipe}, a flexible and modular pipeline for 16S data analysis, that allows us to infer co-occurence networks from 16S sequencing data.
   It incorporates various popular, publicly available tools as well as custom Python modules and scripts to facilitate inference of co-occurrence networks from 16S data.
+  It allows for using any combination of the incorporated tools for intermediate analyses construct the networks.
   The pipeline workflow is as shown in Figure~\ref{fig:figure1}, it supports the conversion of raw 16S sequence data or counts matrices into co-occurrence networks through multiple methods.
@@
 
+  % TODO: Update this table
   \begin{table}[h]
@@
     \begin{tabular}{|c|c|c|}
-      %\begin{tabularx}{\textwidth}{XXXX}
       \hline
@@
       \hline
-      Denoising and Clustering & Dada2 & ?? \\
-      Taxonomy Assignment & \ac{ncbi} + Blast & ?? \\
-      OTU Processing & Dynamic cutoff & ?? \\
-      Network Inference & Consensus & ?? \\
+      Denoising and Clustering & Dada2/Deblur & default \\
+      Taxonomy Assignment & \ac{ncbi} with Blast & RefSeq database \\
+      OTU Processing & Based on statistical power & Dynamic cutoff \\
+      Network Inference & Consensus method & - \\
       \hline
     \end{tabular}
-    %\end{tabularx}
     \caption{Default tools and parameters for the pipeline}

### 2019-08-22 sections/results.tex
Add: Updates for draft v3

@@
     \end{tabular}
-    \caption{Default tools and parameters for the pipeline}
+    \caption{Default tools and parameters for the pipeline. Detailed information can be found in the methods section}
     \label{tab:default_options}
@@
   This indicates that the taxonomy assigned to the reference sequences drastically alters the co-occurrence network and this change is much more significant than how the reference sequences themselves are identified (in the denoise/cluster step).
+  We believe that this clustering is due to the mislabelling of constitutive taxa that are present in high abundance in the community.
   The residual variation can be seen as an artifact that arises due to the effect on the resultant network when multiple steps are changed at the same time.
   Another interesting observation has been that the dissimilarity between the networks decreases when the low abundance OTUs are removed from the network.
+  These results enable us to believe that the most important criterion for accurate comparative analyses of co-occurrence networks is the taxonomy reference database.
 
@@
   Unlike in the previous sections, where were we evaluated the performance of methods on mock datasets, there are no such datasets that contain a set of known interactions for the evaluation of the network inference algorithms.
-  Therefore, we propose the construction of a \hl{consensus network} (Figure~\ref{fig:figure5}C) involving \ac{magma}, \ac{mldm}, \ac{spieceasi} and \ac{sparcc} by merging the p-values generated from bootstraps of the original taxonomy composition table.
+  Therefore, we propose the construction of a \hl{consensus network} (Figure~\ref{fig:figure5}C) involving \ac{magma}, \ac{mldm}, \ac{spieceasi} and \ac{sparcc} by merging the p-values generated from bootstraps of the original taxonomy composition table using the Browns pvalue combining method (ref).
   The resulting network will have weights reported by all four methods.
@@
 
-  \subsection*{The default network}
+  \subsubsection*{The default pipeline}
 
-  The default tools and parameters chosen for the pipeline are given in Table~\ref{tab:default_options}.
-  The default
+  The systematic analysis performed in the previous sections allows us to suggest the set of tools and parameters that produce more accurate and robust co-occurrence networks.
+  These tools and parameters chosen as the defaults for the pipeline and are given in Table~\ref{tab:default_options}.
+  The recommended tool for the \ac{dc} step are \ac{dada2} or Deblur were chosen based on their accuracy in recapitulating the reference sequences in mock communities and synthetic data.
+  The choice of the taxonomy reference database in the \ac{ta} step is dictated largely by the species expected to be present in the sample as well the database used in similar studies if comparison is a goal.
+  Nevertheless, we suggest \ac{ncbi} RefSeq along with blast+ as the query tool since the database is updated regularly and has a broad collection of taxonomies.
+  The abundance threshold at the \ac{op} step is determined automatically based on the number of samples and the required statistical power.
+  Finally, we use the Browns pvalue combining method on the networks generated using \ac{magma}, \ac{mldm}, \ac{spieceasi} and \ac{sparcc} to obtain a final consensus network in the \ac{ni} step.
 

### 2019-11-25 figures.tex
Change: Update figure sizes and labels
Update figures for final draft

@@
 
-\title{ \huge Microbial Interaction Database pipeline - Figures draft \#4 }
+\title{ \huge Inferring microbial co-occurrence networks from 16S data: A systematic evaluation }
 \author{DK, GB, ZH, CD, KK, DS}
@@
     Each step incorporates several processes, each of which in turn have several alternate algorithms for the same task (indicated by the text to the right of the blue boxes).
+    The text along the arrows describes the data that is being passed from one step to another.
   }
@@
   \caption{
-    \textbf{The choice of taxonomic reference database contributes to the most variance in the networks}.
-    \textbf{(A)} The relative variance in the networks contributed by the denoising, taxonomy assignment and filtering steps.
-    \textbf{(B)} The PCA plot shows that the networks can be clearly separated based on the database used for the taxonomy assignment
+    \textbf{The choice of database contributes to the most variance in the networks}.
+    \textbf{(A)} The relative variance in the networks contributed by the DC, TA and OP steps of the pipeline.
+    \textbf{(B)} All combinations of inferred networks are shown as points on a PCA plot.
+    The color of the points corresponds to the taxonomy database, the shape corresponds to the denoising/clustering method and the size corresponds to whether low abundance OTUs were removed or not.
+    \textbf{(B inset)} The network generated using DC=dada2, TA=Greengenes, OP=off and NI=SPARCC and represents the particular point shown (green big triangle).
+    The plot clearly shows that the points can be separated based on the TA step and that the differences due to the DC and OP steps are not as significant.
   }
@@
   \centering
-  \includegraphics[width=\linewidth]{figure3.pdf}
+  \includegraphics[width=0.8\linewidth]{figure3.pdf}
   \caption{
-    \textbf{The representative sequences generated by the different denoising/clustering methods mostly differ in the sequences that are in low abundance.}
+    \textbf{The representative sequences generated by the different denoising/clustering methods are very similar but differ slightly in the sequences that are in low abundance.}
     \textbf{(A)} The average weighted UniFrac distance between the representative sequences shows that the representative sequences and their compositions are fairly identical between the methods,
-    \textbf{(B)} The largers average unweighted UniFrac distance indicates that methods differ in their identifaction of sequences of low abundance,
-    \textbf{(C)} The distributions of the average unweighted UniFrac distance between the expected sequence profile and the calculated sequence profile in mock datasets show that dada2 and deblur were the best performing methods in most of the datasets.
+    \textbf{(B)} The relatively larger average unweighted UniFrac distance indicates that methods differ in their identification of sequences of low abundance,
+    \textbf{(C)} The distributions of the average unweighted UniFrac distance between the expected sequence profile and the calculated sequence profile in mock datasets show that dada2 and Deblur were the best performing methods in most of the datasets.
   }
@@
   \centering
-  \includegraphics[width=\linewidth]{figure4.pdf}
+  \includegraphics[width=0.75\linewidth]{figure4.pdf}
   \caption{
@@
     \textbf{(A)} The assignment of the top 50 representative sequences to their respective taxonomies using the three different reference databases shows how the same sequences are assigned to different genus.
-    \textbf{(B)} The percentage of OTUs assigned to the same taxonomic label when using different reference databases.
-    The percentage of mis-matches decrease at higher taxonomic levels but even at the Phylum level there exists around 10\% of mis-matches.
-    \textbf{(C)} The bray-curtis dissimilarity between the expected taxonomy profile and calculated taxonomy profile in the mock datasets shows that there is no best choice of database for every dataset.
+    \textbf{(B)} The percentage of \ac{OTU}s assigned to the same taxonomic label when using different reference databases.
+    The percentage of mismatches decrease at higher taxonomic levels but even at the Phylum level there exists around 10\% of mismatches.
+    \textbf{(C)} The Bray-Curtis dissimilarity between the expected taxonomy profile and calculated taxonomy profile in the mock datasets shows that there is no best choice of database for every dataset.
   }
@@
   \caption{
-    \textbf{Networks generated using different network inference methods show marked differences both in terms of edge-density and connectivity}.
-    \textbf{(A)} The six different networks generated by the different network inference methods are very dissimilar
+    \textbf{Networks generated using different network inference methods show notable differences both in terms of edge-density and connectivity}.
+    \textbf{(A)} The six different networks generated by the different network inference methods are very dissimilar.
+    The green links are positive associations and the orange links are negative associations.
+    A threshold of 0.3 was set for the methods that infer pairwise correlations (\ac{sparcc}, Spearman, Pearson) and no threshold was set for the other methods.
     \textbf{(B)} The node overlap upset plot indicates that all the networks have a large number of common nodes involved in connections.
@@
   }
-  \label{fig:figure6}
+  \label{fig:figure5}
 \end{figure}
@@
     \textbf{Network inference and taxonomic assignment have the highest influence on the inferred network structures.}
-    \textbf{(A)} The network constructed using the default pipeline parameters (denosing=dada2, database=ncbi, otu\_filtering=on, network\_inference=sparcc) is compared with networks when one of the steps use a different tool.
-    The common connections are in black, connections unique to the network are colored and connections in the default network and but not present in the current network are green.
+    \textbf{(A)} The network constructed using the default pipeline parameters (DC=\ac{dada2}, TA=\ac{ncbi}, OP=on, NI=\ac{sparcc}) is compared with networks generated when one of the steps use a different tool.
+    The common connections (common with default) are in black, connections unique to the network are colored purple and connections in the default network and but not present in the current network are gray.
     \textbf{(B)} The L1 distance of the networks generated by changing one step of the default pipeline against the network generated using the default parameters.
   }
-  \label{fig:figure1}
-\end{figure}
-
-\FloatBarrier
-
-\subsection*{Supplementary}%
-
-\setcounter{figure}{0}
-\renewcommand{\thefigure}{S\arabic{figure}}
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=0.67\linewidth]{figures1.pdf}
-  \caption{
-    \textbf{Comparison of various denoising and clustering algorithms used in the pipeline}.
-    (A, B) Correlation of the abundances of the taxa that are common between the count matrices created by two different methods.
-    (A) The best correlation (most similar methods) is between open-reference and denovo.
-    (B) The worst correlation (least similar methods) is between open-reference and dada2.
-    (C) A heatmap showing the $\mathrm{R}^2$ of all pairwise comparisons of the methods.
-  }
-  \label{fig:figures1}
+  \label{fig:figure6}
 \end{figure}
@@
   \centering
-  \includegraphics[width=0.8\linewidth]{figures2.pdf}
-  \caption{Results of the mock community analysis. (A) Weighted unifrac (B) Unweighted unifrac}
-  \label{fig:figures2}
+  \includegraphics[width=17cm]{figure7.pdf}
+  \caption{
+    The consensus networks generated using the default pipeline settings for the Hard Palate \textbf{(A)} and Stool \textbf{(B)} body-sites.
+}
+  \label{fig:figure7}
 \end{figure}
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=0.8\linewidth]{figures3.pdf}
-  \caption{Results of the synthetic community analysis. (A) Weighted unifrac (B) Unweighted unifrac}
-  \label{fig:figures3}
-\end{figure}
+% \FloatBarrier
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=0.8\linewidth]{pdf/all_denoise_reg.pdf}
-  \caption{All pairwise correlations comparing the similarity between different denoising and clustering methods}
-  \label{fig:figures4}
-\end{figure}
+% \subsection*{Supplementary}%
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=\linewidth]{figures4.pdf}
-  \caption{
-    \textbf{(A)} Taxonomy composition of the 20 most abundant genera predicted using different taxonomy references databases: Greengenes, SILVA and NCBI.
-    The legend shows the common and the unique genera among the taxonomy assignments.
-}
-  \label{fig:figures4}
-\end{figure}
+% \setcounter{figure}{0}
+% \renewcommand{\thefigure}{S\arabic{figure}}
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=\linewidth]{figures5.pdf}
-  \caption{
-    The bray-curtis dissmilarity between the expected taxonomic composition and generated taxonomic composiion for the synthetic datasets.
-}
-  \label{fig:figures4}
-\end{figure}
+% \begin{figure}[h]
+%   \centering
+%   \includegraphics[width=0.67\linewidth]{figureS1.pdf}
+%   \caption{
+%     \textbf{Comparison of various denoising and clustering algorithms used in the pipeline}.
+%     (A, B) Correlation of the abundances of the taxa that are common between the count matrices created by two different methods.
+%     (A) The best correlation (most similar methods) is between open-reference and denovo.
+%     (B) The worst correlation (least similar methods) is between open-reference and dada2.
+%     (C) A heatmap showing the $\mathrm{R}^2$ of all pairwise comparisons of the methods.
+%   }
+%   \label{fig:figureS1}
+% \end{figure}
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=\linewidth]{figures6.pdf}
-  \caption{
-    The effect of OTU processing measurs on the OTU table
-}
-  \label{fig:figures4}
-\end{figure}
+% \begin{figure}[h]
+%   \centering
+%   \includegraphics[width=0.8\linewidth]{figureS2.pdf}
+%   \caption{Results of the mock community analysis. (A) Weighted unifrac (B) Unweighted unifrac}
+%   \label{fig:figureS2}
+% \end{figure}
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=0.8\linewidth]{png/hamming_distance.png}
-  \caption{The hamming distance between the networks is shown. The similarity between various methods was found to vary with the data-source used.}
-  \label{fig:figures4}
-\end{figure}
+% \begin{figure}[h]
+%   \centering
+%   \includegraphics[width=0.8\linewidth]{figureS3.pdf}
+%   \caption{Results of the synthetic community analysis. (A) Weighted unifrac (B) Unweighted unifrac}
+%   \label{fig:figureS3}
+% \end{figure}
 
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=0.8\linewidth]{pdf/denoise_network.pdf}
-  \caption{A network showing union and intersection of networks generated using certain combination of methods}
-  \label{fig:figures5}
-\end{figure}
+% \begin{figure}[h]
+%   \centering
+%   \includegraphics[width=0.8\linewidth]{pdf/all_denoise_reg.pdf}
+%   \caption{All pairwise correlations comparing the similarity between different denoising and clustering methods}
+%   \label{fig:figureS4}
+% \end{figure}
+
+% \begin{figure}[h]
+%   \centering
+%   \includegraphics[width=\linewidth]{figureS4.pdf}
+%   \caption{
+%     \textbf{(A)} Taxonomy composition of the 20 most abundant genera predicted using different taxonomy references databases: Greengenes, SILVA and NCBI.
+%     The legend shows the common and the unique genera among the taxonomy assignments.
+% }
+%   \label{fig:figureS4}
+% \end{figure}
+
+% \begin{figure}[h]
+%   \centering
+%   \includegraphics[width=\linewidth]{figureS5.pdf}
+%   \caption{
+%     The bray-curtis dissmilarity between the expected taxonomic composition and generated taxonomic composiion for the synthetic datasets.
+% }
+%   \label{fig:figureS4}
+% \end{figure}
+
+% \begin{figure}[h]
+%   \centering
+%   \includegraphics[width=\linewidth]{figureS6.pdf}
+%   \caption{
+%     The effect of OTU processing measurs on the OTU table
+% }
+%   \label{fig:figureS4}
+% \end{figure}
+
+% \begin{figure}[h]
+%   \centering
+%   \includegraphics[width=0.8\linewidth]{png/hamming_distance.png}
+%   \caption{The hamming distance between the networks is shown. The similarity between various methods was found to vary with the data-source used.}
+%   \label{fig:figureS4}
+% \end{figure}
+
+% \begin{figure}[h]
+%   \centering
+%   \includegraphics[width=0.8\linewidth]{pdf/denoise_network.pdf}
+%   \caption{A network showing union and intersection of networks generated using certain combination of methods}
+%   \label{fig:figureS5}
+% \end{figure}
 

### 2019-11-25 scripts/figure2/figure2a_plot.R
Update scripts for final draft

@@ -1,0 +1,21 @@
+#!/usr/bin/env Rscript
+
+library(tidyverse)
+library(ggplot2)
+library(ggpubr)
+
+
+group <- c("DC", "TA", "OP", "misc")
+percentage <- c(24.90, 17.97, 5.16, 51.96)
+labels <- paste0(group, " (", percentage, "%)")
+df <- data.frame(group=group, percentage=percentage, labels=labels)
+
+pie_chart <- ggpie(
+                   df,
+                   "percentage",
+                   label="labels",
+                   fill="group",
+                   palette="Set2",
+                   lab.pos="in",lab.font="black"
+)
+ggarrange(pie_chart, labels=c("A"), nrow=1, ncol=1, common.legend=TRUE, legend="right")

### 2019-11-25 scripts/figure2/figure2b_data.py

@@ -1,0 +1,23 @@
+#!/usr/bin/env python3
+
+import click
+import pandas as pd
+from sklearn.decomposition import PCA
+
+
+def calculate_pca(table: pd.DataFrame, components: int = 2):
+    X = table.values.T
+    pca = PCA(n_components=components)
+    X_pca = pca.fit_transform(X)
+    return pca, X_pca
+
+
+@click.command()
+@click.option("--corr_table", help="The file containing the correlation table")
+def main(corr_table):
+    table = pd.read_table(corr_table, index_col=0)
+    pca, X_pca = calculate_pca(table)
+
+
+if __name__ == "__main__":
+    main()

### 2019-11-25 scripts/figure2/figure2b_plot.R

@@ -1,0 +1,26 @@
+#!/usr/bin/env Rscript
+
+library(titdyverse)
+library(ggplot2)
+library(ggpubr)
+
+plot_pca <- function(pca_data) {
+    pca_plot <- ggscatter(pca_data,
+        x="pc1",
+        y="pc2",
+        fill="TA",
+        shape="DC",
+        size="OP",
+        xlab="PC1",
+        ylab="PC2",
+        legend="right",
+    ) +
+    scale_shape_manual(values=c(21, 22, 23, 24, 25)) +
+    scale_size_manual(values=c("no"=5, "yes"=2.5)) +
+    guides(fill=guide_legend(override.aes=list(shape=21)))
+    ggarrange(pca_plot, labels=c("B"), nrow=1, ncol=1, common.legend=TRUE, legend="right")
+}
+
+
+pca_data <- read.table("pca_2comp.csv", sep=",", header=TRUE)
+plot_pca(pca_data)

### 2019-11-25 scripts/figure3/figure3ab_plot.R

@@ -36,7 +36,8 @@ plot_heatmap <- function(data, title) {
             axis.title=element_blank(),
             axis.text.x=element_text(angle=30, hjust=1),
             axis.text.y=element_text(angle=30)
-        )
+        ) +
+        coord_fixed()
 }
 
 unweighted <- read.csv("fmt_unweighted_unifrac.csv", sep=",", header=TRUE)

### 2019-11-25 scripts/figure3/move_otus.py

@@ -18,7 +18,7 @@ MAP = {
 
 @click.command()
 @click.option("--base_dir", help="Location of the directory containing the otus")
-@click.option("--output_dir", help="Location of the output directory")
+@click.option("--output_dir", default=".", help="Location of the output directory")
 def main(base_dir: str, output_dir: str):
     base_path = pathlib.Path(base_dir)
     output_path = pathlib.Path(output_dir)

### 2019-11-25 scripts/figure4/figure4a_plot.R

@@ -8,16 +8,21 @@ library(ggpubr)
 
 notu <- 50
 
-gg <- read.csv("gg.csv", header=TRUE, sep=",", na.strings="")[1:notu,]
-silva <- read.csv("silva.csv", header=TRUE, sep=",", na.strings="")[1:notu,]
-ncbi <- read.csv("ncbi.csv", header=TRUE, sep=",", na.strings="")[1:notu,]
+gg <- read.csv("fmt_gg.csv", header=TRUE, sep=",", na.strings="")[1:notu,]
+silva <- read.csv("fmt_silva.csv", header=TRUE, sep=",", na.strings="")[1:notu,]
+ncbi <- read.csv("fmt_ncbi.csv", header=TRUE, sep=",", na.strings="")[1:notu,]
 
 make_alluvial_plot <- function(db_data, title) {
+    genus_factors <- levels(factor(db_data$Genus))
+    n_genus <- length(genus_factors)
+    palette <- get_palette(, n_genus)
+    palette[n_genus] <- "#d3d3d3"
     alluvial_plot <- ggplot(data=db_data, aes(axis1=OTU, axis2=Genus,  y=Abundance)) +
         scale_x_discrete(limits=c("OTU", "Genus"), expand=c(.1, .05)) +
         xlab("Tax") +
         geom_alluvium(aes(fill=Genus), width=1/12) +
         geom_stratum(width=1/4, aes(fill=Genus)) +
+        scale_fill_manual(values=palette) +
         scale_linetype_manual(values=c("blank", "solid")) +
         # geom_text_repel(aes(label=Genus), stat="stratum", size=2, direction="y", nudge_x=0.5) +
         ggtitle(title) +
@@ -42,7 +47,7 @@ ncbi$OTU <- ncbi$empty
 ncbi_plot <- ncbi_plot +
     geom_text_repel(stat="stratum", label.strata=TRUE, data=ncbi[,c("OTU", "Genus", "Abundance")], nudge_x=1)
 
-ggarrange(
+final_plot <- ggarrange(
     gg_plot,
     silva_plot + rremove("y.axis") + rremove("y.text") + rremove("y.ticks") + rremove("y.title"),
     ncbi_plot + rremove("y.axis") + rremove("y.text") + rremove("y.ticks") + rremove("y.title"),
@@ -51,3 +56,4 @@ ggarrange(
     common.legend=TRUE,
     legend="bottom"
 )
[…]

### 2019-11-25 scripts/figure6/figure6_data.py

@@ -75,6 +75,8 @@ def write_networks(file_paths, default_file, multigraph, color_key_level, output
                         colorkey=colorkey,
                         layer="background",
                     )
+                else:
+                    G.nodes[name]["layer"] = "common"
             for source, target, edge_data in default_graph.edges(data=True):
                 source_name, target_name = id_name_map[source], id_name_map[target]
                 weight = edge_data["weight"]
@@ -87,6 +89,9 @@ def write_networks(file_paths, default_file, multigraph, color_key_level, output
                         pvalue=pvalue,
                         layer="background",
                     )
+                else:
+                    G.edges[(source_name, target_name)]["layer"] = "common"
+                    G.edges[(source_name, target_name)]["weight"] = edge_data["weight"]
         nx.write_gml(G, str(output_path / f"{network_name}.gml"))
     nx.write_gml(combined_graph, str(output_path / "combined.gml"))
     return list(combined_graph.nodes), list(combined_graph.edges)

### 2019-11-25 scripts/figure6/figure6a_plot.R

@@ -15,7 +15,7 @@ get_edgestyle <- function(x) {
 }
 
 
-plot_network <- function(network_file, combined_layout, interaction_threshold, title, edgecolor) {
+plot_network <- function(network_file, combined_layout, interaction_threshold, title, palette, ind) {
   graph_raw <- as_tbl_graph(read_graph(network_file, format="gml"))
   graph <- graph_raw %>%
     activate(edges) %>%
@@ -33,9 +33,10 @@ plot_network <- function(network_file, combined_layout, interaction_threshold, t
     geom_edge_arc(aes(color=factor(layer), edge_linetype=factor(color), edge_alpha=factor(layer))) +
     # geom_node_point(aes(color=factor(colorkey))) +
     geom_node_point() +
-    scale_edge_color_manual(values = c(background="#d3d3d3", foreground=edgecolor)) +
+    #scale_edge_color_manual(values = c(background=palette[[1]], foreground=palette[[ind]], common="black")) +
+    scale_edge_color_manual(values = c(foreground="#7570B3", background="#c1c1c1", common="black")) +
     scale_edge_linetype_manual(values = c(negative="dashed", positive="solid")) +
-    scale_edge_alpha_manual(values = c(foreground=0.6, background=0.7)) +
+    scale_edge_alpha_manual(values = c(foreground=0.4, background=0.4, common=0.7)) +
     labs(title=title) +
     coord_fixed() +
     theme_bw() +
@@ -55,13 +56,15 @@ plot_network <- function(network_file, combined_layout, interaction_threshold, t
 combined_graph <- read_graph("combined.gml", format="gml")
 combined_layout <- create_layout(graph=combined_graph, layout="linear", circular=TRUE)
 
-palette <- brewer.pal(n=6, name="Dark2")
-default_plot <- plot_network("default.gml", combined_layout, 0.3, "default", palette[[1]])
-database_plot <- plot_network("gg.gml", combined_layout, 0.3, "database=gg", palette[[2]])
-clustering_plot <- plot_network("open_ref.gml", combined_layout, 0.3, "clustering=open_ref", palette[[3]])
-otu_filtering_plot <- plot_network("otu_filtering.gml", combined_layout, 0.3, "otu_filtering=off", palette[[4]])
[…]

### 2019-11-25 scripts/figure7/figure7_plot.R

@@ -6,6 +6,9 @@ library(ggraph)
 library(tidygraph)
 library(ggpubr)
 
+# library(dplyr)
+# library(pracma)
+
 get_edgecolor <- function(x) {
   sapply(x, function(y) if(y > 0) "positive" else "negative")
 }
@@ -15,10 +18,11 @@ get_edgestyle <- function(x) {
 }
 
 
-plot_network <- function(network_file, combined_layout, interaction_threshold, title, edgecolor) {
+plot_network <- function(network_file, combined_layout, interaction_threshold, pvalue_threshold, title, edgecolor) {
   graph_raw <- as_tbl_graph(read_graph(network_file, format="gml"))
   graph <- graph_raw %>%
     activate(edges) %>%
+    filter(pvalue < pvalue_threshold) %>%
     filter(abs(weight) > interaction_threshold) %>%
     mutate(color=get_edgecolor(weight)) %>%
     activate(nodes) %>%
@@ -29,10 +33,20 @@ plot_network <- function(network_file, combined_layout, interaction_threshold, t
   graph_layout <- data.frame(temp_layout)
   graph_layout$x <- combined_layout[match_inds,]$x
   graph_layout$y <- combined_layout[match_inds,]$y
+  lo <- data.matrix(graph_layout[, c("x", "y")])
+  angle <- as_tibble(cart2pol(lo)) %>% mutate(degree=phi*180/phi)
   graph_plot <- ggraph(graph=graph, layout="manual", node.positions=graph_layout, circular=TRUE) +
     geom_edge_arc(aes(color=color, edge_linetype=color, edge_alpha=color)) +
     # geom_node_point(aes(color=factor(colorkey))) +
     geom_node_point() +
+    geom_node_text(aes(label = name), size=2, repel=TRUE, angle=0) +
+      # geom_node_text(aes(label = name), 
+      #                size = 2, 
+      #                hjust = ifelse(lo[,1] > 0, -0.2, 1.2),
+      #                angle = case_when(lo[,2] > 0 & lo[,1] > 0 ~ angle$degree, 
+      #                                  lo[,2] < 0 & lo[,1] > 0 ~ angle$degree,
+      #                                  lo[,1] == 1 ~angle$degree,
+      #                                  TRUE ~ angle$degree - 180)) +
     scale_edge_color_manual(values = c(negative="#d95f02", positive="#1b9e77")) +
[…]

### 2019-11-25 sections/introduction.tex
Update introduction based on PI comments

@@
 
-  Microbiomes are ubiquitous and play an important role in biogeochemical transformations in both natural as well as man-made ecosystems.
-  These microbiomes comprise several thousand different microbes interacting with each other and their environment, often through intricate metabolic and signaling relationships.
-  The changes in the microbiome can impact the health of the environmental niche in which they reside \cite{HumanMicrobiomeProjectConsortium2012,Lloyd-Price2016} and this can have major implications in health of the human body.
-  A mechanistic understanding of the key interactions driving the microbe-microbe and microbe-environment dynamics is necessary in order to analyze the dysbiosis in a healthy microbiome and in the design of modifications to these correct these microbiomes.
+  Microbial communities are ubiquitous and play an important role in biogeochemical transformations in both natural as well as man-made ecosystems.
+  These microbiomes comprise several thousand different microbial strains interacting with each other and their environment, often through intricate metabolic and signaling relationships.
+  The changes in the microbiome can impact the environmental niche in which they reside \cite{HumanMicrobiomeProjectConsortium2012,Lloyd-Price2016} and for host-associated microbiomes this can have major implications in health of the human body.
+  A mechanistic understanding of the key interactions between microbes is necessary in order to analyze the dysbiosis in a healthy microbiome and the design interventions.
 
-  Studies have shown the importance of specific microbial interactions in the healthy human microbiome \cite{Lloyd-Price2016} and others have shown how changes in these interactions can lead to dysbiosis \cite{Wang2017,Gilbert2016,Belizario2015}.
-  While direct measurement of interactions is challenging, a lot of effort has gone into measuring correlations.
+  Studies have shown the importance of specific microbial interactions in the healthy microbiome \cite{Lloyd-Price2016} and others have shown how changes in these interactions can lead to dysbiosis \cite{Wang2017,Gilbert2016,Belizario2015}.
+  While direct measurement of interactions is challenging, a significant effort has gone into measuring correlations leading to the correlations between microbial abundances being readily available(?)
   Co-occurrence relationships in microbial communities may be driven by microbial interactions, although the extent and nature of their relationship is highly debated.

### 2019-11-25 sections/coverpage.tex
Update coverpage for final draft

--- 2018-08-08 sections/coverpage.tex
@@
   \vspace*{\fill}
-  { \huge \bfseries Correlating Microbial Abundances for the Microbial Interaction Network Database}\\[0.4cm] % Title of your document
+  {\huge Inferring microbial co-occurrence networks from 16S data: A systematic evaluation}\\[0.4cm]
+  % {\huge  2. Investigating the best practices for inference of microbial co-occurrence networks from 16S data}\\[0.4cm] % Title of your document
+  % {\huge  3. Attempting to find the best practice pipeline for inferring co-occurrence networks from 16S data}\\[0.4cm] % Title of your document
+  % {\huge  4. Deciphering the complexities in co-occurrence network inference from 16S data}\\[0.4cm] % Title of your document
 
@@
 
-  % \vspace{1.5 cm}
-  % Dileep Kishore\textsuperscript{1},
-  % Gabriel Birzu\textsuperscript{2},
-  % Zhenjun Hu\textsuperscript{1},
-  % Charles DeLisi\textsuperscript{1,3},
-  % Kirill Korolev\textsuperscript{1,2},
-  % Daniel Segr\`{e}\textsuperscript{1,3,4}\\
-  % \vspace{1cm}
-  % \textsuperscript{1}Bioinformatics Program, Boston University, Boston, Massachusetts, USA\\
-  % \textsuperscript{2}Department of Physics, Boston University, Boston, Massachusetts, USA\\
-  % \textsuperscript{3}Department of Biomedical Engineering, Boston University, Boston, Massachusetts, USA\\
-  % \textsuperscript{4}Department of Biology, Boston University, Boston, Massachusetts, USA\\
+  \vspace{1.5 cm}
+  Dileep Kishore\textsuperscript{1},
+  Gabriel Birzu\textsuperscript{2},
+  Zhenjun Hu\textsuperscript{1},
+  Charles DeLisi\textsuperscript{1,3},
+  Kirill Korolev\textsuperscript{1,2},\\
+  Daniel Segr\`{e}\textsuperscript{1,3,4}\\
+  \vspace{1cm}
+  \textsuperscript{1}Bioinformatics Program, Boston University, Boston, Massachusetts, USA\\
+  \textsuperscript{2}Department of Physics, Boston University, Boston, Massachusetts, USA\\
+  \textsuperscript{3}Department of Biomedical Engineering, Boston University, Boston, Massachusetts, USA\\
+  \textsuperscript{4}Department of Biology, Boston University, Boston, Massachusetts, USA\\
 

### 2019-11-25 sections/abstract.tex
Update abstract for final draft

--- 2019-04-16 sections/abstract.tex
@@
     Microbes tend to organize into communities consisting of hundreds of species involved in complex interactions with each other.
-    16S ribosomal RNA (16S rRNA) gene profiling provides snapshots that reveal the phylogenies and abundance distributions of these microbial communities.
-    These snapshots, when collected from multiple samples, have the potential to reveal which microbes co-occur, providing a glimpse into the network of inter-dependencies underlying these communities.
-    The inference of networks from 16S data is prone to statistical artifacts, but the extent to which the different steps of the workflow affect the resultant network is still unclear.
+    16S ribosomal RNA (16S rRNA) gene profiling provides snapshots that reveal the phylogenies and abundance profiles of these microbial communities.
+    These snapshots, when collected from multiple samples, have the potential to reveal which microbes co-occur, providing a glimpse into the network of influences in these communities.
+    The inference of networks from 16S data is prone to statistical artifacts.
+    There are many tools for performing each step of the 16S analysis workflow and the results generated by different tools are varied, but the extent to which these steps affect the resultant network is still unclear.
     In this study, we perform a meticulous analysis of each step of a pipeline that processes 16S sequencing data into a network of microbial associations.
-    Through this process, we determine the tools and parameters that generate the most accurate and robust co-occurrence.
+    Through this process, we determine the tools and parameters that generate the most accurate and robust co-occurrence networks.
     Ultimately, we develop a standardized pipeline that follows these default tools and parameters, but that can also help explore the outcome of any other combination of choices.
     We envisage that this standard pipeline for processing 16S sequencing data into networks of microbial co-occurrences could be used for integrating multiple data-sets, and for generating comparative analyses and consensus networks useful for detecting disease-related patterns.
-\todo[size=\footnotesize]{NOTE: Abstract constrained to 200 words} \\
-    The pipeline is available as a Python package and a docker container.
+\\
+    The pipeline is available as a Python package at \href{https://github.com/dileep-kishore/mindpipe}{https://github.com/dileep-kishore/mindpipe}
   }

### 2019-11-25 sections/introduction.tex
Update introduction for final draft

@@
   These microbiomes comprise several thousand different microbial strains interacting with each other and their environment, often through intricate metabolic and signaling relationships.
-  The changes in the microbiome can impact the environmental niche in which they reside \cite{HumanMicrobiomeProjectConsortium2012,Lloyd-Price2016} and for host-associated microbiomes this can have major implications in health of the human body.
-  A mechanistic understanding of the key interactions between microbes is necessary in order to analyze the dysbiosis in a healthy microbiome and the design interventions.
+  Changes in the microbiome can alter the environment in which they reside \cite{HumanMicrobiomeProjectConsortium2012,Lloyd-Price2016} and for host-associated microbiomes this can have major implications in health of the human body.
+  A mechanistic understanding of the key interactions between microbes is necessary in order to analyze the dysbiosis in a healthy microbiome and to design interventions.
 
   Studies have shown the importance of specific microbial interactions in the healthy microbiome \cite{Lloyd-Price2016} and others have shown how changes in these interactions can lead to dysbiosis \cite{Wang2017,Gilbert2016,Belizario2015}.
-  While direct measurement of interactions is challenging, a significant effort has gone into measuring correlations leading to the correlations between microbial abundances being readily available(?)
-  Co-occurrence relationships in microbial communities may be driven by microbial interactions, although the extent and nature of their relationship is highly debated.
+  While direct measurement of interactions is challenging, a significant amount of effort has gone into measuring correlations leading to the correlation information between microbial abundances being readily available.
+  Co-occurrence relationships in microbial communities may be driven by microbial interactions, although the extent and nature of these relationships is highly debated \cite{Zuniga2017}.
   Each interaction within this complex ecological network can have a positive effect, a negative effect or no effect on the species involved.
   Bacteria that compete for the same nutrient will have a negative co-occurrence relationship \cite{Ghoul2016} whereas cross-feeding between species leads to a positive co-occurrence relationship \cite{DSouza2018}.
+  But these co-occurrence relationships are not very straightforward, since, similar species competing for a resource could co-occur if the resource is abundant.
   The microbial interaction networks are highly dynamic, and they constantly reorganize in a varying environment.
@@
   With the advancement in DNA sequencing technologies \cite{Narihiro2017} and data processing methods much more information can be extracted from these microbial community samples than ever before.
-  These advances have led to large-scale data collection efforts involving soil (Earth Microbiome Project) \cite{Thompson2017}, marine (Tara Oceans Project) \cite{Zhang2015} and human-associated microbiota (Human Microbiome Project) \cite{HumanMicrobiomeProjectConsortium2012} and thus led the generation of a wealth of data about these microbial communities.
-  These metagenomic data-sets can be used to infer networks of co-occurrence relationships between microorganisms in the samples.
+  These advances have led to large-scale data collection efforts involving environmental (Earth Microbiome Project) \cite{Thompson2017}, marine (Tara Oceans Project) \cite{Zhang2015} and human-associated microbiota (Human Microbiome Project) \cite{HumanMicrobiomeProjectConsortium2012} and thus led the generation of a wealth of data about these microbial communities.
+  These data-sets can be used to infer networks of co-occurrence relationships between microorganisms in the samples.
   The co-occurrence networks can in turn be used to identify important interactions driving the dynamics of the microbial community, as well as, analyze emergent properties such as robustness and modularity.
-  However, 16S sequencing data is limited by factors such as resolution, sequencing depth, compositional nature, sequencing errors and copy number variations.
-  To counter these limitations and remove statistical artifacts, many different analysis methods have been developed \cite{Callahan2016,Amir2017,Friedman2012,Kurtz2015}.
-  But, the different methods and tools developed to solve these issues infer vastly different community compositions and co-occurrence networks \cite{Golob2017,Weiss2016}.
 
-  The methodology for the analysis of 16S rRNA sequence data is not very straightforward and there are a myriad of tools for each step of the process, making the analysis complicated and not very reproducible.
+  16S sequencing data is limited by many factors such as resolution, sequencing depth, compositional nature, sequencing errors and copy number variations.
+  Many different analysis methods have been developed \cite{Callahan2016,Amir2017,Friedman2012,Kurtz2015} to counter these limitations and remove statistical artifacts, leading to the existence of a myriad of tools for each step of the data analysis workflow.
+  The different methods and tools developed to solve these issues infer vastly different community compositions and co-occurrence networks \cite{Golob2017,Weiss2016}, making the analysis unreproducible and comparisons convoluted.
   But because there does not exist a dataset wherein the microbial interactions have been fully characterized, there is no straightforward way to determine which set of tools or methods generate the most accurate networks.
@@
   We believe that these steps will ensure better reproducibility and easier comparison of networks across data-sets.
+  We call our pipeline mindpipe, which derives its origin from the \ac{mind}.

### 2019-11-25 sections/results.tex
Update results and figures for final draft

@@
 
-  \subsection*{mindpipe}
+  \subsection*{Mindpipe}
 
-  We have developed \hl{mindpipe}, a flexible and modular pipeline for 16S data analysis, that allows us to infer co-occurence networks from 16S sequencing data.
+  We have developed mindpipe, a flexible and modular pipeline for 16S data analysis, that allows us to infer co-occurence networks from 16S sequencing data.
   It incorporates various popular, publicly available tools as well as custom Python modules and scripts to facilitate inference of co-occurrence networks from 16S data.
-  It allows for using any combination of the incorporated tools for intermediate analyses construct the networks.
-  The pipeline workflow is as shown in Figure~\ref{fig:figure1}, it supports the conversion of raw 16S sequence data or counts matrices into co-occurrence networks through multiple methods.
+  It enables the user to process 16S sequence data or count matrices using different combinations of the incorporated tools for intermediate analyses as well as inference of co-occurrence networks.
+  The pipeline workflow is shown in Figure~\ref{fig:figure1}.
   Each process in the pipeline supports alternate tools for performing the same task, users can use the configuration file to change these values.
-  We perform a systematic analysis of each step of the pipeline to determine the set of default options for the pipeline (Table~\ref{tab:default_options}).
+  We perform a systematic analysis of each step of the pipeline to determine the set of default options for the pipeline.
   We also provide the users with a set of guidelines to facilitate tool selection as appropriate for their data.
-
-  % TODO: Update this table
-  \begin{table}[h]
-    \centering
-    \small
-    \begin{tabular}{|c|c|c|}
-      \hline
-      \textbf{Process} & \textbf{Tool} & \textbf{Parameters} \\
-      \hline
-      Denoising and Clustering & Dada2/Deblur & default \\
-      Taxonomy Assignment & \ac{ncbi} with Blast & RefSeq database \\
-      OTU Processing & Based on statistical power & Dynamic cutoff \\
-      Network Inference & Consensus method & - \\
-      \hline
-    \end{tabular}
-    \caption{Default tools and parameters for the pipeline. Detailed information can be found in the methods section}
-    \label{tab:default_options}
-  \end{table}
 
@@
   In order to objectively determine the best method we use both mock data from mockrobiota~\cite{Bokulich2016} as well as, synthetically generated reads from an Illumina read simulator called ART~\cite{Huang2012}.
-  The mock datasets selected were mock4, mock12 and mock16 since they had expected sequences.
-  The synthetic reads were simulated using the three different taxonomy distribution profiles generated from soil, water and hl{risk} datasets.
-  Reference sequences were generated using \ac{ncbi} and the \hl{decard} tool \cite{Golob2017} for the required taxonomy profiles.
+  The mock datasets selected were labelled mock4, mock12 and mock16 in the mockrobiota database and consist of sequenced 16S data from mock communities of various bacteria.
+  These datasets were selected since they contained the expected compositions along with the reference sequences for the organisms in the mock community.
+  The synthetic reads were simulated using the three different taxonomy distribution profiles obtained from soil, water and Crohn's disease datasets.
+  Reference sequences were generated using \ac{ncbi} and the decard tool \cite{Golob2017} for these taxonomy profiles.
   Detailed information on the mock communities and all the other settings used to generate the sequence data are provided in the methods section.
@@
 
-  \subsection*{The choice of reference database and network inference algorithm have the biggest impact on inferred networks}
+  \subsection*{The choice of reference database has the biggest impact on inferred networks}
 
   In order to analyze the effect of different statistical methods on the inferred co-occurrence networks, we generate co-occurrence networks using all possible combinations of methods and estimate the variability in the networks due to each choice.
+  This analysis is performed while keeping the network inference algorithm the same throughout the analysis.
+  The effects of various steps on the final co-occurrence network is estimated by building a linear model of the edges of the network as a function the various step in the analysis pipeline.
   Figure \ref{fig:figure2} shows the fraction of total variation among the co-occurrence networks as contributed by each of the four main steps of the pipeline.
-  The 16S reference database contributes to the most variation ($\sim25\%$) in the networks and the networks can be clearly separated based on the database used.
-  \todo[size=\footnotesize]{Update with full results}
+  The 16S reference database contributes to the most variation ($\sim25\%$) in the networks, and the networks can be clearly separated based on the database used.
   This indicates that the taxonomy assigned to the reference sequences drastically alters the co-occurrence network and this change is much more significant than how the reference sequences themselves are identified (in the denoise/cluster step).
@@
       The color of the points corresponds to the taxonomy database, the shape corresponds to the denoising/clustering method and the size corresponds to whether low abundance OTUs were removed or not.
+      \textbf{(B inset)} The network generated using DC=dada2, TA=GG, OP=no and NI=SPARCC and represents the particular point shown (big red square).
       The plot clearly shows that the points can be separated based on the TA step and that the differences due to the DC and OP steps are not as significant.
@@
 
-  \subsection*{Denoising and clustering methods differ in their identificaiton of less common reference sequences}
+  \subsection*{Denoising and clustering methods differ in their identification of less common reference sequences}
 
-  There is good agreement in the \ac{otu}/\ac{esv} tables when different combinations of methods are used to generate them.
-  We process the same 16S sequencing reads (\hl{FMT} dataset) using 5 different methods \- open-reference clustering, closed-reference clustering, denovo clustering, \ac{dada2}~\cite{Callahan2016} and Deblur~\cite{Amir2017}.
+  Denoising and clustering are commonly carried out to generate representative sequences from the raw 16S sequencing data and to obtain the \ac{otu}/\ac{esv} tables (counts of these representative sequences for each sample).
+  In order to compare the \ac{otu} tables generated by these tools we process the same 16S sequencing reads (\hl{FMT} dataset) using 5 different methods \- open-reference clustering, closed-reference clustering, denovo clustering, \ac{dada2}~\cite{Callahan2016} and Deblur~\cite{Amir2017}.
   The first three methods are from the \ac{qiime1}~\cite{Caporaso2010} package.
+  We find that there is good agreement in the \ac{otu}/\ac{esv} tables when different combinations of methods are used to generate them.
 
@@
   Figure~\ref{fig:figure3}B on the other hand shows an increase in dissimilarity between each pair of methods suggesting that the methods might differ in the sequences of low abundance.
-  In order to verify this claim, we changed the threshold for the sequence count and recalculated the weighted and unweighted UniFrac distance metrics and also looked at the number of representative sequences identified in each case.
-\todo[size=\footnotesize]{In supplementary}
+  In order to verify this claim, we changed the threshold for the sequence count and recalculated the weighted and unweighted UniFrac distance metrics and also looked at the number of representative sequences identified in each case (in the supplementary section).
 
   %% TODO: Check if this is necessary
-  \textcolor{gray}{
-    For each of these methods we use the same taxonomy database (\ac{gg}) to assign taxonomies to the representative sequences.
-    We then correlate the abundances of matching taxonomies between every combination of a pair of methods (Figure \ref{fig:all_denoise_reg}).
-    The \ac{esv} tables generated by methods that perform denoising are very similar to each other $\sim0.9$ and the \ac{otu} tables generated by the clustering methods are very similar to each other $\sim0.9$, but results of denoising and clustering are highly uncorrelated with each other $\sim0.4$ (Figure \ref{fig:figureS2}).
-    The common core network (network made up of edges that are predicted by all methods) is very small compared to the total number of edges predicted by all the methods (Figure \ref{fig:denoise_network}).
-    In these results, only the denoising method was changed (database used: \ac{gg} and network inference: \ac{sparcc}).
-  }
+    % For each of these methods we use the same taxonomy database (\ac{gg}) to assign taxonomies to the representative sequences.
+    % We then correlate the abundances of matching taxonomies between every combination of a pair of methods (Figure \ref{fig:all_denoise_reg}).
+    % The \ac{esv} tables generated by methods that perform denoising are very similar to each other $\sim0.9$ and the \ac{otu} tables generated by the clustering methods are very similar to each other $\sim0.9$, but results of denoising and clustering are highly uncorrelated with each other $\sim0.4$ (Figure \ref{fig:figureS2}).
+    % The common core network (network made up of edges that are predicted by all methods) is very small compared to the total number of edges predicted by all the methods (Figure \ref{fig:denoise_network}).
+    % In these results, only the denoising method was changed (database used: \ac{gg} and network inference: \ac{sparcc}).
 
   These comparisons only elucidate the pairwise similarity or dissimilarity of a pair of methods.
-  In order to determine the tool that most accurately recapitulates the reference sequences in the samples we the mock datasets.
-  The mock community datasets we processed by the various methods and the predicted representative sequences were compared with expected representative sequences and their distribution.
+  In order to determine the tool that most accurately recapitulates the reference sequences in the samples we use the 16S sequences from the mock datasets.
+  The mock community datasets were processed using all the methods and the predicted representative sequences were compared with expected representative sequences and their distribution.
   The results (Figure~\ref{fig:figure3}C and D) show that the performance of the methods are similar to each other as in the case of the real dataset, but, are very different from the expected sequence distribution.
@@
   There is no method that clearly outperforms the rest in all datasets.
-  Based on their slightly better performance on the mock datasets, their \textit{de novo} error correcting nature and other previous studies (\hl{refs}) we recommend either \ac{dada2} or Deblur for users.
+  Based on their slightly better performance on the mock datasets, their \textit{de novo} error correcting nature and other previous studies [refs] we recommend either \ac{dada2} or Deblur for users.
   However, as a result of the unexpected bad performance of Deblur on the synthetic data, the default algorithm in the pipeline was chosen to be \ac{dada2}.
@@
     \centering
-    \includegraphics[width=0.75\textwidth]{figure3.pdf}
+    \includegraphics[width=\textwidth]{figure3.pdf}
     \caption{
-      \textbf{The representative sequences generated by the different denoising/clustering methods are very similar but differ slightly in the sequences that are in low abundance.}
+      \textbf{The representative sequences generated by the different denoising/clustering methods are very similar but differ in the sequences that are in low abundance.}
       \textbf{(A)} The average weighted UniFrac distance between the representative sequences shows that the representative sequences and their compositions are fairly identical between the methods,
@@
 
+  Taxonomy databases are used to assign taxonomic identities to the representative sequences obtained after the DC step.
+  In order to compare the assigned taxonomies from different databases, we use the same reference sequences and assign taxonomies to them using different taxonomy reference databases.
   The three 16S taxonomic reference databases used in this study are SILVA~\cite{Quast2012}, \ac{gg}~\cite{DeSantis2006} and \ac{ncbi} RefSeq.
@@
   There are no two reference databases that are more similar than the other, with \ac{gg} and SILVA producing only marginally similar assignments compared to \ac{ncbi}.
+  This implies that the taxonomy assignments from each reference database are fairly unique and is one of the reasons for a large differences in the resultant co-occurrence networks generated from different taxonomy databases.
 
   As in the previous section, these comparisons only indicate similarity or dissimilarity between methods.
-  In order to obtain an absolute measure of accuracy of the taxonomic assignments we use the expected sequences from the mock datasets as the query sequences for the databases and the expected taxonomic composition as the standard to compare against (Figure~\ref{fig:figure4}C).
+  In order to obtain an absolute measure of accuracy of the taxonomic assignments we use the expected reference sequences from the mock datasets as the query sequences for the databases and the expected taxonomic composition as the standard to compare against (Figure~\ref{fig:figure4}C).
   Again, we observe that none of the databases perform better than the others.
@@
   % TODO: Check if this is necessary
-  \textcolor{gray}{
-    Figure~\ref{fig:figureS3} shows that the top 20 most abundant genera in the three resulting taxonomy composition tables are different.
-    The most abundant genus in the \ac{gg} taxonomy table was \textit{Escherichia} whereas in the SILVA taxonomy table it was \textit{Escherichia-Shigella}.
-    These might seem like minor differences, but when comparing a large number of composition tables these problems are hard to diagnose.
-    Figure~\ref{fig:tax_comparison_all} shows that the result is very similar when you look at all genera instead of just the top 20 most abundant genera.
-    Moreover, Figure~\ref{fig:venn_tax} shows that there does not seem to exist a correlation between abundance and mismatch.
-    This implies that the most abundant sequences are not necessarily the ones that are consistently matched to the same taxonomies in the different reference databases.
-  }
+    % Figure~\ref{fig:figureS3} shows that the top 20 most abundant genera in the three resulting taxonomy composition tables are different.
+    % The most abundant genus in the \ac{gg} taxonomy table was \textit{Escherichia} whereas in the SILVA taxonomy table it was \textit{Escherichia-Shigella}.
+    % These might seem like minor differences, but when comparing a large number of composition tables these problems are hard to diagnose.
+    % Figure~\ref{fig:tax_comparison_all} shows that the result is very similar when you look at all genera instead of just the top 20 most abundant genera.
+    % Moreover, Figure~\ref{fig:venn_tax} shows that there does not seem to exist a correlation between abundance and mismatch.
+    % This implies that the most abundant sequences are not necessarily the ones that are consistently matched to the same taxonomies in the different reference databases.
 
@@
   Our reasons for choosing \ac{ncbi} are that it is regularly maintained and updated and accuracy of the assignment is still comparable to the SILVA and \ac{gg} reference databases that are routinely used as reference databases.
-  \hl{Or combine results from multiple databases?}
+  % \hl{Or combine results from multiple databases?}
 
@@
     \caption{
-      \textbf{Taxonomic reference databases vary widely in terms of their taxonomies.}
-      \textbf{(A)} The assignment of the top 50 representative sequences to their respective taxonomies using the three different reference databases shows how the same sequences are assigned to different genus.
-      \textbf{(B)} The percentage of \ac{OTU}s assigned to the same taxonomic label when using different reference databases.
+      \textbf{Taxonomic reference databases vary widely in terms of their taxonomy assignments.}
+      \textbf{(A)} The assignment of the top 50 representative sequences to their respective taxonomies using the three different reference databases shows how the same sequences are assigned to different Genus.
+      \textbf{(B)} The percentage of \ac{otu}s assigned to the same taxonomic label when using different reference databases.
       The percentage of mismatches decrease at higher taxonomic levels but even at the Phylum level there exists around 10\% of mismatches.
-      \textbf{(C)} The Bray-Curtis dissimilarity between the expected taxonomy profile and calculated taxonomy profile in the mock datasets shows that there is no best choice of database for every dataset.
+      \textbf{(C)} The Bray-Curtis dissimilarity between the expected taxonomy profile and calculated taxonomy profile in the mock datasets shows that there is no singular best choice of database for every dataset.
     }
@@
   To get a more quantitative idea, we can take a look at the nodes and edges (Figure~\ref{fig:figure5}B) in common between the networks using UpSet plots (only \ac{magma}, \ac{mldm}, \ac{spieceasi}, \ac{sparcc} are used in the comparison).
-  The results for the node intersections show that the networks have a large number of edges in common ($63$ out of $67$ nodes in the smallest network - \ac{magma}) and no network possesses any unique node.
+  The results for the node intersections show that the networks have a large number of nodes in common ($63$ out of $67$ nodes in the smallest network - \ac{magma}) and no network possesses any unique node.
   The edge intersections in contrast show that only $19$ edges (out of $98$ edges in the smallest network - \ac{magma}) are in common between all the methods and each network has a large number of unique edges.
@@
   Unlike in the previous sections, where were we evaluated the performance of methods on mock datasets, there are no such datasets that contain a set of known interactions for the evaluation of the network inference algorithms.
-  Therefore, we propose the construction of a \hl{consensus network} (Figure~\ref{fig:figure5}C) involving \ac{magma}, \ac{mldm}, \ac{spieceasi} and \ac{sparcc} by merging the p-values generated from bootstraps of the original taxonomy composition table using the Browns pvalue combining method (ref).
+  Therefore, we propose the construction of a consensus network (Figure~\ref{fig:figure5}C) involving \ac{magma}, \ac{mldm}, \ac{spieceasi} and \ac{sparcc} by merging the p-values generated from bootstraps of the original taxonomy composition table using the Browns pvalue combining method [ref].
   The resulting network will have weights reported by all four methods.
@@
   These tools and parameters chosen as the defaults for the pipeline and are given in Table~\ref{tab:default_options}.
+
+  \begin{table}[h]
+    \centering
+    \small
+    \begin{tabular}{|c|c|c|}
+      \hline
+      \textbf{Process} & \textbf{Tool} & \textbf{Parameters} \\
+      \hline
+      Denoising and Clustering & Dada2/Deblur & default \\
+      Taxonomy Assignment & \ac{ncbi} with Blast & RefSeq database \\
+      OTU Processing & Based on statistical power & Dynamic cutoff \\
+      Network Inference & Consensus method & - \\
+      \hline
+    \end{tabular}
+    \caption{Default tools and parameters for the pipeline}
+    \label{tab:default_options}
+  \end{table}
+
   The recommended tool for the \ac{dc} step are \ac{dada2} or Deblur were chosen based on their accuracy in recapitulating the reference sequences in mock communities and synthetic data.
@@
   Furthermore, the L1 distance of networks generated by altering one of the steps of the pipeline from the default against the default network (Figure~\ref{fig:figure6}B) show that the biggest deviations from the default network occur when the \ac{ta} and \ac{ni} steps are changed, reinforcing the same results observed in Figure~\ref{fig:figure2}.
+Figure~\ref{fig:figure7} shows the co-occurrence networks inferred for the hard palate for healthy subjects in a periodontal disease study [ref] and the healthy stool microbiome in fecal microbial transplant study [ref].
+These consensus networks were generated using the default tools and parameters from Table~\ref{tab:default_options}.
 
@@
       \textbf{Network inference and taxonomic assignment have the highest influence on the inferred network structures.}
-      \textbf{(A)} The network constructed using the default pipeline parameters (DC=\ac{dada2}, TA=\ac{ncbi}, OP=on, NI=\ac{sparcc}) is compared with networks when one of the steps use a different tool.
-      The common connections are in black, connections unique to the network are colored purple and connections in the default network and but not present in the current network are gray.
-      \textbf{(B)} The L1 distance of the networks generated by changing one step of the default pipeline against the network generated using the default parameters.
+      \textbf{(A)} The network constructed using the default pipeline parameters (DC=\ac{dada2}, TA=\ac{ncbi}, OP=on, NI=\ac{sparcc}) is compared with networks generated when one of the steps use a different tool.
+      The common connections (common with the default network) are in black, connections unique to the network are colored purple and connections in the default network but not present in the current network are gray.
+      \textbf{(B)} The L1 distance between the networks generated by changing one step of the default pipeline and the network generated using the default parameters.
     }
@@
     \caption{
-      \textbf{The networks generated using the default pipeline settings.}
+      The consensus networks generated using the default pipeline settings.
+      \textbf{(A)} Co-occurrence network of the Hard Palate microbiome generated from samples of healthy subjects in a periodontal diseases study.
+      \textbf{(B)} Co-occurrence network of the Stool microbiome generated from samples of healthy subjects in a fecal microbiome transplant study.
   }

### 2019-11-25 sections/supplementary_figures.tex
Remove supplementary for now

--- 2018-08-07 sections/supplementary_figures.tex
@@
 
-\section*{Supplementary figures}
+% \newpage
+% \section*{Supplementary figures}
 
-  \renewcommand{\thefigure}{S\arabic{figure}}
-  \setcounter{figure}{0}
+%   \renewcommand{\thefigure}{S\arabic{figure}}
+%   \setcounter{figure}{0}
 
-  \begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.93\textwidth]{all_denoise_reg.pdf}
-    \caption{
-      \textbf{Pairwise regression plots for all the denoising methods}.
-    }
-    \label{fig:all_denoise_reg}
-  \end{figure}
+%   \begin{figure}[h]
+%   \centering
+%   \includegraphics[width=0.67\linewidth]{figureS1.pdf}
+%   \caption{
+%     \textbf{Comparison of various denoising and clustering algorithms used in the pipeline}.
+%     (A, B) Correlation of the abundances of the taxa that are common between the count matrices created by two different methods.
+%     (A) The best correlation (most similar methods) is between open-reference and denovo.
+%     (B) The worst correlation (least similar methods) is between open-reference and dada2.
+%     (C) A heatmap showing the $\mathrm{R}^2$ of all pairwise comparisons of the methods.
+%   }
+%   \label{fig:figureS1}
+% \end{figure}
 
-  \begin{figure}[h!]
-    \centering
-    \includegraphics[width=.9\textwidth]{denoise_network.pdf}
-    \caption{
-      \textbf{Networks generated using various denoising methods.}
-      The denoising methods used in this result are closed-reference clustering, open-reference clustering, denovo clustering and \ac{dada2}.
-      The database (\ac{gg}) and network inference methods (\ac{sparcc}) were kept constant.
-      (A) The set of interactions predicted by using all the methods (union)
-      (B) The set of interactions that were common to all the methods (intersection)
-    }
-    \label{fig:denoise_network}
-  \end{figure}
+%   \begin{figure}[h]
+%     \centering
+%     \includegraphics[width=0.8\linewidth]{figureS2.pdf}
+%     \caption{Results of the mock community analysis. (A) Weighted unifrac (B) Unweighted unifrac}
+%     \label{fig:figureS2}
+%   \end{figure}
+
+%   \begin{figure}[h]
+%     \centering
+%     \includegraphics[width=0.8\linewidth]{figureS3.pdf}
+%     \caption{Results of the synthetic community analysis. (A) Weighted unifrac (B) Unweighted unifrac}
+%     \label{fig:figureS3}
+%   \end{figure}
+
+%   \begin{figure}[h]
+%     \centering
+%     \includegraphics[width=0.8\linewidth]{pdf/all_denoise_reg.pdf}
+%     \caption{All pairwise correlations comparing the similarity between different denoising and clustering methods}
+%     \label{fig:figureS4}
+%   \end{figure}
+
+%   \begin{figure}[h]
+%     \centering
+%     \includegraphics[width=\linewidth]{figureS4.pdf}
+%     \caption{
+%       \textbf{(A)} Taxonomy composition of the 20 most abundant genera predicted using different taxonomy references databases: Greengenes, SILVA and NCBI.
+%       The legend shows the common and the unique genera among the taxonomy assignments.
+%   }
+%     \label{fig:figureS4}
+%   \end{figure}
+
+%   \begin{figure}[h]
+%     \centering
+%     \includegraphics[width=\linewidth]{figureS5.pdf}
+%     \caption{
+%       The bray-curtis dissmilarity between the expected taxonomic composition and generated taxonomic composiion for the synthetic datasets.
+%   }
+%     \label{fig:figureS4}
+%   \end{figure}
+
+%   \begin{figure}[h]
+%     \centering
+%     \includegraphics[width=\linewidth]{figureS6.pdf}
+%     \caption{
+%       The effect of OTU processing measurs on the OTU table
+%   }
+%     \label{fig:figureS4}
+%   \end{figure}
+
+%   \begin{figure}[h]
+%     \centering
+%     \includegraphics[width=0.8\linewidth]{png/hamming_distance.png}
+%     \caption{The hamming distance between the networks is shown. The similarity between various methods was found to vary with the data-source used.}
+%     \label{fig:figureS4}
+%   \end{figure}
+
+%   \begin{figure}[h]
+%     \centering
+%     \includegraphics[width=0.8\linewidth]{pdf/denoise_network.pdf}
+%     \caption{A network showing union and intersection of networks generated using certain combination of methods}
+%     \label{fig:figureS5}
+%   \end{figure}

### 2019-11-25 sections/discussion.tex
Update discussion for final draft

@@
 % General statements
-Co-occurrence associations in microbial communities will identify important interactions that drive microbial community structure and organization.
+Co-occurrence associations in microbial communities help identify important interactions that drive microbial community structure and organization.
 There is still much disagreement between the networks generated using various methods.
 This study makes a significant effort to identify those tools and parameters that enable more accurate inference of these co-occurrence networks.
-The mindpipe pipeline also enables the user to explore the effect of various tools and their combinations on their dataset of interest, thereby enabling one to identify those tools that perform their best on that particular dataset.
+We have developed mindpipe, a pipeline for 16S data analysis and inference of co-occurrence networks.
+This pipeline enables the analysis of 16S sequence data using different combinations of tools and makes comparison of these results straightforward.
+It also enables the user to explore the effect of various tools and their combinations on their dataset of interest, thereby enabling one to identify those tools that perform their best on that particular dataset.
+The pipeline is build to be modular and hence provides for a plug-and-play architecture wherein new tools and steps can be easily added.
 
 The different denoising/clustering methods differ mostly in their identification of sequences that are in low abundances.
-Hence, they do not have much of an impact on the inferred co-occurrence networks when the sequences of abundance are removed.
+Hence, they do not have much of an impact on the inferred co-occurrence networks when the sequences of low abundance are removed.
 However, comparison of inferred and expected reference sequences and their abundances in mock community datasets has allowed us to identify \ac{dada2} as the method which best recapitulates expected the sequence composition.
@@
 
-\hl{Other factors that play an important role in network inference} \\
-Differences in environments of the samples could lead to the inference of spurious interactions.
+% \hl{Other factors that play an important role in network inference} \\
+% Differences in environments of the samples could lead to the inference of spurious interactions.
 
@@
 Thus, in this study we have identified a set of tools and parameters that help infer more accurately and consistently co-occurrence networks from microbial 16S sequencing data.
+
+% TODO: Uses in MIND, exploration, standardization, reproducibility

### 2019-12-11 sections/methods.tex
Updated draft of manuscript

@@
   The mock community 16S datasets were obtained from mockrobiota~\cite{Bokulich2016}.
-  The mock datasets used for this study were labelled mock4, mock12 and mock16 datasets.
+  The mock datasets used for this study were labelled mock4, mock12 and mock16.
+  The mock4 community is composed of 21 bacterial strains represented in equal abundances in two replicate samples, and same strains in unequal abundances in two more replicate samples.
+  The mock12 community is composed of 27 bacterial strains that include closely related taxa with some pairs having only one to two nucleotide difference from another.
+  The mock16 community is composed of equal amounts of 49 bacteria and 10 archea.
   The synthetic datasets were generated using an artificial read simulator called ART~\cite{Huang2012}.
   Three different microbial composition profiles were used as input
-  Synthetic reads were generated using a soil and water microbiome composition profiles from the Earth Microbiome Project~\cite{Thompson2017} and healthy gut microbiome project from the Human Microbiome Project~\cite{HumanMicrobiomeProjectConsortium2012}
-  The scripts used to generate the synthetic data are in the \hl{scripts folder of the respository}
+  Synthetic reads were generated using a soil and water microbiome composition profiles from the Earth Microbiome Project~\cite{Thompson2017} and healthy gut microbiome project from the Human Microbiome Project~\cite{HumanMicrobiomeProjectConsortium2012}.
+  The scripts used to generate the synthetic data are in the scripts folder of the respository (\href{https://github.com/dileep-kishore/mind-pipeline-paper}{https://github.com/dileep-kishore/mind-pipeline-paper})
 
@@
   This also simplifies the process of adding new tools (refer to modules section in the mindpipe documentation).
-  In addition to the Python package, the entire pipeline has been containerized into a Docker~\cite{Merkel1994} image (\hl{dockerhub link}) for easy deployment and setup.
+  In addition to the Python package, the entire pipeline has been containerized into a Docker~\cite{Merkel1994} image (dockerhub link) for easy deployment and setup.
   The main components of the pipeline are detailed in the subsequent sections.
@@
   \end{itemize}
-  The pipeline includes Pearson, Spearman and \ac{sparcc}~\cite{Friedman2012} as the pairwise correlation metrics, and \ac{spieceasi}~\cite{Kurtz2015}, \ac{mldm}~\cite{Yang2017} and \ac{magma} (\hl{in biorxiv}) as the direct association metrics.
+  The pipeline includes Pearson, Spearman and \ac{sparcc}~\cite{Friedman2012} as the pairwise correlation metrics, and \ac{spieceasi}~\cite{Kurtz2015}, \ac{mldm}~\cite{Yang2017} and \ac{magma} (in biorxiv) as the direct association metrics.
 
@@
 
-  \textbf{Power calculations:}
+  % TODO: Variance and PCA calculation
 
-  \textbf{P-value merging}
-  The empirical Browns method is used for combining p-values from the various methods to obtain a consensus pvalue, which is used to create the consensus network.
+  % \textbf{Power calculations:}
+
+\textbf{P-value merging}:
+The empirical Browns method [ref] is used for combining p-values from the various methods to obtain a consensus pvalue, which is used to create the consensus network.
 
   \subsection*{Code and Data Availability}
-
-  Links here

### 2019-12-11 sections/results.tex
Updated draft of manuscript

@@
       \textbf{The choice of database contributes to the most variance in the networks}.
-      \textbf{(A)} The relative variance in the networks contributed by the DC, TA and OP steps of the pipeline.
+      \textbf{(A)} The total relative variance in the networks contributed by the DC, TA and OP steps of the pipeline (right) and the linear model used to calculate the relative variance (left).
       \textbf{(B)} All combinations of inferred networks are shown as points on a PCA plot.
@@
   Denoising and clustering are commonly carried out to generate representative sequences from the raw 16S sequencing data and to obtain the \ac{otu}/\ac{esv} tables (counts of these representative sequences for each sample).
-  In order to compare the \ac{otu} tables generated by these tools we process the same 16S sequencing reads (\hl{FMT} dataset) using 5 different methods \- open-reference clustering, closed-reference clustering, denovo clustering, \ac{dada2}~\cite{Callahan2016} and Deblur~\cite{Amir2017}.
+  In order to compare the \ac{otu} tables generated by these tools we process the same 16S sequencing reads (healthy samples from a fecal microbiome transplant study [ref]) using 5 different methods \- open-reference clustering, closed-reference clustering, denovo clustering, \ac{dada2}~\cite{Callahan2016} and Deblur~\cite{Amir2017}.
   The first three methods are from the \ac{qiime1}~\cite{Caporaso2010} package.
@@
   Based on their slightly better performance on the mock datasets, their \textit{de novo} error correcting nature and other previous studies [refs] we recommend either \ac{dada2} or Deblur for users.
-  However, as a result of the unexpected bad performance of Deblur on the synthetic data, the default algorithm in the pipeline was chosen to be \ac{dada2}.
+  However, as a result of the unexpected poor performance of Deblur on the synthetic data, the default algorithm in the pipeline was chosen to be \ac{dada2}.
 

### 2020-01-23 scripts/figure2/figure2a_plot.R
Update figure size and aspect ratios

@@ -6,7 +6,7 @@ library(ggpubr)
 
 
 group <- c("DC", "TA", "OP", "misc")
-percentage <- c(24.90, 17.97, 5.16, 51.96)
+percentage <- c(17.97, 24.90, 5.16, 51.96)
 labels <- paste0(group, " (", percentage, "%)")
 df <- data.frame(group=group, percentage=percentage, labels=labels)
 

### 2020-01-23 scripts/figure3/figure3ab_plot.R

@@ -48,4 +48,5 @@ weighted_tidy <- tidy_up_data(weighted)
 weighted_plot <- plot_heatmap(weighted_tidy, "weighted unifrac")
 unweighted_plot <- plot_heatmap(unweighted_tidy, "unweighted unifrac")
 
-ggarrange(weighted_plot, unweighted_plot, nrow=1, ncol=2, labels=c("A", "B"), common.legend=TRUE, legend="right")
+heatmap_plot <- ggarrange(weighted_plot, unweighted_plot, nrow=1, ncol=2, labels=c("A", "B"), common.legend=TRUE, legend="right")
+ggsave("figure3ab.pdf", heatmap_plot, width=11, height=8.5)

### 2020-01-23 scripts/figure3/figure3cd_plot.R

@@ -63,4 +63,6 @@ mock_unweighted_tidy <- rbind(mock4_unweighted_tidy, mock12_unweighted_tidy, moc
 mock_unweighted_plot <- make_dot_plot(mock_unweighted_tidy, "unweighted unifrac")
 
 
-ggarrange(mock_weighted_plot, mock_unweighted_plot, labels=c("C", "D"), nrow=1, ncol=2, common.legend=TRUE, legend="right")
+final_plot <- ggarrange(mock_weighted_plot, mock_unweighted_plot, labels=c("C", "D"), nrow=1, ncol=2, common.legend=TRUE, legend="right")
+
+ggsave("figure3cd.pdf", width=11, height=5)

### 2020-01-23 scripts/figure4/figure4a_plot.R

@@ -6,11 +6,20 @@ library(ggplot2)
 library(ggpubr)
 
 
+# TODO: FIXME: Combine all the genus in all the files so that you get full legend
 notu <- 50
 
-gg <- read.csv("fmt_gg.csv", header=TRUE, sep=",", na.strings="")[1:notu,]
-silva <- read.csv("fmt_silva.csv", header=TRUE, sep=",", na.strings="")[1:notu,]
-ncbi <- read.csv("fmt_ncbi.csv", header=TRUE, sep=",", na.strings="")[1:notu,]
+# FIXME: This doesn't work all files get same values
+read_data <- function(data_file, n) {
+    raw_df <- read.csv(data_file, header=TRUE, sep=",", na.strings="")[1:n,]
+    # df <- aggregate(Abundance ~ Genus, raw_df, sum)
+    # df
+    raw_df
+}
+
+gg <- read_data("fmt_gg.csv", notu)
+silva <- read_data("fmt_silva.csv", notu)
+ncbi <- read_data("fmt_ncbi.csv", notu)
 
 make_alluvial_plot <- function(db_data, title) {
     genus_factors <- levels(factor(db_data$Genus))
@@ -36,16 +45,16 @@ silva_plot <- make_alluvial_plot(silva, "SILVA")
 ncbi_plot <- make_alluvial_plot(ncbi, "NCBI")
 
 gg$OTU <- gg$empty
-gg_plot <- gg_plot +
-    geom_text_repel(stat="stratum", label.strata=TRUE, data=gg[,c("OTU", "Genus", "Abundance")], nudge_x=1)
+gg_plot <- gg_plot #+
+    # geom_text_repel(stat="stratum", label.strata=TRUE, data=gg[,c("OTU", "Genus", "Abundance")], nudge_x=1)
 
 silva$OTU <- silva$empty
-silva_plot <- silva_plot +
-    geom_text_repel(stat="stratum", label.strata=TRUE, data=silva[,c("OTU", "Genus", "Abundance")], nudge_x=1)
+silva_plot <- silva_plot #+
+    # geom_text_repel(stat="stratum", label.strata=TRUE, data=silva[,c("OTU", "Genus", "Abundance")], nudge_x=1)
 
 ncbi$OTU <- ncbi$empty
-ncbi_plot <- ncbi_plot +
[…]

### 2020-01-23 scripts/figure4/figure4b_plot.R

@@ -46,13 +46,14 @@ make_bar_plot <- function(data, title) {
         ylab="% of Mismatches",
         palette = c("#00AFBB", "#FC4E07"),
         #palette="Paired"
-    )
+    ) +
+    theme(axis.text.x=element_text(angle=30, hjust=1))
 }
 
 # Importing data
-gg <- tidy_up_data("gg.csv", notu)
-silva <- tidy_up_data("silva.csv", notu)
-ncbi <- tidy_up_data("ncbi.csv", notu)
+gg <- tidy_up_data("fmt_gg.csv", notu)
+silva <- tidy_up_data("fmt_silva.csv", notu)
+ncbi <- tidy_up_data("fmt_ncbi.csv", notu)
 
 # Combinations
 gg_silva <- combine_data(gg, silva)
@@ -63,5 +64,7 @@ gg_silva_plot <- make_bar_plot(gg_silva, "GreenGenes vs. SILVA")
 gg_ncbi_plot <- make_bar_plot(gg_ncbi, "GreenGenes vs. NCBI")
 ncbi_silva_plot <- make_bar_plot(ncbi_silva, "NCBI vs. SILVA")
 
-combined_plot <- ggarrange(gg_silva_plot, gg_ncbi_plot, ncbi_silva_plot, nrow=1, ncol=3, common.legend=TRUE, legend="top")
+combined_plot <- ggarrange(gg_silva_plot, gg_ncbi_plot, ncbi_silva_plot, nrow=1, ncol=3, common.legend=TRUE, legend="right")
 annotate_figure(combined_plot, fig.lab="B", fig.lab.pos="top.left", fig.lab.size=20)
+
+ggsave("figure4b.pdf", width=11, height=3.5)

### 2020-01-23 scripts/figure4/figure4c_plot.R

@@ -42,4 +42,6 @@ mock16_data <- tidy_up_data("mock16_braycurtis.csv", "mock16")
 mock_data <- rbind(mock4_data, mock12_data, mock16_data)
 dot_plot <- make_dot_plot(mock_data)
 dot_plot_facet <- facet(dot_plot, facet.by="tax_level")
-annotate_figure(dot_plot_facet, fig.lab="C", fig.lab.pos="top.left", fig.lab.size=20)
+final_plot <- annotate_figure(dot_plot_facet, fig.lab="C", fig.lab.pos="top.left", fig.lab.size=20)
+
+ggsave("figure4c.pdf", width=11, height=5)

### 2020-01-23 scripts/figure6/figure6a_plot.R

@@ -17,6 +17,11 @@ get_edgestyle <- function(x) {
 
 plot_network <- function(network_file, combined_layout, interaction_threshold, title, palette, ind) {
   graph_raw <- as_tbl_graph(read_graph(network_file, format="gml"))
+  if(title=="default") {
+      foreground = "#c1c1c1"
+  } else {
+      foreground = "#7570B3"
+  }
   graph <- graph_raw %>%
     activate(edges) %>%
     filter(abs(weight) > interaction_threshold) %>%
@@ -34,7 +39,7 @@ plot_network <- function(network_file, combined_layout, interaction_threshold, t
     # geom_node_point(aes(color=factor(colorkey))) +
     geom_node_point() +
     #scale_edge_color_manual(values = c(background=palette[[1]], foreground=palette[[ind]], common="black")) +
-    scale_edge_color_manual(values = c(foreground="#7570B3", background="#c1c1c1", common="black")) +
+    scale_edge_color_manual(values = c(foreground=foreground, background="#c1c1c1", common="black")) +
     scale_edge_linetype_manual(values = c(negative="dashed", positive="solid")) +
     scale_edge_alpha_manual(values = c(foreground=0.4, background=0.4, common=0.7)) +
     labs(title=title) +
@@ -58,13 +63,12 @@ combined_layout <- create_layout(graph=combined_graph, layout="linear", circular
 
 palette <- brewer.pal(n=6, name="Pastel1")
 default_plot <- plot_network("default.gml", combined_layout, 0.3, "default", palette, 1)
-database_plot <- plot_network("gg.gml", combined_layout, 0.3, "TA=Greengenes", palette, 2)
-clustering_plot <- plot_network("open_ref.gml", combined_layout, 0.3, "DC=open_ref", palette, 3)
-otu_filtering_plot <- plot_network("otu_filtering.gml", combined_layout, 0.3, "OP=off", palette, 4)
-network_inference_plot <- plot_network("spieceasi_direct.gml", combined_layout, 0.3, "NI=spieceasi",  palette, 5)
+database_plot <- plot_network("gg.gml", combined_layout, 0.3, "TA=GG", palette, 2)
+clustering_plot <- plot_network("open_ref.gml", combined_layout, 0.3, "DC=open_reference", palette, 3)
+otu_filtering_plot <- plot_network("otu_filtering.gml", combined_layout, 0.3, "OP=no", palette, 4)
[…]

### 2020-01-23 scripts/figure6/figure6b_plot.R

@@ -1,0 +1,11 @@
+#!/usr/bin/env Rscript
+
+library(tidyverse)
+library(ggpubr)
+
+
+l1_data <- read.csv("distance_to_ref.csv", header=TRUE, sep=",")
+
+ggboxplot(l1_data, x="step", y="distance", order=c("DC", "TA", "OP"), color="step", add="jitter",palette="Set2", ylab="L1 distance", xlab="Pipeline step")
+
+ggsave("figure6b.pdf", width=11, height=5)

### 2020-01-23 sections/results.tex
Update methods and results

@@
       \textbf{(B)} The relatively larger average unweighted UniFrac distance indicates that methods differ in their identification of sequences of low abundance,
-      \textbf{(C)} The distributions of the average unweighted UniFrac distance between the expected sequence profile and the calculated sequence profile in mock datasets show that dada2 and Deblur were the best performing methods in most of the datasets.
+      \textbf{(C, D)} The distributions of the avergage weighted UniFrac distance between the expected sequence profile and the calculated sequence profile in mock datasets.
+      \textbf{(D)} The distributions of the average unweighted UniFrac distance show that dada2 and Deblur were the best performing methods in most of the datasets.
     }

### 2020-01-23 sections/supplementary_figures.tex
Update supplementary section

--- 2019-06-05 sections/supplementary_figures.tex
@@
 \newpage
-\section*{Supplementary figures}
+\section*{Supplementary}
 
@@
   \centering
-  \includegraphics[width=0.67\linewidth]{figureS1.pdf}
+  \includegraphics[width=\linewidth]{figureS1.pdf}
   \caption{
     \textbf{Comparison of various denoising and clustering algorithms used in the pipeline}.
-    (A, B) Correlation of the abundances of the taxa that are common between the count matrices created by two different methods.
+    (A, B) Correlation of the abundances of the taxa that are in common between the count matrices created by two different methods.
     (A) The best correlation (most similar methods) is between open-reference and denovo.
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{figureS2.pdf}
-    \caption{Results of the mock community analysis. (A) Weighted unifrac (B) Unweighted unifrac}
+    \includegraphics[width=\linewidth]{figureS2.pdf}
+    \caption{
+      \textbf{Heatmaps showing the weighted and unweighted unifrac distances for the hard palate dataset analysis}.
+      (A) weighted unifrac distances and (B) unweighted unifrac distances between the representative sequences generated by different denoising and clustering algorithms.
+      These results are in agreement with the stool microbiome dataset.
+    }
     \label{fig:figureS2}
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{figureS3.pdf}
-    \caption{Results of the synthetic community analysis. (A) Weighted unifrac (B) Unweighted unifrac}
+    \includegraphics[width=\linewidth]{figureS3.pdf}
+    \caption{
+      \textbf{The distributions of the avergage weighted UniFrac distance between the expected sequence profile and the calculated sequence profile in the synthetic datasets}.
+      We observe no significant difference between the various methods on the synthetic datasets used for this study.
+    }
     \label{fig:figureS3}
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{pdf/all_denoise_reg.pdf}
+    \includegraphics[width=0.9\linewidth]{pdf/all_denoise_reg.pdf}
     \caption{All pairwise correlations comparing the similarity between different denoising and clustering methods}
@@
     \caption{
-      \textbf{(A)} Taxonomy composition of the 20 most abundant genera predicted using different taxonomy references databases: Greengenes, SILVA and NCBI.
+      \textbf{(A)} Taxonomy composition of the 20 most abundant genera predicted for the stool microbiome dataset generated using different taxonomy references databases: Greengenes, SILVA and NCBI.
       The legend shows the common and the unique genera among the taxonomy assignments.
@@
   }
-    \label{fig:figureS4}
+  \label{fig:figureS5}
   \end{figure}
@@
     \caption{
-      The effect of OTU processing measurs on the OTU table
+      Calculation of presence threshold that is applied on the OTU table in the OTU processing (OP) step of the pipeline.
+      This presence threhold $p_t$ is dependent on the number of samples in the dataset and the required correlation stength threshold.
   }
-    \label{fig:figureS4}
+    \label{fig:figureS6}
   \end{figure}
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{png/hamming_distance.png}
-    \caption{The hamming distance between the networks is shown. The similarity between various methods was found to vary with the data-source used.}
-    \label{fig:figureS4}
+    \includegraphics[width=\linewidth]{figureS8.pdf}
+    \caption{
+      The similarity between the networks generated using the different network inference algorithms for stool dataset (A) and the hard palate dataset (B).
+      The similarity between the various methods was found to vary with the dataset used.
+  }
+    \label{fig:figureS8}
   \end{figure}

### 2020-01-23 scripts/figure_s1/figure_s1c.R

@@ -1,0 +1,52 @@
+#!/usr/bin/env Rscript
+
+library(dplyr)
+library(ggplot2)
+library(ggpubr)
+
+methods <- c("closed_reference", "de_novo", "open_reference", "deblur", "dada2")
+
+tidy_up_data <- function(corr_data) {
+    corr_data$method1 <- factor(corr_data$method1, levels=methods)
+    corr_data$method2 <- factor(corr_data$method2, levels=methods)
+    tidy_data <- corr_data %>%
+        group_by(method1, method2) %>%
+        select_at(c("method1", "method2", "corr")) %>%
+        summarize_all(.funs=mean) %>%
+        rowwise() %>%
+        mutate(pair=sort(c(method1, method2)) %>% paste(collapse=",")) %>%
+        group_by(pair) %>%
+        distinct(pair, .keep_all=TRUE)
+}
+
+plot_heatmap <- function(data, title) {
+    ggplot(data=data, aes(x=method1, y=method2, fill=corr)) +
+        geom_tile(color="white") +
+        geom_text(aes(label=abbreviate(corr))) +
+        scale_fill_distiller(
+            palette="Spectral",
+            limits=c(0, 1),
+            breaks=c(0, 0.25, 0.5, 0.75, 1),
+            labels=c("Low", 0.25, 0.5, 0.75, "High")
+        ) +
+        ggtitle(title) +
+        theme_pubr() +
+        theme(
+            plot.title=element_text(hjust=0.5),
+            axis.title=element_blank(),
+            axis.text.x=element_text(angle=30, hjust=1),
+            axis.text.y=element_text(angle=30)
+        ) +
+        coord_fixed()
+}
+
+corr <- read.csv("fmt_corr.csv", sep=",", header=TRUE)
+corr_tidy <- tidy_up_data(corr)
+# weighted <- read.csv("fmt_weighted_unifrac.csv", sep=",", header=TRUE)
+# weighted_tidy <- tidy_up_data(weighted)
+
+corr_plot <- plot_heatmap(corr_tidy, "correlation")
+# unweighted_plot <- plot_heatmap(unweighted_tidy, "unweighted unifrac")
+
+heatmap_plot <- ggarrange(corr_plot, nrow=1, ncol=1, labels=c("C"), common.legend=TRUE, legend="right")
+ggsave("figure_s1c.pdf", heatmap_plot, width=11, height=8.5)

### 2020-01-23 scripts/figure_s2/figure_s2_plot.R

@@ -1,0 +1,52 @@
+#!/usr/bin/env Rscript
+
+library(dplyr)
+library(ggplot2)
+library(ggpubr)
+
+methods <- c("closed_reference", "de_novo", "open_reference", "deblur", "dada2")
+
+tidy_up_data <- function(unifrac_data) {
+    unifrac_data$method1 <- factor(unifrac_data$method1, levels=methods)
+    unifrac_data$method2 <- factor(unifrac_data$method2, levels=methods)
+    tidy_data <- unifrac_data %>%
+        group_by(method1, method2) %>%
+        select_at(c("method1", "method2", "unifrac")) %>%
+        summarize_all(.funs=mean) %>%
+        rowwise() %>%
+        mutate(pair=sort(c(method1, method2)) %>% paste(collapse=",")) %>%
+        group_by(pair) %>%
+        distinct(pair, .keep_all=TRUE)
+}
+
+plot_heatmap <- function(data, title) {
+    ggplot(data=data, aes(x=method1, y=method2, fill=unifrac)) +
+        geom_tile(color="white") +
+        geom_text(aes(label=abbreviate(unifrac))) +
+        scale_fill_distiller(
+            palette="Spectral",
+            limits=c(0, 1),
+            breaks=c(0, 0.25, 0.5, 0.75, 1),
+            labels=c("Similar", 0.25, 0.5, 0.75, "Dissimilar")
+        ) +
+        ggtitle(title) +
+        theme_pubr() +
+        theme(
+            plot.title=element_text(hjust=0.5),
+            axis.title=element_blank(),
+            axis.text.x=element_text(angle=30, hjust=1),
+            axis.text.y=element_text(angle=30)
+        ) +
+        coord_fixed()
+}
+
+unweighted <- read.csv("hp_unweighted_unifrac.csv", sep=",", header=TRUE)
+unweighted_tidy <- tidy_up_data(unweighted)
+weighted <- read.csv("hp_weighted_unifrac.csv", sep=",", header=TRUE)
+weighted_tidy <- tidy_up_data(weighted)
+
+weighted_plot <- plot_heatmap(weighted_tidy, "weighted unifrac")
+unweighted_plot <- plot_heatmap(unweighted_tidy, "unweighted unifrac")
+
+heatmap_plot <- ggarrange(weighted_plot, unweighted_plot, nrow=1, ncol=2, labels=c("A", "B"), common.legend=TRUE, legend="right")
+ggsave("figure_s2.pdf", heatmap_plot, width=11, height=8.5)

### 2020-01-23 scripts/figure_s3/figure3_data.py

@@ -1,0 +1,1 @@
+../figure3/figure3_data.py

### 2020-01-23 scripts/figure_s3/figure_s3_plot.R

@@ -1,0 +1,61 @@
+#!/usr/bin/env Rscript
+
+library(dplyr)
+library(ggplot2)
+library(ggpubr)
+
+methods <- c("input", "closed_reference", "de_novo", "open_reference", "deblur", "dada2")
+
+tidy_up_data <- function(unifrac_data, name) {
+    unifrac_data$dataset <- rep(name, nrow(unifrac_data))
+    unifrac_data$method1 <- factor(unifrac_data$method1, levels=methods)
+    unifrac_data$method2 <- factor(unifrac_data$method2, levels=methods)
+    tidy_data <- unifrac_data %>%
+        filter(method2 == "input") %>%
+        filter(method1 != "input") %>%
+        select_at(c("method1", "sample", "unifrac", "dataset"))
+}
+
+make_dot_plot <- function(unifrac_data, title) {
+    ggstripchart(
+        unifrac_data,
+        x="method1",
+        y="unifrac",
+        color="dataset",
+        title=title,
+        xlab="denoising/clustering method",
+        ylab="unifrac distance",
+        ylim=c(0, 1),
+        size=3,
+        add="jitter",
+    ) +
+    theme(
+        plot.title=element_text(hjust=0.5),
+        axis.text.x=element_text(angle=30, hjust=1),
+    )
+}
+
+# Weighted
+emp_soil_weighted <- read.csv("emp_soil_weighted_unifrac.csv", sep=",", header=TRUE)
+emp_water_weighted <- read.csv("emp_water_weighted_unifrac.csv", sep=",", header=TRUE)
+
+emp_soil_weighted_tidy <- tidy_up_data(emp_soil_weighted, "emp_soil")
+emp_water_weighted_tidy <- tidy_up_data(emp_water_weighted, "emp_water")
+
+weighted_tidy <- rbind(emp_soil_weighted_tidy, emp_water_weighted_tidy)
+weighted_plot <- make_dot_plot(weighted_tidy, "weighted unifrac")
+
+# Unweighted
+emp_soil_unweighted <- read.csv("emp_soil_unweighted_unifrac.csv", sep=",", header=TRUE)
+emp_water_unweighted <- read.csv("emp_water_unweighted_unifrac.csv", sep=",", header=TRUE)
+
+emp_soil_unweighted_tidy <- tidy_up_data(emp_soil_unweighted, "emp_soil")
[…]

### 2020-01-23 scripts/figure_s3/move_otus.py

@@ -1,0 +1,1 @@
+../figure3/move_otus.py

### 2020-01-23 scripts/figure_s3/move_trees.py

@@ -1,0 +1,1 @@
+../figure3/move_trees.py

### 2020-01-23 scripts/figure_s5/figure4ab_data.py

@@ -1,0 +1,1 @@
+../figure4/figure4ab_data.py

### 2020-01-23 scripts/figure_s5/figure4c_data.py

@@ -1,0 +1,1 @@
+../figure4/figure4c_data.py

### 2020-01-23 scripts/figure_s5/figure_s5_plot.R

@@ -1,0 +1,43 @@
+#!/usr/bin/env Rscript
+
+library(tidyverse)
+library(ggplot2)
+library(ggpubr)
+
+levels <- c("Family", "Genus", "Species")
+
+tidy_up_data <- function(data_file, name) {
+    data <- read_csv(data_file)
+    data$dataset <- rep(name, nrow(data))
+    data$tax_level <- factor(data$tax_level, levels=levels)
+    return(data)
+}
+
+
+make_dot_plot <- function(braycurtis_data) {
+    ggstripchart(
+        braycurtis_data,
+        x="database",
+        y="braycurtis",
+        color="dataset",
+        xlab="database",
+        ylab="Braycurtis dissimilarity",
+        ylim=c(0, 1),
+        size=3,
+        add="jitter",
+    ) +
+    theme(
+        plot.title=element_text(hjust=0.5),
+    )
+}
+
+
+emp_soil_data <- tidy_up_data("emp_soil_braycurtis.csv", "emp_soil")
+emp_water_data <- tidy_up_data("emp_water_braycurtis.csv", "emp_water")
+
+synth_data <- rbind(emp_soil_data, emp_water_data)
+dot_plot <- make_dot_plot(synth_data)
+dot_plot_facet <- facet(dot_plot, facet.by="tax_level")
+# final_plot <- annotate_figure(dot_plot_facet, fig.lab="C", fig.lab.pos="top.left", fig.lab.size=20)
+
+ggsave("figure_s5.pdf", width=11, height=5)

### 2020-01-23 scripts/figure_s8/calculate_dist.py

@@ -1,0 +1,84 @@
+#!/usr/bin/env python3
+
+from itertools import product
+import pathlib
+
+import networkx as nx
+from networkx.algorithms.similarity import graph_edit_distance
+import numpy as np
+import pandas as pd
+
+
+def get_similarity(graph1, graph2):
+    # NOTE: https://graph-tool.skewed.de/static/doc/topology.html?highlight=similarity#contents
+    graph1_nodes = list(graph1.nodes)
+    graph2_nodes = list(graph2.nodes)
+    nodes = list(set(graph1_nodes) | set(graph2_nodes))
+    node_dict = dict(zip(nodes, range(len(nodes))))
+    n = len(nodes)
+    adj1 = np.zeros((n, n), dtype=np.float)
+    for u, v, edata in graph1.edges(data=True):
+        u_ind = node_dict[u]
+        v_ind = node_dict[v]
+        adj1[u_ind, v_ind] = edata["weight"]
+        adj1[v_ind, u_ind] = edata["weight"]
+    adj2 = np.zeros((n, n), dtype=np.float)
+    for u, v, edata in graph2.edges(data=True):
+        u_ind = node_dict[u]
+        v_ind = node_dict[v]
+        adj2[u_ind, v_ind] = edata["weight"]
+        adj2[v_ind, u_ind] = edata["weight"]
+    # E = sum(n(graph1_edges), n(graph2_edges))
+    E = n * n
+    # d = sum(|graph1_edges-graph2_edges|)
+    mismatch = np.ones((n, n), dtype=np.int)
+    for u_ind, v_ind in product(range(n), range(n)):
+        w1, w2 = adj1[u_ind, v_ind], adj2[u_ind, v_ind]
+        if (w1 * w2) >= 0:
+            mismatch[u_ind, v_ind] = 0
+    d = mismatch.sum()
+    # S = E - d
+    S = (E - d) / E
+    return S
+
+
+def get_hamming_dist(graph1, graph2, sign=True):
+    adjmat1 = nx.to_numpy_matrix(graph1)
+    adjmat2 = nx.to_numpy_matrix(graph2)
+    if sign:
+        adjmat1[adjmat1 < 0] = -1
+        adjmat1[adjmat1 > 0] = 1
+        adjmat2[adjmat2 < 0] = -1
+        adjmat2[adjmat2 > 0] = 1
+        dtype = int
+    else:
+        dtype = bool
[…]

### 2020-01-23 scripts/figure_s8/figures8_plot.R

@@ -1,0 +1,53 @@
+#!/usr/bin/env Rscript
+
+library(dplyr)
+library(ggplot2)
+library(ggpubr)
+
+methods <- c("magma", "pearson", "mldm", "spearman", "spieceasi", "sparcc")
+
+tidy_up_data <- function(similarity_data) {
+    similarity_data$method1 <- factor(similarity_data$method1, levels=methods)
+    similarity_data$method2 <- factor(similarity_data$method2, levels=methods)
+    tidy_data <- similarity_data %>%
+        group_by(method1, method2) %>%
+        select_at(c("method1", "method2", "similarity")) %>%
+        summarize_all(.funs=mean) %>%
+        rowwise() %>%
+        mutate(pair=sort(c(method1, method2)) %>% paste(collapse=",")) %>%
+        group_by(pair) %>%
+        distinct(pair, .keep_all=TRUE)
+}
+
+plot_heatmap <- function(data, title) {
+    ggplot(data=data, aes(x=method1, y=method2, fill=similarity)) +
+        geom_tile(color="white") +
+        geom_text(aes(label=abbreviate(similarity))) +
+        scale_fill_distiller(
+            palette="Spectral",
+            direction=1,
+            limits=c(0, 1),
+            breaks=c(0, 0.25, 0.5, 0.75, 1),
+            labels=c("Dissimilar", 0.25, 0.5, 0.75, "Similar")
+        ) +
+        ggtitle(title) +
+        theme_pubr() +
+        theme(
+            plot.title=element_text(hjust=0.5),
+            axis.title=element_blank(),
+            axis.text.x=element_text(angle=30, hjust=1),
+            axis.text.y=element_text(angle=30)
+        ) +
+        coord_fixed()
+}
+
+fmt_similarity <- read.csv("fmt_similarity.csv", sep=",", header=TRUE)
+fmt_similarity_tidy <- tidy_up_data(fmt_similarity)
+hp_similarity <- read.csv("hp_similarity.csv", sep=",", header=TRUE)
+hp_similarity_tidy <- tidy_up_data(hp_similarity)
+
+fmt_similarity_plot <- plot_heatmap(fmt_similarity_tidy, "FMT similarity")
+hp_similarity_plot <- plot_heatmap(hp_similarity_tidy, "HP similarity")
+
[…]

### 2020-01-23 sections/coverpage.tex
Add corresponding authors

@@
   Charles DeLisi\textsuperscript{1,3},
-  Kirill Korolev\textsuperscript{1,2},\\
-  Daniel Segr\`{e}\textsuperscript{1,3,4}\\
+  Kirill Korolev\textsuperscript{$\dagger$1,2},\\
+  Daniel Segr\`{e}\textsuperscript{$\dagger$1,3,4}\\
   \vspace{1cm}
@@
   \textsuperscript{4}Department of Biology, Boston University, Boston, Massachusetts, USA\\
+  \textsuperscript{$\dagger$}Correspondence should be sent to \href{mailto:korolev.bu.edu}{korolev.bu.edu} or \href{mailto:dsegre.bu.edu}{dsegre.bu.edu}\\
 

### 2020-03-16 scripts/figure_s3/figure3_data.py
Updates from Overleaf

@@ -1,1 +1,0 @@
-../figure3/figure3_data.py

### 2020-03-16 scripts/figure_s3/move_otus.py

@@ -1,1 +1,0 @@
-../figure3/move_otus.py

### 2020-03-16 sections/results.tex

@@
   Figure~\ref{fig:figure3}B on the other hand shows an increase in dissimilarity between each pair of methods suggesting that the methods might differ in the sequences of low abundance.
-  In order to verify this claim, we changed the threshold for the sequence count and recalculated the weighted and unweighted UniFrac distance metrics and also looked at the number of representative sequences identified in each case (in the supplementary section).
-
-  %% TODO: Check if this is necessary
-    % For each of these methods we use the same taxonomy database (\ac{gg}) to assign taxonomies to the representative sequences.
-    % We then correlate the abundances of matching taxonomies between every combination of a pair of methods (Figure \ref{fig:all_denoise_reg}).
-    % The \ac{esv} tables generated by methods that perform denoising are very similar to each other $\sim0.9$ and the \ac{otu} tables generated by the clustering methods are very similar to each other $\sim0.9$, but results of denoising and clustering are highly uncorrelated with each other $\sim0.4$ (Figure \ref{fig:figureS2}).
+  %% TODO: Update this
+  In order to verify this claim, for each of these methods we use the \ac{gg} taxonomy database to assign taxonomies to the representative sequences.
+  We then correlate the abundances of matching taxonomies between a pair of DC methods (Figure \ref{fig:figureS1}A and B).
+  The \ac{esv} tables generated by methods that perform denoising are very similar to each other $\sim0.91$ and the \ac{otu} tables generated by the clustering methods are very similar to each other $\sim0.9$, but results of denoising and clustering are highly uncorrelated with each other $\sim0.4$ (Figure \ref{fig:figureS1}C).
   % The common core network (network made up of edges that are predicted by all methods) is very small compared to the total number of edges predicted by all the methods (Figure \ref{fig:denoise_network}).
@@
 
-  % TODO: Check if this is necessary
-    % Figure~\ref{fig:figureS3} shows that the top 20 most abundant genera in the three resulting taxonomy composition tables are different.
-    % The most abundant genus in the \ac{gg} taxonomy table was \textit{Escherichia} whereas in the SILVA taxonomy table it was \textit{Escherichia-Shigella}.
-    % These might seem like minor differences, but when comparing a large number of composition tables these problems are hard to diagnose.
-    % Figure~\ref{fig:tax_comparison_all} shows that the result is very similar when you look at all genera instead of just the top 20 most abundant genera.
-    % Moreover, Figure~\ref{fig:venn_tax} shows that there does not seem to exist a correlation between abundance and mismatch.
-    % This implies that the most abundant sequences are not necessarily the ones that are consistently matched to the same taxonomies in the different reference databases.
+  % TODO: Update this
+  Figure~\ref{fig:figureS4} shows that the top 20 most abundant genera in the three resulting taxonomy composition tables are different.
+  For example, the most abundant genus in the \ac{gg} taxonomy table was \textit{Escherichia} whereas in the SILVA taxonomy table it was \textit{Escherichia-Shigella}.
+  Although these are minor differences, when comparing a large number of composition tables these problems are hard to diagnose.
+  The comparison of all assigned genera instead of the just the top 20 contains the same percentage of matches and mismatches, implying that there does not seem to exist a correlation between abundance and mismatch.
+  This suggests that the most abundant sequences are not necessarily the ones that are consistently matched to the same taxonomies in the different reference databases.
 

### 2020-03-19 sections/abstract.tex
Updated draft of manuscript
Edits made by Daniel [03-16-2020]

@@
     Microbes tend to organize into communities consisting of hundreds of species involved in complex interactions with each other.
-    16S ribosomal RNA (16S rRNA) gene profiling provides snapshots that reveal the phylogenies and abundance profiles of these microbial communities.
+    16S ribosomal RNA (16S rRNA) amplicon profiling provides snapshots that reveal the phylogenies and abundance profiles of these microbial communities.
     These snapshots, when collected from multiple samples, have the potential to reveal which microbes co-occur, providing a glimpse into the network of influences in these communities.
     The inference of networks from 16S data is prone to statistical artifacts.
-    There are many tools for performing each step of the 16S analysis workflow and the results generated by different tools are varied, but the extent to which these steps affect the resultant network is still unclear.
+    There are many tools for performing each step of the 16S analysis workflow, but the extent to which these steps affect the final network is still unclear.
     In this study, we perform a meticulous analysis of each step of a pipeline that processes 16S sequencing data into a network of microbial associations.
-    Through this process, we determine the tools and parameters that generate the most accurate and robust co-occurrence networks.
+    % TODO: Instead of this say: We analyze the parts of the process that cause the most variance and what is the result
+    Through this process, we map how different choices of algorithms and parameters affect the co-occurrence network, and estimate steps which contribute most significantly to the variance.
+We further determine the tools and parameters that generate the most accurate and robust co-occurrence networks, based on comparison with mock and synthetic datasets.
     Ultimately, we develop a standardized pipeline that follows these default tools and parameters, but that can also help explore the outcome of any other combination of choices.
+    % TODO: Make this general (do not confine it to disease-related) - niche-related(?)
     We envisage that this standard pipeline for processing 16S sequencing data into networks of microbial co-occurrences could be used for integrating multiple data-sets, and for generating comparative analyses and consensus networks useful for detecting disease-related patterns.
 \\
+    % TODO: Move package to segrelab github
     The pipeline is available as a Python package at \href{https://github.com/dileep-kishore/mindpipe}{https://github.com/dileep-kishore/mindpipe}

### 2020-03-19 sections/introduction.tex
Updated draft of manuscript
Edits made by Daniel [03-16-2020]

@@
   These advances have led to large-scale data collection efforts involving environmental (Earth Microbiome Project) \cite{Thompson2017}, marine (Tara Oceans Project) \cite{Zhang2015} and human-associated microbiota (Human Microbiome Project) \cite{HumanMicrobiomeProjectConsortium2012} and thus led the generation of a wealth of data about these microbial communities.
-  These data-sets can be used to infer networks of co-occurrence relationships between microorganisms in the samples.
-  The co-occurrence networks can in turn be used to identify important interactions driving the dynamics of the microbial community, as well as, analyze emergent properties such as robustness and modularity.
 
-  16S sequencing data is limited by many factors such as resolution, sequencing depth, compositional nature, sequencing errors and copy number variations.
+  These data-sets can be used to infer networks of co-occurrence relationships between microorganisms in the samples.These networks have microbial taxa as nodes, and edges that represent the frequent co-occurrence (or negative correlations) across different datasets.
+Co-occurrence networks can in principle be used to identify important interactions driving the dynamics of the microbial community, including emergent ecosystem-level properties such as ecosystem robustness and modularity.
+
+  One of the limitations of this type of analysis is that 16S sequencing data is limited by many factors such as resolution, sequencing depth, compositional nature, sequencing errors and copy number variations.
+
   Many different analysis methods have been developed \cite{Callahan2016,Amir2017,Friedman2012,Kurtz2015} to counter these limitations and remove statistical artifacts, leading to the existence of a myriad of tools for each step of the data analysis workflow.
-  The different methods and tools developed to solve these issues infer vastly different community compositions and co-occurrence networks \cite{Golob2017,Weiss2016}, making the analysis unreproducible and comparisons convoluted.
-  But because there does not exist a dataset wherein the microbial interactions have been fully characterized, there is no straightforward way to determine which set of tools or methods generate the most accurate networks.
+  The different methods and tools developed to solve these issues infer vastly different community compositions and co-occurrence networks \cite{Golob2017,Weiss2016}, making it difficult to reliably compare networks across different publications and studies.
+Conversely, given the lack of comprehensive comparisons between directly observed microbial interactions (e.g.
+from co-culture experiments) and co-occurence networks, there is no straightforward way to determine which set of tools or methods generate the most accurate networks.
+
+ Previous work has developed popular platforms (like MG-RAST \cite{Keegan2016}, Qiita \cite{qiita}) and tools (such as QIIME \cite{Caporaso2010}) that provide pipelines for 16S data analysis.
+However, existing tools typically are not focused on the effects of upstream statistical methods on the inferred co-occurrence networks.
+Furthermore, no organized framework currently exist to systematically analyze and compare existing components of the data-analysis from amplicons to networks.
+
+  In this study, we present a standardized 16S data-analysis pipeline that produces robust and reproducible co-occurrence networks from community 16S sequence data, and allow users to interactively explore how the network would change upon using different alternative tools and parameters at each step.
+Our pipeline is coupled to an online integrative tool for the organization, visualization and analysis of inter-microbial networks.
+Given that this
+  In addition to making this tool freely available, we implemented a systematic comparative analysis to determine which steps of the pipeline have the largest influence on the final network, and what choice seems to provide best agreement with the tested mock and synthetic datasets.
+  We believe that these steps will ensure better reproducibility and easier comparison of co-occurrence networks across data-sets.
+We expect that our tool will also be useful for benchmarking future alternative methods, and for ensuring a transparent evaluation of the possible biases introduced by the use of specific tools.
+
+  % TODO: Study the diversity of methods and explore the possiblity of a standardized approach to 16S analysis
   In this study, we propose to build a standardized 16S data-analysis pipeline that will produce robust and reproducible co-occurrence networks from community 16S sequence data.
@@
 
-  There exist many platforms like MG-RAST \cite{Keegan2016}, Qiita \cite{qiita} and tools such as QIIME \cite{Caporaso2010} that provide pipelines for 16S data analysis, but the main motivations behind the construction of our own pipeline are two-fold.
-  Firstly, the existing tools do not focus their analysis on co-occurrence networks and the effect of upstream statistical methods on the inferred networks.
   Secondly, we undertake a comprehensive study of existing methods in order to suggest optimal methods, tools and parameters for users of the pipeline.

### 2020-03-19 sections/methods.tex
Update methods and results
Edits made by Daniel [03-16-2020]

@@
 
+% TODO: Make the methods more mathematically inclined
 \section*{Materials and Methods}
@@
   Real datasets are those sequencing reads obtained from naturally occurring microbial community samples.
-  The current study used healthy stool samples from a fecal microbiome transplant study \cite{Kang2017} and healhty saliva samples from a periodontal disease study~\cite{Chen2018} as real datasets for analysis.
+  The current study used healthy stool samples from a fecal microbiome transplant study \cite{Kang2017} and healthy saliva samples from a periodontal disease study~\cite{Chen2018} as real datasets for analysis.
   The mock community 16S datasets were obtained from mockrobiota~\cite{Bokulich2016}.

### 2020-05-29 sections/abstract.tex
Updates from Overleaf

@@
     In this study, we perform a meticulous analysis of each step of a pipeline that processes 16S sequencing data into a network of microbial associations.
-    % TODO: Instead of this say: We analyze the parts of the process that cause the most variance and what is the result
     Through this process, we map how different choices of algorithms and parameters affect the co-occurrence network, and estimate steps which contribute most significantly to the variance.
@@
     Ultimately, we develop a standardized pipeline that follows these default tools and parameters, but that can also help explore the outcome of any other combination of choices.
-    % TODO: Make this general (do not confine it to disease-related) - niche-related(?)
-    We envisage that this standard pipeline for processing 16S sequencing data into networks of microbial co-occurrences could be used for integrating multiple data-sets, and for generating comparative analyses and consensus networks useful for detecting disease-related patterns.
-\\
-    % TODO: Move package to segrelab github
+    We envisage that this pipeline for processing 16S sequencing data into networks of microbial co-occurrences could be used for integrating multiple data-sets, and for generating comparative analyses and consensus networks useful for detecting disease-related patterns.
+    % TODO: Move package to segrelab github and change name to MiCoNE
     The pipeline is available as a Python package at \href{https://github.com/dileep-kishore/mindpipe}{https://github.com/dileep-kishore/mindpipe}

### 2020-05-29 sections/acknowledgments.tex
Add contributions section
Remove supplementary for now
Updates from Overleaf

@@
 
-  We are grateful for ...
+We are grateful to members of the Segrè lab for helpful discussions and for feedback on the manuscript.
+This work was partially funded by grants from the Defense Advanced Research Projects Agency (Purchase Request No.
+HR0011515303, Contract No.
+HR0011\-15\-C\-0091), the U.S.
+Department of Energy (DE\-SC0012627), the NIH (T32GM100842, 5R01DE024468, R01GM121950 and Sub\_P30DK036836\_P\&F), the National Science Foundation (1457695), the Human Frontiers Science Program (RGP0020/2016), and the Boston University Interdisciplinary Biomedical Research Office.
+[DANIEL, KIRILL: MISSING GRANTS???]
+
+\section*{Contributions}

### 2020-05-29 sections/coverpage.tex
Updates from Overleaf

@@
   \vspace*{\fill}
-  {\huge Inferring microbial co-occurrence networks from 16S data: A systematic evaluation}\\[0.4cm]
+  {\huge Inferring microbial co-occurrence networks from amplicon data: a systematic evaluation}\\[0.4cm]
   % {\huge  2. Investigating the best practices for inference of microbial co-occurrence networks from 16S data}\\[0.4cm] % Title of your document
@@
   \textsuperscript{4}Department of Biology, Boston University, Boston, Massachusetts, USA\\
-  \textsuperscript{$\dagger$}Correspondence should be sent to \href{mailto:korolev.bu.edu}{korolev.bu.edu} or \href{mailto:dsegre.bu.edu}{dsegre.bu.edu}\\
+  \textsuperscript{$\dagger$}Correspondence should be sent to \href{mailto:korolev@bu.edu}{korolev@bu.edu} or \href{mailto:dsegre@bu.edu}{dsegre@bu.edu}\\
 

### 2020-05-29 sections/discussion.tex
Updated draft of manuscript
Updates from Overleaf

--- 2018-07-27 sections/discussion.tex
@@
 
-  Discussion
+% General statements
+Co-occurrence associations in microbial communities help identify important interactions that drive microbial community structure and organization.
+Our analysis shows that networks generated using different combinations of tools and approaches can look significantly different from each other, highlighting the importance of a clear assessment of the source of variability and of tools that provide the most robust and accurate results.
+Our newly developed integrated software for the inference of co-occurrence networks from 16S rRNA data, \ac{micone}, constitutes a freely customizable and user friendly pipeline that allows users to easily test combinations of tools and to compare networks generated by multiple possible choices.
+Importantly, in addition to revisiting the test cases presented in this work, users will be able to explore the effect of various tool combinations on their own datasets of interest.
+The \ac{micone} pipeline is built in a modular fashion.
+Its plug-and-play architecture will make it possible for users to add new tools and steps, either from existing packages, or from those that were not examined in the present work and those that will be developed in the future.
+
+The main outcome of this work is thus two-fold: on one hand we transparently reveal the dependence of co-occurrence networks on tool and parameter choices, making it possible to more rigorously assess and compare existing networks.
+On the other hand, we take advantage of our spectrum of computational options and the availability of mock and synthetic datasets, to suggest a default standard setting, and a consensus approach, likely to yield networks that are robust across multiple tool/parameter choices.
+
+An important caveat related to this last point is the fact that our conclusions are based on the specific datasets used in our analysis.
+While our datasets cover a relatively broad spectrum of biomes and sequencing pipelines, datasets that have drastically different distributions may require a re-assessment of the best settings through our pipeline.
+
+It is worth pointing out some additional more specific conclusions stemming from the individual steps of our analysis.
+
+The different denoising/clustering methods differ mostly in their identification of sequences that are in low abundances.
+Hence, they do not have much of an impact on the inferred co-occurrence networks when the sequences of low abundance are removed.
+However, comparison of inferred and expected reference sequences and their abundances in mock community datasets has allowed us to identify \ac{dada2} as the method which best recapitulates the expected sequence composition.
+There are newer recently published methods like dbOTU3~\cite{Olesen2017} as well as older popular methods like mothur~\cite{Schloss2009} that have not been included in the study, and are possible avenues for future investigation.
+
+The choice of taxonomy database was found to be the most important factor in the inference of a microbial co-occurrence network, contributing $\sim20\%$ of the total variance.
+The frequent changes in the taxonomy nomenclature coupled with the frequency of updates to the various 16S reference databases create inherent differences \cite{Balvociute2017} in taxonomy hierarchies in these databases.
+Our analysis did not reveal one particular reference database to perform better in all scenarios.
+We suggest that that choice of the database should be made based on possible reported or inferred biases in the representation of given biomes in a specific databases \cite{Balvociute2017}.
+The default reference database in the pipeline is the NCBI 16S RefSeq database as it is more frequently updated and is most compatible with the blast+ query tool.
+We also enable users to use custom databases \cite{Griffen2011,Ritari2015} with the blast and naive bayes classifiers that are incorporated into the pipeline (from \ac{qiime2}).
+
+Filtering out taxa that are present in low abundances in all samples did not increase (in most datasets tested) the proportion of taxa in common between taxonomy tables generated using different reference databases.
+However, we do observe that the reduction in the number of taxa leads to better agreement in the networks inferred through different methods.
+Moreover, filtering is necessary in order to increase the power in tests of significance when the number of taxa is much greater than the number of samples.
+
+The networks generated by different network inference methods show considerable differences in edge-density and connectivity.
+One reason for this is the underlying assumptions regarding sparsity, distribution and compositionality that the algorithms make.
+The consensus network created by merging the networks inferred using the different network inference methods enables the creation of a network whose links have evidence based on multiple inference algorithms.
+
+% \hl{Other factors that play an important role in network inference} \\
+% Differences in environments of the samples could lead to the inference of spurious interactions [ref].
+
+Exploring the effects of these combinations of methods on the resultant networks is difficult and inconvenient since different tools differ in their input and output formats and require inter-converting between the various formats.
+The pipeline facilitates this comparative exploration by providing a variety of modules for inter-conversion between various formats, and by allowing easy incorporation of new tools as modules.
+
+We envision that \ac{micone}, and the underlying tools and databases that help process amplicon sequencing data into co-occurrence networks, will be increasingly useful towards building large comparative analyses across studies.
+By having a unified transparent tool to compute networks, it will be possible to reprocess available 16S datasets to obtain networks that are directly comparable to each other.
+Furthermore, even in the analysis of published networks across studies and processing methods, \ac{micone} could help understand underlying biases of each network, which could in turn be taken into account upon making cross-study comparisons.
+
+% TODO: Uses in MIND, exploration, standardization, reproducibility

### 2020-05-29 sections/introduction.tex
Updates from Overleaf

@@
 
-  These data-sets can be used to infer networks of co-occurrence relationships between microorganisms in the samples.These networks have microbial taxa as nodes, and edges that represent the frequent co-occurrence (or negative correlations) across different datasets.
+  These data-sets can be used to infer networks of co-occurrence relationships between microorganisms in the samples.
+  These networks have microbial taxa as nodes, and edges that represent the frequent co-occurrence (or negative correlations) across different datasets.
   Co-occurrence networks can in principle be used to identify important interactions driving the dynamics of the microbial community, including emergent ecosystem-level properties such as ecosystem robustness and modularity.
@@
   One of the limitations of this type of analysis is that 16S sequencing data is limited by many factors such as resolution, sequencing depth, compositional nature, sequencing errors and copy number variations.
-
   Many different analysis methods have been developed \cite{Callahan2016,Amir2017,Friedman2012,Kurtz2015} to counter these limitations and remove statistical artifacts, leading to the existence of a myriad of tools for each step of the data analysis workflow.
@@
 
-  In this study, we present a standardized 16S data-analysis pipeline that produces robust and reproducible co-occurrence networks from community 16S sequence data, and allow users to interactively explore how the network would change upon using different alternative tools and parameters at each step.
+  In this study, we present a standardized 16S data-analysis pipeline called \ac{micone} that produces robust and reproducible co-occurrence networks from community 16S sequence data, and allow users to interactively explore how the network would change upon using different alternative tools and parameters at each step.
+  % TODO: Link or describe MIND here
   Our pipeline is coupled to an online integrative tool for the organization, visualization and analysis of inter-microbial networks.
-Given that this
   In addition to making this tool freely available, we implemented a systematic comparative analysis to determine which steps of the pipeline have the largest influence on the final network, and what choice seems to provide best agreement with the tested mock and synthetic datasets.
@@
   We expect that our tool will also be useful for benchmarking future alternative methods, and for ensuring a transparent evaluation of the possible biases introduced by the use of specific tools.
-
-  % TODO: Study the diversity of methods and explore the possiblity of a standardized approach to 16S analysis
-  In this study, we propose to build a standardized 16S data-analysis pipeline that will produce robust and reproducible co-occurrence networks from community 16S sequence data.
-  We do this by conducting a systematic study of the various methods currently available.
-  We compare different methods to each other and additionally use mock microbial community datasets and synthetically generated reads to benchmark them.
-
-  Secondly, we undertake a comprehensive study of existing methods in order to suggest optimal methods, tools and parameters for users of the pipeline.
-  We believe that these steps will ensure better reproducibility and easier comparison of networks across data-sets.
-  We call our pipeline mindpipe, which derives its origin from the \ac{mind}.

### 2020-05-29 sections/methods.tex
Updates from Overleaf

--- 2019-06-05 sections/methods.tex
@@
   \vspace{-5mm}
-  The study uses three kinds of 16S rRNA sequencing datasets \- real datasets, mock datasets and synthetic datasets.
-  Real datasets are those sequencing reads obtained from naturally occurring microbial community samples.
-  The current study used healthy stool samples from a fecal microbiome transplant study \cite{Kang2017} and healhty saliva samples from a periodontal disease study~\cite{Chen2018} as real datasets for analysis.
-  The mock community 16S datasets were obtained from mockrobiota~\cite{Bokulich2016}.
-  The mock datasets used for this study were labelled mock4, mock12 and mock16 datasets.
+  The study uses three kinds of 16S rRNA sequencing datasets: real datasets, mock datasets and synthetic datasets.
+  Real datasets are collections of sequencing reads obtained from naturally occurring microbial community samples.
+  The current study used healthy stool samples from a fecal microbiome transplant study \cite{Kang2017} and healthy saliva samples from a periodontal disease study~\cite{Chen2018} as real datasets for analysis.
+  The mock community 16S datasets are real sequencing data obtained for artificially assembled collections of species in known proportions.
+The mock datasets used for this study, obtained from mockrobiota~\cite{Bokulich2016}, were labelled mock4, mock12 and mock16.
+  The mock4 community is composed of 21 bacterial strains.
+Two replicate samples from mock4  contain all species in equal abundances, and two additional replicate samples contain the same species in unequal abundances.
+  The mock12 community is composed of 27 bacterial strains that include closely related taxa with some pairs having only one to two nucleotide difference from another.
+  The mock16 community is composed of 49 bacteria and 10 archea, all represented in equal amount.
   The synthetic datasets were generated using an artificial read simulator called ART~\cite{Huang2012}.
-  Three different microbial composition profiles were used as input
-  Synthetic reads were generated using a soil and water microbiome composition profiles from the Earth Microbiome Project~\cite{Thompson2017} and healthy gut microbiome project from the Human Microbiome Project~\cite{HumanMicrobiomeProjectConsortium2012}
-  The scripts used to generate the synthetic data are in the \hl{scripts folder of the respository}
+  Three different microbial composition profiles were used as input; reads were generated using a soil and water microbiome composition profiles from the Earth Microbiome Project~\cite{Thompson2017} and healthy gut microbiome project from the Human Microbiome Project~\cite{HumanMicrobiomeProjectConsortium2012}.
+  The reads are simulated using the NCBI RefSeq database as the reference sequence pool and the "art\_illumina" sequence profile with a mutation rate of 2\%.
+  The scripts used to generate the synthetic data are in the scripts folder of the respository (\href{https://github.com/dileep-kishore/mind-pipeline-paper}{https://github.com/dileep-kishore/mind-pipeline-paper}).
 
-  \subsection*{mindpipe}
+  \subsection*{\ac{micone}}
 
   \vspace{-5mm}
-  The flowchart describing the workflow of the 16S data-analysis pipeline is as shown in Figure \ref{fig:figure1}.
-  The pipeline integrates many publicly available tools as well as custom Python modules and scripts to extract co-occurrence associations from 16S sequence data.
-  The inputs to the pipeline by default are the raw community 16S rRNA sequence reads, but it can also additionally be configured to use trimmed sequences, \ac{otu} tables and other types of intermediate data.
+  The flowchart describing the workflow of \ac{micone} (\acl{micone}), our complete 16S data-analysis pipeline, is shown in Figure \ref{fig:figure1}.
+  The pipeline integrates many publicly available tools as well as custom R or Python modules and scripts to extract co-occurrence associations from 16S sequence data.
+Each of these tools corresponds to a distinct R or python module that recapitulates the relevant analyses.
+All such individual modules are available as part of the \ac{micone} package.
+  The inputs to the pipeline by default are the raw community 16S rRNA sequence reads, but the software can be alternatively configured to use trimmed sequences, \ac{otu} tables and other types of intermediate data.
   The final output of the pipeline is the inferred network of co-occurrence relationships among the microbes present in the samples.
 
-  The mindpipe pipeline provides both a Python API as well as a command-line interface and only requires a single configuration file.
+  The \ac{micone} pipeline provides both a Python API as well as a command-line interface and only requires a single configuration file.
   The configuration file lists the inputs, output and the steps to be performed during runtime, along with the parameters to be used (if different from defaults) for the various steps.
   Since the entire pipeline run-through is stored in the form of a text file (the configuration file), subsequent runs are highly reproducible and changes can be easily tracked using version control.
-  It uses the nextflow workflow manager~\cite{Tommaso2015} under the hood and facilitates the usage of the pipeline on the local machine, cluster or cloud with minimal configuration change.
+  It uses the nextflow workflow manager~\cite{Tommaso2015} under the hood, making it readily usable on local machines, cluster or cloud with minimal configuration change.
   It also allows for automatic parallelization of all possible processes, both within and across samples.
-  The pipeline is designed to be modular, where each tool or method is organized into modules which can be easily modified or replaced.
-  This also simplifies the process of adding new tools (refer to modules section in the mindpipe documentation).
+  The pipeline is designed to be modular: each tool or method is organized into modules which can be easily modified or replaced.
+  This modular architecture simplifies the process of adding new tools (refer to modules section in the \ac{micone} documentation).
   In addition to the Python package, the entire pipeline has been containerized into a Docker~\cite{Merkel1994} image (\hl{dockerhub link}) for easy deployment and setup.
@@
 
-  \subsection*{Denoising and Clustering}
+  \subsection*{Denoising and Clustering (DC)}
   \vspace{-5mm}
   This module deals with processing the raw 16S sequence data into \ac{otu} or \ac{esv} count tables.
-  It consists of three processes:
-  \begin{itemize}
-    \item The quality control process handles the demultiplexing and quality control steps such as trimming adapters and trimming low-quality nucleotide stretches from the sequences.
-    \item The denoise/cluster process handles the conversion of the demultiplexed, trimmed sequences into \ac{otu} or \ac{esv} count tables (some methods perform clustering and taxonomy assignment in the same step).
-    \item The chimera checking process handles the removal of chimeric sequences created during the \ac{pcr} step.
-  \end{itemize}
-  The output of this module is a counts matrix that describes the number of reads of a particular \ac{otu} or \ac{esv} (rows of the matrix) present in each sample (columns of the matrix).
-  The pipeline incorporates open reference clustering, closed reference clustering and de novo clustering methods from \ac{qiime1}~\cite{Caporaso2010} and denoising methods from \ac{dada2}~\cite{Callahan2016} and Deblur~\cite{Amir2017}.
+  It consists of the following processes: quality control, denoising (or clustering) and chimera checking.
+  The quality control process handles the demultiplexing and quality control steps such as trimming adapters and trimming low-quality nucleotide stretches from the sequences.
+  The denoise/cluster process handles the conversion of the demultiplexed, trimmed sequences into \ac{otu} or \ac{esv} count tables (some methods, like closed reference and open reference clustering, perform clustering and taxonomy assignment in the same step).
+  The chimera checking process handles the removal of chimeric sequences created during the \ac{pcr} step.
+  The output of this module is a matrix of counts, that describes the number of reads of a particular \ac{otu} or \ac{esv} (rows of the matrix) present in each sample (columns of the matrix).
+  The options currently available in the pipeline for denoising and clustering: open reference clustering, closed reference clustering and de novo clustering methods from \ac{qiime1}~\cite{Caporaso2010} and denoising methods from \ac{dada2}~\cite{Callahan2016} and Deblur~\cite{Amir2017}.
+  The quality filtering and chimera checking tools are derived from those used in \ac{qiime2} and \ac{dada2}.
 
-  \subsection*{Taxonomy Assignment}
+  \subsection*{Taxonomy Assignment (TA)}
   \vspace{-5mm}
   This module deals with assigning taxonomies to either the representative sequences of the \ac{otu}s or directly to the \ac{esv}s.
-  In order to assign taxonomies to a particular sequence we use:
-  \begin{itemize}
-    \item A taxonomy database, which is a collection of 16S sequences of known organisms.
-    \item A query tool, which allows one to compare a sequence of interest to all the sequences in the database to find the best match.
-    \item Consensus method
-  \end{itemize}
+  In order to assign taxonomies to a particular sequence we need a taxonomy database and a query tool.
+  The taxonomy database contains the collection of 16S sequences of micro-organisms of interest and the query tool allows one to compare a sequence of interest to all the sequences in the database to identify the best matches.
+Finally, a consensus method is used to identify the most probable match from the list of best matches.
   The pipeline incorporates \ac{gg}~\cite{DeSantis2006}, SILVA~\cite{Quast2012} and the \ac{ncbi}~\cite{Sayers2009} databases for taxonomy assignment and the Naive Bayes classifier from \ac{qiime2} and \ac{ncbi} blast as the query tools (from \ac{qiime2}).
-  The consensus algorithm used is also defaults used by the classifiers in \ac{qiime2}.
+  The consensus algorithm used is the default method used by the classifiers in \ac{qiime2}.
 
-  % TODO: Add Gabriel's new stuff here
-  \subsection*{OTU and ESV Processing}
+  % TODO: Add references and basic equations
+  \subsection*{OTU and ESV Processing (OP)}
   \vspace{-5mm}
-  This module deals with normalization, filtering and applying transformations on the \ac{otu} or \ac{esv} counts matrix.
-  \begin{itemize}
-    \item Rarefaction is a normalization technique used to overcome the bias that might arise due to variable sampling depth in different samples.
+  This module deals with normalization, filtering and applying transformations to the \ac{otu} or \ac{esv} counts matrix.
+  Rarefaction is a normalization technique used to overcome the bias that might arise due to variable sampling depth in different samples.
 This is performed either by sub-sampling or by normalization of the matrix to the lowest sampling depth.
-    \item Filtering is performed to remove samples or features (\ac{otu}s or \ac{esv}s) from the count matrix that are sparse.
+  Rarefaction is usually followed by filtering, which is performed to remove samples or features (\ac{otu}s or \ac{esv}s) from the count matrix that are sparse.
 In order to determine the filtering threshold we fix the number of samples and correlation detection power needed and determine the number of features to be used.
-    \item Transformations are performed in order to correct for and overcome the compositional bias that is inherent in a counts matrix (in most cases this is handled by the network inference algorithm).
-  \end{itemize}
+  Finally, transformations are performed in order to correct for and overcome the compositional bias that is inherent in a counts matrix (in most cases this is handled by the network inference algorithm).
 
-  \subsection*{Network Inference}
+  % TODO: Add reference for bootstrapping
+  \subsection*{Network Inference (NI)}
   \vspace{-5mm}
-  This module deals with the inference of co-occurrence associations from the \ac{otu} or \ac{otu} counts matrix.
-  These associations can be represented as a network, with nodes representing taxonomies and edges the association between them.
-  \begin{itemize}
-    \item A null model is created by re-sampling and bootstrapping the correlation/interaction matrix
-    \item Significance of these associations is determined by calculating pvalues against this null model.
-  \end{itemize}
-  The pipeline includes Pearson, Spearman and \ac{sparcc}~\cite{Friedman2012} as the pairwise correlation metrics, and \ac{spieceasi}~\cite{Kurtz2015}, \ac{mldm}~\cite{Yang2017} and \ac{magma} (\hl{in biorxiv}) as the direct association metrics.
+  This module deals with the inference of co-occurrence associations from the \ac{otu} or \ac{esv} counts matrix.
+  These associations can be represented as a network, with nodes representing taxonomies of the micro-organisms and edges representing the association between them.
+  A null model is created by re-sampling and bootstrapping the correlation/interaction matrix and is used to calculate the significance of the inferred associations by calculating the p-values against this null model.
+  The pipeline includes Pearson, Spearman and \ac{sparcc}~\cite{Friedman2012} as the pairwise correlation metrics, and \ac{spieceasi}~\cite{Kurtz2015}, \ac{mldm}~\cite{Yang2017} and \ac{magma} \hl{in biorxiv [ADD REF]} as the direct association metrics.
+  The empirical Browns method \hl{ref ADD REF} is used for combining p-values from the various methods to obtain a consensus p-value, which is used to create the consensus network.
 
-  \subsection*{Statistical Analysis}
+  \subsection*{Network Variability}
+  \vspace{-5mm}
+  In order to compare across different networks, and analyze the degree of variability induced by the choice of different modules and parameters, we organized multiple networks into a single mathematical structure that we could use for linear regression.
+In particular, we transformed the adjacency matrix of each co-occurrence network into a vector.
+We then merged the networks generated from all possible combinations of tools into a table (N, see below) in which each column represents one network.
 
-  \textbf{Network Variability}:
-  The adjacency matrices of the networks generated from all possible combinations of tools are merged.
-  The variance contributed by each step of the pipeline is calculated for every connection in the merged network using ANOVA.
+  \begin{equation*}
+    N = \begin{bmatrix}
+     edge_{1,1} & edge_{2,1} & \cdots  & edge_{n, 1} \\ 
+     edge_{1,2} & edge_{2,2} & \cdots  & edge_{n, 2} \\ 
+     \vdots & \vdots & \vdots  & \vdots \\ 
+     edge_{1,n} & edge_{2,n} & \cdots  & edge_{n, n}
+    \end{bmatrix}      
+  \end{equation*}
+
+In other words, $N$ is the merged table and each column $N_i$ is the vector representation of one of the networks.
+
+We next use linear regression to express each network Ni as a linear function of categorical variables that describe the possible options in each of the first three steps of the pipeline.
+
+  % TODO: Explain the categorical linear model and ANOVA better
+
+In particular, we infer parameters $\alpha_i$ such that:
+   \begin{equation*}
+       N_i = \sum_{j=1}^5 \left( \alpha^{DC(j)}_i.\delta^{DC(j)}_i \right) +
+             \sum_{j=1}^3 \left( \alpha^{TA(j)}_i.\delta^{TA(j)}_i \right) +
+             \sum_{j=1}^2 \left( \alpha^{OP(j)}_i.\delta^{OP(j)}_i \right) +
+             \epsilon_i
+   \end{equation*}
+
+   $N_i = \mathbf{\alpha_i^{DC}} \cdot \mathbf{ \delta_i^{DC}}$
+
+   where, $\alpha_i$ are the coefficients of the regression, $\epsilon_i$ are the residuals and $\delta_i$ are the indicator variables that correspond to the processes utilized in the pipeline used to create the network $N_i$; for example, $\delta^{DC(1)}_i = 1$ if the DC(1) process was used in the generation of the network $N_i$ .
+   Here, (i) DC(1) = "closed reference", DC(2) = "open reference", DC(3) = "de novo", DC(4) = "dada2", DC(5) = "deblur"; (ii)  TA(1) = "GreenGenes", TA(2) = "SILVA", TA(3) = "NCBI"; (iii) OP(1) = "no filtering", OP(2) = "filtering".
+
+  The variance contributed by each step of the pipeline is calculated for every connection in the merged table through ANOVA using the Python statsmodels package and is shown in Figure~\ref{fig:figure2}B.
   The total variance for the network is calculated by adding the variances for each connection.
-  The PCA analysis is also performed on the merged network.
-
-  \textbf{Power calculations:}
-
-  \textbf{P-value merging}
-  The empirical Browns method is used for combining p-values from the various methods to obtain a consensus pvalue, which is used to create the consensus network.
+  The PCA analysis is also performed on the merged table to generate Figure~\ref{fig:figure2}C.
 
   \subsection*{Code and Data Availability}
-
-  Links here
+  pipeline: \href{}{} \\
+  data and scripts related to the paper: \href{}{}

### 2020-05-29 sections/results.tex
Updates from Overleaf

@@
 
-  \subsection*{Mindpipe}
+  \subsection*{\acl{micone} (\acs{micone})}
 
-  We have developed mindpipe, a flexible and modular pipeline for 16S data analysis, that allows us to infer co-occurence networks from 16S sequencing data.
+  We have developed \ac{micone}, a flexible and modular pipeline for 16S amplicon sequencing rRNA data (hereafter mentioned simply as 16S data) analysis, that allows us to infer microbial co-occurence networks.
   It incorporates various popular, publicly available tools as well as custom Python modules and scripts to facilitate inference of co-occurrence networks from 16S data.
-  It enables the user to process 16S sequence data or count matrices using different combinations of the incorporated tools for intermediate analyses as well as inference of co-occurrence networks.
-  The pipeline workflow is shown in Figure~\ref{fig:figure1}.
-  Each process in the pipeline supports alternate tools for performing the same task, users can use the configuration file to change these values.
-  We perform a systematic analysis of each step of the pipeline to determine the set of default options for the pipeline.
-  We also provide the users with a set of guidelines to facilitate tool selection as appropriate for their data.
+Using \ac{micone} one can obtain co-occurrence networks by applying to 16S data (or to already processed taxonomic count matrices) any combination of the available tools.
+The effects of changing any of the intermediate step can be monitored and evaluated in terms of its final network outcome, as well as on any of the intermediate metrics and data outputs.
+  The \ac{micone} pipeline workflow is shown in Figure~\ref{fig:figure1}.
+  The different steps for going from 16S data to co-occurrence networks can be grouped into four major modules; (i) the denoising and clustering (DC) step, which handles denoising of the raw 16S sequencing data into representative sequences; (ii) the taxonomy assignment (TA) step that assigns taxonomic labels to the representative sequences; (iii) the OTU processing (OP) step that filters and transforms the taxonomy abundance table; and finally (iv) the network inferences (NI) step which infers the microbial co-occurrence network.
+  Each process in the pipeline supports alternate tools for performing the same task (see Methods and Figure~\ref{fig:figure1}).
+A centralized configuration file contains all the specifications for what modules to be used in the pipeline, and can be modified by the user to choose the desired set of tools.
+  In what follows, we perform a systematic analysis of each step of the pipeline to estimate how much the final co-occurrence network depends on the possible choices at each step.
+We also evaluate a large number of tool combinations to determine a set of recommended default options for the pipeline and provide the users with a set of guidelines to facilitate tool selection as appropriate for their data.
 
@@
     \centering
-    \includegraphics[width=0.7\linewidth]{figure1.pdf}
+    \includegraphics[width=0.69\linewidth]{figure1.pdf}
     \caption{
@@
       Each step incorporates several processes, each of which in turn have several alternate algorithms for the same task (indicated by the text to the right of the blue boxes).
-      The text along the arrows describes the data that is being passed from one step to another.
+      The text along the arrows describes the data that is being passed from one step to another. For details on each process and data types, see Methods.
     }
@@
   % TODO: Talk about the datasets: real, mock and synthetic (with references)
-  In order to objectively determine the best method we use both mock data from mockrobiota~\cite{Bokulich2016} as well as, synthetically generated reads from an Illumina read simulator called ART~\cite{Huang2012}.
+  In addition to the algorithmic pipeline itself, our analysis involves two types of data: The first type consists of sets of 16S sequencing data for real communities (see Methods).
+The second are datasets artificially created for the specific goal of helping evaluate computational analysis tools.
+In particular, in order to objectively compare, to the extent possible, how well each step in \ac{micone} best captures the underlying data, we use both mock data from mockrobiota~\cite{Bokulich2016} as well as, synthetically generated reads from an Illumina read simulator called ART~\cite{Huang2012}.
   The mock datasets selected were labelled mock4, mock12 and mock16 in the mockrobiota database and consist of sequenced 16S data from mock communities of various bacteria.
   These datasets were selected since they contained the expected compositions along with the reference sequences for the organisms in the mock community.
-  The synthetic reads were simulated using the three different taxonomy distribution profiles obtained from soil, water and Crohn's disease datasets.
+  The synthetic reads were simulated using the three different taxonomy distribution profiles obtained from soil, water and healthy stool microbiome datasets \hl{[ref]}.
   Reference sequences were generated using \ac{ncbi} and the decard tool \cite{Golob2017} for these taxonomy profiles.
-  Detailed information on the mock communities and all the other settings used to generate the sequence data are provided in the methods section.
+  Detailed information on the mock communities and the settings used to generate the synthetic data are provided in the Methods section.
 
@@
 
-  In order to analyze the effect of different statistical methods on the inferred co-occurrence networks, we generate co-occurrence networks using all possible combinations of methods and estimate the variability in the networks due to each choice.
-  This analysis is performed while keeping the network inference algorithm the same throughout the analysis.
-  The effects of various steps on the final co-occurrence network is estimated by building a linear model of the edges of the network as a function the various step in the analysis pipeline.
-  Figure \ref{fig:figure2} shows the fraction of total variation among the co-occurrence networks as contributed by each of the four main steps of the pipeline.
-  The 16S reference database contributes to the most variation ($\sim25\%$) in the networks, and the networks can be clearly separated based on the database used.
-  This indicates that the taxonomy assigned to the reference sequences drastically alters the co-occurrence network and this change is much more significant than how the reference sequences themselves are identified (in the denoise/cluster step).
-  We believe that this clustering is due to the mislabelling of constitutive taxa that are present in high abundance in the community.
+  % TODO: Summarized sections
+  In order to analyze the effect of different statistical methods on the inferred co-occurrence networks, we generated co-occurrence networks using all possible combinations of methods and estimated the variability in the networks due to each choice.
+  This analysis is performed while keeping the network inference algorithm (NI step) the same throughout the analysis.
+  The effects of various steps on the final co-occurrence network is estimated by building a linear model of the edges of the network as a function the various step in the analysis pipeline (see Methods).
+  Figure \ref{fig:figure2}B, shows the fraction of total variation among the co-occurrence networks as contributed by the first three steps of the pipeline.
+In other words, each point corresponds to a different combination of tools, and captures how much the final network is affected by such choice.
+  The 16S reference database contributes the most ($\sim25\%$) to variation in the networks.
+This is also reflected in the fact that the networks can be clearly separated based on the database used (Figure \ref{fig:figure2}B).
+  This indicates that the taxonomy assigned to the reference sequences drastically alters the co-occurrence network and this change is much more significant than how the reference sequences themselves are identified (in the DC step).
+  We believe that the grouping by taxonomy assignment into clusters (..... in Fig. XXX) is due to the mislabelling of constitutive taxa that are present in high abundance in the community.
   The residual variation can be seen as an artifact that arises due to the effect on the resultant network when multiple steps are changed at the same time.
-  Another interesting observation has been that the dissimilarity between the networks decreases when the low abundance OTUs are removed from the network.
+  Another interesting observation is that the dissimilarity between the networks decreases when the low abundance \ac{otu}s are removed from the network. This is visible in terms of ...
   These results enable us to believe that the most important criterion for accurate comparative analyses of co-occurrence networks is the taxonomy reference database.
 
-  \begin{figure}
+  \begin{figure}[H]
     \centering
     \includegraphics[width=0.85\linewidth]{figure2.pdf}
+  \end{figure}
+  \begin{figure}[!t]
+    \centering
       \caption{
@@
   Denoising and clustering are commonly carried out to generate representative sequences from the raw 16S sequencing data and to obtain the \ac{otu}/\ac{esv} tables (counts of these representative sequences for each sample).
-  In order to compare the \ac{otu} tables generated by these tools we process the same 16S sequencing reads (healthy samples from a fecal microbiome transplant study [ref]) using 5 different methods \- open-reference clustering, closed-reference clustering, denovo clustering, \ac{dada2}~\cite{Callahan2016} and Deblur~\cite{Amir2017}.
+  In order to compare the \ac{otu} tables generated by different tools we processed the same 16S sequencing reads (healthy samples from a fecal microbiome transplant study~\cite{Kang2017}) using 5 different methods:  open-reference clustering, closed-reference clustering, denovo clustering, \ac{dada2}~\cite{Callahan2016} and Deblur~\cite{Amir2017}.
   The first three methods are from the \ac{qiime1}~\cite{Caporaso2010} package.
-  We find that there is good agreement in the \ac{otu}/\ac{esv} tables when different combinations of methods are used to generate them.
+  We find that there is good agreement in the \ac{otu}/\ac{esv} tables when different combinations of methods are used to generate them (Supplementary Figure~\ref{fig:figureS1}).
 
@@
   The weighted UniFrac distance metric takes into account the counts of the representative sequences, whereas the unweighted UniFrac distance metric does not and hence gives equal weights to each sequence.
-  From Figure~\ref{fig:figure3}A we can observe that the representative sequences generated by the different methods are similar to each other when weighted by their abundance.
-  Figure~\ref{fig:figure3}B on the other hand shows an increase in dissimilarity between each pair of methods suggesting that the methods might differ in the sequences of low abundance.
-  %% TODO: Update this
+  From Figure~\ref{fig:figure3}A one can see that the representative sequences generated by the different methods are similar to each other when weighted by their abundance.
+  Figure~\ref{fig:figure3}B on the other hand shows an increase in dissimilarity between each pair of methods suggesting that the methods might differ in the treatment of sequences of low abundance.
   In order to verify this claim, for each of these methods we use the \ac{gg} taxonomy database to assign taxonomies to the representative sequences.
@@
   The \ac{esv} tables generated by methods that perform denoising are very similar to each other $\sim0.91$ and the \ac{otu} tables generated by the clustering methods are very similar to each other $\sim0.9$, but results of denoising and clustering are highly uncorrelated with each other $\sim0.4$ (Figure \ref{fig:figureS1}C).
-  % The common core network (network made up of edges that are predicted by all methods) is very small compared to the total number of edges predicted by all the methods (Figure \ref{fig:denoise_network}).
-  % In these results, only the denoising method was changed (database used: \ac{gg} and network inference: \ac{sparcc}).
 
   These comparisons only elucidate the pairwise similarity or dissimilarity of a pair of methods.
-  In order to determine the tool that most accurately recapitulates the reference sequences in the samples we use the 16S sequences from the mock datasets.
-  The mock community datasets were processed using all the methods and the predicted representative sequences were compared with expected representative sequences and their distribution.
-  The results (Figure~\ref{fig:figure3}C and D) show that the performance of the methods are similar to each other as in the case of the real dataset, but, are very different from the expected sequence distribution.
-  This result is more exaggerated in the case of the unweighted UniFrac metric where some of the datasets show a very high deviation from the expected sequences.
+  In order to determine the tool that most accurately recapitulates the reference sequences in the samples, we use the 16S sequences from the mock datasets.
+  In particular, we used the pipeline to process mock community datasets using each of the possible methods included for this step.
+We next compared predicted representative sequences with expected representative sequences and their distribution.
+  The results (Figure~\ref{fig:figure3}C and D) show that, for the mock datasets, the different methods perform similar to each other, exactly as observed in the case of the real dataset.
+However, the mock predicted sequence distributions are substantially different from the expected sequence distribution.
+  This result is more exaggerated in the case of the unweighted UniFrac metric, where some of the datasets show a very high deviation from the expected sequences.
   These high deviations are primarily in two of the three datasets that were analyzed and show that the datasets themselves play a big role in the performance of these methods.
@@
   There is no method that clearly outperforms the rest in all datasets.
-  Based on their slightly better performance on the mock datasets, their \textit{de novo} error correcting nature and other previous studies [refs] we recommend either \ac{dada2} or Deblur for users.
-  However, as a result of the unexpected poor performance of Deblur on the synthetic data, the default algorithm in the pipeline was chosen to be \ac{dada2}.
+  Based on their slightly better performance on the mock datasets, their \textit{de novo} error correcting nature and other previous studies [refs], \ac{dada2} and Deblur seem to be in general the most reliable.
+  Given the unexpected poor performance of Deblur on the synthetic data, the default algorithm in the pipeline was chosen to be \ac{dada2} (Supplementary Figure~\ref{fig:figureS3}).
 
@@
       \textbf{(B)} The relatively larger average unweighted UniFrac distance indicates that methods differ in their identification of sequences of low abundance,
-      \textbf{(C, D)} The distributions of the avergage weighted UniFrac distance between the expected sequence profile and the calculated sequence profile in mock datasets.
+      \textbf{(C, D)} The distributions of the avegage weighted UniFrac distance between the expected sequence profile and the calculated sequence profile in mock datasets.
       \textbf{(D)} The distributions of the average unweighted UniFrac distance show that dada2 and Deblur were the best performing methods in most of the datasets.
@@
   Figure~\ref{fig:figure4}A depicts a flow diagram that shows how the top 50 representative sequences (sorted by abundance) are assigned a Genus according to the three different databases.
-  We observe that not only does the assigned Genus composition vary significantly, but, the percentage of unassigned representative sequences (gray) also differ.
+  We observe that not only does the assigned Genus composition vary significantly, but the percentage of unassigned representative sequences (gray) also differ.
   Even the most abundant representative sequence is assigned to an "unknown" Genus in two of the three databases.
-  A representative sequence might be get assigned an "unknown" Genus if either the taxonomy identifier associated with the sequence in the database did not contain a Genus or the more likely case where the database contains multiple sequences that are very similar to the query (representative) sequence and the consensus algorithm (from \ac{qiime2}) is unable to assign one particular Genus at the required confidence.
+  A representative sequence might be assigned an "unknown" Genus for one of two reasons: the first is if the taxonomy identifier associated with the sequence in the database did not contain a Genus; the second (more likely) reason is that the database contains multiple sequences that are very similar to the query (representative) sequence and the consensus algorithm (from \ac{qiime2}) is unable to assign one particular Genus at the required confidence.
   After assigning all the representative sequences to taxonomies we perform a pairwise comparison of the similarity between assignments from different databases at every taxonomic level (Figure~\ref{fig:figure4}B).
@@
   This implies that the taxonomy assignments from each reference database are fairly unique and is one of the reasons for a large differences in the resultant co-occurrence networks generated from different taxonomy databases.
+   % TODO: Update this
+  Supplementary Figure~\ref{fig:figureS4} shows that the top 20 most abundant genera in the three resulting taxonomy composition tables are different.
+  For example, the most abundant genus in the \ac{gg} taxonomy table was \textit{Escherichia} whereas in the SILVA taxonomy table it was \textit{Escherichia-Shigella}.
+  Although these are minor differences, when comparing a large number of taxonomy composition tables these problems are hard to diagnose.
+%   The comparison of all assigned genera instead of the just the top 20 contains the same percentage of matches and mismatches, implying that there does not seem to exist a correlation between abundance and mismatch.
+%   This suggests that the most abundant sequences are not necessarily the ones that are consistently matched to the same taxonomies in the different reference databases.
 
@@
   In order to obtain an absolute measure of accuracy of the taxonomic assignments we use the expected reference sequences from the mock datasets as the query sequences for the databases and the expected taxonomic composition as the standard to compare against (Figure~\ref{fig:figure4}C).
-  Again, we observe that none of the databases perform better than the others.
+  Again, we observe that none of the databases perform better than the others in absolute terms.
 
-  % TODO: Update this
-  Figure~\ref{fig:figureS4} shows that the top 20 most abundant genera in the three resulting taxonomy composition tables are different.
-  For example, the most abundant genus in the \ac{gg} taxonomy table was \textit{Escherichia} whereas in the SILVA taxonomy table it was \textit{Escherichia-Shigella}.
-  Although these are minor differences, when comparing a large number of composition tables these problems are hard to diagnose.
-  The comparison of all assigned genera instead of the just the top 20 contains the same percentage of matches and mismatches, implying that there does not seem to exist a correlation between abundance and mismatch.
-  This suggests that the most abundant sequences are not necessarily the ones that are consistently matched to the same taxonomies in the different reference databases.
+  Given that no database performs better than others against mock datasets, and that databases are almost equally distant from each other in terms of final output, the choice of which database to use should be driven by other reason.
+One user-specific way to choose, would be based on the known representation of taxa for the microbiome of interest (see also Discussion).
+Another reason could be the frequency of updates and the potential for future growth, which prompted us to set \ac{ncbi} as the \ac{micone} standard for taxonomy assignment.
+ In addition to being regularly maintained and updated the \ac{ncbi} database has the advantage that its accuracy of assignments is still comparable to the SILVA and \ac{gg} reference databases that are routinely used as reference databases.
 
-  These results prompted us to turn to \ac{ncbi} for taxonomy assignment.
-  Our reasons for choosing \ac{ncbi} are that it is regularly maintained and updated and accuracy of the assignment is still comparable to the SILVA and \ac{gg} reference databases that are routinely used as reference databases.
-  % \hl{Or combine results from multiple databases?}
-
-  \begin{figure}[h]
+  \begin{figure}[H]
     \centering
-    \includegraphics[width=0.72\textwidth]{figure4.pdf}
+    \includegraphics[width=0.9\textwidth,height=1.2\textwidth]{figure4.pdf}
+  \end{figure}
+  \begin{figure}[!t]
+    \centering
     \caption{
@@
 
-  % TODO: Add reference here
-  The six different network inference methods used in this study are \ac{magma}, \ac{mldm}, \ac{spieceasi}, \ac{sparcc}, Spearman and Pearson.
+  % TODO: Talk about the difference between correlations and associations
+  The six different network inference methods used in this study are \ac{magma}~\hl{in biorxiv [ADD REF]}, \ac{mldm}~\cite{Yang2017}, \ac{spieceasi}~\cite{Kurtz2015}, \ac{sparcc}~\cite{Friedman2012}, Spearman and Pearson.
   These network inference methods into two groups, the first set of methods (Pearson, Spearman, \ac{sparcc}) infer pairwise correlations while the second set infer direct associations (\ac{spieceasi}, \ac{mldm}, \ac{magma}).
 
-  The taxonomy composition table obtained using the \ac{ncbi} reference database is used as the input to these algorithms to infer co-occurrence associations between the microbes.
+  For the analysis presented in this section, we used the taxonomy composition table obtained using the \ac{ncbi} reference database as the input for algorithms that infer co-occurrence associations between the microbes.
   Figure~\ref{fig:figure5}A shows the networks inferred from this dataset using the different inference algorithms.
-  An edge weight threshold of 0.3 was applied to the networks generated from the methods that infer pairwise correlations and no threshold was applied in case of the other networks.
   We can clearly see that the different networks differ vastly in their edge-density and connectivity and even some of the edges in common to these networks have their signs inverted.
+Note, however, that some of these comparisons depend on the threshold that has to be applied to the pairwise correlations methods (currently 0.3, based on [REF]).
   To get a more quantitative idea, we can take a look at the nodes and edges (Figure~\ref{fig:figure5}B) in common between the networks using UpSet plots (only \ac{magma}, \ac{mldm}, \ac{spieceasi}, \ac{sparcc} are used in the comparison).
@@
   Unlike in the previous sections, where were we evaluated the performance of methods on mock datasets, there are no such datasets that contain a set of known interactions for the evaluation of the network inference algorithms.
-  Therefore, we propose the construction of a consensus network (Figure~\ref{fig:figure5}C) involving \ac{magma}, \ac{mldm}, \ac{spieceasi} and \ac{sparcc} by merging the p-values generated from bootstraps of the original taxonomy composition table using the Browns pvalue combining method [ref].
-  The resulting network will have weights reported by all four methods.
+  Therefore, we propose the construction of a consensus network (Figure~\ref{fig:figure5}C) involving \ac{magma}, \ac{mldm}, \ac{spieceasi} and \ac{sparcc} by merging the p-values generated from bootstraps of the original taxonomy composition table using the Browns p-value combining method [ref] (see Methods section).
+  Based on this approach, \ac{micone} reports as default output the consensus network, annotated with weights (correlations for \ac{sparcc} and direct associations for the other methods) for all four methods.
 
-  \begin{figure}[h]
+  \begin{figure}[H]
     \centering
-    \includegraphics[width=0.85\textwidth]{figure5.pdf}
+    \includegraphics[width=0.9\textwidth]{figure5.pdf}
+  \end{figure}
+  \begin{figure}[t!]
+    \centering
     \caption{
@@
 
-  The systematic analysis performed in the previous sections allows us to suggest the set of tools and parameters that produce more accurate and robust co-occurrence networks.
-  These tools and parameters chosen as the defaults for the pipeline and are given in Table~\ref{tab:default_options}.
+  The systematic analyses performed in the previous sections clearly show that the choice of tools and parameters can have a big impact on the final co-occurrence network.
+For some of these choices (e.g.
+\ac{dada2} vs.
+deblur) there is no clear metric to establish a best protocol.
+  For other choices, the mock communities provide an opportunity to select combination of parameters that yield more accurate and robust results.
+  Despite this partial degree of assessment, we wish to suggest a combination of tools and parameters that produce networks that are derived from the combination of tools which performed best on the mock communities, and displayed highest robustness to switching to alternative methods.
+  These tools and parameters are chosen as the defaults for the pipeline and are given in Table~\ref{tab:default_options}.
 
@@
 
-  The recommended tool for the \ac{dc} step are \ac{dada2} or Deblur were chosen based on their accuracy in recapitulating the reference sequences in mock communities and synthetic data.
+  The recommended tool for the \ac{dc} step (\ac{dada2} or Deblur) were chosen based on their accuracy in recapitulating the reference sequences in mock communities and synthetic data.
   The choice of the taxonomy reference database in the \ac{ta} step is dictated largely by the species expected to be present in the sample as well the database used in similar studies if comparison is a goal.
@@
   The abundance threshold at the \ac{op} step is determined automatically based on the number of samples and the required statistical power.
-  Finally, we use the Browns pvalue combining method on the networks generated using \ac{magma}, \ac{mldm}, \ac{spieceasi} and \ac{sparcc} to obtain a final consensus network in the \ac{ni} step.
+  Finally, we use the Browns p-value combining method on the networks generated using \ac{magma}, \ac{mldm}, \ac{spieceasi} and \ac{sparcc} to obtain a final consensus network in the \ac{ni} step.
 
@@
   These results indicate that the biggest differences in networks occur when the reference database or the network inference algorithm are changed.
-  Furthermore, the L1 distance of networks generated by altering one of the steps of the pipeline from the default against the default network (Figure~\ref{fig:figure6}B) show that the biggest deviations from the default network occur when the \ac{ta} and \ac{ni} steps are changed, reinforcing the same results observed in Figure~\ref{fig:figure2}.
-Figure~\ref{fig:figure7} shows the co-occurrence networks inferred for the hard palate for healthy subjects in a periodontal disease study [ref] and the healthy stool microbiome in fecal microbial transplant study [ref].
+  Furthermore, the L1 distance of networks generated by altering one of the steps of the pipeline from the default against the default network (Figure~\ref{fig:figure6}B) shows that the biggest deviations from the default network occur when the \ac{ta} and \ac{ni} steps are changed, reinforcing the same results observed in Figure~\ref{fig:figure2}.
+Figure~\ref{fig:figure7} shows the co-occurrence networks inferred for the hard palate for healthy subjects in a periodontal disease study~\cite{Chen2018} and the healthy stool microbiome in fecal microbial transplant study~\cite{Kang2017}.
 These consensus networks were generated using the default tools and parameters from Table~\ref{tab:default_options}.
@@
     \centering
-    \includegraphics[width=0.9\linewidth]{figure6.pdf}
+    \includegraphics[width=0.85\linewidth]{figure6.pdf}
     \caption{

### 2020-05-29 sections/supplementary_figures.tex
Updates from Overleaf

@@
     \caption{
-      \textbf{The distributions of the avergage weighted UniFrac distance between the expected sequence profile and the calculated sequence profile in the synthetic datasets}.
+      \textbf{The distributions of the average weighted UniFrac distance between the expected sequence profile and the calculated sequence profile in the synthetic datasets}.
       We observe no significant difference between the various methods on the synthetic datasets used for this study.
@@
 
-  \begin{figure}[h]
-    \centering
-    \includegraphics[width=0.9\linewidth]{pdf/all_denoise_reg.pdf}
-    \caption{All pairwise correlations comparing the similarity between different denoising and clustering methods}
-    \label{fig:figureS4}
-  \end{figure}
+%   \begin{figure}[h]
+%     \centering
+%     \includegraphics[width=0.9\linewidth]{pdf/all_denoise_reg.pdf}
+%     \caption*{All pairwise correlations comparing the similarity between different denoising and clustering methods}
+%     \label{fig:figureS4}
+%   \end{figure}
 

### 2020-06-03 sections/results.tex
Changes from overleaf

@@
   These datasets were selected since they contained the expected compositions along with the reference sequences for the organisms in the mock community.
-  The synthetic reads were simulated using the three different taxonomy distribution profiles obtained from soil, water and healthy stool microbiome datasets \hl{[ref]}.
-  Reference sequences were generated using \ac{ncbi} and the decard tool \cite{Golob2017} for these taxonomy profiles.
+  The synthetic reads were simulated using the three different taxonomy distribution profiles obtained from soil, water and healthy stool microbiome datasets \hl{[refs]}.
+  Reference sequences were generated using \ac{ncbi} and the Decard tool \cite{Golob2017} for these taxonomy profiles.
   Detailed information on the mock communities and the settings used to generate the synthetic data are provided in the Methods section.

### 2020-06-18 sections/abstract.tex
Updates to draft

@@
     We envisage that this pipeline for processing 16S sequencing data into networks of microbial co-occurrences could be used for integrating multiple data-sets, and for generating comparative analyses and consensus networks useful for detecting disease-related patterns.
-    % TODO: Move package to segrelab github and change name to MiCoNE
-    The pipeline is available as a Python package at \href{https://github.com/dileep-kishore/mindpipe}{https://github.com/dileep-kishore/mindpipe}
+        The pipeline is available as a Python package at \href{https://github.com/segrelab/MiCoNE}{https://github.com/segrelab/MiCoNE}
   }
@@
 % Insert keywords here
-\keywords{Microbiome, 16S rRNA, Pipeline, Interaction, Network, Correlations, Qiime, Co-occurrence}
+\keywords{Microbiome, 16S rRNA, Pipeline, Interaction, Denoising, Taxonomy, Network Inference, Correlations, Qiime, Co-occurrence, Networks}

### 2020-06-18 sections/discussion.tex
Updates to draft

@@
 We suggest that that choice of the database should be made based on possible reported or inferred biases in the representation of given biomes in a specific databases \cite{Balvociute2017}.
-The default reference database in the pipeline is the NCBI 16S RefSeq database as it is more frequently updated and is most compatible with the blast+ query tool.
+The default reference database in the pipeline is the \ac{ncbi} 16S RefSeq database as it is more frequently updated and is most compatible with the blast+ query tool.
 We also enable users to use custom databases \cite{Griffen2011,Ritari2015} with the blast and naive bayes classifiers that are incorporated into the pipeline (from \ac{qiime2}).
@@
 
-% TODO: Uses in MIND, exploration, standardization, reproducibility
+% TODO: Uses in MIND (?)

### 2020-06-18 sections/introduction.tex
Updates to draft

@@
   With the advancement in DNA sequencing technologies \cite{Narihiro2017} and data processing methods much more information can be extracted from these microbial community samples than ever before.
-  These advances have led to large-scale data collection efforts involving environmental (Earth Microbiome Project) \cite{Thompson2017}, marine (Tara Oceans Project) \cite{Zhang2015} and human-associated microbiota (Human Microbiome Project) \cite{HumanMicrobiomeProjectConsortium2012} and thus led the generation of a wealth of data about these microbial communities.
+  These advances have led to large-scale data collection efforts involving environmental (\acl{emp}) \cite{Thompson2017}, marine (Tara Oceans Project) \cite{Zhang2015} and human-associated microbiota (Human Microbiome Project) \cite{HumanMicrobiomeProjectConsortium2012} and thus led the generation of a wealth of data about these microbial communities.
 
@@
   Conversely, given the lack of comprehensive comparisons between directly observed microbial interactions (e.g.
-from co-culture experiments) and co-occurence networks, there is no straightforward way to determine which set of tools or methods generate the most accurate networks.
+from co-culture experiments) and co-occurrence networks, there is no straightforward way to determine which set of tools or methods generate the most accurate networks.
 
  Previous work has developed popular platforms (like MG-RAST \cite{Keegan2016}, Qiita \cite{qiita}) and tools (such as QIIME \cite{Caporaso2010}) that provide pipelines for 16S data analysis.
-However, existing tools typically are not focused on the effects of upstream statistical methods on the inferred co-occurrence networks.
+ However, existing tools typically are focused on \ac{otu} generation and not on the effects of upstream statistical methods on the inferred co-occurrence networks.
  Furthermore, no organized framework currently exist to systematically analyze and compare existing components of the data-analysis from amplicons to networks.
@@
   In this study, we present a standardized 16S data-analysis pipeline called \ac{micone} that produces robust and reproducible co-occurrence networks from community 16S sequence data, and allow users to interactively explore how the network would change upon using different alternative tools and parameters at each step.
-  % TODO: Link or describe MIND here
+  % TODO: Link or describe MIND here (?)
   Our pipeline is coupled to an online integrative tool for the organization, visualization and analysis of inter-microbial networks.

### 2020-06-18 sections/methods.tex
Updates to draft

@@
   Real datasets are collections of sequencing reads obtained from naturally occurring microbial community samples.
-  The current study used healthy stool samples from a fecal microbiome transplant study \cite{Kang2017} and healthy saliva samples from a periodontal disease study~\cite{Chen2018} as real datasets for analysis.
+  The current study used healthy stool samples from a fecal microbiome transplant study~\cite{Kang2017} and healthy saliva samples from a periodontal disease study~\cite{Chen2018} as real datasets for analysis.
   The mock community 16S datasets are real sequencing data obtained for artificially assembled collections of species in known proportions.
@@
   The synthetic datasets were generated using an artificial read simulator called ART~\cite{Huang2012}.
-  Three different microbial composition profiles were used as input; reads were generated using a soil and water microbiome composition profiles from the Earth Microbiome Project~\cite{Thompson2017} and healthy gut microbiome project from the Human Microbiome Project~\cite{HumanMicrobiomeProjectConsortium2012}.
+  Three different microbial composition profiles were used as input; reads were generated using a soil and water microbiome composition profiles from the \ac{emp}~\cite{Thompson2017} and healthy gut microbiome project from the fecal microbiome transplant study~\cite{Kang2017}.
   The reads are simulated using the NCBI RefSeq database as the reference sequence pool and the "art\_illumina" sequence profile with a mutation rate of 2\%.
-  The scripts used to generate the synthetic data are in the scripts folder of the respository (\href{https://github.com/dileep-kishore/mind-pipeline-paper}{https://github.com/dileep-kishore/mind-pipeline-paper}).
+  The scripts used to generate the synthetic data are in the scripts folder of the respository (\href{https://github.com/segrelab/mind-pipeline-paper}{https://github.com/segrelab/mind-pipeline-paper}).
 
@@
   The output of this module is a matrix of counts, that describes the number of reads of a particular \ac{otu} or \ac{esv} (rows of the matrix) present in each sample (columns of the matrix).
-  The options currently available in the pipeline for denoising and clustering: open reference clustering, closed reference clustering and de novo clustering methods from \ac{qiime1}~\cite{Caporaso2010} and denoising methods from \ac{dada2}~\cite{Callahan2016} and Deblur~\cite{Amir2017}.
-  The quality filtering and chimera checking tools are derived from those used in \ac{qiime2} and \ac{dada2}.
+  The options currently available in the pipeline for denoising and clustering: open reference clustering, closed reference clustering and de novo clustering methods from \ac{qiime1} v1.9.1~\cite{Caporaso2010} and denoising methods from \ac{dada2} v1.14~\cite{Callahan2016} and Deblur v1.1.0~\cite{Amir2017}.
+  The quality filtering and chimera checking tools are derived from those used in \ac{qiime2} v2019.10.0 and \ac{dada2}.
 
@@
   Finally, a consensus method is used to identify the most probable match from the list of best matches.
-  The pipeline incorporates \ac{gg}~\cite{DeSantis2006}, SILVA~\cite{Quast2012} and the \ac{ncbi}~\cite{Sayers2009} databases for taxonomy assignment and the Naive Bayes classifier from \ac{qiime2} and \ac{ncbi} blast as the query tools (from \ac{qiime2}).
+  The pipeline incorporates \ac{gg} 13\_8~\cite{DeSantis2006}, SILVA 132~\cite{Quast2012} and the \ac{ncbi} (16S RefSeq as of Oct 2019)~\cite{Sayers2009} databases for taxonomy assignment and the Naive Bayes classifier from \ac{qiime2} and \ac{ncbi} blast as the query tools (from \ac{qiime2}).
   The consensus algorithm used is the default method used by the classifiers in \ac{qiime2}.
@@
   Rarefaction is a normalization technique used to overcome the bias that might arise due to variable sampling depth in different samples.
-This is performed either by sub-sampling or by normalization of the matrix to the lowest sampling depth.
+  This is performed either by sub-sampling or by normalization of the matrix to the lowest sampling depth \cite{Weiss2015}.
   Rarefaction is usually followed by filtering, which is performed to remove samples or features (\ac{otu}s or \ac{esv}s) from the count matrix that are sparse.
@@
 
-  % TODO: Add reference for bootstrapping
   \subsection*{Network Inference (NI)}
@@
   These associations can be represented as a network, with nodes representing taxonomies of the micro-organisms and edges representing the association between them.
-  A null model is created by re-sampling and bootstrapping the correlation/interaction matrix and is used to calculate the significance of the inferred associations by calculating the p-values against this null model.
-  The pipeline includes Pearson, Spearman and \ac{sparcc}~\cite{Friedman2012} as the pairwise correlation metrics, and \ac{spieceasi}~\cite{Kurtz2015}, \ac{mldm}~\cite{Yang2017} and \ac{magma} \hl{in biorxiv [ADD REF]} as the direct association metrics.
-  The empirical Browns method \hl{ref ADD REF} is used for combining p-values from the various methods to obtain a consensus p-value, which is used to create the consensus network.
+  A null model is created by re-sampling and bootstrapping the correlation/interaction matrix and is used to calculate the significance of the inferred associations by calculating the p-values against this null model \cite{Watts2018}.
+  The pipeline includes Pearson, Spearman and FastSpar v0.0.10 (a faster implementation of \ac{sparcc})~\cite{Watts2018} as the pairwise correlation metrics, and \ac{spieceasi} v1.0.7~\cite{Kurtz2015}, \ac{mldm} v1.1~\cite{Yang2017} and \ac{magma}~\cite{Cougoul2019} as the direct association metrics.
+  The empirical Browns method is used for combining p-values from the various methods to obtain a consensus p-value, which is used to create the consensus network.
 
@@
 
-In other words, $N$ is the merged table and each column $N_i$ is the vector representation of one of the networks.
+  In other words, $N$ is the merged table and each column $N_i$ is the vector representation of one of the networks and each row $L_i$ represents the one particular edge in all networks (assigned 0 if the edge does not exist in the network).
 
-We next use linear regression to express each network Ni as a linear function of categorical variables that describe the possible options in each of the first three steps of the pipeline.
+  We next use linear regression to express each link $L_i$ as a linear function of categorical variables that describe the possible options in each of the first three steps of the pipeline.
 
   % TODO: Explain the categorical linear model and ANOVA better
-
   In particular, we infer parameters $\alpha_i$ such that:
    \begin{equation*}
-       N_i = \sum_{j=1}^5 \left( \alpha^{DC(j)}_i.\delta^{DC(j)}_i \right) +
+       L_i = \sum_{j=1}^5 \left( \alpha^{DC(j)}_i.\delta^{DC(j)}_i \right) +
              \sum_{j=1}^3 \left( \alpha^{TA(j)}_i.\delta^{TA(j)}_i \right) +
@@
    \end{equation*}
-
-   $N_i = \mathbf{\alpha_i^{DC}} \cdot \mathbf{ \delta_i^{DC}}$
 
@@
   \subsection*{Code and Data Availability}
-  pipeline: \href{}{} \\
+  pipeline: \href{https://github.com/segrelab/MiCoNE}{https://github.com/segrelab/MiCoNE} \\
   data and scripts related to the paper: \href{}{}

### 2020-06-18 sections/results.tex
Updates to draft

--- 2020-05-29 sections/results.tex
@@
 
-  We have developed \ac{micone}, a flexible and modular pipeline for 16S amplicon sequencing rRNA data (hereafter mentioned simply as 16S data) analysis, that allows us to infer microbial co-occurence networks.
+  We have developed \ac{micone}, a flexible and modular pipeline for 16S amplicon sequencing rRNA data (hereafter mentioned simply as 16S data) analysis, that allows us to infer microbial co-occurrence networks.
   It incorporates various popular, publicly available tools as well as custom Python modules and scripts to facilitate inference of co-occurrence networks from 16S data.
@@
   The \ac{micone} pipeline workflow is shown in Figure~\ref{fig:figure1}.
-  The different steps for going from 16S data to co-occurrence networks can be grouped into four major modules; (i) the denoising and clustering (DC) step, which handles denoising of the raw 16S sequencing data into representative sequences; (ii) the taxonomy assignment (TA) step that assigns taxonomic labels to the representative sequences; (iii) the OTU processing (OP) step that filters and transforms the taxonomy abundance table; and finally (iv) the network inferences (NI) step which infers the microbial co-occurrence network.
+  The different steps for going from 16S data to co-occurrence networks can be grouped into four major modules; (i) the denoising and clustering (DC) step, which handles denoising of the raw 16S sequencing data into representative sequences; (ii) the taxonomy assignment (TA) step that assigns taxonomic labels to the representative sequences; (iii) the \ac{otu} processing (OP) step that filters and transforms the taxonomy abundance table; and finally (iv) the network inferences (NI) step which infers the microbial co-occurrence network.
   Each process in the pipeline supports alternate tools for performing the same task (see Methods and Figure~\ref{fig:figure1}).
@@
 
-  % TODO: Talk about the datasets: real, mock and synthetic (with references)
-  In addition to the algorithmic pipeline itself, our analysis involves two types of data: The first type consists of sets of 16S sequencing data for real communities (see Methods).
+  In addition to the algorithmic pipeline itself, our analysis involves two types of data: The first type consists of sets of 16S sequencing data for real communities consisting of Stool and Oral microbiomes (see Methods).
   The second are datasets artificially created for the specific goal of helping evaluate computational analysis tools.
-In particular, in order to objectively compare, to the extent possible, how well each step in \ac{micone} best captures the underlying data, we use both mock data from mockrobiota~\cite{Bokulich2016} as well as, synthetically generated reads from an Illumina read simulator called ART~\cite{Huang2012}.
-  The mock datasets selected were labelled mock4, mock12 and mock16 in the mockrobiota database and consist of sequenced 16S data from mock communities of various bacteria.
-  These datasets were selected since they contained the expected compositions along with the reference sequences for the organisms in the mock community.
-  The synthetic reads were simulated using the three different taxonomy distribution profiles obtained from soil, water and healthy stool microbiome datasets \hl{[ref]}.
-  Reference sequences were generated using \ac{ncbi} and the decard tool \cite{Golob2017} for these taxonomy profiles.
+  In particular, in order to objectively compare, to the extent possible, how well each step in \ac{micone} best captures the underlying data, we use both mock data (labelled mock4, mock12 and mock16) from mockrobiota~\cite{Bokulich2016} as well as, synthetically generated reads from an Illumina read simulator called ART~\cite{Huang2012}.
+  These mock datasets were selected since they contained the expected compositions along with the reference sequences for the organisms in the mock community.
+  The synthetic reads were simulated using three different taxonomy distribution profiles, namely soil and water microbiomes obtained \ac{emp}~\cite{Thompson2017} and Stool microbiome that is used in our real community analysis~\cite{Kang2017}.
+  Reference sequences were generated using \ac{ncbi} and the Decard tool~\cite{Golob2017} for these taxonomy profiles.
   Detailed information on the mock communities and the settings used to generate the synthetic data are provided in the Methods section.
@@
 
-  % TODO: Summarized sections
   In order to analyze the effect of different statistical methods on the inferred co-occurrence networks, we generated co-occurrence networks using all possible combinations of methods and estimated the variability in the networks due to each choice.
@@
   This indicates that the taxonomy assigned to the reference sequences drastically alters the co-occurrence network and this change is much more significant than how the reference sequences themselves are identified (in the DC step).
-  We believe that the grouping by taxonomy assignment into clusters (..... in Fig. XXX) is due to the mislabelling of constitutive taxa that are present in high abundance in the community.
-  The residual variation can be seen as an artifact that arises due to the effect on the resultant network when multiple steps are changed at the same time.
-  Another interesting observation is that the dissimilarity between the networks decreases when the low abundance \ac{otu}s are removed from the network. This is visible in terms of ...
+  We believe that the grouping of the networks by taxonomy assignment into clusters (Figure~\ref{fig:figure2}B) is due to the mislabelling of constitutive taxa that are present in high abundance in the community, which drastically alters the nodes and hence the underlying network topology.
+  The residual variation (Figure \ref{fig:figure2}A) can be seen as an artifact that arises when multiple steps are changed at the same time.
+  Another interesting observation is that the dissimilarity between the networks decreases when the low abundance \ac{otu}s are removed from the network.
+This result is elaborated in detail in the denoising and clustering section.
   These results enable us to believe that the most important criterion for accurate comparative analyses of co-occurrence networks is the taxonomy reference database.
@@
   This implies that the taxonomy assignments from each reference database are fairly unique and is one of the reasons for a large differences in the resultant co-occurrence networks generated from different taxonomy databases.
-   % TODO: Update this
+
   Supplementary Figure~\ref{fig:figureS4} shows that the top 20 most abundant genera in the three resulting taxonomy composition tables are different.
@@
   % TODO: Talk about the difference between correlations and associations
-  The six different network inference methods used in this study are \ac{magma}~\hl{in biorxiv [ADD REF]}, \ac{mldm}~\cite{Yang2017}, \ac{spieceasi}~\cite{Kurtz2015}, \ac{sparcc}~\cite{Friedman2012}, Spearman and Pearson.
+  The six different network inference methods used in this study are \ac{magma}~\cite{Cougoul2019}, \ac{mldm}~\cite{Yang2017}, \ac{spieceasi}~\cite{Kurtz2015}, \ac{sparcc}~\cite{Friedman2012}, Spearman and Pearson.
   These network inference methods into two groups, the first set of methods (Pearson, Spearman, \ac{sparcc}) infer pairwise correlations while the second set infer direct associations (\ac{spieceasi}, \ac{mldm}, \ac{magma}).
+  Pairwise correlation methods involve calculating the correlation coefficient between every pair of \ac{OTU}/\ac{esv}s leading to the detection of spurious indirect connections.
+  On the other hand, direct association methods use conditional independence to avoid the detection of correlated but indirectly connected \ac{otu}s.
 

### 2020-07-21 sections/methods.tex
Updates from Overleaf

@@
   The reads are simulated using the NCBI RefSeq database as the reference sequence pool and the "art\_illumina" sequence profile with a mutation rate of 2\%.
-  The scripts used to generate the synthetic data are in the scripts folder of the respository (\href{https://github.com/segrelab/mind-pipeline-paper}{https://github.com/segrelab/mind-pipeline-paper}).
+  The scripts used to generate the synthetic data are in the scripts folder of the respository (\href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper}).
 

### 2020-07-21 sections/results.tex
Updates from Overleaf

@@
   In particular, in order to objectively compare, to the extent possible, how well each step in \ac{micone} best captures the underlying data, we use both mock data (labelled mock4, mock12 and mock16) from mockrobiota~\cite{Bokulich2016} as well as, synthetically generated reads from an Illumina read simulator called ART~\cite{Huang2012}.
-  These mock datasets were selected since they contained the expected compositions along with the reference sequences for the organisms in the mock community.
+  These mock datasets consist of fake sequencing reads generated from reads obtained from synthetic microbial isolates mixed in know proportions.
+They contain the expected compositions along with the reference sequences for the organisms in the mock community.
   The synthetic reads were simulated using three different taxonomy distribution profiles, namely soil and water microbiomes obtained \ac{emp}~\cite{Thompson2017} and Stool microbiome that is used in our real community analysis~\cite{Kang2017}.
@@
   The effects of various steps on the final co-occurrence network is estimated by building a linear model of the edges of the network as a function the various step in the analysis pipeline (see Methods).
-  Figure \ref{fig:figure2}B, shows the fraction of total variation among the co-occurrence networks as contributed by the first three steps of the pipeline.
+  Figure \ref{fig:figure2}B, shows the fraction of total variation among the co-occurrence networks due to the first three steps of the pipeline.
 In other words, each point corresponds to a different combination of tools, and captures how much the final network is affected by such choice.
@@
       \textbf{The choice of database contributes to the most variance in the networks}.
-      \textbf{(A)} The total relative variance in the networks contributed by the DC, TA and OP steps of the pipeline (right) and the linear model used to calculate the relative variance (left).
+      \textbf{(A)} The total relative variance in the networks contributed by the DC, TA and OP steps of the pipeline (right) and the linear model used to calculate the relative variance (left), see the Methods section for details.
       \textbf{(B)} All combinations of inferred networks are shown as points on a PCA plot.

### 2020-07-24 sections/discussion.tex
Updates from Overleaf

@@
 Our analysis shows that networks generated using different combinations of tools and approaches can look significantly different from each other, highlighting the importance of a clear assessment of the source of variability and of tools that provide the most robust and accurate results.
-Our newly developed integrated software for the inference of co-occurrence networks from 16S rRNA data, \ac{micone}, constitutes a freely customizable and user friendly pipeline that allows users to easily test combinations of tools and to compare networks generated by multiple possible choices.
+Our newly developed integrated software for the inference of co-occurrence networks from 16S rRNA data, \ac{micone}, constitutes a freely customizable and user friendly pipeline that allows users to easily test combinations of tools and to compare networks generated by multiple possible choices (refer to the Methods section).
 Importantly, in addition to revisiting the test cases presented in this work, users will be able to explore the effect of various tool combinations on their own datasets of interest.

### 2020-07-24 sections/results.tex
Updates from Overleaf

@@
       \textbf{The workflow of the microbial co-occurrence analysis pipeline}.
-      The processes can be grouped into four major steps: \textbf{(DC)} denoising and clustering, \textbf{(TA)} taxonomy assignment, \textbf{(OP)} OTU/ESV processing, and \textbf{(NI)} network inference.
+      The steps can be grouped into four major groups: \textbf{(DC)} \textbf{D}enoising and \textbf{C}lustering, \textbf{(TA)} \textbf{T}axonomy \textbf{A}ssignment, \textbf{(OP)} \textbf{O}TU or ESV \textbf{P}rocessing, and \textbf{(NI)} \textbf{N}etwork \textbf{I}nference.
       Each step incorporates several processes, each of which in turn have several alternate algorithms for the same task (indicated by the text to the right of the blue boxes).
@@
   The three databases vastly differ in terms of their last update status - \ac{gg} was last updated on May 2013, SILVA was last updated on December 2017 at the time of writing and \ac{ncbi} is updated as new sequences are curated.
-  Since updates to taxonomic classifications are frequent, these databases vary significantly in terms of taxonomy hierarchies~\cite{Balvociute2017}.
+  Since updates to taxonomic classifications are frequent, these databases vary significantly in terms of taxonomy hierarchies including species names and phylogenetic relationships~\cite{Balvociute2017}.
 
@@
 Another reason could be the frequency of updates and the potential for future growth, which prompted us to set \ac{ncbi} as the \ac{micone} standard for taxonomy assignment.
- In addition to being regularly maintained and updated the \ac{ncbi} database has the advantage that its accuracy of assignments is still comparable to the SILVA and \ac{gg} reference databases that are routinely used as reference databases.
+ In addition to being regularly maintained and updated the \ac{ncbi} database already has the advantage that its accuracy of assignments is still comparable to the SILVA and \ac{gg} reference databases that are routinely used as reference databases.
 
@@
   The six different network inference methods used in this study are \ac{magma}~\cite{Cougoul2019}, \ac{mldm}~\cite{Yang2017}, \ac{spieceasi}~\cite{Kurtz2015}, \ac{sparcc}~\cite{Friedman2012}, Spearman and Pearson.
-  These network inference methods into two groups, the first set of methods (Pearson, Spearman, \ac{sparcc}) infer pairwise correlations while the second set infer direct associations (\ac{spieceasi}, \ac{mldm}, \ac{magma}).
-  Pairwise correlation methods involve calculating the correlation coefficient between every pair of \ac{OTU}/\ac{esv}s leading to the detection of spurious indirect connections.
-  On the other hand, direct association methods use conditional independence to avoid the detection of correlated but indirectly connected \ac{otu}s.
+  These network inference methods fall into two groups, the first set of methods (Pearson, Spearman, \ac{sparcc}) infer pairwise correlations while the second set infer direct associations (\ac{spieceasi}, \ac{mldm}, \ac{magma}).
+  Pairwise correlation methods involve calculating the correlation coefficient between every pair of \ac{otu}/\ac{esv}s leading to the detection of spurious indirect connections.
+  On the other hand, direct association methods use conditional independence to avoid the detection of correlated but indirectly connected \ac{otu}s~\cite{Kurtz2015,Menon2018}.
 
@@
 Note, however, that some of these comparisons depend on the threshold that has to be applied to the pairwise correlations methods (currently 0.3, based on [REF]).
-  To get a more quantitative idea, we can take a look at the nodes and edges (Figure~\ref{fig:figure5}B) in common between the networks using UpSet plots (only \ac{magma}, \ac{mldm}, \ac{spieceasi}, \ac{sparcc} are used in the comparison).
+  To get a more quantitative idea, we can take a look at the nodes and edges (Figure~\ref{fig:figure5}B) in common between the networks using UpSet plots~\cite{Lex} (only \ac{magma}, \ac{mldm}, \ac{spieceasi}, \ac{sparcc} are used in the comparison since Pearson and Spearman add a large number of spurious edges since they aren't intended for compositional datasets).
   The results for the node intersections show that the networks have a large number of nodes in common ($63$ out of $67$ nodes in the smallest network - \ac{magma}) and no network possesses any unique node.
@@
       A threshold of 0.3 was set for the methods that infer pairwise correlations (\ac{sparcc}, Spearman, Pearson) and no threshold was set for the other methods.
-      \textbf{(B)} The node overlap upset plot indicates that all the networks have a large number of common nodes involved in connections.
-      Whereas, \textbf{(C)} The edge overlap upset plot shows that a very small fraction of these connections are actually shared.
+      \textbf{(B)} The node overlap Upset plot~\cite{Lex} indicates that all the networks have a large number of common nodes involved in connections.
+      Whereas, \textbf{(C)} The edge overlap Upset plot shows that a very small fraction of these connections are actually shared.
     }

### 2020-07-27 sections/acknowledgments.tex
Updates to draft
Updates from Overleaf

--- 2018-07-27 sections/acknowledgments.tex
@@
 
+% TODO: Update this section
 \section*{Acknowledgments}
 
-  We are grateful for ...
+We are grateful to members of the Segrè lab for helpful discussions and for feedback on the manuscript.
+This work was partially funded by grants from the NIH (T32GM100842, 5R01DE024468, R01GM121950 and NIA UH2AG064704), NSFOCE-BSF grant 1635070, the Human Frontiers Science Program (RGP0020/2016), and the Boston University Interdisciplinary Biomedical Research Office.
+[DANIEL, KIRILL: MISSING GRANTS???]
+
+\section*{Contributions}

### 2020-07-27 sections/methods.tex
Updates from Overleaf

@@
   pipeline: \href{https://github.com/segrelab/MiCoNE}{https://github.com/segrelab/MiCoNE} \\
-  data and scripts related to the paper: \href{}{}
+  data and scripts related to the paper: \href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper}

### 2020-07-31 sections/results.tex
Updates from Overleaf

@@
   The synthetic reads were simulated using three different taxonomy distribution profiles, namely soil and water microbiomes obtained \ac{emp}~\cite{Thompson2017} and Stool microbiome that is used in our real community analysis~\cite{Kang2017}.
-  Reference sequences were generated using \ac{ncbi} and the Decard tool~\cite{Golob2017} for these taxonomy profiles.
+  Reference sequences were generated using \ac{ncbi} and the Decard package~\cite{Golob2017} for these taxonomy profiles.
   Detailed information on the mock communities and the settings used to generate the synthetic data are provided in the Methods section.
@@
   There is no method that clearly outperforms the rest in all datasets.
-  Based on their slightly better performance on the mock datasets, their \textit{de novo} error correcting nature and other previous studies [refs], \ac{dada2} and Deblur seem to be in general the most reliable.
+  Based on their slightly better performance on the mock datasets, their \textit{de novo} error correcting nature and other previous studies~\cite{Nearing2018}, \ac{dada2} and Deblur seem to be in general the most reliable.
   Given the unexpected poor performance of Deblur on the synthetic data, the default algorithm in the pipeline was chosen to be \ac{dada2} (Supplementary Figure~\ref{fig:figureS3}).
@@
   We can clearly see that the different networks differ vastly in their edge-density and connectivity and even some of the edges in common to these networks have their signs inverted.
-Note, however, that some of these comparisons depend on the threshold that has to be applied to the pairwise correlations methods (currently 0.3, based on [REF]).
-  To get a more quantitative idea, we can take a look at the nodes and edges (Figure~\ref{fig:figure5}B) in common between the networks using UpSet plots~\cite{Lex} (only \ac{magma}, \ac{mldm}, \ac{spieceasi}, \ac{sparcc} are used in the comparison since Pearson and Spearman add a large number of spurious edges since they aren't intended for compositional datasets).
+Note, however, that some of these comparisons depend on the threshold that has to be applied to the pairwise correlations methods (currently 0.3, based on~\cite{Friedman2012}).
+  To get a more quantitative idea, we can take a look at the nodes and edges (Figure~\ref{fig:figure5}B) in common between the networks using UpSet plots~\cite{Lex} (only \ac{magma}, \ac{mldm}, \ac{spieceasi}, \ac{sparcc} are used in the comparison since Pearson and Spearman add a large number of spurious edges since they are not intended for compositional datasets).
   The results for the node intersections show that the networks have a large number of nodes in common ($63$ out of $67$ nodes in the smallest network - \ac{magma}) and no network possesses any unique node.
@@
   Unlike in the previous sections, where were we evaluated the performance of methods on mock datasets, there are no such datasets that contain a set of known interactions for the evaluation of the network inference algorithms.
-  Therefore, we propose the construction of a consensus network (Figure~\ref{fig:figure5}C) involving \ac{magma}, \ac{mldm}, \ac{spieceasi} and \ac{sparcc} by merging the p-values generated from bootstraps of the original taxonomy composition table using the Browns p-value combining method [ref] (see Methods section).
+  Therefore, we propose the construction of a consensus network (Figure~\ref{fig:figure5}C) involving \ac{magma}, \ac{mldm}, \ac{spieceasi} and \ac{sparcc} by merging the p-values generated from bootstraps of the original taxonomy composition table using the Browns p-value combining method~\cite{Poole} (see Methods section).
   Based on this approach, \ac{micone} reports as default output the consensus network, annotated with weights (correlations for \ac{sparcc} and direct associations for the other methods) for all four methods.

### 2020-08-07 sections/methods.tex
Updates from Overleaf

@@
   The mock community 16S datasets are real sequencing data obtained for artificially assembled collections of species in known proportions.
-  The mock datasets used for this study, obtained from mockrobiota~\cite{Bokulich2016}, were labelled mock4, mock12 and mock16.
+  The mock datasets used for this study, obtained from mockrobiota~\cite{Bokulich2016}, are labelled mock4, mock12 and mock16.
   The mock4 community is composed of 21 bacterial strains.
@@
   The reads are simulated using the NCBI RefSeq database as the reference sequence pool and the "art\_illumina" sequence profile with a mutation rate of 2\%.
-  The scripts used to generate the synthetic data are in the scripts folder of the respository (\href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper}).
+  The scripts used to generate the synthetic data are in the scripts folder of the repository (\href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper}).
 

### 2020-08-07 sections/results.tex
Updates from Overleaf

@@
       \textbf{(B)} The relatively larger average unweighted UniFrac distance indicates that methods differ in their identification of sequences of low abundance,
-      \textbf{(C, D)} The distributions of the avegage weighted UniFrac distance between the expected sequence profile and the calculated sequence profile in mock datasets.
+      \textbf{(C, D)} The distributions of the average weighted UniFrac distance between the expected sequence profile and the calculated sequence profile in mock datasets.
       \textbf{(D)} The distributions of the average unweighted UniFrac distance show that dada2 and Deblur were the best performing methods in most of the datasets.

### 2020-08-21 sections/discussion.tex
Updates from Overleaf

--- 2020-06-18 sections/discussion.tex
@@
 Our analysis shows that networks generated using different combinations of tools and approaches can look significantly different from each other, highlighting the importance of a clear assessment of the source of variability and of tools that provide the most robust and accurate results.
-Our newly developed integrated software for the inference of co-occurrence networks from 16S rRNA data, \ac{micone}, constitutes a freely customizable and user friendly pipeline that allows users to easily test combinations of tools and to compare networks generated by multiple possible choices.
+Our newly developed integrated software for the inference of co-occurrence networks from 16S rRNA data, \ac{micone}, constitutes a freely customizable and user friendly pipeline that allows users to easily test combinations of tools and to compare networks generated by multiple possible choices (see Methods).
 Importantly, in addition to revisiting the test cases presented in this work, users will be able to explore the effect of various tool combinations on their own datasets of interest.
 The \ac{micone} pipeline is built in a modular fashion.
-Its plug-and-play architecture will make it possible for users to add new tools and steps, either from existing packages, or from those that were not examined in the present work and those that will be developed in the future.
+Its plug-and-play architecture will make it possible for users to add new tools and steps, either from existing packages, or from packages that were not examined in the present work, as well as future ones.
 
@@
 However, comparison of inferred and expected reference sequences and their abundances in mock community datasets has allowed us to identify \ac{dada2} as the method which best recapitulates the expected sequence composition.
-There are newer recently published methods like dbOTU3~\cite{Olesen2017} as well as older popular methods like mothur~\cite{Schloss2009} that have not been included in the study, and are possible avenues for future investigation.
+For the current work we have decided to focus on the tools most widely used at the time of the analysis.
+Some tools that we recently published (e.g.
+dbOTU3~\cite{Olesen2017}) as well as older popular methods like mothur~\cite{Schloss2009} have not been included in the study, but could be added into the pipelines in future updated analyses.
 
@@
 The frequent changes in the taxonomy nomenclature coupled with the frequency of updates to the various 16S reference databases create inherent differences \cite{Balvociute2017} in taxonomy hierarchies in these databases.
-Our analysis did not reveal one particular reference database to perform better in all scenarios.
+Our analysis revealed that no particular reference database performs better than the others across all scenarios.
 We suggest that that choice of the database should be made based on possible reported or inferred biases in the representation of given biomes in a specific databases \cite{Balvociute2017}.

### 2020-08-21 sections/introduction.tex
Updates from Overleaf

@@
 
-  Microbial communities are ubiquitous and play an important role in biogeochemical transformations in both natural as well as man-made ecosystems.
-  These microbiomes comprise several thousand different microbial strains interacting with each other and their environment, often through intricate metabolic and signaling relationships.
-  Changes in the microbiome can alter the environment in which they reside \cite{HumanMicrobiomeProjectConsortium2012,Lloyd-Price2016} and for host-associated microbiomes this can have major implications in health of the human body.
-  A mechanistic understanding of the key interactions between microbes is necessary in order to analyze the dysbiosis in a healthy microbiome and to design interventions.
+  Microbial communities are ubiquitous and play an important role in marine and terrestrial environments, urban ecosystems, metabolic engineering, and human health [REFs].
+  These microbial communities, or microbiomes, often comprise several hundreds of different microbial strains interacting with each other and their environment, often through intricate metabolic and signaling relationships.
+  Understanding how these interconnections shape community structure and functionalities is a fundamental challenge in microbial ecology, with applications in the study of microbial ecosystems across different biomes.
+  With the advancement in DNA sequencing technologies \cite{Narihiro2017} and data processing methods,  more information can be extracted from these microbial community samples than ever before.
+  In particular, high-throughput sequencing, including community metagenomic sequencing and sequencing of 16S rRNA gene amplicons, has the potential to help detect, identify and quantify a large portion of the constitutive microorganisms of a microbiome \cite{Jovel2016} [MORE REFs, Reviews, keystone papers...].
+  These advances have led to large-scale data collection efforts involving environmental (\acl{emp}) \cite{Thompson2017}, marine (Tara Oceans Project) \cite{Zhang2015} and human-associated microbiota (Human Microbiome Project) \cite{HumanMicrobiomeProjectConsortium2012}.
 
-  Studies have shown the importance of specific microbial interactions in the healthy microbiome \cite{Lloyd-Price2016} and others have shown how changes in these interactions can lead to dysbiosis \cite{Wang2017,Gilbert2016,Belizario2015}.
-  While direct measurement of interactions is challenging, a significant amount of effort has gone into measuring correlations leading to the correlation information between microbial abundances being readily available.
-  Co-occurrence relationships in microbial communities may be driven by microbial interactions, although the extent and nature of these relationships is highly debated \cite{Zuniga2017}.
-  Each interaction within this complex ecological network can have a positive effect, a negative effect or no effect on the species involved.
-  Bacteria that compete for the same nutrient will have a negative co-occurrence relationship \cite{Ghoul2016} whereas cross-feeding between species leads to a positive co-occurrence relationship \cite{DSouza2018}.
-  But these co-occurrence relationships are not very straightforward, since, similar species competing for a resource could co-occur if the resource is abundant.
-  The microbial interaction networks are highly dynamic, and they constantly reorganize in a varying environment.
-  Even under static conditions, these interactions are often difficult to predict due to their non-linear nature \cite{Konopka2015}.
-  Compounding on this complexity is the fact that these interaction networks usually involve thousands of species.
+ This wealth of information on the composition and functions of a community at different times and under different environmental conditions has the potential to help us understand how communities assemble and operate.
+ A powerful tool for translating microbiome data into knowledge is the construction of possible inter-dependence networks across species.
+ The importance of these networks of relationships is two fold: first, such networks can serve as maps that help identify hubs of keystone species [REF], or basic microbiome changes that occur as a consequence of environmental perturbations or underlying host conditions [REF]; second, networks of inter-dependencies can serve as a key bridge towards building mechanistic models of microbial communities, greatly enhancing our capacity to understand and control them.
+ For example, multiple studies have shown the importance of specific microbial interactions in the healthy microbiome \cite{Lloyd-Price2016} and others have shown how changes in these interactions can lead to dysbiosis \cite{Wang2017,Gilbert2016,Belizario2015}.
+ In the context of terrestrial biogeochemistry..... [ANOTHER EXAMPLE].
 
-  High-throughput metagenomic sequencing techniques (such as 16S rRNA gene profiling) help detect, identify and quantify a large part of the constitutive microorganisms of a microbiome \cite{Jovel2016}.
-  Knowledge of the community composition at a particular instance in time would enable us to derive partial insights into the underlying dynamics.
-  With the advancement in DNA sequencing technologies \cite{Narihiro2017} and data processing methods much more information can be extracted from these microbial community samples than ever before.
-  These advances have led to large-scale data collection efforts involving environmental (\acl{emp}) \cite{Thompson2017}, marine (Tara Oceans Project) \cite{Zhang2015} and human-associated microbiota (Human Microbiome Project) \cite{HumanMicrobiomeProjectConsortium2012} and thus led the generation of a wealth of data about these microbial communities.
+ Direct measurement of interactions, e.g. through co-culture microdroplet experiments [REF Blainey], or spatial visualization of natural communities [REF Borisy]  is possible, but still challenging.
+ In parallel, sequencing data across multiple samples can be used for estimating co-occurrence relationships between taxa.
+ While the the relationship between directly measured interactions and statistically inferred co-occurrence is still poorly understood \cite{Zuniga2017}, a significant amount of effort has gone into estimating correlations from large microbiome sequence datasets.
+ Co-occurrence networks have microbial taxa as nodes, and edges that represent the frequent co-occurrence (or negative correlations) across different datasets.
 
-  These data-sets can be used to infer networks of co-occurrence relationships between microorganisms in the samples.
-  These networks have microbial taxa as nodes, and edges that represent the frequent co-occurrence (or negative correlations) across different datasets.
-  Co-occurrence networks can in principle be used to identify important interactions driving the dynamics of the microbial community, including emergent ecosystem-level properties such as ecosystem robustness and modularity.
+One of the most frequently used avenues for inferring co-occurrence networks is the parsing and analysis of 16S sequencing data.
+A large number of software tools and pipelines have been developed to analyze 16S sequencing data, often focused on addressing the many known limitations of this methodology, including resolution, sequencing depth, compositional nature, sequencing errors and copy number variations.
+Popular methods for different phases of the analysis of 16S data include; (i) tools for removing statistical artifacts...; (ii),
 
-  One of the limitations of this type of analysis is that 16S sequencing data is limited by many factors such as resolution, sequencing depth, compositional nature, sequencing errors and copy number variations.
   Many different analysis methods have been developed \cite{Callahan2016,Amir2017,Friedman2012,Kurtz2015} to counter these limitations and remove statistical artifacts, leading to the existence of a myriad of tools for each step of the data analysis workflow.
@@
  Previous work has developed popular platforms (like MG-RAST \cite{Keegan2016}, Qiita \cite{qiita}) and tools (such as QIIME \cite{Caporaso2010}) that provide pipelines for 16S data analysis.
+
  However, existing tools typically are focused on \ac{otu} generation and not on the effects of upstream statistical methods on the inferred co-occurrence networks.

### 2020-08-21 sections/methods.tex
Updates from Overleaf

@@
   The output of this module is a matrix of counts, that describes the number of reads of a particular \ac{otu} or \ac{esv} (rows of the matrix) present in each sample (columns of the matrix).
-  The options currently available in the pipeline for denoising and clustering: open reference clustering, closed reference clustering and de novo clustering methods from \ac{qiime1} v1.9.1~\cite{Caporaso2010} and denoising methods from \ac{dada2} v1.14~\cite{Callahan2016} and Deblur v1.1.0~\cite{Amir2017}.
+  The options currently available in the pipeline for denoising and clustering are: open reference clustering, closed reference clustering and de novo clustering methods from \ac{qiime1} v1.9.1~\cite{Caporaso2010} and denoising methods from \ac{dada2} v1.14~\cite{Callahan2016} and Deblur v1.1.0~\cite{Amir2017}.
   The quality filtering and chimera checking tools are derived from those used in \ac{qiime2} v2019.10.0 and \ac{dada2}.
@@
 
-  % TODO: Add references and basic equations
+  % TODO: Add references and basic equations or details
   \subsection*{OTU and ESV Processing (OP)}
@@
   The pipeline includes Pearson, Spearman and FastSpar v0.0.10 (a faster implementation of \ac{sparcc})~\cite{Watts2018} as the pairwise correlation metrics, and \ac{spieceasi} v1.0.7~\cite{Kurtz2015}, \ac{mldm} v1.1~\cite{Yang2017} and \ac{magma}~\cite{Cougoul2019} as the direct association metrics.
-  The empirical Browns method is used for combining p-values from the various methods to obtain a consensus p-value, which is used to create the consensus network.
+  The empirical Browns method~\cite{Poole} is used for combining p-values from the various methods to obtain a consensus p-value, which is used to create the consensus network.
 
@@
 
-  In other words, $N$ is the merged table and each column $N_i$ is the vector representation of one of the networks and each row $L_i$ represents the one particular edge in all networks (assigned 0 if the edge does not exist in the network).
+  In other words, $N$ is the merged table, each column $N_i$ is the vector representation of one of the networks, and each row $L_i$ represents the one particular edge in all networks (assigned 0 if the edge does not exist in the network).
 
-  We next use linear regression to express each link $L_i$ as a linear function of categorical variables that describe the possible options in each of the first three steps of the pipeline.
+  We use linear regression to express each link $L_i$ as a linear function of categorical variables that describe the possible options in each of the first three steps of the pipeline.
 
-  % TODO: Explain the categorical linear model and ANOVA better
+  % TODO: Explain the categorical linear model and ANOVA better (text)
   In particular, we infer parameters $\alpha_i$ such that:
@@
   \subsection*{Code and Data Availability}
-  pipeline: \href{https://github.com/segrelab/MiCoNE}{https://github.com/segrelab/MiCoNE} \\
-  data and scripts related to the paper: \href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper}
+  Pipeline: \href{https://github.com/segrelab/MiCoNE}{https://github.com/segrelab/MiCoNE} \\
+  Data and scripts: \href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper}

### 2020-08-25 sections/abstract.tex
Updates from Overleaf

@@
     Ultimately, we develop a standardized pipeline that follows these default tools and parameters, but that can also help explore the outcome of any other combination of choices.
-    We envisage that this pipeline for processing 16S sequencing data into networks of microbial co-occurrences could be used for integrating multiple data-sets, and for generating comparative analyses and consensus networks useful for detecting disease-related patterns.
+    We envisage that this pipeline for processing 16S sequencing data into networks of microbial co-occurrences could be used for integrating multiple data-sets, and for generating comparative analyses and consensus networks that can help understand and control microbial community assembly in different biomes.
         The pipeline is available as a Python package at \href{https://github.com/segrelab/MiCoNE}{https://github.com/segrelab/MiCoNE}

### 2020-08-25 sections/acknowledgments.tex
Updates from Overleaf

--- 2018-07-27 sections/acknowledgments.tex
@@
 
+% TODO: Update this section
 \section*{Acknowledgments}
 
-  We are grateful for ...
+We are grateful to members of the Segrè lab for helpful discussions and for feedback on the manuscript.
+This work was partially funded by grants from the National Institutes of Health (National Institute of General Medical Sciences, award R01GM121950; National Institute of Dental and Craniofacial Research, award number R01DE024468; and National Institute on Aging, award number UH2AG064704), the U.S.
+Department of Energy, Office of Science, Office of Biological & Environmental Research through the Microbial Community Analysis and Functional Evaluation in Soils SFA Program (m-CAFEs) under contract number DE-AC02-05CH11231 to Lawrence Berkeley National Laboratory, the National Science Foundation (grant NSFOCE-BSF 1635070), the Human Frontiers Science Program (RGP0020/2016), and the Boston University Interdisciplinary Biomedical Research Office.
+
+\section*{Contributions}
+Designed the research project: DK, KK, DS, ZH, CDL. Performed analysis: DK, GB. Wrote the first draft of the manuscript: DK. Revised and wrote final version of the manuscript: DK, DS, KK.

### 2020-08-25 sections/abstract.tex
Updates from Overleaf

@@
     16S ribosomal RNA (16S rRNA) amplicon profiling provides snapshots that reveal the phylogenies and abundance profiles of these microbial communities.
-    These snapshots, when collected from multiple samples, have the potential to reveal which microbes co-occur, providing a glimpse into the network of influences in these communities.
+    These snapshots, when collected from multiple samples, have the potential to reveal which microbes co-occur, providing a glimpse into the network of associations in these communities.
     The inference of networks from 16S data is prone to statistical artifacts.
     There are many tools for performing each step of the 16S analysis workflow, but the extent to which these steps affect the final network is still unclear.
-    In this study, we perform a meticulous analysis of each step of a pipeline that processes 16S sequencing data into a network of microbial associations.
+    In this study, we perform a meticulous analysis of each step of a pipeline that can convert 16S sequencing data into a network of microbial associations.
     Through this process, we map how different choices of algorithms and parameters affect the co-occurrence network, and estimate steps which contribute most significantly to the variance.

### 2020-08-25 sections/acknowledgments.tex
Updates from Overleaf

@@
 
-% TODO: Update this section
 \section*{Acknowledgments}
@@
 This work was partially funded by grants from the National Institutes of Health (National Institute of General Medical Sciences, award R01GM121950; National Institute of Dental and Craniofacial Research, award number R01DE024468; and National Institute on Aging, award number UH2AG064704), the U.S.
-Department of Energy, Office of Science, Office of Biological & Environmental Research through the Microbial Community Analysis and Functional Evaluation in Soils SFA Program (m-CAFEs) under contract number DE-AC02-05CH11231 to Lawrence Berkeley National Laboratory, the National Science Foundation (grant NSFOCE-BSF 1635070), the Human Frontiers Science Program (RGP0020/2016), and the Boston University Interdisciplinary Biomedical Research Office.
+Department of Energy, Office of Science, Office of Biological \& Environmental Research through the Microbial Community Analysis and Functional Evaluation in Soils SFA Program (m-CAFEs) under contract number DE-AC02-05CH11231 to Lawrence Berkeley National Laboratory, the National Science Foundation (grant NSFOCE-BSF 1635070), the Human Frontiers Science Program (RGP0020/2016), and the Boston University Interdisciplinary Biomedical Research Office.
 

### 2020-08-25 sections/coverpage.tex
Updates from Overleaf

@@
   \vspace{1.5 cm}
-  Dileep Kishore\textsuperscript{1},
-  Gabriel Birzu\textsuperscript{2},
+  Dileep Kishore\textsuperscript{1,2},
+  Gabriel Birzu\textsuperscript{3},
   Zhenjun Hu\textsuperscript{1},
   Charles DeLisi\textsuperscript{1,3},
-  Kirill Korolev\textsuperscript{$\dagger$1,2},\\
-  Daniel Segr\`{e}\textsuperscript{$\dagger$1,3,4}\\
+  Kirill Korolev\textsuperscript{$\dagger$1,2,3},\\
+  Daniel Segr\`{e}\textsuperscript{$\dagger$1,2,4,5}\\
   \vspace{1cm}
   \textsuperscript{1}Bioinformatics Program, Boston University, Boston, Massachusetts, USA\\
-  \textsuperscript{2}Department of Physics, Boston University, Boston, Massachusetts, USA\\
-  \textsuperscript{3}Department of Biomedical Engineering, Boston University, Boston, Massachusetts, USA\\
-  \textsuperscript{4}Department of Biology, Boston University, Boston, Massachusetts, USA\\
+  \textsuperscript{2}Biological Design Center, Boston University, Boston, Massachusetts, USA\\
+  \textsuperscript{3}Department of Physics, Boston University, Boston, Massachusetts, USA\\
+  \textsuperscript{4}Department of Biomedical Engineering, Boston University, Boston, Massachusetts, USA\\
+  \textsuperscript{5}Department of Biology, Boston University, Boston, Massachusetts, USA\\
   \textsuperscript{$\dagger$}Correspondence should be sent to \href{mailto:korolev@bu.edu}{korolev@bu.edu} or \href{mailto:dsegre@bu.edu}{dsegre@bu.edu}\\

### 2020-08-25 sections/introduction.tex
Updates from Overleaf

@@
 
-  Microbial communities are ubiquitous and play an important role in marine and terrestrial environments, urban ecosystems, metabolic engineering, and human health [REFs].
+  Microbial communities are ubiquitous and play an important role in marine and terrestrial environments, urban ecosystems, metabolic engineering, and human health \cite{Ghoul2016,Thompson2017}.
   These microbial communities, or microbiomes, often comprise several hundreds of different microbial strains interacting with each other and their environment, often through intricate metabolic and signaling relationships.
@@
   With the advancement in DNA sequencing technologies \cite{Narihiro2017} and data processing methods,  more information can be extracted from these microbial community samples than ever before.
-  In particular, high-throughput sequencing, including community metagenomic sequencing and sequencing of 16S rRNA gene amplicons, has the potential to help detect, identify and quantify a large portion of the constitutive microorganisms of a microbiome \cite{Jovel2016} [MORE REFs, Reviews, keystone papers...].
+  In particular, high-throughput sequencing, including community metagenomic sequencing and sequencing of 16S rRNA gene amplicons, has the potential to help detect, identify and quantify a large portion of the constitutive microorganisms of a microbiome \cite{Jovel2016,Lloyd-Price2016}.
   These advances have led to large-scale data collection efforts involving environmental (\acl{emp}) \cite{Thompson2017}, marine (Tara Oceans Project) \cite{Zhang2015} and human-associated microbiota (Human Microbiome Project) \cite{HumanMicrobiomeProjectConsortium2012}.
@@
  A powerful tool for translating microbiome data into knowledge is the construction of possible inter-dependence networks across species.
- The importance of these networks of relationships is two fold: first, such networks can serve as maps that help identify hubs of keystone species [REF], or basic microbiome changes that occur as a consequence of environmental perturbations or underlying host conditions [REF]; second, networks of inter-dependencies can serve as a key bridge towards building mechanistic models of microbial communities, greatly enhancing our capacity to understand and control them.
+ The importance of these networks of relationships is two fold: first, such networks can serve as maps that help identify hubs of keystone species \cite{Menon2018,Rottjers2018}, or basic microbiome changes that occur as a consequence of environmental perturbations or underlying host conditions \cite{Gilbert2016}; second, networks of inter-dependencies can serve as a key bridge towards building mechanistic models of microbial communities, greatly enhancing our capacity to understand and control them.
  For example, multiple studies have shown the importance of specific microbial interactions in the healthy microbiome \cite{Lloyd-Price2016} and others have shown how changes in these interactions can lead to dysbiosis \cite{Wang2017,Gilbert2016,Belizario2015}.
- In the context of terrestrial biogeochemistry..... [ANOTHER EXAMPLE].
+ In the context of terrestrial biogeochemistry, co-occurrence networks have proposed as a valuable approach towards reconstructing the processes leading to microbiome assembly [REF Fierer], and understanding the response of microbial communities to environmental perturbations [REF].
 
- Direct measurement of interactions, e.g. through co-culture microdroplet experiments [REF Blainey], or spatial visualization of natural communities [REF Borisy]  is possible, but still challenging.
+ Direct high-throughput measurement of interactions, e.g.
+through co-culture microdroplet experiments [REF Blainey], or spatial visualization of natural communities [REF Borisy] is possible, but it requires specific technological capabilities, and has not been used very extensively yet.
  In parallel, sequencing data across multiple samples can be used for estimating co-occurrence relationships between taxa.
@@
 
-One of the most frequently used avenues for inferring co-occurrence networks is the parsing and analysis of 16S sequencing data.
-A large number of software tools and pipelines have been developed to analyze 16S sequencing data, often focused on addressing the many known limitations of this methodology, including resolution, sequencing depth, compositional nature, sequencing errors and copy number variations.
-Popular methods for different phases of the analysis of 16S data include; (i) tools for removing statistical artifacts...; (ii),
-
-  Many different analysis methods have been developed \cite{Callahan2016,Amir2017,Friedman2012,Kurtz2015} to counter these limitations and remove statistical artifacts, leading to the existence of a myriad of tools for each step of the data analysis workflow.
-  The different methods and tools developed to solve these issues infer vastly different community compositions and co-occurrence networks \cite{Golob2017,Weiss2016}, making it difficult to reliably compare networks across different publications and studies.
-  Conversely, given the lack of comprehensive comparisons between directly observed microbial interactions (e.g.
+One of the most frequently used avenues for inferring co-occurrence networks is the parsing and analysis of 16S sequencing data [REF].
+A large number of software tools and pipelines have been developed to analyze 16S sequencing data, often focused on addressing the many known limitations of this methodology, including resolution, sequencing depth, compositional nature, sequencing errors and copy number variations [REF].
+Popular methods for different phases of the analysis of 16S data include tools for: (i) denoising and clustering sequencing reads \cite{Caporaso2010,Callahan2016}; (ii) assigning taxonomy to the reads \cite{DeSantis2006,Quast2012}; (iii) processing and transforming the taxonomy count matrices \cite{Weiss2015}; and (iv) inferring the co-occurrence network \cite{Cougoul2019,Kurtz2015}.
+Different specific algorithms are often aggregated into popular platforms (like MG-RAST \cite{Keegan2016}, Qiita \cite{qiita}) and packages (such as QIIME \cite{Caporaso2010}) that provide pipelines for 16S data analysis.
+The different methods and tools developed to solve issues arising in 16S analysis can lead to vastly different inferences of community compositions and co-occurrence networks \cite{Golob2017,Weiss2016}, making it difficult to reliably compare networks across different publications and studies.
+ This is partially due to the fact that existing platforms are typically focused on \ac{otu} generation and not on the effects of upstream statistical methods on the inferred co-occurrence networks.
+ Furthermore, no organized framework currently exist to systematically analyze and compare existing components of the data-analysis from amplicons to networks.
+  More broadly, given the lack of comprehensive comparisons between directly observed microbial interactions (e.g.
 from co-culture experiments) and co-occurrence networks, there is no straightforward way to determine which set of tools or methods generate the most accurate networks.
-
- Previous work has developed popular platforms (like MG-RAST \cite{Keegan2016}, Qiita \cite{qiita}) and tools (such as QIIME \cite{Caporaso2010}) that provide pipelines for 16S data analysis.
-
- However, existing tools typically are focused on \ac{otu} generation and not on the effects of upstream statistical methods on the inferred co-occurrence networks.
- Furthermore, no organized framework currently exist to systematically analyze and compare existing components of the data-analysis from amplicons to networks.
 

### 2020-08-25 sections/results.tex
Updates from Overleaf

@@
   We have developed \ac{micone}, a flexible and modular pipeline for 16S amplicon sequencing rRNA data (hereafter mentioned simply as 16S data) analysis, that allows us to infer microbial co-occurrence networks.
-  It incorporates various popular, publicly available tools as well as custom Python modules and scripts to facilitate inference of co-occurrence networks from 16S data.
+  It incorporates various popular, publicly available tools as well as custom Python modules and scripts to facilitate inference of co-occurrence networks from 16S data (see Methods).
   Using \ac{micone} one can obtain co-occurrence networks by applying to 16S data (or to already processed taxonomic count matrices) any combination of the available tools.
@@
   Each process in the pipeline supports alternate tools for performing the same task (see Methods and Figure~\ref{fig:figure1}).
-  A centralized configuration file contains all the specifications for what modules to be used in the pipeline, and can be modified by the user to choose the desired set of tools.
+  A centralized configuration file contains all the specifications for what modules are used in the pipeline, and can be modified by the user to choose the desired set of tools.
   In what follows, we perform a systematic analysis of each step of the pipeline to estimate how much the final co-occurrence network depends on the possible choices at each step.
@@
 
-  In addition to the algorithmic pipeline itself, our analysis involves two types of data: The first type consists of sets of 16S sequencing data for real communities consisting of Stool and Oral microbiomes (see Methods).
-  The second are datasets artificially created for the specific goal of helping evaluate computational analysis tools.
+  Our analysis involves two types of data: The first type consists of sets of 16S sequencing data for real communities consisting of Stool and Oral microbiomes.
+  The second are datasets synthetically or artificially created for the specific goal of helping evaluate computational analysis tools (see Methods).
   In particular, in order to objectively compare, to the extent possible, how well each step in \ac{micone} best captures the underlying data, we use both mock data (labelled mock4, mock12 and mock16) from mockrobiota~\cite{Bokulich2016} as well as, synthetically generated reads from an Illumina read simulator called ART~\cite{Huang2012}.
@@
 
-  In order to analyze the effect of different statistical methods on the inferred co-occurrence networks, we generated co-occurrence networks using all possible combinations of methods and estimated the variability in the networks due to each choice.
+  In order to analyze the effect of different statistical methods on the inferred co-occurrence networks, we generated co-occurrence networks using all possible combinations of methods and estimated the variability in the networks due to each choice (Figure \ref{fig:figure1}).
   This analysis is performed while keeping the network inference algorithm (NI step) the same throughout the analysis.
@@
 This is also reflected in the fact that the networks can be clearly separated based on the database used (Figure \ref{fig:figure2}B).
-  This indicates that the taxonomy assigned to the reference sequences drastically alters the co-occurrence network and this change is much more significant than how the reference sequences themselves are identified (in the DC step).
-  We believe that the grouping of the networks by taxonomy assignment into clusters (Figure~\ref{fig:figure2}B) is due to the mislabelling of constitutive taxa that are present in high abundance in the community, which drastically alters the nodes and hence the underlying network topology.
+  This indicates that the taxonomy assigned to the reference sequences drastically alters the co-occurrence network.
+  In fact the variability induced by taxonomy assignment is much more significant than that due to the variability induced based on how the reference sequences themselves are identified (in the DC step).
+  The grouping of the networks by taxonomy assignment into clusters (Figure~\ref{fig:figure2}B) seems to derive from the mislabelling of constitutive taxa that are present in high abundance in the community, which drastically alter the nodes and hence the underlying network topology.
   The residual variation (Figure \ref{fig:figure2}A) can be seen as an artifact that arises when multiple steps are changed at the same time.
-  Another interesting observation is that the dissimilarity between the networks decreases when the low abundance \ac{otu}s are removed from the network.
-This result is elaborated in detail in the denoising and clustering section.
-  These results enable us to believe that the most important criterion for accurate comparative analyses of co-occurrence networks is the taxonomy reference database.
+  Another interesting observation (elaborated in detail in the denoising and clustering section) is that the dissimilarity between the networks decreases when the low abundance \ac{otu}s are removed from the network.
+  These results suggest that the most important criterion for accurate comparative analyses of co-occurrence networks is the taxonomy reference database.
 
@@
   We then correlate the abundances of matching taxonomies between a pair of DC methods (Figure \ref{fig:figureS1}A and B).
-  The \ac{esv} tables generated by methods that perform denoising are very similar to each other $\sim0.91$ and the \ac{otu} tables generated by the clustering methods are very similar to each other $\sim0.9$, but results of denoising and clustering are highly uncorrelated with each other $\sim0.4$ (Figure \ref{fig:figureS1}C).
+  The \ac{esv} tables generated by methods that perform denoising are very similar to each other ($\sim0.91$) and the \ac{otu} tables generated by the clustering methods are very similar to each other ($\sim0.9$), but results of denoising and clustering are highly uncorrelated with each other ($\sim0.4$) (Figure \ref{fig:figureS1}C).
 
   These comparisons only elucidate the pairwise similarity or dissimilarity of a pair of methods.
-  In order to determine the tool that most accurately recapitulates the reference sequences in the samples, we use the 16S sequences from the mock datasets.
+  In order to determine the tool that most accurately recapitulates the reference sequences in the samples, we used the 16S sequences from the mock datasets.
   In particular, we used the pipeline to process mock community datasets using each of the possible methods included for this step.
@@
   In order to compare the assigned taxonomies from different databases, we use the same reference sequences and assign taxonomies to them using different taxonomy reference databases.
-  The three 16S taxonomic reference databases used in this study are SILVA~\cite{Quast2012}, \ac{gg}~\cite{DeSantis2006} and \ac{ncbi} RefSeq.
+  The three 16S taxonomic reference databases used in this study are SILVA~\cite{Quast2012}, \ac{gg}~\cite{DeSantis2006} and \ac{ncbi} RefSeq~\cite{Sayers2009}.
   SILVA and \ac{gg} are two popular 16S databases used for taxonomy identification.
@@
   The assignments beyond Family level (Family, Genus and Species) are very dissimilar with $<70\%$ similarity between any pair of databases.
-  There are no two reference databases that are more similar than the other, with \ac{gg} and SILVA producing only marginally similar assignments compared to \ac{ncbi}.
-  This implies that the taxonomy assignments from each reference database are fairly unique and is one of the reasons for a large differences in the resultant co-occurrence networks generated from different taxonomy databases.
+  There are no two reference databases that are more similar than the other pairs, with \ac{gg} and SILVA producing only marginally similar assignments compared to \ac{ncbi}.
+  This implies that the taxonomy assignments from each reference database are fairly unique and are largely responsible for the differences observed in the co-occurrence networks generated from different taxonomy databases.
 
@@
   Figure~\ref{fig:figure5}A shows the networks inferred from this dataset using the different inference algorithms.
-  We can clearly see that the different networks differ vastly in their edge-density and connectivity and even some of the edges in common to these networks have their signs inverted.
+  The different networks differ vastly in their edge-density and connectivity; even some of the edges in common to these networks have their signs inverted.
 Note, however, that some of these comparisons depend on the threshold that has to be applied to the pairwise correlations methods (currently 0.3, based on~\cite{Friedman2012}).
@@
 
-  Unlike in the previous sections, where were we evaluated the performance of methods on mock datasets, there are no such datasets that contain a set of known interactions for the evaluation of the network inference algorithms.
-  Therefore, we propose the construction of a consensus network (Figure~\ref{fig:figure5}C) involving \ac{magma}, \ac{mldm}, \ac{spieceasi} and \ac{sparcc} by merging the p-values generated from bootstraps of the original taxonomy composition table using the Browns p-value combining method~\cite{Poole} (see Methods section).
+  Unlike the previous steps of the pipeline, where were we evaluated the performance of methods on mock datasets, there is no equivalent dataset that contain a set of known interactions for the evaluation of the network inference algorithms.
+  Therefore, we propose the construction of a consensus network (Figure~\ref{fig:figure5}C) involving \ac{magma}, \ac{mldm}, \ac{spieceasi} and \ac{sparcc}.
+This consensus network is built by merging the p-values generated from bootstraps of the original taxonomy composition table using the Browns p-value combining method~\cite{Poole} (see Methods section).
   Based on this approach, \ac{micone} reports as default output the consensus network, annotated with weights (correlations for \ac{sparcc} and direct associations for the other methods) for all four methods.

### 2020-08-27 sections/introduction.tex
Updates from Overleaf

@@
  For example, multiple studies have shown the importance of specific microbial interactions in the healthy microbiome \cite{Lloyd-Price2016} and others have shown how changes in these interactions can lead to dysbiosis \cite{Wang2017,Gilbert2016,Belizario2015}.
- In the context of terrestrial biogeochemistry, co-occurrence networks have proposed as a valuable approach towards reconstructing the processes leading to microbiome assembly [REF Fierer], and understanding the response of microbial communities to environmental perturbations [REF].
+ In the context of terrestrial biogeochemistry, co-occurrence networks have proposed as a valuable approach towards reconstructing the processes leading to microbiome assembly \cite{Fierer2017}, and understanding the response of microbial communities to environmental perturbations \cite{Jiao2019}.
 
  Direct high-throughput measurement of interactions, e.g.
-through co-culture microdroplet experiments [REF Blainey], or spatial visualization of natural communities [REF Borisy] is possible, but it requires specific technological capabilities, and has not been used very extensively yet.
+through co-culture microdroplet experiments \cite{Hsu2019,Jian2020}, or spatial visualization of natural communities \cite{Wilbert2020} is possible, but it requires specific technological capabilities, and has not been used very extensively yet.
  In parallel, sequencing data across multiple samples can be used for estimating co-occurrence relationships between taxa.
@@
 
-One of the most frequently used avenues for inferring co-occurrence networks is the parsing and analysis of 16S sequencing data [REF].
-A large number of software tools and pipelines have been developed to analyze 16S sequencing data, often focused on addressing the many known limitations of this methodology, including resolution, sequencing depth, compositional nature, sequencing errors and copy number variations [REF].
+One of the most frequently used avenues for inferring co-occurrence networks is the parsing and analysis of 16S sequencing data \cite{Rottjers2018,Friedman2012}.
+A large number of software tools and pipelines have been developed to analyze 16S sequencing data, often focused on addressing the many known limitations of this methodology, including resolution, sequencing depth, compositional nature, sequencing errors and copy number variations \cite{Bharti2019,Pollock2018}.
 Popular methods for different phases of the analysis of 16S data include tools for: (i) denoising and clustering sequencing reads \cite{Caporaso2010,Callahan2016}; (ii) assigning taxonomy to the reads \cite{DeSantis2006,Quast2012}; (iii) processing and transforming the taxonomy count matrices \cite{Weiss2015}; and (iv) inferring the co-occurrence network \cite{Cougoul2019,Kurtz2015}.

### 2021-01-11 sections/abstract.tex
Merge with files from overleaf

@@
     In this study, we perform a meticulous analysis of each step of a pipeline that can convert 16S sequencing data into a network of microbial associations.
-    Through this process, we map how different choices of algorithms and parameters affect the co-occurrence network, and estimate steps which contribute most significantly to the variance.
-We further determine the tools and parameters that generate the most accurate and robust co-occurrence networks, based on comparison with mock and synthetic datasets.
-    Ultimately, we develop a standardized pipeline that follows these default tools and parameters, but that can also help explore the outcome of any other combination of choices.
-    We envisage that this pipeline for processing 16S sequencing data into networks of microbial co-occurrences could be used for integrating multiple data-sets, and for generating comparative analyses and consensus networks that can help understand and control microbial community assembly in different biomes.
-    The pipeline is available as a Python package at \href{https://github.com/segrelab/MiCoNE}{https://github.com/segrelab/MiCoNE}
+    Through this process, we map how different choices of algorithms and parameters affect the co-occurrence network and estimate steps that contribute most significantly to the variance.
+We further determine the tools and parameters that generate the most accurate and robust co-occurrence networks based on comparison with mock and synthetic datasets.
+    Ultimately, we develop a standardized pipeline (available at \href{https://github.com/segrelab/MiCoNE}{https://github.com/segrelab/MiCoNE}) that follows these default tools and parameters, but that can also help explore the outcome of any other combination of choices.
+    We envisage that this pipeline could be used for integrating multiple data-sets, and for generating comparative analyses and consensus networks that can help understand and control microbial community assembly in different biomes.
 }  
@@
 \keywords{Microbiome, 16S rRNA, Pipeline, Interaction, Denoising, Taxonomy, Network Inference, Correlations, Qiime, Co-occurrence, Networks}
+
+\section*{Importance}
+
+To understand and control the mechanisms that determine the structure and function of microbial communities, it is important to map the interrelationships between its constituent microbial species.
+The surge in the high-throughput sequencing of microbial communities has led to the creation of thousands of datasets containing information about microbial abundances.
+These abundances can be transformed into networks of co-occurrences across multiple samples, providing a glimpse into the structure of microbiomes.
+However, processing these datasets to obtain co-occurrence information relies on several complex steps, each of which involves multiple choices of tools and corresponding parameters.
+These multiple options pose questions about the accuracy and uniqueness of the inferred networks.
+In this study, we address this workflow and provide a systematic analysis of how these choices of tools and parameters affect the final network, and on how to select those that are most appropriate for a particular dataset.

### 2021-01-11 sections/acknowledgments.tex
Merge with files from overleaf

@@
 This work was partially funded by grants from the National Institutes of Health (National Institute of General Medical Sciences, award R01GM121950; National Institute of Dental and Craniofacial Research, award number R01DE024468; and National Institute on Aging, award number UH2AG064704), the U.S.
-Department of Energy, Office of Science, Office of Biological \& Environmental Research through the Microbial Community Analysis and Functional Evaluation in Soils SFA Program (m-CAFEs) under contract number DE-AC02-05CH11231 to Lawrence Berkeley National Laboratory, the National Science Foundation (grant NSFOCE-BSF 1635070), the Human Frontiers Science Program (RGP0020/2016), and the Boston University Interdisciplinary Biomedical Research Office.
+Department of Energy, Office of Science, Office of Biological \& Environmental Research through the Microbial Community Analysis and Functional Evaluation in Soils SFA Program (m-CAFEs) under contract number DE-AC02-05CH11231 to Lawrence Berkeley National Laboratory, the National Science Foundation (grants 1457695 and NSFOCE-BSF 1635070), the Human Frontiers Science Program (RGP0020/2016), and the Boston University Interdisciplinary Biomedical Research Office.
+KSK was supported by Simons Foundation Grant \#409704, by the Research Corporation for Science Advancement through Cottrell Scholar Award \#24010, by the Scialog grant \#26119, and by the Gordon and Betty Moore Foundation grant \#6790.08.
 

### 2021-01-11 sections/coverpage.tex
Merge with files from overleaf

@@
   Dileep Kishore\textsuperscript{1,2},
-  Gabriel Birzu\textsuperscript{3},
+  Gabriel Birzu\textsuperscript{3,6},
   Zhenjun Hu\textsuperscript{1},
   Charles DeLisi\textsuperscript{1,3},
-  Kirill Korolev\textsuperscript{$\dagger$1,2,3},\\
+  Kirill S. Korolev\textsuperscript{$\dagger$1,2,3},\\
   Daniel Segr\`{e}\textsuperscript{$\dagger$1,2,4,5}\\
@@
   \textsuperscript{5}Department of Biology, Boston University, Boston, Massachusetts, USA\\
+  \textsuperscript{6}Department of Applied Physics, Stanford University, Stanford, California, USA\\
   \textsuperscript{$\dagger$}Correspondence should be sent to \href{mailto:korolev@bu.edu}{korolev@bu.edu} or \href{mailto:dsegre@bu.edu}{dsegre@bu.edu}\\

### 2021-01-11 sections/discussion.tex
Merge with files from overleaf

@@
 The default reference database in the pipeline is the \ac{ncbi} 16S RefSeq database as it is more frequently updated and is most compatible with the blast+ query tool.
-We also enable users to use custom databases \cite{Griffen2011,Ritari2015} with the blast and naive bayes classifiers that are incorporated into the pipeline (from \ac{qiime2}).
+We also enable users to use custom databases \cite{Ritari2015} with the blast+ and naive bayes classifiers that are incorporated into the pipeline (from \ac{qiime2}).
 

### 2021-01-11 sections/introduction.tex
Merge with files from overleaf

@@
  For example, multiple studies have shown the importance of specific microbial interactions in the healthy microbiome \cite{Lloyd-Price2016} and others have shown how changes in these interactions can lead to dysbiosis \cite{Wang2017,Gilbert2016,Belizario2015}.
- In the context of terrestrial biogeochemistry, co-occurrence networks have proposed as a valuable approach towards reconstructing the processes leading to microbiome assembly \cite{Fierer2017}, and understanding the response of microbial communities to environmental perturbations \cite{Jiao2019}.
+ In the context of terrestrial bio-geochemistry, co-occurrence networks have been proposed as a valuable approach towards reconstructing the processes leading to microbiome assembly \cite{Fierer2017}, and understanding the response of microbial communities to environmental perturbations \cite{Jiao2019}.
 
  Direct high-throughput measurement of interactions, e.g.
-through co-culture microdroplet experiments \cite{Hsu2019,Jian2020}, or spatial visualization of natural communities \cite{Wilbert2020} is possible, but it requires specific technological capabilities, and has not been used very extensively yet.
+through co-culture micro-droplet experiments \cite{Hsu2019,Jian2020}, or spatial visualization of natural communities \cite{Wilbert2020} is possible, but it requires specific technological capabilities, and has yet to be extensively used.
  In parallel, sequencing data across multiple samples can be used for estimating co-occurrence relationships between taxa.
@@
  This is partially due to the fact that existing platforms are typically focused on \ac{otu} generation and not on the effects of upstream statistical methods on the inferred co-occurrence networks.
- Furthermore, no organized framework currently exist to systematically analyze and compare existing components of the data-analysis from amplicons to networks.
+ Furthermore, no organized framework currently exist to systematically analyze and compare existing components of the data analysis from amplicons to networks.
   More broadly, given the lack of comprehensive comparisons between directly observed microbial interactions (e.g.
@@
 
- In this study, we present a standardized 16S data-analysis pipeline called \ac{micone} that produces robust and reproducible co-occurrence networks from community 16S sequence data, and allow users to interactively explore how the network would change upon using different alternative tools and parameters at each step.
+ In this study, we present a standardized 16S data analysis pipeline called \ac{micone} that produces robust and reproducible co-occurrence networks from community 16S sequence data, and allow users to interactively explore how the network would change upon using different alternative tools and parameters at each step.
   % TODO: Link or describe MIND here (?)
@@
   In addition to making this tool freely available, we implemented a systematic comparative analysis to determine which steps of the pipeline have the largest influence on the final network, and what choice seems to provide best agreement with the tested mock and synthetic datasets.
-  We believe that these steps will ensure better reproducibility and easier comparison of co-occurrence networks across data-sets.
+  We believe that these steps will ensure better reproducibility and easier comparison of co-occurrence networks across datasets.
   We expect that our tool will also be useful for benchmarking future alternative methods, and for ensuring a transparent evaluation of the possible biases introduced by the use of specific tools.

### 2021-01-11 sections/methods.tex
Merge with files from overleaf

@@
   This modular architecture simplifies the process of adding new tools (refer to modules section in the \ac{micone} documentation).
-  In addition to the Python package, the entire pipeline has been containerized into a Docker~\cite{Merkel1994} image (\hl{dockerhub link}) for easy deployment and setup.
+%   In addition to the Python package, the entire pipeline has been containerized into a Docker~\cite{Merkel1994} image (\hl{dockerhub link}) for easy deployment and setup.
   The main components of the pipeline are detailed in the subsequent sections.
@@
 
-  In other words, $N$ is the merged table, each column $N_i$ is the vector representation of one of the networks, and each row $L_i$ represents the one particular edge in all networks (assigned 0 if the edge does not exist in the network).
+  In other words, $N$ is the merged table, each column $N_i$ is the vector representation of one of the networks, and each row $L_i$ represents one particular edge in all networks (assigned 0 if the edge does not exist in the network).
 

### 2021-01-11 sections/results.tex
Merge with files from overleaf

@@
 
-  \begin{figure}[h]
-    \centering
-    \includegraphics[width=0.69\linewidth]{figure1.pdf}
-    \caption{
-      \textbf{The workflow of the microbial co-occurrence analysis pipeline}.
-      The steps can be grouped into four major groups: \textbf{(DC)} \textbf{D}enoising and \textbf{C}lustering, \textbf{(TA)} \textbf{T}axonomy \textbf{A}ssignment, \textbf{(OP)} \textbf{O}TU or ESV \textbf{P}rocessing, and \textbf{(NI)} \textbf{N}etwork \textbf{I}nference.
-      Each step incorporates several processes, each of which in turn have several alternate algorithms for the same task (indicated by the text to the right of the blue boxes).
-      The text along the arrows describes the data that is being passed from one step to another. For details on each process and data types, see Methods.
-    }
-    \label{fig:figure1}
-  \end{figure}
-
-  Our analysis involves two types of data: The first type consists of sets of 16S sequencing data for real communities consisting of Stool and Oral microbiomes.
+  Our analysis involves two types of data: The first type consists of sets of 16S sequencing data from real communities sampled from human Stool and Oral microbiomes.
   The second are datasets synthetically or artificially created for the specific goal of helping evaluate computational analysis tools (see Methods).
@@
   These results suggest that the most important criterion for accurate comparative analyses of co-occurrence networks is the taxonomy reference database.
-
-  \begin{figure}[H]
-    \centering
-    \includegraphics[width=0.85\linewidth]{figure2.pdf}
-  \end{figure}
-  \begin{figure}[!t]
-    \centering
-      \caption{
-      \textbf{The choice of database contributes to the most variance in the networks}.
-      \textbf{(A)} The total relative variance in the networks contributed by the DC, TA and OP steps of the pipeline (right) and the linear model used to calculate the relative variance (left), see the Methods section for details.
-      \textbf{(B)} All combinations of inferred networks are shown as points on a PCA plot.
-      The color of the points corresponds to the taxonomy database, the shape corresponds to the denoising/clustering method and the size corresponds to whether low abundance OTUs were removed or not.
-      \textbf{(B inset)} The network generated using DC=dada2, TA=GG, OP=no and NI=SPARCC and represents the particular point shown (big red square).
-      The plot clearly shows that the points can be separated based on the TA step and that the differences due to the DC and OP steps are not as significant.
-    }
-    \label{fig:figure2}
-  \end{figure}
 
@@
   Given the unexpected poor performance of Deblur on the synthetic data, the default algorithm in the pipeline was chosen to be \ac{dada2} (Supplementary Figure~\ref{fig:figureS3}).
-
-  \begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{figure3.pdf}
-    \caption{
-      \textbf{The representative sequences generated by the different denoising/clustering methods are very similar but differ in the sequences that are in low abundance.}
-      \textbf{(A)} The average weighted UniFrac distance between the representative sequences shows that the representative sequences and their compositions are fairly identical between the methods,
-      \textbf{(B)} The relatively larger average unweighted UniFrac distance indicates that methods differ in their identification of sequences of low abundance,
-      \textbf{(C, D)} The distributions of the average weighted UniFrac distance between the expected sequence profile and the calculated sequence profile in mock datasets.
-      \textbf{(D)} The distributions of the average unweighted UniFrac distance show that dada2 and Deblur were the best performing methods in most of the datasets.
-    }
-    \label{fig:figure3}
-  \end{figure}
 
@@
 
-  \begin{figure}[H]
-    \centering
-    \includegraphics[width=0.9\textwidth,height=1.2\textwidth]{figure4.pdf}
-  \end{figure}
-  \begin{figure}[!t]
-    \centering
-    \caption{
-      \textbf{Taxonomic reference databases vary widely in terms of their taxonomy assignments.}
-      \textbf{(A)} The assignment of the top 50 representative sequences to their respective taxonomies using the three different reference databases shows how the same sequences are assigned to different Genus.
-      \textbf{(B)} The percentage of \ac{otu}s assigned to the same taxonomic label when using different reference databases.
-      The percentage of mismatches decrease at higher taxonomic levels but even at the Phylum level there exists around 10\% of mismatches.
-      \textbf{(C)} The Bray-Curtis dissimilarity between the expected taxonomy profile and calculated taxonomy profile in the mock datasets shows that there is no singular best choice of database for every dataset.
-    }
-    \label{fig:figure4}
-  \end{figure}
-
   \FloatBarrier
@@
 Note, however, that some of these comparisons depend on the threshold that has to be applied to the pairwise correlations methods (currently 0.3, based on~\cite{Friedman2012}).
-  To get a more quantitative idea, we can take a look at the nodes and edges (Figure~\ref{fig:figure5}B) in common between the networks using UpSet plots~\cite{Lex} (only \ac{magma}, \ac{mldm}, \ac{spieceasi}, \ac{sparcc} are used in the comparison since Pearson and Spearman add a large number of spurious edges since they are not intended for compositional datasets).
+  To get a more quantitative picture of the differences between the inferred networks, we checked the distribution of common nodes and edges (Figure~\ref{fig:figure5}B) using UpSet plots~\cite{Lex} (only \ac{magma}, \ac{mldm}, \ac{spieceasi}, \ac{sparcc} are used in the comparison since Pearson and Spearman add a large number of spurious edges since they are not intended for compositional datasets).
   The results for the node intersections show that the networks have a large number of nodes in common ($63$ out of $67$ nodes in the smallest network - \ac{magma}) and no network possesses any unique node.
@@
   Based on this approach, \ac{micone} reports as default output the consensus network, annotated with weights (correlations for \ac{sparcc} and direct associations for the other methods) for all four methods.
-
-  \begin{figure}[H]
-    \centering
-    \includegraphics[width=0.9\textwidth]{figure5.pdf}
-  \end{figure}
-  \begin{figure}[t!]
-    \centering
-    \caption{
-      \textbf{Networks generated using different network inference methods show notable differences both in terms of edge-density and connectivity}.
-      \textbf{(A)} The six different networks generated by the different network inference methods are very dissimilar.
-      The green links are positive associations and the orange links are negative associations.
-      A threshold of 0.3 was set for the methods that infer pairwise correlations (\ac{sparcc}, Spearman, Pearson) and no threshold was set for the other methods.
-      \textbf{(B)} The node overlap Upset plot~\cite{Lex} indicates that all the networks have a large number of common nodes involved in connections.
-      Whereas, \textbf{(C)} The edge overlap Upset plot shows that a very small fraction of these connections are actually shared.
-    }
-    \label{fig:figure5}
-  \end{figure}
 
@@
 
-  \begin{table}[h]
-    \centering
-    \small
-    \begin{tabular}{|c|c|c|}
-      \hline
-      \textbf{Process} & \textbf{Tool} & \textbf{Parameters} \\
-      \hline
-      Denoising and Clustering & Dada2/Deblur & default \\
-      Taxonomy Assignment & \ac{ncbi} with Blast & RefSeq database \\
-      OTU Processing & Based on statistical power & Dynamic cutoff \\
-      Network Inference & Consensus method & - \\
-      \hline
-    \end{tabular}
-    \caption{Default tools and parameters for the pipeline}
-    \label{tab:default_options}
-  \end{table}
-
   The recommended tool for the \ac{dc} step (\ac{dada2} or Deblur) were chosen based on their accuracy in recapitulating the reference sequences in mock communities and synthetic data.
@@
 These consensus networks were generated using the default tools and parameters from Table~\ref{tab:default_options}.
-
-  \begin{figure}[h]
-    \centering
-    \includegraphics[width=0.85\linewidth]{figure6.pdf}
-    \caption{
-      \textbf{Network inference and taxonomic assignment have the highest influence on the inferred network structures.}
-      \textbf{(A)} The network constructed using the default pipeline parameters (DC=\ac{dada2}, TA=\ac{ncbi}, OP=on, NI=\ac{sparcc}) is compared with networks generated when one of the steps use a different tool.
-      The common connections (common with the default network) are in black, connections unique to the network are colored purple and connections in the default network but not present in the current network are gray.
-      \textbf{(B)} The L1 distance between the networks generated by changing one step of the default pipeline and the network generated using the default parameters.
-    }
-    \label{fig:figure6}
-  \end{figure}
-
-  \begin{figure}[h]
-    \centering
-    \includegraphics[width=17cm]{figure7.pdf}
-    \caption{
-      The consensus networks generated using the default pipeline settings.
-      \textbf{(A)} Co-occurrence network of the Hard Palate microbiome generated from samples of healthy subjects in a periodontal diseases study.
-      \textbf{(B)} Co-occurrence network of the Stool microbiome generated from samples of healthy subjects in a fecal microbiome transplant study.
-  }
-    \label{fig:figure7}
-  \end{figure}

### 2021-01-11 sections/supplementary.tex
Add supplementary table
Update supplementary table
Merge with files from overleaf

--- 2020-05-29 sections/supplementary_figures.tex
@@
 
+  \renewcommand{\thetable}{S\arabic{table}}
+  \setcounter{table}{0}
+
+  \input{sections/supplementary_table.tex}
+
   \begin{figure}[h]
@@
     (A, B) Correlation of the abundances of the taxa that are in common between the count matrices created by two different methods.
-    (A) The best correlation (most similar methods) is between open-reference and denovo.
-    (B) The worst correlation (least similar methods) is between open-reference and dada2.
+    (A) The worst correlation (least similar methods) is between open-reference and dada2.
+    (B) The best correlation (most similar methods) is between open-reference and denovo.
     (C) A heatmap showing the $\mathrm{R}^2$ of all pairwise comparisons of the methods.
@@
 
-  \begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{figureS6.pdf}
-    \caption{
-      Calculation of presence threshold that is applied on the OTU table in the OTU processing (OP) step of the pipeline.
-      This presence threhold $p_t$ is dependent on the number of samples in the dataset and the required correlation stength threshold.
-  }
-    \label{fig:figureS6}
-  \end{figure}
+%   \begin{figure}[h]
+%     \centering
+%     \includegraphics[width=\linewidth]{figureS6.pdf}
+%     \caption{
+%       Calculation of presence threshold that is applied on the OTU table in the OTU processing (OP) step of the pipeline.
+%       This presence threhold $p_t$ is dependent on the number of samples in the dataset and the required correlation stength threshold.
+%   }
+%     \label{fig:figureS6}
+%   \end{figure}
 
@@
     \includegraphics[width=0.8\linewidth]{pdf/denoise_network.pdf}
-    \caption{A network showing union and intersection of networks generated using certain combination of methods}
+    \caption{A network showing union (A) and intersection (B) of networks generated using different denoising and clustering tools on the Stool dataset.}
     \label{fig:figureS5}

### 2021-01-11 sections/tables_figures.tex
Merge with files from overleaf

--- 2019-11-25 figures.tex
@@
-%--------------------------------------------------------%
+%!TEX root = ../main.tex
+
+\newpage
+\section*{Tables and Figures}
+
+% Tables
+
+  \begin{table}[h]
+    \centering
+    \small
+    \begin{tabular}{|c|c|c|}
+      \hline
+      \textbf{Process} & \textbf{Tool} & \textbf{Parameters} \\
+      \hline
+      Denoising and Clustering & Dada2/Deblur & default \\
+      Taxonomy Assignment & \ac{ncbi} with Blast & RefSeq database \\
+      OTU Processing & Based on statistical power & Dynamic cutoff \\
+      Network Inference & Consensus method & - \\
+      \hline
+    \end{tabular}
+    \caption{Default tools and parameters for the pipeline}
+    \label{tab:default_options}
+  \end{table}
+
 % Figures
-%--------------------------------------------------------%
-
-\input{sections/preamble}
-
-% \titleformat{\subsection}[runin]{}{}{}{}[]
-
-\title{ \huge Inferring microbial co-occurrence networks from 16S data: A systematic evaluation }
-\author{DK, GB, ZH, CD, KK, DS}
-\date{\today}
-
-\begin{document}
-
-\maketitle
 
@@
       \textbf{The workflow of the microbial co-occurrence analysis pipeline}.
-    The processes can be grouped into four major steps: \textbf{(DC)} denoising and clustering, \textbf{(TA)} taxonomy assignment, \textbf{(OP)} OTU/ESV processing, and \textbf{(NI)} network inference.
+      The steps can be grouped into four major groups: \textbf{(DC)} \textbf{D}enoising and \textbf{C}lustering, \textbf{(TA)} \textbf{T}axonomy \textbf{A}ssignment, \textbf{(OP)} \textbf{O}TU or ESV \textbf{P}rocessing, and \textbf{(NI)} \textbf{N}etwork \textbf{I}nference.
       Each step incorporates several processes, each of which in turn have several alternate algorithms for the same task (indicated by the text to the right of the blue boxes).
-    The text along the arrows describes the data that is being passed from one step to another.
+      The text along the arrows describes the data that is being passed from one step to another. For details on each process and data types, see Methods.
     }
@@
   \end{figure}
+  \FloatBarrier
+  \newpage
 
-\begin{figure}[h]
+  \begin{figure}[H]
     \centering
-  \includegraphics[width=0.75\linewidth]{figure2.pdf}
+    \includegraphics[width=0.85\linewidth]{figure2.pdf}
+  \end{figure}
+  \begin{figure}[H]
+    \centering
       \caption{
       \textbf{The choice of database contributes to the most variance in the networks}.
-    \textbf{(A)} The relative variance in the networks contributed by the DC, TA and OP steps of the pipeline.
+      \textbf{(A)} The total relative variance in the networks contributed by the DC, TA and OP steps of the pipeline (right) and the linear model used to calculate the relative variance (left), see the Methods section for details.
       \textbf{(B)} All combinations of inferred networks are shown as points on a PCA plot.
       The color of the points corresponds to the taxonomy database, the shape corresponds to the denoising/clustering method and the size corresponds to whether low abundance OTUs were removed or not.
-    \textbf{(B inset)} The network generated using DC=dada2, TA=Greengenes, OP=off and NI=SPARCC and represents the particular point shown (green big triangle).
+      \textbf{(B inset)} The network generated using DC=dada2, TA=GG, OP=no and NI=SPARCC and represents the particular point shown (big red square).
       The plot clearly shows that the points can be separated based on the TA step and that the differences due to the DC and OP steps are not as significant.
@@
   \end{figure}
+  \FloatBarrier
+  \newpage
 
-\begin{figure}[h]
+  \begin{figure}
     \centering
-  \includegraphics[width=0.8\linewidth]{figure3.pdf}
+    \includegraphics[width=\textwidth]{figure3.pdf}
     \caption{
-    \textbf{The representative sequences generated by the different denoising/clustering methods are very similar but differ slightly in the sequences that are in low abundance.}
+      \textbf{The representative sequences generated by the different denoising/clustering methods are very similar but differ in the sequences that are in low abundance.}
       \textbf{(A)} The average weighted UniFrac distance between the representative sequences shows that the representative sequences and their compositions are fairly identical between the methods,
       \textbf{(B)} The relatively larger average unweighted UniFrac distance indicates that methods differ in their identification of sequences of low abundance,
-    \textbf{(C)} The distributions of the average unweighted UniFrac distance between the expected sequence profile and the calculated sequence profile in mock datasets show that dada2 and Deblur were the best performing methods in most of the datasets.
+      \textbf{(C, D)} The distributions of the average weighted UniFrac distance between the expected sequence profile and the calculated sequence profile in mock datasets.
+      \textbf{(D)} The distributions of the average unweighted UniFrac distance show that dada2 and Deblur were the best performing methods in most of the datasets.
     }
@@
   \end{figure}
+  \FloatBarrier
+  \newpage
 
-\begin{figure}[h]
+  \begin{figure}[H]
     \centering
-  \includegraphics[width=0.75\linewidth]{figure4.pdf}
+    \includegraphics[width=0.9\textwidth,height=1.2\textwidth]{figure4.pdf}
+  \end{figure}
+  \begin{figure}[H]
+    \centering
     \caption{
-    \textbf{Taxonomic reference databases vary widely in terms of their taxonomies.}
-    \textbf{(A)} The assignment of the top 50 representative sequences to their respective taxonomies using the three different reference databases shows how the same sequences are assigned to different genus.
-    \textbf{(B)} The percentage of \ac{OTU}s assigned to the same taxonomic label when using different reference databases.
+      \textbf{Taxonomic reference databases vary widely in terms of their taxonomy assignments.}
+      \textbf{(A)} The assignment of the top 50 representative sequences to their respective taxonomies using the three different reference databases shows how the same sequences are assigned to different Genus.
+      \textbf{(B)} The percentage of \ac{otu}s assigned to the same taxonomic label when using different reference databases.
       The percentage of mismatches decrease at higher taxonomic levels but even at the Phylum level there exists around 10\% of mismatches.
-    \textbf{(C)} The Bray-Curtis dissimilarity between the expected taxonomy profile and calculated taxonomy profile in the mock datasets shows that there is no best choice of database for every dataset.
+      \textbf{(C)} The Bray-Curtis dissimilarity between the expected taxonomy profile and calculated taxonomy profile in the mock datasets shows that there is no singular best choice of database for every dataset.
     }
@@
   \end{figure}
+  \FloatBarrier
+  \newpage
 
-\begin{figure}[h]
+  \begin{figure}[H]
     \centering
-  \includegraphics[width=0.85\linewidth]{figure5.pdf}
+    \includegraphics[width=0.9\textwidth]{figure5.pdf}
+  \end{figure}
+  \begin{figure}[H]
+    \centering
     \caption{
@@
       A threshold of 0.3 was set for the methods that infer pairwise correlations (\ac{sparcc}, Spearman, Pearson) and no threshold was set for the other methods.
-    \textbf{(B)} The node overlap upset plot indicates that all the networks have a large number of common nodes involved in connections.
-    Whereas, \textbf{(C)} The edge overlap upset plot shows that a very small fraction of these connections are actually shared.
+      \textbf{(B)} The node overlap Upset plot~\cite{Lex} indicates that all the networks have a large number of common nodes involved in connections.
+      Whereas, \textbf{(C)} The edge overlap Upset plot shows that a very small fraction of these connections are actually shared.
     }
@@
   \end{figure}
+  \FloatBarrier
+  \newpage
 
@@
     \centering
-  \includegraphics[width=0.9\linewidth]{figure6.pdf}
+    \includegraphics[width=0.85\linewidth]{figure6.pdf}
     \caption{
@@
       \textbf{(A)} The network constructed using the default pipeline parameters (DC=\ac{dada2}, TA=\ac{ncbi}, OP=on, NI=\ac{sparcc}) is compared with networks generated when one of the steps use a different tool.
-    The common connections (common with default) are in black, connections unique to the network are colored purple and connections in the default network and but not present in the current network are gray.
-    \textbf{(B)} The L1 distance of the networks generated by changing one step of the default pipeline against the network generated using the default parameters.
+      The common connections (common with the default network) are in black, connections unique to the network are colored purple and connections in the default network but not present in the current network are gray.
+      \textbf{(B)} The L1 distance between the networks generated by changing one step of the default pipeline and the network generated using the default parameters.
     }
@@
     \caption{
-    The consensus networks generated using the default pipeline settings for the Hard Palate \textbf{(A)} and Stool \textbf{(B)} body-sites.
+      The consensus networks generated using the default pipeline settings.
+      \textbf{(A)} Co-occurrence network of the Hard Palate microbiome generated from samples of healthy subjects in a periodontal diseases study.
+      \textbf{(B)} Co-occurrence network of the Stool microbiome generated from samples of healthy subjects in a fecal microbiome transplant study.
   }
@@
   \end{figure}
-
-% \FloatBarrier
-
-% \subsection*{Supplementary}%
-
-% \setcounter{figure}{0}
-% \renewcommand{\thefigure}{S\arabic{figure}}
-
-% \begin{figure}[h]
-%   \centering
-%   \includegraphics[width=0.67\linewidth]{figureS1.pdf}
-%   \caption{
-%     \textbf{Comparison of various denoising and clustering algorithms used in the pipeline}.
-%     (A, B) Correlation of the abundances of the taxa that are common between the count matrices created by two different methods.
-%     (A) The best correlation (most similar methods) is between open-reference and denovo.
-%     (B) The worst correlation (least similar methods) is between open-reference and dada2.
-%     (C) A heatmap showing the $\mathrm{R}^2$ of all pairwise comparisons of the methods.
-%   }
-%   \label{fig:figureS1}
-% \end{figure}
-
-% \begin{figure}[h]
-%   \centering
-%   \includegraphics[width=0.8\linewidth]{figureS2.pdf}
-%   \caption{Results of the mock community analysis. (A) Weighted unifrac (B) Unweighted unifrac}
-%   \label{fig:figureS2}
-% \end{figure}
-
-% \begin{figure}[h]
-%   \centering
-%   \includegraphics[width=0.8\linewidth]{figureS3.pdf}
-%   \caption{Results of the synthetic community analysis. (A) Weighted unifrac (B) Unweighted unifrac}
-%   \label{fig:figureS3}
-% \end{figure}
-
-% \begin{figure}[h]
-%   \centering
-%   \includegraphics[width=0.8\linewidth]{pdf/all_denoise_reg.pdf}
-%   \caption{All pairwise correlations comparing the similarity between different denoising and clustering methods}
-%   \label{fig:figureS4}
-% \end{figure}
-
-% \begin{figure}[h]
-%   \centering
-%   \includegraphics[width=\linewidth]{figureS4.pdf}
-%   \caption{
-%     \textbf{(A)} Taxonomy composition of the 20 most abundant genera predicted using different taxonomy references databases: Greengenes, SILVA and NCBI.
-%     The legend shows the common and the unique genera among the taxonomy assignments.
-% }
-%   \label{fig:figureS4}
-% \end{figure}
-
-% \begin{figure}[h]
-%   \centering
-%   \includegraphics[width=\linewidth]{figureS5.pdf}
-%   \caption{
-%     The bray-curtis dissmilarity between the expected taxonomic composition and generated taxonomic composiion for the synthetic datasets.
-% }
-%   \label{fig:figureS4}
-% \end{figure}
-
-% \begin{figure}[h]
-%   \centering
-%   \includegraphics[width=\linewidth]{figureS6.pdf}
-%   \caption{
-%     The effect of OTU processing measurs on the OTU table
-% }
-%   \label{fig:figureS4}
-% \end{figure}
-
-% \begin{figure}[h]
-%   \centering
-%   \includegraphics[width=0.8\linewidth]{png/hamming_distance.png}
-%   \caption{The hamming distance between the networks is shown. The similarity between various methods was found to vary with the data-source used.}
-%   \label{fig:figureS4}
-% \end{figure}
-
-% \begin{figure}[h]
-%   \centering
-%   \includegraphics[width=0.8\linewidth]{pdf/denoise_network.pdf}
-%   \caption{A network showing union and intersection of networks generated using certain combination of methods}
-%   \label{fig:figureS5}
-% \end{figure}
-
-\end{document}

### 2021-01-20 sections/methods.tex
Add: "pvalue merging" details to methods section

@@
 
+ The merged table $N$ with $p$ edges and $n$ networks, where, each column $N_j$ is the vector representation of one of the networks, each row $L_i$ is the vector representation of one particular edge in all networks (assigned a value of 0 if the edge did not exist in the network but in other networks), and each element $E_{i,j}$ belongs to edge $L_i$ and network $N_j$.
+
   \begin{equation*}
-    N = \begin{bmatrix}
-     edge_{1,1} & edge_{2,1} & \cdots  & edge_{n, 1} \\ 
-     edge_{1,2} & edge_{2,2} & \cdots  & edge_{n, 2} \\ 
+   N_{p \times n} =
+   \begin{bmatrix}
+     L_1 \\
+     L_2 \\
+     \vdots \\
+     L_p
+   \end{bmatrix}
+   =
+   \begin{bmatrix}
+     E_{1,1} & E_{1,2} & \cdots  & E_{1, n} \\
+     E_{2,1} & E_{2,2} & \cdots  & E_{2, n} \\
      \vdots & \vdots & \vdots  & \vdots \\
-     edge_{1,n} & edge_{2,n} & \cdots  & edge_{n, n}
+     E_{p,1} & E_{p,2} & \cdots  & E_{p, n}
    \end{bmatrix}
@@
 
+  \subsection*{Consensus Network and p-value merging}
+  \vspace{-5mm}
+  Let there be $k$ p-values, denoted by $P_i$, generated from $k$ different methods.
+
+  Fisher~\cite{fisher_224a_1948} proposed that for $k$ independent p-values, the statistic $\Psi$:
+  \begin{equation*}
+    \Psi = \sum_{i=1}^k -2 \log \left( P_i \right)
+    \Psi \sim \chi^2_{2k}
+  \end{equation*}
+
+  Brown~\cite{brown_400_1975} extended Fisher's method to dependent p-values by using a re-scaled $\chi^2$ distribution:
+  \begin{equation*}
+    \Psi \sim c \chi^2_{2f}
+  \end{equation*}
+  where, $f$ is the degrees of freedom and $c$ is the scale factor and are given by:
+  \begin{equation*}
+    f = \frac{\mathrm{E}[\Psi]^2}{\mathrm{Var}[\Psi]} ~~~\text{and}~~~ c = \frac{\mathrm{Var}[\Psi]}{2\mathrm{E}[\Psi]} = \frac{k}{f}
+  \end{equation*}
+
+  Furthermore, Brown showed that $\mathrm{E}[\Psi]$ and $\mathrm{Var}[\Psi]$ can be calculated directly via a numerical integration:
+  \begin{equation*}
+    \mathrm{E}[\Psi] = 2k ~~~\text{and}~~~ \mathrm{Var}[\Psi] = 4k + 2\sum_{i<j} \mathrm{Cov} \left( -2\log(P_i), -2\log(P_j) \right)
+  \end{equation*}
+
+  Kost and McDermott~\cite{kost_combining_2002} further fit a third-order polynomial to approximate the covariance
+  \begin{equation*}
+    \mathrm{Cov} \left( -2\log(P_i), -2\log(P_j) \right) \approx 3.263 \rho_{ij} + 0.710 \rho_{ij}^2 + 0.027 \rho_{ij}^3
+  \end{equation*}
+  where, $\rho_{ij}$ is the correlation between method $i$ and method $j$
+
+  The final combined p-value is then given by:
+  \begin{equation*}
+    P_{combined} = 1.0 - \Phi_{2f}\left( \psi / c \right)
+    \text{where},~ $\psi = -2 \sum_{i=1}^k \log(P_i) ~~~\text{and}~~~ \Phi_{2f} = \mathrm{CDF}\left( \chi^2_{2f} \right)
+  \end{equation*}
+
   \subsection*{Code and Data Availability}

### 2021-01-25 sections/methods.tex
Update on Overleaf.
Update datasets and consensus network in methods section

@@
   The current study used healthy stool samples from a fecal microbiome transplant study~\cite{Kang2017} and healthy saliva samples from a periodontal disease study~\cite{Chen2018} as real datasets for analysis.
+  % TODO: How many samples are there for each and average number of reads.
   The mock community 16S datasets are real sequencing data obtained for artificially assembled collections of species in known proportions.
@@
   Three different microbial composition profiles were used as input; reads were generated using a soil and water microbiome composition profiles from the \ac{emp}~\cite{Thompson2017} and healthy gut microbiome project from the fecal microbiome transplant study~\cite{Kang2017}.
-  The reads are simulated using the NCBI RefSeq database as the reference sequence pool and the "art\_illumina" sequence profile with a mutation rate of 2\%.
+  % TODO: Create a section in the supplementary that describes what the original profiles were and how we used ART (what settings) to generate the synthetic data.
+  The reads are simulated using the NCBI RefSeq database as the reference sequence pool and the `art\_illumina' sequence profile with a mutation rate of 2\%.
   The scripts used to generate the synthetic data are in the scripts folder of the repository (\href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper}).
@@
   \vspace{-5mm}
-  Let there be $k$ p-values, denoted by $P_i$, generated from $k$ different methods.
-
-  Fisher~\cite{fisher_224a_1948} proposed that for $k$ independent p-values, the statistic $\Psi$:
+  Fisher~\cite{fisher_224a_1948} proposed that for $k$ independent p-values, each generated by $k$ different methods and denoted by $P_i$, the statistic $\Psi$:
   \begin{equation*}
-    \Psi = \sum_{i=1}^k -2 \log \left( P_i \right)
-    \Psi \sim \chi^2_{2k}
+    \begin{aligned}
+        \Psi &= \sum_{i=1}^k -2 \log \left( P_i \right) \\
+        \Psi &\sim \chi^2_{2k}
+    \end{aligned}
   \end{equation*}
@@
   \begin{equation*}
-    \mathrm{E}[\Psi] = 2k ~~~\text{and}~~~ \mathrm{Var}[\Psi] = 4k + 2\sum_{i<j} \mathrm{Cov} \left( -2\log(P_i), -2\log(P_j) \right)
+    \mathrm{E}[\Psi] = 2k ~~~\text{and}~~~ \mathrm{Var}[\Psi] = 4k + 2\sum_{i<j} \mathrm{Cov}\left( -2\log(P_i), -2\log(P_j) \right)
   \end{equation*}
@@
   Kost and McDermott~\cite{kost_combining_2002} further fit a third-order polynomial to approximate the covariance
-  \begin{equation*}
-    \mathrm{Cov} \left( -2\log(P_i), -2\log(P_j) \right) \approx 3.263 \rho_{ij} + 0.710 \rho_{ij}^2 + 0.027 \rho_{ij}^3
-  \end{equation*}
+  \begin{equation}
+    \mathrm{Cov}\left( -2\log(P_i), -2\log(P_j) \right) \approx 3.263 \rho_{ij} + 0.710 \rho_{ij}^2 + 0.027 \rho_{ij}^3
+    \label{eqn:covariance-pvalues}
+  \end{equation}
   where, $\rho_{ij}$ is the correlation between method $i$ and method $j$
 
-  The final combined p-value is then given by:
-  \begin{equation*}
-    P_{combined} = 1.0 - \Phi_{2f}\left( \psi / c \right)
-    \text{where},~ $\psi = -2 \sum_{i=1}^k \log(P_i) ~~~\text{and}~~~ \Phi_{2f} = \mathrm{CDF}\left( \chi^2_{2f} \right)
-  \end{equation*}
+  The final combined p-value~\cite{Poole} is then given by:
+  \begin{equation}
+    \begin{aligned}
+        & P_{combined} = 1.0 - \Phi_{2f}\left( \psi / c \right) \\
+        \text{where},~ &\psi = -2 \sum_{i=1}^k \log(P_i) ~~~\text{and}~~~ \Phi_{2f} = \mathrm{CDF}\left( \chi^2_{2f} \right)
+    \end{aligned}
+    \label{eqn:pvalue-combined}
+  \end{equation}
+
+  The consensus method in \ac{micone} uses Equation~\ref{eqn:covariance-pvalues} to estimate the covariance of the pvalues and Equation~\ref{eqn:pvalue-combined} to merge the p-values (obtained from bootstrapping) from different methods.
+The combined p-values are used to threshold for significance during the consensus network step.
 

### 2021-02-03 sections/methods.tex
Update manuscript to address comment R1.1

@@
   \vspace{-5mm}
+
+  \subsubsection*{Bootstrapping}
+  For each network inference method (both correlation and direct association based methods), $1000$ permutations of the original \ac{otu} counts data were generated~\cite{Watts2018}.
+  We then recalculate the associations in these permuted \ac{otu} tables using the different network inference algorithms.
+  Finally, we calculate the p-value based on how often a more extreme association is observed for randomly permuted data.
+
+  \subsubsection*{p-value merging}
   Fisher~\cite{fisher_224a_1948} proposed that for $k$ independent p-values, each generated by $k$ different methods and denoted by $P_i$, the statistic $\Psi$:
@@
 
-  The consensus method in \ac{micone} uses Equation~\ref{eqn:covariance-pvalues} to estimate the covariance of the pvalues and Equation~\ref{eqn:pvalue-combined} to merge the p-values (obtained from bootstrapping) from different methods.
+  The consensus method in \ac{micone} (refer Documentation) uses Equation~\ref{eqn:covariance-pvalues} to estimate the covariance of the pvalues and Equation~\ref{eqn:pvalue-combined} to merge the p-values (obtained from bootstrapping) from different methods.
 The combined p-values are used to threshold for significance during the consensus network step.
@@
   Pipeline: \href{https://github.com/segrelab/MiCoNE}{https://github.com/segrelab/MiCoNE} \\
+  Documentation: \href{https://micone.readthedocs.io}{https://micone.readthedocs.io} \\
   Data and scripts: \href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper}

### 2021-02-03 sections/results.tex
Update manuscript to address comment R1.1

@@
   Therefore, we propose the construction of a consensus network (Figure~\ref{fig:figure5}C) involving \ac{magma}, \ac{mldm}, \ac{spieceasi} and \ac{sparcc}.
-This consensus network is built by merging the p-values generated from bootstraps of the original taxonomy composition table using the Browns p-value combining method~\cite{Poole} (see Methods section).
+This consensus network is built by merging the p-values generated from bootstraps of the original taxonomy counts table using the Browns p-value combining method~\cite{brown_400_1975} (see Methods section).
   Based on this approach, \ac{micone} reports as default output the consensus network, annotated with weights (correlations for \ac{sparcc} and direct associations for the other methods) for all four methods.

### 2021-02-03 sections/methods.tex
Update manuscript to address minor comments from Reviewer 1

@@
   The pipeline includes Pearson, Spearman and FastSpar v0.0.10 (a faster implementation of \ac{sparcc})~\cite{Watts2018} as the pairwise correlation metrics, and \ac{spieceasi} v1.0.7~\cite{Kurtz2015}, \ac{mldm} v1.1~\cite{Yang2017} and \ac{magma}~\cite{Cougoul2019} as the direct association metrics.
-  The empirical Browns method~\cite{Poole} is used for combining p-values from the various methods to obtain a consensus p-value, which is used to create the consensus network.
+  The Brown's p-value merging method~\cite{brown_400_1975} is used for combining p-values from the various methods to obtain a consensus p-value, which is used to create the consensus network.
 
@@
 
-  The variance contributed by each step of the pipeline is calculated for every connection in the merged table through ANOVA using the Python statsmodels package and is shown in Figure~\ref{fig:figure2}B.
+  The variance contributed by each step of the pipeline is calculated for every connection in the merged table through ANOVA using the Python statsmodels package and is shown in Figure~\ref{fig:figure2}A.
   The total variance for the network is calculated by adding the variances for each connection.
-  The PCA analysis is also performed on the merged table to generate Figure~\ref{fig:figure2}C.
+  The merged network table $N_{p \times n}$ is used as the input to the PCA analysis to generate Figure~\ref{fig:figure2}B.
 
@@
 
-  The final combined p-value~\cite{Poole} is then given by:
+  The final combined p-value~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016} is then given by:
   \begin{equation}
@@
   The consensus method in \ac{micone} (refer Documentation) uses Equation~\ref{eqn:covariance-pvalues} to estimate the covariance of the pvalues and Equation~\ref{eqn:pvalue-combined} to merge the p-values (obtained from bootstrapping) from different methods.
+  Note that we do not use Pearson and Spearman methods in the p-value merging step and these algorithms are only used for demonstration and comparison.
   The combined p-values are used to threshold for significance during the consensus network step.

### 2021-02-03 sections/methods.tex
Update manuscript to address comment R2.5

@@
   This module deals with normalization, filtering and applying transformations to the \ac{otu} or \ac{esv} counts matrix.
-  Rarefaction is a normalization technique used to overcome the bias that might arise due to variable sampling depth in different samples.
-  This is performed either by sub-sampling or by normalization of the matrix to the lowest sampling depth \cite{Weiss2015}.
-  Rarefaction is usually followed by filtering, which is performed to remove samples or features (\ac{otu}s or \ac{esv}s) from the count matrix that are sparse.
+  The module also supports rarefaction, which is a normalization technique used to overcome the bias that might arise due to variable sampling depth in different samples.
+  This is performed either by sub-sampling or by normalization of the matrix to the lowest sampling depth~\cite{Weiss2015}.
+  Although the pipeline supports rarefaction, the analyses reported in the paper do not rareify the counts matrices.
+  Filtering, is performed to remove samples or features (\ac{otu}s or \ac{esv}s) from the count matrix that are sparse.
   In order to determine the filtering threshold we fix the number of samples and correlation detection power needed and determine the number of features to be used.

### 2021-02-03 sections/results.tex
Update manuscript to address minor comments from Reviewer 1
Update manuscript to address comment R2.1.2

@@
 
-  % TODO: Talk about the difference between correlations and associations
   The six different network inference methods used in this study are \ac{magma}~\cite{Cougoul2019}, \ac{mldm}~\cite{Yang2017}, \ac{spieceasi}~\cite{Kurtz2015}, \ac{sparcc}~\cite{Friedman2012}, Spearman and Pearson.
@@
 
-  For the analysis presented in this section, we used the taxonomy composition table obtained using the \ac{ncbi} reference database as the input for algorithms that infer co-occurrence associations between the microbes.
+  For the analysis presented in this section, we used the taxonomy counts table obtained using the \ac{ncbi} reference database as the input for algorithms that infer co-occurrence associations between the microbes.
+  The taxonomy count tables are first collapsed to the Genus level (or the required taxonomy level) by summing up the counts of the lower taxonomy levels that map to the same higher taxonomy level name.
   Figure~\ref{fig:figure5}A shows the networks inferred from this dataset using the different inference algorithms.
@@
 This consensus network is built by merging the p-values generated from bootstraps of the original taxonomy counts table using the Browns p-value combining method~\cite{brown_400_1975} (see Methods section).
+  Note that we do not use Pearson and Spearman methods in the p-value merging step and these algorithms are only used for demonstration and comparison.
   Based on this approach, \ac{micone} reports as default output the consensus network, annotated with weights (correlations for \ac{sparcc} and direct associations for the other methods) for all four methods.

### 2021-02-10 sections/methods.tex
Update manuscript to address comment R2.1.2
Fix minor typos and sentence structure improvement

@@
   This module deals with the inference of co-occurrence associations from the \ac{otu} or \ac{esv} counts matrix.
-  These associations can be represented as a network, with nodes representing taxonomies of the micro-organisms and edges representing the association between them.
-  A null model is created by re-sampling and bootstrapping the correlation/interaction matrix and is used to calculate the significance of the inferred associations by calculating the p-values against this null model \cite{Watts2018}.
+  The counts matrices are collapsed to the Genus level (or the required taxonomy level) by adding up the counts of \ac{otu} or \ac{esv} that map to the same taxa.
+  These collapsed matrices are as used as input to the network inferrence methods to produce association matrices at the appropriate taxonomy level.
+  These associations can be represented as a network, with nodes representing taxonomies of the micro-organisms and edges representing the associations between them.
+  A null model is created by re-sampling and bootstrapping the correlation/interaction matrix and is used to calculate the significance of the inferred associations by calculating the p-values against this null model~\cite{Watts2018}.
   The pipeline includes Pearson, Spearman and FastSpar v0.0.10 (a faster implementation of \ac{sparcc})~\cite{Watts2018} as the pairwise correlation metrics, and \ac{spieceasi} v1.0.7~\cite{Kurtz2015}, \ac{mldm} v1.1~\cite{Yang2017} and \ac{magma}~\cite{Cougoul2019} as the direct association metrics.

### 2021-02-10 sections/results.tex
Fix minor typos and sentence structure improvement

@@
 
-  The representative sequences obtained from the \ac{dada2} method in DC step were used for taxonomic assignment using the three reference databases.
+  The representative sequences obtained from the \ac{dada2} method in the DC step were used for taxonomic assignment using the three reference databases.
   Figure~\ref{fig:figure4}A depicts a flow diagram that shows how the top 50 representative sequences (sorted by abundance) are assigned a Genus according to the three different databases.
+  % TOOD: Update with a more quantitative description
   We observe that not only does the assigned Genus composition vary significantly, but the percentage of unassigned representative sequences (gray) also differ.
@@
 
-  Given that no database performs better than others against mock datasets, and that databases are almost equally distant from each other in terms of final output, the choice of which database to use should be driven by other reason.
+  Given that no database performs better than others against mock datasets, and that databases are almost equally distant from each other in terms of final output, the choice of which database to use should be driven by other reasons.
   One user-specific way to choose, would be based on the known representation of taxa for the microbiome of interest (see also Discussion).

### 2021-02-10 sections/tables_figures.tex
Update manuscript to address comment R2.5
Fix minor typos and sentence structure improvement

@@
       \textbf{(B inset)} The network generated using DC=dada2, TA=GG, OP=no and NI=SPARCC and represents the particular point shown (big red square).
-      The plot clearly shows that the points can be separated based on the TA step and that the differences due to the DC and OP steps are not as significant.
+      The points on the PCA plot can be separated based on the TA step and the differences due to the DC and OP steps are comparatively not as significant.
     }

### 2021-04-09 sections/methods.tex
Add: Subsection about synthetic interaction data

@@
 
+  \subsection*{Synthetic interaction data}
+  \vspace{-5mm}
+
+  We generated synthetic interaction data using two methodologies previously used for benchmarking network inference methods.
+
+  The first method, hereby referred to as ``seqtime'', used generalized Lotka-Volterra (gLV) equations to model the microbial community dynamics and utilized the Klemm–Eguı́luz algorithm to generate a clique-based interaction network \cite{rottjersHairballsHypothesesBiological2018}.
+We used the \href{https://github.com/hallucigenia-sparsa/seqtime}{seqtime} R package to simulate communities with number of species (\(N\)) varying from 10 to 150 and number of samples (\(S\)) varying from 50 to 500.
+The abundance values of the species in the community at the end of the simulation time are used to create the OTU table.
+
+  The second method, hereby referred to as ``NorTA'', used the Normal to Anything (NorTA) approach coupled with a given interaction network topology to generate the abundance distribution of the microbial community \cite{kurtzSparseCompositionallyRobust2015}.
+We used the \href{https://github.com/zdk123/SpiecEasi}{spieceasi} R package to simulate communities with different network topologies (scale-free, cluster, block, Erdos-Renyi, band and hub) and target abundance distributions (Negative Binomial, Poisson, Zero-Inflated Negative Binomial).
+The OTU table was generated using the American Gut Project example in the \texttt{spieceasi} package (amgut1.filt) with the default parameter options.
+
+  For each method, we generated the OTU table depicting the abundances of species and used this to generate association matrices using MiCoNE.
+The interaction matrix was used as the source of expected (true) interactions and the associations predicted using MiCoNE were the source of predicted interactions.
+We evaluated the precision and sensitivity of the associations predicted by the individual network inference methods as well as the consensus (Figure XXX).
+
   \subsection*{Code and Data Availability}

### 2021-04-09 sections/methods.tex
Add: Subsection for the consensus method

@@
   The Brown's p-value merging method~\cite{brown_400_1975} is used for combining p-values from the various methods to obtain a consensus p-value, which is used to create the consensus network.
+
+  \subsection{Consensus method}
+  \vspace{-5mm}
+
+  The consensus algorithm was designed to increase the number of true positives reported at the end of the network inference step.
+For this purpose, we designed two simple algorithms that combine the edges reported by the different network inference tools for each pair of nodes using a consensus based on a parameter \(p\).
+The inputs to both the algorithms are the interaction networks (association matrices) generated by each method \(N_{i}\) and the threshold parameter \(p\).
+Here, the \(N_{i}\) each have the same set of nodes and only differ by the value of the association inferred between every pair.
+
+  \subsection{Algorithm 1: Simple voting}
+  \label{sec:org34a8cdc}
+
+  The simple voting method performs a voting-based consensus to determine whether an edge will exist between a given node-pair in the final consensus network.
+For each pair of nodes \((\text{node}_{i}, \text{node}_{j})\), we determine the number of network inference methods that report an edge between them (\(\text{edges}(\text{node}_{i}, \text{node}_{j})\)).
+Each node-pair will have an edge in the final consensus network if the number of reported edges is more than a threshold (Equation \ref{eqn:simple-voting})
+
+  \begin{equation}
+  \text{edges}(\text{node}_{i}, \text{node}_{j}) \geq \lfloor p * x \rfloor
+  \label{eqn:simple-voting}
+  \end{equation}
+  where, \(x\) is the number of network inference methods used in the consensus.
+
+  The simple voting method gives the union of the networks when \(p \in (0, \frac{1}{x})\) and will return the intersection when \(p \in ((x - 1), 1)\).
+In general, if \(p \in ((n - 1), \frac{n}{x})\), this algorithm will report an edge in the consensus network when at least \(n\) network inference methods report this edge.
+
+  \subsection{Algorithm 2: Scaled-sum method}
+  \label{sec:orgb0ddb1f}
+
+  This algorithm performs a consensus based on sum of all edges reported between a node-pair.
+First, the network generated by each network inference method is re-scaled so that \(\max(\mid N_{i} \mid) = 1\) (Equation \ref{eqn:scaled-sum-rescaling}).
+
+  \begin{equation}
+  N^{scaled}_{i} = \frac{N_{i}}{\max(\mid N_{i} \mid)}
+  \label{eqn:scaled-sum-rescaling}
+  \end{equation}
+
+  The re-scaling is performed because the associations reported by the direct association metrics are not necessarily correlations and are not directly comparable to the associations from the correlation-based methods due to their lower weights.
+This problem can also be alleviated by weighing each network inference method differently.
+Next, for each pair of nodes, we sum the weights of all reported edges (Equation \ref{eqn:scaled-sum})
+
+  \begin{equation}
+  \text{scaled\_sum}(\text{node}_{m}, \text{node}_{n}) = \sum_{i} N^{scaled}_{i}(\text{node}_{m}, \text{node}_{n})
+  \label{eqn:scaled-sum}
+  \end{equation}
+
+  An edge will exist between a given node-pair in the consensus network if \(\text{scaled\_sum} > (x - 1) * p\), where \(x\) is the number of network inference methods used in the consensus.
+The advantage of this method is that it also takes into account how the strength of the association reported for that particular node is in the inferred networks.
 

### 2021-09-15 sections/supplementary.tex
Update on Overleaf.

@@
   \input{sections/supplementary_table.tex}
+
+  \subsection{Configuration used for MiCoNE runs}
+
+  \subsection{Generation of sequence data}
+
+  \subsubsection{Mock sequence data}
+  Mock 4: A mock community composed of 21 bacterial strains represented in equal abundances in two replicate samples, and the same strains represented in uneven abundances in two replicate samples.
+These are the same community members present in mock-3 and mock-5, which consists of the same mock community samples analyzed on separate Illumina sequencing runs.
+This was generated by Dr.
+Dirk Gevers at Broad Institute in 2012.
+Previously called dataset-3 in [Bokulich et al.
+2013](https://dx.doi.org/10.1038/nmeth.2276) and dataset B4 in [Bokulich et al.
+2015](https://dx.doi.org/10.7287/peerj.preprints.934v2).
+
+  Mock 12: Composed of 27 bacterial strains amplified with 515f/806r primers.
+Intended to include more closely related taxa than the other mock communities, the members of which were chosen in part for their well-separated 16S rRNA gene sequences.
+These strains are all distinguishable over the sequenced region of the 16S rRNA gene, but some pairs of strains differ by as little as one nucleotide.
+Generated by Benjamin Callahan at Stanford University in 2015.
+Known as "Extreme" dataset in the original publication, doi:10.1038/nmeth.3869.
+
+  Mock 16: Mock community generated by Melanie Schirmer at University of Glasgow in 2015.
+Composed of even amounts of purified genomic DNA from 49 bacteria and 10 archaea.
+Known as dataset 35 in the original publication, doi:10.1093/nar/gku1341.
+
+  \subsubsection{Synthetic sequence data}
+  - We first create a sequence pool for various samples. This will have mutated (2\% mutation rate) reference sequences at abundances that are close to real abundances
+  - We then use `art_illumina` with 1x fold_coverage to generate fastq read files for each sample.
+
+    We created reference sequences that are 1/10th or 1/100th such that we have repeating sequences.
+Eg.
+If we have 3500 total reads, and 90\% are from same taxonomy, we have 32 or 315 reads of this sequence then we can fix `-f` as 3500/multiplier.
+To get multiplier we  could look at low abundant sequence and how much do we need to have atleast one of these sequences.
+OR should I just have fold coverage of 1x and 3500 references?
+Make 2\% of the positions in each sequence randomly mutate to create nucleotide diversity [source](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5967554/)
+
+  \subsection{Generation of synthetic interaction data}
+  Figures describing the original OTU tables or the interactions used to generate the community. Also include the seed and scripts for reproducing the networks.
 

### 2021-11-10 scripts/figure2/cca_script.py
Chore: Commit before updates

@@ -1,0 +1,110 @@
+#!/usr/bin/env python3
+
+import json
+import pathlib
+
+import numpy as np
+import pandas as pd
+import rcca
+from sklearn.cross_decomposition import CCA, PLSCanonical
+from sklearn.preprocessing import OneHotEncoder
+
+import statsmodels.formula.api as smf
+from statsmodels.stats.anova import anova_lm
+
+
+def read_correlations(corr_table_file: pathlib.Path) -> pd.DataFrame:
+    corr_table = pd.read_table(corr_table_file, index_col=0, sep="\t")
+    return corr_table
+
+
+def read_metadata(metadata_file: pathlib.Path) -> dict:
+    with open(metadata_file) as fid:
+        metadata = json.load(fid)
+    return metadata
+
+
+def create_matrices(corr_table: pd.DataFrame, metadata: dict):
+    # NOTE: X = nxq and Y = nxp
+    Y_full = corr_table.values.T
+    n = Y_full.shape[0]
+    # NOTE: We have 3 higher variables
+    # DC = 5 # TA = 3 # OP = 2
+    # Totally, 10 variables after one-hot-encoding
+    X_features = []
+    # Ensure order is of samples is correct
+    for ind in range(n):
+        X_features.append(metadata[str(ind + 1)])
+    enc = OneHotEncoder()
+    enc.fit(X_features)
+    X = enc.transform(X_features).toarray().astype(int)
+    Y_var = np.var(Y_full, axis=0)
+    max_var_inds = np.argsort(Y_var)[-1000:]
+    Y = Y_full[:, max_var_inds]
+    assert X.shape[0] == Y.shape[0]
+    return X, Y, enc
+
+
+# TODO: These functions should return variances instead of CCA objects
+def perform_sklearn_cca(X: np.ndarray, Y: np.ndarray, method: str = "CCA"):
+    assert X.shape[0] == Y.shape[0]
+    n, p, q = X.shape[0], X.shape[1], Y.shape[1]
+    n_components = min(n, p, q)
+    if method == "CCA":
+        cca = CCA(n_components=4, scale=False)
+    elif method == "PLSCanonical":
+        cca = PLSCanonical(n_components=4, scale=False)
+    else:
+        raise ValueError(f"Unsupport method: {method}")
+    cca.fit(X, Y)
+    return cca
+
+
+def perform_pyrcca_cca(X: np.ndarray, Y: np.ndarray):
+    assert X.shape[0] == Y.shape[0]
[…]

### 2021-11-10 scripts/figure2/pca_script.py

@@ -1,0 +1,94 @@
+#!/usr/bin/env python3
+
+import pathlib
+import json
+
+import numpy as np
+import pandas as pd
+from sklearn.decomposition import PCA, SparsePCA
+from statsmodels.formula.api import ols
+import statsmodels.api as sm
+
+
+def read_correlations(corr_table_file: pathlib.Path) -> pd.DataFrame:
+    corr_table = pd.read_table(corr_table_file, index_col=0, sep="\t")
+    corr_table.index.name = "links"
+    return corr_table
+
+
+def create_ydf(corr_table: pd.DataFrame) -> pd.DataFrame:
+    # NOTE: X = nxq and Y = nxp
+    Y_full = corr_table.T
+    # Y_var = np.var(Y_full, axis=0)
+    # max_var_inds = np.argsort(Y_var)[-1000:]
+    # Y = Y_full[:, max_var_inds]
+    return Y_full
+
+
+def read_metadata(metadata_file: pathlib.Path) -> dict:
+    with open(metadata_file) as fid:
+        metadata = json.load(fid)
+    return metadata
+
+
+def create_xdf(metadata: dict) -> pd.DataFrame:
+    # NOTE: X = nxq and Y = nxp
+    # NOTE: We have 3 factors
+    # DC = 5 # TA = 3 # OP = 2
+    # Totally, 10 variables after one-hot-encoding
+    X = pd.DataFrame(metadata, index=["DC", "TA", "OP"])
+    return X
+
+
+def perform_PCA(Y: pd.DataFrame) -> PCA:
+    pca = PCA()
+    pca.fit(Y)
+    return pca
+
+
+def perform_anova(X: pd.DataFrame, Y: pd.DataFrame) -> dict:
+    anova_dict = dict()
+    for component in Y.columns:
+        data = [
+            {
+                "Y": Y.loc[i, component],
+                "DC": X.loc["DC", i],
+                "TA": X.loc["TA", i],
+                "OP": X.loc["OP", i],
+            }
+            for i in Y.index
+        ]
+        df = pd.DataFrame(data)
+        lm = ols("Y ~ C(DC) + C(TA) + C(OP)", data=df).fit()
+        anova_dict[component] = sm.stats.anova_lm(lm, typ=2)
+    return anova_dict
+
+
+def normalize_anova(anova_dict: dict, pca: PCA) -> list:
+    var_ratio = pca.explained_variance_ratio_
+    assert sum(var_ratio) == 1
+    variance_list = []
+    for component in anova_dict:
[…]

### 2021-11-23 scripts/figure4/figure4a_plot.R
Change: Update figure 4 scripts to support new pipeline

@@ -17,9 +17,9 @@ read_data <- function(data_file, n) {
     raw_df
 }
 
-gg <- read_data("fmt_gg.csv", notu)
-silva <- read_data("fmt_silva.csv", notu)
-ncbi <- read_data("fmt_ncbi.csv", notu)
+gg <- read_data("../../data/figure4/output/moving_pictures/gg.csv", notu)
+silva <- read_data("../../data/figure4/output/moving_pictures/silva.csv", notu)
+ncbi <- read_data("../../data/figure4/output/moving_pictures/ncbi.csv", notu)
 
 make_alluvial_plot <- function(db_data, title) {
     genus_factors <- levels(factor(db_data$Genus))

### 2021-11-23 scripts/figure4/figure4ab_data.py

@@ -8,7 +8,7 @@ import click
 
 def write_tables(file_paths, notus, output_path):
     for file in file_paths:
-        db_name = file.stem
+        db_name = file.parent.parent.parent.stem
         otu = load_table(str(file))
         otu_df = otu.to_dataframe(dense=True)
         obs_metadata = otu.metadata_to_dataframe(axis="observation")
@@ -22,7 +22,9 @@ def write_tables(file_paths, notus, output_path):
         final_df.index = range(1, notus + 1)
         final_df.index.name = "OTU"
         final_df["empty"] = [""] * notus
-        final_df.to_csv(output_path / f"{db_name}.csv", sep=",", index=True)
+        output_file = output_path / f"{db_name}.csv"
+        print(f"Writing file {output_file}")
+        final_df.to_csv(output_file, sep=",", index=True)
 
 
 @click.command()

### 2021-11-23 scripts/figure4/figure4b_plot.R

@@ -27,7 +27,9 @@ combine_data <- function(db1, db2) {
     final_tbl <- comb_tbl %>%
         gather(levels, key="tax_level", value="value") %>%
         select_at(.vars=c("tax_level", "value", "type"))
-    final_tbl$value = final_tbl$value / nrow(db1) * 100
+    # final_tbl$value = final_tbl$value / nrow(db1) * 100
+    # if we want to display the numbers instead
+    final_tbl$value = final_tbl$value
     return(final_tbl)
 }
 
@@ -51,9 +53,9 @@ make_bar_plot <- function(data, title) {
 }
 
 # Importing data
-gg <- tidy_up_data("fmt_gg.csv", notu)
-silva <- tidy_up_data("fmt_silva.csv", notu)
-ncbi <- tidy_up_data("fmt_ncbi.csv", notu)
+gg <- tidy_up_data("../../data/figure4/output/moving_pictures/gg.csv", notu)
+silva <- tidy_up_data("../../data/figure4/output/moving_pictures/silva.csv", notu)
+ncbi <- tidy_up_data("../../data/figure4/output/moving_pictures/ncbi.csv", notu)
 
 # Combinations
 gg_silva <- combine_data(gg, silva)

### 2021-11-23 scripts/figure4/make_input_taxonomy.py

@@ -4,7 +4,7 @@ import click
 from ete3 import NCBITaxa
 import pandas as pd
 
-from mindpipe import Lineage
+from micone import Lineage
 
 NCBI = NCBITaxa()
 TAX_LEVELS = Lineage._fields

### 2021-12-11 sections/methods.tex

@@
 
-  \subsection{Consensus method}
+  \subsection*{Consensus Network and p-value merging}
+  \vspace{-5mm}
+
+  \subsubsection*{Bootstrapping}
+  For each network inference method (both correlation and direct association based methods), $1000$ permutations of the original \ac{otu} counts data were generated~\cite{Watts2018}.
+  We then recalculate the associations in these permuted \ac{otu} tables using the different network inference algorithms.
+  Finally, we calculate the p-value based on how often a more extreme association is observed for randomly permuted data.
+
+  \subsubsection*{p-value merging (can be moved to the supplementary section)}
+  Fisher~\cite{fisher_224a_1948} proposed that for $k$ independent p-values, each generated by $k$ different methods and denoted by $P_i$, the statistic $\Psi$:
+  \begin{equation*}
+    \begin{aligned}
+        \Psi &= \sum_{i=1}^k -2 \log \left( P_i \right) \\
+        \Psi &\sim \chi^2_{2k}
+    \end{aligned}
+  \end{equation*}
+
+  Brown~\cite{brown_400_1975} extended Fisher's method to dependent p-values by using a re-scaled $\chi^2$ distribution:
+  \begin{equation*}
+    \Psi \sim c \chi^2_{2f}
+  \end{equation*}
+  where, $f$ is the degrees of freedom and $c$ is the scale factor and are given by:
+  \begin{equation*}
+    f = \frac{\mathrm{E}[\Psi]^2}{\mathrm{Var}[\Psi]} ~~~\text{and}~~~ c = \frac{\mathrm{Var}[\Psi]}{2\mathrm{E}[\Psi]} = \frac{k}{f}
+  \end{equation*}
+
+  Furthermore, Brown showed that $\mathrm{E}[\Psi]$ and $\mathrm{Var}[\Psi]$ can be calculated directly via a numerical integration:
+  \begin{equation*}
+    \mathrm{E}[\Psi] = 2k ~~~\text{and}~~~ \mathrm{Var}[\Psi] = 4k + 2\sum_{i<j} \mathrm{Cov}\left( -2\log(P_i), -2\log(P_j) \right)
+  \end{equation*}
+
+  Kost and McDermott~\cite{kost_combining_2002} further fit a third-order polynomial to approximate the covariance
+  \begin{equation}
+    \mathrm{Cov}\left( -2\log(P_i), -2\log(P_j) \right) \approx 3.263 \rho_{ij} + 0.710 \rho_{ij}^2 + 0.027 \rho_{ij}^3
+    \label{eqn:covariance-pvalues}
+  \end{equation}
+  where, $\rho_{ij}$ is the correlation between method $i$ and method $j$
+
+  The final combined p-value~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016} is then given by:
+  \begin{equation}
+    \begin{aligned}
+        & P_{combined} = 1.0 - \Phi_{2f}\left( \psi / c \right) \\
+        \text{where},~ &\psi = -2 \sum_{i=1}^k \log(P_i) ~~~\text{and}~~~ \Phi_{2f} = \mathrm{CDF}\left( \chi^2_{2f} \right)
+    \end{aligned}
+    \label{eqn:pvalue-combined}
+  \end{equation}
+
+  The consensus method in \ac{micone} (refer Documentation) uses Equation~\ref{eqn:covariance-pvalues} to estimate the covariance of the pvalues and Equation~\ref{eqn:pvalue-combined} to merge the p-values (obtained from bootstrapping) from different methods.
+  Note that we do not use Pearson and Spearman methods in the p-value merging step and these algorithms are only used for demonstration and comparison.
+  The combined p-values are used to threshold for significance during the consensus network step.
+
+ \subsection*{NOTATION}
+
+Note this section defines the notation for the network comparison component of our pipeline, which assumes that all networks to be compared have the same number of nodes
+
+q = number of nodes
+
+w = number of different networks from different approaches (previously x?)
+
+$N^i$ = network k (currently Ni)
+
+$N^i_{a,b}$ = edge (a,b) in network i
+
+It useful to define a "stretched" version of each adjacency matrix $N^i$ into a column vector, where all columns are stacked onto each other into an $q^2$ long vector:
+
+$\bar{N}^i_{j}$ , $j=1:q^2$
+
+  \subsection*{Consensus method}
   \vspace{-5mm}
@@
 For this purpose, we designed two simple algorithms that combine the edges reported by the different network inference tools for each pair of nodes using a consensus based on a parameter \(p\).
-The inputs to both the algorithms are the interaction networks (association matrices) generated by each method \(N_{i}\) and the threshold parameter \(p\).
+The inputs to both the algorithms are the interaction networks (association matrices) generated by each method \(N_{i}\) and the threshold parameter (p).
 Here, the \(N_{i}\) each have the same set of nodes and only differ by the value of the association inferred between every pair.
 
-  \subsection{Algorithm 1: Simple voting}
-  \label{sec:org34a8cdc}
+  \subsection*{Algorithm 1: Simple voting}
 
@@
 
+ For each edge (i,j), we count number $M_{i,j}$ of networks in which that edge exists, i.e.:
+
  \begin{equation}
-  \text{edges}(\text{node}_{i}, \text{node}_{j}) \geq \lfloor p * x \rfloor
+ \begin{aligned}
+ M_{i,j} &= | \left{ a | N^{(a)}_{i,j}>0 \right} | \\
+ M_{i,j} &\geq \lfloor \theta * w \rfloor
+ \end{aligned}
   \label{eqn:simple-voting}
   \end{equation}
-  where, \(x\) is the number of network inference methods used in the consensus.
+
+  \begin{equation}
+  \text{edges}(\text{node}_{i}, \text{node}_{j}) \geq \lfloor \theta * w \rfloor
+  \label{eqn:simple-voting}
+  \end{equation}
+
+   \begin{equation}
+   \sum_k N^k_{i,j} \geq \lfloor \theta * w \rfloor
+  \label{eqn:simple-voting2}
+  \end{equation}
+
+  where, \(w\) is the number of network inference methods used in the consensus.
+
+ where, \theta is the user-defined threshold....
 
@@
 
-  \subsection{Algorithm 2: Scaled-sum method}
-  \label{sec:orgb0ddb1f}
+  \subsubsection*{Algorithm 2: Scaled-sum method}
 
   This algorithm performs a consensus based on sum of all edges reported between a node-pair.
-First, the network generated by each network inference method is re-scaled so that \(\max(\mid N_{i} \mid) = 1\) (Equation \ref{eqn:scaled-sum-rescaling}).
+First, the network generated by each network inference method is re-scaled so that (\max(\mid N_{i} \mid) = 1) (Equation \ref{eqn:scaled-sum-rescaling}).
+
+\begin{equation}
+  NS^{a}_{i,j} = \frac{N^{(a)}_{i,j}}{\max(\mid N_{i} \mid)}
+  \label{eqn:scaled-sum-rescaling}
+  \end{equation}
 
@@
 This problem can also be alleviated by weighing each network inference method differently.
-Next, for each pair of nodes, we sum the weights of all reported edges (Equation \ref{eqn:scaled-sum})
+Next, for each pair of nodes, we sum the weights of all reported edges (Equation \ref{eqn:scaled-sum}).
 
@@
 
-  \subsection*{Network Variability}
+  \subsubsection*{Network Variability}
   \vspace{-5mm}
@@
 
-  \subsection*{Consensus Network and p-value merging}
-  \vspace{-5mm}
-
-  \subsubsection*{Bootstrapping}
-  For each network inference method (both correlation and direct association based methods), $1000$ permutations of the original \ac{otu} counts data were generated~\cite{Watts2018}.
-  We then recalculate the associations in these permuted \ac{otu} tables using the different network inference algorithms.
-  Finally, we calculate the p-value based on how often a more extreme association is observed for randomly permuted data.
-
-  \subsubsection*{p-value merging}
-  Fisher~\cite{fisher_224a_1948} proposed that for $k$ independent p-values, each generated by $k$ different methods and denoted by $P_i$, the statistic $\Psi$:
-  \begin{equation*}
-    \begin{aligned}
-        \Psi &= \sum_{i=1}^k -2 \log \left( P_i \right) \\
-        \Psi &\sim \chi^2_{2k}
-    \end{aligned}
-  \end{equation*}
-
-  Brown~\cite{brown_400_1975} extended Fisher's method to dependent p-values by using a re-scaled $\chi^2$ distribution:
-  \begin{equation*}
-    \Psi \sim c \chi^2_{2f}
-  \end{equation*}
-  where, $f$ is the degrees of freedom and $c$ is the scale factor and are given by:
-  \begin{equation*}
-    f = \frac{\mathrm{E}[\Psi]^2}{\mathrm{Var}[\Psi]} ~~~\text{and}~~~ c = \frac{\mathrm{Var}[\Psi]}{2\mathrm{E}[\Psi]} = \frac{k}{f}
-  \end{equation*}
-
-  Furthermore, Brown showed that $\mathrm{E}[\Psi]$ and $\mathrm{Var}[\Psi]$ can be calculated directly via a numerical integration:
-  \begin{equation*}
-    \mathrm{E}[\Psi] = 2k ~~~\text{and}~~~ \mathrm{Var}[\Psi] = 4k + 2\sum_{i<j} \mathrm{Cov}\left( -2\log(P_i), -2\log(P_j) \right)
-  \end{equation*}
-
-  Kost and McDermott~\cite{kost_combining_2002} further fit a third-order polynomial to approximate the covariance
-  \begin{equation}
-    \mathrm{Cov}\left( -2\log(P_i), -2\log(P_j) \right) \approx 3.263 \rho_{ij} + 0.710 \rho_{ij}^2 + 0.027 \rho_{ij}^3
-    \label{eqn:covariance-pvalues}
-  \end{equation}
-  where, $\rho_{ij}$ is the correlation between method $i$ and method $j$
-
-  The final combined p-value~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016} is then given by:
-  \begin{equation}
-    \begin{aligned}
-        & P_{combined} = 1.0 - \Phi_{2f}\left( \psi / c \right) \\
-        \text{where},~ &\psi = -2 \sum_{i=1}^k \log(P_i) ~~~\text{and}~~~ \Phi_{2f} = \mathrm{CDF}\left( \chi^2_{2f} \right)
-    \end{aligned}
-    \label{eqn:pvalue-combined}
-  \end{equation}
-
-  The consensus method in \ac{micone} (refer Documentation) uses Equation~\ref{eqn:covariance-pvalues} to estimate the covariance of the pvalues and Equation~\ref{eqn:pvalue-combined} to merge the p-values (obtained from bootstrapping) from different methods.
-  Note that we do not use Pearson and Spearman methods in the p-value merging step and these algorithms are only used for demonstration and comparison.
-  The combined p-values are used to threshold for significance during the consensus network step.
+  The possible correlations between the edges of the network lead to inaccurate estimates of the variance if a GLM were used to directly model the relationships between the links and steps in the workflow.
+Therefore, we transformed the network matrix $N$ using a Principal Component Analysis into the $N_{reduced}$ matrix.
+The dimensions of the $N_{reduced}$ matrix are be denoted by $(d \times n)$ where d is the number of principal components and n is the number of network in the $N$ matrix.
 

### 2021-12-13 sections/methods.tex
Change: Update datasets subsection in methods section

@@
 
-  \subsection*{Datasets}
+  \subsection*{16S rRNA sequencing Datasets}
 
   \vspace{-5mm}
-  The study uses three kinds of 16S rRNA sequencing datasets: real datasets, mock datasets and synthetic datasets.
-  Real datasets are collections of sequencing reads obtained from naturally occurring microbial community samples.
-  The current study used healthy stool samples from a fecal microbiome transplant study~\cite{Kang2017} and healthy saliva samples from a periodontal disease study~\cite{Chen2018} as real datasets for analysis.
+    The study uses two kinds of 16S rRNA sequencing datasets: biological datasets and mock datasets.
+    Biological datasets are collections of sequencing reads obtained from naturally occurring microbial community samples.
+    The current analysis used stool samples from a fecal microbiome transplant study of autism~\cite{Kang2017} as biological datasets.
+    The study was composed of multiple sequencing runs and those runs that contained paired end reads (runs 2, 3 and 4), were downloaded from Qiita~\cite{qiita} from https://qiita.ucsd.edu/study/description/10532 and used as input sequences for the \ac{micone} pipeline.
     % TODO: How many samples are there for each and average number of reads.
-  The mock community 16S datasets are real sequencing data obtained for artificially assembled collections of species in known proportions.
+    Sequences from both control (x samples containing xx reads) and autistic (y samples containing yy reads) patients were included in the analyses but the groups were analyzed independently in the pipeline.
+    The mock community 16S datasets are sequencing data obtained for artificially assembled collections of species in known proportions.
     The mock datasets used for this study, obtained from mockrobiota~\cite{Bokulich2016}, are labelled mock4, mock12 and mock16.
@@
     The mock16 community is composed of 49 bacteria and 10 archea, all represented in equal amount.
-  The synthetic datasets were generated using an artificial read simulator called ART~\cite{Huang2012}.
-  Three different microbial composition profiles were used as input; reads were generated using a soil and water microbiome composition profiles from the \ac{emp}~\cite{Thompson2017} and healthy gut microbiome project from the fecal microbiome transplant study~\cite{Kang2017}.
-  % TODO: Create a section in the supplementary that describes what the original profiles were and how we used ART (what settings) to generate the synthetic data.
-  The reads are simulated using the NCBI RefSeq database as the reference sequence pool and the `art\_illumina' sequence profile with a mutation rate of 2\%.
-  The scripts used to generate the synthetic data are in the scripts folder of the repository (\href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper}).
 
@@
 
-  The first method, hereby referred to as ``seqtime'', used generalized Lotka-Volterra (gLV) equations to model the microbial community dynamics and utilized the Klemm–Eguı́luz algorithm to generate a clique-based interaction network \cite{rottjersHairballsHypothesesBiological2018}.
-We used the \href{https://github.com/hallucigenia-sparsa/seqtime}{seqtime} R package to simulate communities with number of species (\(N\)) varying from 10 to 150 and number of samples (\(S\)) varying from 50 to 500.
+  The first method, hereby referred to as ``seqtime'', used generalized Lotka-Volterra (gLV) equations to model the microbial community dynamics and utilized the Klemm–Eguı́luz algorithm to generate a clique-based interaction network~\cite{rottjersHairballsHypothesesBiological2018}.
+  We used the \href{https://github.com/hallucigenia-sparsa/seqtime}{seqtime} R package to simulate communities with number of species ($N$) varying from 10 to 150 (10, 25, 50, 100, 150 and 200).
+  The initial species concentrations are randomly sampled from a Poisson distribution and the simulation is rerun to generate a number of samples ($S$) varying from 50 to 500 (50, 100, 200, 500) for different communities.
   The abundance values of the species in the community at the end of the simulation time are used to create the OTU table.
@@
 
-  For each method, we generated the OTU table depicting the abundances of species and used this to generate association matrices using MiCoNE.
-The interaction matrix was used as the source of expected (true) interactions and the associations predicted using MiCoNE were the source of predicted interactions.
-We evaluated the precision and sensitivity of the associations predicted by the individual network inference methods as well as the consensus (Figure XXX).
+  For each method, we generated the OTU table depicting the abundances of species and used this to generate association networks using \ac{micone}.
+  The interaction matrix was used as the source of expected (true) interactions and the associations predicted using \ac{micone} were the source of predicted interactions.
+  Finally, for each dataset we evaluated the precision and sensitivity of the associations predicted by the individual network inference methods as well as the consensus (Figure \ref{fig:figure7}).
 

### 2021-12-13 sections/methods.tex
Add: Sequence processing subsection and update DC subsection

@@
   All such individual modules are available as part of the \ac{micone} package.
-  The inputs to the pipeline by default are the raw community 16S rRNA sequence reads, but the software can be alternatively configured to use trimmed sequences, \ac{otu} tables and other types of intermediate data.
+  The inputs to the pipeline by default are raw untrimmed 16S rRNA sequence reads, but the software can be alternatively configured to use trimmed sequences, \ac{otu} tables and other types of intermediate data.
+  The pipeline supports both paired end and single end reads, and additionally supports independently processing multiple runs and merging the OTU tables in the DC step.
   The final output of the pipeline is the inferred network of co-occurrence relationships among the microbes present in the samples.
 
-  The \ac{micone} pipeline provides both a Python API as well as a command-line interface and only requires a single configuration file.
-  The configuration file lists the inputs, output and the steps to be performed during runtime, along with the parameters to be used (if different from defaults) for the various steps.
+  The \ac{micone} pipeline provides both a Python API as well as a command-line interface and only uses a single configuration file (nextflow.config) to contain the configuration parameters.
+  The configuration file along with the run file (main.nf) lists the inputs, output and the steps to be performed during runtime, along with the parameters to be used (if different from defaults) for the various steps.
   Since the entire pipeline run-through is stored in the form of a text file (the configuration file), subsequent runs are highly reproducible and changes can be easily tracked using version control.
@@
 
+  \subsection*{Sequence Processing (SP)}
+  This module deals with processing the raw multiplexed 16S sequence data into demultiplexed, quality-controlled, trimmed sequences.
+  It consists of the following processes: demulitplexing and trimming (includes quality control and filtering).
+  The trimming process handles the quality control steps such as trimming adapters and trimming low-quality nucleotide stretches from the sequences.
+  The parameters and tools in this process are fixed and are not available for user customization.
+  The various tools used for the processes were adapted from \ac{qiime2} v2021.8.0~\cite{bolyenReproducibleInteractiveScalable2019}.
+
   \subsection*{Denoising and Clustering (DC)}
   \vspace{-5mm}
-  This module deals with processing the raw 16S sequence data into \ac{otu} or \ac{esv} count tables.
-  It consists of the following processes: quality control, denoising (or clustering) and chimera checking.
-  The quality control process handles the demultiplexing and quality control steps such as trimming adapters and trimming low-quality nucleotide stretches from the sequences.
+  This module deals with processing the quality-controlled, trimmed 16S sequence data into \ac{otu} or \ac{esv} count tables.
+  It consists of the following processes: denoising (or clustering) and chimera checking.
   The denoise/cluster process handles the conversion of the demultiplexed, trimmed sequences into \ac{otu} or \ac{esv} count tables (some methods, like closed reference and open reference clustering, perform clustering and taxonomy assignment in the same step).
@@
   The output of this module is a matrix of counts, that describes the number of reads of a particular \ac{otu} or \ac{esv} (rows of the matrix) present in each sample (columns of the matrix).
-  The options currently available in the pipeline for denoising and clustering are: open reference clustering, closed reference clustering and de novo clustering methods from \ac{qiime1} v1.9.1~\cite{Caporaso2010} and denoising methods from \ac{dada2} v1.14~\cite{Callahan2016} and Deblur v1.1.0~\cite{Amir2017}.
-  The quality filtering and chimera checking tools are derived from those used in \ac{qiime2} v2019.10.0 and \ac{dada2}.
+  The options currently available in the pipeline for denoising and clustering are: open reference clustering, closed reference clustering and de novo clustering methods from the vsearch plugin of \ac{qiime2} v2021.8.0~\cite{bolyenReproducibleInteractiveScalable2019} and denoising methods from \ac{dada2} v1.14~\cite{Callahan2016} and Deblur v1.1.0~\cite{Amir2017} (from the deblur plugin of \ac{qiime2}).
+  The quality filtering and chimera checking tools are derived from those used in \ac{qiime2} v2021.8.0 (uchime method) and \ac{dada2} (remove bimera method).
 
@@
  \begin{aligned}
- M_{i,j} &= | \left{ a | N^{(a)}_{i,j}>0 \right} | \\
+ M_{i,j} &= | \left{ a \mid N^{a}_{i,j}>0 \right} | \\
  M_{i,j} &\geq \lfloor \theta * w \rfloor

### 2021-12-15 sections/methods.tex
Add: DC and TA step tables

@@
   This module deals with processing the raw multiplexed 16S sequence data into demultiplexed, quality-controlled, trimmed sequences.
-  It consists of the following processes: demulitplexing and trimming (includes quality control and filtering).
+  It consists of the following processes: demultiplexing and trimming (includes quality control and filtering).
   The trimming process handles the quality control steps such as trimming adapters and trimming low-quality nucleotide stretches from the sequences.
@@
 
+  \begin{table}[h]
+    \centering
+    \small
+    \begin{tabular}{|c|c|c|}
+      \hline
+      \textbf{Module} & \textbf{Tool} & \textbf{References} \\
+      \hline
+      Denoising and Clustering & Closed reference & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
+      Denoising and Clustering & Open reference & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
+      Denoising and Clustering & De novo & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
+      Denoising and Clustering & Dada2 & \cite{Callahan2016} \\
+      Denoising and Clustering & Deblur & \cite{Amir2017,bolyenReproducibleInteractiveScalable2019} \\
+      \hline
+      Chimera checking & Uchime & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
+      Chimera checking & Remove bimera & \cite{Callahan2016} \\
+    \end{tabular}
+    \caption{Tools used in the DC module}
+    \label{tab:dc_tools}
+  \end{table}
+
   \subsection*{Taxonomy Assignment (TA)}
@@
   Finally, a consensus method is used to identify the most probable match from the list of best matches.
-  The pipeline incorporates \ac{gg} 13\_8~\cite{DeSantis2006}, SILVA 132~\cite{Quast2012} and the \ac{ncbi} (16S RefSeq as of Oct 2019)~\cite{Sayers2009} databases for taxonomy assignment and the Naive Bayes classifier from \ac{qiime2} and \ac{ncbi} blast as the query tools (from \ac{qiime2}).
+  The pipeline incorporates \ac{gg} 13\_8~\cite{DeSantis2006}, SILVA 138~\cite{Quast2012} and the \ac{ncbi} (16S RefSeq as of Oct 2021)~\cite{Sayers2009} databases for taxonomy assignment.
+  These databases were downloaded and built using the RESCRIPt qiime2 plugin~\cite{iiRESCRIPtReproducibleSequence2021}.
+  The Naive Bayes classifier from \ac{qiime2} and \ac{ncbi} blast as the query tools (from \ac{qiime2}).
   The consensus algorithm used is the default method used by the classifiers in \ac{qiime2}.
+
+  \begin{table}[h]
+    \centering
+    \small
+    \begin{tabular}{|c|c|c|}
+      \hline
+      \textbf{Module} & \textbf{Tool/Database} & \textbf{References} \\
+      \hline
+      Query tool & Blast & \cite{camachoBLASTArchitectureApplications2009} \\
+      Query tool & Naive bayes classifier & \cite{bokulichOptimizingTaxonomicClassification2018} \\
+      \hline
+      Database & Greengenes & \cite{DeSantis2006} \\
+      Database & SILVA & \cite{Quast2012} \\
+      Database & NCBI RefSeq & \cite{Sayers2009} \\
+    \end{tabular}
+    \caption{Tools used in the TA module}
+    \label{tab:ta_tools}
+  \end{table}
 

### 2022-01-12 sections/results.tex
Add: TODOs to results section

@@
 
+  % TODO: This is the most important statement (look at the dada2 paper for better arguments)
+  We prefer dada2 because it peforms marginally better (value??) than the other methods and it is also performs independent of the taxonomy database (so does denovo why are you suggesting this then?).
+  Denovo also performs better but the denoising helps when there are more sequences (where is the proof?) [ref]
+
   \FloatBarrier
@@
  In addition to being regularly maintained and updated the \ac{ncbi} database already has the advantage that its accuracy of assignments is still comparable to the SILVA and \ac{gg} reference databases that are routinely used as reference databases.
+
+ % TODO: This is the most important statement here
+ From previous studies, (refer to the Bokulich2018 paper here) which have done more quantitative benchmarks on the subject of which tool is the best (for classifer)...
+ There are also other studies, [ref] that focus on the database aspect and they too have found that...
 

### 2022-01-14 sections/acronyms.tex
Do not print acronyms at the end
Fix: Greengenes acronym
Add: Acronym for MAGMA
Add: Pipeline steps to acronym list
Updates from Overleaf
Updates to draft
Add: OP, NI steps and consensus network to methods section

@@
 %!TEX root = ../main.tex
-
-\section*{Acronyms}
-
-\vspace{1em}
 
@@
     \acro{mind}[MIND]{Microbial Interaction Network Database}
+    \acro{micone}[MiCoNE]{Microbial Co-occurrence Network Explorer}
     \acro{otu}[OTU]{Operational Taxanomic Unit}
@@
     \acro{pcr}[PCR]{Polymerase Chain Reaction}
-    \acro{gg}[GG]{Green Genes}
+    \acro{emp}[EMP]{Earth Microbiome Project}
+    \acro{gg}[GG]{Greengenes}
     \acro{ncbi}[NCBI]{National Center for Biotechnology Information}
@@
     \acro{mldm}[mLDM]{metagenomic Lognormal-Dirichlet-Multinomial}
-    \acro{daa}[DAA]{Direct Association Analysis}
+    \acro{cozine}[COZINE]{COmpositional Zero-Inflated Network Estimation}
+    \acro{spring}[SPRING]{Semi-Parametric Rank-based approach for INference in Graphical model}
+    \acro{harmonies}[HARMONIES]{A Hybrid Approach for Microbiome Networks Inference via Exploiting Sparsity}
     \acro{kegg}[KEGG]{Kyoto Encyclopedia of Genes and Genomes}
     \acro{fba}[FBA]{Flux Balance Analysis}
+    \acro{dc}[DC]{Denoising and Clustering}
+    \acro{ta}[TA]{Taxonomy Assignment}
+    \acro{op}[OP]{OTU Processing}
+    \acro{ni}[NI]{Network Inference}
 \end{acronym}

### 2022-01-14 sections/methods.tex
Add: TODOs to results section
Fix: Add absolute function for scaled_sum formula
Add: OP, NI steps and consensus network to methods section

@@
     The mock12 community is composed of 27 bacterial strains that include closely related taxa with some pairs having only one to two nucleotide difference from another.
-    The mock16 community is composed of 49 bacteria and 10 archea, all represented in equal amount.
+    The mock16 community is composed of 49 bacteria and 10 Archea, all represented in equal amount.
 
@@
       Chimera checking & Remove bimera & \cite{Callahan2016} \\
+      \hline
     \end{tabular}
@@
   These databases were downloaded and built using the RESCRIPt qiime2 plugin~\cite{iiRESCRIPtReproducibleSequence2021}.
-  The Naive Bayes classifier from \ac{qiime2} and \ac{ncbi} blast as the query tools (from \ac{qiime2}).
+  The Naive Bayes classifier and the \ac{ncbi} blast used as the query tools in this study were from the \ac{qiime2} package and the parameters used were also the defaults in the package.
   The consensus algorithm used is the default method used by the classifiers in \ac{qiime2}.
@@
       \hline
-      Query tool & Blast & \cite{camachoBLASTArchitectureApplications2009} \\
+      Query tool & Blast & \cite{camachoBLASTArchitectureApplications2009,bokulichOptimizingTaxonomicClassification2018} \\
       Query tool & Naive bayes classifier & \cite{bokulichOptimizingTaxonomicClassification2018} \\
@@
       Database & NCBI RefSeq & \cite{Sayers2009} \\
+      \hline
     \end{tabular}
@@
 
-  % TODO: Add references and basic equations or details
   \subsection*{OTU and ESV Processing (OP)}
   \vspace{-5mm}
-  This module deals with normalization, filtering and applying transformations to the \ac{otu} or \ac{esv} counts matrix.
+  This module deals with normalization, filtering, forking, grouping and applying transformations to the \ac{otu} or \ac{esv} counts matrix.
+  Normalization of the count matrix involves converting the count matrix of read counts into a count matrix containing relative abundances.
   The module also supports rarefaction, which is a normalization technique used to overcome the bias that might arise due to variable sampling depth in different samples.
   This is performed either by sub-sampling or by normalization of the matrix to the lowest sampling depth~\cite{Weiss2015}.
-  Although the pipeline supports rarefaction, the analyses reported in the paper do not rareify the counts matrices.
+  Although the pipeline supports rarefaction, these modules are turned off by default and the analyses reported in the paper do not normalize or rareify the counts matrices.
+  However, most of the network inference methods perform normalization and other transformation operations on the counts matrix as a part of their workflow.
   Filtering, is performed to remove samples or features (\ac{otu}s or \ac{esv}s) from the count matrix that are sparse.
-  In order to determine the filtering threshold we fix the number of samples and correlation detection power needed and determine the number of features to be used.
-  Finally, transformations are performed in order to correct for and overcome the compositional bias that is inherent in a counts matrix (in most cases this is handled by the network inference algorithm).
+  By default, samples are filtered out if the total reads in a sample are less than 500 and features are filtered out if the abundance of the 1\% and prevalence is less than 5\%.
+  The forking operation splits the count matrix into multiple matrices based on sample metadata column.
+  The group operation transforms the \ac{otu} or \ac{esv} count matrix into a taxonomic count matrix at the requested level by adding up counts that map to the same taxonomy.
+  Finally, transformations are performed in order to correct for and overcome the compositional bias that is inherent in a counts matrix (in analyses peformed in the study these are handled by the network inference algorithm).
+  All the modules in this step were implemented using functions from the \href{http://biom-format.org/}{biom-format} Python package~\cite{mcdonaldBiologicalObservationMatrix2012}.
 
@@
   \vspace{-5mm}
-  This module deals with the inference of co-occurrence associations from the \ac{otu} or \ac{esv} counts matrix.
-  The counts matrices are collapsed to the Genus level (or the required taxonomy level) by adding up the counts of \ac{otu} or \ac{esv} that map to the same taxa.
-  These collapsed matrices are as used as input to the network inferrence methods to produce association matrices at the appropriate taxonomy level.
+  This module deals with the inference of co-occurrence associations from the processed \ac{otu} or \ac{esv} counts matrix.
+  The input count matrices are collapsed to the Genus level (or any other required taxonomy level) using the group module at the OP step.
+  These collapsed matrices are as used as input to the network inference methods to produce association matrices at the appropriate taxonomy level.
   These associations can be represented as a network, with nodes representing taxonomies of the micro-organisms and edges representing the associations between them.
-  A null model is created by re-sampling and bootstrapping the correlation/interaction matrix and is used to calculate the significance of the inferred associations by calculating the p-values against this null model~\cite{Watts2018}.
-  The pipeline includes Pearson, Spearman and FastSpar v0.0.10 (a faster implementation of \ac{sparcc})~\cite{Watts2018} as the pairwise correlation metrics, and \ac{spieceasi} v1.0.7~\cite{Kurtz2015}, \ac{mldm} v1.1~\cite{Yang2017} and \ac{magma}~\cite{Cougoul2019} as the direct association metrics.
-  The Brown's p-value merging method~\cite{brown_400_1975} is used for combining p-values from the various methods to obtain a consensus p-value, which is used to create the consensus network.
+  A null model is created by re-sampling and bootstrapping the counts matrix and recalculating the correlations/associations.
+  These bootstrapped association matrices are used to calculate the significance of the inferred associations by calculating the p-values against this null model~\cite{Watts2018}.
+  The pipeline includes 4 methods for pairwise correlation metrics, and 6 methods for direct association metrics (refer Table~\ref{tab:ni_tools}).
+  The Brown's p-value merging method~\cite{brown_400_1975} is used for combining p-values from the pairwise correlations methods to obtain a consensus p-value, which can be used to filter for significance.
+  In the final module of this step, the consensus algorithms are used to create the final consensus network using associations from all the network inference methods.
+  The output of this step are co-occurrence association networks encoded in the JSON format (refer Supplementary section).
+
+  \begin{table}[h]
+    \centering
+    \small
+    \begin{tabular}{|c|c|c|}
+      \hline
+      \textbf{Module} & \textbf{Tool/Database} & \textbf{References} \\
+      \hline
+      Booststrap & \texttt{fastspar\_bootstraps} v1.0 & \cite{Watts2018} \\
+      Booststrap & \texttt{fastspar\_pvalues} v1.0 & \cite{Watts2018} \\
+      \hline
+      Direct & \ac{spieceasi} v1.1.2 & \cite{Kurtz2015} \\
+      Direct & FlashWeave.jl v0.18.1 & \cite{tackmannRapidInferenceDirect2019} \\
+      Direct & \ac{cozine} v1.0 & \cite{haCompositionalZeroinflatedNetwork2020a} \\
+      Direct & \ac{harmonies} v1.0 & \cite{jiangHARMONIESHybridApproach2020} \\
+      Direct & \ac{spring} v1.0.4 & \cite{yoonMicrobialNetworksSPRING2019} \\
+      Direct & \ac{mldm} v1.1 & \cite{Yang2017} \\
+      \hline
+      Correlation & FastSpar (\ac{sparcc}) v1.0 & \cite{Watts2018} \\
+      Correlation & Pearson & - \\
+      Correlation & Spearman & - \\
+      Correlation & propr v2.1.2 & \cite{quinnProprRpackageIdentifying2017} \\
+      \hline
+    \end{tabular}
+    \caption{Tools used in the NI module}
+    \label{tab:ni_tools}
+  \end{table}
 
@@
 
-  The consensus method in \ac{micone} (refer Documentation) uses Equation~\ref{eqn:covariance-pvalues} to estimate the covariance of the pvalues and Equation~\ref{eqn:pvalue-combined} to merge the p-values (obtained from bootstrapping) from different methods.
+  The p-value merging method in \ac{micone} (refer Documentation) uses Equation~\ref{eqn:covariance-pvalues} to estimate the covariance of the pvalues and Equation~\ref{eqn:pvalue-combined} to merge the p-values (obtained from bootstrapping) from the different correlation methods.
+  % FIXME: Not sure what our stance on this is?
   Note that we do not use Pearson and Spearman methods in the p-value merging step and these algorithms are only used for demonstration and comparison.
@@
 
- \subsection*{NOTATION}
+ \subsection*{Notation}
 
-  Note this section defines the notation for the network comparison component of our pipeline, which assumes that all networks to be compared have the same number of nodes
+  This section defines the notation for the consensus network component of our pipeline, which assumes that all networks to be compared have the same number of nodes.
 
-  q = number of nodes
+  q = the number of nodes
 
-  w = number of different networks from different approaches (previously x?)
+  w = the number of co-occurrence networks generated from different approaches
 
-  $N^i$ = network k (currently Ni)
+  $N^i$ = The $i^{th}$ occurrence network represented as a $q \times q$ adjacency matrix, where $i \in {1, q}$
 
-  $N^i_{a,b}$ = edge (a,b) in network i
+  $N^i_{a,b}$ = edge $(a,b)$ in network $i$
 
-  It useful to define a "stretched" version of each adjacency matrix $N^i$ into a column vector, where all columns are stacked onto each other into an $q^2$ long vector:
-
-  $\bar{N}^i_{j}$ , $j=1:q^2$
+  $\bar{N}^i$ is a "flattened" version of the adjacency matrix $N^i$ into a $q^2 \times 1$ column vector, where all columns are stacked onto each other into a $q^2$ long vector:
 
@@
 
-  The consensus algorithm was designed to increase the number of true positives reported at the end of the network inference step.
-For this purpose, we designed two simple algorithms that combine the edges reported by the different network inference tools for each pair of nodes using a consensus based on a parameter \(p\).
-The inputs to both the algorithms are the interaction networks (association matrices) generated by each method \(N_{i}\) and the threshold parameter (p).
-Here, the \(N_{i}\) each have the same set of nodes and only differ by the value of the association inferred between every pair.
+  The consensus algorithm was designed to increase the precision (number of true positives) at the end of the network inference step.
+  For this purpose, we designed two simple algorithms that combine the edges reported by the different network inference tools for each pair of nodes using a consensus based on a parameter $p$.
+  The inputs to both the algorithms are the co-occurrence networks (association matrices) $\bar{N}^i$ (flattened version of $N^{i}$) generated by each method $i$ and the threshold parameter $p$.
+  Here, the $\bar{N}^{i}$ each have the same set of nodes $q$ and only differ by the value of the association inferred between every pair.
+  Networks which do not have a particular node, are updated such that the node is added as an isolated component.
+  In this manner, $\bar{N}^{i}_j$ represents edge $j$ in network $i$.
 
   \subsection*{Algorithm 1: Simple voting}
+   \vspace{-5mm}
 
   The simple voting method performs a voting-based consensus to determine whether an edge will exist between a given node-pair in the final consensus network.
-For each pair of nodes \((\text{node}_{i}, \text{node}_{j})\), we determine the number of network inference methods that report an edge between them  (\(\text{edges}(\text{node}_{i}, \text{node}_{j})\)).
-Each node-pair will have an edge in the final consensus network if the number of reported edges is more than a threshold (Equation \ref{eqn:simple-voting})
+  For each pair of nodes, we determine the number of network inference methods that report an edge $j=k$ between them, i.e. $\bar{N}^{i}_{j=k}, \forall i \in 1:w$.
+  Each node-pair will have an edge in the final consensus network if the number of reported edges is more than the threshold (Equation~\ref{eqn:simple-voting2})
 
- For each edge (i,j), we count number $M_{i,j}$ of networks in which that edge exists, i.e.:
+ For each edge $j$, we obtain $M_j$ which represents the number of networks in which edge $j$ is reported, i.e.:
 
   \begin{equation}
- \begin{aligned}
- M_{i,j} &= | \left{ a \mid N^{a}_{i,j}>0 \right} | \\
- M_{i,j} &\geq \lfloor \theta * w \rfloor
- \end{aligned}
-  \label{eqn:simple-voting}
+      M_j = f(g(\bar{N}^{i=1}_j), \dots, g(\bar{N}^{i=w}_j)) \\
+    \label{eq:simple-voting}
   \end{equation}
 
+  where,
+  $w$ is the number of network inference methods used in the consensus,
+  \begin{equation*}
+    g(x) =
+    \begin{cases}
+       & 0 \text{ if } x=0, \\
+       & -1 \text{ if } x<0, \\
+       & 1 \text{ if } x>0
+    \end{cases}
+  \end{equation*}
+  and
+  \begin{equation*}
+    f(x_1,\dots,x_w) = max
+    \begin{cases}
+       & \sum_i \mid x \in x_i \mid x_i = -1 \mid, \\
+       & \sum_i \left( x \in x_i \mid x_i = 1 \right) \\
+    \end{cases}
+  \end{equation*}
+
+  The edge $j$ appears in the final consensus network if the reported edges is greater than the threshold.
   \begin{equation}
-  \text{edges}(\text{node}_{i}, \text{node}_{j}) \geq \lfloor \theta * w \rfloor
-  \label{eqn:simple-voting}
+     M_j \geq \lfloor p * w \rfloor
+    \label{eqn:simple-voting2}
   \end{equation}
+  where,
+  $p$ is the user-defined parameter threshold
+
+  The simple voting method gives the union of the networks when $p \in (0, \frac{1}{w})$ and will return the intersection when $p \in ((w - 1), 1)$.
+  In general, if $p \in ((n - 1), \frac{n}{w})$, this algorithm will report an edge in the consensus network when at least $n$ network inference methods report this edge.
+
+  \subsubsection*{Algorithm 2: Scaled-sum method}
+   \vspace{-5mm}
+
+  This algorithm performs a consensus based on the sum of all edges reported between a node-pair.
+First, the network generated by each network inference method is re-scaled so that $\max_{j}(\mid \bar{N}^i_j \mid) = 1$ (Equation~\ref{eqn:scaled-sum-rescaling})
 
   \begin{equation}
-   \sum_k N^k_{i,j} \geq \lfloor \theta * w \rfloor
-  \label{eqn:simple-voting2}
-  \end{equation}
-
-  where, \(w\) is the number of network inference methods used in the consensus.
-
- where, \theta is the user-defined threshold....
-
-  The simple voting method gives the union of the networks when \(p \in (0, \frac{1}{x})\) and will return the intersection when \(p \in ((x - 1), 1)\).
-In general, if \(p \in ((n - 1), \frac{n}{x})\), this algorithm will report an edge in the consensus network when at least \(n\) network inference methods report this edge.
-
-  \subsubsection*{Algorithm 2: Scaled-sum method}
-
-  This algorithm performs a consensus based on sum of all edges reported between a node-pair.
-First, the network generated by each network inference method is re-scaled so that (\max(\mid N_{i} \mid) = 1) (Equation \ref{eqn:scaled-sum-rescaling}).
-
-\begin{equation}
-  NS^{a}_{i,j} = \frac{N^{(a)}_{i,j}}{\max(\mid N_{i} \mid)}
-  \label{eqn:scaled-sum-rescaling}
-  \end{equation}
-
-  \begin{equation}
-  N^{scaled}_{i} = \frac{N_{i}}{\max(\mid N_{i} \mid)}
+    S^{i} = \frac{N^{i}}{\max(\mid N^{i} \mid)}
     \label{eqn:scaled-sum-rescaling}
@@
   The re-scaling is performed because the associations reported by the direct association metrics are not necessarily correlations and are not directly comparable to the associations from the correlation-based methods due to their lower weights.
-This problem can also be alleviated by weighing each network inference method differently.
-Next, for each pair of nodes, we sum the weights of all reported edges (Equation \ref{eqn:scaled-sum}).
+  This problem can also be alleviated by weighing each network inference method differently, but for simplicity we chose the current approach.
+  Next, for each edge $j$, we sum the weights of all reported edges from the different networks (Equation \ref{eqn:scaled-sum}).
 
   \begin{equation}
-  \text{scaled\_sum}(\text{node}_{m}, \text{node}_{n}) = \sum_{i} N^{scaled}_{i}(\text{node}_{m}, \text{node}_{n})
+    s_j = \left\mid \sum_i S^i_j \right\mid
     \label{eqn:scaled-sum}
@@
 
-  An edge will exist between a given node-pair in the consensus network if \(\text{scaled\_sum} > (x - 1) * p\), where \(x\) is the number of network inference methods used in the consensus.
-The advantage of this method is that it also takes into account how the strength of the association reported for that particular node is in the inferred networks.
+  An edge $j$ will exist between a given node-pair in the consensus network if:
+  \begin{equation}
+    s_j > (w - 1) \times p
+    \label{eq:scaled-sum2}
+  \end{equation}
+  where, $w$ is the number of network inference methods used in the consensus.
+
+  The advantage of this method is that it also takes into account the strength of the association reported for that particular node in the inferred networks.
 
@@
 
-  The second method, hereby referred to as ``NorTA'', used the Normal to Anything (NorTA) approach coupled with a given interaction network topology to generate the abundance distribution of the microbial community \cite{kurtzSparseCompositionallyRobust2015}.
+  The second method, hereby referred to as ``NorTA'', used the Normal to Anything (NorTA) approach coupled with a given interaction network topology to generate the abundance distribution of the microbial community \cite{Kurtz2015}.
   We used the \href{https://github.com/zdk123/SpiecEasi}{spieceasi} R package to simulate communities with different network topologies (scale-free, cluster, block, Erdos-Renyi, band and hub) and target abundance distributions (Negative Binomial, Poisson, Zero-Inflated Negative Binomial).

### 2022-01-17 sections/methods.tex
Change: Update consensus algorithm subsection in the Methods section

@@
 
-  \subsubsection*{Bootstrapping}
-  For each network inference method (both correlation and direct association based methods), $1000$ permutations of the original \ac{otu} counts data were generated~\cite{Watts2018}.
-  We then recalculate the associations in these permuted \ac{otu} tables using the different network inference algorithms.
-  Finally, we calculate the p-value based on how often a more extreme association is observed for randomly permuted data.
-
-  \subsubsection*{p-value merging (can be moved to the supplementary section)}
-  Fisher~\cite{fisher_224a_1948} proposed that for $k$ independent p-values, each generated by $k$ different methods and denoted by $P_i$, the statistic $\Psi$:
-  \begin{equation*}
-    \begin{aligned}
-        \Psi &= \sum_{i=1}^k -2 \log \left( P_i \right) \\
-        \Psi &\sim \chi^2_{2k}
-    \end{aligned}
-  \end{equation*}
-
-  Brown~\cite{brown_400_1975} extended Fisher's method to dependent p-values by using a re-scaled $\chi^2$ distribution:
-  \begin{equation*}
-    \Psi \sim c \chi^2_{2f}
-  \end{equation*}
-  where, $f$ is the degrees of freedom and $c$ is the scale factor and are given by:
-  \begin{equation*}
-    f = \frac{\mathrm{E}[\Psi]^2}{\mathrm{Var}[\Psi]} ~~~\text{and}~~~ c = \frac{\mathrm{Var}[\Psi]}{2\mathrm{E}[\Psi]} = \frac{k}{f}
-  \end{equation*}
-
-  Furthermore, Brown showed that $\mathrm{E}[\Psi]$ and $\mathrm{Var}[\Psi]$ can be calculated directly via a numerical integration:
-  \begin{equation*}
-    \mathrm{E}[\Psi] = 2k ~~~\text{and}~~~ \mathrm{Var}[\Psi] = 4k + 2\sum_{i<j} \mathrm{Cov}\left( -2\log(P_i), -2\log(P_j) \right)
-  \end{equation*}
-
-  Kost and McDermott~\cite{kost_combining_2002} further fit a third-order polynomial to approximate the covariance
-  \begin{equation}
-    \mathrm{Cov}\left( -2\log(P_i), -2\log(P_j) \right) \approx 3.263 \rho_{ij} + 0.710 \rho_{ij}^2 + 0.027 \rho_{ij}^3
-    \label{eqn:covariance-pvalues}
-  \end{equation}
-  where, $\rho_{ij}$ is the correlation between method $i$ and method $j$
-
-  The final combined p-value~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016} is then given by:
-  \begin{equation}
-    \begin{aligned}
-        & P_{combined} = 1.0 - \Phi_{2f}\left( \psi / c \right) \\
-        \text{where},~ &\psi = -2 \sum_{i=1}^k \log(P_i) ~~~\text{and}~~~ \Phi_{2f} = \mathrm{CDF}\left( \chi^2_{2f} \right)
-    \end{aligned}
-    \label{eqn:pvalue-combined}
-  \end{equation}
-
-  The p-value merging method in \ac{micone} (refer Documentation) uses Equation~\ref{eqn:covariance-pvalues} to estimate the covariance of the pvalues and Equation~\ref{eqn:pvalue-combined} to merge the p-values (obtained from bootstrapping) from the different correlation methods.
-  % FIXME: Not sure what our stance on this is?
-  Note that we do not use Pearson and Spearman methods in the p-value merging step and these algorithms are only used for demonstration and comparison.
-  The combined p-values are used to threshold for significance during the consensus network step.
-
- \subsection*{Notation}
+ \subsubsection*{Notation}
 
@@
 
-  $N^i$ = The $i^{th}$ occurrence network represented as a $q \times q$ adjacency matrix, where $i \in {1, q}$
+  $N^i$ = The weights of the $i^{th}$ co-occurrence network represented as a $q \times q$ adjacency matrix, where $i \in {1, q}$
+  $P^i$ = The pvalues of the $i^{th}$ co-occurrence network represented as a $q \times q$ adjacency matrix, where $i \in {1, q}$
 
@@
   $\bar{N}^i$ is a "flattened" version of the adjacency matrix $N^i$ into a $q^2 \times 1$ column vector, where all columns are stacked onto each other into a $q^2$ long vector:
+  $\bar{P}^i$ is a "flattened" version of the adjacency matrix $P^i$ into a $q^2 \times 1$ column vector, where all columns are stacked onto each other into a $q^2$ long vector:
 
-  \subsection*{Consensus method}
+  \subsubsection*{Bootstrapping and p-value calculation}
+  For all correlation-based methods $k \in 1:w$, $1000$ permutations of the original \ac{otu} counts data were generated~\cite{Watts2018}.
+  All the direct association based methods used in the study have their own regularization methods built-in and hence do not need to undergo bootstrapping.
+  We then recalculate the correlations in these permuted \ac{otu} tables using the different correlation-based algorithms.
+  Finally, we calculate the p-value based on how often a more extreme association is observed for randomly permuted data.
+
+  \subsubsection*{p-value merging}
+
+  The first part of the consensus algorithm workflow is to merge the p-values for the networks generated by the correlation-based methods.
+  This step is performed using the Brown's p-value merging method~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016,faustCoNetAppInference2016} (refer Supplementary).
+
+  % TODO: Is this necessary here, should this be elaborated?
+  The final combined p-value~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016} is then given by:
+  \begin{equation}
+    \begin{aligned}
+        & \hat{P}_j = 1.0 - \Phi_{2f}\left( \psi / c \right) \\
+        \text{where},~ &\psi = -2 \sum_{i=1}^k \log(\bar{P}^i_j) ~~~\text{and}~~~ \Phi_{2f} = \mathrm{CDF}\left( \chi^2_{2f} \right)
+    \end{aligned}
+    \label{eqn:pvalue-combined}
+  \end{equation}
+  where, $\hat{P}_j$ is the combined p-value for the edge $j$
+
+  % FIXME: Not sure what our stance on this is?
+  Note that we do not use Pearson and Spearman methods in the p-value merging step and these algorithms are only used for demonstration and comparison.
+  The combined p-values are used to threshold for significance right before the consensus algorithm is applied to the inferred networks.
+
+  \subsubsection*{Consensus method}
   \vspace{-5mm}
@@
 
-  \subsection*{Algorithm 1: Simple voting}
-   \vspace{-5mm}
+  \paragraph*{Algorithm 1: Simple voting}
 
@@
 
-  \subsubsection*{Algorithm 2: Scaled-sum method}
-   \vspace{-5mm}
+  \paragraph*{Algorithm 2: Scaled-sum method}
 
@@
   \begin{equation}
-    S^{i} = \frac{N^{i}}{\max(\mid N^{i} \mid)}
+    S^{i} = \frac{\bar{N}^{i}}{\max(\mid \bar{N}^{i} \mid)}
     \label{eqn:scaled-sum-rescaling}

### 2022-01-17 sections/supplementary.tex
Change: Update consensus algorithm subsection in the Methods section

@@
   Figures describing the original OTU tables or the interactions used to generate the community. Also include the seed and scripts for reproducing the networks.
+
+  \subsection{Consensus and p-value}
+
+  % FIXME: Should this section be rewritten? Ask Daniel
+  \subsubsection*{p-value merging}
+  Fisher~\cite{fisher_224a_1948} proposed that for $k$ independent p-values, each generated by $k$ different methods and denoted by $P_i$, the statistic $\Psi$:
+  \begin{equation*}
+    \begin{aligned}
+        \Psi &= \sum_{i=1}^k -2 \log \left( P_i \right) \\
+        \Psi &\sim \chi^2_{2k}
+    \end{aligned}
+  \end{equation*}
+
+  Brown~\cite{brown_400_1975} extended Fisher's method to dependent p-values by using a re-scaled $\chi^2$ distribution:
+  \begin{equation*}
+    \Psi \sim c \chi^2_{2f}
+  \end{equation*}
+  where, $f$ is the degrees of freedom and $c$ is the scale factor and are given by:
+  \begin{equation*}
+    f = \frac{\mathrm{E}[\Psi]^2}{\mathrm{Var}[\Psi]} ~~~\text{and}~~~ c = \frac{\mathrm{Var}[\Psi]}{2\mathrm{E}[\Psi]} = \frac{k}{f}
+  \end{equation*}
+
+  Furthermore, Brown showed that $\mathrm{E}[\Psi]$ and $\mathrm{Var}[\Psi]$ can be calculated directly via a numerical integration:
+  \begin{equation*}
+    \mathrm{E}[\Psi] = 2k ~~~\text{and}~~~ \mathrm{Var}[\Psi] = 4k + 2\sum_{i<j} \mathrm{Cov}\left( -2\log(P_i), -2\log(P_j) \right)
+  \end{equation*}
+
+  Kost and McDermott~\cite{kost_combining_2002} further fit a third-order polynomial to approximate the covariance
+  \begin{equation}
+    \mathrm{Cov}\left( -2\log(P_i), -2\log(P_j) \right) \approx 3.263 \rho_{ij} + 0.710 \rho_{ij}^2 + 0.027 \rho_{ij}^3
+    \label{eqn:covariance-pvalues}
+  \end{equation}
+  where, $\rho_{ij}$ is the correlation between method $i$ and method $j$
+
+  The final combined p-value~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016} is then given by:
+  \begin{equation}
+    \begin{aligned}
+        & P_{combined} = 1.0 - \Phi_{2f}\left( \psi / c \right) \\
+        \text{where},~ &\psi = -2 \sum_{i=1}^k \log(P_i) ~~~\text{and}~~~ \Phi_{2f} = \mathrm{CDF}\left( \chi^2_{2f} \right)
+    \end{aligned}
+    \label{eqn:pvalue-combined}
+  \end{equation}
+
+  The p-value merging method in \ac{micone} (refer Documentation) uses Equation~\ref{eqn:covariance-pvalues} to estimate the covariance of the pvalues and Equation~\ref{eqn:pvalue-combined} to merge the p-values (obtained from bootstrapping) from the different correlation methods.
+  % FIXME: Not sure what our stance on this is?
+  Note that we do not use Pearson and Spearman methods in the p-value merging step and these algorithms are only used for demonstration and comparison.
+  The combined p-values are used to threshold for significance during the consensus network step.
 

### 2022-01-18 scripts/figure2/figure2_data.py
Fix: ANOVA variance calculation
Variance estimate should be the mean_sq not the sum_sq

@@ -111,7 +111,10 @@ def normalize_anova(anova_dict: dict, pca: PCA) -> list:
     assert np.isclose(sum(var_ratio), 1)
     variance_list = []
     for component in anova_dict:
-        variance = anova_dict[component]["sum_sq"]
+        anova_dict[component]["mean_sq"] = (
+            anova_dict[component]["sum_sq"] / anova_dict[component]["df"]
+        )
+        variance = anova_dict[component]["mean_sq"]
         ratio = var_ratio[component]
         variance_list.append(variance * ratio)
     return variance_list

### 2022-01-18 scripts/figure2/figure2_plot.R

@@ -20,9 +20,9 @@ percentage_variance <- read.csv("../../data/figure2/output/moving_pictures/perce
 
 percentage_variance <-
   percentage_variance %>%
-  mutate(labels = scales::percent(sum_sq, scale=1)) %>%
+  mutate(labels = scales::percent(mean_sq, scale=1)) %>%
   arrange(desc(X)) %>%
-  mutate(text_y = cumsum(sum_sq) - sum_sq / 2)
+  mutate(text_y = cumsum(mean_sq) - mean_sq / 2)
 
 
 # df <- y_reduced %>% select("hash", "X0", "X1") %>% filter((abs(X0) < 5) & (abs(X1) < 10)) %>% left_join(x, by="hash")
@@ -30,7 +30,7 @@ df <- y_reduced %>% select("hash", "X0", "X1") %>% left_join(x, by="hash")
 scatter_ta <- plot_scatter(df, "TA")
 scatter_ni <- plot_scatter(df, "NI")
 
-pie_chart <- ggplot(data = percentage_variance, aes(x = "", y = sum_sq, fill = X)) +
+pie_chart <- ggplot(data = percentage_variance, aes(x = "", y = mean_sq, fill = X)) +
                        geom_bar(stat = "identity") + 
                        geom_label_repel(aes(label = labels, y = text_y), nudge_x = 1.6) +
                        coord_polar(theta = "y") +

### 2022-01-18 sections/methods.tex
Add: Updated variance calculation details to Methods section

@@
 
+ % TODO: Use bold for matrices and lower-case bold for vectors
   This section defines the notation for the consensus network component of our pipeline, which assumes that all networks to be compared have the same number of nodes.
 
-  q = the number of nodes
+  $w$ = the number of co-occurrence network algorithms
 
-  w = the number of co-occurrence networks generated from different approaches
+  $q$ = the number of combined nodes from all networks in $w$
 
-  $N^i$ = The weights of the $i^{th}$ co-occurrence network represented as a $q \times q$ adjacency matrix, where $i \in {1, q}$
+  $N^i$ = The weights of the $i^{th}$ co-occurrence network represented as a $q \times q$ adjacency matrix, where $i \in {1, w}$
+
   $P^i$ = The pvalues of the $i^{th}$ co-occurrence network represented as a $q \times q$ adjacency matrix, where $i \in {1, q}$
@@
   $\bar{N}^i$ is a "flattened" version of the adjacency matrix $N^i$ into a $q^2 \times 1$ column vector, where all columns are stacked onto each other into a $q^2$ long vector:
+
   $\bar{P}^i$ is a "flattened" version of the adjacency matrix $P^i$ into a $q^2 \times 1$ column vector, where all columns are stacked onto each other into a $q^2$ long vector:
@@
 
-  \subsubsection*{Network Variability}
+  \subsection*{Network Variability}
   \vspace{-5mm}
- In order to compare across different networks, and analyze the degree of variability induced by the choice of different modules and parameters, we organized multiple networks into a single mathematical structure that we could use for linear regression.
- In particular, we transformed the adjacency matrix of each co-occurrence network into a vector.
- We then merged the networks generated from all possible combinations of tools into a table (N, see below) in which each column represents one network.
 
- The merged table $N$ with $p$ edges and $n$ networks, where, each column $N_j$ is the vector representation of one of the networks, each row $L_i$ is the vector representation of one particular edge in all networks (assigned a value of 0 if the edge did not exist in the network but in other networks), and each element $E_{i,j}$ belongs to edge $L_i$ and network $N_j$.
+  \subsubsection*{Notation}
+  \vspace{-5mm}
+  This section defines notation for the network variability analysis in our study\dots
+
+  $w^\prime$ = the number of co-occurrence networks generated from all possible combinations of tools and parameters in the workflow.
+
+  $q^\prime$ = the number of combined nodes from all networks in $w^\prime$
+
+  % TODO: Should we use different variable names here?
+  $N^i$ = The weights of the $i^{th}$ co-occurrence network represented as a $q^\prime \times q^\prime$ adjacency matrix, where $i \in {1, w^\prime}$
+
+  $N^i_{a,b}$ = edge $(a,b)$ in network $i$
+
+  $\bar{N}_i$ is a "flattened" version of the adjacency matrix $N^i$ into a $q^\prime^2 \times 1$ column vector, where all columns are stacked onto each other into a $q^\prime^2$ long vector:
+
+  \subsubsection*{Principal Component Analysis and variability calculation}
+  \vspace{-5mm}
+ In order to compare across different networks, and analyze the degree of variability induced by the choice of different modules, we organized multiple networks into a single mathematical structure that we could use for linear regression.
+ First, we obtain co-occurrence network $\bar{N}_i$ for each possible tool and parameter combination in the workflow.
+ Then, we obtain the merged matrix $\mathbf{\bar{N}}$ with $q^{\prime2}$ edges from $w^\prime$ networks, where, each column $i$ is the vector representation of one of the networks $\bar{N}^i$, each row $j$ is the vector representation of one particular edge in all networks $\bar{N}_j$ (assigned a value of 0 if the edge did not exist in the network but in other networks), and each element $n^i_j$ belongs to edge $j$ and network $i$.
 
   \begin{equation*}
-   N_{p \times n} =
-   \begin{bmatrix}
-     L_1 \\
-     L_2 \\
-     \vdots \\
-     L_p
-   \end{bmatrix}
-   =
-   \begin{bmatrix}
-     E_{1,1} & E_{1,2} & \cdots  & E_{1, n} \\
-     E_{2,1} & E_{2,2} & \cdots  & E_{2, n} \\
-     \vdots & \vdots & \vdots  & \vdots \\
-     E_{p,1} & E_{p,2} & \cdots  & E_{p, n}
-   \end{bmatrix}
+   \mathbf{\bar{N}} =
+     \begin{blockarray}{*{4}{c} l}
+      \begin{block}{*{4}{>{$\footnotesize}c<{$}} l}
+        $\bar{N}$^1 & $\bar{N}$^2 & \dots & $\bar{N}$^{w^\prime} & \\
+      \end{block}
+      \begin{block}{[*{4}{c}]>{$\footnotesize}l<{$}}
+       n^1_1 & n^2_1 & \cdots  & n^{w^\prime}_1 \bigstrut[t]& $\bar{N}$_1 \\
+       n^1_2 & n^2_2 & \cdots  & n^{w^\prime}_2 & $\bar{N}$_2 \\
+       \vdots & \vdots & \vdots  & \vdots & \vdots \\
+       n^1_{q^\prime^2} & n^2_{q^\prime^2} & \cdots  & n^{w^\prime}_{q^\prime^2} & $\bar{N}$_{q^\prime^2}
+      \end{block}
+     \end{blockarray}
   \end{equation*}
 
-  In other words, $N$ is the merged table, each column $N_i$ is the vector representation of one of the networks, and each row $L_i$ represents one particular edge in all networks (assigned 0 if the edge does not exist in the network).
+  To infer the variability contributed due to different steps in the pipeline we perform a linear regression on each edge in $\mathbf{\bar{N}}$ and a subsequent ANOVA to extract the group variances.
+  One issue with this approach is that the possible correlations between the edges of the network could lead to inaccurate estimates of the variance if a GLM were used to directly model the relationships between edges and steps in the workflow.
+  In order to remedy this issue, we perform Principal Component Analysis on the matrix $\mathbf{\bar{N}}$ to obtain the $\mathbf{C}$ matrix ($w^\prime \times c$) of components for each network, such that we reduce the dimensions from the $q^\prime^2$ dimensional edge space to a $c$ dimensional component space.
 
-  We use linear regression to express each link $L_i$ as a linear function of categorical variables that describe the possible options in each of the first three steps of the pipeline.
+  We then use linear regression to express each component $C_j$ (where $j \in 1:c$) as a linear function of categorical variables that describe the possible options in each of the steps of the pipeline.
 
-  % TODO: Explain the categorical linear model and ANOVA better (text)
+  % TODO: Should we remove CC from the LM model?
   In particular, we infer parameters $\alpha_i$ such that:
    \begin{equation*}
-       L_i = \sum_{j=1}^5 \left( \alpha^{DC(j)}_i.\delta^{DC(j)}_i \right) +
-             \sum_{j=1}^3 \left( \alpha^{TA(j)}_i.\delta^{TA(j)}_i \right) +
-             \sum_{j=1}^2 \left( \alpha^{OP(j)}_i.\delta^{OP(j)}_i \right) +
-             \epsilon_i
+       C_j = \sum_{k=1}^5 \left( \alpha^{DC(k)}_j.\delta^{DC(k)}_j \right) +
+             \sum_{k=1}^2 \left( \alpha^{CC(k)}_j.\delta^{CC(k)}_j \right) +
+             \sum_{k=1}^3 \left( \alpha^{TA(k)}_j.\delta^{TA(k)}_j \right) +
+             \sum_{k=1}^2 \left( \alpha^{OP(k)}_j.\delta^{OP(k)}_j \right) +
+             \sum_{k=1}^{10} \left( \alpha^{NI(k)}_j.\delta^{NI(k)}_j \right) +
+             \epsilon_j
    \end{equation*}
@@
    where, $\alpha_i$ are the coefficients of the regression, $\epsilon_i$ are the residuals and $\delta_i$ are the indicator variables that correspond to the processes utilized in the pipeline used to create the network $N_i$; for example, $\delta^{DC(1)}_i = 1$ if the DC(1) process was used in the generation of the network $N_i$ .
-   Here, (i) DC(1) = "closed reference", DC(2) = "open reference", DC(3) = "de novo", DC(4) = "dada2", DC(5) = "deblur"; (ii)  TA(1) = "GreenGenes", TA(2) = "SILVA", TA(3) = "NCBI"; (iii) OP(1) = "no filtering", OP(2) = "filtering".
+   Here, (i) DC(1) = "closed reference", DC(2) = "open reference", DC(3) = "de novo", DC(4) = "dada2", DC(5) = "deblur";
+   (ii) CC(1) = "remove bimera", CC(2) = "uchime";
+   (iii)  TA(1) = "GreenGenes", TA(2) = "SILVA", TA(3) = "NCBI";
+   (iv) OP(1) = "no filtering", OP(2) = "filtering";
+   (v) NI(1) = "SparCC", NI(2) = "Pearson", NI(5) = "Spearman", NI(4) = "Propr", NI(5) = "SpiecEasi", NI(6) = "COZINE", NI(7) = "HARMONIES", NI(8) = "SPRING", NI(9) = "mLDM", NI(10) = "FlashWeave";
 
-  The variance contributed by each step of the pipeline is calculated for every connection in the merged table through ANOVA using the Python statsmodels package and is shown in Figure~\ref{fig:figure2}A.
+  The variance contributed by each step of the pipeline is calculated for every component in $\mathbf{C}$ matrix through ANOVA using the Python \texttt{statsmodels} package and is shown in Figure~\ref{fig:figure2}A.
   The total variance for the network is calculated by adding the variances for each connection.
-  The merged network table $N_{p \times n}$ is used as the input to the PCA analysis to generate Figure~\ref{fig:figure2}B.
+  The merged network table $\mathbf{\bar{N}}$ is used as the input to the PCA analysis to generate Figure~\ref{fig:figure2}B.
 
-  The possible correlations between the edges of the network lead to inaccurate estimates of the variance if a GLM were used to directly model the relationships between the links and steps in the workflow.
-Therefore, we transformed the network matrix $N$ using a Principal Component Analysis into the $N_{reduced}$ matrix.
-The dimensions of the $N_{reduced}$ matrix are be denoted by $(d \times n)$ where d is the number of principal components and n is the number of network in the $N$ matrix.
+  % TODO: This section
+  \subsubsection*{Canonical Correlation Analysis}
 

### 2022-01-18 sections/methods.tex
Add: Update synthetic data methods and repository link

@@
 
-  % TODO: This section
-  \subsubsection*{Canonical Correlation Analysis}
-
   \subsection*{Synthetic interaction data}
@@
 
-  The first method, hereby referred to as ``seqtime'', used generalized Lotka-Volterra (gLV) equations to model the microbial community dynamics and utilized the Klemm–Eguı́luz algorithm to generate a clique-based interaction network~\cite{rottjersHairballsHypothesesBiological2018}.
+  The first method, hereby referred to as ``seqtime'', used generalized Lotka-Volterra (gLV) equations to model the microbial community dynamics and utilized the Klemm–Eguı́luz algorithm to generate a clique-based interaction network~\cite{Rottjers2018}.
   We used the \href{https://github.com/hallucigenia-sparsa/seqtime}{seqtime} R package to simulate communities with number of species ($N$) varying from 10 to 150 (10, 25, 50, 100, 150 and 200).
@@
 
-  For each method, we generated the OTU table depicting the abundances of species and used this to generate association networks using \ac{micone}.
+  For each method, we generated the OTU table depicting the abundances of species and use this as input to generate association networks using \ac{micone} pipeline.
   The interaction matrix was used as the source of expected (true) interactions and the associations predicted using \ac{micone} were the source of predicted interactions.
@@
   Data and scripts: \href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper}
+  Synthetic data and scripts: \href{https://github.com/segrelab/MiCoNE-synthetic-data}{https://github.com/segrelab/MiCoNE-synthetic-data}

### 2022-01-19 sections/methods.tex
Update on Overleaf.

@@
  % TODO: Use bold for matrices and lower-case bold for vectors
-  This section defines the notation for the consensus network component of our pipeline, which assumes that all networks to be compared have the same number of nodes.
+  This section defines the notation used below to describe the consensus network algorithm used in the our pipeline. Note that all networks to be compared have the same number of nodes.
 
-  $w$ = the number of co-occurrence network algorithms
+  $w$ = number of co-occurrence networks to be integrated in the consensus network (corresponding to the number of the different approaches used to infer co-occurrence networks)
 
-  $q$ = the number of combined nodes from all networks in $w$
+  $q$ = number of nodes in any co-occurrence network (assumed to be the same across all $w$ networks)
 
-  $N^i$ = The weights of the $i^{th}$ co-occurrence network represented as a $q \times q$ adjacency matrix, where $i \in {1, w}$
+  $N^i$ = matrix of network edge Weights for the $i^{th}$ co-occurrence network. This is a $q \times q$ matrix, where $i \in \{1,\dots,w\}$. $N^i_{a,b}$ represents edge $(a,b)$ in network $i$
 
-  $P^i$ = The pvalues of the $i^{th}$ co-occurrence network represented as a $q \times q$ adjacency matrix, where $i \in {1, q}$
+  $P^i$ = matrix of p-values for all edges of the $i^{th}$ co-occurrence network. This is a $q \times q$ matrix, where $i \in \{1,\dots,q\}$
 
-  $N^i_{a,b}$ = edge $(a,b)$ in network $i$
+  $\bar{N}^i$ is a "flattened" version of the adjacency matrix $N^i$ into a $q^2 \times 1$ column vector, where all columns are stacked onto each other into a $q^2$ long vector.
+Element $\bar{N}^i_j$ corresponds to the $j^{th}$ edge in the $i^{th}$ network.
 
-  $\bar{N}^i$ is a "flattened" version of the adjacency matrix $N^i$ into a $q^2 \times 1$ column vector, where all columns are stacked onto each other into a $q^2$ long vector:
-
-  $\bar{P}^i$ is a "flattened" version of the adjacency matrix $P^i$ into a $q^2 \times 1$ column vector, where all columns are stacked onto each other into a $q^2$ long vector:
+  $\bar{P}^i$ is a "flattened" version of the adjacency matrix $P^i$ into a $q^2 \times 1$ column vector, where all columns are stacked onto each other into a $q^2$ long vector.
 
   \subsubsection*{Bootstrapping and p-value calculation}
-  For all correlation-based methods $k \in 1:w$, $1000$ permutations of the original \ac{otu} counts data were generated~\cite{Watts2018}.
+  For all correlation-based methods $k \in \{1,\dots,w\}$, $1000$ permutations of the original \ac{otu} counts data were generated~\cite{Watts2018}.
   All the direct association based methods used in the study have their own regularization methods built-in and hence do not need to undergo bootstrapping.
@@
   The first part of the consensus algorithm workflow is to merge the p-values for the networks generated by the correlation-based methods.
-  This step is performed using the Brown's p-value merging method~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016,faustCoNetAppInference2016} (refer Supplementary).
+  This step is performed using the Brown's p-value merging method~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016,faustCoNetAppInference2016}.
 
   % TODO: Is this necessary here, should this be elaborated?
-  The final combined p-value~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016} is then given by:
+  As described in more detail in the Supplementary Material and in the original reference~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016}, the final combined p-value is given by:
   \begin{equation}
@@
   \end{equation}
-  where, $\hat{P}_j$ is the combined p-value for the edge $j$
+  where, $\hat{P}_j$ is the combined p-value for the edge $j$, f is the number of degrees of freedom, and c is a scale factor.
 
@@
 
+    Note that this method is used to filter relevant interactions.
+However, if a given pair of nodes is inferred to have edges that satisfy the above relationships, all corresponding edges from the $w$ networks will be returned by the algorithm.
+
   \paragraph*{Algorithm 1: Simple voting}
+  % TODO: Add a reference
+  The simple voting method performs a voting-based consensus to determine whether an edge will exist between a given node-pair in the final consensus network.
+  For each pair of nodes, we determine the number of network inference methods that report an edge $j=k$ between them, i.e. $\bar{N}^{i}_{j=k}, \forall i \in \{1,\dots,w\}$.
+  Each node-pair will have an edge in the final consensus network if the number of reported edges is larger than a user-defined threshold.
+  (Equation~\ref{eqn:simple-voting2})
 
-  The simple voting method performs a voting-based consensus to determine whether an edge will exist between a given node-pair in the final consensus network.
-  For each pair of nodes, we determine the number of network inference methods that report an edge $j=k$ between them, i.e. $\bar{N}^{i}_{j=k}, \forall i \in 1:w$.
-  Each node-pair will have an edge in the final consensus network if the number of reported edges is more than the threshold (Equation~\ref{eqn:simple-voting2})
-
- For each edge $j$, we obtain $M_j$ which represents the number of networks in which edge $j$ is reported, i.e.:
+ The number of reported edges is computed as follows: for each edge $j$, we obtain $M_j$ which represents the number of networks in which edge $j$ is reported, i.e.:
 
@@
 
-  where,
-  $w$ is the number of network inference methods used in the consensus,
+where, $g$ and $f$ are defined as follows:
+
   \begin{equation*}
@@
     \begin{cases}
-       & 0 \text{ if } x=0, \\
-       & -1 \text{ if } x<0, \\
-       & 1 \text{ if } x>0
+       & 0, \text{ if } x=0, \\
+       & -1, \text{ if } x<0, \\
+       & 1, \text{ if } x>0
     \end{cases}
@@
   \begin{equation*}
-    f(x_1,\dots,x_w) = max
-    \begin{cases}
-       & \sum_i \mid x \in x_i \mid x_i = -1 \mid, \\
-       & \sum_i \left( x \in x_i \mid x_i = 1 \right) \\
-    \end{cases}
+    f(x_1,\dots,x_w) = max \left( \#(i \mid x_i=-1),\#(i \mid x_i=1)
+       \right)
   \end{equation*}
 
-  The edge $j$ appears in the final consensus network if the reported edges is greater than the threshold.
+ The edge $j$ appears in the final consensus network if the reported edges is greater than the threshold:
   \begin{equation}
-     M_j \geq \lfloor p * w \rfloor
+     M_j \geq \lfloor p \times w \rfloor
     \label{eqn:simple-voting2}
@@
   where,
-  $p$ is the user-defined parameter threshold
+  $p$ is the user-defined parameter threshold.
 
   The simple voting method gives the union of the networks when $p \in (0, \frac{1}{w})$ and will return the intersection when $p \in ((w - 1), 1)$.
-  In general, if $p \in ((n - 1), \frac{n}{w})$, this algorithm will report an edge in the consensus network when at least $n$ network inference methods report this edge.
+  In general, if $p \in ((n - 1), \frac{n}{w})$, this algorithm will report an edge in the consensus network when at least $n$ network inference methods report this edge. [REFERENCE]
 
@@
 
-  This algorithm performs a consensus based on the sum of all edges reported between a node-pair.
-First, the network generated by each network inference method is re-scaled so that $\max_{j}(\mid \bar{N}^i_j \mid) = 1$ (Equation~\ref{eqn:scaled-sum-rescaling})
+  This algorithm generates a consensus network based on the sum of all edges reported between a pair of nodes.
+Since in generating this consensus network we combine direct association metrics with correlation-based methods, summing of the edges is preceded by a pre-processing step, in which all networks are re-scaled.
+
+  First, the network generated by each network inference method ($\bar{N}^i$) is re-scaled into a normalized version ($\bar{S}^i$), as follows  (Equation~\ref{eqn:scaled-sum-rescaling}):
 
   \begin{equation}
-    S^{i} = \frac{\bar{N}^{i}}{\max(\mid \bar{N}^{i} \mid)}
+    \bar{S}^{i} = \frac{\bar{N}^{i}}{\max(\mid \bar{N}^{i} \mid)}
     \label{eqn:scaled-sum-rescaling}
@@
 
-  The re-scaling is performed because the associations reported by the direct association metrics are not necessarily correlations and are not directly comparable to the associations from the correlation-based methods due to their lower weights.
-  This problem can also be alleviated by weighing each network inference method differently, but for simplicity we chose the current approach.
-  Next, for each edge $j$, we sum the weights of all reported edges from the different networks (Equation \ref{eqn:scaled-sum}).
+    In this way, it is guaranteed that $\max_{j}(\mid \bar{S}^i_j \mid) = 1$.
+
+  Next, for each edge $(a,b)$, we sum the weights of all reported edges from the different networks (Equation \ref{eqn:scaled-sum}).
 
   \begin{equation}
-    s_j = \left\mid \sum_i S^i_j \right\mid
+    s_j = \sum_i \bar{S}^i_j 
     \label{eqn:scaled-sum}
@@
 
-  An edge $j$ will exist between a given node-pair in the consensus network if:
+  An edge $j$ will be included in the consensus network if:
   \begin{equation}
-    s_j > (w - 1) \times p
+    \mid s_j \mid > (w - 1) \times p
     \label{eq:scaled-sum2}
-  \end{equation}
-  where, $w$ is the number of network inference methods used in the consensus.
+  \end{equation}.
 
-  The advantage of this method is that it also takes into account the strength of the association reported for that particular node in the inferred networks.
+  The advantage of this method over the simple voting method is that it also takes into account the strength of the association reported for that particular node in the inferred networks.
 
@@
   This section defines notation for the network variability analysis in our study\dots
+  % TODO: Update W and Q in the entire subsection
+  $W$ =  number of co-occurrence networks generated from all possible combinations of tools and parameters in the workflow.
+Note that this is different from $w$, which counted only the possible network inference modules.
 
-  $w^\prime$ = the number of co-occurrence networks generated from all possible combinations of tools and parameters in the workflow.
-
-  $q^\prime$ = the number of combined nodes from all networks in $w^\prime$
+  $Q$ =  number of combined nodes from all $W$ networks
 
@@
  In order to compare across different networks, and analyze the degree of variability induced by the choice of different modules, we organized multiple networks into a single mathematical structure that we could use for linear regression.
- First, we obtain co-occurrence network $\bar{N}_i$ for each possible tool and parameter combination in the workflow.
- Then, we obtain the merged matrix $\mathbf{\bar{N}}$ with $q^{\prime2}$ edges from $w^\prime$ networks, where, each column $i$ is the vector representation of one of the networks $\bar{N}^i$, each row $j$ is the vector representation of one particular edge in all networks $\bar{N}_j$ (assigned a value of 0 if the edge did not exist in the network but in other networks), and each element $n^i_j$ belongs to edge $j$ and network $i$.
+ First, we obtain co-occurrence network $\bar{N}_i$ for each of the $W$ possible tool and parameter combinations in the workflow.
+ Then, we construct a matrix $\mathbf{\bar{N}}$ whose $i^{th}$ column is the flattened version of the $i^{th}$ network, i.e.
+the column vector $\bar{N}^i$.
+Therefore $\bar{N}^i_j$ is the weight of edge $j$ in network $i$.
+$\bar{N}^i_j$ is assigned a value of 0 if edge $j$ did not exist in network $i$ but was present in other networks.
+Note that row $j$ of $\mathbf{\bar{N}}$, $\bar{N}_j$ is the vector that encodes the values of edge $j$ across all the networks.
 
@@
 
-  To infer the variability contributed due to different steps in the pipeline we perform a linear regression on each edge in $\mathbf{\bar{N}}$ and a subsequent ANOVA to extract the group variances.
-  One issue with this approach is that the possible correlations between the edges of the network could lead to inaccurate estimates of the variance if a GLM were used to directly model the relationships between edges and steps in the workflow.
+  To infer the variability contributed due to different steps in the pipeline we perform a linear regression on each edge in $\mathbf{\bar{N}}$ and a subsequent ANOVA to extract the within group variances.
+  One issue with this approach is that the possible correlations between the edges of the network could lead to inaccurate estimates of the variance if a linear model were used to directly model the relationships between edges and steps in the workflow.
   In order to remedy this issue, we perform Principal Component Analysis on the matrix $\mathbf{\bar{N}}$ to obtain the $\mathbf{C}$ matrix ($w^\prime \times c$) of components for each network, such that we reduce the dimensions from the $q^\prime^2$ dimensional edge space to a $c$ dimensional component space.
@@
   Documentation: \href{https://micone.readthedocs.io}{https://micone.readthedocs.io} \\
-  Data and scripts: \href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper}
+  Data and scripts: \href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper} \\
   Synthetic data and scripts: \href{https://github.com/segrelab/MiCoNE-synthetic-data}{https://github.com/segrelab/MiCoNE-synthetic-data}

### 2022-04-11 sections/methods.tex
Change: Update the pipeline section of Results

@@
 
-  The second method, hereby referred to as ``NorTA'', used the Normal to Anything (NorTA) approach coupled with a given interaction network topology to generate the abundance distribution of the microbial community \cite{Kurtz2015}.
+  The second method, hereby referred to as ``NorTA'', used the Normal to Anything (NorTA) approach coupled with a given interaction network topology to generate the abundance distribution of the microbial community~\cite{Kurtz2015}.
   We used the \href{https://github.com/zdk123/SpiecEasi}{spieceasi} R package to simulate communities with different network topologies (scale-free, cluster, block, Erdos-Renyi, band and hub) and target abundance distributions (Negative Binomial, Poisson, Zero-Inflated Negative Binomial).

### 2022-04-11 sections/results.tex
Add: OP, NI steps and consensus network to methods section
Change: Update the pipeline section of Results

@@
   The \ac{micone} pipeline workflow is shown in Figure~\ref{fig:figure1}.
-  The different steps for going from 16S data to co-occurrence networks can be grouped into four major modules; (i) the denoising and clustering (DC) step, which handles denoising of the raw 16S sequencing data into representative sequences; (ii) the taxonomy assignment (TA) step that assigns taxonomic labels to the representative sequences; (iii) the \ac{otu} processing (OP) step that filters and transforms the taxonomy abundance table; and finally (iv) the network inferences (NI) step which infers the microbial co-occurrence network.
+  The different steps for going from 16S data to co-occurrence networks can be grouped into five major modules; (i) the sequence processing (SP) step, which handles the demultiplexing, quality control and trimming of the raw 16S sequencing data into demultiplexed and trimmed sequences; (ii) the denoising and clustering (DC) step, which handles denoising of the demultiplexed and trimmed sequences into representative sequences; (iii) the taxonomy assignment (TA) step that assigns taxonomic labels to the representative sequences; (iv) the \ac{otu} processing (OP) step that filters and transforms the taxonomy abundance table; and finally (v) the network inferences (NI) step which infers the microbial co-occurrence network.
   Each process in the pipeline supports alternate tools for performing the same task (see Methods and Figure~\ref{fig:figure1}).
   A centralized configuration file contains all the specifications for what modules are used in the pipeline, and can be modified by the user to choose the desired set of tools.
+  The default tools and parameters for the configuration file can be found in the documentation (see Methods).
   In what follows, we perform a systematic analysis of each step of the pipeline to estimate how much the final co-occurrence network depends on the possible choices at each step.
@@
 
-  Our analysis involves two types of data: The first type consists of sets of 16S sequencing data from real communities sampled from human Stool and Oral microbiomes.
-  The second are datasets synthetically or artificially created for the specific goal of helping evaluate computational analysis tools (see Methods).
-  In particular, in order to objectively compare, to the extent possible, how well each step in \ac{micone} best captures the underlying data, we use both mock data (labelled mock4, mock12 and mock16) from mockrobiota~\cite{Bokulich2016} as well as, synthetically generated reads from an Illumina read simulator called ART~\cite{Huang2012}.
-  These mock datasets consist of fake sequencing reads generated from reads obtained from synthetic microbial isolates mixed in know proportions.
-They contain the expected compositions along with the reference sequences for the organisms in the mock community.
-  The synthetic reads were simulated using three different taxonomy distribution profiles, namely soil and water microbiomes obtained \ac{emp}~\cite{Thompson2017} and Stool microbiome that is used in our real community analysis~\cite{Kang2017}.
-  Reference sequences were generated using \ac{ncbi} and the Decard package~\cite{Golob2017} for these taxonomy profiles.
+  Our analysis involves two types of data: The first type consists of sets of 16S sequencing data from real communities sampled from human Stool microbiomes.
+  The second are datasets synthetically or artificially created for the specific goal of helping evaluate computational analysis tools.
+  In particular, in order to objectively compare, to the extent possible, how well each step in \ac{micone} best captures the underlying data, we use both mock data (labelled mock4, mock12 and mock16) from mockrobiota~\cite{Bokulich2016} as well as, synthetically generated networks using the NorTA~\cite{Kurtz2015} and seqtime~\cite{Rottjers2018}.
   Detailed information on the mock communities and the settings used to generate the synthetic data are provided in the Methods section.
@@
  From previous studies, (refer to the Bokulich2018 paper here) which have done more quantitative benchmarks on the subject of which tool is the best (for classifer)...
- There are also other studies, [ref] that focus on the database aspect and they too have found that...
+ There are also other studies, [ref, OTT paper] that focus on the database aspect and they too have found that...
 

### 2022-04-11 sections/results.tex
Change: Update the Network Variability section of Results

@@
 
-  Our analysis involves two types of data: The first type consists of sets of 16S sequencing data from real communities sampled from human Stool microbiomes.
+  Our analysis involves two types of data: The first type consists of sets of 16S sequencing data from real communities sampled from human stool microbiomes from a fecal microbiome transplant study of autism~\cite{Kang2017}.
   The second are datasets synthetically or artificially created for the specific goal of helping evaluate computational analysis tools.
@@
 
-  In order to analyze the effect of different statistical methods on the inferred co-occurrence networks, we generated co-occurrence networks using all possible combinations of methods and estimated the variability in the networks due to each choice (Figure \ref{fig:figure1}).
-  This analysis is performed while keeping the network inference algorithm (NI step) the same throughout the analysis.
+  % Main takeaway -> TA is most important
+  % Talk about OP also
+  % What does PCA tell us?
+  % Conclusion
+
+  In order to analyze the effect of different statistical methods on the inferred co-occurrence networks, we generated co-occurrence networks using all possible combinations of methods and estimated the variability in the networks due to each choice (Figure \ref{fig:figure2}A).
+  This analysis and all subsequent analysis, unless explicitly mentioned, are peformed only on the control samples in the FMT study.
+  The networks are filtered on both association strength (0.1) and pvalue (0.05, only correlation methods) before the networks are subjected to the analyses.
   The effects of various steps on the final co-occurrence network is estimated by building a linear model of the edges of the network as a function the various step in the analysis pipeline (see Methods).
-  Figure \ref{fig:figure2}B, shows the fraction of total variation among the co-occurrence networks due to the first three steps of the pipeline.
-In other words, each point corresponds to a different combination of tools, and captures how much the final network is affected by such choice.
-  The 16S reference database contributes the most ($\sim25\%$) to variation in the networks.
-This is also reflected in the fact that the networks can be clearly separated based on the database used (Figure \ref{fig:figure2}B).
-  This indicates that the taxonomy assigned to the reference sequences drastically alters the co-occurrence network.
+  Figure \ref{fig:figure2}A, shows the percentage of total variation among the co-occurrence networks due to the different steps of the pipeline.
+  The TA step, more specifically, the choice of 16S reference database contributes the most ($\sim65.4\%$) to the variation in the networks, followed by the OP step ($\sim26.8\%$).
+  This result highlights the importance of the taxonomy assignment step in the 16S data analysis workflow, implying that a change in the reference database will result in drastically different inferred networks.
+  % TODO: Verify this
+  The reason for this seems to derive from the mislabelling of constitutive taxa that are present in high abundance in the community, which drastically alter the nodes and hence the underlying network topology.
   In fact the variability induced by taxonomy assignment is much more significant than that due to the variability induced based on how the reference sequences themselves are identified (in the DC step).
-  The grouping of the networks by taxonomy assignment into clusters (Figure~\ref{fig:figure2}B) seems to derive from the mislabelling of constitutive taxa that are present in high abundance in the community, which drastically alter the nodes and hence the underlying network topology.
-  The residual variation (Figure \ref{fig:figure2}A) can be seen as an artifact that arises when multiple steps are changed at the same time.
-  Another interesting observation (elaborated in detail in the denoising and clustering section) is that the dissimilarity between the networks decreases when the low abundance \ac{otu}s are removed from the network.
-  These results suggest that the most important criterion for accurate comparative analyses of co-occurrence networks is the taxonomy reference database.
+  The residual variation (Figure \ref{fig:figure2}A) can be seen as an artifact that arises when multiple steps are changed at the same time, however this is small.
+
+  Figure \ref{fig:figure2}B is the PCA plot of all the generated co-occurrence networks, colored by the databases in the TA step (left) and the network inference algorihtms in the NI step (right).
+  In other words, each point corresponds to a network generated using different combination of tools, and captures how much the final network is affected by such choice.
+  The importance of the TA step indicated in Figure \ref{fig:figure2}A is also reflected here, in the fact that most of the networks can be clearly separated based on the database used (Figure \ref{fig:figure2}B).
+  However, it is also evident that some of the networks, especially those inferred using the direct association based network infernce methods, are much closer in the PCA plot regardless of the reference database used.
+  % TODO: Figure out the how the OP step plays a role in the PCA plot
+  It can also be seen that the dissimilarity between the networks decreases when the low abundance \ac{otu}s are removed from the network.
+  These results suggest that the most important criterion for accurate comparative analyses of co-occurrence networks is the taxonomy reference database and the level of filtering of the taxonomy tables prior to network inference.
 

### 2022-04-12 sections/methods.tex
Change: Update the TA section of the Results

@@
   The output of this module is a matrix of counts, that describes the number of reads of a particular \ac{otu} or \ac{esv} (rows of the matrix) present in each sample (columns of the matrix).
-  The options currently available in the pipeline for denoising and clustering are: open reference clustering, closed reference clustering and de novo clustering methods from the vsearch plugin of \ac{qiime2} v2021.8.0~\cite{bolyenReproducibleInteractiveScalable2019} and denoising methods from \ac{dada2} v1.14~\cite{Callahan2016} and Deblur v1.1.0~\cite{Amir2017} (from the deblur plugin of \ac{qiime2}).
+  The options currently available in the pipeline for denoising and clustering are: open reference clustering, closed reference clustering and de novo clustering methods from the vsearch plugin of \ac{qiime2} v2021.8.0~\cite{bolyenReproducibleInteractiveScalable2019} and denoising methods from \ac{dada2} v1.14~\cite{Callahan2016} (from the \ac{dada2} R package) and Deblur v1.1.0~\cite{Amir2017} (from the deblur plugin of \ac{qiime2}).
   The quality filtering and chimera checking tools are derived from those used in \ac{qiime2} v2021.8.0 (uchime method) and \ac{dada2} (remove bimera method).

### 2022-04-12 sections/results.tex
Change: Update the TA section of the Results

@@
 
-  % Main takeaway -> TA is most important
-  % Talk about OP also
-  % What does PCA tell us?
-  % Conclusion
-
   In order to analyze the effect of different statistical methods on the inferred co-occurrence networks, we generated co-occurrence networks using all possible combinations of methods and estimated the variability in the networks due to each choice (Figure \ref{fig:figure2}A).
-  This analysis and all subsequent analysis, unless explicitly mentioned, are peformed only on the control samples in the FMT study.
+  This analysis and all subsequent analysis, unless explicitly mentioned, are performed on all the samples in the fecal microbiome transplant study.
   The networks are filtered on both association strength (0.1) and pvalue (0.05, only correlation methods) before the networks are subjected to the analyses.
@@
   % TODO: Figure out the how the OP step plays a role in the PCA plot
+  % TODO: Supplementary Figure S2
   It can also be seen that the dissimilarity between the networks decreases when the low abundance \ac{otu}s are removed from the network.
@@
 
-  Denoising and clustering are commonly carried out to generate representative sequences from the raw 16S sequencing data and to obtain the \ac{otu}/\ac{esv} tables (counts of these representative sequences for each sample).
-  In order to compare the \ac{otu} tables generated by different tools we processed the same 16S sequencing reads (healthy samples from a fecal microbiome transplant study~\cite{Kang2017}) using 5 different methods:  open-reference clustering, closed-reference clustering, denovo clustering, \ac{dada2}~\cite{Callahan2016} and Deblur~\cite{Amir2017}.
-  The first three methods are from the \ac{qiime1}~\cite{Caporaso2010} package.
-  We find that there is good agreement in the \ac{otu}/\ac{esv} tables when different combinations of methods are used to generate them (Supplementary Figure~\ref{fig:figureS1}).
+  Denoising and clustering are commonly carried out to generate representative sequences from the demultiplexed and trimmed 16S sequencing data to obtain the \ac{otu}/\ac{esv} tables (counts of these representative sequences for each sample).
+  In order to compare the representative sequence count tables generated by different tools we processed the same 16S sequencing reads (samples from a fecal microbiome transplant study~\cite{Kang2017}) using 5 different methods: open-reference clustering, closed-reference clustering, denovo clustering, \ac{dada2}~\cite{Callahan2016} and Deblur~\cite{Amir2017}.
+  The first three methods are from the vsearch plugin and Deblur is from the deblur plugin of \ac{qiime2} v2021.8.0~\cite{bolyenReproducibleInteractiveScalable2019}.
+  The \ac{dada2} method is from the \ac{dada2} R package.
 
   To compare the representative sequences generated by these methods we employ both the weighted~\cite{Lozupone2007} (Figure~\ref{fig:figure3}A) and unweighted UniFrac method~\cite{Lozupone2005} (Figure~\ref{fig:figure3}B).
+  In both the comparisons, we set a count threshold of 10, such that if the count of the representative sequences in a particular sample is less than the threshold, it is omitted from the analysis.
+  The values shown in the figures correspond to the mean UniFrac values across all samples in the study.
   The weighted UniFrac distance metric takes into account the counts of the representative sequences, whereas the unweighted UniFrac distance metric does not and hence gives equal weights to each sequence.
-  From Figure~\ref{fig:figure3}A one can see that the representative sequences generated by the different methods are similar to each other when weighted by their abundance.
-  Figure~\ref{fig:figure3}B on the other hand shows an increase in dissimilarity between each pair of methods suggesting that the methods might differ in the treatment of sequences of low abundance.
+  From Figure~\ref{fig:figure3}A one can see that the representative sequences generated by the different methods, with the exception of Deblur, are similar to each other when weighted by their abundance.
+  Figure~\ref{fig:figure3}B on the other hand shows an increase in dissimilarity between each pair of methods suggesting that the methods might differ in the assignment of sequences of low abundance.
+
+  % TODO: Supplementary work - Figure S3
   In order to verify this claim, for each of these methods we use the \ac{gg} taxonomy database to assign taxonomies to the representative sequences.
-  We then correlate the abundances of matching taxonomies between a pair of DC methods (Figure \ref{fig:figureS1}A and B).
+  We then correlate the abundances of matching taxonomies between every pairwise combination of DC methods (Figure \ref{fig:figureS1}A and B).
   The \ac{esv} tables generated by methods that perform denoising are very similar to each other ($\sim0.91$) and the \ac{otu} tables generated by the clustering methods are very similar to each other ($\sim0.9$), but results of denoising and clustering are highly uncorrelated with each other ($\sim0.4$) (Figure \ref{fig:figureS1}C).
 
+  % TODO: Update thresholds (10) on the mock datasets and check again
   These comparisons only elucidate the pairwise similarity or dissimilarity of a pair of methods.
-  In order to determine the tool that most accurately recapitulates the reference sequences in the samples, we used the 16S sequences from the mock datasets.
+  In order to determine the tool that most accurately recapitulates the reference sequences in the samples, we used the information about the known 16S sequences from the mock datasets.
   In particular, we used the pipeline to process mock community datasets using each of the possible methods included for this step.
-  We next compared predicted representative sequences with expected representative sequences and their distribution.
-  The results (Figure~\ref{fig:figure3}C and D) show that, for the mock datasets, the different methods perform similar to each other, exactly as observed in the case of the real dataset.
-However, the mock predicted sequence distributions are substantially different from the expected sequence distribution.
+  We next compared the predicted representative sequences with the expected representative sequences and their distribution.
+  The results (Figure~\ref{fig:figure3}C and D) show that, for the mock datasets, the predicted sequence distributions are substantially different from the expected sequence distribution.
+  This can be seen in the large values for both unweighted and weighted UniFrac distance values when compared to Figure~\ref{fig:figure3}A and B.
   This result is more exaggerated in the case of the unweighted UniFrac metric, where some of the datasets show a very high deviation from the expected sequences.
-  These high deviations are primarily in two of the three datasets that were analyzed and show that the datasets themselves play a big role in the performance of these methods.
-  This can be clearly seen in the performance (weighted UniFrac distance) of \ac{dada2} and Deblur on mock12 and mock16 datasets, where, Deblur outperforms \ac{dada2} on mock12 but the under-performs on mock16.
-
+  These high deviations are primarily in one or two of the three datasets that were analyzed and show that the datasets themselves play a big role in the performance of these methods.
+  Open-reference and denovo clustering methods perform the best under the weighted UniFrac metric and the worst under the unweighted UniFrac metric.
+  This result can be attributed to the large number of low abundance representative sequences that are generated by these methods.
+  Deblur performs poorly under weighted Unifrac and although its performance on the mock4 dataset is the best under unweighted UniFrac, its performance on the other datasets is average.
   There is no method that clearly outperforms the rest in all datasets.
   Based on their slightly better performance on the mock datasets, their \textit{de novo} error correcting nature and other previous studies~\cite{Nearing2018}, \ac{dada2} and Deblur seem to be in general the most reliable.
-  Given the unexpected poor performance of Deblur on the synthetic data, the default algorithm in the pipeline was chosen to be \ac{dada2} (Supplementary Figure~\ref{fig:figureS3}).
+  Overall, since \ac{dada2} as compared to Deblur, displays better performance on all the mock datasets and under the weighted UniFrac metric, we have decided to recommend this tool for the DC step of the pipeline.
 
-  % TODO: This is the most important statement (look at the dada2 paper for better arguments)
-  We prefer dada2 because it peforms marginally better (value??) than the other methods and it is also performs independent of the taxonomy database (so does denovo why are you suggesting this then?).
-  Denovo also performs better but the denoising helps when there are more sequences (where is the proof?) [ref]
+  % TODO: Supplementary work - Figure S4
+  After the denoising, the sequences are subject to chimera checking.
+  The \ac{micone} pipeline supports two different chimera checking methods, uchime and remove bimera.
+  The uchime method is from \ac{qiime2}, whereas the remove bimera method is from the \ac{dada2} R package.
+  We did not notice any significant difference between the two chimera checking methods, implying that they identify and remove mostly the same set of sequences as chimeras (Figure ???).
 
@@
   \subsection*{Taxonomy databases vary widely in taxonomy hierarchy and update frequency}
+
+  % How are the top assignments when compared
+  % How does these assignments vary across level
+  % What does this mean for us, conclusions
+  % What does the mock data show?
+  % Is there a best tool/database? What would you choose as the default?
 
@@
   The \ac{ncbi} RefSeq nucleotide database contains 16S rRNA sequences as a part of two BioProjects - 33175 and 33317.
-  The three databases vastly differ in terms of their last update status - \ac{gg} was last updated on May 2013, SILVA was last updated on December 2017 at the time of writing and \ac{ncbi} is updated as new sequences are curated.
+  The three databases vastly differ in terms of their last update status - \ac{gg} was last updated on May 2013, SILVA was last updated on August 2020 at the time of writing and \ac{ncbi} is updated regularly as new sequences are curated.
   Since updates to taxonomic classifications are frequent, these databases vary significantly in terms of taxonomy hierarchies including species names and phylogenetic relationships~\cite{Balvociute2017}.
+  Assignment using a particular database also requires a query tool, we use the NaiveBayes classifier from \ac{qiime2} for the \ac{gg} and SILVA databases and the blast tool (also included as a \ac{qiime2} plugin) for the \ac{ncbi} database.
+  These tools have been well quantified and optimized~\cite{bokulichOptimizingTaxonomicClassification2018}, hence, we make use of the default parameters in our analyses.
 
-  The representative sequences obtained from the \ac{dada2} method in the DC step were used for taxonomic assignment using the three reference databases.
+  The representative sequences obtained from the \ac{dada2} method (with remove bimera for chimera checking) in the DC step were used for taxonomic assignment using the three reference databases.
   Figure~\ref{fig:figure4}A depicts a flow diagram that shows how the top 50 representative sequences (sorted by abundance) are assigned a Genus according to the three different databases.
-  % TOOD: Update with a more quantitative description
-  We observe that not only does the assigned Genus composition vary significantly, but the percentage of unassigned representative sequences (gray) also differ.
-  Even the most abundant representative sequence is assigned to an "unknown" Genus in two of the three databases.
+  We observe that the top 50 representative sequences are assigned to 17 Genus in \ac{gg}, 23 in SILVA and 17 in \ac{ncbi}, however the sequences that are assigned to a particular Genus vary significantly based on the taxonomy database.
+  % TODO: Annotate Figure 4A with bacteria names
+  We observe that not only does the assigned Genus composition vary significantly, but the percentage of unassigned representative sequences (pink) also differ.
+  Even the most abundant representative sequence is assigned to an "unknown" Genus in two of the three databases (\ac{ncbi} and \ac{gg}).
   A representative sequence might be assigned an "unknown" Genus for one of two reasons: the first is if the taxonomy identifier associated with the sequence in the database did not contain a Genus; the second (more likely) reason is that the database contains multiple sequences that are very similar to the query (representative) sequence and the consensus algorithm (from \ac{qiime2}) is unable to assign one particular Genus at the required confidence.
-  After assigning all the representative sequences to taxonomies we perform a pairwise comparison of the similarity between assignments from different databases at every taxonomic level (Figure~\ref{fig:figure4}B).
-  The assignments beyond Family level (Family, Genus and Species) are very dissimilar with $<70\%$ similarity between any pair of databases.
-  There are no two reference databases that are more similar than the other pairs, with \ac{gg} and SILVA producing only marginally similar assignments compared to \ac{ncbi}.
-  This implies that the taxonomy assignments from each reference database are fairly unique and are largely responsible for the differences observed in the co-occurrence networks generated from different taxonomy databases.
-
-  Supplementary Figure~\ref{fig:figureS4} shows that the top 20 most abundant genera in the three resulting taxonomy composition tables are different.
+  We also observe that other abundant Genus like Bacteroides and Escherichia are also assigned to different representative sequences.
   For example, the most abundant genus in the \ac{gg} taxonomy table was \textit{Escherichia} whereas in the SILVA taxonomy table it was \textit{Escherichia-Shigella}.
   Although these are minor differences, when comparing a large number of taxonomy composition tables these problems are hard to diagnose.
-%   The comparison of all assigned genera instead of the just the top 20 contains the same percentage of matches and mismatches, implying that there does not seem to exist a correlation between abundance and mismatch.
-%   This suggests that the most abundant sequences are not necessarily the ones that are consistently matched to the same taxonomies in the different reference databases.
 
-  As in the previous section, these comparisons only indicate similarity or dissimilarity between methods.
+  After assigning all the representative sequences to taxonomies we perform a pairwise comparison of the similarity between the top 100 assignments (by abundance) from different databases at every taxonomic level (Figure~\ref{fig:figure4}B).
+  The assignments beyond Family level (Family, Genus and Species) are very dissimilar with $<67\%$ similarity between any pair of databases.
+  % TODO: Verify this
+  The assignment between SILVA and the other two databases is significantly different ($40\%$ mismatch) even at the Phylum level, indicating that 40 of the top 100 assignments at the Phylum level in the SILVA database are different in SILVA.
+  The \ac{gg} and \ac{ncbi} assignments are more similar compared to the other pairwise comparisons, mainly due to the poor performance of SIVLA at the Phylum level.
+  However, the number of mismatches for taxonomy assignments beyond the Family level are very high ($>66\%$), implying that the taxonomy assignments from each reference database are fairly unique and are largely responsible for the differences observed in the co-occurrence networks generated from different taxonomy databases.
+
+  % TODO: Supplementary work - Figure S5
+  The comparison of all assigned genera instead of the just the top 20 contains the same percentage of matches and mismatches, implying that there does not seem to exist a correlation between abundance and mismatch.
+  This suggests that the most abundant sequences are not necessarily the ones that are consistently matched to the same taxonomies in the different reference databases.
+
+  As in the previous section, the comparisons so far only indicate similarity or dissimilarity between methods.
   In order to obtain an absolute measure of accuracy of the taxonomic assignments we use the expected reference sequences from the mock datasets as the query sequences for the databases and the expected taxonomic composition as the standard to compare against (Figure~\ref{fig:figure4}C).
-  Again, we observe that none of the databases perform better than the others in absolute terms.
+  Here, we observe that none of the databases perform better than the others in absolute terms and the dissimilarity is high, indicating that all the databases perform poorly when trying to recapture the expected taxonomic composition.
 
-  Given that no database performs better than others against mock datasets, and that databases are almost equally distant from each other in terms of final output, the choice of which database to use should be driven by other reasons.
+  Given that no database performs better than others against mock datasets, the choice of which database to use should be driven by other reasons.
   One user-specific way to choose, would be based on the known representation of taxa for the microbiome of interest (see also Discussion).
-Another reason could be the frequency of updates and the potential for future growth, which prompted us to set \ac{ncbi} as the \ac{micone} standard for taxonomy assignment.
- In addition to being regularly maintained and updated the \ac{ncbi} database already has the advantage that its accuracy of assignments is still comparable to the SILVA and \ac{gg} reference databases that are routinely used as reference databases.
+Another reason could be the frequency of updates and the potential for future growth.
+  % TODO: Find reference for this
+  Since the \ac{gg} database is more widely used in the field~[ref], to enable easier comparison with published networks in literature, we set \ac{gg} as the \ac{micone} standard for taxonomy assignment.
+  On the other hand, since the degree of similarity of assignments between \ac{ncbi} and \ac{gg} is fairly high, and due to its regular updation and maintenance the \ac{ncbi} database is also a good choice for the user.
 
- % TODO: This is the most important statement here
- From previous studies, (refer to the Bokulich2018 paper here) which have done more quantitative benchmarks on the subject of which tool is the best (for classifer)...
- There are also other studies, [ref, OTT paper] that focus on the database aspect and they too have found that...
+  % TODO: More refs to support our argument?
+ % There are also other studies, [ref, OTT paper] that focus on the database aspect and they too have found that...
 

### 2022-04-12 scripts/figure3/figure3_data.py

@@ -26,12 +26,12 @@ def get_vectors(otu_1: pd.DataFrame, otu_2: pd.DataFrame, threshold: int):
     common_samples = samples_1 & samples_2
     for col in common_samples:
         otu_1_col, otu_2_col = otu_1[[col]], otu_2[[col]]
-        otu_1_col = otu_1_col[otu_1_col[col] > threshold]
-        otu_2_col = otu_2_col[otu_2_col[col] > threshold]
+        otu_1_col.loc[otu_1_col[col] <= threshold, :] = 0
+        otu_2_col.loc[otu_2_col[col] <= threshold, :] = 0
         otu_1_col.columns = ["1"]
         otu_2_col.columns = ["2"]
         joint_df = otu_1_col.join(otu_2_col, how="outer")
-        joint_df.fillna(0.0, inplace=True)
+        joint_df.fillna(0 inplace=True)
         otu_ids = list(joint_df.index)
         u = list(joint_df["1"])
         v = list(joint_df["2"])

### 2022-04-12 sections/results.tex
Change: Update the NI section of the Results

@@
 
-  % How are the top assignments when compared
-  % How does these assignments vary across level
-  % What does this mean for us, conclusions
-  % What does the mock data show?
-  % Is there a best tool/database? What would you choose as the default?
-
   Taxonomy databases are used to assign taxonomic identities to the representative sequences obtained after the DC step.
@@
 
-  The six different network inference methods used in this study are \ac{magma}~\cite{Cougoul2019}, \ac{mldm}~\cite{Yang2017}, \ac{spieceasi}~\cite{Kurtz2015}, \ac{sparcc}~\cite{Friedman2012}, Spearman and Pearson.
-  These network inference methods fall into two groups, the first set of methods (Pearson, Spearman, \ac{sparcc}) infer pairwise correlations while the second set infer direct associations (\ac{spieceasi}, \ac{mldm}, \ac{magma}).
-  Pairwise correlation methods involve calculating the correlation coefficient between every pair of \ac{otu}/\ac{esv}s leading to the detection of spurious indirect connections.
-  On the other hand, direct association methods use conditional independence to avoid the detection of correlated but indirectly connected \ac{otu}s~\cite{Kurtz2015,Menon2018}.
+  The ten different network inference methods used in this study are: \ac{spieceasi}~\cite{Kurtz2015}, FlashWeave.jl~\cite{tackmannRapidInferenceDirect2019}, \ac{cozine}~\cite{haCompositionalZeroinflatedNetwork2020a}, \ac{harmonies}~\cite{jiangHARMONIESHybridApproach2020}, \ac{spring}~\cite{yoonMicrobialNetworksSPRING2019}, \ac{mldm}~\cite{Yang2017}, \ac{sparcc}~\cite{Friedman2012,Watts2018}, propr~\cite{quinnProprRpackageIdentifying2017}, Spearman and Pearson.
+  These network inference methods fall into two groups, the first set of methods (Pearson, Spearman, \ac{sparcc}, propr) infer pairwise correlations while the second set infer direct associations (\ac{spieceasi}, FlashWeave, \ac{cozine}, \ac{harmonies}, \ac{spring}, \ac{mldm}).
+  Pairwise correlation methods involve calculation of the correlation coefficient between every pair of node (taxonomic entity like Genus) leading to the detection of spurious indirect connections.
+  On the other hand, direct association methods use conditional independence to avoid the detection of correlated but indirectly connected taxonomic entities~\cite{Kurtz2015,Menon2018}.
 
-  For the analysis presented in this section, we used the taxonomy counts table obtained using the \ac{ncbi} reference database as the input for algorithms that infer co-occurrence associations between the microbes.
-  The taxonomy count tables are first collapsed to the Genus level (or the required taxonomy level) by summing up the counts of the lower taxonomy levels that map to the same higher taxonomy level name.
+  For the analysis presented in this section, we used the filtered (filtering during the OP step) Genus level counts table obtained using the \ac{gg} (dada2 and remove bimera methods in the DC step) reference database from the TA step as the input for the network inference algorithms.
+  The Genus count tables were obtained by summing up the counts of the lower taxonomy levels (Species and \ac{otu}) that map to the same higher taxonomy level name.
   Figure~\ref{fig:figure5}A shows the networks inferred from this dataset using the different inference algorithms.
-  The different networks differ vastly in their edge-density and connectivity; even some of the edges in common to these networks have their signs inverted.
-Note, however, that some of these comparisons depend on the threshold that has to be applied to the pairwise correlations methods (currently 0.3, based on~\cite{Friedman2012}).
-  To get a more quantitative picture of the differences between the inferred networks, we checked the distribution of common nodes and edges (Figure~\ref{fig:figure5}B) using UpSet plots~\cite{Lex} (only \ac{magma}, \ac{mldm}, \ac{spieceasi}, \ac{sparcc} are used in the comparison since Pearson and Spearman add a large number of spurious edges since they are not intended for compositional datasets).
-  The results for the node intersections show that the networks have a large number of nodes in common ($63$ out of $67$ nodes in the smallest network - \ac{magma}) and no network possesses any unique node.
-  The edge intersections in contrast show that only $19$ edges (out of $98$ edges in the smallest network - \ac{magma}) are in common between all the methods and each network has a large number of unique edges.
+  The different networks differ vastly in their edge-density and connectivity; even some of the edges in common to these networks can have their signs inverted.
+  Note, however, that some of these comparisons depend on the threshold that has to be applied to the networks (currently 0.3 for the correlation based methods, based on~\cite{Friedman2012} and 0.01 for the direct association methods).
+  \ac{mldm} is not shown in the comparisons because the algorithm failed to converge for most of the network combinations, hence, it is only used in the network variation analysis.
+
+  % TODO: Update after getting SparCC results
+  In order to obtain a quantitative picture of the differences between the inferred networks, we analyzed the distribution of common nodes and edges (Figure~\ref{fig:figure5}B) using UpSet plots~\cite{Lex}.
+  Only \ac{harmonies}, \ac{cozine},\ac{spieceasi}, \ac{spring}, propr, and \ac{sparcc} are used in the comparison since Pearson and Spearman add a large number of spurious edges since they are not intended for compositional datasets).
+  The results for the node intersections show that the networks have a large number of nodes in common ($33$ out of $68$ total nodes) and no network possesses any unique node.
+  The edge intersections in contrast show that only $8$ edges (out of $202$ total edges) are in common between all the methods and each network has a large number of unique edges.
   These results indicate that there is a substantial rewiring of connections in the inferred networks.
 
-  Unlike the previous steps of the pipeline, where were we evaluated the performance of methods on mock datasets, there is no equivalent dataset that contain a set of known interactions for the evaluation of the network inference algorithms.
-  Therefore, we propose the construction of a consensus network (Figure~\ref{fig:figure5}C) involving \ac{magma}, \ac{mldm}, \ac{spieceasi} and \ac{sparcc}.
+  % TODO: What is the final concluding statement here? How to segway to the next section
+
+  \FloatBarrier
+
+  \subsubsection*{The consensus network}
+
+  Therefore, we propose the construction of a consensus network (Figure~\ref{fig:figure5}C) involving \ac{mldm}, \ac{spieceasi} and \ac{sparcc}.
 This consensus network is built by merging the p-values generated from bootstraps of the original taxonomy counts table using the Browns p-value combining method~\cite{brown_400_1975} (see Methods section).
+
+  % FIXME: We have synthetic data
+  Similar to the previous steps of the pipeline, where were we evaluated the performance of methods on mock datasets, we can evaluate the network inference algorithms using synthetic interaction data.
+  Specifically, we use the ...
+
   Note that we do not use Pearson and Spearman methods in the p-value merging step and these algorithms are only used for demonstration and comparison.
@@
   \subsubsection*{The default pipeline}
+  % TODO: Change the order of Figure 6 and 7
 
@@
 
-  The recommended tool for the \ac{dc} step (\ac{dada2} or Deblur) were chosen based on their accuracy in recapitulating the reference sequences in mock communities and synthetic data.
+  The recommended tool for the \ac{dc} step (\ac{dada2}) was chosen based on their accuracy in recapitulating the reference sequences in mock communities and synthetic data.
   The choice of the taxonomy reference database in the \ac{ta} step is dictated largely by the species expected to be present in the sample as well the database used in similar studies if comparison is a goal.
-  Nevertheless, we suggest \ac{ncbi} RefSeq along with blast+ as the query tool since the database is updated regularly and has a broad collection of taxonomies.
-  The abundance threshold at the \ac{op} step is determined automatically based on the number of samples and the required statistical power.
-  Finally, we use the Browns p-value combining method on the networks generated using \ac{magma}, \ac{mldm}, \ac{spieceasi} and \ac{sparcc} to obtain a final consensus network in the \ac{ni} step.
+  Nevertheless, we suggest \ac{gg} database along with the NaiveBayes classifier as the query tool since the database is used most often in taxonomic studies.
+  % TODO: Default OP step values and references
+  The abundance threshold at the \ac{op} step is set at ... because ....
+  Finally, we make use of the consensus methods developed in this paper ...
+on the networks generated using \ac{mldm}, \ac{spieceasi} and \ac{sparcc} to obtain a final consensus network in the \ac{ni} step.
 

### 2022-04-13 sections/methods.tex
Change: Update the consensus network section of the Results

@@
 
+  % TODO: Add reference: Karoline Faust, Franziska Bauchinger, Béatrice Laroche, Sophie de Buyl, Leo Lahti, Alex D Washburne, Didier Gonze, Stefanie Widder, "Signatures of ecological processes in microbial community time series", Microbiome (2018).
   The first method, hereby referred to as ``seqtime'', used generalized Lotka-Volterra (gLV) equations to model the microbial community dynamics and utilized the Klemm–Eguı́luz algorithm to generate a clique-based interaction network~\cite{Rottjers2018}.
@@
   The second method, hereby referred to as ``NorTA'', used the Normal to Anything (NorTA) approach coupled with a given interaction network topology to generate the abundance distribution of the microbial community~\cite{Kurtz2015}.
-  We used the \href{https://github.com/zdk123/SpiecEasi}{spieceasi} R package to simulate communities with different network topologies (scale-free, cluster, block, Erdos-Renyi, band and hub) and target abundance distributions (Negative Binomial, Poisson, Zero-Inflated Negative Binomial).
+  We used the \href{https://github.com/zdk123/SpiecEasi}{spieceasi} R package~\cite{Kurtz2015} to simulate communities with different network topologies (scale-free, cluster, block, Erdos-Renyi, band and hub) and target abundance distributions (Negative Binomial, Poisson, Zero-Inflated Negative Binomial).
   The OTU table was generated using the American Gut Project example in the \texttt{spieceasi} package (amgut1.filt) with the default parameter options.

### 2022-04-13 sections/results.tex
Change: Update the consensus network section of the Results

@@
   % TODO: Update after getting SparCC results
-  In order to obtain a quantitative picture of the differences between the inferred networks, we analyzed the distribution of common nodes and edges (Figure~\ref{fig:figure5}B) using UpSet plots~\cite{Lex}.
+  In order to obtain a quantitative picture of the differences between the inferred networks, we analyzed the distribution of common nodes and edges (Figure~\ref{fig:figure5} B and C) using UpSet plots~\cite{Lex}.
   Only \ac{harmonies}, \ac{cozine},\ac{spieceasi}, \ac{spring}, propr, and \ac{sparcc} are used in the comparison since Pearson and Spearman add a large number of spurious edges since they are not intended for compositional datasets).
@@
 
-  \subsubsection*{The consensus network}
+  \subsection*{The consensus network}
 
-  Therefore, we propose the construction of a consensus network (Figure~\ref{fig:figure5}C) involving \ac{mldm}, \ac{spieceasi} and \ac{sparcc}.
-This consensus network is built by merging the p-values generated from bootstraps of the original taxonomy counts table using the Browns p-value combining method~\cite{brown_400_1975} (see Methods section).
+  We observe a large number of rewirings in the inferred association networks generated by the different methods (Figure \ref{fig:figure5}C), implying that there is a lack of agreement on the associations that should be a part of the final network.
+  Therefore, we have developed a consensus method that would take into consideration the evidence offered by each network inference algorithm and generate a consensus network that would contain the most probable interactions for the input data.
+  % TODO: Should we include Pearson and Spearman in the consensus (currently we do)!
+  We have developed two algorithms for the generation of the consensus network, the simple voting and the scaled-sum methods (see Methods).
+  First, for the correlation-based methods we merge the p-values using the Brown's p-value merging method~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016,faustCoNetAppInference2016} (see Methods).
+  Second, we filter all the inferred networks based on an association strength threshold of 0.1 and a p-value cutoff of 0.05.
+  Finally, we apply the consensus algorithms we have developed on these filtered networks.
+  The simple voting method does a consensus based on whether the number of inferred networks that have an edge between a pair of nodes is greater than a threshold.
+  The scaled-sum method does a consensus based on whether the normalized sum of an edge between a pair of nodes from all inferred networks is greater than a threshold.
+  The default algorithm for the pipeline is the scaled-sum method, since this method takes into account the weights of the association in the determination of the final consensus.
+  Additionally, by default we include all the network inference methods in the consensus calculation by default, but a subset of methods can be considered if it is required.
+  Based on this approach, \ac{micone} reports as default output the consensus network, where is each edge is annotated with weights (correlations for the correlation-based methods and direct associations for the other methods) from all the methods used in the consensus algorithm.
 
-  % FIXME: We have synthetic data
-  Similar to the previous steps of the pipeline, where were we evaluated the performance of methods on mock datasets, we can evaluate the network inference algorithms using synthetic interaction data.
-  Specifically, we use the ...
+  Similar to the previous steps of the pipeline, where were we evaluated the performance of methods on mock datasets, we can evaluate the network inference algorithms and the final consensus network using synthetic interaction data.
+  For this purpose, we generated synthetic interaction data using the ``NorTA''[ref] and ``seqtime''[ref] methods (see Methods), which have been previously used for benchmarking network inference methods.
+  For each method, an \ac{otu} counts table is generated based on the selected interaction network topology and properties.
+  This counts table is used as the input to the \ac{micone} pipeline to generate predicted associations.
+  The interaction network used to generate the counts table is used as the source of true interactions to calculate the precision and sensitivity for each network inference algorithm for the NorTA datasets (Figure \ref{fig:figure6}).
+  Each point in the scatter plot represents a network with different input network topology and OTU abundance distributions.
+  Overall, we see that the scaled-sum method for a parameter value of 0.5 performs the best in terms of average precision ($0.985$).
+  Among the individual inference methods, we observe that \ac{spieceasi} has the best average precision ($0.944$)
+  The goal of the consensus network algorithms is to increase the precision of the resultant consensus network by removing edges that are either not present in most of the inference methods or whose association strength is low.
+  From Figure \ref{fig:figure6} we can confirm that the consensus algorithm, especially the scaled-sum method, does indeed produce networks with more precision, but as a consequence the sensitivity of the networks is reduced.
+  Therefore, the consensus networks provide the means to obtain a short list of associations that would have a high probability of being present in the real interaction network.
+  This small set of associations could be further tested and validated experimentally to verify their plausibility.
 
-  Note that we do not use Pearson and Spearman methods in the p-value merging step and these algorithms are only used for demonstration and comparison.
-  Based on this approach, \ac{micone} reports as default output the consensus network, annotated with weights (correlations for \ac{sparcc} and direct associations for the other methods) for all four methods.
+  % TODO: Supplementary work: Figure S6
+  The precision and sensitivity graphs on the seqtime data can be found in the Supplementary section.
 
@@
 
-  \subsubsection*{The default pipeline}
-  % TODO: Change the order of Figure 6 and 7
+  \subsection*{The default pipeline}
 

### 2022-04-13 sections/tables_figures.tex
Add: TODO about figure 3 captions
Change: Update the pipeline section of Results
Change: Update the Network Variability section of Results
Change: Update the consensus network section of the Results

@@
 
-  \begin{figure}[h]
+  \begin{figure}[H]
     \centering
-    \includegraphics[width=0.67\linewidth]{figure1.pdf}
+    \includegraphics[width=0.75\linewidth]{figure1.pdf}
+  \end{figure}
+  \begin{figure}[H]
+    \centering
     \caption{
@@
   \end{figure}
+
   \FloatBarrier
@@
     \centering
-    \includegraphics[width=0.85\linewidth]{figure2.pdf}
+    \includegraphics[width=1.0\linewidth]{figure2.pdf}
   \end{figure}
@@
 
+  % TODO: Describe the datasets being shown here clearly. In the caption + figure legends (if necessary) and in the text
   \begin{figure}
@@
     \caption{
-      \textbf{Network inference and taxonomic assignment have the highest influence on the inferred network structures.}
-      \textbf{(A)} The network constructed using the default pipeline parameters (DC=\ac{dada2}, TA=\ac{ncbi}, OP=on, NI=\ac{sparcc}) is compared with networks generated when one of the steps use a different tool.
-      The common connections (common with the default network) are in black, connections unique to the network are colored purple and connections in the default network but not present in the current network are gray.
-      \textbf{(B)} The L1 distance between the networks generated by changing one step of the default pipeline and the network generated using the default parameters.
+      The precision and sensitivity of the inferred networks on the synthetic data.
     }
@@
     \caption{
-      The consensus networks generated using the default pipeline settings.
-      \textbf{(A)} Co-occurrence network of the Hard Palate microbiome generated from samples of healthy subjects in a periodontal diseases study.
-      \textbf{(B)} Co-occurrence network of the Stool microbiome generated from samples of healthy subjects in a fecal microbiome transplant study.
+      \textbf{Network inference and taxonomic assignment have the highest influence on the inferred network structures.}
+      \textbf{(A)} The network constructed using the default pipeline parameters (DC=\ac{dada2}, TA=\ac{ncbi}, OP=on, NI=\ac{sparcc}) is compared with networks generated when one of the steps use a different tool.
+      The common connections (common with the default network) are in black, connections unique to the network are colored purple and connections in the default network but not present in the current network are gray.
+      \textbf{(B)} The L1 distance between the networks generated by changing one step of the default pipeline and the network generated using the default parameters.
     }

### 2022-04-13 sections/results.tex
Change: Update default pipeline section of Results

@@
   The systematic analyses performed in the previous sections clearly show that the choice of tools and parameters can have a big impact on the final co-occurrence network.
-For some of these choices (e.g.
-\ac{dada2} vs.
-deblur) there is no clear metric to establish a best protocol.
-  For other choices, the mock communities provide an opportunity to select combination of parameters that yield more accurate and robust results.
+  For some of these choices (e.g. \ac{dada2} vs. Deblur) there is no clear metric to establish a best protocol.
+  For other choices, the mock communities and synthetic data provide an opportunity to select combination of tools that yield more accurate and robust results.
   Despite this partial degree of assessment, we wish to suggest a combination of tools and parameters that produce networks that are derived from the combination of tools which performed best on the mock communities, and displayed highest robustness to switching to alternative methods.
@@
 
-  The recommended tool for the \ac{dc} step (\ac{dada2}) was chosen based on their accuracy in recapitulating the reference sequences in mock communities and synthetic data.
-  The choice of the taxonomy reference database in the \ac{ta} step is dictated largely by the species expected to be present in the sample as well the database used in similar studies if comparison is a goal.
+  The recommended tool for the \ac{dc} step (\ac{dada2}) was chosen based on its accuracy in recapitulating the reference sequences in mock communities.
+  For the chimera checking module of the \ac{dc} step, both uchime and remove bimera perform similarly, we suggest using remove bimera since it was developed specifically for \ac{dada2} and should provide the best performance for that denoising method.
+  The choice of the taxonomy reference database in the \ac{ta} step is dictated largely by the species expected to be present in the sample as well the database used in similar studies, if comparison is a goal.
   Nevertheless, we suggest \ac{gg} database along with the NaiveBayes classifier as the query tool since the database is used most often in taxonomic studies.
@@
   The abundance threshold at the \ac{op} step is set at ... because ....
-  Finally, we make use of the consensus methods developed in this paper ...
-on the networks generated using \ac{mldm}, \ac{spieceasi} and \ac{sparcc} to obtain a final consensus network in the \ac{ni} step.
+  For the \ac{op} step, we recommend that the filtering be turned on and used at the default settings, unless a taxonomy entity of interest might be filtered out at the default settings.
+  This is necessary since a large number of rows in the counts matrix would reduce the accuracy of the inference methods and increase the computational time.
+  Finally for the \ac{ni}, we recommend the scaled-sum consensus method along with p-value merging (for the correlation-based methods) to generate a final consensus network for the input data.
 
-  Figure~\ref{fig:figure6}A shows the default network compared against networks generated by altering one of the steps of the pipeline from the default.
-  These results indicate that the biggest differences in networks occur when the reference database or the network inference algorithm are changed.
-  Furthermore, the L1 distance of networks generated by altering one of the steps of the pipeline from the default against the default network (Figure~\ref{fig:figure6}B) shows that the biggest deviations from the default network occur when the \ac{ta} and \ac{ni} steps are changed, reinforcing the same results observed in Figure~\ref{fig:figure2}.
-Figure~\ref{fig:figure7} shows the co-occurrence networks inferred for the hard palate for healthy subjects in a periodontal disease study~\cite{Chen2018} and the healthy stool microbiome in fecal microbial transplant study~\cite{Kang2017}.
+  Figure~\ref{fig:figure7}A shows the default network compared against networks generated by altering one of the steps of the pipeline from the default.
+  These results indicate that the biggest rewiring in networks occur when the reference database or the network inference algorithm are changed.
+  % TODO: Show proof??
+  The filtering in the \ac{OP} step when turned off also adds a large number of new nodes, but the edges between the common nodes (with the default network) are fairly conserved.
+  % TODO: Finalize this figure
+  Furthermore, the L1 distance of networks generated by altering one of the steps of the pipeline from the default against the default network (Figure~\ref{fig:figure7}B) shows that the biggest deviations from the default network occur when the \ac{ta} and \ac{ni} steps are changed, reinforcing the same results observed in Figure~\ref{fig:figure2}.
+  Figure~\ref{fig:figure7} shows the co-occurrence networks inferred for the healthy subjects (control) and autistic subjects in the fecal microbiome transplant study~\cite{Kang2017}.
   These consensus networks were generated using the default tools and parameters from Table~\ref{tab:default_options}.

### 2022-04-14 sections/discussion.tex
Change: Update the discussion section

@@
 
-% General statements
+  \subsection*{Why \ac{micone}?}
+
   Co-occurrence associations in microbial communities help identify important interactions that drive microbial community structure and organization.
@@
   Importantly, in addition to revisiting the test cases presented in this work, users will be able to explore the effect of various tool combinations on their own datasets of interest.
-The \ac{micone} pipeline is built in a modular fashion.
-Its plug-and-play architecture will make it possible for users to add new tools and steps, either from existing packages, or from packages that were not examined in the present work, as well as future ones.
+  The \ac{micone} pipeline is built in a modular fashion; its plug-and-play architecture will make it possible for users to add new tools and steps, either from existing packages, or from packages that were not examined in the present work, as well as those developed in the future.
+  The easy-to-use nature, the configurability and modular nature of the \ac{micone} package enables users to start and end the pipeline at any point in the worflow, enabling users to run parts of the pipeline in isolation.
+  The \ac{micone} Python package provides functions and methods to perform detailed analysis of the count matrices and the inferred networks.
+  % TODO: Ref and supplementary
+  Additionally, the inferred networks by default are created using a custom JSON format (see Supplementary), but are interchangeable to Cytoscape[ref], GML[ref], and many other popular formats via the \ac{micone} Python package.
 
-The main outcome of this work is thus two-fold: on one hand we transparently reveal the dependence of co-occurrence networks on tool and parameter choices, making it possible to more rigorously assess and compare existing networks.
+  % TODO: Refs
+  There have been several tools/workflows such as \ac{qiime2}~\cite{bolyenReproducibleInteractiveScalable2019} and NetCoMi[ref] that have been developed for analysis of 16S sequencing data and to infer networks from these datasets.
+  However, there has been no single tool that completely integrates the process of inferring microbial interaction networks from 16S sequencing reads.
+  \ac{micone} is unique as it offers this functionality packaged in a simple workflow that can be run locally, on the compute cluster or the cloud.
+  For an end user who is interested in identifying the microbial interactions present in a specific sequencing dataset, \ac{micone} offers a fast and reproducible way to compare and test novel hypotheses related to interactions.
+
+  \subsection*{The default pipeline and suggested tools}
+
+  The main outcome of this work is two-fold: on one hand we transparently reveal the dependence of co-occurrence networks on tool and parameter choices, and built a tool for this purpose, enabling to more rigorously assess and compare existing networks.
   On the other hand, we take advantage of our spectrum of computational options and the availability of mock and synthetic datasets, to suggest a default standard setting, and a consensus approach, likely to yield networks that are robust across multiple tool/parameter choices.
 
-An important caveat related to this last point is the fact that our conclusions are based on the specific datasets used in our analysis.
-While our datasets cover a relatively broad spectrum of biomes and sequencing pipelines, datasets that have drastically different distributions may require a re-assessment of the best settings through our pipeline.
+  An important caveat related to this last point is the fact that due to the lack of a universal standard for microbial interaction data, our conclusions are based on the specific datasets used in our analysis.
+  While our datasets cover several mock and synthetic datasets that cover a fairly diverse range of abundance distributions and network topologies, datasets that have drastically different distributions may require a re-assessment of the best settings through our pipeline.
 
   It is worth pointing out some additional more specific conclusions stemming from the individual steps of our analysis.
-
   The different denoising/clustering methods differ mostly in their identification of sequences that are in low abundances.
@@
   However, comparison of inferred and expected reference sequences and their abundances in mock community datasets has allowed us to identify \ac{dada2} as the method which best recapitulates the expected sequence composition.
+  For the chimera checking module, we suggest using the remove bimera method since it was developed specifically for \ac{dada2} and performs equally as well as uchime.
   For the current work we have decided to focus on the tools most widely used at the time of the analysis.
-Some tools that we recently published (e.g.
+  Some tools there were not as widely used (e.g.
 dbOTU3~\cite{Olesen2017}) as well as older popular methods like mothur~\cite{Schloss2009} have not been included in the study, but could be added into the pipelines in future updated analyses.
 
-The choice of taxonomy database was found to be the most important factor in the inference of a microbial co-occurrence network, contributing $\sim20\%$ of the total variance.
+  The choice of taxonomy database was found to be the most important factor in the inference of a microbial co-occurrence network, contributing $\sim65.4\%$ of the total variance.
   The frequent changes in the taxonomy nomenclature coupled with the frequency of updates to the various 16S reference databases create inherent differences \cite{Balvociute2017} in taxonomy hierarchies in these databases.
@@
 We suggest that that choice of the database should be made based on possible reported or inferred biases in the representation of given biomes in a specific databases \cite{Balvociute2017}.
-The default reference database in the pipeline is the \ac{ncbi} 16S RefSeq database as it is more frequently updated and is most compatible with the blast+ query tool.
+  The default reference database in the pipeline is the \ac{gg} reference database along with the NaiveBayes classifier as the query tool.
+  The reason for our choice stem from the popularity of the \ac{gg} database in taxonomic studies, which would enable easy comparision across datasets.
   We also enable users to use custom databases \cite{Ritari2015} with the blast+ and naive bayes classifiers that are incorporated into the pipeline (from \ac{qiime2}).
 
-Filtering out taxa that are present in low abundances in all samples did not increase (in most datasets tested) the proportion of taxa in common between taxonomy tables generated using different reference databases.
+  The \ac{op} step of the pipeline the step that second in its contribution to total network variance.
+  This can be attributed to the large number of nodes that are added to the final networks when filtering is turned off.
+  % TODO: Refs
+  Additionally, the large number of nodes also greatly decrease the accuracy of the network inference algorithms [ref] as well as increase the computational complexity.
+  % TODO: Supplementary work - Figure S5
+  We observe that filtering out taxa that are present in low abundances in all samples did not necessarily increase the proportion of taxa in common between taxonomy tables generated using different reference databases.
+  % TODO: Supplementary work - Figure S2
   However, we do observe that the reduction in the number of taxa leads to better agreement in the networks inferred through different methods.
@@
   One reason for this is the underlying assumptions regarding sparsity, distribution and compositionality that the algorithms make.
-The consensus network created by merging the networks inferred using the different network inference methods enables the creation of a network whose links have evidence based on multiple inference algorithms.
-
-% \hl{Other factors that play an important role in network inference} \\
-% Differences in environments of the samples could lead to the inference of spurious interactions [ref].
+  The consensus network created by merging the networks inferred from the different network inference methods enables the creation of a network whose links have evidence based on multiple inference algorithms.
+  We have developed two consensus algorithms to support this effort, the simple voting method and the scaled-sum method.
+  Ultimately, we find that the scaled-sum method along with pvalue merging (applied to the correlation-based methods) performs the best on synthetic datasets, therefore this is the default for the \ac{ni} step of the pipeline.
 
@@
   The pipeline facilitates this comparative exploration by providing a variety of modules for inter-conversion between various formats, and by allowing easy incorporation of new tools as modules.
-
+  Since there exists no ``best method'' for the various steps of 16S data analysis, especially for microbial network inference, \ac{micone} is intended to help researchers to identify the methods and algorithms that are most suitable for their datasets in a easy-to-use and reproducible manner.
   We envision that \ac{micone}, and the underlying tools and databases that help process amplicon sequencing data into co-occurrence networks, will be increasingly useful towards building large comparative analyses across studies.
@@
 
-% TODO: Uses in MIND (?)
+  \subsection*{Future directions}
+
+  The current analyses include one fecal microbiome transplant dataset with healthy and autistic samples, three mock community datasets and several datasets generated by two synthetic interaction methods.
+  The analyses presented in this study are specific to the datasets utilized, and due to the lack of a gold standard reference dataset for microbial interaction data, it is not possible to generalize these results to datasets with widely different microbiome profiles.
+  Incorporating datasets from a broad spectrum of biomes with varying microbial distributions into \ac{micone} will likely increase the robustness and generalizability of the results from these analyses.
+  Furthermore, expanding the scope of the \ac{micone} pipeline to also include metagenomics sequencing data will enable an even broader access to a diverse range of datasets as well as domains of life that can be incorporated into the analysis.
+
+  Although the pipeline currently supports changing of parameter values for the most commonly used parameters for several tools, we only use the those default parameter values that are recommended by the tool creators.
+  The exploration of the effects of parameter values on the inferred association networks might help optimize these values for better network inference, but these are not part of this current study.
+  Additionally, despite incorporating a wide range of tools and algorithsm into the \ac{micone} pipeline, there are several other tools that were omitted due time and computational constraints.
+  However, the modular nature of the \ac{micone} pipeline enables easy addition of new tools and methods at any step of the workflow.
+
+  % TODO: Should we move the last paragraph of the previous section here?
+
+  % TODO: Health and disease
+  In summary, we believe that \ac{micone} is a useful tool for the exploration of microbial networks generated from 16S data, allowing for rapid, configurable and reproducible inference of microbial networks and further formulation of hypotheses about the role of these interactions on community composition and stability.

### 2022-04-14 sections/results.tex
Change: Update the discussion section

@@
   % TODO: Default OP step values and references
-  The abundance threshold at the \ac{op} step is set at ... because ....
+  The abundance threshold at the \ac{op} step is set at ... because large number of nodes are added to networks and decrease the performance of network inference algorithms
   For the \ac{op} step, we recommend that the filtering be turned on and used at the default settings, unless a taxonomy entity of interest might be filtered out at the default settings.

### 2022-04-19 sections/methods.tex
Change: Update the discussion section
Change: Move tables to the end of the manuscript

@@
 
-  \subsection*{16S rRNA sequencing Datasets}
+  \subsection*{16S rRNA sequencing datasets}
 
@@
 
+  % TODO: Write about the JSON network format (supplementary)
   The \ac{micone} pipeline provides both a Python API as well as a command-line interface and only uses a single configuration file (nextflow.config) to contain the configuration parameters.
+  The \ac{micone} Python API several biom file and network related function and methods, enabling detailed comparison of counts tables and inferred networks if desired.
   The configuration file along with the run file (main.nf) lists the inputs, output and the steps to be performed during runtime, along with the parameters to be used (if different from defaults) for the various steps.
@@
 
-  \begin{table}[h]
-    \centering
-    \small
-    \begin{tabular}{|c|c|c|}
-      \hline
-      \textbf{Module} & \textbf{Tool} & \textbf{References} \\
-      \hline
-      Denoising and Clustering & Closed reference & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
-      Denoising and Clustering & Open reference & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
-      Denoising and Clustering & De novo & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
-      Denoising and Clustering & Dada2 & \cite{Callahan2016} \\
-      Denoising and Clustering & Deblur & \cite{Amir2017,bolyenReproducibleInteractiveScalable2019} \\
-      \hline
-      Chimera checking & Uchime & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
-      Chimera checking & Remove bimera & \cite{Callahan2016} \\
-      \hline
-    \end{tabular}
-    \caption{Tools used in the DC module}
-    \label{tab:dc_tools}
-  \end{table}
-
   \subsection*{Taxonomy Assignment (TA)}
@@
 
-  \begin{table}[h]
-    \centering
-    \small
-    \begin{tabular}{|c|c|c|}
-      \hline
-      \textbf{Module} & \textbf{Tool/Database} & \textbf{References} \\
-      \hline
-      Query tool & Blast & \cite{camachoBLASTArchitectureApplications2009,bokulichOptimizingTaxonomicClassification2018} \\
-      Query tool & Naive bayes classifier & \cite{bokulichOptimizingTaxonomicClassification2018} \\
-      \hline
-      Database & Greengenes & \cite{DeSantis2006} \\
-      Database & SILVA & \cite{Quast2012} \\
-      Database & NCBI RefSeq & \cite{Sayers2009} \\
-      \hline
-    \end{tabular}
-    \caption{Tools used in the TA module}
-    \label{tab:ta_tools}
-  \end{table}
-
+  % TODO : Add information about filtering when OP=off
   \subsection*{OTU and ESV Processing (OP)}
@@
   The output of this step are co-occurrence association networks encoded in the JSON format (refer Supplementary section).
-
-  \begin{table}[h]
-    \centering
-    \small
-    \begin{tabular}{|c|c|c|}
-      \hline
-      \textbf{Module} & \textbf{Tool/Database} & \textbf{References} \\
-      \hline
-      Booststrap & \texttt{fastspar\_bootstraps} v1.0 & \cite{Watts2018} \\
-      Booststrap & \texttt{fastspar\_pvalues} v1.0 & \cite{Watts2018} \\
-      \hline
-      Direct & \ac{spieceasi} v1.1.2 & \cite{Kurtz2015} \\
-      Direct & FlashWeave.jl v0.18.1 & \cite{tackmannRapidInferenceDirect2019} \\
-      Direct & \ac{cozine} v1.0 & \cite{haCompositionalZeroinflatedNetwork2020a} \\
-      Direct & \ac{harmonies} v1.0 & \cite{jiangHARMONIESHybridApproach2020} \\
-      Direct & \ac{spring} v1.0.4 & \cite{yoonMicrobialNetworksSPRING2019} \\
-      Direct & \ac{mldm} v1.1 & \cite{Yang2017} \\
-      \hline
-      Correlation & FastSpar (\ac{sparcc}) v1.0 & \cite{Watts2018} \\
-      Correlation & Pearson & - \\
-      Correlation & Spearman & - \\
-      Correlation & propr v2.1.2 & \cite{quinnProprRpackageIdentifying2017} \\
-      \hline
-    \end{tabular}
-    \caption{Tools used in the NI module}
-    \label{tab:ni_tools}
-  \end{table}
 
@@
              \sum_{k=1}^3 \left( \alpha^{TA(k)}_j.\delta^{TA(k)}_j \right) +
-             \sum_{k=1}^2 \left( \alpha^{OP(k)}_j.\delta^{OP(k)}_j \right) +
+             \sum_{k=1}^2 \left( \alpha^{OP(k)}_j.\delta^{OP(k)}_j \right) + \\
              \sum_{k=1}^{10} \left( \alpha^{NI(k)}_j.\delta^{NI(k)}_j \right) +
@@
   Synthetic data and scripts: \href{https://github.com/segrelab/MiCoNE-synthetic-data}{https://github.com/segrelab/MiCoNE-synthetic-data}
+
+  \subsection*{Statistical analyses}
+  \vspace{-5mm}
+
+  \subsubsection*{DC step}
+  % Write about UniFrac
+
+  \subsubsection*{TA step}
+  % Write about Braycurtis
+
+  \subsubsection*{NI step}
+  % Write about precision and sensitivity equations used
+  % L1 distance equations used

### 2022-04-19 sections/tables_figures.tex
Change: Figure 6 <-> Figure 7
Change: Move tables to the end of the manuscript

@@
     \label{tab:default_options}
+  \end{table}
+
+  \begin{table}[h]
+    \centering
+    \small
+    \begin{tabular}{|c|c|c|}
+      \hline
+      \textbf{Module} & \textbf{Tool} & \textbf{References} \\
+      \hline
+      Denoising and Clustering & Closed reference & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
+      Denoising and Clustering & Open reference & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
+      Denoising and Clustering & De novo & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
+      Denoising and Clustering & Dada2 & \cite{Callahan2016} \\
+      Denoising and Clustering & Deblur & \cite{Amir2017,bolyenReproducibleInteractiveScalable2019} \\
+      \hline
+      Chimera checking & Uchime & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
+      Chimera checking & Remove bimera & \cite{Callahan2016} \\
+      \hline
+    \end{tabular}
+    \caption{Tools used in the DC module}
+    \label{tab:dc_tools}
+  \end{table}
+
+  \begin{table}[h]
+    \centering
+    \small
+    \begin{tabular}{|c|c|c|}
+      \hline
+      \textbf{Module} & \textbf{Tool/Database} & \textbf{References} \\
+      \hline
+      Query tool & Blast & \cite{camachoBLASTArchitectureApplications2009,bokulichOptimizingTaxonomicClassification2018} \\
+      Query tool & Naive bayes classifier & \cite{bokulichOptimizingTaxonomicClassification2018} \\
+      \hline
+      Database & Greengenes & \cite{DeSantis2006} \\
+      Database & SILVA & \cite{Quast2012} \\
+      Database & NCBI RefSeq & \cite{Sayers2009} \\
+      \hline
+    \end{tabular}
+    \caption{Tools used in the TA module}
+    \label{tab:ta_tools}
+  \end{table}
+
+  \begin{table}[h]
+    \centering
+    \small
+    \begin{tabular}{|c|c|c|}
+      \hline
+      \textbf{Module} & \textbf{Tool/Database} & \textbf{References} \\
+      \hline
+      Booststrap & \texttt{fastspar\_bootstraps} v1.0 & \cite{Watts2018} \\
+      Booststrap & \texttt{fastspar\_pvalues} v1.0 & \cite{Watts2018} \\
+      \hline
+      Direct & \ac{spieceasi} v1.1.2 & \cite{Kurtz2015} \\
+      Direct & FlashWeave.jl v0.18.1 & \cite{tackmannRapidInferenceDirect2019} \\
+      Direct & \ac{cozine} v1.0 & \cite{haCompositionalZeroinflatedNetwork2020a} \\
+      Direct & \ac{harmonies} v1.0 & \cite{jiangHARMONIESHybridApproach2020} \\
+      Direct & \ac{spring} v1.0.4 & \cite{yoonMicrobialNetworksSPRING2019} \\
+      Direct & \ac{mldm} v1.1 & \cite{Yang2017} \\
+      \hline
+      Correlation & FastSpar (\ac{sparcc}) v1.0 & \cite{Watts2018} \\
+      Correlation & Pearson & - \\
+      Correlation & Spearman & - \\
+      Correlation & propr v2.1.2 & \cite{quinnProprRpackageIdentifying2017} \\
+      \hline
+    \end{tabular}
+    \caption{Tools used in the NI module}
+    \label{tab:ni_tools}
   \end{table}
@@
     \centering
-    \includegraphics[width=0.85\linewidth]{figure6.pdf}
+    \includegraphics[width=1.0\linewidth]{figure6.pdf}
     \caption{
@@
     \centering
-    \includegraphics[width=17cm]{figure7.pdf}
+    \includegraphics[width=0.85\linewidth]{figure7.pdf}
     \caption{

### 2022-04-19 sections/methods.tex
Add: Section about DC statistical analyses

@@
   \subsubsection*{DC step}
-  % Write about UniFrac
+  In order to compare the representative sequences generated by the various methods in the DC step, we employ both the weighted~\cite{Lozupone2007} (Figure~\ref{fig:figure3}A) and unweighted UniFrac method~\cite{Lozupone2005} (Figure~\ref{fig:figure3}B).
+  The UniFrac distance metric (unique fraction metric) is a beta-diversity measure that computes the distance between two sets of taxa as the fraction of the branch length of the tree that leads to descendants from either one environment or the other, but not both~\cite{Lozupone2005}
+  The weighted UniFrac distance metric takes into account the abundances of the representative sequences when calculating shared and unshared branch lengths, whereas the unweighted UniFrac distance metric does not and hence gives equal weights to each sequence.
+  In Figure~\ref{fig:figure3} the distances between methods are the average distance calculated between the reference sequence distribution generated for every corresponding sample in the dataset.
+  All UniFrac calculations were performed using the \texttt{scikit-bio}~\cite{thescikit-biodevelopmentteamScikitbioBioinformaticsLibrary2022} v0.5.7 Python package.
 

### 2022-04-19 sections/methods.tex
Add: Section about TA statistical analyses

@@
   In Figure~\ref{fig:figure3} the distances between methods are the average distance calculated between the reference sequence distribution generated for every corresponding sample in the dataset.
-  All UniFrac calculations were performed using the \texttt{scikit-bio}~\cite{thescikit-biodevelopmentteamScikitbioBioinformaticsLibrary2022} v0.5.7 Python package.
+  All UniFrac calculations were performed using the \texttt{scikit-bio}~\cite{thescikit-biodevelopmentteamScikitbioBioinformaticsLibrary2022} v0.5.6 Python package.
 
   \subsubsection*{TA step}
-  % Write about Braycurtis
+  In Figure~\ref{fig:figure4}C we use the Bray-Curtis distance metric to calculate the distance between the predicted (using the TA step) and expected taxonomic distribution.
+  The Bray-Curtis distance is used to quantify the compositional dissimilarity between two different taxonomic distributions.
+  It is defined as: $\sum{|u_i-v_i|} / \sum{|u_i+v_i|}$
+  The Bray-Curtis distance calculations were performed using the \texttt{scipy}~\cite{virtanenSciPyFundamentalAlgorithms2020} v1.8.0 Python package.
 

### 2022-04-19 sections/methods.tex
Add: Section about NI statistical analyses

@@
   % Write about precision and sensitivity equations used
+  In Figure~\ref{fig:figure6} we calculate the accuracy and robustness of the inferred association networks (using the various network inference algorithms and the consensus methods) against the original interaction used to create the taxonomic distribution.
+  We use precision and sensitivity to calculate the accuracy of the predictions.
+  Precision is given by, $\frac{TP}{TP + FP}$
+  Sensitivity is given by, $\frac{TP}{FN + TP}$
   % L1 distance equations used
+  In Figure~\ref{fig:figure7}B we use the L1 distance metric to measure to the distance between the network inferred using different tools at the same step of the workflow.
+  The L1 distance between networks $x$ and $y$ is calculated using the following formula:
+  $$ \frac{|A_{x}-A_{y}|}{2} $$, where $A$ is the adjacency matrix of the network.

### 2022-04-20 sections/results.tex
Change: Update results with insights from Figure 7B

@@
   The filtering in the \ac{OP} step when turned off also adds a large number of new nodes, but the edges between the common nodes (with the default network) are fairly conserved.
-  % TODO: Finalize this figure
-  Furthermore, the L1 distance of networks generated by altering one of the steps of the pipeline from the default against the default network (Figure~\ref{fig:figure7}B) shows that the biggest deviations from the default network occur when the \ac{ta} and \ac{ni} steps are changed, reinforcing the same results observed in Figure~\ref{fig:figure2}.
-  Figure~\ref{fig:figure7} shows the co-occurrence networks inferred for the healthy subjects (control) and autistic subjects in the fecal microbiome transplant study~\cite{Kang2017}.
+  Furthermore, the L1 distance of networks generated by altering one of the steps of the pipeline from the default against the default network (Figure~\ref{fig:figure7}B) shows that the biggest deviations from the default network occur when the \ac{ni} steps are changed.
+  This result is different from that observed in Figure~\ref{fig:figure2} since in this particular analyses all the steps other than the \ac{ni} step have the consensus algorithm applied to them, which drastically reduces the edges in the graph that have lower probabilities.
+  This result reinforces the importance of the consensus network in increasing the robustness of the inferred networks to different processing methods.
+  Figure~\ref{fig:figure7}C shows the co-occurrence networks inferred for the healthy subjects (control) and autistic subjects in the fecal microbiome transplant study~\cite{Kang2017}.
   These consensus networks were generated using the default tools and parameters from Table~\ref{tab:default_options}.

### 2022-04-20 scripts/synthetic_data/art_wrapper.sh
Add: Run config and metrics for all pipeline runs

@@ -1,6 +1,0 @@
-#!/usr/bin/env bash
-
-for i in ../outputs/seq_pools/*.fasta; do
-	fname=$(basename -s .fasta "$i")
-	art_illumina -ss HS25 -amp -na -i "$i" -l 150 -f 1 -o "../outputs/reads/$fname"
-done

### 2022-04-20 scripts/synthetic_data/fastq_header_mapper.py

@@ -1,35 +1,0 @@
-#!/usr/bin/env python3
-
-from collections import defaultdict
-import csv
-import pathlib
-
-
-def main(fastq_files, output_folder):
-    num_size = 6
-    count = 0
-    map_dict = defaultdict(list)
-    for fastq_file in fastq_files:
-        output_file = output_folder / fastq_file.name
-        with open(fastq_file, 'r') as in_fid:
-            with open(output_file, 'w') as out_fid:
-                for line in in_fid:
-                    if line.startswith('@'):
-                        tax = line[1:].split('-', 1)[0]
-                        id_ = str(count).zfill(num_size)
-                        map_dict[tax].append(id_)
-                        count += 1
-                        out_fid.write(f"@{id_}\n")
-                    else:
-                        out_fid.write(line)
-    csv_file = output_folder / "map_file.csv"
-    with open(csv_file, 'w') as fid:
-        csv_writer = csv.writer(fid, delimiter=',')
-        for k, v in map_dict.items():
-            csv_writer.writerow([k, *v])
-
-
-if __name__ == "__main__":
-    FILES = pathlib.Path("../seqs/art/original").glob("*.fq")
-    OUT_FOL = pathlib.Path("../seqs/art")
-    main(FILES, OUT_FOL)

### 2022-04-20 scripts/synthetic_data/generate_seq_pools.py

@@ -1,43 +1,0 @@
-#!/usr/bin/env python3
-
-from Bio import SeqIO
-from Bio.Seq import Seq
-from Bio.SeqRecord import SeqRecord
-import pandas as pd
-import random
-
-ALPHABET = {'A', 'T', 'G', 'C'}
-
-
-def mutate(record: SeqRecord, id_suffix: str, rate: float = 0.02) -> SeqRecord:
-    new_id = f"{record.id}-{id_suffix}"
-    sequence = record.seq
-    change = lambda x: random.choice(list(ALPHABET - {x}))
-    mutated_seq = ''.join([change(s) if random.random() <= rate else s for s in sequence])
-    return SeqRecord(seq=Seq(mutated_seq), id=new_id, name='', description='')
-
-
-def main(reference_file, abundance_file, read_depth, out_folder) -> None:
-    abundances = pd.read_csv(abundance_file, index_col=0)
-    counts = (abundances * read_depth / 100).astype(int)
-    seq_dict = {}
-    for record in SeqIO.parse(reference_file, "fasta"):
-        seq_dict[record.id] = record
-    for sample_name, sample_counts in counts.items():
-        print(f"Processing {sample_name}")
-        seq_list = []
-        valid_counts = sample_counts[sample_counts > 0]
-        for tax_name, count in valid_counts.items():
-            tax_seq = seq_dict[tax_name]
-            for i in range(count):
-                seq_list.append(mutate(tax_seq, str(i)))
-        fname = out_folder + sample_name + ".fasta"
-        SeqIO.write(seq_list, fname, "fasta")
-
-
-if __name__ == "__main__":
-    REFERENCE = "../inputs/references.fasta"
-    ABUNDANCE = "../inputs/abundance_file.txt"
-    DEPTH = 5000
-    OUT_FOLDER = "../outputs/seq_pools/"
-    main(REFERENCE, ABUNDANCE, DEPTH, OUT_FOLDER)

### 2022-04-20 scripts/synthetic_data/real_comp_calculator.py

@@ -1,30 +1,0 @@
-#!/usr/bin/env python3
-
-from collections import defaultdict
-import pathlib
-
-from Bio import SeqIO
-import pandas as pd
-
-
-def main(files):
-    counts_dict = {}
-    for file in files:
-        key = file.stem
-        counts_dict[key] = defaultdict(int)
-        for record in SeqIO.parse(file, "fastq"):
-            tax = record.id.split('-', 1)[0]
-            counts_dict[key][tax] += 1
-    counts_df = pd.DataFrame(counts_dict)
-    counts_df.fillna(value=0.0, inplace=True)
-    counts_df.index.name = "tax"
-    counts_df.to_csv("../real_composition/art/counts.csv")
-    abund_df = counts_df / counts_df.sum(axis=0)
-    abund_df.to_csv("../real_composition/art/abundances.csv")
-    comp_df = abund_df * 100
-    comp_df.to_csv("../real_composition/art/composition.csv")
-
-
-if __name__ == "__main__":
-    FILES = pathlib.Path("../seqs/art/original/").glob("*.fq")
-    main(FILES)

### 2022-04-20 sections/methods.tex
Add: Data processing section to supplementary

@@
     The current analysis used stool samples from a fecal microbiome transplant study of autism~\cite{Kang2017} as biological datasets.
-    The study was composed of multiple sequencing runs and those runs that contained paired end reads (runs 2, 3 and 4), were downloaded from Qiita~\cite{qiita} from https://qiita.ucsd.edu/study/description/10532 and used as input sequences for the \ac{micone} pipeline.
-    % TODO: How many samples are there for each and average number of reads.
-    Sequences from both control (x samples containing xx reads) and autistic (y samples containing yy reads) patients were included in the analyses but the groups were analyzed independently in the pipeline.
+    The study was composed of multiple sequencing runs and those runs that contained paired end reads (runs 2 (10M reads), 3 (750K reads) and 4 (16M reads)), were downloaded from Qiita~\cite{qiita} from https://qiita.ucsd.edu/study/description/10532 and used as input sequences for the \ac{micone} pipeline.
+    Sequences from both control (212 samples including neurotypical and donors) and autistic (126 samples) patients were included in the analyses but the groups were analyzed independently in the pipeline.
     The mock community 16S datasets are sequencing data obtained for artificially assembled collections of species in known proportions.
@@
 
-  \subsection*{Code and Data Availability}
-  Pipeline: \href{https://github.com/segrelab/MiCoNE}{https://github.com/segrelab/MiCoNE} \\
-  Documentation: \href{https://micone.readthedocs.io}{https://micone.readthedocs.io} \\
-  Data and scripts: \href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper} \\
-  Synthetic data and scripts: \href{https://github.com/segrelab/MiCoNE-synthetic-data}{https://github.com/segrelab/MiCoNE-synthetic-data}
-
   \subsection*{Statistical analyses}
@@
   $$ \frac{|A_{x}-A_{y}|}{2} $$, where $A$ is the adjacency matrix of the network.
+
+  \subsection*{Code and Data Availability}
+  Pipeline: \href{https://github.com/segrelab/MiCoNE}{https://github.com/segrelab/MiCoNE} \\
+  Documentation: \href{https://micone.readthedocs.io}{https://micone.readthedocs.io} \\
+  Data and scripts: \href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper} \\
+  Synthetic data and scripts: \href{https://github.com/segrelab/MiCoNE-synthetic-data}{https://github.com/segrelab/MiCoNE-synthetic-data}

### 2022-04-20 sections/supplementary.tex
Update on Overleaf.
Add: Data processing section to supplementary

--- 2021-01-11 sections/supplementary.tex
@@
   \input{sections/supplementary_table.tex}
+
+  \subsection*{Processing the FMT data}
+
+    \subsubsection*{Data download and pre-processing}
+    The main dataset used in this study contained stool samples (healthy and autistic individuals) from a fecal microbiome transplant dataset~\cite{Kang2017}.
+    The data containing the 16S sequencing reads (V4 region) was downloaded from Qiita~\cite{qiita} (study id: 10532).
+    Only runs 2, 3, and 4 were used for the subsequent analysis because these runs were paired-end sequencing data whereas run 1 was single-end.
+    The sample metadata was updated so as to contain only bmi, sex, height, weight and experimental group.
+    This was necessary as some of the network inference algorithms (like \ac{mldm}) required information about envioronmental heterogeneity.
+    However, these results were not included in the current analyses.
+
+    \subsubsection*{Processing using the \ac{micone} pipeline}
+    The data was then processed using the \ac{micone} pipeline starting at the \ac{sp} step and ending at the \ac{ni} step with the consensus algorithm.
+    The configuration files (main.nf and nextflow.config) used to run the \ac{micone} pipeline as well the details of the pipeline execution (dag, report, timeline and trace) are in the "runs/FMT" directory of the data and scripts repository (\href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper})
+    % TODO: Store pipeline results
+    The results of the pipeline execution are stored on Zenodo[REF] at ...
+
+  \subsection*{Processing the mock data}
+
+    \subsubsection*{Data download and pre-processing}
+    The mock datasets, mock4, mock12 and mock16 used for this study, were obtained from mockrobiota~\cite{Bokulich2016}.
+    Mock 4 is a mock community composed of 21 bacterial strains represented in equal abundances in two replicate samples, and the same strains represented in uneven abundances in two replicate samples.
+    Mock 12 is composed of 27 bacterial strains containing closely related taxa, the members of which were chosen in part for their well-separated 16S rRNA gene sequences.
+Some pairs of strains differ by as little as one nucleotide, but all the strains are distinguishable over the sequenced region of the 16S rRNA gene.
+    Mock 16 is a mock community composed of even amounts of purified genomic DNA from 49 bacteria and 10 archaea.
+    The datasets did not require any preprocessing and could be directly used as input to the pipeline
+
+    \subsubsection*{Processing using the \ac{micone} pipeline}
+    The data was processed using the \ac{micone} pipeline starting at the \ac{sp} step and ending at the \ac{op} step with the filtered taxonomic tables as the final output.
+    The configuration files (main.nf and nextflow.config) used to run the \ac{micone} pipeline as well the details of the pipeline execution (dag, report, timeline and trace) are in the "runs/mock*" directory of the data and scripts repository (\href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper})
+    % TODO: Store pipeline results
+    The results of the pipeline execution are stored on Zenodo[REF] at ...
+
+  \subsection*{Synthetic interaction data}
+
+    \subsubsection*{Data generation}
+    Figures describing the original OTU tables or the interactions used to generate the community. Also include the seed and scripts for reproducing the networks.
+
+    \subsubsection*{Processing using the \ac{micone} pipeline}
+
+  \subsection*{Consensus and p-value}
+
+  % FIXME: Update according to the new notation in methods section
+  \subsubsection*{p-value merging}
+  Fisher~\cite{fisher_224a_1948} proposed that for $k$ independent p-values, each generated by $k$ different methods and denoted by $P_i$, the statistic $\Psi$:
+  \begin{equation*}
+    \begin{aligned}
+        \Psi &= \sum_{i=1}^k -2 \log \left( P_i \right) \\
+        \Psi &\sim \chi^2_{2k}
+    \end{aligned}
+  \end{equation*}
+
+  Brown~\cite{brown_400_1975} extended Fisher's method to dependent p-values by using a re-scaled $\chi^2$ distribution:
+  \begin{equation*}
+    \Psi \sim c \chi^2_{2f}
+  \end{equation*}
+  where, $f$ is the degrees of freedom and $c$ is the scale factor and are given by:
+  \begin{equation*}
+    f = \frac{\mathrm{E}[\Psi]^2}{\mathrm{Var}[\Psi]} ~~~\text{and}~~~ c = \frac{\mathrm{Var}[\Psi]}{2\mathrm{E}[\Psi]} = \frac{k}{f}
+  \end{equation*}
+
+  Furthermore, Brown showed that $\mathrm{E}[\Psi]$ and $\mathrm{Var}[\Psi]$ can be calculated directly via a numerical integration:
+  \begin{equation*}
+    \mathrm{E}[\Psi] = 2k ~~~\text{and}~~~ \mathrm{Var}[\Psi] = 4k + 2\sum_{i<j} \mathrm{Cov}\left( -2\log(P_i), -2\log(P_j) \right)
+  \end{equation*}
+
+  Kost and McDermott~\cite{kost_combining_2002} further fit a third-order polynomial to approximate the covariance
+  \begin{equation}
+    \mathrm{Cov}\left( -2\log(P_i), -2\log(P_j) \right) \approx 3.263 \rho_{ij} + 0.710 \rho_{ij}^2 + 0.027 \rho_{ij}^3
+    \label{eqn:covariance-pvalues}
+  \end{equation}
+  where, $\rho_{ij}$ is the correlation between method $i$ and method $j$
+
+  The final combined p-value~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016} is then given by:
+  \begin{equation}
+    \begin{aligned}
+        & P_{combined} = 1.0 - \Phi_{2f}\left( \psi / c \right) \\
+        \text{where},~ &\psi = -2 \sum_{i=1}^k \log(P_i) ~~~\text{and}~~~ \Phi_{2f} = \mathrm{CDF}\left( \chi^2_{2f} \right)
+    \end{aligned}
+    \label{eqn:pvalue-combined}
+  \end{equation}
+
+  The p-value merging method in \ac{micone} (refer Documentation) uses Equation~\ref{eqn:covariance-pvalues} to estimate the covariance of the pvalues and Equation~\ref{eqn:pvalue-combined} to merge the p-values (obtained from bootstrapping) from the different correlation methods.
+  % FIXME: Not sure what our stance on this is?
+  Note that we do not use Pearson and Spearman methods in the p-value merging step and these algorithms are only used for demonstration and comparison.
+  The combined p-values are used to threshold for significance during the consensus network step.
 

### 2022-04-20 sections/methods.tex
Add: Synth data section to supplementary

@@
 
-  % TODO: Add reference: Karoline Faust, Franziska Bauchinger, Béatrice Laroche, Sophie de Buyl, Leo Lahti, Alex D Washburne, Didier Gonze, Stefanie Widder, "Signatures of ecological processes in microbial community time series", Microbiome (2018).
-  The first method, hereby referred to as ``seqtime'', used generalized Lotka-Volterra (gLV) equations to model the microbial community dynamics and utilized the Klemm–Eguı́luz algorithm to generate a clique-based interaction network~\cite{Rottjers2018}.
+  The first method, hereby referred to as ``seqtime''~\cite{faustSignaturesEcologicalProcesses2018}, used generalized Lotka-Volterra (gLV) equations to model the microbial community dynamics and utilized the Klemm–Eguı́luz algorithm to generate a clique-based interaction network~\cite{Rottjers2018}.
   We used the \href{https://github.com/hallucigenia-sparsa/seqtime}{seqtime} R package to simulate communities with number of species ($N$) varying from 10 to 150 (10, 25, 50, 100, 150 and 200).

### 2022-04-20 sections/supplementary.tex
Add: Synth data section to supplementary

@@
     \subsubsection*{Data generation}
-    Figures describing the original OTU tables or the interactions used to generate the community. Also include the seed and scripts for reproducing the networks.
+    The synthetic interaction data for the study were generated using two methods.
+    The first method, ``seqtime''~\cite{faustSignaturesEcologicalProcesses2018} utilized generalized Lotka-Volterra (gLV) equations to model the microbial community dynamics and made use of the Klemm–Eguı́luz algorithm to generate clique-based interaction networks~\cite{Rottjers2018}.
+    We used the seqtime R package to simulate communities with different numbers of species and samples (see Methods for details).
+    The second method, ``NorTA'' used the Normal to Anything (NorTA) approach coupled with a given interaction network topology to generate the abundance distribution of the microbial community~\cite{Kurtz2015}.
+    We used the spieceasi R package to simulate communities with different abundance distributions and network topologies (see Methods for details).
+    The scripts to generate these datasets can be found in the synthetic data and scripts repository (\href{https://github.com/segrelab/MiCoNE-synthetic-data}{https://github.com/segrelab/MiCoNE-synthetic-data})
 
     \subsubsection*{Processing using the \ac{micone} pipeline}
+    The data was processed using the \ac{micone} pipeline using only the \ac{ni} step with the consensus networks as the final output.
+    The configuration files (main.nf and nextflow.config) used to run the \ac{micone} pipeline as well the details of the pipeline execution (dag, report, timeline and trace) are in the "runs/norta" and "runs/seqtime" directories of the data and scripts repository (\href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper})
+    % TODO: Store pipeline results
+    The results of the pipeline execution are stored on Zenodo[REF] at ...
 
@@
   The combined p-values are used to threshold for significance during the consensus network step.
+
+  \subsection*{Performance of network inference algorithms on seqtime data}
 

### 2022-04-20 scripts/data_processing/process_figures_data.sh
Add: Script to generate Figure S1

@@ -122,3 +122,5 @@ python -W ignore figure3_data.py \
   --weighted True \
   --threshold 10 \
   --output ../../data/figure3/output/$DATASET
+
+#### Supplementary figures

### 2022-04-20 scripts/figure_generation/generate_figures.sh

@@ -56,3 +56,10 @@ Rscript figure6_plot.R "../../data/figure6/output/norta/" "$OUTPUT_DIR"
 echo "Generating Figure 7"
 cd ../figure7/
 Rscript figure7_plot.R "../../data/figure7/output/$DATASET/" "fmt-control" "fmt-autism" "$OUTPUT_DIR"
+
+#### Supplementary figures
+
+# Figure S1
+echo "Generating Figure S1"
+cd ../figure_s1/
+Rscript figure_s1_plot.R "../../data/figure2/output/$DATASET/" "$OUTPUT_DIR"

### 2022-04-20 scripts/figure_s1/figure_s1_plot.R

@@ -1,0 +1,46 @@
+#!/usr/bin/env Rscript
+
+library(dplyr)
+library(ggplot2)
+library(ggpubr)
+library(ggrepel)
+
+# inputs
+args <- commandArgs(trailingOnly = TRUE)
+if (length(args) == 0) {
+    data_folder <- "../../data/figure2/output/moving_pictures/"
+    output_folder <- "../../figures/"
+} else if (length(args) == 2) {
+    data_folder <- args[1]
+    output_folder <- args[2]
+} else {
+    stop("Required number of arguments must equal 2")
+}
+x_csv <- paste0(data_folder, "x.csv")
+y_reduced_csv <- paste0(data_folder, "y_reduced2.csv")
+percentage_variance_csv <- paste0(data_folder, "percentage_variance.csv")
+output_file <- paste0(output_folder, "figure_s1.pdf")
+
+plot_scatter <- function(data, color_var) {
+    ggscatter(data,
+        x = "X0", y = "X1",
+        color = color_var,
+        # ellipse = TRUE
+    ) +
+    theme(
+        text = element_text(size = 16),
+    )
+}
+
+x <- read.csv(x_csv)
+y_reduced <- read.csv(y_reduced_csv)
+names(y_reduced)[names(y_reduced) == "X"] <- "hash"
+
+df <- y_reduced %>%
+    select("hash", "X0", "X1") %>%
+    left_join(x, by = "hash")
+scatter_dc <- plot_scatter(df, "DC")
+scatter_op <- plot_scatter(df, "OP")
+
+ggarrange(scatter_dc, scatter_op, ncol = 2, labels = c("A", "B"))
+ggsave(output_file, width = 14, height = 12)

### 2022-04-20 scripts/figure_s2/figure_s2_plot.R

@@ -1,52 +1,0 @@
-#!/usr/bin/env Rscript
-
-library(dplyr)
-library(ggplot2)
-library(ggpubr)
-
-methods <- c("closed_reference", "de_novo", "open_reference", "deblur", "dada2")
-
-tidy_up_data <- function(unifrac_data) {
-    unifrac_data$method1 <- factor(unifrac_data$method1, levels=methods)
-    unifrac_data$method2 <- factor(unifrac_data$method2, levels=methods)
-    tidy_data <- unifrac_data %>%
-        group_by(method1, method2) %>%
-        select_at(c("method1", "method2", "unifrac")) %>%
-        summarize_all(.funs=mean) %>%
-        rowwise() %>%
-        mutate(pair=sort(c(method1, method2)) %>% paste(collapse=",")) %>%
-        group_by(pair) %>%
-        distinct(pair, .keep_all=TRUE)
-}
-
-plot_heatmap <- function(data, title) {
-    ggplot(data=data, aes(x=method1, y=method2, fill=unifrac)) +
-        geom_tile(color="white") +
-        geom_text(aes(label=abbreviate(unifrac))) +
-        scale_fill_distiller(
-            palette="Spectral",
-            limits=c(0, 1),
-            breaks=c(0, 0.25, 0.5, 0.75, 1),
-            labels=c("Similar", 0.25, 0.5, 0.75, "Dissimilar")
-        ) +
-        ggtitle(title) +
-        theme_pubr() +
-        theme(
-            plot.title=element_text(hjust=0.5),
-            axis.title=element_blank(),
-            axis.text.x=element_text(angle=30, hjust=1),
-            axis.text.y=element_text(angle=30)
-        ) +
-        coord_fixed()
-}
-
-unweighted <- read.csv("hp_unweighted_unifrac.csv", sep=",", header=TRUE)
-unweighted_tidy <- tidy_up_data(unweighted)
-weighted <- read.csv("hp_weighted_unifrac.csv", sep=",", header=TRUE)
-weighted_tidy <- tidy_up_data(weighted)
-
-weighted_plot <- plot_heatmap(weighted_tidy, "weighted unifrac")
-unweighted_plot <- plot_heatmap(unweighted_tidy, "unweighted unifrac")
-
-heatmap_plot <- ggarrange(weighted_plot, unweighted_plot, nrow=1, ncol=2, labels=c("A", "B"), common.legend=TRUE, legend="right")
-ggsave("figure_s2.pdf", heatmap_plot, width=11, height=8.5)

### 2022-04-20 scripts/figure_s3/figure_s3_plot.R

@@ -1,66 +1,0 @@
-#!/usr/bin/env Rscript
-
-library(dplyr)
-library(ggplot2)
-library(ggpubr)
-
-methods <- c("input", "closed_reference", "de_novo", "open_reference", "deblur", "dada2")
-
-tidy_up_data <- function(unifrac_data, name) {
-    unifrac_data$dataset <- rep(name, nrow(unifrac_data))
-    unifrac_data$method1 <- factor(unifrac_data$method1, levels=methods)
-    unifrac_data$method2 <- factor(unifrac_data$method2, levels=methods)
-    tidy_data <- unifrac_data %>%
-        filter(method2 == "input") %>%
-        filter(method1 != "input") %>%
-        select_at(c("method1", "sample", "unifrac", "dataset"))
-}
-
-make_dot_plot <- function(unifrac_data, title) {
-    ggstripchart(
-        unifrac_data,
-        x="method1",
-        y="unifrac",
-        color="dataset",
-        title=title,
-        xlab="denoising/clustering method",
-        ylab="unifrac distance",
-        ylim=c(0, 1),
-        size=3,
-        add="jitter",
-    ) +
-    theme(
-        plot.title=element_text(hjust=0.5),
-        axis.text.x=element_text(angle=30, hjust=1),
-    )
-}
-
-# Weighted
-emp_soil_weighted <- read.csv("emp_soil_weighted_unifrac.csv", sep=",", header=TRUE)
-emp_water_weighted <- read.csv("emp_water_weighted_unifrac.csv", sep=",", header=TRUE)
-fmt_weighted <- read.csv("mock12_weighted_unifrac.csv", sep=",", header=TRUE)
-
-emp_soil_weighted_tidy <- tidy_up_data(emp_soil_weighted, "emp_soil")
-emp_water_weighted_tidy <- tidy_up_data(emp_water_weighted, "emp_water")
-fmt_weighted_tidy <- tidy_up_data(fmt_weighted, "stool")
-
-weighted_tidy <- rbind(emp_soil_weighted_tidy, emp_water_weighted_tidy, fmt_weighted_tidy)
-weighted_plot <- make_dot_plot(weighted_tidy, "weighted unifrac")
-
-# Unweighted
-emp_soil_unweighted <- read.csv("emp_soil_unweighted_unifrac.csv", sep=",", header=TRUE)
[…]

### 2022-04-20 scripts/figure_s5/figure_s5_plot.R

@@ -1,44 +1,0 @@
-#!/usr/bin/env Rscript
-
-library(tidyverse)
-library(ggplot2)
-library(ggpubr)
-
-levels <- c("Family", "Genus", "Species")
-
-tidy_up_data <- function(data_file, name) {
-    data <- read_csv(data_file)
-    data$dataset <- rep(name, nrow(data))
-    data$tax_level <- factor(data$tax_level, levels=levels)
-    return(data)
-}
-
-
-make_dot_plot <- function(braycurtis_data) {
-    ggstripchart(
-        braycurtis_data,
-        x="database",
-        y="braycurtis",
-        color="dataset",
-        xlab="database",
-        ylab="Braycurtis dissimilarity",
-        ylim=c(0, 1),
-        size=3,
-        add="jitter",
-    ) +
-    theme(
-        plot.title=element_text(hjust=0.5),
-    )
-}
-
-
-emp_soil_data <- tidy_up_data("emp_soil_braycurtis.csv", "emp_soil")
-emp_water_data <- tidy_up_data("emp_water_braycurtis.csv", "emp_water")
-fmt_data <- tidy_up_data("mock12_braycurtis.csv", "stool")
-
-synth_data <- rbind(emp_soil_data, emp_water_data, fmt_data)
-dot_plot <- make_dot_plot(synth_data)
-dot_plot_facet <- facet(dot_plot, facet.by="tax_level")
-# final_plot <- annotate_figure(dot_plot_facet, fig.lab="C", fig.lab.pos="top.left", fig.lab.size=20)
-
-ggsave("figure_s5.pdf", width=11, height=5)

### 2022-04-20 scripts/figure_s5/make_input_taxonomy.py

@@ -1,59 +1,0 @@
-#!/usr/bin/env python
-
-import click
-from ete3 import NCBITaxa
-import pandas as pd
-
-from mindpipe import Lineage
-
-NCBI = NCBITaxa()
-TAX_LEVELS = Lineage._fields
-TAX_MAP = {
-    "Kingdom": "superkingdom",
-    "Phylum": "phylum",
-    "Class": "class",
-    "Order": "order",
-    "Family": "family",
-    "Genus": "genus",
-    "Species": "species",
-}
-
-
-def get_lineage(species_name):
-    taxid = NCBI.get_name_translator([species_name])[species_name][0]
-    lineage_taxids = NCBI.get_lineage(taxid)
-    lineage_names = NCBI.get_taxid_translator(lineage_taxids)
-    lineage_ranks = {v: k for k, v in NCBI.get_rank(lineage_taxids).items()}
-    lineage_dict = dict()
-    for tax_level in TAX_LEVELS:
-        try:
-            rank_taxid = lineage_ranks[TAX_MAP[tax_level]]
-            rank_name = lineage_names[rank_taxid]
-            lineage_dict[tax_level] = rank_name
-        except KeyError:
-            if tax_level == "Genus":
-                lineage_dict["Genus"] = species_name.split(" ")[0]
-            elif tax_level == "Species":
-                lineage_dict["Species"] = species_name
-            else:
-                print(f"Warning unknown {tax_level} for {species_name}")
-                lineage_dict[tax_level] = ""
-    return Lineage(**lineage_dict)
-
-
-@click.command()
-@click.option("--expected_tax", help="Location of the input file")
-@click.option("--output_file", help="Location of the output file")
-def main(expected_tax: str, output_file: str):
-    tax_table = pd.read_csv(expected_tax, index_col=0, sep="\t")
-    species = [" ".join(t.split("_")[:2]) for t in tax_table.index]
-    lineages = map(get_lineage, species)
-    lineage_data = [lineage.to_dict(level="Species") for lineage in lineages]
-    lineage_table = pd.DataFrame(data=lineage_data, index=tax_table.index)
-    lineage_table = lineage_table[list(TAX_LEVELS)]
-    final_table = pd.concat([lineage_table, tax_table], axis=1)
-    final_table.to_csv(output_file, sep=",", index=False)
-
-
-if __name__ == "__main__":
-    main()

### 2022-04-20 sections/results.tex
Add: Script to generate Figure S1
Add: Section about effect of OP on variability

@@
   This result highlights the importance of the taxonomy assignment step in the 16S data analysis workflow, implying that a change in the reference database will result in drastically different inferred networks.
-  % TODO: Verify this
+  % TODO: Verify this: Figure S4
   The reason for this seems to derive from the mislabelling of constitutive taxa that are present in high abundance in the community, which drastically alter the nodes and hence the underlying network topology.
@@
   However, it is also evident that some of the networks, especially those inferred using the direct association based network infernce methods, are much closer in the PCA plot regardless of the reference database used.
-  % TODO: Figure out the how the OP step plays a role in the PCA plot
-  % TODO: Supplementary Figure S2
-  It can also be seen that the dissimilarity between the networks decreases when the low abundance \ac{otu}s are removed from the network.
+  It can also be seen that the variation (on the first two principal axes) between the networks decreases when the low abundance \ac{otu}s are removed from the network (Figure~\ref{fig:figure_s1}).
   These results suggest that the most important criterion for accurate comparative analyses of co-occurrence networks is the taxonomy reference database and the level of filtering of the taxonomy tables prior to network inference.

### 2022-04-20 sections/supplementary.tex
Change: Split up the parameter table
Add: Script to generate Figure S1
Add: Section about effect of OP on variability

@@
 
+  \subsection*{Effects of the \ac{dc} and \ac{op} steps on network variability}
+
+    % TODO: Elaborate on this
+    Figure~\ref{fig:figure_s1} describes the variation in the inferrred networks in the first two principal axes.
+    This Figure is the similar to Figure~\ref{fig:figure2}B and further confirms that the variability in the networks decreases upon filtering out the taxonomic entities at low abundance.
+
+    \begin{figure}[H]
+      \centering
+      \includegraphics[width=1.0\linewidth]{figure_s1.pdf}
+    \end{figure}
+    \begin{figure}[H]
+      \centering
+        \caption{
+          \textbf{The variation in the inferred networks in the first two principal axes}
+          The plots are colored by the tools or parameters used in the \ac{dc} step (A) and the \ac{op} step (B)
+        }
+      \label{fig:figure_s1}
+    \end{figure}
+    \FloatBarrier
+    \newpage
+
   \subsection*{Consensus and p-value}
@@
 
+  % TODO: Write all these sections
   \subsection*{Performance of network inference algorithms on seqtime data}
+  \subsection*{The JSON network format}
 

### 2022-04-20 scripts/figure_s3/figure_s3_data.py
Add: Script to generate Figure S3

@@ -1,0 +1,155 @@
+#!/usr/bin/env python3
+
+import pathlib
+from typing import List
+
+from biom import load_table
+import click
+import numpy as np
+import pandas as pd
+from skbio.tree import TreeNode
+from skbio.diversity.beta import weighted_unifrac, unweighted_unifrac
+
+
+def get_files(path_glob: str) -> List[pathlib.Path]:
+    base_dir, glob = path_glob.split("*", 1)
+    glob = "*" + glob
+    base_path = pathlib.Path(base_dir)
+    return list(base_path.glob(glob))
+
+
+def get_vectors(otu_1: pd.DataFrame, otu_2: pd.DataFrame, threshold: int):
+    samples_1 = set(otu_1.columns)
+    samples_2 = set(otu_2.columns)
+    removed_samples = len(samples_1 - samples_2) + len(samples_2 - samples_1)
+    print(f"Warning {removed_samples} samples have been removed")
+    common_samples = samples_1 & samples_2
+    for col in common_samples:
+        otu_1_col, otu_2_col = otu_1[[col]], otu_2[[col]]
+        otu_1_col.loc[otu_1_col[col] <= threshold, :] = 0
+        otu_2_col.loc[otu_2_col[col] <= threshold, :] = 0
+        otu_1_col.columns = ["1"]
+        otu_2_col.columns = ["2"]
+        joint_df = otu_1_col.join(otu_2_col, how="outer")
+        joint_df.fillna(0, inplace=True)
+        otu_ids = list(joint_df.index)
+        u = list(joint_df["1"])
+        v = list(joint_df["2"])
+        yield u, v, otu_ids, col
+
+
+def get_unifrac(
+    otu_file_1: pathlib.Path,
+    otu_file_2: pathlib.Path,
+    tree_file: pathlib.Path,
+    weighted: bool,
+    threshold: int,
+):
+    otu_1 = load_table(str(otu_file_1)).to_dataframe(dense=True)
+    otu_2 = load_table(str(otu_file_2)).to_dataframe(dense=True)
+    tree = TreeNode.read(str(tree_file))
+    unifrac_data = dict()
+    for u, v, otu_ids, col in get_vectors(otu_1, otu_2, threshold):
+        if weighted:
+            unifrac_value = weighted_unifrac(
+                u, v, otu_ids, tree, normalized=True, validate=True
+            )
+        else:
+            unifrac_value = unweighted_unifrac(u, v, otu_ids, tree, validate=True)
[…]

### 2022-04-20 scripts/figure_s3/figure_s3_plot.R

@@ -1,0 +1,74 @@
+#!/usr/bin/env Rscript
+
+library(dplyr)
+library(ggplot2)
+library(ggpubr)
+
+# inputs
+args <- commandArgs(trailingOnly = TRUE)
+if (length(args) == 0) {
+    data_folder <- "../../data/figure3/output/moving_pictures/"
+    mock_folder <- "../../data/figure3/output/"
+    output_folder <- "../../figures/"
+} else if (length(args) == 3) {
+    data_folder <- args[1]
+    mock_folder <- args[2]
+    output_folder <- args[3]
+} else {
+    stop("Required number of arguments must equal 3")
+}
+unweighted_unifrac_csv <- paste0(data_folder, "threshold_unweighted_unifrac.csv")
+weighted_unifrac_csv <- paste0(data_folder, "threshold_weighted_unifrac.csv")
+output_file <- paste0(output_folder, "figure_s3.pdf")
+
+#########################################
+# ab
+#########################################
+
+tidy_up_data_ab <- function(unifrac_data) {
+    unifrac_data$method1 <- factor(unifrac_data$method1)
+    unifrac_data$method2 <- factor(unifrac_data$method2)
+    tidy_data <- unifrac_data %>%
+        group_by(method1, method2) %>%
+        select_at(c("method1", "method2", "unifrac")) %>%
+        summarize_all(.funs = mean) %>%
+        rowwise() %>%
+        mutate(pair = sort(c(method1, method2)) %>% paste(collapse = ",")) %>%
+        group_by(pair) %>%
+        distinct(pair, .keep_all = TRUE)
+}
+
+plot_heatmap <- function(data, title) {
+    ggplot(data = data, aes(x = method1, y = method2, fill = unifrac)) +
+        geom_tile(color = "white") +
+        geom_text(aes(label = abbreviate(unifrac))) +
+        scale_fill_distiller(
+            palette = "Spectral",
+            limits = c(0, 1),
+            breaks = c(0, 0.25, 0.5, 0.75, 1),
+            labels = c("Similar", 0.25, 0.5, 0.75, "Dissimilar")
+        ) +
+        ggtitle(title) +
+        theme_pubr() +
+        theme(
+            plot.title = element_text(hjust = 0.5),
+            axis.title = element_blank(),
+            axis.text.x = element_text(angle = 30, hjust = 1),
[…]

### 2022-04-21 sections/results.tex
Add: Section for DC step supplementary

@@
   % TODO: Supplementary work - Figure S3
-  In order to verify this claim, for each of these methods we use the \ac{gg} taxonomy database to assign taxonomies to the representative sequences.
-  We then correlate the abundances of matching taxonomies between every pairwise combination of DC methods (Figure \ref{fig:figureS1}A and B).
-  The \ac{esv} tables generated by methods that perform denoising are very similar to each other ($\sim0.91$) and the \ac{otu} tables generated by the clustering methods are very similar to each other ($\sim0.9$), but results of denoising and clustering are highly uncorrelated with each other ($\sim0.4$) (Figure \ref{fig:figureS1}C).
+  In order to verify this claim, for each of these methods we use only the top 100 reference sequences and recalculate the distance metrics (Figure \ref{fig:figureS3}).
+  We observe that the both the weighted and unweighted UniFrac distances are decreased considerably implying that the top reference sequences generated by the different methods are similar to each other.
 

### 2022-04-21 sections/supplementary.tex
Add: Script to generate Figure S3
Add: Section for DC step supplementary

@@
       \label{fig:figure_s1}
+    \end{figure}
+    \FloatBarrier
+    \newpage
+
+    % TODO: Write something about TSNE
+    \begin{figure}[H]
+      \centering
+      \includegraphics[width=1.0\linewidth]{figure_s2.pdf}
+    \end{figure}
+    \begin{figure}[H]
+      \centering
+        \caption{
+          \textbf{The TSNE plot of the networks inferred by various methods}.
+        }
+      \label{fig:figure_s2}
+    \end{figure}
+    \FloatBarrier
+    \newpage
+
+  \subsection*{Effects of low abundant reference sequences in the \ac{dc} step}
+
+    Figure~\ref{fig:figureS3} is similar to Figure~\ref{fig:figure3} shows the average UniFrac distance between reference sequences generated by the various methods in the \ac{dc} step.
+    In Figure~\ref{fig:figureS3} however, we use only the top  100 reference sequences (by abundance) for this calculation.
+    We observe that both the weighted and unweighted UniFrac distances are reduced compared to the figure with the all the reference sequences.
+    This shows that the most abundant reference sequences generated by the methods are similar to each other.
+
+    \begin{figure}[H]
+      \centering
+      \includegraphics[width=1.0\linewidth]{figure_s3.pdf}
+    \end{figure}
+    \begin{figure}[H]
+      \centering
+        \caption{
+          \textbf{The UniFrac distance between the 100 most abundant reference sequences generated by various method in the \ac{dc} step}
+        }
+      \label{fig:figure_s3}
     \end{figure}

### 2022-04-21 scripts/figure_s3/figure_s3_data.py

@@ -25,9 +25,12 @@ def get_vectors(otu_1: pd.DataFrame, otu_2: pd.DataFrame, threshold: int):
     print(f"Warning {removed_samples} samples have been removed")
     common_samples = samples_1 & samples_2
     for col in common_samples:
-        otu_1_col, otu_2_col = otu_1[[col]], otu_2[[col]]
-        otu_1_col.loc[otu_1_col[col] <= threshold, :] = 0
-        otu_2_col.loc[otu_2_col[col] <= threshold, :] = 0
+        otu_1_col = (
+            otu_1[[col]].sort_values(by=[col], ascending=False).iloc[: threshold + 1, :]
+        )
+        otu_2_col = (
+            otu_2[[col]].sort_values(by=[col], ascending=False).iloc[: threshold + 1, :]
+        )
         otu_1_col.columns = ["1"]
         otu_2_col.columns = ["2"]
         joint_df = otu_1_col.join(otu_2_col, how="outer")
@@ -86,7 +89,9 @@ def abbr_name(method: str) -> str:
 @click.option(
     "--asv", default=True, type=bool, help="To display ASV along with method name"
 )
-@click.option("--threshold", default=10, type=int, help="Threshold for sequence count")
+@click.option(
+    "--threshold", default=100, type=int, help="Number of top OTUs used in calculation"
+)
 @click.option("--output", default=".", help="The path to the output directory")
 def main(trees: str, otus: str, weighted: bool, asv: bool, threshold: int, output: str):
     output_path = pathlib.Path(output)

### 2022-04-21 scripts/figure_s3/figure_s3_plot.R
Change: Figure S3 threshold to 1000

@@ -7,8 +7,8 @@ library(ggpubr)
 # inputs
 args <- commandArgs(trailingOnly = TRUE)
 if (length(args) == 0) {
-    data_folder <- "../../data/figure3/output/moving_pictures/"
-    mock_folder <- "../../data/figure3/output/"
+    data_folder <- "../../data/figure_s3/output/moving_pictures/"
+    mock_folder <- "../../data/figure_s3/output/"
     output_folder <- "../../figures/"
 } else if (length(args) == 3) {
     data_folder <- args[1]

### 2022-04-21 scripts/data_extraction/extract_figure_s4_data.py
Add: Preliminary scripts for Figure S4

@@ -1,0 +1,39 @@
+import pathlib
+import sys
+
+from data_extraction import extract_data
+
+CC_METHODS = ("remove_bimera", "uchime")
+
+
+def extract_figure_s4_data(
+    input_folder: pathlib.Path, output_folder: pathlib.Path, meta_id_list: list
+) -> None:
+    # STEP1: Loop through each TA method
+    for cc_method in CC_METHODS:
+        workflow = "denoise_cluster"
+        module = cc_method
+        process = "filtered_output"
+        # TODO: We need to loop over all dc methods
+        # TODO: We also need to change the folder names to dada2-remove_bimera or something
+        # for the tree algorithm
+        prev_process = "dada2"
+        # STEP2: Extract the data
+        output_subfolder = output_folder / module
+        extract_data(
+            input_folder,
+            workflow,
+            module,
+            process,
+            prev_process,
+            output_subfolder,
+            meta_id_list,
+        )
+
+
+if __name__ == "__main__":
+    INPUT_FOLDER = pathlib.Path(sys.argv[1])
+    DATASET = sys.argv[2]
+    META_ID_LIST = tuple(sys.argv[3:])
+    OUTPUT_FOLDER = pathlib.Path(f"../../data/figure_s4/input/{DATASET}")
+    extract_figure_s4_data(INPUT_FOLDER, OUTPUT_FOLDER, META_ID_LIST)

### 2022-04-21 scripts/figure_generation/generate_figures.sh

@@ -63,3 +63,16 @@ Rscript figure7_plot.R "../../data/figure7/output/$DATASET/" "fmt-control" "fmt-
 echo "Generating Figure S1"
 cd ../figure_s1/
 Rscript figure_s1_plot.R "../../data/figure2/output/$DATASET/" "$OUTPUT_DIR"
+
+# Figure S2
+echo "Generating Figure S2"
+cd ../figure_s2/
+Rscript figure_s2_plot.R "../../data/figure2/output/$DATASET/" "$OUTPUT_DIR"
+
+# Figure S3
+echo "Generating Figure S3"
+cd ../figure_s3/
+Rscript figure_s3_plot.R \
+  "../../data/figure3/output/$DATASET/" \
+  "../../data/figure3/output/" \
+  "$OUTPUT_DIR"

### 2022-04-21 scripts/figure_s4/figure_s4_extract_process.sh

@@ -1,0 +1,61 @@
+#!/bin/bash -l
+
+#$ -l h_rt=10:00:00
+#$ -N micone_figure_s4
+#$ -P visant
+#$ -o figure_s4_outputs.txt
+#$ -e figure_s4_errors.txt
+#$ -m e
+#$ -pe omp 8
+
+INPUT_FOLDER="/rprojectnb/visant/dkishore/FMT/outputs"
+DATASET="FMT"
+PLATFORM="scc"
+
+# Data extraction
+if [ $PLATFORM == "scc" ]; then
+  module load miniconda
+  conda activate micone-figures-py
+fi
+
+cd ../data_extraction/
+mkdir -pv ../../data/figure_s4/input/$DATASET
+python extract_figure_s4_data.py $INPUT_FOLDER $DATASET "*"
+
+# Data processing
+# TODO: Here we must change all the processing and figure scripts
+# We only need trees for <dc_method>_remove_bimera vs. <dc_method>_uchime
+# Then we just need a bar graph or something instead of heatmap
+if [ $PLATFORM == "scc" ]; then
+  module load miniconda
+  conda activate micone-qiime2
+fi
+
+cd ../figure3/
+mkdir -pv ../../data/figure_s4/intermediate/$DATASET
+mkdir -pv ../../data/figure_s4/output/$DATASET
+# Make trees
+python make_trees.py \
+  --base_dir ../../data/figure_s4/input/$DATASET \
+  --output_dir ../../data/figure_s4/intermediate/$DATASET
+# Unweighted unifrac
+python -W ignore figure_s4_data.py \
+  --trees ../../data/figure_s4/intermediate/$DATASET/trees/**/*.nwk \
+  --otus ../../data/figure_s4/input/$DATASET/**/*_filtered.biom \
+  --weighted False \
+  --threshold 10 \
+  --output ../../data/figure_s4/output/$DATASET
+# Weighted unifrac
+python -W ignore figure_s4_data.py \
+  --trees ../../data/figure_s4/intermediate/$DATASET/trees/**/*.nwk \
+  --otus ../../data/figure_s4/input/$DATASET/**/*_filtered.biom \
+  --weighted True \
+  --threshold 10 \
+  --output ../../data/figure_s4/output/$DATASET
+
+# Figure creation
+OUTPUT_DIR="../../figures/FMT/"
+Rscript figure_s4_plot.R \
[…]

### 2022-04-21 scripts/figure_s4/figure_s4_plot.R

@@ -1,0 +1,74 @@
+#!/usr/bin/env Rscript
+
+library(dplyr)
+library(ggplot2)
+library(ggpubr)
+
+# inputs
+args <- commandArgs(trailingOnly = TRUE)
+if (length(args) == 0) {
+    data_folder <- "../../data/figure_s4/output/moving_pictures/"
+    mock_folder <- "../../data/figure_s4/output/"
+    output_folder <- "../../figures/"
+} else if (length(args) == 3) {
+    data_folder <- args[1]
+    mock_folder <- args[2]
+    output_folder <- args[3]
+} else {
+    stop("Required number of arguments must equal 3")
+}
+unweighted_unifrac_csv <- paste0(data_folder, "unweighted_unifrac.csv")
+weighted_unifrac_csv <- paste0(data_folder, "weighted_unifrac.csv")
+output_file <- paste0(output_folder, "figure_s4.pdf")
+
+#########################################
+# ab
+#########################################
+
+tidy_up_data_ab <- function(unifrac_data) {
+    unifrac_data$method1 <- factor(unifrac_data$method1)
+    unifrac_data$method2 <- factor(unifrac_data$method2)
+    tidy_data <- unifrac_data %>%
+        group_by(method1, method2) %>%
+        select_at(c("method1", "method2", "unifrac")) %>%
+        summarize_all(.funs = mean) %>%
+        rowwise() %>%
+        mutate(pair = sort(c(method1, method2)) %>% paste(collapse = ",")) %>%
+        group_by(pair) %>%
+        distinct(pair, .keep_all = TRUE)
+}
+
+plot_heatmap <- function(data, title) {
+    ggplot(data = data, aes(x = method1, y = method2, fill = unifrac)) +
+        geom_tile(color = "white") +
+        geom_text(aes(label = abbreviate(unifrac))) +
+        scale_fill_distiller(
+            palette = "Spectral",
+            limits = c(0, 1),
+            breaks = c(0, 0.25, 0.5, 0.75, 1),
+            labels = c("Similar", 0.25, 0.5, 0.75, "Dissimilar")
+        ) +
+        ggtitle(title) +
+        theme_pubr() +
+        theme(
+            plot.title = element_text(hjust = 0.5),
+            axis.title = element_blank(),
+            axis.text.x = element_text(angle = 30, hjust = 1),
[…]

### 2022-04-21 sections/results.tex
Fix: Figure 3 apply threshold to mock data

@@
 
-  % TODO: Supplementary work - Figure S3
+  % FIXME: Supplementary work - Figure S3
   In order to verify this claim, for each of these methods we use only the top 100 reference sequences and recalculate the distance metrics (Figure \ref{fig:figureS3}).
@@
 
-  % TODO: Update thresholds (10) on the mock datasets and check again
   These comparisons only elucidate the pairwise similarity or dissimilarity of a pair of methods.
@@
   These high deviations are primarily in one or two of the three datasets that were analyzed and show that the datasets themselves play a big role in the performance of these methods.
-  Open-reference and denovo clustering methods perform the best under the weighted UniFrac metric and the worst under the unweighted UniFrac metric.
+  Open-reference and denovo clustering methods perform the best under the weighted UniFrac metric and the worst (marginally) under the unweighted UniFrac metric.
   This result can be attributed to the large number of low abundance representative sequences that are generated by these methods.
   Deblur performs poorly under weighted Unifrac and although its performance on the mock4 dataset is the best under unweighted UniFrac, its performance on the other datasets is average.
+  The Deblur method returns a very small number of reference sequences (2388) and this could account for the reason for the high dissimilarity with the other methods as well as irregular performance on the mock data.
   There is no method that clearly outperforms the rest in all datasets.

### 2022-04-21 scripts/figure3/figure3_data.py

@@ -72,6 +72,7 @@ def abbr_name(method: str) -> str:
         return "D2"
     if method.startswith("deblur"):
         return "DB"
+    return method
 
 
 @click.command()

### 2022-04-21 scripts/figure3/figure3_mockdata.sh

@@ -2,7 +2,7 @@
 
 DATASET=("mock4" "mock12" "mock16")
 for i in "${DATASET[@]}"; do
-  python figure3_data.py \
+  python -W ignore figure3_data.py \
     --trees "../../data/figure3/intermediate/$i/trees/**/*.nwk" \
     --otus "../../data/figure3/input/$i/**/*.biom" \
     --weighted False \
@@ -10,7 +10,7 @@ for i in "${DATASET[@]}"; do
     --threshold 10 \
     --output "../../data/figure3/output/$i"
 
-  python figure3_data.py \
+  python -W ignore figure3_data.py \
     --trees "../../data/figure3/intermediate/$i/trees/**/*.nwk" \
     --otus "../../data/figure3/input/$i/**/*.biom" \
     --weighted True \

### 2022-04-21 scripts/figure4/figure4ab_data.py
Fix: Inconsistencies in Figure 4A and 4B
1. Create labels for Figure 4A
2. Fix inconsistencies due to SILVA in Figure 4B

@@ -11,6 +11,26 @@ import pandas as pd
 LEVELS = ["Phylum", "Class", "Order", "Family", "Genus", "Species"]
 
 
+SILVA_PHYLUM_MAP = {
+    "Acidobacteriota": "Acidobacteria",
+    "Actinobacteriota": "Actinobacteria",
+    "Armatimonadota": "Armatimonadetes",
+    "Bacteroidota": "Bacteroidetes",
+    "Caldisericota": "Caldiserica",
+    "Deferribacterota": "Deferribacteres",
+    "Elusimicrobiota": "Elusimicrobia",
+    "Fibrobacterota": "Fibrobacteres",
+    "Fusobacteriota": "Fusobacteria",
+    "Gemmatimonadota": "Gemmatimonadetes",
+    "Nitrospirota": "Nitrospirae",
+    "Planctomycetota": "Planctomycetes",
+    "Spirochaetota": "Spirochaetes",
+    "Synergistota": "Synergistetes",
+    "Thermotogota": "Thermotogae",
+    "Verrucomicrobiota": "Verrucomicrobia",
+}
+
+
 def create_tables(file_paths, notus, output_path):
     tables = dict()
     for file in file_paths:
@@ -25,16 +45,27 @@ def create_tables(file_paths, notus, output_path):
             :notus
         ]
         final_df.replace("", "unknown", inplace=True)
-        final_df.index = range(1, notus + 1)
+        # # FIXME: What happened to the original OTU hash?
+        # final_df.index = range(1, notus + 1)
         final_df.index.name = "OTU"
         final_df["empty"] = [""] * notus
         tables[db_name] = final_df
     return tables
 
 
+def check_fix_tables(tables):
+    assert all(
+        tables["naive_bayes(gg_13_8_99)"].index
+        == tables["naive_bayes(silva_138_99)"].index
+    )
+    assert all(tables["naive_bayes(gg_13_8_99)"].index == tables["blast(ncbi)"].index)
+    tables["naive_bayes(silva_138_99)"].Phylum.replace(SILVA_PHYLUM_MAP, inplace=True)
+
+
 # Table headers for combined dataframes:
 # tax_level, value, assignment(match/mismatch)
 def get_mismatches(db1: pd.DataFrame, db2: pd.DataFrame) -> pd.DataFrame:
+    # FIXME: These files aren't in the same order!
     db1_lin_df = db1.drop(["empty", "Abundance"], axis=1)
[…]

### 2022-04-21 sections/results.tex
Fix: Section related to Figure 4

@@
   From Figure~\ref{fig:figure3}A one can see that the representative sequences generated by the different methods, with the exception of Deblur, are similar to each other when weighted by their abundance.
-  Figure~\ref{fig:figure3}B on the other hand shows an increase in dissimilarity between each pair of methods suggesting that the methods might differ in the assignment of sequences of low abundance.
+  Figure~\ref{fig:figure3}B on the other hand shows an increase in dissimilarity between each pair of methods suggesting that the methods might differ in the assignment of sequences of lower abundance.
 
-  % FIXME: Supplementary work - Figure S3
-  In order to verify this claim, for each of these methods we use only the top 100 reference sequences and recalculate the distance metrics (Figure \ref{fig:figureS3}).
-  We observe that the both the weighted and unweighted UniFrac distances are decreased considerably implying that the top reference sequences generated by the different methods are similar to each other.
+  In order to verify this claim, for each of these methods we use only the top 1000 representative sequences and recalculate the distance metrics (Figure \ref{fig:figureS3}).
+  We observe that the both the weighted and unweighted UniFrac distances are increased implying that the top representative sequences generated by the different methods are not similar to each other.
+  % TODO: Can we make this conclusion?
+  This leads us to the conclude that the representative sequences in the middle range of the abundance distribution are those that are most similar between the methods.
 
@@
   Deblur performs poorly under weighted Unifrac and although its performance on the mock4 dataset is the best under unweighted UniFrac, its performance on the other datasets is average.
-  The Deblur method returns a very small number of reference sequences (2388) and this could account for the reason for the high dissimilarity with the other methods as well as irregular performance on the mock data.
+  The Deblur method returns a very small number of representative sequences (2388) and this could account for the reason for the high dissimilarity with the other methods as well as irregular performance on the mock data.
   There is no method that clearly outperforms the rest in all datasets.
@@
   The representative sequences obtained from the \ac{dada2} method (with remove bimera for chimera checking) in the DC step were used for taxonomic assignment using the three reference databases.
-  Figure~\ref{fig:figure4}A depicts a flow diagram that shows how the top 50 representative sequences (sorted by abundance) are assigned a Genus according to the three different databases.
+  Figure~\ref{fig:figure4}A depicts a flow diagram (alluvial plot) that shows how the top 50 representative sequences (sorted by abundance) are assigned a Genus according to the three different databases.
   We observe that the top 50 representative sequences are assigned to 17 Genus in \ac{gg}, 23 in SILVA and 17 in \ac{ncbi}, however the sequences that are assigned to a particular Genus vary significantly based on the taxonomy database.
-  % TODO: Annotate Figure 4A with bacteria names
   We observe that not only does the assigned Genus composition vary significantly, but the percentage of unassigned representative sequences (pink) also differ.
-  Even the most abundant representative sequence is assigned to an "unknown" Genus in two of the three databases (\ac{ncbi} and \ac{gg}).
+  The large percentage of unassigned sequences indicate that even among the most abundant representative sequences a large fraction is assigned an "unknown" Genus in two of the three databases (\ac{ncbi} and \ac{gg}).
   A representative sequence might be assigned an "unknown" Genus for one of two reasons: the first is if the taxonomy identifier associated with the sequence in the database did not contain a Genus; the second (more likely) reason is that the database contains multiple sequences that are very similar to the query (representative) sequence and the consensus algorithm (from \ac{qiime2}) is unable to assign one particular Genus at the required confidence.
-  We also observe that other abundant Genus like Bacteroides and Escherichia are also assigned to different representative sequences.
-  For example, the most abundant genus in the \ac{gg} taxonomy table was \textit{Escherichia} whereas in the SILVA taxonomy table it was \textit{Escherichia-Shigella}.
-  Although these are minor differences, when comparing a large number of taxonomy composition tables these problems are hard to diagnose.
+  Additionally, we observe that other abundant Genus like Bacteroides and Escherichia are also assigned to different representative sequences (indicated by the different proporitions in the alluvial plot).
+  Furthermore, the most abundant Genus (among the top 50 representative sequences) in the \ac{gg} taxonomy table was \textit{Escherichia} whereas it was assigned to "unknown" in NCBI and in the SILVA taxonomy table it was assigned to \textit{Escherichia-Shigella}.
+  Although these are minor differences (in the case of \ac{gg} and SILVA), when comparing a large number of taxonomy composition tables these problems are hard to diagnose.
 
   After assigning all the representative sequences to taxonomies we perform a pairwise comparison of the similarity between the top 100 assignments (by abundance) from different databases at every taxonomic level (Figure~\ref{fig:figure4}B).
-  The assignments beyond Family level (Family, Genus and Species) are very dissimilar with $<67\%$ similarity between any pair of databases.
-  % TODO: Verify this
-  The assignment between SILVA and the other two databases is significantly different ($40\%$ mismatch) even at the Phylum level, indicating that 40 of the top 100 assignments at the Phylum level in the SILVA database are different in SILVA.
-  The \ac{gg} and \ac{ncbi} assignments are more similar compared to the other pairwise comparisons, mainly due to the poor performance of SIVLA at the Phylum level.
-  However, the number of mismatches for taxonomy assignments beyond the Family level are very high ($>66\%$), implying that the taxonomy assignments from each reference database are fairly unique and are largely responsible for the differences observed in the co-occurrence networks generated from different taxonomy databases.
+  The assignments beyond Family level (Family, Genus and Species) are very dissimilar with $<45\%$ similarity between any pair of databases.
+  The matches and mismatches in the assignments from all three databases are comparable, with \ac{ncbi} and SILVA being the most similar across most of the taxonomic levels, followed closed by \ac{gg} vs.
+SILVA.
+  However, the number of mismatches for taxonomy assignments beyond the Genus level (Genus, Species) are very high ($>67\%$), implying that the taxonomy assignments from each reference database are fairly unique and are largely responsible for the differences observed in the co-occurrence networks generated from different taxonomy databases.
+  The assignment between SILVA and the other two databases were originally significantly different ($40\%$ mismatch) even at the Phylum level, indicating that 40 of the top 100 assignments at the Phylum level in the SILVA database are different in SILVA.
+  However, this was fixed by making minor adjustments to the taxonomic names, such as changing Bacteroidota to Bacteroidetes in the SILVA Phylum assignments.
 

### 2022-04-21 sections/supplementary.tex
Add: Preliminary scripts for Figure S4
Fix: Section related to Figure 4

--- 2022-04-20 sections/supplementary.tex
@@
     \newpage
+
+    % TODO: Write something about TSNE
+    \begin{figure}[H]
+      \centering
+      \includegraphics[width=1.0\linewidth]{figure_s2.pdf}
+    \end{figure}
+    \begin{figure}[H]
+      \centering
+        \caption{
+          \textbf{The TSNE plot of the networks inferred by various methods}.
+        }
+      \label{fig:figure_s2}
+    \end{figure}
+    \FloatBarrier
+    \newpage
+
+  \subsection*{Similarity of high abundance reference sequences}
+
+    Figure~\ref{fig:figureS3} is similar to Figure~\ref{fig:figure3} shows the average UniFrac distance between reference sequences generated by the various methods in the \ac{dc} step.
+    In Figure~\ref{fig:figureS3} however, we use only the top 1000 representative sequences (by abundance) for this calculation.
+    We observe that both the weighted and unweighted UniFrac distances are increased compared to the figure with the all the representative sequences.
+    This shows that the most abundant representative sequences generated by the methods are not similar to each other.
+
+    \begin{figure}[H]
+      \centering
+      \includegraphics[width=1.0\linewidth]{figure_s3.pdf}
+    \end{figure}
+    \begin{figure}[H]
+      \centering
+        \caption{
+          \textbf{The UniFrac distance between the 100 most abundant representative sequences generated by various method in the \ac{dc} step}
+        }
+      \label{fig:figure_s3}
+    \end{figure}
+    \FloatBarrier
+    \newpage
+
+    % TODO: Figure S4 - Unifrac bar plot for CC methods
+    \begin{figure}[H]
+      \centering
+      \includegraphics[width=1.0\linewidth]{figure_s4.pdf}
+    \end{figure}
+    \begin{figure}[H]
+      \centering
+        \caption{
+        }
+      \label{fig:figure_s4}
+    \end{figure}
+    \FloatBarrier
+    \newpage
+
+  \subsection*{}
 

### 2022-04-21 sections/results.tex
Add: Script to generate Figure S5

@@
   However, this was fixed by making minor adjustments to the taxonomic names, such as changing Bacteroidota to Bacteroidetes in the SILVA Phylum assignments.
-
-  % TODO: Supplementary work - Figure S5
-  The comparison of all assigned genera instead of the just the top 20 contains the same percentage of matches and mismatches, implying that there does not seem to exist a correlation between abundance and mismatch.
-  This suggests that the most abundant sequences are not necessarily the ones that are consistently matched to the same taxonomies in the different reference databases.
+  The comparison of all assigned genera (Figure~\ref{fig:figure_s5}) instead of the just the top 20 contains a higher percentage of mismatches.
+  This suggests that the most abundant sequences are the ones that are consistently matched to the same taxonomies in the different reference databases, at least for the dataset tested in the current analysis.
 

### 2022-04-21 scripts/figure4/figure4_plot.R

@@ -89,31 +89,6 @@ ggsave(output_file_a, width = 11, height = 8.5)
 notu <- 50
 levels <- c("Phylum", "Class", "Order", "Family", "Genus", "Species")
 
-
-# tidy_up_data <- function(db_file, notu) {
-#     db_data <- read_csv(db_file)[1:notu, ]
-#     for (i in 1:length(levels)) {
-#         cols <- levels[1:i]
-#         db_data[levels[i]] <- apply(db_data[, cols], 1, paste, collapse = "-")
-#     }
-#     return(db_data)
-# }
-#
-# combine_data <- function(db1, db2) {
-#     comb_tbl <- as_tibble(
-#         rbind(t(colSums(db1 == db2)), t(colSums(db1 != db2)))
-#     )
-#     comb_tbl$type <- c("matches", "mismatches")
-#     final_tbl <- comb_tbl %>%
-#         gather(levels, key = "tax_level", value = "value") %>%
-#         select_at(.vars = c("tax_level", "value", "type"))
-#     # final_tbl$value = final_tbl$value / nrow(db1) * 100
-#     # if we want to display the numbers instead
-#     final_tbl$value <- final_tbl$value
-#     return(final_tbl)
-# }
-
-
 read_paired_data <- function(data_file) {
     raw_df <- read.csv(data_file, header = TRUE, sep = ",", na.strings = "")
     raw_df

### 2022-04-21 scripts/figure_s5/figure_s5_plot.R

@@ -1,0 +1,76 @@
+#!/usr/bin/env Rscript
+
+library(tidyverse)
+library(ggplot2)
+library(ggpubr)
+library(ggalluvial)
+library(ggrepel)
+
+# inputs
+args <- commandArgs(trailingOnly = TRUE)
+if (length(args) == 0) {
+    data_folder <- "../../data/figure_s5/output/moving_pictures/"
+    mock_folder <- "../../data/figure_s5/output/"
+    output_folder <- "../../figures/"
+} else if (length(args) == 3) {
+    data_folder <- args[1]
+    mock_folder <- args[2]
+    output_folder <- args[3]
+} else {
+    stop("Required number of arguments must equal 3")
+}
+gg_silva_csv <- paste0(data_folder, "gg_silva.csv")
+gg_ncbi_csv <- paste0(data_folder, "gg_ncbi.csv")
+ncbi_silva_csv <- paste0(data_folder, "ncbi_silva.csv")
+output_file <- paste0(output_folder, "figure_s5.pdf")
+
+
+################################################################################
+# Figure 4b
+################################################################################
+# Parameters
+notu <- 50
+levels <- c("Phylum", "Class", "Order", "Family", "Genus", "Species")
+
+read_paired_data <- function(data_file) {
+    raw_df <- read.csv(data_file, header = TRUE, sep = ",", na.strings = "")
+    raw_df
+}
+
+
+make_bar_plot <- function(data, title) {
+    ggbarplot(
+        data,
+        x = "tax_level",
+        y = "value",
+        fill = "assignment",
+        color = "assignment",
+        label = TRUE,
+        lab.pos = "in",
+        title = title,
+        xlab = "Taxonomy level",
+        ylab = "Number of Assignments",
+        palette = c("#00AFBB", "#FC4E07"),
+        # palette="Paired"
+    ) +
+        theme(
+        plot.title = element_text(size=10),
+        axis.text.x = element_text(angle = 30, hjust = 1)
+    )
+}
+
+# Combinations
+gg_silva <- read_paired_data(gg_silva_csv)
+gg_ncbi <- read_paired_data(gg_ncbi_csv)
+ncbi_silva <- read_paired_data(ncbi_silva_csv)
+
[…]

### 2022-04-21 scripts/figure_s5/figure_s5_process.sh

@@ -1,0 +1,10 @@
+#!/usr/bin/env bash
+
+DATASET="FMT"
+
+cd ../figure4/
+mkdir -pv ../../data/figure_s5/output/$DATASET
+python figure4ab_data.py \
+  --files "../../data/figure4/input/$DATASET/**/otu_table_wtax.biom" \
+  --notus -1 \
+  --output "../../data/figure_s5/output/$DATASET"

### 2022-04-22 sections/results.tex
Change: Update TA section with supplementary info

@@
   This result highlights the importance of the taxonomy assignment step in the 16S data analysis workflow, implying that a change in the reference database will result in drastically different inferred networks.
-  % TODO: Verify this: Figure S4
-  The reason for this seems to derive from the mislabelling of constitutive taxa that are present in high abundance in the community, which drastically alter the nodes and hence the underlying network topology.
+  The reason for this seems to derive from the differential labeling of constitutive representative sequences into taxonomic entities (Figure~\ref{fig:figure4} and Figure~\ref{fig:figure_s5}), which drastically alter the nodes and hence the underlying network topology.
   In fact the variability induced by taxonomy assignment is much more significant than that due to the variability induced based on how the reference sequences themselves are identified (in the DC step).
@@
   However, this was fixed by making minor adjustments to the taxonomic names, such as changing Bacteroidota to Bacteroidetes in the SILVA Phylum assignments.
-  The comparison of all assigned genera (Figure~\ref{fig:figure_s5}) instead of the just the top 20 contains a higher percentage of mismatches.
+  The comparison of all assigned genera (Figure~\ref{fig:figure_s5}) instead of the just the top 100 contains a higher percentage of mismatches.
   This suggests that the most abundant sequences are the ones that are consistently matched to the same taxonomies in the different reference databases, at least for the dataset tested in the current analysis.
@@
   Since the \ac{gg} database is more widely used in the field~[ref], to enable easier comparison with published networks in literature, we set \ac{gg} as the \ac{micone} standard for taxonomy assignment.
-  On the other hand, since the degree of similarity of assignments between \ac{ncbi} and \ac{gg} is fairly high, and due to its regular updation and maintenance the \ac{ncbi} database is also a good choice for the user.
+  On the other hand, since the degree of similarity of assignments between \ac{ncbi} and \ac{gg} is fairly high (Figure~\ref{fig:figure_s5}), and due to its regular updation and maintenance the \ac{ncbi} database is also a good choice for the user.
 

### 2022-04-22 sections/supplementary.tex
Add: Script to generate Figure S5
Change: Update TA section with supplementary info

@@
 
-  \subsection*{}
+  \subsection*{Mismatches in all assignments at the \ac{TA} step}
+
+    % TODO: Elaborate on this and work on the caption
+    Figure~\ref{fig:figure_s5} is similar to Figure~\ref{fig:figure4}B, but instead of the matching only the top 100 taxonomic entities (by abundance), we match all the assignments from one database with those from the other two databases.
+    We observe that the percentage of mismatches are indeed higher when considering all the assignments, implying that matching of the taxonomies in the more abundant sequences are more consistent.
+
+    \begin{figure}[H]
+      \centering
+      \includegraphics[width=1.0\linewidth]{figure_s5.pdf}
+    \end{figure}
+    \begin{figure}[H]
+      \centering
+        \caption{
+        }
+      \label{fig:figure_s5}
+    \end{figure}
+    \FloatBarrier
+    \newpage
 

### 2022-04-22 sections/methods.tex
Add: Supplementary section for consensus network

@@
 
- % TODO: Use bold for matrices and lower-case bold for vectors
   This section defines the notation used below to describe the consensus network algorithm used in the our pipeline. Note that all networks to be compared have the same number of nodes.
 
-  $w$ = number of co-occurrence networks to be integrated in the consensus network (corresponding to the number of the different approaches used to infer co-occurrence networks)
+  $w$, the number of co-occurrence networks to be integrated in the consensus network (corresponding to the number of the different approaches used to infer co-occurrence networks)
 
-  $q$ = number of nodes in any co-occurrence network (assumed to be the same across all $w$ networks)
+  $q$, the number of nodes in any co-occurrence network (same across all $w$ networks)
 
-  $N^i$ = matrix of network edge Weights for the $i^{th}$ co-occurrence network. This is a $q \times q$ matrix, where $i \in \{1,\dots,w\}$. $N^i_{a,b}$ represents edge $(a,b)$ in network $i$
+  $N^i$, the matrix of network edge weights for the $i^{th}$ co-occurrence network. This is a $q \times q$ matrix, where $i \in \{1,\dots,w\}$. $N^i_{a,b}$ represents edge $(a,b)$ in network $i$
 
-  $P^i$ = matrix of p-values for all edges of the $i^{th}$ co-occurrence network. This is a $q \times q$ matrix, where $i \in \{1,\dots,q\}$
+  $P^i$, the matrix of p-values for all edges of the $i^{th}$ co-occurrence network. This is a $q \times q$ matrix, where $i \in \{1,\dots,q\}$
 
@@
 
-  % TODO: Is this necessary here, should this be elaborated?
-  As described in more detail in the Supplementary Material and in the original reference~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016}, the final combined p-value is given by:
+  As described in more detail in the Supplementary section and in the original reference~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016}, the final combined p-value is given by:
   \begin{equation}
@@
         & \hat{P}_j = 1.0 - \Phi_{2f}\left( \psi / c \right) \\
-        \text{where},~ &\psi = -2 \sum_{i=1}^k \log(\bar{P}^i_j) ~~~\text{and}~~~ \Phi_{2f} = \mathrm{CDF}\left( \chi^2_{2f} \right)
+        \text{where},~ &\psi = -2 \sum_{i=1}^w \log(\bar{P}^i_j) ~~~\text{and}~~~ \Phi_{2f} = \mathrm{CDF}\left( \chi^2_{2f} \right)
     \end{aligned}
@@
 
-  % FIXME: Not sure what our stance on this is?
-  Note that we do not use Pearson and Spearman methods in the p-value merging step and these algorithms are only used for demonstration and comparison.
+  Note that we do not use Pearson and Spearman methods in the p-value merging step to calculate the consensus network, these algorithms are only used for demonstration and comparison.
   The combined p-values are used to threshold for significance right before the consensus algorithm is applied to the inferred networks.
@@
   \end{equation*}
+  where, $#$ refers to the cardinality of the set.
 
@@
   The simple voting method gives the union of the networks when $p \in (0, \frac{1}{w})$ and will return the intersection when $p \in ((w - 1), 1)$.
-  In general, if $p \in ((n - 1), \frac{n}{w})$, this algorithm will report an edge in the consensus network when at least $n$ network inference methods report this edge. [REFERENCE]
+  In general, if $p \in ((n - 1), \frac{n}{w})$, this algorithm will report an edge in the consensus network when at least $n$ network inference methods report this edge.
 
@@
   \begin{equation}
-    s_j = \sum_i \bar{S}^i_j
+    s_j = \sum_{i=1}^{w} \bar{S}^i_j
     \label{eqn:scaled-sum}
@@
   This section defines notation for the network variability analysis in our study\dots
+
   % TODO: Update W and Q in the entire subsection

### 2022-04-22 sections/results.tex
Add: Supplementary section for consensus network

@@
 
-  % TODO: Update after getting SparCC results
   In order to obtain a quantitative picture of the differences between the inferred networks, we analyzed the distribution of common nodes and edges (Figure~\ref{fig:figure5} B and C) using UpSet plots~\cite{Lex}.
   Only \ac{harmonies}, \ac{cozine},\ac{spieceasi}, \ac{spring}, propr, and \ac{sparcc} are used in the comparison since Pearson and Spearman add a large number of spurious edges since they are not intended for compositional datasets).
-  The results for the node intersections show that the networks have a large number of nodes in common ($33$ out of $68$ total nodes) and no network possesses any unique node.
-  The edge intersections in contrast show that only $8$ edges (out of $202$ total edges) are in common between all the methods and each network has a large number of unique edges.
+  The results for the node intersections show that the networks have a large number of nodes in common ($33$ out of $68$ total unique nodes) and no network possesses any unique node.
+  The edge intersections in contrast show that only $8$ edges (out of $202$ total unique edges) are in common between all the methods and each network has a large number of unique edges.
   These results indicate that there is a substantial rewiring of connections in the inferred networks.
-
-  % TODO: What is the final concluding statement here? How to segway to the next section
+  Hence, in order to determine the interactions that are most likely to exist in the dataset we advocate for the construction of a consensus algorithm.
 
@@
   Therefore, we have developed a consensus method that would take into consideration the evidence offered by each network inference algorithm and generate a consensus network that would contain the most probable interactions for the input data.
-  % TODO: Should we include Pearson and Spearman in the consensus (currently we do)!
   We have developed two algorithms for the generation of the consensus network, the simple voting and the scaled-sum methods (see Methods).
+  We only include \ac{sparcc} and propr in the consensus network calculations, for consensus networks that include Spearman and Pearson see Figure~\ref{fig:figure_s6} in the Supplementary section.
   First, for the correlation-based methods we merge the p-values using the Brown's p-value merging method~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016,faustCoNetAppInference2016} (see Methods).
@@
 
-  % TODO: Supplementary work: Figure S6
+  % TODO: Supplementary work: Figure S7
   The precision and sensitivity graphs on the seqtime data can be found in the Supplementary section.
@@
   The filtering in the \ac{OP} step when turned off also adds a large number of new nodes, but the edges between the common nodes (with the default network) are fairly conserved.
+  % TODO: Say that this shows how important the consensus network is in the whole pipeline
   Furthermore, the L1 distance of networks generated by altering one of the steps of the pipeline from the default against the default network (Figure~\ref{fig:figure7}B) shows that the biggest deviations from the default network occur when the \ac{ni} steps are changed.

### 2022-04-22 sections/supplementary.tex
Add: Supplementary section for consensus network

@@
 
+  % QUESTION: Should the CCA section be only included in the response letter?
+
   \subsection*{Similarity of high abundance reference sequences}
@@
 
+  % TODO: Create table for this section
+  % Should this be in the main manuscript?
+  \subsection*{Network metrics}
+
   \subsection*{Consensus and p-value}
 
-  % FIXME: Update according to the new notation in methods section
-  \subsubsection*{p-value merging}
-  Fisher~\cite{fisher_224a_1948} proposed that for $k$ independent p-values, each generated by $k$ different methods and denoted by $P_i$, the statistic $\Psi$:
+  \subsection*{p-value merging}
+
+  Fisher~\cite{fisher_224a_1948} proposed that for $w$ independent p-values, each generated by $w$ different methods and denoted by $\bar{P}^i$, the following will hold true for the statistic $\Psi$:
   \begin{equation*}
     \begin{aligned}
-        \Psi &= \sum_{i=1}^k -2 \log \left( P_i \right) \\
-        \Psi &\sim \chi^2_{2k}
+      \Psi &= \sum_{i=1}^w -2 \log \left( \bar{P}^i \right) \\
+        \Psi &\sim \chi^2_{2w}
     \end{aligned}
@@
   \begin{equation*}
-    f = \frac{\mathrm{E}[\Psi]^2}{\mathrm{Var}[\Psi]} ~~~\text{and}~~~ c = \frac{\mathrm{Var}[\Psi]}{2\mathrm{E}[\Psi]} = \frac{k}{f}
+    f = \frac{\mathrm{E}[\Psi]^2}{\mathrm{Var}[\Psi]} ~~~\text{and}~~~ c = \frac{\mathrm{Var}[\Psi]}{2\mathrm{E}[\Psi]} = \frac{w}{f}
   \end{equation*}
 
-  Furthermore, Brown showed that $\mathrm{E}[\Psi]$ and $\mathrm{Var}[\Psi]$ can be calculated directly via a numerical integration:
+  Furthermore, Brown showed that $\mathrm{E}[\Psi]$ and $\mathrm{Var}[\Psi]$ can be calculated via a numerical integration:
   \begin{equation*}
-    \mathrm{E}[\Psi] = 2k ~~~\text{and}~~~ \mathrm{Var}[\Psi] = 4k + 2\sum_{i<j} \mathrm{Cov}\left( -2\log(P_i), -2\log(P_j) \right)
+    \mathrm{E}[\Psi] = 2w ~~~\text{and}~~~ \mathrm{Var}[\Psi] = 4w + 2\sum_{i<j} \mathrm{Cov}\left( -2\log(\bar{P}^i), -2\log(\bar{P}^j) \right)
   \end{equation*}
@@
   \begin{equation}
-    \mathrm{Cov}\left( -2\log(P_i), -2\log(P_j) \right) \approx 3.263 \rho_{ij} + 0.710 \rho_{ij}^2 + 0.027 \rho_{ij}^3
+    \mathrm{Cov}\left( -2\log(\bar{P}^i), -2\log(\bar{P}^j) \right) \approx 3.263 \rho_{ij} + 0.710 \rho_{ij}^2 + 0.027 \rho_{ij}^3
     \label{eqn:covariance-pvalues}
@@
     \begin{aligned}
-        & P_{combined} = 1.0 - \Phi_{2f}\left( \psi / c \right) \\
-        \text{where},~ &\psi = -2 \sum_{i=1}^k \log(P_i) ~~~\text{and}~~~ \Phi_{2f} = \mathrm{CDF}\left( \chi^2_{2f} \right)
+        & \hat{P}_j = 1.0 - \Phi_{2f}\left( \psi / c \right) \\
+        \text{where},~ &\psi = -2 \sum_{i=1}^w \log(\bar{P}^i_j) ~~~\text{and}~~~ \Phi_{2f} = \mathrm{CDF}\left( \chi^2_{2f} \right)
     \end{aligned}
@@
 
-  The p-value merging method in \ac{micone} (refer Documentation) uses Equation~\ref{eqn:covariance-pvalues} to estimate the covariance of the pvalues and Equation~\ref{eqn:pvalue-combined} to merge the p-values (obtained from bootstrapping) from the different correlation methods.
-  % FIXME: Not sure what our stance on this is?
+  The p-value merging and consensus method in \ac{micone} (refer Documentation) uses Equation~\ref{eqn:covariance-pvalues} to estimate the covariance of the pvalues and Equation~\ref{eqn:pvalue-combined} to merge the p-values (obtained from bootstrapping) from the different correlation methods.
   Note that we do not use Pearson and Spearman methods in the p-value merging step and these algorithms are only used for demonstration and comparison.
   The combined p-values are used to threshold for significance during the consensus network step.
+
+  \subsection*{Performance of consensus network on NorTA data}
+
+    \begin{figure}[H]
+      \centering
+      \includegraphics[width=1.0\linewidth]{figure_s6.pdf}
+    \end{figure}
+    \begin{figure}[H]
+      \centering
+        \caption{
+        }
+      \label{fig:figure_s6}
+    \end{figure}
+    \FloatBarrier
+    \newpage
+
+  \subsection*{Performance of consensus network on seqtime data}
+
+    \begin{figure}[H]
+      \centering
+      \includegraphics[width=1.0\linewidth]{figure_s7.pdf}
+    \end{figure}
+    \begin{figure}[H]
+      \centering
+        \caption{
+        }
+      \label{fig:figure_s7}
+    \end{figure}
+    \FloatBarrier
+    \newpage
 

### 2022-04-22 sections/results.tex
Add: Supplementary section for seqtime data

@@
   Similar to the previous steps of the pipeline, where were we evaluated the performance of methods on mock datasets, we can evaluate the network inference algorithms and the final consensus network using synthetic interaction data.
-  For this purpose, we generated synthetic interaction data using the ``NorTA''[ref] and ``seqtime''[ref] methods (see Methods), which have been previously used for benchmarking network inference methods.
+  For this purpose, we generated synthetic interaction data using the ``NorTA''~\cite{Kurtz2015} and ``seqtime''~\cite{faustSignaturesEcologicalProcesses2018} methods (see Methods), which have been previously used for benchmarking network inference methods.
   For each method, an \ac{otu} counts table is generated based on the selected interaction network topology and properties.
@@
 
-  % TODO: Supplementary work: Figure S7
-  The precision and sensitivity graphs on the seqtime data can be found in the Supplementary section.
+  Furthermore, we also find that the consensus algorithm performs drastically better in terms of precision (0.820 for scaled-sum method vs.
+0.624 for \ac{spieceasi}) in recapturing the orginal interactions in the data.
+  The precision and sensitivity graphs on the seqtime data can be found in the Supplementary section (Figure~\ref{fig:figure_s7}).
 
@@
   Furthermore, the L1 distance of networks generated by altering one of the steps of the pipeline from the default against the default network (Figure~\ref{fig:figure7}B) shows that the biggest deviations from the default network occur when the \ac{ni} steps are changed.
+  This shows us that not only does the consensus help reconcile the variability in the edges due to multiple network inference algorithms, it also helps reconcile the variability to the other steps of the pipline as well.
   This result is different from that observed in Figure~\ref{fig:figure2} since in this particular analyses all the steps other than the \ac{ni} step have the consensus algorithm applied to them, which drastically reduces the edges in the graph that have lower probabilities.

### 2022-04-22 sections/supplementary.tex
Add: Supplementary section for seqtime data

@@
 
-  \subsection*{Performance of consensus network on NorTA data}
+  \subsection*{Performance of alternate consensus network on NorTA data}
+
+    In Figure~\ref{fig:figure_s6} we calculate the consensus network by also taking into account the networks inferred by the Pearson and Spearman methods.
+    We see that the consensus networks in this case do not perform as well as those networks that did not incorporate these two algorithms.
+    However, we still observe that the consensus network at some parameter values still outperforms the best network inference method (\ac{spieceasi}).
 
@@
 
-  \subsection*{Performance of consensus network on seqtime data}
+  \subsection*{Performance of network inference algorithms on seqtime data}
+
+    % TODO: Elaborate on this
+    In Figure~\ref{fig:figure_s7} we perform the same analysis we did with the NorTA data using the seqtime data.
+    We observe that the best performing network inference algorithm is \ac{spieceasi}, but the consensus algorithms, both the scaled-sum and simple voting methods outperform it in terms of precision.
+    These results show that the scaled-sum method is much better suited for inferring robust and accurate interactions from count data regardless of the distributions, topologies and number of samples in the data.
 
@@
 
-  % TODO: Write all these sections
-  \subsection*{Performance of network inference algorithms on seqtime data}
+  % TODO: Write more about this and link to documentation
   \subsection*{The JSON network format}
 
-  \begin{figure}[h]
-  \centering
-  \includegraphics[width=\linewidth]{figureS1.pdf}
-  \caption{
-    \textbf{Comparison of various denoising and clustering algorithms used in the pipeline}.
-    (A, B) Correlation of the abundances of the taxa that are in common between the count matrices created by two different methods.
-    (A) The worst correlation (least similar methods) is between open-reference and dada2.
-    (B) The best correlation (most similar methods) is between open-reference and denovo.
-    (C) A heatmap showing the $\mathrm{R}^2$ of all pairwise comparisons of the methods.
-  }
-  \label{fig:figureS1}
-\end{figure}
-
-  \begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{figureS2.pdf}
-    \caption{
-      \textbf{Heatmaps showing the weighted and unweighted unifrac distances for the hard palate dataset analysis}.
-      (A) weighted unifrac distances and (B) unweighted unifrac distances between the representative sequences generated by different denoising and clustering algorithms.
-      These results are in agreement with the stool microbiome dataset.
-    }
-    \label{fig:figureS2}
-  \end{figure}
-
-  \begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{figureS3.pdf}
-    \caption{
-      \textbf{The distributions of the average weighted UniFrac distance between the expected sequence profile and the calculated sequence profile in the synthetic datasets}.
-      We observe no significant difference between the various methods on the synthetic datasets used for this study.
-    }
-    \label{fig:figureS3}
-  \end{figure}
-
-%   \begin{figure}[h]
-%     \centering
-%     \includegraphics[width=0.9\linewidth]{pdf/all_denoise_reg.pdf}
-%     \caption*{All pairwise correlations comparing the similarity between different denoising and clustering methods}
-%     \label{fig:figureS4}
-%   \end{figure}
-
-  \begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{figureS4.pdf}
-    \caption{
-      \textbf{(A)} Taxonomy composition of the 20 most abundant genera predicted for the stool microbiome dataset generated using different taxonomy references databases: Greengenes, SILVA and NCBI.
-      The legend shows the common and the unique genera among the taxonomy assignments.
-  }
-    \label{fig:figureS4}
-  \end{figure}
-
-  \begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{figureS5.pdf}
-    \caption{
-      The bray-curtis dissmilarity between the expected taxonomic composition and generated taxonomic composiion for the synthetic datasets.
-  }
-  \label{fig:figureS5}
-  \end{figure}
-
-%   \begin{figure}[h]
-%     \centering
-%     \includegraphics[width=\linewidth]{figureS6.pdf}
-%     \caption{
-%       Calculation of presence threshold that is applied on the OTU table in the OTU processing (OP) step of the pipeline.
-%       This presence threhold $p_t$ is dependent on the number of samples in the dataset and the required correlation stength threshold.
-%   }
-%     \label{fig:figureS6}
-%   \end{figure}
-
-  \begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{figureS8.pdf}
-    \caption{
-      The similarity between the networks generated using the different network inference algorithms for stool dataset (A) and the hard palate dataset (B).
-      The similarity between the various methods was found to vary with the dataset used.
-  }
-    \label{fig:figureS8}
-  \end{figure}
-
-  \begin{figure}[h]
-    \centering
-    \includegraphics[width=0.8\linewidth]{pdf/denoise_network.pdf}
-    \caption{A network showing union (A) and intersection (B) of networks generated using different denoising and clustering tools on the Stool dataset.}
-    \label{fig:figureS5}
-  \end{figure}
+    The default format

### 2022-04-22 sections/results.tex
Add: Section for CC step in the supplementary

@@
 
-  % TODO: Supplementary work - Figure S4
   After the denoising, the sequences are subject to chimera checking.
@@
   The uchime method is from \ac{qiime2}, whereas the remove bimera method is from the \ac{dada2} R package.
-  We did not notice any significant difference between the two chimera checking methods, implying that they identify and remove mostly the same set of sequences as chimeras (Figure ???).
+  We did not notice any significant difference between the two chimera checking methods (Figure~\ref{fig:figure_s4}), implying that they identify and remove mostly the same set of sequences as chimeras.
 

### 2022-04-22 sections/supplementary.tex
Add: Section for CC step in the supplementary

@@
 
-    % TODO: Figure S4 - Unifrac bar plot for CC methods
+    We find that the two chimera checking methods, uchime and remove bimera, produce similar outputs.
+    This can be verified by comparing the output (representative sequence distributions) of the two chimera checking methods for each denoising method (Figure~\ref{fig:figure_s4}).
+    We find that with the exception of de novo and open-reference under the unweighted UniFrac metric, all the other methods have low dissimilarity.
+    This is especially true for the \ac{dada2} and Deblur methods which are the recommended denoising methods in the \ac{micone} pipeline.
+    % TODO: Can we verify this?
+    Therefore, we recommended remove bimera as the default chimera method if one is using \ac{dada2} and uchime when one is using Deblur, since these methods were developed for these respective algorithms.
+
     \begin{figure}[H]

### 2022-04-22 scripts/figure_s4/figure_s4_data.py

@@ -112,7 +112,7 @@ def main(trees: str, otus: str, weighted: bool, asv: bool, threshold: int, outpu
         abbr = abbr_name(dc_method)
         for col in unifrac_data.index:
             if asv:
-                label = f"{abbr} ({otu_count})"
+                label = f"{abbr}"
             else:
                 label = f"{abbr}"
             data.append(

### 2022-04-22 scripts/figure_s4/figure_s4_extract_process.sh

@@ -37,15 +37,15 @@ python figure_s4_make_trees.py \
   --output_dir ../../data/figure_s4/intermediate/$DATASET
 # Unweighted unifrac
 python -W ignore figure_s4_data.py \
-  --trees ../../data/figure_s4/intermediate/$DATASET/trees/**/*.nwk \
-  --otus ../../data/figure_s4/input/$DATASET/**/*_filtered.biom \
+  --trees "../../data/figure_s4/intermediate/$DATASET/trees/**/*.nwk" \
+  --otus "../../data/figure_s4/input/$DATASET/**/*_filtered.biom" \
   --weighted False \
   --threshold 10 \
-  --output ../../data/figure_s4/output/$DATASET
+  --output "../../data/figure_s4/output/$DATASET"
 # Weighted unifrac
 python -W ignore figure_s4_data.py \
-  --trees ../../data/figure_s4/intermediate/$DATASET/trees/**/*.nwk \
-  --otus ../../data/figure_s4/input/$DATASET/**/*_filtered.biom \
+  --trees "../../data/figure_s4/intermediate/$DATASET/trees/**/*.nwk" \
+  --otus "../../data/figure_s4/input/$DATASET/**/*_filtered.biom" \
   --weighted True \
   --threshold 10 \
-  --output ../../data/figure_s4/output/$DATASET
+  --output "../../data/figure_s4/output/$DATASET"

### 2022-04-22 scripts/figure_s4/figure_s4_plot.R

@@ -7,12 +7,10 @@ library(ggpubr)
 # inputs
 args <- commandArgs(trailingOnly = TRUE)
 if (length(args) == 0) {
-    data_folder <- "../../data/figure_s4/output/moving_pictures/"
-    mock_folder <- "../../data/figure_s4/output/"
-    output_folder <- "../../figures/"
+    data_folder <- "../../data/figure_s4/output/FMT/"
+    output_folder <- "../../figures/FMT/"
 } else if (length(args) == 3) {
     data_folder <- args[1]
-    mock_folder <- args[2]
     output_folder <- args[3]
 } else {
     stop("Required number of arguments must equal 3")
@@ -26,8 +24,7 @@ output_file <- paste0(output_folder, "figure_s4.pdf")
 #########################################
 
 tidy_up_data <- function(unifrac_data) {
-    unifrac_data$method1 <- factor(unifrac_data$method1)
-    unifrac_data$method2 <- factor(unifrac_data$method2)
+    unifrac_data$method <- factor(unifrac_data$method)
     tidy_data <- unifrac_data %>%
         group_by(method) %>%
         select_at(c("method", "unifrac")) %>%
@@ -36,26 +33,10 @@ tidy_up_data <- function(unifrac_data) {
 }
 
 plot_dotplot <- function(data, title) {
-    ggdotchart(data, x = "method", y ="unifrac")
-    # TODO: Update from here
-    ggplot(data = data, aes(x = method1, y = method2, fill = unifrac)) +
-        geom_tile(color = "white") +
-        geom_text(aes(label = abbreviate(unifrac))) +
-        scale_fill_distiller(
-            palette = "Spectral",
-            limits = c(0, 1),
-            breaks = c(0, 0.25, 0.5, 0.75, 1),
-            labels = c("Similar", 0.25, 0.5, 0.75, "Dissimilar")
-        ) +
+    ggbarplot(data, x = "method", y ="unifrac", fill="steelblue", label = TRUE, lab.pos = "out", lab.nb.digits = 3) +
+        ylim(0, 1) +
         ggtitle(title) +
-        theme_pubr() +
-        theme(
-            plot.title = element_text(hjust = 0.5),
-            axis.title = element_blank(),
-            axis.text.x = element_text(angle = 30, hjust = 1),
-            axis.text.y = element_text(angle = 30)
-        ) +
[…]

### 2022-04-25 sections/supplementary.tex
Add: Section about network metrics to supplementary

@@
 
-  % TODO: Create table for this section
-  % Should this be in the main manuscript?
   \subsection*{Network metrics}
 
-  \subsection*{Consensus and p-value}
+  In Table~\ref{tab:network_metrics} we show various global network metrics calculated for each tool in the pipeline.
+  All the networks that make use of a particular tool are grouped together and the average metrics are calculated for each group.
+  All the metrics were calculated using the \texttt{networkx} Python package~\cite{hagbergExploringNetworkStructure2008}.
+  \begin{enumerate}
+    \item The average shortest path length describes the average of all the shortest paths in the graph.
+No number is reported if the graph is not connected, thereforce, none of the networks that make use of \ac{harmonies}, \ac{cozine}, \ac{spring}, \ac{spieceasi} and Pearson are connected.
+    \item The average clustering is the average clustering coefficient of the graph.
+The closer the value is to 1.0, the more densely connected is the graph.
+We can observe that the networks that use correlation-based methods have the highest values while the direct-association based methods have the lowest.
+    \item The number of connected components are the highest for the direct-association based methods and the lowest for the correlation-based methods.
+In the case of propr, all the networks have only one giant component.
+    \item Modularity metric is the average modularity over all partitions in a graph calculated using a label propagation algorithm.
+Positive values imply that there are more edges between vertices of the same type than we would expect by chance, and negative means that there are less.
+The networks inferred by \ac{mldm} report very few edges, and skew the average modularity scores.
+    \item Node connectivity refers to the minimum number of nodes that must be removed from the graph to make it disconnected.
+We observe that only the networks generated using propr have a high value since most of these networks are connected.
+    \item Edge connectivity refers to the minimum number of edges that must be removed from the graph to make it disconnected. These values are similar to node connectivity.
+    \item Degree assortativity coefficient measures the similarity of connections in the graph with respect to the node degree.
+Again we observe that the direct-association based methods have negative degree assortativity, meaning that there are many hubs in these networks.
+The correlation-based methods have positive values implying that in these networks nodes with similar degrees attach to one another.
+  \end{enumerate}
+
+  \begin{table}[]
+  \begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|l|}
+  \hline
+  DC & CC            & TA                & OP          & NI         & Average shortest path length & Average clustering & No.
+of connected components & Modularity & Node connectivity & Edge connectivity & Degree assortativity coefficient \\ \hline
+  DB & -             & -                 & -           & -          & 2.066                        & 0.265              & 27.088                      & 0.387      & 1.868             & 1.868             & 0.070                            \\
+  OR & -             & -                 & -           & -          & 1.904                        & 0.273              & 24.868                      & -17.060    & 1.877             & 1.877             & 0.099                            \\
+  D2 & -             & -                 & -           & -          & 2.263                        & 0.267              & 28.228                      & 0.326      & 1.789             & 1.789             & 0.078                            \\
+  DN & -             & -                 & -           & -          & 2.065                        & 0.272              & 25.132                      & 0.302      & 1.754             & 1.754             & 0.093                            \\
+  CR & -             & -                 & -           & -          & 1.937                        & 0.276              & 24.289                      & -1.865     & 2.079             & 2.079             & 0.097                            \\
+  -  & remove bimera & -                 & -           & -          & 2.070                        & 0.272              & 25.968                      & -7.439     & 1.863             & 1.863             & 0.086                            \\
+  -  & uchime        & -                 & -           & -          & 2.066                        & 0.269              & 25.874                      & 0.275      & 1.884             & 1.884             & 0.089                            \\
+  -  & -             & BLAST(NCBI)       & -           & -          & 2.218                        & 0.262              & 22.784                      & -3.683     & 1.495             & 1.495             & 0.100                            \\
+  -  & -             & NaiveBayes(SILVA) & -           & -          & 1.872                        & 0.275              & 30.716                      & 0.415      & 2.674             & 2.674             & 0.074                            \\
+  -  & -             & NaiveBayes(GG)    & -           & -          & 2.086                        & 0.274              & 24.263                      & -7.478     & 1.453             & 1.453             & 0.088                            \\
+  -  & -             & -                 & Filter(on)  & -          & 1.975                        & 0.252              & 18.493                      & -7.194     & 1.880             & 1.880             & 0.073                            \\
+  -  & -             & -                 & Filter(off) & -          & 2.193                        & 0.291              & 34.174                      & 0.431      & 1.867             & 1.867             & 0.104                            \\
+  -  & -             & -                 & -           & propr      & 1.650                        & 0.528              & 1.000                       & 0.000      & 10.583            & 10.583            & 0.036                            \\
+  -  & -             & -                 & -           & harmonies  &                              & 0.108              & 82.633                      & 0.757      & 0.000             & 0.000             & 0.340                            \\
+  -  & -             & -                 & -           & flashweave & 3.878                        & 0.079              & 1.800                       & 0.458      & 0.383             & 0.383             & -0.032                           \\
+  -  & -             & -                 & -           & cozine     &                              & 0.090              & 82.167                      & 0.694      & 0.000             & 0.000             & -0.072                           \\
+  -  & -             & -                 & -           & spring     &                              & 0.092              & 11.767                      & 0.566      & 0.000             & 0.000             & -0.035                           \\
+  -  & -             & -                 & -           & sparcc     & 1.747                        & 0.525              & 1.033                       & 0.009      & 5.500             & 5.500             & -0.011                           \\
+  -  & -             & -                 & -           & spieceasi  &                              & 0.076              & 39.383                      & 0.717      & 0.000             & 0.000             & -0.031                           \\
+  -  & -             & -                 & -           & pearson    &                              & 0.457              & 23.883                      & 0.551      & 0.000             & 0.000             & 0.478                            \\
+  -  & -             & -                 & -           & spearman   & 1.875                        & 0.551              & 1.450                       & 0.044      & 1.117             & 1.117             & 0.170                            \\
+  -  & -             & -                 & -           & mldm       & 2.760                        & 0.130              & 2.267                       & -75.648    & 0.433             & 0.433             & -0.028                           \\ \hline
+  \end{tabular}
+  \caption{
+    \textbf{Table of global network metrics for all combinations of networks}.
+    The metrics are averaged for each tool in every step of the pipeline, implying that all combinations of networks that use a particular tool are grouped and averaged and their metrics are reported.
+  }
+  \label{tab:network_metrics}
+  \end{table}
 

### 2022-04-25 scripts/network_metrics/calculate_network_metrics.py

@@ -171,7 +171,7 @@ def main(
 
     print(df.head())
     df_summary = summarize_metrics(df)
-    df_summary.to_csv(output_path / "metrics_summary.csv", sep=",")
+    df_summary.to_csv(output_path / "metrics_summary.csv", sep=",", float_format="%.3f")
     print(df_summary.head())
 
 

### 2022-04-25 sections/methods.tex
Add: Section about JSON network format

@@
 
-  % TODO: Write about the JSON network format (supplementary)
   The \ac{micone} pipeline provides both a Python API as well as a command-line interface and only uses a single configuration file (nextflow.config) to contain the configuration parameters.
-  The \ac{micone} Python API several biom file and network related function and methods, enabling detailed comparison of counts tables and inferred networks if desired.
+  The \ac{micone} Python API provides several biom file and network related functions and methods, enabling detailed comparison of counts tables and inferred networks if desired.
+  It also contains helper functions that can help in parsing taxonomies and communicate with the \ac{ncbi} taxonomy database to query taxonomy by name or taxonomic ids.
   The configuration file along with the run file (main.nf) lists the inputs, output and the steps to be performed during runtime, along with the parameters to be used (if different from defaults) for the various steps.
@@
   This modular architecture simplifies the process of adding new tools (refer to modules section in the \ac{micone} documentation).
+  % TODO: Write a DockerFile for the package
   % In addition to the Python package, the entire pipeline has been containerized into a Docker~\cite{Merkel1994} image (\hl{dockerhub link}) for easy deployment and setup.
@@
   In the final module of this step, the consensus algorithms are used to create the final consensus network using associations from all the network inference methods.
-  The output of this step are co-occurrence association networks encoded in the JSON format (refer Supplementary section).
+  The output of this step are co-occurrence association networks encoded in the JSON format (refer Supplementary section) and are also exportable to a variety of network formats like .gml.
 

### 2022-04-25 sections/supplementary.tex
Add: Section about JSON network format

@@
 
-  % TODO: Write more about this and link to documentation
-  \subsection*{The JSON network format}
+  % TODO: Should we mention the MIND database here?
+  \subsection*{The JSON network format and network exports}
 
-    The default format
+    The default format \ac{micone} uses for storing the network files is the JSON format.
+    The custom JSON schema we have designed is able to store all network related information from nodes and links to metadata related to links and datasets.
+    Additionally, \ac{micone} also supports exporting of networks into a variety of other formats such as edge lists, .gml and cytoscape formats.
+    Since, we make use of \texttt{networkx}~\cite{hagbergExploringNetworkStructure2008} for the export functionality, networks can be exported to all supported formats but the corresponding metadata might not be exported appropriately.
+    The details of the format and information about importing/exporting it and other network formats can be found in the \ac{micone} documentation.

### 2022-04-25 sections/methods.tex
Add: Table about OP parameter values

@@
   The quality filtering and chimera checking tools are derived from those used in \ac{qiime2} v2021.8.0 (uchime method) and \ac{dada2} (remove bimera method).
+  The list of tools used in this step, along with their modules and references are provided in Table~\ref{tab:dc_tools}.
 
@@
   The consensus algorithm used is the default method used by the classifiers in \ac{qiime2}.
+  The list of tools used in this step, along with their modules and references are provided in Table~\ref{tab:ta_tools}.
 
-  % TODO : Add information about filtering when OP=off
   \subsection*{OTU and ESV Processing (OP)}
@@
   Filtering, is performed to remove samples or features (\ac{otu}s or \ac{esv}s) from the count matrix that are sparse.
-  By default, samples are filtered out if the total reads in a sample are less than 500 and features are filtered out if the abundance of the 1\% and prevalence is less than 5\%.
+  By default, when the OP module is ``on'', the samples are filtered out if the total reads in a sample are less than 500 and features are filtered out if the abundance is less than 1\%, prevalence is less than 5\% and count sum across all the samples is less than 100.
+  When the OP module is ``off'', the filtering is still performed but only to remove those features that are very low in presence.
+  The parameters used are given in Table~\ref{tab:op_tools}.
   The forking operation splits the count matrix into multiple matrices based on sample metadata column.
@@
   The output of this step are co-occurrence association networks encoded in the JSON format (refer Supplementary section) and are also exportable to a variety of network formats like .gml.
+  The list of tools used in this step, along with their modules and references are provided in Table~\ref{tab:ni_tools}.
 

### 2022-04-25 sections/tables_figures.tex
Add: Table about OP parameter values

@@
     \label{tab:ta_tools}
+  \end{table}
+
+  \begin{table}[h]
+  \begin{tabular}{|c|c|c|}
+  \hline
+  Condition                    & Parameter                 & Value          \\ \hline
+  \multirow{3}{*}{Filter(off)} & Prevalence threshold      & 2 / n\_samples \\ \cline{2-3}
+                               & Abundance threshold       & 0.001          \\ \cline{2-3}
+                               & Observation sum threshold & 10             \\ \hline
+  \multirow{3}{*}{Filter(on)}  & Prevalence threshold      & 0.05           \\ \cline{2-3}
+                               & Abundance threshold       & 0.01           \\ \cline{2-3}
+                               & Observation sum threshold & 100            \\ \hline
+  \end{tabular}
+  \caption{Parameters used in the OP module}
+  \label{tab:ni_tools}
   \end{table}

### 2022-04-25 sections/methods.tex
Add: References for scaled-sum and simple voting methods

@@
   \paragraph*{Algorithm 1: Simple voting}
-  % TODO: Add a reference
-  The simple voting method performs a voting-based consensus to determine whether an edge will exist between a given node-pair in the final consensus network.
+  The simple voting method performs a voting-based consensus to determine whether an edge will exist between a given node-pair in the final consensus network~\cite{bustinceFuzzySetsTheir2008,tsarevApplicationMajorityVoting2018}.
   For each pair of nodes, we determine the number of network inference methods that report an edge $j=k$ between them, i.e. $\bar{N}^{i}_{j=k}, \forall i \in \{1,\dots,w\}$.
@@
 
-  This algorithm generates a consensus network based on the sum of all edges reported between a pair of nodes.
+  This algorithm generates a consensus network based on the sum of all edges reported between a pair of nodes~\cite{bustinceFuzzySetsTheir2008,tsarevApplicationMajorityVoting2018}.
   Since in generating this consensus network we combine direct association metrics with correlation-based methods, summing of the edges is preceded by a pre-processing step, in which all networks are re-scaled.

### 2022-04-25 sections/methods.tex
Fix: Notation in Network Variability section

@@
   \vspace{-5mm}
-  This section defines notation for the network variability analysis in our study\dots
+  This section defines notation for the network variability analysis in our study.
 
-  % TODO: Update W and Q in the entire subsection
-  $W$ =  number of co-occurrence networks generated from all possible combinations of tools and parameters in the workflow.
+  $W$, the number of co-occurrence networks generated from all possible combinations of tools and parameters in the workflow.
 Note that this is different from $w$, which counted only the possible network inference modules.
 
-  $Q$ =  number of combined nodes from all $W$ networks
+  $Q$, the number of combined nodes from all $W$ networks
 
-  % TODO: Should we use different variable names here?
-  $N^i$ = The weights of the $i^{th}$ co-occurrence network represented as a $q^\prime \times q^\prime$ adjacency matrix, where $i \in {1, w^\prime}$
+  $N^i$ = The weights of the $i^{th}$ co-occurrence network represented as a $Q \times Q$ adjacency matrix, where $i \in {1, \dots, W}$
 
@@
 
-  $\bar{N}_i$ is a "flattened" version of the adjacency matrix $N^i$ into a $q^\prime^2 \times 1$ column vector, where all columns are stacked onto each other into a $q^\prime^2$ long vector:
+  $\bar{N}_i$ is a "flattened" version of the adjacency matrix $N^i$ into a $Q^2 \times 1$ column vector, where all columns are stacked onto each other into a $Q^2$ long vector:
 
   \subsubsection*{Principal Component Analysis and variability calculation}
+
   \vspace{-5mm}
@@
    First, we obtain co-occurrence network $\bar{N}_i$ for each of the $W$ possible tool and parameter combinations in the workflow.
- Then, we construct a matrix $\mathbf{\bar{N}}$ whose $i^{th}$ column is the flattened version of the $i^{th}$ network, i.e.
-the column vector $\bar{N}^i$.
+   Then, we construct a matrix $\mathbf{\bar{N}}$ whose $i^{th}$ column is the flattened version of the $i^{th}$ network, i.e. the column vector $\bar{N}^i$.
    Therefore $\bar{N}^i_j$ is the weight of edge $j$ in network $i$.
@@
       \begin{block}{*{4}{>{$\footnotesize}c<{$}} l}
-        $\bar{N}$^1 & $\bar{N}$^2 & \dots & $\bar{N}$^{w^\prime} & \\
+        $\bar{N}$^1 & $\bar{N}$^2 & \dots & $\bar{N}$^{W} & \\
       \end{block}
       \begin{block}{[*{4}{c}]>{$\footnotesize}l<{$}}
-       n^1_1 & n^2_1 & \cdots  & n^{w^\prime}_1 \bigstrut[t]& $\bar{N}$_1 \\
-       n^1_2 & n^2_2 & \cdots  & n^{w^\prime}_2 & $\bar{N}$_2 \\
+       n^1_1 & n^2_1 & \cdots  & n^{W}_1 \bigstrut[t]& $\bar{N}$_1 \\
+       n^1_2 & n^2_2 & \cdots  & n^{W}_2 & $\bar{N}$_2 \\
        \vdots & \vdots & \vdots  & \vdots & \vdots \\
-       n^1_{q^\prime^2} & n^2_{q^\prime^2} & \cdots  & n^{w^\prime}_{q^\prime^2} & $\bar{N}$_{q^\prime^2}
+       n^1_{Q^2} & n^2_{Q^2} & \cdots  & n^{W}_{Q^2} & $\bar{N}$_{Q^2}
       \end{block}
@@
   One issue with this approach is that the possible correlations between the edges of the network could lead to inaccurate estimates of the variance if a linear model were used to directly model the relationships between edges and steps in the workflow.
-  In order to remedy this issue, we perform Principal Component Analysis on the matrix $\mathbf{\bar{N}}$ to obtain the $\mathbf{C}$ matrix ($w^\prime \times c$) of components for each network, such that we reduce the dimensions from the $q^\prime^2$ dimensional edge space to a $c$ dimensional component space.
+  In order to remedy this issue, we perform Principal Component Analysis on the matrix $\mathbf{\bar{N}}$ to obtain the $\mathbf{C}$ matrix ($W \times c$) of components for each network, such that we reduce the dimensions from the $Q^2$ dimensional edge space to a $c$ dimensional component space.
 
@@
 
-  % TODO: Should we remove CC from the LM model?
   In particular, we infer parameters $\alpha_i$ such that:
-   \begin{equation*}
-       C_j = \sum_{k=1}^5 \left( \alpha^{DC(k)}_j.\delta^{DC(k)}_j \right) +
+     \begin{multline*}
+      C_j = &\sum_{k=1}^5 \left( \alpha^{DC(k)}_j.\delta^{DC(k)}_j \right) +
              \sum_{k=1}^2 \left( \alpha^{CC(k)}_j.\delta^{CC(k)}_j \right) +
@@
              \sum_{k=1}^2 \left( \alpha^{OP(k)}_j.\delta^{OP(k)}_j \right) + \\
-             \sum_{k=1}^{10} \left( \alpha^{NI(k)}_j.\delta^{NI(k)}_j \right) +
+             &\sum_{k=1}^{10} \left( \alpha^{NI(k)}_j.\delta^{NI(k)}_j \right) +
              \epsilon_j
-   \end{equation*}
+     \end{multline*}
 
@@
 
-  The variance contributed by each step of the pipeline is calculated for every component in $\mathbf{C}$ matrix through ANOVA using the Python \texttt{statsmodels} package and is shown in Figure~\ref{fig:figure2}A.
+   The variance contributed by each step of the pipeline is calculated for every component in $\mathbf{C}$ matrix through ANOVA using the Python \texttt{statsmodels}~\cite{seaboldStatsmodelsEconometricStatistical2010} package and is shown in Figure~\ref{fig:figure2}A.
   The total variance for the network is calculated by adding the variances for each connection.

### 2022-04-25 sections/discussion.tex
Change: Update discussion section

@@
   The \ac{micone} Python package provides functions and methods to perform detailed analysis of the count matrices and the inferred networks.
-  % TODO: Ref and supplementary
-  Additionally, the inferred networks by default are created using a custom JSON format (see Supplementary), but are interchangeable to Cytoscape[ref], GML[ref], and many other popular formats via the \ac{micone} Python package.
+  Additionally, the inferred networks by default are exported to a custom JSON format (see Supplementary), but are interchangeable to Cytoscape~\cite{shannonCytoscapeSoftwareEnvironment2003}, GML~\cite{himsoltGMLPortableGraph2010}, and many other popular formats via the \ac{micone} Python package.
 
-  % TODO: Refs
-  There have been several tools/workflows such as \ac{qiime2}~\cite{bolyenReproducibleInteractiveScalable2019} and NetCoMi[ref] that have been developed for analysis of 16S sequencing data and to infer networks from these datasets.
+  There have been several tools/workflows such as \ac{qiime2}~\cite{bolyenReproducibleInteractiveScalable2019} and NetCoMi~\cite{peschelNetCoMiNetworkConstruction2020} that have been developed for analysis of 16S sequencing data and to infer networks from these datasets.
   However, there has been no single tool that completely integrates the process of inferring microbial interaction networks from 16S sequencing reads.
@@
   The default reference database in the pipeline is the \ac{gg} reference database along with the NaiveBayes classifier as the query tool.
-  The reason for our choice stem from the popularity of the \ac{gg} database in taxonomic studies, which would enable easy comparision across datasets.
+  The reason for our choice stem from the popularity of the \ac{gg} database in taxonomic studies, which would enable easy comparison across datasets.
   We also enable users to use custom databases \cite{Ritari2015} with the blast+ and naive bayes classifiers that are incorporated into the pipeline (from \ac{qiime2}).
 
-  The \ac{op} step of the pipeline the step that second in its contribution to total network variance.
+  The \ac{op} step of the pipeline is second in its contribution to total network variance.
   This can be attributed to the large number of nodes that are added to the final networks when filtering is turned off.
-  % TODO: Refs
-  Additionally, the large number of nodes also greatly decrease the accuracy of the network inference algorithms [ref] as well as increase the computational complexity.
-  % TODO: Supplementary work - Figure S5
-  We observe that filtering out taxa that are present in low abundances in all samples did not necessarily increase the proportion of taxa in common between taxonomy tables generated using different reference databases.
-  % TODO: Supplementary work - Figure S2
-  However, we do observe that the reduction in the number of taxa leads to better agreement in the networks inferred through different methods.
+  Additionally, a very large number of nodes also decreases the accuracy of the network inference algorithms for the same sample size~\cite{peschelNetCoMiNetworkConstruction2020} and increases the computational complexity~\cite{tackmannRapidInferenceDirect2019}.
+  We observe that filtering out taxa that are present in low abundances in all samples decrease the proportion of taxa in common between taxonomy tables generated using different reference databases, but not significantly (Figure~\cite{fig:figure_s5}).
+  However, we do observe that the reduction in the number of taxa leads to better agreement in the networks inferred through different methods (Figure~\ref{fig:figure_s1}).
   Moreover, filtering is necessary in order to increase the power in tests of significance when the number of taxa is much greater than the number of samples.

### 2022-04-25 sections/results.tex
Fix: Update the captions on the figures

@@
 
-  Figure \ref{fig:figure2}B is the PCA plot of all the generated co-occurrence networks, colored by the databases in the TA step (left) and the network inference algorihtms in the NI step (right).
+  Figure \ref{fig:figure2}B is the PCA plot of all the generated co-occurrence networks, colored by the databases in the TA step (left) and the network inference algorithms in the NI step (right).
   In other words, each point corresponds to a network generated using different combination of tools, and captures how much the final network is affected by such choice.
@@
   The \ac{dada2} method is from the \ac{dada2} R package.
+  The closed and open reference methods in this analysis use the \ac{gg} database for reference sequence alignment.
 
@@
   Additionally, we observe that other abundant Genus like Bacteroides and Escherichia are also assigned to different representative sequences (indicated by the different proporitions in the alluvial plot).
-  Furthermore, the most abundant Genus (among the top 50 representative sequences) in the \ac{gg} taxonomy table was \textit{Escherichia} whereas it was assigned to "unknown" in NCBI and in the SILVA taxonomy table it was assigned to \textit{Escherichia-Shigella}.
-  Although these are minor differences (in the case of \ac{gg} and SILVA), when comparing a large number of taxonomy composition tables these problems are hard to diagnose.
+  Furthermore, the most abundant Genus (among the top 50 representative sequences) in the \ac{ncbi} taxonomy table was \textit{Escherichia} whereas it was assigned to "unknown" in \ac{gg} and in the SILVA taxonomy table it was assigned to \textit{Escherichia-Shigella}.
+  Although these are minor differences (in the case of \ac{ncbi} and SILVA), when comparing a large number of taxonomy composition tables these problems are hard to diagnose.
 

### 2022-04-25 sections/tables_figures.tex
Add: FMT figures to manuscript
Change: Update discussion section
Fix: Update the captions on the figures

@@
       \hline
-      \textbf{Process} & \textbf{Tool} & \textbf{Parameters} \\
+      \textbf{Process} & \textbf{Tool} \\
       \hline
-      Denoising and Clustering & Dada2/Deblur & default \\
-      Taxonomy Assignment & \ac{ncbi} with Blast & RefSeq database \\
-      OTU Processing & Based on statistical power & Dynamic cutoff \\
-      Network Inference & Consensus method & - \\
+      Denoising and Clustering & Dada2 and remove bimera \\
+      Taxonomy Assignment & \ac{gg} with NaiveBayes classifier \\
+      OTU Processing & Thresholds on prevalence and abundance \\
+      Network Inference & scaled-sum consensus method \\
       \hline
@@
   \begin{table}[h]
+    \centering
+    \small
     \begin{tabular}{|c|c|c|}
@@
 % Figures
+  \FloatBarrier
+  \newpage
 
-  \begin{figure}[H]
+  \begin{figure}
     \centering
@@
   \end{figure}
-  \begin{figure}[H]
+  \begin{figure}[h]
     \centering
     \caption{
-      \textbf{The workflow of the microbial co-occurrence analysis pipeline}.
-      The steps can be grouped into four major groups: \textbf{(DC)} \textbf{D}enoising and \textbf{C}lustering, \textbf{(TA)} \textbf{T}axonomy \textbf{A}ssignment, \textbf{(OP)} \textbf{O}TU or ESV \textbf{P}rocessing, and \textbf{(NI)} \textbf{N}etwork \textbf{I}nference.
+      \textbf{The workflow of the \ac{micone} pipeline}.
+      The steps can be grouped into five major groups: \textbf{(SP)} \textbf{S}equence \textbf{P}rocessig, \textbf{(DC)} \textbf{D}enoising and \textbf{C}lustering, \textbf{(TA)} \textbf{T}axonomy \textbf{A}ssignment, \textbf{(OP)} \textbf{O}TU or ESV \textbf{P}rocessing, and \textbf{(NI)} \textbf{N}etwork \textbf{I}nference.
       Each step incorporates several processes, each of which in turn have several alternate algorithms for the same task (indicated by the text to the right of the blue boxes).
-      The text along the arrows describes the data that is being passed from one step to another. For details on each process and data types, see Methods.
+      The text along the arrows describes the data that is being passed from one step to another.
+      The final output of the pipeline are consensus networks generated from the co-occurrence networks.
+      For details on each process and data types, see Methods.
     }
@@
 
-  \begin{figure}[H]
+  \begin{figure}
     \centering
@@
   \end{figure}
-  \begin{figure}[H]
+  \begin{figure}
     \centering
       \caption{
-      \textbf{The choice of database contributes to the most variance in the networks}.
-      \textbf{(A)} The total relative variance in the networks contributed by the DC, TA and OP steps of the pipeline (right) and the linear model used to calculate the relative variance (left), see the Methods section for details.
+      \textbf{The choice of reference database contributes to the most variance in the networks}.
+      \textbf{(A)} The total relative variance in the networks contributed by the DC, CC (chimera checking), TA and OP steps of the pipeline (right) and the linear model used to calculate the relative variance (left), see the Methods section for details.
+      The analyses was performed on the healthy stool samples in a fecal microbiome transplant dataset~\ref{Kang2017}.
+      The taxonomy database contributes most to the variance between the networks (65.4\%) followed by filtering of the counts matrix (26.8\%) in the OP step.
       \textbf{(B)} All combinations of inferred networks are shown as points on a PCA plot.
-      The color of the points corresponds to the taxonomy database, the shape corresponds to the denoising/clustering method and the size corresponds to whether low abundance OTUs were removed or not.
-      \textbf{(B inset)} The network generated using DC=dada2, TA=GG, OP=no and NI=SPARCC and represents the particular point shown (big red square).
-      The points on the PCA plot can be separated based on the TA step and the differences due to the DC and OP steps are comparatively not as significant.
+      Each point on the PCA plot represents a network inferred using different combinations of tools and parameters that are available in the \ac{micone} pipeline.
+      The color of the points corresponds to the reference taxonomy database used to reconstruct those networks (left), and the network inference algorithm used to infer the networks (right).
+      % TODO: The points aren't really separated
+      The points on the PCA plot can be separated based on the TA step and the differences due to the DC and CC steps are comparatively not as significant.
     }
@@
 
-  % TODO: Describe the datasets being shown here clearly. In the caption + figure legends (if necessary) and in the text
   \begin{figure}
@@
     \caption{
-      \textbf{The representative sequences generated by the different denoising/clustering methods are very similar but differ in the sequences that are in low abundance.}
-      \textbf{(A)} The average weighted UniFrac distance between the representative sequences shows that the representative sequences and their compositions are fairly identical between the methods,
-      \textbf{(B)} The relatively larger average unweighted UniFrac distance indicates that methods differ in their identification of sequences of low abundance,
-      \textbf{(C, D)} The distributions of the average weighted UniFrac distance between the expected sequence profile and the calculated sequence profile in mock datasets.
-      \textbf{(D)} The distributions of the average unweighted UniFrac distance show that dada2 and Deblur were the best performing methods in most of the datasets.
+      \textbf{The representative sequences generated by the different denoising/clustering methods are very similar but differ in the sequences that are lower in abundance.}
+      \textbf{(A)} The average weighted UniFrac distance between the representative sequences shows that the representative sequences and their compositions are fairly identical between the methods (with the exception of Deblur (DB)).
+      \textbf{(B)} The relatively larger average unweighted UniFrac distance indicates that methods differ in their identification of sequences that are lower in abundance.
+      \textbf{(C, D)} The distributions of the average weighted and unweighted UniFrac distance between the expected sequence profile and the calculated sequence profile in the mock datasets.
+      The distributions of the average weighted UniFrac distance show that de novo (DN) and open reference (OR) were the best performing methods in most of the datasets, while they are worst performing methods under the unweighted UniFrac metric.
+      The above average performance of dada2 (D2) under both distance metrics combined with its approach of identifying \ac{esv}s using de novo methods, prompts us to propose it as the default method for the DC step.
     }
@@
     \caption{
-      \textbf{Taxonomic reference databases vary widely in terms of their taxonomy assignments.}
-      \textbf{(A)} The assignment of the top 50 representative sequences to their respective taxonomies using the three different reference databases shows how the same sequences are assigned to different Genus.
-      \textbf{(B)} The percentage of \ac{otu}s assigned to the same taxonomic label when using different reference databases.
-      The percentage of mismatches decrease at higher taxonomic levels but even at the Phylum level there exists around 10\% of mismatches.
-      \textbf{(C)} The Bray-Curtis dissimilarity between the expected taxonomy profile and calculated taxonomy profile in the mock datasets shows that there is no singular best choice of database for every dataset.
+      \textbf{Taxonomic reference databases vary widely in terms of their taxonomy assignments beyond the Order level.}
+      \textbf{(A)} The assignments of the top 50 representative sequences to their respective taxonomies using the three different reference databases show how the same sequences are assigned to different Genus under the different databases.
+      A significant portion of the representative sequences are assigned to an ``unknown'' Genus in two of three databases (\ac{gg} and \ac{ncbi}).
+      \textbf{(B)} The number of representative sequences assigned to the same taxonomic label when using different reference databases, shown for the top 100 sequences.
+      The number of mismatches are fewer at higher taxonomic levels but even at the Order level there exists greater than 51\% of mismatches, demonstrating the poor agreement in taxonomic labels assigned by the different databases.
+      \textbf{(C)} The Bray-Curtis dissimilarity between the expected taxonomy profile and calculated taxonomy profile in the mock datasets shows that there is no singular best choice of database for every dataset, as all the databases have similar performance.
+      However, we recommend the \ac{gg} database due to its popularity and if the goal is to compare against existing networks in literature.
     }
@@
       \textbf{Networks generated using different network inference methods show notable differences both in terms of edge-density and connectivity}.
-      \textbf{(A)} The six different networks generated by the different network inference methods are very dissimilar.
+      \textbf{(A)} The nine different networks generated by the different network inference methods (excluding \ac{mldm}) are very dissimilar.
       The green links are positive associations and the orange links are negative associations.
-      A threshold of 0.3 was set for the methods that infer pairwise correlations (\ac{sparcc}, Spearman, Pearson) and no threshold was set for the other methods.
-      \textbf{(B)} The node overlap Upset plot~\cite{Lex} indicates that all the networks have a large number of common nodes involved in connections.
-      Whereas, \textbf{(C)} The edge overlap Upset plot shows that a very small fraction of these connections are actually shared.
+      A threshold of 0.3 was set for the correlation-based methods (\ac{sparcc}, Propr, Spearman and Pearson) and a threshold of 0.01 was set for the other direct association methods.
+      \textbf{(B)} The node overlap Upset plot indicates that all the networks have a large number of common nodes involved in connections (33 out of 68).
+      Whereas, \textbf{(C)} the edge overlap Upset plot shows that a very small fraction of these connections are actually shared (8 out of 202).
     }
@@
     \caption{
-      The precision and sensitivity of the inferred networks on the synthetic data.
+      \textbf{The precision and sensitivity of the inferred networks on the NorTA synthetic interaction data.}
+      The ``I'' prefix indicates the individual network inference methods and the ``X'' prefix denotes the different consensus based methods.
+      The different consensus based methods used are: pvalue merging, scaled-sum (SS) and simple voting (SV) method.
+      Pearson and Spearman methods are not used in the calculation of the consensus.
+      Among all the independent network inference methods, \ac{spieceasi} has the best average precision (0.944), but the overall best precision was consistently obtained by the scaled-sum method (0.956, 0.985 and 1.000).
+      The simple voting method when using presence of edges in all inferred networks as a requirement (parameter value 1.000), also outperforms \ac{spieceasi} on average precision (0.969).
+      Therefore, we recommend the scaled-sum consensus method as the default tool for network inference.
     }
@@
 
+  % TODO: Verify the validity of Figure 7A and 7B again
+  % Are they both using consensus?
   \begin{figure}[h]
@@
     \caption{
-      \textbf{Network inference and taxonomic assignment have the highest influence on the inferred network structures.}
-      \textbf{(A)} The network constructed using the default pipeline parameters (DC=\ac{dada2}, TA=\ac{ncbi}, OP=on, NI=\ac{sparcc}) is compared with networks generated when one of the steps use a different tool.
+      \textbf{The consensus algorithm drastically reduces variability in the networks and increases robustness of inferred networks}
+      \textbf{(A)} The network constructed using the default pipeline parameters (DC=\ac{dada2}, CC=remove bimera, TA=\ac{gg}, OP=Filter(on), NI=scaled-sum consensus) is compared with networks generated when one of the steps use a different tool.
       The common connections (common with the default network) are in black, connections unique to the network are colored purple and connections in the default network but not present in the current network are gray.
-      \textbf{(B)} The L1 distance between the networks generated by changing one step of the default pipeline and the network generated using the default parameters.
+      Changing the NI, TA and OP steps leads to the creation of the most number of new nodes and edges.
+      \textbf{(B)} The boxplot showing the L1 distance between the networks generated by changing one step of the default pipeline and the network generated using the default parameters.
+      The high value of L1 distances for the network inference step as compared to the other steps demonstrates how the consensus method is able to reduce the variability in the networks in the other steps of the pipeline.
+      \textbf{(C)} The networks generated for the control (left) and autistic (right) samples in the fecal microbiome transplant dataset.
+      These networks were generated using the default tools and parameters recommended by the \ac{micone} pipeline.
     }

### 2022-04-25 scripts/figure7/figure7_plot.R

@@ -198,7 +198,7 @@ c_plot <- ggarrange(default_plot_1, default_plot_2, ncol=2, nrow=1, common.legen
 
 ggsave(output_file_c, width = 11, height = 5)
 
-annotate_figure(c_plot, fig.lab="C", fig.lab.pos="top.left", fig.lab.size=20)
+c_plot <- annotate_figure(c_plot, fig.lab="C", fig.lab.pos="top.left", fig.lab.size=20)
 
 
 final_plot <- ggarrange(a_plot, b_plot, c_plot, ncol = 1)

### 2022-04-25 sections/methods.tex
Add: Add captions and update supplementary figures

@@
   The options currently available in the pipeline for denoising and clustering are: open reference clustering, closed reference clustering and de novo clustering methods from the vsearch plugin of \ac{qiime2} v2021.8.0~\cite{bolyenReproducibleInteractiveScalable2019} and denoising methods from \ac{dada2} v1.14~\cite{Callahan2016} (from the \ac{dada2} R package) and Deblur v1.1.0~\cite{Amir2017} (from the deblur plugin of \ac{qiime2}).
-  The quality filtering and chimera checking tools are derived from those used in \ac{qiime2} v2021.8.0 (uchime method) and \ac{dada2} (remove bimera method).
+  The quality filtering and chimera checking tools are derived from those used in \ac{qiime2} v2021.8.0 (uchime-denovo method) and \ac{dada2} (remove bimera method).
   The list of tools used in this step, along with their modules and references are provided in Table~\ref{tab:dc_tools}.
@@
   \end{equation*}
-  where, $#$ refers to the cardinality of the set.
+  where, # refers to the cardinality of the set.
 
@@
       \begin{block}{*{4}{>{$\footnotesize}c<{$}} l}
-        $\bar{N}$^1 & $\bar{N}$^2 & \dots & $\bar{N}$^{W} & \\
+        \bar{N}^1 & \bar{N}^2 & \dots & \bar{N}^{W} & \\
       \end{block}
       \begin{block}{[*{4}{c}]>{$\footnotesize}l<{$}}
-       n^1_1 & n^2_1 & \cdots  & n^{W}_1 \bigstrut[t]& $\bar{N}$_1 \\
-       n^1_2 & n^2_2 & \cdots  & n^{W}_2 & $\bar{N}$_2 \\
+       n^1_1 & n^2_1 & \cdots  & n^{W}_1 \bigstrut[t]& \bar{N}_1 \\
+       n^1_2 & n^2_2 & \cdots  & n^{W}_2 & \bar{N}_2 \\
        \vdots & \vdots & \vdots  & \vdots & \vdots \\
-       n^1_{Q^2} & n^2_{Q^2} & \cdots  & n^{W}_{Q^2} & $\bar{N}$_{Q^2}
+       n^1_{Q^2} & n^2_{Q^2} & \cdots  & n^{W}_{Q^2} & \bar{N}_{Q^2}
       \end{block}

### 2022-04-25 sections/results.tex
Add: Add captions and update supplementary figures

@@
   The recommended tool for the \ac{dc} step (\ac{dada2}) was chosen based on its accuracy in recapitulating the reference sequences in mock communities.
-  For the chimera checking module of the \ac{dc} step, both uchime and remove bimera perform similarly, we suggest using remove bimera since it was developed specifically for \ac{dada2} and should provide the best performance for that denoising method.
+  For the chimera checking module of the \ac{dc} step, both uchime-denovo and remove bimera perform similarly, we suggest using remove bimera since it was developed specifically for \ac{dada2} and should provide the best performance for that denoising method.
   The choice of the taxonomy reference database in the \ac{ta} step is dictated largely by the species expected to be present in the sample as well the database used in similar studies, if comparison is a goal.

### 2022-04-25 sections/supplementary.tex
Add: Add captions and update supplementary figures

@@
 
-    % TODO: Elaborate on this
-    Figure~\ref{fig:figure_s1} describes the variation in the inferrred networks in the first two principal axes.
-    This Figure is the similar to Figure~\ref{fig:figure2}B and further confirms that the variability in the networks decreases upon filtering out the taxonomic entities at low abundance.
+    Figure~\ref{fig:figure_s1} describes the variation in the inferred networks in the first two principal axes and is similar to Figure~\ref{fig:figure2}B.
+    The \ac{dc} step does not seem to have any correlation with the similarity of networks on the PCA plot, but all the networks with filtering at the \ac{op} step are similar to each other.
+    This further confirms that the variability in the networks decreases upon filtering out the taxonomic entities at low abundance.
 
@@
           \textbf{The variation in the inferred networks in the first two principal axes}
-          The plots are colored by the tools or parameters used in the \ac{dc} step (A) and the \ac{op} step (B)
+          All combinations of inferred networks are shown as points on a PCA plot.
+          Each point on the PCA plot represents a network inferred using different combinations of tools and parameters that are available in the \ac{micone} pipeline.
+          The points are colored by the tools or parameters used in the \ac{dc} step (A) and the \ac{op} step (B).
         }
@@
 
-    % TODO: Write something about TSNE
+    % TODO: Verify this about t-SNE
     \begin{figure}[H]
@@
         \caption{
-          \textbf{The TSNE plot of the networks inferred by various methods}.
+          \textbf{The t-SNE plot of the networks inferred by various methods}.
+          The points are colored by the tools and parameters used in \ac{dc} step (A), \ac{ta} step (B), \ac{op} step (C) and \ac{ni} step (D).
+          The separation of the points based on taxonomy reference database shows that the points might cluster based on reference database in high-dimensional space.
         }
@@
 
-  % QUESTION: Should the CCA section be only included in the response letter?
-
   \subsection*{Similarity of high abundance reference sequences}
 
-    Figure~\ref{fig:figureS3} is similar to Figure~\ref{fig:figure3} shows the average UniFrac distance between reference sequences generated by the various methods in the \ac{dc} step.
-    In Figure~\ref{fig:figureS3} however, we use only the top 1000 representative sequences (by abundance) for this calculation.
+    Figure~\ref{fig:figure_s3} is similar to Figure~\ref{fig:figure3} shows the average UniFrac distance between reference sequences generated by the various methods in the \ac{dc} step.
+    In Figure~\ref{fig:figure_s3} however, we use only the top 1000 representative sequences (by abundance) for this calculation.
     We observe that both the weighted and unweighted UniFrac distances are increased compared to the figure with the all the representative sequences.
-    This shows that the most abundant representative sequences generated by the methods are not similar to each other.
+    This shows that the 1000 most abundant representative sequences generated by the methods are not similar to each other.
+    And since the weighted UniFrac is much smaller than the unweighted UniFrac distance, we can conclude that those reference sequences that are present in the middle of the abundance distribution are dissimilar.
 
@@
         \caption{
-          \textbf{The UniFrac distance between the 100 most abundant representative sequences generated by various method in the \ac{dc} step}
+          \textbf{The UniFrac distance between the 1000 most abundant representative sequences generated by various method in the \ac{dc} step}
         }
@@
     This is especially true for the \ac{dada2} and Deblur methods which are the recommended denoising methods in the \ac{micone} pipeline.
-    % TODO: Can we verify this?
-    Therefore, we recommended remove bimera as the default chimera method if one is using \ac{dada2} and uchime when one is using Deblur, since these methods were developed for these respective algorithms.
+    Therefore, we recommended remove bimera as the default chimera method if one is using \ac{dada2} and uchime-denovo when one is using Deblur, since these methods were developed for these respective algorithms (\ac{qiime2} uses uchime-denovo in the Deblur workflow).
 
@@
         \caption{
+          \textbf{The weighted and unweighted UniFrac distance between representative sequences generated using remove bimera and uchime for each denoising method.}
         }
@@
 
-    % TODO: Elaborate on this and work on the caption
     Figure~\ref{fig:figure_s5} is similar to Figure~\ref{fig:figure4}B, but instead of the matching only the top 100 taxonomic entities (by abundance), we match all the assignments from one database with those from the other two databases.
@@
         \caption{
+          \textbf{The pairwise comparison of assignments generated using different databases for each representative sequence}
         }
@@
   \begin{table}[]
-  \begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|l|}
+    \centering
+    \small
+    \begin{tabular}{|l|l|l|l|l|p{2cm}|p{2cm}|p{2cm}|p{2cm}|p{2cm}|p{2cm}|p{2cm}|}
   \hline
@@
 
+    % TODO: Rerun this figure
     \begin{figure}[H]
@@
         \caption{
+          \textbf{The precision and sensitivity of the inferred networks on the NorTA synthetic interaction data with Pearson and Spearman added to the consensus.}
+            The ``I'' prefix indicates the individual network inference methods and the ``X'' prefix denotes the different consensus based methods.
+            The different consensus based methods used are: pvalue merging, scaled-sum (SS) and simple voting (SV) method.
+            With the addition of Pearson and Spearman in the consensus calculation we observe a decrease in the average precision scores of both the consensus algorithms.
+            The overall best precision was still consistently obtained by the scaled-sum method (0.985, 0.963 and 1.000), but the performance was not stable to changes in parameter value.
+            The simple voting method when using presence of edges in all inferred networks as a requirement (parameter value 1.000), again outperforms \ac{spieceasi} on average precision (0.961).
         }
@@
 
-    % TODO: Elaborate on this
     In Figure~\ref{fig:figure_s7} we perform the same analysis we did with the NorTA data using the seqtime data.
@@
         \caption{
+          \textbf{The precision and sensitivity of the inferred networks on the seqtime synthetic interaction data.}
+          The ``I'' prefix indicates the individual network inference methods and the ``X'' prefix denotes the different consensus based methods.
+          The different consensus based methods used are: pvalue merging, scaled-sum (SS) and simple voting (SV) method.
+          Pearson and Spearman methods are not used in the calculation of the consensus.
+          Among all the independent network inference methods, \ac{spieceasi} has the best average precision (0.624), but the overall best precision was consistently obtained by the scaled-sum method (0.688, 0.820 and 1.000).
+          The simple voting method when using presence of edges in all inferred networks as a requirement (parameter value 1.000), also outperforms \ac{spieceasi} on average precision (0.692).
         }

### 2022-04-25 sections/tables_figures.tex
Fix: Update arrangement of figures and tables

@@
       \hline
-      Chimera checking & Uchime & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
+      Chimera checking & Uchime-denovo & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
       Chimera checking & Remove bimera & \cite{Callahan2016} \\
@@
 
-  \begin{figure}
+  \begin{figure}[h]
     \centering
@@
   \end{figure}
-  \begin{figure}[h]
+  \begin{figure}[ht!]
     \centering
@@
     \includegraphics[width=\textwidth]{figure3.pdf}
+  \end{figure}
+  \begin{figure}
+    \centering
     \caption{
@@
     \centering
-    \includegraphics[width=0.9\textwidth,height=1.2\textwidth]{figure4.pdf}
+    \includegraphics[width=\textwidth,height=1.2\textwidth]{figure4.pdf}
   \end{figure}
@@
     \centering
-    \includegraphics[width=0.9\textwidth]{figure5.pdf}
+    \includegraphics[width=\textwidth]{figure5.pdf}
   \end{figure}
@@
   % Are they both using consensus?
-  \begin{figure}[h]
+  \begin{figure}[H]
     \centering
-    \includegraphics[width=0.85\linewidth]{figure7.pdf}
+    \includegraphics[width=1.0\linewidth]{figure7.pdf}
+  \end{figure}
+  \begin{figure}[H]
+    \centering
     \caption{

### 2022-04-25 sections/supplementary.tex
Fix: Update arrangement of figures and tables for suppl

@@
 
-  \subsection*{Mismatches in all assignments at the \ac{TA} step}
+  \subsection*{Mismatches in all assignments at the \ac{ta} step}
 
@@
 We observe that only the networks generated using propr have a high value since most of these networks are connected.
-    \item Edge connectivity refers to the minimum number of edges that must be removed from the graph to make it disconnected. These values are similar to node connectivity.
     \item Degree assortativity coefficient measures the similarity of connections in the graph with respect to the node degree.
@@
 
-  \begin{table}[]
+  \begin{landscape}
+  \begin{table}[H]
     \centering
     \small
-    \begin{tabular}{|l|l|l|l|l|p{2cm}|p{2cm}|p{2cm}|p{2cm}|p{2cm}|p{2cm}|p{2cm}|}
+    \begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{2cm}|p{1.5cm}|p{1.5cm}|p{1.5cm}|p{1.5cm}|p{1.5cm}|p{1.5cm}|p{1.5cm}|}
   \hline
   DC & CC            & TA                & OP          & NI         & Average shortest path length & Average clustering & No.
-of connected components & Modularity & Node connectivity & Edge connectivity & Degree assortativity coefficient \\ \hline
-  DB & -             & -                 & -           & -          & 2.066                        & 0.265              & 27.088                      & 0.387      & 1.868             & 1.868             & 0.070                            \\
-  OR & -             & -                 & -           & -          & 1.904                        & 0.273              & 24.868                      & -17.060    & 1.877             & 1.877             & 0.099                            \\
-  D2 & -             & -                 & -           & -          & 2.263                        & 0.267              & 28.228                      & 0.326      & 1.789             & 1.789             & 0.078                            \\
-  DN & -             & -                 & -           & -          & 2.065                        & 0.272              & 25.132                      & 0.302      & 1.754             & 1.754             & 0.093                            \\
-  CR & -             & -                 & -           & -          & 1.937                        & 0.276              & 24.289                      & -1.865     & 2.079             & 2.079             & 0.097                            \\
-  -  & remove bimera & -                 & -           & -          & 2.070                        & 0.272              & 25.968                      & -7.439     & 1.863             & 1.863             & 0.086                            \\
-  -  & uchime        & -                 & -           & -          & 2.066                        & 0.269              & 25.874                      & 0.275      & 1.884             & 1.884             & 0.089                            \\
-  -  & -             & BLAST(NCBI)       & -           & -          & 2.218                        & 0.262              & 22.784                      & -3.683     & 1.495             & 1.495             & 0.100                            \\
-  -  & -             & NaiveBayes(SILVA) & -           & -          & 1.872                        & 0.275              & 30.716                      & 0.415      & 2.674             & 2.674             & 0.074                            \\
-  -  & -             & NaiveBayes(GG)    & -           & -          & 2.086                        & 0.274              & 24.263                      & -7.478     & 1.453             & 1.453             & 0.088                            \\
-  -  & -             & -                 & Filter(on)  & -          & 1.975                        & 0.252              & 18.493                      & -7.194     & 1.880             & 1.880             & 0.073                            \\
-  -  & -             & -                 & Filter(off) & -          & 2.193                        & 0.291              & 34.174                      & 0.431      & 1.867             & 1.867             & 0.104                            \\
-  -  & -             & -                 & -           & propr      & 1.650                        & 0.528              & 1.000                       & 0.000      & 10.583            & 10.583            & 0.036                            \\
-  -  & -             & -                 & -           & harmonies  &                              & 0.108              & 82.633                      & 0.757      & 0.000             & 0.000             & 0.340                            \\
-  -  & -             & -                 & -           & flashweave & 3.878                        & 0.079              & 1.800                       & 0.458      & 0.383             & 0.383             & -0.032                           \\
-  -  & -             & -                 & -           & cozine     &                              & 0.090              & 82.167                      & 0.694      & 0.000             & 0.000             & -0.072                           \\
-  -  & -             & -                 & -           & spring     &                              & 0.092              & 11.767                      & 0.566      & 0.000             & 0.000             & -0.035                           \\
-  -  & -             & -                 & -           & sparcc     & 1.747                        & 0.525              & 1.033                       & 0.009      & 5.500             & 5.500             & -0.011                           \\
-  -  & -             & -                 & -           & spieceasi  &                              & 0.076              & 39.383                      & 0.717      & 0.000             & 0.000             & -0.031                           \\
-  -  & -             & -                 & -           & pearson    &                              & 0.457              & 23.883                      & 0.551      & 0.000             & 0.000             & 0.478                            \\
-  -  & -             & -                 & -           & spearman   & 1.875                        & 0.551              & 1.450                       & 0.044      & 1.117             & 1.117             & 0.170                            \\
-  -  & -             & -                 & -           & mldm       & 2.760                        & 0.130              & 2.267                       & -75.648    & 0.433             & 0.433             & -0.028                           \\ \hline
+of connected components & Modularity & Node connectivity & Degree assortativity coefficient \\ \hline
+  DB & -             & -                 & -           & -          & 2.066                        & 0.265              & 27.088                      & 0.387      & 1.868             & 0.070                            \\
+  OR & -             & -                 & -           & -          & 1.904                        & 0.273              & 24.868                      & -17.060    & 1.877             & 0.099                            \\
+  D2 & -             & -                 & -           & -          & 2.263                        & 0.267              & 28.228                      & 0.326      & 1.789             & 0.078                            \\
+  DN & -             & -                 & -           & -          & 2.065                        & 0.272              & 25.132                      & 0.302      & 1.754             & 0.093                            \\
+  CR & -             & -                 & -           & -          & 1.937                        & 0.276              & 24.289                      & -1.865     & 2.079             & 0.097                            \\
+  -  & remove bimera & -                 & -           & -          & 2.070                        & 0.272              & 25.968                      & -7.439     & 1.863             & 0.086                            \\
+  -  & uchime        & -                 & -           & -          & 2.066                        & 0.269              & 25.874                      & 0.275      & 1.884             & 0.089                            \\
+  -  & -             & NCBI              & -           & -          & 2.218                        & 0.262              & 22.784                      & -3.683     & 1.495             & 0.100                            \\
+  -  & -             & SILVA             & -           & -          & 1.872                        & 0.275              & 30.716                      & 0.415      & 2.674             & 0.074                            \\
+  -  & -             & GG                & -           & -          & 2.086                        & 0.274              & 24.263                      & -7.478     & 1.453             & 0.088                            \\
+  -  & -             & -                 & On          & -          & 1.975                        & 0.252              & 18.493                      & -7.194     & 1.880             & 0.073                            \\
+  -  & -             & -                 & Off         & -          & 2.193                        & 0.291              & 34.174                      & 0.431      & 1.867             & 0.104                            \\
+  -  & -             & -                 & -           & propr      & 1.650                        & 0.528              & 1.000                       & 0.000      & 10.583            & 0.036                            \\
+  -  & -             & -                 & -           & harmonies  &                              & 0.108              & 82.633                      & 0.757      & 0.000             & 0.340                            \\
+  -  & -             & -                 & -           & flashweave & 3.878                        & 0.079              & 1.800                       & 0.458      & 0.383             & -0.032                           \\
+  -  & -             & -                 & -           & cozine     &                              & 0.090              & 82.167                      & 0.694      & 0.000             & -0.072                           \\
+  -  & -             & -                 & -           & spring     &                              & 0.092              & 11.767                      & 0.566      & 0.000             & -0.035                           \\
+  -  & -             & -                 & -           & sparcc     & 1.747                        & 0.525              & 1.033                       & 0.009      & 5.500             & -0.011                           \\
+  -  & -             & -                 & -           & spieceasi  &                              & 0.076              & 39.383                      & 0.717      & 0.000             & -0.031                           \\
+  -  & -             & -                 & -           & pearson    &                              & 0.457              & 23.883                      & 0.551      & 0.000             & 0.478                            \\
+  -  & -             & -                 & -           & spearman   & 1.875                        & 0.551              & 1.450                       & 0.044      & 1.117             & 0.170                            \\
+  -  & -             & -                 & -           & mldm       & 2.760                        & 0.130              & 2.267                       & -75.648    & 0.433             & -0.028                           \\ \hline
   \end{tabular}
   \caption{
-    \textbf{Table of global network metrics for all combinations of networks}.
+    \textbf{Table of global network metrics for all combinations of networks.}
     The metrics are averaged for each tool in every step of the pipeline, implying that all combinations of networks that use a particular tool are grouped and averaged and their metrics are reported.
@@
   \end{table}
+  \end{landscape}
 

### 2022-05-04 scripts/figure2/figure2_plot.R
Change: Update font size and scaling of main figures

@@ -25,11 +25,10 @@ plot_scatter <- function(data, color_var) {
     ggscatter(data,
         x = "X0", y = "X1",
         color = color_var,
-        # ellipse = TRUE
     ) +
-    theme(
-        text = element_text(size = 16),
-    )
+        theme(
+            text = element_text(size = 15.5),
+        )
 }
 
 x <- read.csv(x_csv)
@@ -61,9 +60,9 @@ pie_chart <- ggplot(data = percentage_variance, aes(x = "", y = mean_sq, fill = 
         axis.title = element_blank(),
         axis.text.y = element_blank(),
         axis.text.x = element_blank(),
-        legend.position="right",
-        legend.justification="right",
-        text = element_text(size = 16),
+        legend.position = "right",
+        legend.justification = "right",
+        text = element_text(size = 20),
     )
 
 scatter_facet <- ggarrange(scatter_ta, scatter_ni, ncol = 2)

### 2022-05-04 scripts/figure3/figure3_plot.R

@@ -62,7 +62,8 @@ plot_heatmap <- function(data, title) {
             plot.title = element_text(hjust = 0.5),
             axis.title = element_blank(),
             axis.text.x = element_text(angle = 30, hjust = 1),
-            axis.text.y = element_text(angle = 30)
+            axis.text.y = element_text(angle = 30),
+            text = element_text(size = 15),
         ) +
         coord_fixed()
 }
@@ -115,9 +116,10 @@ make_dot_plot <- function(unifrac_data, title) {
             theme(
                 plot.title = element_text(hjust = 0.5),
                 axis.text.x = element_text(angle = 30, hjust = 1),
+                text = element_text(size = 15),
             ),
         fun = "mean_sd"
-    ) #+ stat_compare_means()
+    )
 }
 
 # Weighted

### 2022-05-04 scripts/figure4/figure4_plot.R

@@ -62,24 +62,28 @@ make_alluvial_plot2 <- function(db_data, title) {
     n_genus <- length(genus_factors)
     palette <- get_palette(, n_genus)
     alluvial_plot <- ggplot(db_data, aes(x = database, y = Abundance, stratum = Genus, alluvium = OTU, fill = Genus)) +
-        stat_alluvium(geom="flow", lode.guidance="forward") +
+        stat_alluvium(geom = "flow", lode.guidance = "forward") +
         stat_stratum() +
         scale_fill_manual(values = palette) +
-        geom_text_repel(aes(label=Genus), stat="stratum", size=2, direction="y", nudge_x=0.5) +
+        geom_text_repel(aes(label = Genus), stat = "stratum", size = 2, direction = "y", nudge_x = 0.5) +
         ggtitle(title) +
-        theme_pubr()
+        theme_pubr() +
+        theme(
+            text = element_text(size = 15),
+            legend.position = "none",
+        )
 }
 
 
-gg_n_genus = length(unique(gg$Genus))
-silva_n_genus = length(unique(silva$Genus))
-ncbi_n_genus = length(unique(ncbi$Genus))
+gg_n_genus <- length(unique(gg$Genus))
+silva_n_genus <- length(unique(silva$Genus))
+ncbi_n_genus <- length(unique(ncbi$Genus))
 
 combined_plot <- make_alluvial_plot2(combined, paste("NaiveBayes(GG)=", gg_n_genus, ", NaiveBayes(SILVA)=", silva_n_genus, ", BLAST(NCBI)=", ncbi_n_genus))
 
-final_plot_a <- ggarrange(combined_plot, nrow=1, ncol=1, common.legend=TRUE, legend="right")
+final_plot_a <- ggarrange(combined_plot, nrow = 1, ncol = 1, common.legend = FALSE)
 annotate_figure(final_plot_a, fig.lab = "A", fig.lab.pos = "top.left", fig.lab.size = 20)
-ggsave(output_file_a, width = 11, height = 8.5)
+ggsave(output_file_a, width = 11, height = 10)
 
 
 ################################################################################
@@ -111,9 +115,10 @@ make_bar_plot <- function(data, title) {
         # palette="Paired"
     ) +
         theme(
-        plot.title = element_text(size=10),
[…]

### 2022-05-04 scripts/figure5/figure5_plot.R

@@ -82,7 +82,8 @@ plot_network <- function(network_file, combined_layout, interaction_threshold, t
             axis.ticks = element_blank(),
             panel.background = element_blank(),
             panel.border = element_blank(),
-            panel.grid = element_blank()
+            panel.grid = element_blank(),
+            text = element_text(size = 15),
         )
 }
 
@@ -124,7 +125,7 @@ plot_upset <- function(bmatrix, sets, kind) {
         mainbar.y.label = paste0("Intersection of ", kind),
         sets.x.label = paste0("Number of ", kind),
         queries = list(list(query = intersects, params = sets, color = "red", active = TRUE, query.name = "Common to all")),
-        text.scale = c(2, 1.3, 1.3, 1, 1.3, 1.3),
+        text.scale = c(2.2, 1.5, 1.5, 1, 1.5, 1.5),
     )
 }
 
@@ -180,5 +181,5 @@ bc_plot <- ggarrange(nmatrix_gg, ematrix_gg, labels = c("B", "C"))
 ggsave(output_file_bc, width = 11, height = 5)
 
 
-final_plot <- ggarrange(a_plot, bc_plot, ncol = 1, heights=c(1.3,1.0))
+final_plot <- ggarrange(a_plot, bc_plot, ncol = 1, heights = c(1.3, 1.0))
 ggsave(output_file, final_plot, width = 11, height = 12)

### 2022-05-04 scripts/figure6/figure6_plot.R

@@ -19,12 +19,14 @@ performance_csv <- paste0(data_folder, "performance.csv")
 output_file <- paste0(output_folder, "figure6.pdf")
 
 plot_scatter <- function(data) {
-  ggscatter(data,
-    x = "precision", y = "sensitivity",
-    color = "factor1", shape = "factor2"
-  ) +
-    theme_pubr() #+
-  # stat_stars(aes(color = "factor1"))
+    ggscatter(data,
+        x = "precision", y = "sensitivity",
+        color = "factor1", shape = "factor2"
+    ) +
+        theme_bw() +
+        theme(
+            text = element_text(size = 15),
+        )
 }
 
 norta <- read.csv(performance_csv, sep = ",", header = TRUE)

### 2022-05-04 scripts/figure7/figure7_plot.R

@@ -72,7 +72,7 @@ plot_network <- function(network_file, combined_layout, interaction_threshold, t
     graph_layout$x <- combined_layout[match_inds, ]$x
     graph_layout$y <- combined_layout[match_inds, ]$y
     graph_plot <- ggraph(graph = graph, layout = "manual", x = graph_layout$x, y = graph_layout$y, circular = TRUE) +
-        geom_edge_arc(aes(color = factor(layer), edge_linetype = factor(color), edge_alpha = factor(layer))) +
+        geom_edge_arc(aes(color = layer, edge_linetype = color, edge_alpha = layer)) +
         # geom_node_point(aes(color=factor(colorkey))) +
         geom_node_point() +
         # scale_edge_color_manual(values = c(background=palette[[1]], foreground=palette[[ind]], common="black")) +
@@ -92,6 +92,7 @@ plot_network <- function(network_file, combined_layout, interaction_threshold, t
             panel.background = element_blank(),
             panel.border = element_blank(),
             panel.grid = element_blank(),
+            text = element_text(size = 15),
         )
 }
 
@@ -100,17 +101,16 @@ combined_layout <- create_layout(graph = combined_graph, layout = "linear", circ
 
 palette <- brewer.pal(n = 6, name = "Pastel1")
 default_plot <- plot_network(default_gml, combined_layout, 0.3, "default", palette, 1)
-clustering_plot <- plot_network(clustering_gml, combined_layout, 0.3, "DC=closed_reference(gg_97)", palette, 3)
+clustering_plot <- plot_network(clustering_gml, combined_layout, 0.3, "DC=CR", palette, 3)
 chimera_checking_plot <- plot_network(chimera_checking_gml, combined_layout, 0.3, "CC=uchime", palette, 4)
-database_plot <- plot_network(database_gml, combined_layout, 0.3, "TA=blast(NCBI)", palette, 2)
-otu_filtering_plot <- plot_network(otu_filtering_gml, combined_layout, 0.3, "OP=off", palette, 4)
+database_plot <- plot_network(database_gml, combined_layout, 0.3, "TA=BLAST(NCBI)", palette, 2)
+otu_filtering_plot <- plot_network(otu_filtering_gml, combined_layout, 0.3, "OP=Filter(off)", palette, 4)
 network_inference_plot <- plot_network(network_inference_gml, combined_layout, 0.3, "NI=SparCC", palette, 5)
[…]

### 2022-05-04 scripts/figure_generation/generate_figures.sh

@@ -76,3 +76,5 @@ Rscript figure_s3_plot.R \
   "../../data/figure3/output/$DATASET/" \
   "../../data/figure3/output/" \
   "$OUTPUT_DIR"
+
+# NOTE: Other supplementary figures should be generated locally

### 2022-05-04 scripts/figure6/figure6_data.py
Change: Supl figure font sizes

@@ -169,7 +169,7 @@ def fix_name(name: str) -> str:
         return f"X:SV[{parameter_value}]"
     elif name.startswith("pvalue_merging"):
         parameter_value = name.rsplit("_", 1)[-1]
-        return "X:Pvalue Merging"
+        return "X:PM"
     else:
         return f"I:{name}"
 

### 2022-05-04 scripts/figure_s3/figure_s3_plot.R

@@ -54,7 +54,8 @@ plot_heatmap <- function(data, title) {
             plot.title = element_text(hjust = 0.5),
             axis.title = element_blank(),
             axis.text.x = element_text(angle = 30, hjust = 1),
-            axis.text.y = element_text(angle = 30)
+            axis.text.y = element_text(angle = 30),
+            text = element_text(size = 15),
         ) +
         coord_fixed()
 }

### 2022-05-04 scripts/figure_s4/figure_s4_plot.R

@@ -32,11 +32,14 @@ tidy_up_data <- function(unifrac_data) {
         distinct(method, .keep_all = TRUE)
 }
 
-plot_dotplot <- function(data, title) {
+plot_bar <- function(data, title) {
     ggbarplot(data, x = "method", y ="unifrac", fill="steelblue", label = TRUE, lab.pos = "out", lab.nb.digits = 3) +
         ylim(0, 1) +
         ggtitle(title) +
-        theme_pubr()
+        theme_pubr() +
+        theme(
+            text = element_text(size = 15),
+        )
 }
 
 unweighted <- read.csv(unweighted_unifrac_csv, sep = ",", header = TRUE)
@@ -44,11 +47,11 @@ unweighted_tidy <- tidy_up_data(unweighted)
 weighted <- read.csv(weighted_unifrac_csv, sep = ",", header = TRUE)
 weighted_tidy <- tidy_up_data(weighted)
 
-weighted_plot <- plot_dotplot(weighted_tidy, "weighted unifrac")
-unweighted_plot <- plot_dotplot(unweighted_tidy, "unweighted unifrac")
+weighted_plot <- plot_bar(weighted_tidy, "weighted unifrac")
+unweighted_plot <- plot_bar(unweighted_tidy, "unweighted unifrac")
 
-dotplot_plot <- ggarrange(
+bar_plot <- ggarrange(
     weighted_plot, unweighted_plot,
     nrow = 1, ncol = 2, labels = c("A", "B"), common.legend = TRUE, legend = "right"
 )
-ggsave(output_file, dotplot_plot, width = 11, height = 5.5)
+ggsave(output_file, bar_plot, width = 11, height = 5.5)

### 2022-05-04 scripts/figure_s5/figure_s5_plot.R

@@ -11,7 +11,7 @@ args <- commandArgs(trailingOnly = TRUE)
 if (length(args) == 0) {
     data_folder <- "../../data/figure_s5/output/moving_pictures/"
     mock_folder <- "../../data/figure_s5/output/"
-    output_folder <- "../../figures/"
+    output_folder <- "../../figures/FMT/"
 } else if (length(args) == 3) {
     data_folder <- args[1]
     mock_folder <- args[2]
@@ -55,7 +55,8 @@ make_bar_plot <- function(data, title) {
     ) +
         theme(
         plot.title = element_text(size=10),
-        axis.text.x = element_text(angle = 30, hjust = 1)
+        axis.text.x = element_text(angle = 30, hjust = 1),
+        text = element_text(size = 15),
     )
 }
 
@@ -73,4 +74,4 @@ combined_plot_b <- ggarrange(
     nrow = 1, ncol = 3, common.legend = TRUE, legend = "right"
 )
 # annotate_figure(combined_plot_b, fig.lab = "A", fig.lab.pos = "top.left", fig.lab.size = 20)
-ggsave(output_file, width = 11, height = 4)
+ggsave(output_file, width = 11, height = 6)

### 2022-05-04 scripts/figure_s6/figure_s6_plot.R

@@ -8,7 +8,7 @@ library(ggpubr)
 args <- commandArgs(trailingOnly = TRUE)
 if (length(args) == 0) {
     data_folder <- "../../data/figure_s6/output/norta/"
-    output_folder <- "../../figures/"
+    output_folder <- "../../figures/FMT/"
 } else if (length(args) == 2) {
     data_folder <- args[1]
     output_folder <- args[2]
@@ -23,8 +23,10 @@ plot_scatter <- function(data) {
     x = "precision", y = "sensitivity",
     color = "factor1", shape = "factor2"
   ) +
-    theme_pubr() #+
-  # stat_stars(aes(color = "factor1"))
+    theme_bw() +
+    theme(
+        text = element_text(size = 15),
+    )
 }
 
 norta <- read.csv(performance_csv, sep = ",", header = TRUE)

### 2022-05-04 scripts/figure_s7/figure_s7_plot.R

@@ -8,7 +8,7 @@ library(ggpubr)
 args <- commandArgs(trailingOnly = TRUE)
 if (length(args) == 0) {
     data_folder <- "../../data/figure_s7/output/seqtime/"
-    output_folder <- "../../figures/"
+    output_folder <- "../../figures/FMT/"
 } else if (length(args) == 2) {
     data_folder <- args[1]
     output_folder <- args[2]
@@ -23,8 +23,10 @@ plot_scatter <- function(data) {
     x = "precision", y = "sensitivity",
     color = "factor1", shape = "factor2"
   ) +
-    theme_pubr() #+
-  # stat_stars(aes(color = "factor1"))
+    theme_bw() +
+    theme(
+        text = element_text(size = 15),
+    )
 }
 
 seqtime <- read.csv(performance_csv, sep = ",", header = TRUE)

### 2022-05-05 sections/methods.tex
Change: Merge the tool tables in the methods section

@@
   The quality filtering and chimera checking tools are derived from those used in \ac{qiime2} v2021.8.0 (uchime-denovo method) and \ac{dada2} (remove bimera method).
-  The list of tools used in this step, along with their modules and references are provided in Table~\ref{tab:dc_tools}.
+  The list of tools used in this step, along with their modules and references are provided in Table~\ref{tab:micone_tools}.
 
@@
   The consensus algorithm used is the default method used by the classifiers in \ac{qiime2}.
-  The list of tools used in this step, along with their modules and references are provided in Table~\ref{tab:ta_tools}.
+  The list of tools used in this step, along with their modules and references are provided in Table~\ref{tab:micone_tools}.
 
@@
   When the OP module is ``off'', the filtering is still performed but only to remove those features that are very low in presence.
-  The parameters used are given in Table~\ref{tab:op_tools}.
+  The parameters used are given in Table~\ref{tab:micone_tools}.
   The forking operation splits the count matrix into multiple matrices based on sample metadata column.
@@
   These bootstrapped association matrices are used to calculate the significance of the inferred associations by calculating the p-values against this null model~\cite{Watts2018}.
-  The pipeline includes 4 methods for pairwise correlation metrics, and 6 methods for direct association metrics (refer Table~\ref{tab:ni_tools}).
+  The pipeline includes 4 methods for pairwise correlation metrics, and 6 methods for direct association metrics (refer Table~\ref{tab:micone_tools}).
   The Brown's p-value merging method~\cite{brown_400_1975} is used for combining p-values from the pairwise correlations methods to obtain a consensus p-value, which can be used to filter for significance.
@@
   The output of this step are co-occurrence association networks encoded in the JSON format (refer Supplementary section) and are also exportable to a variety of network formats like .gml.
-  The list of tools used in this step, along with their modules and references are provided in Table~\ref{tab:ni_tools}.
+  The list of tools used in this step, along with their modules and references are provided in Table~\ref{tab:micone_tools}.
 

### 2022-05-05 sections/supplementary.tex
Change: Supl figure font sizes
Change: Merge the tool tables in the methods section

@@
             The ``I'' prefix indicates the individual network inference methods and the ``X'' prefix denotes the different consensus based methods.
-            The different consensus based methods used are: pvalue merging, scaled-sum (SS) and simple voting (SV) method.
+            The different consensus based methods used are: pvalue merging (PM), scaled-sum (SS) and simple voting (SV) method.
             With the addition of Pearson and Spearman in the consensus calculation we observe a decrease in the average precision scores of both the consensus algorithms.
@@
           The ``I'' prefix indicates the individual network inference methods and the ``X'' prefix denotes the different consensus based methods.
-          The different consensus based methods used are: pvalue merging, scaled-sum (SS) and simple voting (SV) method.
+          The different consensus based methods used are: pvalue merging (PM), scaled-sum (SS) and simple voting (SV) method.
           Pearson and Spearman methods are not used in the calculation of the consensus.
@@
 
-    The default format \ac{micone} uses for storing the network files is the JSON format.
+    The default format \ac{micone} uses for storing the network files is the JSON (JavaScript Object Notation) format.
     The custom JSON schema we have designed is able to store all network related information from nodes and links to metadata related to links and datasets.

### 2022-05-05 sections/tables_figures.tex
Change: Update font size and scaling of main figures
Change: Supl figure font sizes
Change: Merge the tool tables in the methods section

@@
     \small
+    \begin{tabular}{|c|c|c|c|}
+      \hline
+      \textbf{Workflow step} & \textbf{Module/Condition} & \textbf{Tool/Parameter} & \textbf{References/Value} \\
+      \hline
+      \multirow{7}{*}{Denoising and Clustering} & \multirow{5}{*}{Denoising and Clustering} & Closed reference & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
+                                                &  & Open reference & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
+                                                &  & De novo & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
+                                                &  & Dada2 & \cite{Callahan2016} \\
+                                                &  & Deblur & \cite{Amir2017,bolyenReproducibleInteractiveScalable2019} \\
+                                                \cline{2-4}
+                                                & \multirow{2}{*}{Chimera checking} & Uchime-denovo & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
+                                                & & Remove bimera & \cite{Callahan2016} \\
+                                                \hline
+      \multirow{5}{*}{Taxonomy assignment} &  \multirow{2}{*}{Query tool} & Blast & \cite{camachoBLASTArchitectureApplications2009,bokulichOptimizingTaxonomicClassification2018} \\
+                                           &  & Naive bayes classifier & \cite{bokulichOptimizingTaxonomicClassification2018} \\
+                                           \cline{2-4}
+                                           & \multirow{3}{*}{Database} & Greengenes & \cite{DeSantis2006} \\
+                                           & & SILVA & \cite{Quast2012} \\
+                                           & & NCBI RefSeq & \cite{Sayers2009} \\
+      \hline
+      \multirow{6}{*}{OTU processing} & \multirow{3}{*}{Filter(off)} & Prevalence threshold      & 2 / n\_samples \\
+                                      & & Abundance threshold       & 0.001          \\
+                                      & & Observation sum threshold & 10             \\ \cline{2-4}
+                                      & \multirow{3}{*}{Filter(on)}  & Prevalence threshold      & 0.05           \\
+                                      & & Abundance threshold       & 0.01           \\
+                                      & & Observation sum threshold & 100            \\
+      \hline
+      \multirow{12}{*}{Network Inference} & \multirow{2}{*}{Bootstrapping}& \texttt{fastspar\_bootstraps} v1.0 & \cite{Watts2018} \\
+                                          & & \texttt{fastspar\_pvalues} v1.0 & \cite{Watts2018} \\
+                                          \cline{2-4}
+                                          & \multirow{6}{*}{Direct association} & \ac{spieceasi} v1.1.2 & \cite{Kurtz2015} \\
+                                          & & FlashWeave.jl v0.18.1 & \cite{tackmannRapidInferenceDirect2019} \\
+                                          & & \ac{cozine} v1.0 & \cite{haCompositionalZeroinflatedNetwork2020a} \\
+                                          & & \ac{harmonies} v1.0 & \cite{jiangHARMONIESHybridApproach2020} \\
+                                          & & \ac{spring} v1.0.4 & \cite{yoonMicrobialNetworksSPRING2019} \\
+                                          & & \ac{mldm} v1.1 & \cite{Yang2017} \\
+                                          \cline{2-4}
+                                          & \multirow{2}{*}{Correlation-based} & FastSpar (\ac{sparcc}) v1.0 & \cite{Watts2018} \\
+                                          & & Pearson & - \\
+                                          & & Spearman & - \\
+                                          & & propr v2.1.2 & \cite{quinnProprRpackageIdentifying2017} \\
+      \hline
+    \end{tabular}
+    \caption{Tools used in the \ac{micone} pipeline}
+    \label{tab:micone_tools}
+  \end{table}
+
+  \begin{table}[h]
+    \centering
+    \small
     \begin{tabular}{|c|c|c|}
       \hline
-      \textbf{Process} & \textbf{Tool} \\
+      \textbf{Workflow step} & \textbf{Default tool} \\
       \hline
@@
     \end{tabular}
-    \caption{Default tools and parameters for the pipeline}
+    \caption{Default tools and parameters of the \ac{micone} pipeline}
     \label{tab:default_options}
-  \end{table}
-
-  \begin{table}[h]
-    \centering
-    \small
-    \begin{tabular}{|c|c|c|}
-      \hline
-      \textbf{Module} & \textbf{Tool} & \textbf{References} \\
-      \hline
-      Denoising and Clustering & Closed reference & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
-      Denoising and Clustering & Open reference & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
-      Denoising and Clustering & De novo & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
-      Denoising and Clustering & Dada2 & \cite{Callahan2016} \\
-      Denoising and Clustering & Deblur & \cite{Amir2017,bolyenReproducibleInteractiveScalable2019} \\
-      \hline
-      Chimera checking & Uchime-denovo & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
-      Chimera checking & Remove bimera & \cite{Callahan2016} \\
-      \hline
-    \end{tabular}
-    \caption{Tools used in the DC module}
-    \label{tab:dc_tools}
-  \end{table}
-
-  \begin{table}[h]
-    \centering
-    \small
-    \begin{tabular}{|c|c|c|}
-      \hline
-      \textbf{Module} & \textbf{Tool/Database} & \textbf{References} \\
-      \hline
-      Query tool & Blast & \cite{camachoBLASTArchitectureApplications2009,bokulichOptimizingTaxonomicClassification2018} \\
-      Query tool & Naive bayes classifier & \cite{bokulichOptimizingTaxonomicClassification2018} \\
-      \hline
-      Database & Greengenes & \cite{DeSantis2006} \\
-      Database & SILVA & \cite{Quast2012} \\
-      Database & NCBI RefSeq & \cite{Sayers2009} \\
-      \hline
-    \end{tabular}
-    \caption{Tools used in the TA module}
-    \label{tab:ta_tools}
-  \end{table}
-
-  \begin{table}[h]
-    \centering
-    \small
-    \begin{tabular}{|c|c|c|}
-    \hline
-    Condition                    & Parameter                 & Value          \\ \hline
-    \multirow{3}{*}{Filter(off)} & Prevalence threshold      & 2 / n\_samples \\ \cline{2-3}
-                                 & Abundance threshold       & 0.001          \\ \cline{2-3}
-                                 & Observation sum threshold & 10             \\ \hline
-    \multirow{3}{*}{Filter(on)}  & Prevalence threshold      & 0.05           \\ \cline{2-3}
-                                 & Abundance threshold       & 0.01           \\ \cline{2-3}
-                                 & Observation sum threshold & 100            \\ \hline
-    \end{tabular}
-    \caption{Parameters used in the OP module}
-    \label{tab:ni_tools}
-  \end{table}
-
-  \begin{table}[h]
-    \centering
-    \small
-    \begin{tabular}{|c|c|c|}
-      \hline
-      \textbf{Module} & \textbf{Tool/Database} & \textbf{References} \\
-      \hline
-      Booststrap & \texttt{fastspar\_bootstraps} v1.0 & \cite{Watts2018} \\
-      Booststrap & \texttt{fastspar\_pvalues} v1.0 & \cite{Watts2018} \\
-      \hline
-      Direct & \ac{spieceasi} v1.1.2 & \cite{Kurtz2015} \\
-      Direct & FlashWeave.jl v0.18.1 & \cite{tackmannRapidInferenceDirect2019} \\
-      Direct & \ac{cozine} v1.0 & \cite{haCompositionalZeroinflatedNetwork2020a} \\
-      Direct & \ac{harmonies} v1.0 & \cite{jiangHARMONIESHybridApproach2020} \\
-      Direct & \ac{spring} v1.0.4 & \cite{yoonMicrobialNetworksSPRING2019} \\
-      Direct & \ac{mldm} v1.1 & \cite{Yang2017} \\
-      \hline
-      Correlation & FastSpar (\ac{sparcc}) v1.0 & \cite{Watts2018} \\
-      Correlation & Pearson & - \\
-      Correlation & Spearman & - \\
-      Correlation & propr v2.1.2 & \cite{quinnProprRpackageIdentifying2017} \\
-      \hline
-    \end{tabular}
-    \caption{Tools used in the NI module}
-    \label{tab:ni_tools}
   \end{table}
@@
     \centering
-    \includegraphics[width=0.75\linewidth]{figure1.pdf}
+    \includegraphics[width=0.74\linewidth]{figure1.pdf}
   \end{figure}
@@
       The ``I'' prefix indicates the individual network inference methods and the ``X'' prefix denotes the different consensus based methods.
-      The different consensus based methods used are: pvalue merging, scaled-sum (SS) and simple voting (SV) method.
+      The different consensus based methods used are: pvalue merging (PM), scaled-sum (SS) and simple voting (SV) method.
       Pearson and Spearman methods are not used in the calculation of the consensus.

### 2022-05-05 sections/tables_figures.tex
Add: Updated parameter table for SP step

@@
       \hline
-      \multirow{7}{*}{Denoising and Clustering} & \multirow{5}{*}{Denoising and Clustering} & Closed reference & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
+      \multirow{7}{*}{Denoising and Clustering} & \multirow{5}{*}{Denoise and Cluster} & Closed reference & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
                                                 &  & Open reference & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\

### 2022-05-06 sections/abstract.tex
Change: Update Abstract + Importance for draft v2

@@
     Through this process, we map how different choices of algorithms and parameters affect the co-occurrence network and estimate steps that contribute most significantly to the variance.
-We further determine the tools and parameters that generate the most accurate and robust co-occurrence networks based on comparison with mock and synthetic datasets.
+    We further determine the tools and parameters that generate the most accurate and robust co-occurrence networks and develop consensus network algorithms based on benchmarks with mock and synthetic datasets.
     Ultimately, we develop a standardized pipeline (available at \href{https://github.com/segrelab/MiCoNE}{https://github.com/segrelab/MiCoNE}) that follows these default tools and parameters, but that can also help explore the outcome of any other combination of choices.
-    We envisage that this pipeline could be used for integrating multiple data-sets, and for generating comparative analyses and consensus networks that can help understand and control microbial community assembly in different biomes.
+    We envisage that this pipeline could be used for integrating multiple data-sets, and for generating comparative analyses and consensus networks that can help understand microbial community assembly in different biomes.
   }
@@
 % Insert keywords here
-\keywords{Microbiome, 16S rRNA, Pipeline, Interaction, Denoising, Taxonomy, Network Inference, Correlations, Qiime, Co-occurrence, Networks}
+\keywords{Microbiome, 16S rRNA, Pipeline, Interaction, Denoising, Taxonomy, Network Inference, Correlations, Qiime2, Co-occurrence, Networks, Consensus algorithm}
 
 \section*{Importance}
-
   To understand and control the mechanisms that determine the structure and function of microbial communities, it is important to map the interrelationships between its constituent microbial species.
@@
   These abundances can be transformed into networks of co-occurrences across multiple samples, providing a glimpse into the structure of microbiomes.
-However, processing these datasets to obtain co-occurrence information relies on several complex steps, each of which involves multiple choices of tools and corresponding parameters.
+  However, processing these datasets to obtain co-occurrence information relies on several complex steps, each of which involves numerous choices of tools and corresponding parameters.
   These multiple options pose questions about the accuracy and uniqueness of the inferred networks.
   In this study, we address this workflow and provide a systematic analysis of how these choices of tools and parameters affect the final network, and on how to select those that are most appropriate for a particular dataset.
+  We also develop a consensus network algorithm that helps generate more robust and accurate occurrence networks based on benchmarks synthetic datasets.

### 2022-05-06 sections/introduction.tex
Change: Update Introduction for draft v2

@@
 
-  Microbial communities are ubiquitous and play an important role in marine and terrestrial environments, urban ecosystems, metabolic engineering, and human health \cite{Ghoul2016,Thompson2017}.
+  Microbial communities are ubiquitous and play an important role in marine and terrestrial environments, urban ecosystems, and human health \cite{Ghoul2016,Thompson2017}.
   These microbial communities, or microbiomes, often comprise several hundreds of different microbial strains interacting with each other and their environment, often through intricate metabolic and signaling relationships.
-  Understanding how these interconnections shape community structure and functionalities is a fundamental challenge in microbial ecology, with applications in the study of microbial ecosystems across different biomes.
+  Understanding how these interconnections shape community structure and functionalities is a fundamental challenge in microbial ecology, and has applications in the study of microbial ecosystems across different biomes.
   With the advancement in DNA sequencing technologies \cite{Narihiro2017} and data processing methods,  more information can be extracted from these microbial community samples than ever before.
-  In particular, high-throughput sequencing, including community metagenomic sequencing and sequencing of 16S rRNA gene amplicons, has the potential to help detect, identify and quantify a large portion of the constitutive microorganisms of a microbiome \cite{Jovel2016,Lloyd-Price2016}.
+  In particular, high-throughput sequencing, including metagenomic sequencing and sequencing of 16S rRNA gene amplicons of microbial communities, has the potential to help detect, identify and quantify a large portion of the constitutive microorganisms of a microbiome \cite{Jovel2016,Lloyd-Price2016}.
   These advances have led to large-scale data collection efforts involving environmental (\acl{emp}) \cite{Thompson2017}, marine (Tara Oceans Project) \cite{Zhang2015} and human-associated microbiota (Human Microbiome Project) \cite{HumanMicrobiomeProjectConsortium2012}.
@@
  This wealth of information on the composition and functions of a community at different times and under different environmental conditions has the potential to help us understand how communities assemble and operate.
- A powerful tool for translating microbiome data into knowledge is the construction of possible inter-dependence networks across species.
- The importance of these networks of relationships is two fold: first, such networks can serve as maps that help identify hubs of keystone species \cite{Menon2018,Rottjers2018}, or basic microbiome changes that occur as a consequence of environmental perturbations or underlying host conditions \cite{Gilbert2016}; second, networks of inter-dependencies can serve as a key bridge towards building mechanistic models of microbial communities, greatly enhancing our capacity to understand and control them.
+ A powerful tool for translating microbiome composition data into knowledge is the construction of possible inter-dependence networks across species.
+ The importance of these networks of relationships is two fold: first, such networks can serve as maps that help identify hubs of keystone species \cite{Menon2018,Rottjers2018}, and rewiring of interactions that occur as a consequence of environmental perturbations or underlying host conditions \cite{Gilbert2016}; second, networks of interdependencies can serve as a key bridge towards building mechanistic models of microbial communities, greatly enhancing our capacity to understand and control them.
  For example, multiple studies have shown the importance of specific microbial interactions in the healthy microbiome \cite{Lloyd-Price2016} and others have shown how changes in these interactions can lead to dysbiosis \cite{Wang2017,Gilbert2016,Belizario2015}.
@@
 through co-culture micro-droplet experiments \cite{Hsu2019,Jian2020}, or spatial visualization of natural communities \cite{Wilbert2020} is possible, but it requires specific technological capabilities, and has yet to be extensively used.
- In parallel, sequencing data across multiple samples can be used for estimating co-occurrence relationships between taxa.
- While the the relationship between directly measured interactions and statistically inferred co-occurrence is still poorly understood \cite{Zuniga2017}, a significant amount of effort has gone into estimating correlations from large microbiome sequence datasets.
- Co-occurrence networks have microbial taxa as nodes, and edges that represent the frequent co-occurrence (or negative correlations) across different datasets.
+ However, sequencing data across multiple samples, which is more readily available, can be used for estimating co-occurrence relationships between taxa.
+ While the relationship between directly measured interactions and statistically inferred co-occurrence is still poorly understood \cite{Zuniga2017}, a significant amount of effort has gone into estimating correlations from large microbiome sequence datasets.
+ Co-occurrence networks have microbial taxa as nodes, and edges that represent the frequent co-occurrences (or negative correlations) across different datasets.
 
 One of the most frequently used avenues for inferring co-occurrence networks is the parsing and analysis of 16S sequencing data \cite{Rottjers2018,Friedman2012}.
-A large number of software tools and pipelines have been developed to analyze 16S sequencing data, often focused on addressing the many known limitations of this methodology, including resolution, sequencing depth, compositional nature, sequencing errors and copy number variations \cite{Bharti2019,Pollock2018}.
-Popular methods for different phases of the analysis of 16S data include tools for: (i) denoising and clustering sequencing reads \cite{Caporaso2010,Callahan2016}; (ii) assigning taxonomy to the reads \cite{DeSantis2006,Quast2012}; (iii) processing and transforming the taxonomy count matrices \cite{Weiss2015}; and (iv) inferring the co-occurrence network \cite{Cougoul2019,Kurtz2015}.
-Different specific algorithms are often aggregated into popular platforms (like MG-RAST \cite{Keegan2016}, Qiita \cite{qiita}) and packages (such as QIIME \cite{Caporaso2010}) that provide pipelines for 16S data analysis.
+Numerous software tools and pipelines have been developed to analyze 16S sequencing data, often focused on addressing the many known limitations of this methodology, including resolution, sequencing depth, compositional nature, sequencing errors and copy number variations \cite{Bharti2019,Pollock2018}.
+Popular methods for different phases of the analysis of 16S data include tools for: (i) quality checking and trimming the sequencing reads; (ii) denoising and clustering the trimmed reads \cite{Caporaso2010,Callahan2016}; (iii) assigning taxonomy to the denoised reads \cite{DeSantis2006,Quast2012}; (iv) processing and transforming the taxonomy count matrices \cite{Weiss2015}; and (v) inferring the co-occurrence network \cite{Cougoul2019,Kurtz2015}.
+Different specific algorithms are often aggregated into popular platforms (like MG-RAST~\cite{Keegan2016}, Qiita~\cite{qiita}) and packages (such as QIIME2~\cite{bolyenReproducibleInteractiveScalable2019}) that provide pipelines for 16S data analysis.
 The different methods and tools developed to solve issues arising in 16S analysis can lead to vastly different inferences of community compositions and co-occurrence networks \cite{Golob2017,Weiss2016}, making it difficult to reliably compare networks across different publications and studies.
@@
   Our pipeline is coupled to an online integrative tool for the organization, visualization and analysis of inter-microbial networks.
-  In addition to making this tool freely available, we implemented a systematic comparative analysis to determine which steps of the pipeline have the largest influence on the final network, and what choice seems to provide best agreement with the tested mock and synthetic datasets.
+  In addition to making this tool freely available, we completed a systematic comparative analysis to determine which steps of the pipeline have the largest influence on the final network, and what choice seems to have the best agreement with the tested mock and synthetic datasets.
   We believe that these steps will ensure better reproducibility and easier comparison of co-occurrence networks across datasets.

### 2022-05-06 sections/introduction.tex
Change: Update the Results for draft v2

@@
  A powerful tool for translating microbiome composition data into knowledge is the construction of possible inter-dependence networks across species.
- The importance of these networks of relationships is two fold: first, such networks can serve as maps that help identify hubs of keystone species \cite{Menon2018,Rottjers2018}, and rewiring of interactions that occur as a consequence of environmental perturbations or underlying host conditions \cite{Gilbert2016}; second, networks of interdependencies can serve as a key bridge towards building mechanistic models of microbial communities, greatly enhancing our capacity to understand and control them.
+ The importance of these networks of relationships is two-fold: first, such networks can serve as maps that help identify hubs of keystone species \cite{Menon2018,Rottjers2018}, and rewiring of interactions that occur as a consequence of environmental perturbations or underlying host conditions \cite{Gilbert2016}; second, networks of interdependencies can serve as a key bridge towards building mechanistic models of microbial communities, greatly enhancing our capacity to understand and control them.
  For example, multiple studies have shown the importance of specific microbial interactions in the healthy microbiome \cite{Lloyd-Price2016} and others have shown how changes in these interactions can lead to dysbiosis \cite{Wang2017,Gilbert2016,Belizario2015}.

### 2022-05-06 sections/results.tex
Change: Update the Results for draft v2

--- 2022-04-25 sections/results.tex
@@
 
-  We have developed \ac{micone}, a flexible and modular pipeline for 16S amplicon sequencing rRNA data (hereafter mentioned simply as 16S data) analysis, that allows us to infer microbial co-occurrence networks.
+  We developed \ac{micone}, a flexible and modular pipeline for 16S amplicon sequencing rRNA data (hereafter mentioned simply as 16S data) analysis.
   It incorporates various popular, publicly available tools as well as custom Python modules and scripts to facilitate inference of co-occurrence networks from 16S data (see Methods).
-  Using \ac{micone} one can obtain co-occurrence networks by applying to 16S data (or to already processed taxonomic count matrices) any combination of the available tools.
-  The effects of changing any of the intermediate step can be monitored and evaluated in terms of its final network outcome, as well as on any of the intermediate metrics and data outputs.
-  The \ac{micone} pipeline workflow is shown in Figure~\ref{fig:figure1}.
-  The different steps for going from 16S data to co-occurrence networks can be grouped into five major modules; (i) the sequence processing (SP) step, which handles the demultiplexing, quality control and trimming of the raw 16S sequencing data into demultiplexed and trimmed sequences; (ii) the denoising and clustering (DC) step, which handles denoising of the demultiplexed and trimmed sequences into representative sequences; (iii) the taxonomy assignment (TA) step that assigns taxonomic labels to the representative sequences; (iv) the \ac{otu} processing (OP) step that filters and transforms the taxonomy abundance table; and finally (v) the network inferences (NI) step which infers the microbial co-occurrence network.
+  Using \ac{micone} one can obtain co-occurrence networks by applying to 16S data (or to already processed taxonomic count matrices) any combinations of the supported tools.
+  The effects of changing any intermediate step can be monitored and evaluated in terms of its final network outcome, as well as on any of the intermediate metrics and data outputs.
+  The main workflow of the \ac{micone} pipeline is shown in Figure~\ref{fig:figure1}.
+  The different steps that are a part of the co-occurrence network inference workflow can be grouped into five major modules; (i) the sequence processing (SP) step, which handles the demultiplexing, quality control and trimming of the raw 16S sequencing data into demultiplexed and trimmed sequences; (ii) the denoising and clustering (DC) step, which handles denoising of the demultiplexed and trimmed sequences into representative sequences; (iii) the taxonomy assignment (TA) step that assigns taxonomic labels to the representative sequences; (iv) the \ac{otu} processing (OP) step that filters and transforms the taxonomy abundance table; and finally (v) the network inferences (NI) step which infers the microbial co-occurrence network.
   Each process in the pipeline supports alternate tools for performing the same task (see Methods and Figure~\ref{fig:figure1}).
-  A centralized configuration file contains all the specifications for what modules are used in the pipeline, and can be modified by the user to choose the desired set of tools.
-  The default tools and parameters for the configuration file can be found in the documentation (see Methods).
-  In what follows, we perform a systematic analysis of each step of the pipeline to estimate how much the final co-occurrence network depends on the possible choices at each step.
-  We also evaluate a large number of tool combinations to determine a set of recommended default options for the pipeline and provide the users with a set of guidelines to facilitate tool selection as appropriate for their data.
+  A centralized configuration file contains all the specifications regarding the modules used in the pipeline, and can be modified by the user to choose the desired set of tools.
+  The default tools and parameters for the configuration file can be found in the \ac{micone} documentation.
+  In the current study, we performed a systematic analysis of each step of the pipeline to estimate how much the final co-occurrence network depends on the possible choices at each step.
+  We also evaluated numerous tool combinations to determine a set of recommended default options for the pipeline and provide the users with a set of guidelines to facilitate tool selection as appropriate for their data.
 
@@
   The second are datasets synthetically or artificially created for the specific goal of helping evaluate computational analysis tools.
-  In particular, in order to objectively compare, to the extent possible, how well each step in \ac{micone} best captures the underlying data, we use both mock data (labelled mock4, mock12 and mock16) from mockrobiota~\cite{Bokulich2016} as well as, synthetically generated networks using the NorTA~\cite{Kurtz2015} and seqtime~\cite{Rottjers2018}.
+  In particular, in order to objectively compare, to the extent possible, how well each step in \ac{micone} best captures the underlying data, we use both mock data (labelled mock4, mock12 and mock16) from mockrobiota~\cite{Bokulich2016} as well as, synthetically networks generated using the NorTA~\cite{Kurtz2015} and seqtime~\cite{Rottjers2018} methods.
   Detailed information on the mock communities and the settings used to generate the synthetic data are provided in the Methods section.
@@
 
-  In order to analyze the effect of different statistical methods on the inferred co-occurrence networks, we generated co-occurrence networks using all possible combinations of methods and estimated the variability in the networks due to each choice (Figure \ref{fig:figure2}A).
-  This analysis and all subsequent analysis, unless explicitly mentioned, are performed on all the samples in the fecal microbiome transplant study.
+  In order to analyze the effect of different processing methods on the inferred co-occurrence networks, we generated co-occurrence networks using all possible combinations of methods and estimated the variability in the networks due to each choice (Figure \ref{fig:figure2}A).
+  All the network analysis in the study, unless explicitly mentioned, are performed only on the healthy samples in the fecal microbiome transplant study.
+  However, all other analysis are performed on all the samples in the fecal microbiome transplant study.
   The networks are filtered on both association strength (0.1) and pvalue (0.05, only correlation methods) before the networks are subjected to the analyses.
-  The effects of various steps on the final co-occurrence network is estimated by building a linear model of the edges of the network as a function the various step in the analysis pipeline (see Methods).
+  The effects of various steps on the final co-occurrence network is estimated by building a linear model of the edges of the network as a function the various steps in the pipeline workflow (see Methods).
   Figure \ref{fig:figure2}A, shows the percentage of total variation among the co-occurrence networks due to the different steps of the pipeline.
-  The TA step, more specifically, the choice of 16S reference database contributes the most ($\sim65.4\%$) to the variation in the networks, followed by the OP step ($\sim26.8\%$).
+  The \ac{ta} step, more specifically, the choice of 16S reference database contributes the most ($\sim65.4\%$) to the variation in the networks, followed by the \ac{op} step ($\sim26.8\%$).
   This result highlights the importance of the taxonomy assignment step in the 16S data analysis workflow, implying that a change in the reference database will result in drastically different inferred networks.
   The reason for this seems to derive from the differential labeling of constitutive representative sequences into taxonomic entities (Figure~\ref{fig:figure4} and Figure~\ref{fig:figure_s5}), which drastically alter the nodes and hence the underlying network topology.
-  In fact the variability induced by taxonomy assignment is much more significant than that due to the variability induced based on how the reference sequences themselves are identified (in the DC step).
-  The residual variation (Figure \ref{fig:figure2}A) can be seen as an artifact that arises when multiple steps are changed at the same time, however this is small.
+  In fact the variability induced by taxonomy assignment is much more significant than that due to the variability introduced based on how the reference sequences themselves are identified (in the DC step).
+  The residual variation (Figure \ref{fig:figure2}A) can be seen as an artifact that arises when multiple steps are changed at the same time, however this is negligible.
 
   Figure \ref{fig:figure2}B is the PCA plot of all the generated co-occurrence networks, colored by the databases in the TA step (left) and the network inference algorithms in the NI step (right).
-  In other words, each point corresponds to a network generated using different combination of tools, and captures how much the final network is affected by such choice.
-  The importance of the TA step indicated in Figure \ref{fig:figure2}A is also reflected here, in the fact that most of the networks can be clearly separated based on the database used (Figure \ref{fig:figure2}B).
-  However, it is also evident that some of the networks, especially those inferred using the direct association based network infernce methods, are much closer in the PCA plot regardless of the reference database used.
+  In other words, each point corresponds to a network generated using different combinations of tools, and captures how much the final network is affected by such a choice.
+  The importance of the TA step which was implied in Figure \ref{fig:figure2}A is also reflected here, since most of the networks can be clearly separated based on the database used (Figure \ref{fig:figure2}B).
+  However, it is also evident that some networks, especially those inferred using the direct association based network inference methods, are much closer in the PCA plot regardless of the reference database used.
   It can also be seen that the variation (on the first two principal axes) between the networks decreases when the low abundance \ac{otu}s are removed from the network (Figure~\ref{fig:figure_s1}).
-  These results suggest that the most important criterion for accurate comparative analyses of co-occurrence networks is the taxonomy reference database and the level of filtering of the taxonomy tables prior to network inference.
+  These results suggest that the most important criterion for accurate comparative analysis of co-occurrence networks is the taxonomy reference database followed by the level of filtering of the taxonomy tables prior to network inference.
 
@@
 
-  Denoising and clustering are commonly carried out to generate representative sequences from the demultiplexed and trimmed 16S sequencing data to obtain the \ac{otu}/\ac{esv} tables (counts of these representative sequences for each sample).
-  In order to compare the representative sequence count tables generated by different tools we processed the same 16S sequencing reads (samples from a fecal microbiome transplant study~\cite{Kang2017}) using 5 different methods: open-reference clustering, closed-reference clustering, denovo clustering, \ac{dada2}~\cite{Callahan2016} and Deblur~\cite{Amir2017}.
+  The \ac{dc} step is commonly carried out to generate representative sequences from the demultiplexed and trimmed 16S sequencing data to obtain the \ac{otu}/\ac{esv} tables (counts of representative sequences for each sample).
+  In order to compare the representative sequence count tables generated by different tools, we processed the same 16S sequencing reads (samples from a fecal microbiome transplant study~\cite{Kang2017}) using 5 different methods: open-reference clustering, closed-reference clustering, denovo clustering, \ac{dada2}~\cite{Callahan2016} and Deblur~\cite{Amir2017}.
   The first three methods are from the vsearch plugin and Deblur is from the deblur plugin of \ac{qiime2} v2021.8.0~\cite{bolyenReproducibleInteractiveScalable2019}.
@@
   From Figure~\ref{fig:figure3}A one can see that the representative sequences generated by the different methods, with the exception of Deblur, are similar to each other when weighted by their abundance.
-  Figure~\ref{fig:figure3}B on the other hand shows an increase in dissimilarity between each pair of methods suggesting that the methods might differ in the assignment of sequences of lower abundance.
-
-  In order to verify this claim, for each of these methods we use only the top 1000 representative sequences and recalculate the distance metrics (Figure \ref{fig:figureS3}).
-  We observe that the both the weighted and unweighted UniFrac distances are increased implying that the top representative sequences generated by the different methods are not similar to each other.
+  Figure~\ref{fig:figure3}B on the other hand shows an increase in dissimilarity between each pair of methods suggesting that the methods might differ in the assignment of sequences lower in abundance.
+  In order to verify this claim, for each of these methods we filtered out only the top 1000 representative sequences and recalculated the distance metrics for each method (Figure \ref{fig:figureS3}).
+  We observe that the both the weighted and unweighted UniFrac distances are increased implying that the top representative sequences generated by the different methods are not as similar to each other.
   % TODO: Can we make this conclusion?
-  This leads us to the conclude that the representative sequences in the middle range of the abundance distribution are those that are most similar between the methods.
+  This leads us to conclude that the representative sequences in the middle range of the abundance distribution are those that must be the most similar between the methods.
 
   These comparisons only elucidate the pairwise similarity or dissimilarity of a pair of methods.
-  In order to determine the tool that most accurately recapitulates the reference sequences in the samples, we used the information about the known 16S sequences from the mock datasets.
+  In order to determine the tool that most accurately recapitulates the reference sequences in the samples, we used the information about the known 16S sequences from the mock datasets (mock4, mock12, and mock16).
   In particular, we used the pipeline to process mock community datasets using each of the possible methods included for this step.
-  We next compared the predicted representative sequences with the expected representative sequences and their distribution.
+  Subsequently, we compared the predicted representative sequences with the expected representative sequences and their distribution.
   The results (Figure~\ref{fig:figure3}C and D) show that, for the mock datasets, the predicted sequence distributions are substantially different from the expected sequence distribution.
   This can be seen in the large values for both unweighted and weighted UniFrac distance values when compared to Figure~\ref{fig:figure3}A and B.
-  This result is more exaggerated in the case of the unweighted UniFrac metric, where some of the datasets show a very high deviation from the expected sequences.
+  This result is more exaggerated in the case of the unweighted UniFrac metric, where some processed mock datasets show a very high deviation from the expected sequences.
   These high deviations are primarily in one or two of the three datasets that were analyzed and show that the datasets themselves play a big role in the performance of these methods.
-  Open-reference and denovo clustering methods perform the best under the weighted UniFrac metric and the worst (marginally) under the unweighted UniFrac metric.
+  Open-reference and de novo clustering methods perform the best under the weighted UniFrac metric and the worst (marginally) under the unweighted UniFrac metric.
   This result can be attributed to the large number of low abundance representative sequences that are generated by these methods.
@@
   There is no method that clearly outperforms the rest in all datasets.
-  Based on their slightly better performance on the mock datasets, their \textit{de novo} error correcting nature and other previous studies~\cite{Nearing2018}, \ac{dada2} and Deblur seem to be in general the most reliable.
+  Based on their slightly better performance on the mock datasets, their \textit{de novo} error correcting nature and other previous studies~\cite{Nearing2018}, \ac{dada2} and Deblur appear to be the most reliable.
   Overall, since \ac{dada2} as compared to Deblur, displays better performance on all the mock datasets and under the weighted UniFrac metric, we have decided to recommend this tool for the DC step of the pipeline.
@@
   After the denoising, the sequences are subject to chimera checking.
-  The \ac{micone} pipeline supports two different chimera checking methods, uchime and remove bimera.
-  The uchime method is from \ac{qiime2}, whereas the remove bimera method is from the \ac{dada2} R package.
+  The \ac{micone} pipeline supports two different chimera checking methods, uchime-denovo and remove bimera.
+  The uchime-denovo method is from \ac{qiime2}, whereas the remove bimera method is from the \ac{dada2} R package.
   We did not notice any significant difference between the two chimera checking methods (Figure~\ref{fig:figure_s4}), implying that they identify and remove mostly the same set of sequences as chimeras.
+  Since the remove bimera method was originally developed in conjunction with dada2 we use this method as the default for the chimera checking module.
 
@@
   The three 16S taxonomic reference databases used in this study are SILVA~\cite{Quast2012}, \ac{gg}~\cite{DeSantis2006} and \ac{ncbi} RefSeq~\cite{Sayers2009}.
-  SILVA and \ac{gg} are two popular 16S databases used for taxonomy identification.
-  The \ac{ncbi} RefSeq nucleotide database contains 16S rRNA sequences as a part of two BioProjects - 33175 and 33317.
+  SILVA and \ac{gg} are two popular 16S databases used for taxonomy identification and the \ac{ncbi} RefSeq nucleotide database contains 16S rRNA sequences as a part of two BioProjects - 33175 and 33317.
   The three databases vastly differ in terms of their last update status - \ac{gg} was last updated on May 2013, SILVA was last updated on August 2020 at the time of writing and \ac{ncbi} is updated regularly as new sequences are curated.
   Since updates to taxonomic classifications are frequent, these databases vary significantly in terms of taxonomy hierarchies including species names and phylogenetic relationships~\cite{Balvociute2017}.
-  Assignment using a particular database also requires a query tool, we use the NaiveBayes classifier from \ac{qiime2} for the \ac{gg} and SILVA databases and the blast tool (also included as a \ac{qiime2} plugin) for the \ac{ncbi} database.
+  Assignment using a particular database also requires a query tool, we use the ``Naive Bayes'' classifier from \ac{qiime2} for the \ac{gg} and SILVA databases and the ``BLAST'' tool (also included as a \ac{qiime2} plugin) for the \ac{ncbi} database.
   These tools have been well quantified and optimized~\cite{bokulichOptimizingTaxonomicClassification2018}, hence, we make use of the default parameters in our analyses.
@@
   The representative sequences obtained from the \ac{dada2} method (with remove bimera for chimera checking) in the DC step were used for taxonomic assignment using the three reference databases.
-  Figure~\ref{fig:figure4}A depicts a flow diagram (alluvial plot) that shows how the top 50 representative sequences (sorted by abundance) are assigned a Genus according to the three different databases.
+  Figure~\ref{fig:figure4}A depicts a flow diagram (alluvial plot) that shows how the top 50 representative sequences (sorted by abundance) are assigned a Genus according to the three databases.
   We observe that the top 50 representative sequences are assigned to 17 Genus in \ac{gg}, 23 in SILVA and 17 in \ac{ncbi}, however the sequences that are assigned to a particular Genus vary significantly based on the taxonomy database.
@@
   A representative sequence might be assigned an "unknown" Genus for one of two reasons: the first is if the taxonomy identifier associated with the sequence in the database did not contain a Genus; the second (more likely) reason is that the database contains multiple sequences that are very similar to the query (representative) sequence and the consensus algorithm (from \ac{qiime2}) is unable to assign one particular Genus at the required confidence.
-  Additionally, we observe that other abundant Genus like Bacteroides and Escherichia are also assigned to different representative sequences (indicated by the different proporitions in the alluvial plot).
-  Furthermore, the most abundant Genus (among the top 50 representative sequences) in the \ac{ncbi} taxonomy table was \textit{Escherichia} whereas it was assigned to "unknown" in \ac{gg} and in the SILVA taxonomy table it was assigned to \textit{Escherichia-Shigella}.
+  Additionally, we observe that other abundant Genus like Bacteroides and Escherichia are also assigned to different representative sequences (indicated by the different proportions in the alluvial plot).
+  Furthermore, the most abundant Genus (among the top 50 representative sequences) in the \ac{ncbi} taxonomy table was \textit{Escherichia}, but these sequences were assigned to "unknown" in \ac{gg} and in the SILVA taxonomy table they were assigned to \textit{Escherichia-Shigella}.
   Although these are minor differences (in the case of \ac{ncbi} and SILVA), when comparing a large number of taxonomy composition tables these problems are hard to diagnose.
@@
   However, this was fixed by making minor adjustments to the taxonomic names, such as changing Bacteroidota to Bacteroidetes in the SILVA Phylum assignments.
-  The comparison of all assigned genera (Figure~\ref{fig:figure_s5}) instead of the just the top 100 contains a higher percentage of mismatches.
+  The comparison of all assigned genera (Figure~\ref{fig:figure_s5}), instead of just the top 100, contains a higher percentage of mismatches.
   This suggests that the most abundant sequences are the ones that are consistently matched to the same taxonomies in the different reference databases, at least for the dataset tested in the current analysis.
@@
   In order to obtain an absolute measure of accuracy of the taxonomic assignments we use the expected reference sequences from the mock datasets as the query sequences for the databases and the expected taxonomic composition as the standard to compare against (Figure~\ref{fig:figure4}C).
-  Here, we observe that none of the databases perform better than the others in absolute terms and the dissimilarity is high, indicating that all the databases perform poorly when trying to recapture the expected taxonomic composition.
+  Here, we observe that none of the databases perform better than the others in absolute terms and that the dissimilarity with the expected compositon is high, indicating that all the databases perform poorly when trying to recapture the expected taxonomic composition.
 
   Given that no database performs better than others against mock datasets, the choice of which database to use should be driven by other reasons.
-  One user-specific way to choose, would be based on the known representation of taxa for the microbiome of interest (see also Discussion).
+  One user-specific way to choose, would be based on the known representation of taxa for the microbiome of interest (see Discussion).
 Another reason could be the frequency of updates and the potential for future growth.
-  % TODO: Find reference for this
-  Since the \ac{gg} database is more widely used in the field~[ref], to enable easier comparison with published networks in literature, we set \ac{gg} as the \ac{micone} standard for taxonomy assignment.
-  On the other hand, since the degree of similarity of assignments between \ac{ncbi} and \ac{gg} is fairly high (Figure~\ref{fig:figure_s5}), and due to its regular updation and maintenance the \ac{ncbi} database is also a good choice for the user.
-
-  % TODO: More refs to support our argument?
- % There are also other studies, [ref, OTT paper] that focus on the database aspect and they too have found that...
+  Since the \ac{gg} database is more widely used in the field~\cite{Balvociute2017}, to enable easier comparison with published networks in literature, we set \ac{gg} as the \ac{micone} standard for taxonomy assignment.
+  On the other hand, since the degree of similarity of assignments between \ac{ncbi} and \ac{gg} is fairly high (Figure~\ref{fig:figure_s5}), and due to its regular revision and maintenance the \ac{ncbi} database is also a good choice for the user.
+  % TODO: Do we need more refs to support our argument?
 
@@
 
-  The ten different network inference methods used in this study are: \ac{spieceasi}~\cite{Kurtz2015}, FlashWeave.jl~\cite{tackmannRapidInferenceDirect2019}, \ac{cozine}~\cite{haCompositionalZeroinflatedNetwork2020a}, \ac{harmonies}~\cite{jiangHARMONIESHybridApproach2020}, \ac{spring}~\cite{yoonMicrobialNetworksSPRING2019}, \ac{mldm}~\cite{Yang2017}, \ac{sparcc}~\cite{Friedman2012,Watts2018}, propr~\cite{quinnProprRpackageIdentifying2017}, Spearman and Pearson.
-  These network inference methods fall into two groups, the first set of methods (Pearson, Spearman, \ac{sparcc}, propr) infer pairwise correlations while the second set infer direct associations (\ac{spieceasi}, FlashWeave, \ac{cozine}, \ac{harmonies}, \ac{spring}, \ac{mldm}).
+  The ten different network inference methods used in this study are: \ac{spieceasi}~\cite{Kurtz2015}, FlashWeave~\cite{tackmannRapidInferenceDirect2019}, \ac{cozine}~\cite{haCompositionalZeroinflatedNetwork2020a}, \ac{harmonies}~\cite{jiangHARMONIESHybridApproach2020}, \ac{spring}~\cite{yoonMicrobialNetworksSPRING2019}, \ac{mldm}~\cite{Yang2017}, \ac{sparcc}~\cite{Friedman2012,Watts2018}, propr~\cite{quinnProprRpackageIdentifying2017}, Spearman and Pearson.
+  These network inference methods fall into two groups, the first set of methods (Pearson, Spearman, \ac{sparcc}, and propr) infer pairwise correlations while the second set infer direct associations (\ac{spieceasi}, FlashWeave, \ac{cozine}, \ac{harmonies}, \ac{spring}, and \ac{mldm}).
   Pairwise correlation methods involve calculation of the correlation coefficient between every pair of node (taxonomic entity like Genus) leading to the detection of spurious indirect connections.
@@
 
-  For the analysis presented in this section, we used the filtered (filtering during the OP step) Genus level counts table obtained using the \ac{gg} (dada2 and remove bimera methods in the DC step) reference database from the TA step as the input for the network inference algorithms.
+  For the analysis presented in this section, we used the filtered (filtering performed during the \ac{op} step) Genus level counts table obtained using the \ac{gg} (dada2 and remove bimera methods were used in the DC step) reference database from the \ac{ta} step as the input for the network inference algorithms.
   The Genus count tables were obtained by summing up the counts of the lower taxonomy levels (Species and \ac{otu}) that map to the same higher taxonomy level name.
   Figure~\ref{fig:figure5}A shows the networks inferred from this dataset using the different inference algorithms.
-  The different networks differ vastly in their edge-density and connectivity; even some of the edges in common to these networks can have their signs inverted.
+  The different networks differ vastly in their edge-density and connectivity; even the edges in common to these networks can have their signs inverted.
   Note, however, that some of these comparisons depend on the threshold that has to be applied to the networks (currently 0.3 for the correlation based methods, based on~\cite{Friedman2012} and 0.01 for the direct association methods).
@@
   In order to obtain a quantitative picture of the differences between the inferred networks, we analyzed the distribution of common nodes and edges (Figure~\ref{fig:figure5} B and C) using UpSet plots~\cite{Lex}.
-  Only \ac{harmonies}, \ac{cozine},\ac{spieceasi}, \ac{spring}, propr, and \ac{sparcc} are used in the comparison since Pearson and Spearman add a large number of spurious edges since they are not intended for compositional datasets).
-  The results for the node intersections show that the networks have a large number of nodes in common ($33$ out of $68$ total unique nodes) and no network possesses any unique node.
+  Only \ac{harmonies}, \ac{cozine},\ac{spieceasi}, \ac{spring}, propr, and \ac{sparcc} are used in the comparison since Pearson and Spearman add a large number of spurious edges as they are not intended for compositional datasets).
+  The results for the node intersections show that the networks have a large proportion of nodes in common ($33$ out of $68$ total unique nodes) and no network possesses any unique node.
   The edge intersections in contrast show that only $8$ edges (out of $202$ total unique edges) are in common between all the methods and each network has a large number of unique edges.
   These results indicate that there is a substantial rewiring of connections in the inferred networks.
-  Hence, in order to determine the interactions that are most likely to exist in the dataset we advocate for the construction of a consensus algorithm.
+  Hence, in order to determine the interactions that are most likely to exist in the dataset we advocate for the development of a consensus algorithm.
 
@@
 
-  \subsection*{The consensus network}
+  \subsection*{The consensus network algorithms}
 
-  We observe a large number of rewirings in the inferred association networks generated by the different methods (Figure \ref{fig:figure5}C), implying that there is a lack of agreement on the associations that should be a part of the final network.
+  We observe a high degree of rewiring in the inferred association networks generated by the different methods (Figure \ref{fig:figure5}C), implying that there is a lack of agreement on the associations that should be a part of the final network.
   Therefore, we have developed a consensus method that would take into consideration the evidence offered by each network inference algorithm and generate a consensus network that would contain the most probable interactions for the input data.
   We have developed two algorithms for the generation of the consensus network, the simple voting and the scaled-sum methods (see Methods).
-  We only include \ac{sparcc} and propr in the consensus network calculations, for consensus networks that include Spearman and Pearson see Figure~\ref{fig:figure_s6} in the Supplementary section.
   First, for the correlation-based methods we merge the p-values using the Brown's p-value merging method~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016,faustCoNetAppInference2016} (see Methods).
@@
   The scaled-sum method does a consensus based on whether the normalized sum of an edge between a pair of nodes from all inferred networks is greater than a threshold.
-  The default algorithm for the pipeline is the scaled-sum method, since this method takes into account the weights of the association in the determination of the final consensus.
-  Additionally, by default we include all the network inference methods in the consensus calculation by default, but a subset of methods can be considered if it is required.
-  Based on this approach, \ac{micone} reports as default output the consensus network, where is each edge is annotated with weights (correlations for the correlation-based methods and direct associations for the other methods) from all the methods used in the consensus algorithm.
+  The default algorithm for the pipeline is the scaled-sum method, since this method takes into account the weights of the associations in the determination of the final consensus.
+  By default, of the correlation-based methods only \ac{sparcc} and propr are included in the calculation of consensus network, for consensus networks that include Spearman and Pearson see Figure~\ref{fig:figure_s6} in the Supplementary section.
+  Additionally, the pipeline also allows for a different subset of methods to be used in the calculation of the consensus if it is required.
+  Based on this approach, \ac{micone} reports as default output, the consensus network, where is each edge is annotated with weights (correlations for the correlation-based methods and direct associations for the other methods) from all the methods used in the consensus algorithm.
 
@@
   Each point in the scatter plot represents a network with different input network topology and OTU abundance distributions.
-  Overall, we see that the scaled-sum method for a parameter value of 0.5 performs the best in terms of average precision ($0.985$).
-  Among the individual inference methods, we observe that \ac{spieceasi} has the best average precision ($0.944$)
   The goal of the consensus network algorithms is to increase the precision of the resultant consensus network by removing edges that are either not present in most of the inference methods or whose association strength is low.
   From Figure \ref{fig:figure6} we can confirm that the consensus algorithm, especially the scaled-sum method, does indeed produce networks with more precision, but as a consequence the sensitivity of the networks is reduced.
+  Overall, we see that the scaled-sum method for a parameter value of 1.0 performs the best in terms of average precision ($1.000$).
+  However, the scaled-sum method for a parameter value of 0.667 has the second-highest average precision ($0.985$) and does not underperform in terms of sensitivity.
+  Among the individual inference methods, we observe that \ac{spieceasi} has the best average precision ($0.944$)
   Therefore, the consensus networks provide the means to obtain a short list of associations that would have a high probability of being present in the real interaction network.
   This small set of associations could be further tested and validated experimentally to verify their plausibility.
-
   Furthermore, we also find that the consensus algorithm performs drastically better in terms of precision (0.820 for scaled-sum method vs.
-0.624 for \ac{spieceasi}) in recapturing the orginal interactions in the data.
+0.624 for \ac{spieceasi}) in recapturing the original interactions in the ``seqtime'' data.
   The precision and sensitivity graphs on the seqtime data can be found in the Supplementary section (Figure~\ref{fig:figure_s7}).
@@
   The systematic analyses performed in the previous sections clearly show that the choice of tools and parameters can have a big impact on the final co-occurrence network.
-  For some of these choices (e.g. \ac{dada2} vs. Deblur) there is no clear metric to establish a best protocol.
-  For other choices, the mock communities and synthetic data provide an opportunity to select combination of tools that yield more accurate and robust results.
-  Despite this partial degree of assessment, we wish to suggest a combination of tools and parameters that produce networks that are derived from the combination of tools which performed best on the mock communities, and displayed highest robustness to switching to alternative methods.
+  For some of these choices (e.g. \ac{dada2} vs. Deblur), there is no clear metric to establish the best protocol.
+  For other choices, the mock communities and synthetic data provide an opportunity to select the combinations of tools that yield more accurate and robust results.
+  Despite this partial degree of assessment, we wish to suggest a combination of tools and parameters that produce networks that are derived from the combination of tools which performed best on the mock communities, and displayed the highest robustness to switching to alternative methods.
   These tools and parameters are chosen as the defaults for the pipeline and are given in Table~\ref{tab:default_options}.
@@
   The recommended tool for the \ac{dc} step (\ac{dada2}) was chosen based on its accuracy in recapitulating the reference sequences in mock communities.
-  For the chimera checking module of the \ac{dc} step, both uchime and remove bimera perform similarly, we suggest using remove bimera since it was developed specifically for \ac{dada2} and should provide the best performance for that denoising method.
+  For the chimera checking module of the \ac{dc} step, since both uchime-denovo and remove bimera perform similarly, we suggest using remove bimera since it was developed specifically for \ac{dada2} and should provide the best performance in conjunction with that denoising method.
   The choice of the taxonomy reference database in the \ac{ta} step is dictated largely by the species expected to be present in the sample as well the database used in similar studies, if comparison is a goal.
-  Nevertheless, we suggest \ac{gg} database along with the NaiveBayes classifier as the query tool since the database is used most often in taxonomic studies.
-  % TODO: Default OP step values and references
-  The abundance threshold at the \ac{op} step is set at ... because large number of nodes are added to networks and decrease the performance of network inference algorithms
+  Nevertheless, we suggest \ac{gg} database along with the Naive Bayes classifier as the query tool since the database is used most often in taxonomic studies.
+  The filtering thresholds at the \ac{op} step are set at values described in Table~\ref{tab:micone_tools} in order to prevent the addition of large numbers of nodes to networks and the subsequent decrease in the performance of network inference algorithms.
   For the \ac{op} step, we recommend that the filtering be turned on and used at the default settings, unless a taxonomy entity of interest might be filtered out at the default settings.
   This is necessary since a large number of rows in the counts matrix would reduce the accuracy of the inference methods and increase the computational time.
-  Finally for the \ac{ni}, we recommend the scaled-sum consensus method along with p-value merging (for the correlation-based methods) to generate a final consensus network for the input data.
+  Finally, for the \ac{ni} step, we recommend the scaled-sum consensus method along with p-value merging (for the correlation-based methods) to generate a final consensus network for the input data.
 
-  Figure~\ref{fig:figure7}A shows the default network compared against networks generated by altering one of the steps of the pipeline from the default.
-  These results indicate that the biggest rewiring in networks occur when the reference database or the network inference algorithm are changed.
+  Figure~\ref{fig:figure7}A show the default network compared against networks generated by altering one of the steps of the pipeline from the default.
+  This result suggests that the most significant changes (many new nodes and edges are added) occur when the \ac{ta} step, \ac{op} step and the \ac{ni} step are changed.
+  However, the L1 distance of networks generated by altering one of the steps of the pipeline from the default against the default network (Figure~\ref{fig:figure7}B) shows that the biggest deviations from the default network occur when the \ac{ni} step is changed.
+  These results indicate that the biggest rewiring in networks occur when the network inference algorithm is changed and that the changes in the others steps do not create as significant of a difference.
   % TODO: Show proof??
-  The filtering in the \ac{OP} step when turned off also adds a large number of new nodes, but the edges between the common nodes (with the default network) are fairly conserved.
+  The change of taxonomy database and the filtering in the \ac{OP} step when turned off also adds a large number of new nodes, but the edges between the common nodes (with the default network) are fairly conserved.
+
   % TODO: Say that this shows how important the consensus network is in the whole pipeline
-  Furthermore, the L1 distance of networks generated by altering one of the steps of the pipeline from the default against the default network (Figure~\ref{fig:figure7}B) shows that the biggest deviations from the default network occur when the \ac{ni} steps are changed.
-  This shows us that not only does the consensus help reconcile the variability in the edges due to multiple network inference algorithms, it also helps reconcile the variability to the other steps of the pipline as well.
-  This result is different from that observed in Figure~\ref{fig:figure2} since in this particular analyses all the steps other than the \ac{ni} step have the consensus algorithm applied to them, which drastically reduces the edges in the graph that have lower probabilities.
-  This result reinforces the importance of the consensus network in increasing the robustness of the inferred networks to different processing methods.
+  The low L1 distances show us that not only does the consensus help reconcile the variability in the edges due to multiple network inference algorithms, it also helps reconcile the variability to the other steps of the pipeline as well.
+  This result is different from that observed in Figure~\ref{fig:figure2} since in this particular analysis all the steps other than the \ac{ni} step have the consensus algorithm applied to them, which drastically reduces the edges in the graph that have lower probabilities.
+  This reinforces the importance of the consensus network in increasing the robustness of the inferred networks to different processing methods.
   Figure~\ref{fig:figure7}C shows the co-occurrence networks inferred for the healthy subjects (control) and autistic subjects in the fecal microbiome transplant study~\cite{Kang2017}.

### 2022-05-06 sections/methods.tex
Change: Equation for N bar matrix
Change: Update the Methods for draft v2

@@
   \subsection*{16S rRNA sequencing datasets}
-
   \vspace{-5mm}
-    The study uses two kinds of 16S rRNA sequencing datasets: biological datasets and mock datasets.
+    This study utilized two kinds of 16S rRNA sequencing datasets: biological datasets and mock datasets.
     Biological datasets are collections of sequencing reads obtained from naturally occurring microbial community samples.
     The current analysis used stool samples from a fecal microbiome transplant study of autism~\cite{Kang2017} as biological datasets.
-    The study was composed of multiple sequencing runs and those runs that contained paired end reads (runs 2 (10M reads), 3 (750K reads) and 4 (16M reads)), were downloaded from Qiita~\cite{qiita} from https://qiita.ucsd.edu/study/description/10532 and used as input sequences for the \ac{micone} pipeline.
-    Sequences from both control (212 samples including neurotypical and donors) and autistic (126 samples) patients were included in the analyses but the groups were analyzed independently in the pipeline.
+    The study was composed of multiple sequencing runs and those runs that contained paired end reads (run 2 (10M reads), run 3 (750K reads) and run 4 (16M reads)), were downloaded from Qiita~\cite{qiita} (study ID 10532) and used as input sequences for the \ac{micone} pipeline.
+    Sequences from both control (212 samples including neurotypical and donors) and autistic (126 samples) patients were included in the analyses, but the groups were analyzed independently in the pipeline after the \ac{op} step.
     The mock community 16S datasets are sequencing data obtained for artificially assembled collections of species in known proportions.
@@
     Two replicate samples from mock4 contain all species in equal abundances, and two additional replicate samples contain the same species in unequal abundances.
-    The mock12 community is composed of 27 bacterial strains that include closely related taxa with some pairs having only one to two nucleotide difference from another.
-    The mock16 community is composed of 49 bacteria and 10 Archea, all represented in equal amount.
+    The mock12 community is composed of 27 bacterial strains that include closely related taxa with some pairs having only one to two nucleotide difference from one another.
+    The mock16 community is composed of 49 bacteria and 10 Archaea, all represented in equal amount.
 
   \subsection*{\ac{micone}}
-
   \vspace{-5mm}
@@
   The pipeline integrates many publicly available tools as well as custom R or Python modules and scripts to extract co-occurrence associations from 16S sequence data.
-  Each of these tools corresponds to a distinct R or python module that recapitulates the relevant analyses.
+  Each of these tools corresponds to a distinct module that recapitulates the relevant analyses.
   All such individual modules are available as part of the \ac{micone} package.
-  The inputs to the pipeline by default are raw untrimmed 16S rRNA sequence reads, but the software can be alternatively configured to use trimmed sequences, \ac{otu} tables and other types of intermediate data.
-  The pipeline supports both paired end and single end reads, and additionally supports independently processing multiple runs and merging the OTU tables in the DC step.
+  The inputs to the pipeline by default are raw untrimmed 16S rRNA sequence reads, but the software can be alternatively configured to use trimmed sequences, \ac{otu} tables and other types of intermediate data (see documentation).
+  The pipeline supports both paired end and single end reads, and additionally supports independently processing reads from multiple runs and merging the OTU tables in the DC step.
   The final output of the pipeline is the inferred network of co-occurrence relationships among the microbes present in the samples.
 
-  The \ac{micone} pipeline provides both a Python API as well as a command-line interface and only uses a single configuration file (nextflow.config) to contain the configuration parameters.
-  The \ac{micone} Python API provides several biom file and network related functions and methods, enabling detailed comparison of counts tables and inferred networks if desired.
-  It also contains helper functions that can help in parsing taxonomies and communicate with the \ac{ncbi} taxonomy database to query taxonomy by name or taxonomic ids.
+  The \ac{micone} pipeline provides both a Python API together with a command-line interface and only uses a single configuration file (nextflow.config) to contain the configuration parameters.
+  The \ac{micone} Python API provides several biom (\ac{otu} table) and network related functions and methods, enabling detailed comparison of counts tables and inferred networks if desired.
+  It also contains helper functions that can help in parsing taxonomies and communicate with the \ac{ncbi} taxonomy database to query taxonomy by name or taxonomic IDs.
   The configuration file along with the run file (main.nf) lists the inputs, output and the steps to be performed during runtime, along with the parameters to be used (if different from defaults) for the various steps.
-  Since the entire pipeline run-through is stored in the form of a text file (the configuration file), subsequent runs are highly reproducible and changes can be easily tracked using version control.
-  It uses the nextflow workflow manager~\cite{Tommaso2015} under the hood, making it readily usable on local machines, cluster or cloud with minimal configuration change.
+  Since the entire pipeline run is stored in the form of a text file (the configuration file), subsequent runs are highly reproducible and changes can be easily tracked using version control.
+  It makes use the nextflow workflow manager~\cite{Tommaso2015} under the hood, making it readily usable on the local machine, cluster or cloud with minimal configuration change.
   It also allows for automatic parallelization of all possible processes, both within and across samples.
@@
   \subsection*{Sequence Processing (SP)}
+  \vspace{-5mm}
   This module deals with processing the raw multiplexed 16S sequence data into demultiplexed, quality-controlled, trimmed sequences.
-  It consists of the following processes: demultiplexing and trimming (includes quality control and filtering).
+  It consists of the demultiplexing and trimming processes.
   The trimming process handles the quality control steps such as trimming adapters and trimming low-quality nucleotide stretches from the sequences.
@@
   It consists of the following processes: denoising (or clustering) and chimera checking.
-  The denoise/cluster process handles the conversion of the demultiplexed, trimmed sequences into \ac{otu} or \ac{esv} count tables (some methods, like closed reference and open reference clustering, perform clustering and taxonomy assignment in the same step).
+  The denoise/cluster process handles the conversion of the demultiplexed, trimmed sequences into \ac{otu} or \ac{esv} count tables (some methods, like closed reference and open reference clustering, make use of a taxonomy reference database for clustering).
   The chimera checking process handles the removal of chimeric sequences created during the \ac{pcr} step.
@@
   \vspace{-5mm}
-  This module deals with assigning taxonomies to either the representative sequences of the \ac{otu}s or directly to the \ac{esv}s.
-  In order to assign taxonomies to a particular sequence we need a taxonomy database and a query tool.
-  The taxonomy database contains the collection of 16S sequences of micro-organisms of interest and the query tool allows one to compare a sequence of interest to all the sequences in the database to identify the best matches.
+  This module deals with assigning taxonomies to either the representative sequences of the \ac{otu}s or the \ac{esv}s.
+  In order to assign taxonomies to a particular sequence a taxonomy database and a query tool are necessary.
+  The taxonomy database contains the collection of 16S sequences of microorganisms of interest and the query tool allows one to compare a sequence of interest to all the sequences in the database to identify the best matches.
   Finally, a consensus method is used to identify the most probable match from the list of best matches.
@@
   These databases were downloaded and built using the RESCRIPt qiime2 plugin~\cite{iiRESCRIPtReproducibleSequence2021}.
-  The Naive Bayes classifier and the \ac{ncbi} blast used as the query tools in this study were from the \ac{qiime2} package and the parameters used were also the defaults in the package.
+  The Naive Bayes classifier and the \ac{ncbi} blast used as the query tools in this study were from the \ac{qiime2} package and the parameters used were the defaults of the package.
   The consensus algorithm used is the default method used by the classifiers in \ac{qiime2}.
@@
   This is performed either by sub-sampling or by normalization of the matrix to the lowest sampling depth~\cite{Weiss2015}.
-  Although the pipeline supports rarefaction, these modules are turned off by default and the analyses reported in the paper do not normalize or rareify the counts matrices.
+  Although the pipeline supports normalization and rarefaction, these modules are turned off by default and the analysis reported in the paper do not normalize or rareify the counts matrices.
   However, most of the network inference methods perform normalization and other transformation operations on the counts matrix as a part of their workflow.
-  Filtering, is performed to remove samples or features (\ac{otu}s or \ac{esv}s) from the count matrix that are sparse.
+  Filtering, is performed to remove samples or features (\ac{otu}s or \ac{esv}s) from the counts matrix that are sparse.
   By default, when the OP module is ``on'', the samples are filtered out if the total reads in a sample are less than 500 and features are filtered out if the abundance is less than 1\%, prevalence is less than 5\% and count sum across all the samples is less than 100.
@@
   The group operation transforms the \ac{otu} or \ac{esv} count matrix into a taxonomic count matrix at the requested level by adding up counts that map to the same taxonomy.
-  Finally, transformations are performed in order to correct for and overcome the compositional bias that is inherent in a counts matrix (in analyses peformed in the study these are handled by the network inference algorithm).
+  Finally, transformations are performed in order to correct for and overcome the compositional bias that is inherent in the counts matrix (in the analysis performed in the study these are handled by the network inference algorithm).
   All the modules in this step were implemented using functions from the \href{http://biom-format.org/}{biom-format} Python package~\cite{mcdonaldBiologicalObservationMatrix2012}.
@@
   The input count matrices are collapsed to the Genus level (or any other required taxonomy level) using the group module at the OP step.
-  These collapsed matrices are as used as input to the network inference methods to produce association matrices at the appropriate taxonomy level.
-  These associations can be represented as a network, with nodes representing taxonomies of the micro-organisms and edges representing the associations between them.
-  A null model is created by re-sampling and bootstrapping the counts matrix and recalculating the correlations/associations.
-  These bootstrapped association matrices are used to calculate the significance of the inferred associations by calculating the p-values against this null model~\cite{Watts2018}.
+  These collapsed matrices are used as input to the network inference methods to produce association matrices at the appropriate taxonomy level.
+  These associations can be represented as a network, with nodes representing taxonomies of the microorganisms and edges representing the associations between them.
+  A null model is created by re-sampling and bootstrapping the counts matrix and recalculating the correlations.
+  These bootstrapped association matrices are used to calculate the significance of the inferred correlations by calculating the p-values against this null model~\cite{Watts2018}.
   The pipeline includes 4 methods for pairwise correlation metrics, and 6 methods for direct association metrics (refer Table~\ref{tab:micone_tools}).
   The Brown's p-value merging method~\cite{brown_400_1975} is used for combining p-values from the pairwise correlations methods to obtain a consensus p-value, which can be used to filter for significance.
-  In the final module of this step, the consensus algorithms are used to create the final consensus network using associations from all the network inference methods.
-  The output of this step are co-occurrence association networks encoded in the JSON format (refer Supplementary section) and are also exportable to a variety of network formats like .gml.
+  The bootstrapping and p-value calculations are only performed on the correlations-based methods.
+  In the final module of this step, the consensus algorithms are used to create the final consensus network using associations from all the network inference methods (except Pearson and Spearman, by default).
+  The output of this step are co-occurrence association networks encoded in the JSON format (refer Supplementary section) and which can also be exportable to a variety of network formats.
   The list of tools used in this step, along with their modules and references are provided in Table~\ref{tab:micone_tools}.
@@
   \subsection*{Consensus Network and p-value merging}
-  \vspace{-5mm}
 
  \subsubsection*{Notation}
+  \vspace{-5mm}
+ This section defines the notation used below to describe the consensus network algorithm used in the \ac{micone} pipeline.
+ Note that all networks to be compared have the same number of nodes.
 
-  This section defines the notation used below to describe the consensus network algorithm used in the our pipeline. Note that all networks to be compared have the same number of nodes.
+  $w$, the number of co-occurrence networks to be integrated into the consensus network (corresponding to the number of the different approaches used to infer co-occurrence networks excluding Spearman and Pearson)
 
-  $w$, the number of co-occurrence networks to be integrated in the consensus network (corresponding to the number of the different approaches used to infer co-occurrence networks)
+  $q$, the number of unique nodes across all $w$ co-occurrence networks
 
-  $q$, the number of nodes in any co-occurrence network (same across all $w$ networks)
+  $N^i$, the matrix of edge weights for the $i^{th}$ co-occurrence network.
+  This is a $q \times q$ matrix, where $i \in \{1,\dots,w\}$.
+  $N^i_{a,b}$ represents edge $(a,b)$ in network $i$
 
-  $N^i$, the matrix of network edge weights for the $i^{th}$ co-occurrence network. This is a $q \times q$ matrix, where $i \in \{1,\dots,w\}$. $N^i_{a,b}$ represents edge $(a,b)$ in network $i$
+  $P^i$, the matrix of p-values for all edges of the $i^{th}$ co-occurrence network.
+  This is a $q \times q$ matrix, where $i \in \{1,\dots,q\}$
 
-  $P^i$, the matrix of p-values for all edges of the $i^{th}$ co-occurrence network. This is a $q \times q$ matrix, where $i \in \{1,\dots,q\}$
-
-  $\bar{N}^i$ is a "flattened" version of the adjacency matrix $N^i$ into a $q^2 \times 1$ column vector, where all columns are stacked onto each other into a $q^2$ long vector.
+  $\bar{N}^i$, the ``flattened'' version of the adjacency matrix $N^i$ into a $q^2 \times 1$ column vector, where all columns are stacked onto each other into a $q^2$ long vector.
   Element $\bar{N}^i_j$ corresponds to the $j^{th}$ edge in the $i^{th}$ network.
 
-  $\bar{P}^i$ is a "flattened" version of the adjacency matrix $P^i$ into a $q^2 \times 1$ column vector, where all columns are stacked onto each other into a $q^2$ long vector.
+  $\bar{P}^i$, the ``flattened'' version of the adjacency matrix $P^i$ into a $q^2 \times 1$ column vector, where all columns are stacked onto each other into a $q^2$ long vector.
 
   \subsubsection*{Bootstrapping and p-value calculation}
-  For all correlation-based methods $k \in \{1,\dots,w\}$, $1000$ permutations of the original \ac{otu} counts data were generated~\cite{Watts2018}.
+  \vspace{-5mm}
+  For all correlation-based methods $k \subset w$, $1000$ permutations of the original \ac{otu} counts data were generated~\cite{Watts2018}.
   All the direct association based methods used in the study have their own regularization methods built-in and hence do not need to undergo bootstrapping.
-  We then recalculate the correlations in these permuted \ac{otu} tables using the different correlation-based algorithms.
-  Finally, we calculate the p-value based on how often a more extreme association is observed for randomly permuted data.
+  The correlations in the permuted \ac{otu} tables are recalculated using the different correlation-based algorithms.
+  Finally, the p-value is determined based on how often a more extreme association is observed for randomly permuted data.
 
   \subsubsection*{p-value merging}
-
+  \vspace{-5mm}
   The first part of the consensus algorithm workflow is to merge the p-values for the networks generated by the correlation-based methods.
@@
         & \hat{P}_j = 1.0 - \Phi_{2f}\left( \psi / c \right) \\
-        \text{where},~ &\psi = -2 \sum_{i=1}^w \log(\bar{P}^i_j) ~~~\text{and}~~~ \Phi_{2f} = \mathrm{CDF}\left( \chi^2_{2f} \right)
+        \text{where},~ &\psi = -2 \sum_{i=1}^k \log(\bar{P}^i_j) ~~~\text{and}~~~ \Phi_{2f} = \mathrm{CDF}\left( \chi^2_{2f} \right)
     \end{aligned}
@@
   \end{equation}
-  where, $\hat{P}_j$ is the combined p-value for the edge $j$, f is the number of degrees of freedom, and c is a scale factor.
+  where, $\hat{P}_j$ is the combined p-value for the edge $j$, $f$ is the number of degrees of freedom, and $c$ is a scale factor.
 
-  Note that we do not use Pearson and Spearman methods in the p-value merging step to calculate the consensus network, these algorithms are only used for demonstration and comparison.
+  Note that we do not use Pearson and Spearman methods in the p-value merging step to determine the consensus network, these methods are only used for demonstration and comparison.
   The combined p-values are used to threshold for significance right before the consensus algorithm is applied to the inferred networks.
@@
   \vspace{-5mm}
-
   The consensus algorithm was designed to increase the precision (number of true positives) at the end of the network inference step.
-  For this purpose, we designed two simple algorithms that combine the edges reported by the different network inference tools for each pair of nodes using a consensus based on a parameter $p$.
+  For this purpose, we developed two simple algorithms that combine the edges reported by the different network inference tools using a consensus based on a parameter $p$.
   The inputs to both the algorithms are the co-occurrence networks (association matrices) $\bar{N}^i$ (flattened version of $N^{i}$) generated by each method $i$ and the threshold parameter $p$.
@@
 
-    Note that this method is used to filter relevant interactions.
-However, if a given pair of nodes is inferred to have edges that satisfy the above relationships, all corresponding edges from the $w$ networks will be returned by the algorithm.
+    Note that this method is only used to filter relevant interactions.
+    If a given pair of nodes is inferred to have edges that satisfy the above relationships, all corresponding edges from the $w$ networks will be returned by the algorithm, as a multigraph.
 
-  \paragraph*{Algorithm 1: Simple voting}
+  \paragraph*{Algorithm 1 - Simple voting:}
   The simple voting method performs a voting-based consensus to determine whether an edge will exist between a given node-pair in the final consensus network~\cite{bustinceFuzzySetsTheir2008,tsarevApplicationMajorityVoting2018}.
-  For each pair of nodes, we determine the number of network inference methods that report an edge $j=k$ between them, i.e. $\bar{N}^{i}_{j=k}, \forall i \in \{1,\dots,w\}$.
-  Each node-pair will have an edge in the final consensus network if the number of reported edges is larger than a user-defined threshold.
-  (Equation~\ref{eqn:simple-voting2})
+  For each pair of nodes, we determine the number of network inference methods that report an edge $j=l$ between them, i.e. $\bar{N}^{i}_{j=l}, \forall i \in \{1,\dots,w\}$.
+  Each node-pair will have an edge in the final consensus network if the number of reported edges is larger than a user-defined threshold (Equation~\ref{eqn:simple-voting2}).
 
- The number of reported edges is computed as follows: for each edge $j$, we obtain $M_j$ which represents the number of networks in which edge $j$ is reported, i.e.:
+ The number of reported edges is computed as follows:
 
+ For each edge $j$, we obtain $M_j$ which represents the number of networks in which edge $j$ is reported.
   \begin{equation}
@@
   \end{equation*}
-  where, # refers to the cardinality of the set.
+  where, $\#$ refers to the cardinality of the set.
 
- The edge $j$ appears in the final consensus network if the reported edges is greater than the threshold:
+ The edge $j$ appears in the final consensus network if the number of reported edges is greater than the threshold.
   \begin{equation}
@@
   \end{equation}
-  where,
-  $p$ is the user-defined parameter threshold.
+  where, $p$ is the user-defined threshold parameter.
 
-  The simple voting method gives the union of the networks when $p \in (0, \frac{1}{w})$ and will return the intersection when $p \in ((w - 1), 1)$.
-  In general, if $p \in ((n - 1), \frac{n}{w})$, this algorithm will report an edge in the consensus network when at least $n$ network inference methods report this edge.
+  The simple voting method returns the union of the networks when $0 \leq p \leq \frac{1}{w}$ and will return the intersection when $(w - 1) \leq p \leq 1$.
+  In general, if $(n - 1) \leq p \leq \frac{n}{w}$, this algorithm will report an edge in the consensus network when at least $n$ network inference methods report this edge.
 
-  \paragraph*{Algorithm 2: Scaled-sum method}
+  \paragraph*{Algorithm 2 - Scaled-sum method:}
 
   This algorithm generates a consensus network based on the sum of all edges reported between a pair of nodes~\cite{bustinceFuzzySetsTheir2008,tsarevApplicationMajorityVoting2018}.
-  Since in generating this consensus network we combine direct association metrics with correlation-based methods, summing of the edges is preceded by a pre-processing step, in which all networks are re-scaled.
+  Since in generating a consensus network using this method we sum the edges reported by direct association methods with those from correlation-based methods, summing of the edges is preceded by a pre-processing step, in which all networks are re-scaled.
 
-  First, the network generated by each network inference method ($\bar{N}^i$) is re-scaled into a normalized version ($\bar{S}^i$), as follows  (Equation~\ref{eqn:scaled-sum-rescaling}):
-
+  First, the network generated by each network inference method ($\bar{N}^i$) is re-scaled into a normalized version ($\bar{S}^i$), as follows:
   \begin{equation}
-    \bar{S}^{i} = \frac{\bar{N}^{i}}{\max(\mid \bar{N}^{i} \mid)}
+    \bar{S}^{i} = \frac{\bar{N}^{i}}{\max(\mid \bar{N}^{i} \mid)},~~\forall i \in {1, \dots, w}
     \label{eqn:scaled-sum-rescaling}
@@
 
-  Next, for each edge $(a,b)$, we sum the weights of all reported edges from the different networks (Equation \ref{eqn:scaled-sum}).
-
+  Next, for each edge $j$, we sum the weights of all reported edges from the different networks.
   \begin{equation}
@@
 
-  An edge $j$ will be included in the consensus network if:
+  An edge $j$ will be included in the consensus network if $s_j$ passes a threshold.
   \begin{equation}
@@
     \label{eq:scaled-sum2}
-  \end{equation}.
+  \end{equation}
 
@@
   \subsection*{Network Variability}
-  \vspace{-5mm}
 
@@
   \vspace{-5mm}
-  This section defines notation for the network variability analysis in our study.
+  This section defines notation for the network variability analysis performed for Figure~\ref{fig:figure2}.
 
   $W$, the number of co-occurrence networks generated from all possible combinations of tools and parameters in the workflow.
-Note that this is different from $w$, which counted only the possible network inference modules.
+  Note that this is different from $w$, which counted only the different network inference modules.
 
-  $Q$, the number of combined nodes from all $W$ networks
+  $Q$, the number of unique nodes across all $W$ networks.
 
-  $N^i$ = The weights of the $i^{th}$ co-occurrence network represented as a $Q \times Q$ adjacency matrix, where $i \in {1, \dots, W}$
+  $N^i$, the edge weights of the $i^{th}$ co-occurrence network represented as a $Q \times Q$ adjacency matrix, where $i \in {1, \dots, W}$.
+  $N^i_{a,b}$ represents the edge $(a,b)$ in network $i$
 
-  $N^i_{a,b}$ = edge $(a,b)$ in network $i$
-
-  $\bar{N}_i$ is a "flattened" version of the adjacency matrix $N^i$ into a $Q^2 \times 1$ column vector, where all columns are stacked onto each other into a $Q^2$ long vector:
+  $\bar{N}_i$, the ``flattened'' version of the adjacency matrix $N^i$ into a $Q^2 \times 1$ column vector, where all columns are stacked onto each other into a $Q^2$ long vector.
 
   \subsubsection*{Principal Component Analysis and variability calculation}
-
   \vspace{-5mm}
-   In order to compare across different networks, and analyze the degree of variability induced by the choice of different modules, we organized multiple networks into a single mathematical structure that we could use for linear regression.
-   First, we obtain co-occurrence network $\bar{N}_i$ for each of the $W$ possible tool and parameter combinations in the workflow.
-   Then, we construct a matrix $\mathbf{\bar{N}}$ whose $i^{th}$ column is the flattened version of the $i^{th}$ network, i.e. the column vector $\bar{N}^i$.
-   Therefore $\bar{N}^i_j$ is the weight of edge $j$ in network $i$.
-   $\bar{N}^i_j$ is assigned a value of 0 if edge $j$ did not exist in network $i$ but was present in other networks.
+   In order to compare across different networks and calculate the degree of variability induced by the choice of different modules, we organized multiple networks into a single mathematical structure that we could use for linear regression.
+   First, we obtained the co-occurrence network $\bar{N}_i$ for each of the $W$ possible tool and parameter combinations in the workflow.
+   We then constructed a matrix $\mathbf{\bar{N}}$ whose $i^{th}$ column is the flattened version of the $i^{th}$ network, i.e. the column vector $\bar{N}^i$.
+   Therefore, $\bar{N}^i_j$ is the weight of edge $j$ in network $i$.
+   $\bar{N}^i_j$ is assigned a value of 0 if edge $j$ did not exist in network $i$ but was present in one of the other networks.
    Note that row $j$ of $\mathbf{\bar{N}}$, $\bar{N}_j$ is the vector that encodes the values of edge $j$ across all the networks.
@@
    \mathbf{\bar{N}} =
-     \begin{blockarray}{*{4}{c} l}
-      \begin{block}{*{4}{>{$\footnotesize}c<{$}} l}
-        \bar{N}^1 & \bar{N}^2 & \dots & \bar{N}^{W} & \\
-      \end{block}
-      \begin{block}{[*{4}{c}]>{$\footnotesize}l<{$}}
-       n^1_1 & n^2_1 & \cdots  & n^{W}_1 \bigstrut[t]& \bar{N}_1 \\
-       n^1_2 & n^2_2 & \cdots  & n^{W}_2 & \bar{N}_2 \\
-       \vdots & \vdots & \vdots  & \vdots & \vdots \\
-       n^1_{Q^2} & n^2_{Q^2} & \cdots  & n^{W}_{Q^2} & \bar{N}_{Q^2}
-      \end{block}
-     \end{blockarray}
+      \begin{bmatrix}
+       \bar{N}^1_1 & \bar{N}^2_1 & \cdots  & \bar{N}^{W}_1  \\
+       \bar{N}^1_2 & \bar{N}^2_2 & \cdots  & \bar{N}^{W}_2 \\
+       \vdots & \vdots & \vdots  & \vdots \\
+       \bar{N}^1_{Q^2} & \bar{N}^2_{Q^2} & \cdots  & \bar{N}^{W}_{Q^2}
+      \end{bmatrix}
   \end{equation*}
 
-  To infer the variability contributed due to different steps in the pipeline we perform a linear regression on each edge in $\mathbf{\bar{N}}$ and a subsequent ANOVA to extract the within group variances.
-  One issue with this approach is that the possible correlations between the edges of the network could lead to inaccurate estimates of the variance if a linear model were used to directly model the relationships between edges and steps in the workflow.
-  In order to remedy this issue, we perform Principal Component Analysis on the matrix $\mathbf{\bar{N}}$ to obtain the $\mathbf{C}$ matrix ($W \times c$) of components for each network, such that we reduce the dimensions from the $Q^2$ dimensional edge space to a $c$ dimensional component space.
+  To infer the variability contributed due to the different steps in the pipeline we performed a linear regression on each edge in $\mathbf{\bar{N}}$ and a subsequent ANOVA to extract the within group variances.
+  One issue with this approach is that the possibility of correlations existing between the edges of the network could lead to inaccurate estimates of the variance if a linear model were used to directly model the relationships between edges and steps in the workflow.
+  In order to remedy this issue, we performed a PCA (Principal Component Analysis) on the matrix $\mathbf{\bar{N}}$ to obtain the $\mathbf{C}$ matrix ($W \times c$) of components for each network, such that we reduce the dimensions from the $Q^2$ dimensional edge space to a $c$ dimensional component space.
 
@@
   In particular, we infer parameters $\alpha_i$ such that:
-     \begin{multline*}
+  \begin{multline}
   C_j = &\sum_{k=1}^5 \left( \alpha^{DC(k)}_j.\delta^{DC(k)}_j \right) +
@@
          \epsilon_j
-     \end{multline*}
+  \end{multline}
 
-   where, $\alpha_i$ are the coefficients of the regression, $\epsilon_i$ are the residuals and $\delta_i$ are the indicator variables that correspond to the processes utilized in the pipeline used to create the network $N_i$; for example, $\delta^{DC(1)}_i = 1$ if the DC(1) process was used in the generation of the network $N_i$ .
-   Here, (i) DC(1) = "closed reference", DC(2) = "open reference", DC(3) = "de novo", DC(4) = "dada2", DC(5) = "deblur";
-   (ii) CC(1) = "remove bimera", CC(2) = "uchime";
-   (iii)  TA(1) = "GreenGenes", TA(2) = "SILVA", TA(3) = "NCBI";
-   (iv) OP(1) = "no filtering", OP(2) = "filtering";
-   (v) NI(1) = "SparCC", NI(2) = "Pearson", NI(5) = "Spearman", NI(4) = "Propr", NI(5) = "SpiecEasi", NI(6) = "COZINE", NI(7) = "HARMONIES", NI(8) = "SPRING", NI(9) = "mLDM", NI(10) = "FlashWeave";
+   where, $\alpha_i$ are the coefficients of the regression, $\epsilon_i$ are the residuals and $\delta_i$ are the indicator variables that correspond to the processes utilized in the pipeline used to create the network $N_i$; for example, $\delta^{DC(1)}_i = 1$ if the DC(1) process was used in the generation of the network $N_i$.
 
-   The variance contributed by each step of the pipeline is calculated for every component in $\mathbf{C}$ matrix through ANOVA using the Python \texttt{statsmodels}~\cite{seaboldStatsmodelsEconometricStatistical2010} package and is shown in Figure~\ref{fig:figure2}A.
+   Here,
+   \begin{enumerate}
+     \item $DC(k) \subset$ \{CR, OR, DN, D2, DB\}
+     \item $CC(k) \subset$ \{remove bimera, uchime-denovo\}
+     \item $TA(k) \subset$ \{NaiveBayes(GG), NaiveBayes(SILVA), BLAST(NCBI)\}
+     \item $OP(k) \subset$ \{Filter(on), Filter(off)\}
+     \item $NI(k) \subset$ \{\ac{sparcc}, propr, Spearman, Pearson, \ac{spieceasi}, \ac{cozine}, \ac{harmonies}, \ac{spring}, \ac{mldm}, FlashWeave\}
+   \end{enumerate}
+
+   The variance contributed by each step of the pipeline is calculated for every component in $\mathbf{C}$ matrix through ANOVA using the Python statsmodels~\cite{seaboldStatsmodelsEconometricStatistical2010} package and is shown in Figure~\ref{fig:figure2}A.
   The total variance for the network is calculated by adding the variances for each connection.
@@
   \vspace{-5mm}
-
   We generated synthetic interaction data using two methodologies previously used for benchmarking network inference methods.
@@
   We used the \href{https://github.com/hallucigenia-sparsa/seqtime}{seqtime} R package to simulate communities with number of species ($N$) varying from 10 to 150 (10, 25, 50, 100, 150 and 200).
-  The initial species concentrations are randomly sampled from a Poisson distribution and the simulation is rerun to generate a number of samples ($S$) varying from 50 to 500 (50, 100, 200, 500) for different communities.
-  The abundance values of the species in the community at the end of the simulation time are used to create the OTU table.
+  The initial species concentrations were randomly sampled from a Poisson distribution and the simulation was rerun to generate a number of samples ($S$) varying from 50 to 500 (50, 100, 200, 500) for different communities.
+  The abundance values of the species in the community at the end of the simulation time were used to create the OTU table.
 
@@
   We used the \href{https://github.com/zdk123/SpiecEasi}{spieceasi} R package~\cite{Kurtz2015} to simulate communities with different network topologies (scale-free, cluster, block, Erdos-Renyi, band and hub) and target abundance distributions (Negative Binomial, Poisson, Zero-Inflated Negative Binomial).
-  The OTU table was generated using the American Gut Project example in the \texttt{spieceasi} package (amgut1.filt) with the default parameter options.
+  The OTU table was generated using the American Gut Project example in the spieceasi package (amgut1.filt) with the default parameter options.
 
-  For each method, we generated the OTU table depicting the abundances of species and use this as input to generate association networks using \ac{micone} pipeline.
+  For each method, we generated the OTU table depicting the abundances of species and used this as input to generate association networks using \ac{micone} pipeline.
   The interaction matrix was used as the source of expected (true) interactions and the associations predicted using \ac{micone} were the source of predicted interactions.
-  Finally, for each dataset we evaluated the precision and sensitivity of the associations predicted by the individual network inference methods as well as the consensus (Figure \ref{fig:figure7}).
+  Finally, for each dataset we evaluated the precision and sensitivity of the associations predicted by the individual network inference methods as well as the consensus (Figure \ref{fig:figure6}).
 
   \subsection*{Statistical analyses}
-  \vspace{-5mm}
 
   \subsubsection*{DC step}
+  \vspace{-5mm}
   In order to compare the representative sequences generated by the various methods in the DC step, we employ both the weighted~\cite{Lozupone2007} (Figure~\ref{fig:figure3}A) and unweighted UniFrac method~\cite{Lozupone2005} (Figure~\ref{fig:figure3}B).
@@
   The weighted UniFrac distance metric takes into account the abundances of the representative sequences when calculating shared and unshared branch lengths, whereas the unweighted UniFrac distance metric does not and hence gives equal weights to each sequence.
-  In Figure~\ref{fig:figure3} the distances between methods are the average distance calculated between the reference sequence distribution generated for every corresponding sample in the dataset.
+  In Figure~\ref{fig:figure3} the distances between methods are the average distance calculated between the reference sequence distribution for a pair of methods generated for every corresponding sample in the dataset.
   All UniFrac calculations were performed using the \texttt{scikit-bio}~\cite{thescikit-biodevelopmentteamScikitbioBioinformaticsLibrary2022} v0.5.6 Python package.
@@
   \subsubsection*{TA step}
-  In Figure~\ref{fig:figure4}C we use the Bray-Curtis distance metric to calculate the distance between the predicted (using the TA step) and expected taxonomic distribution.
+  \vspace{-5mm}
+  In Figure~\ref{fig:figure4}C we use the Bray-Curtis distance metric to calculate the distance between the predicted (using the taxonomy databases in the TA step) and expected taxonomic distribution.
   The Bray-Curtis distance is used to quantify the compositional dissimilarity between two different taxonomic distributions.
-  It is defined as: $\sum{|u_i-v_i|} / \sum{|u_i+v_i|}$
+  It is defined as:
+  $$\sum{|u_i-v_i|} / \sum{|u_i+v_i|}$$
   The Bray-Curtis distance calculations were performed using the \texttt{scipy}~\cite{virtanenSciPyFundamentalAlgorithms2020} v1.8.0 Python package.
@@
   \subsubsection*{NI step}
-  % Write about precision and sensitivity equations used
+  \vspace{-5mm}
   In Figure~\ref{fig:figure6} we calculate the accuracy and robustness of the inferred association networks (using the various network inference algorithms and the consensus methods) against the original interaction used to create the taxonomic distribution.
   We use precision and sensitivity to calculate the accuracy of the predictions.
-  Precision is given by, $\frac{TP}{TP + FP}$
-  Sensitivity is given by, $\frac{TP}{FN + TP}$
-  % L1 distance equations used
+
+  Precision = $\frac{TP}{TP + FP}$
+
+  Sensitivity = $\frac{TP}{FN + TP}$
+
   In Figure~\ref{fig:figure7}B we use the L1 distance metric to measure to the distance between the network inferred using different tools at the same step of the workflow.
   The L1 distance between networks $x$ and $y$ is calculated using the following formula:
-  $$ \frac{|A_{x}-A_{y}|}{2} $$, where $A$ is the adjacency matrix of the network.
+  $$\frac{|A_{x} - A_{y}|}{2}$$
+  where, $A$ is the adjacency matrix of the network.
 

### 2022-05-06 sections/supplementary.tex
Change: Update the Methods for draft v2

@@
 
-  Fisher~\cite{fisher_224a_1948} proposed that for $w$ independent p-values, each generated by $w$ different methods and denoted by $\bar{P}^i$, the following will hold true for the statistic $\Psi$:
+  Fisher~\cite{fisher_224a_1948} proposed that for $k$ independent p-values, each generated by $k$ different methods and denoted by $\bar{P}^i$, the following will hold true for the statistic $\Psi$:
   \begin{equation*}
     \begin{aligned}
-      \Psi &= \sum_{i=1}^w -2 \log \left( \bar{P}^i \right) \\
-        \Psi &\sim \chi^2_{2w}
+      \Psi &= \sum_{i=1}^k -2 \log \left( \bar{P}^i \right) \\
+        \Psi &\sim \chi^2_{2k}
     \end{aligned}
@@
   \begin{equation*}
-    f = \frac{\mathrm{E}[\Psi]^2}{\mathrm{Var}[\Psi]} ~~~\text{and}~~~ c = \frac{\mathrm{Var}[\Psi]}{2\mathrm{E}[\Psi]} = \frac{w}{f}
+    f = \frac{\mathrm{E}[\Psi]^2}{\mathrm{Var}[\Psi]} ~~~\text{and}~~~ c = \frac{\mathrm{Var}[\Psi]}{2\mathrm{E}[\Psi]} = \frac{k}{f}
   \end{equation*}
@@
   \begin{equation*}
-    \mathrm{E}[\Psi] = 2w ~~~\text{and}~~~ \mathrm{Var}[\Psi] = 4w + 2\sum_{i<j} \mathrm{Cov}\left( -2\log(\bar{P}^i), -2\log(\bar{P}^j) \right)
+    \mathrm{E}[\Psi] = 2k ~~~\text{and}~~~ \mathrm{Var}[\Psi] = 4k + 2\sum_{i<j} \mathrm{Cov}\left( -2\log(\bar{P}^i), -2\log(\bar{P}^j) \right)
   \end{equation*}
@@
     \mathrm{Cov}\left( -2\log(\bar{P}^i), -2\log(\bar{P}^j) \right) \approx 3.263 \rho_{ij} + 0.710 \rho_{ij}^2 + 0.027 \rho_{ij}^3
-    \label{eqn:covariance-pvalues}
+    \label{eqn:suppl_covariance-pvalues}
   \end{equation}
@@
         & \hat{P}_j = 1.0 - \Phi_{2f}\left( \psi / c \right) \\
-        \text{where},~ &\psi = -2 \sum_{i=1}^w \log(\bar{P}^i_j) ~~~\text{and}~~~ \Phi_{2f} = \mathrm{CDF}\left( \chi^2_{2f} \right)
+        \text{where},~ &\psi = -2 \sum_{i=1}^k \log(\bar{P}^i_j) ~~~\text{and}~~~ \Phi_{2f} = \mathrm{CDF}\left( \chi^2_{2f} \right)
     \end{aligned}
-    \label{eqn:pvalue-combined}
+    \label{eqn:suppl_pvalue-combined}
   \end{equation}
 
-  The p-value merging and consensus method in \ac{micone} (refer Documentation) uses Equation~\ref{eqn:covariance-pvalues} to estimate the covariance of the pvalues and Equation~\ref{eqn:pvalue-combined} to merge the p-values (obtained from bootstrapping) from the different correlation methods.
+  The p-value merging and consensus method in \ac{micone} (refer Documentation) uses Equation~\ref{eqn:suppl_covariance-pvalues} to estimate the covariance of the p-values and Equation~\ref{eqn:suppl_pvalue-combined} to merge the p-values (obtained from bootstrapping) from the different correlation methods.
   Note that we do not use Pearson and Spearman methods in the p-value merging step and these algorithms are only used for demonstration and comparison.

### 2022-05-07 sections/discussion.tex
Change: Update the Discussion for draft v2

@@
   Co-occurrence associations in microbial communities help identify important interactions that drive microbial community structure and organization.
-  Our analysis shows that networks generated using different combinations of tools and approaches can look significantly different from each other, highlighting the importance of a clear assessment of the source of variability and of tools that provide the most robust and accurate results.
-  Our newly developed integrated software for the inference of co-occurrence networks from 16S rRNA data, \ac{micone}, constitutes a freely customizable and user friendly pipeline that allows users to easily test combinations of tools and to compare networks generated by multiple possible choices (see Methods).
+  A myriad of tools and methods have been developed for different parts of the workflow for inference of co-occurrence networks from 16S rRNA data.
+  Our analysis shows that networks generated using different combinations of tools and approaches are significantly different from each other, highlighting the need for a clear assessment of the source of variability and of tools that provide the most robust and accurate results.
+  Our newly developed integrated software for the inference of co-occurrence networks from 16S rRNA data, \ac{micone}, is a freely customizable and user-friendly pipeline that allows users to easily test combinations of tools and to compare networks generated by multiple possible choices.
   Importantly, in addition to revisiting the test cases presented in this work, users will be able to explore the effect of various tool combinations on their own datasets of interest.
-  The \ac{micone} pipeline is built in a modular fashion; its plug-and-play architecture will make it possible for users to add new tools and steps, either from existing packages, or from packages that were not examined in the present work, as well as those developed in the future.
-  The easy-to-use nature, the configurability and modular nature of the \ac{micone} package enables users to start and end the pipeline at any point in the worflow, enabling users to run parts of the pipeline in isolation.
-  The \ac{micone} Python package provides functions and methods to perform detailed analysis of the count matrices and the inferred networks.
-  Additionally, the inferred networks by default are exported to a custom JSON format (see Supplementary), but are interchangeable to Cytoscape~\cite{shannonCytoscapeSoftwareEnvironment2003}, GML~\cite{himsoltGMLPortableGraph2010}, and many other popular formats via the \ac{micone} Python package.
+  The \ac{micone} pipeline is built in a modular fashion; its plug-and-play architecture will make it possible for users to add new tools and steps, either using existing packages that were not examined in the present work or those developed in the future.
+  The easy-to-use nature, the configurability and modular nature of the \ac{micone} package enables users to start and end the pipeline at any point in the workflow, enabling users to run parts of the pipeline in isolation.
+  This would allow for the user to bring in their own data at any step of the pipeline and explore the effects of the desired downstream data processing tools.
+  The \ac{micone} Python package provides functions and methods to perform detailed analysis of the count matrices and the co-occurrence networks.
+  Additionally, the inferred networks by default are exported to a custom JSON format (see Supplementary), but are easily exported to Cytoscape~\cite{shannonCytoscapeSoftwareEnvironment2003}, GML~\cite{himsoltGMLPortableGraph2010}, and many other popular formats via the \ac{micone} Python package.
 
@@
 
-  \subsection*{The default pipeline and suggested tools}
+  \subsection*{The default pipeline and recommended tools}
 
-  The main outcome of this work is two-fold: on one hand we transparently reveal the dependence of co-occurrence networks on tool and parameter choices, and built a tool for this purpose, enabling to more rigorously assess and compare existing networks.
-  On the other hand, we take advantage of our spectrum of computational options and the availability of mock and synthetic datasets, to suggest a default standard setting, and a consensus approach, likely to yield networks that are robust across multiple tool/parameter choices.
+  The main outcome of this work is two-fold.
+  On one hand, we transparently reveal the dependence of co-occurrence networks on tool and parameter choices, and built a tool for this purpose, enabling to more rigorously assess and compare existing networks.
+  On the other hand, we take advantage of our spectrum of computational options and the availability of mock and synthetic datasets, to suggest a default standard setting, and a consensus approach, that is more likely to yield networks that are robust across multiple tool/parameter choices.
+  An important caveat related to this last point is the fact that due to the lack of a universal standard for microbial interaction data, our conclusions are based on the specific datasets used in our analysis.
+  While our datasets comprise several mock and synthetic datasets that cover a fairly diverse range of abundance distributions and network topologies, datasets that have drastically different distributions may require a re-assessment of the best settings through our pipeline.
 
-  An important caveat related to this last point is the fact that due to the lack of a universal standard for microbial interaction data, our conclusions are based on the specific datasets used in our analysis.
-  While our datasets cover several mock and synthetic datasets that cover a fairly diverse range of abundance distributions and network topologies, datasets that have drastically different distributions may require a re-assessment of the best settings through our pipeline.
-
-  It is worth pointing out some additional more specific conclusions stemming from the individual steps of our analysis.
-  The different denoising/clustering methods differ mostly in their identification of sequences that are in low abundances.
-  Hence, they do not have much of an impact on the inferred co-occurrence networks when the sequences of low abundance are removed.
-  However, comparison of inferred and expected reference sequences and their abundances in mock community datasets has allowed us to identify \ac{dada2} as the method which best recapitulates the expected sequence composition.
-  For the chimera checking module, we suggest using the remove bimera method since it was developed specifically for \ac{dada2} and performs equally as well as uchime.
+  Additionally, it is worth pointing out some additional more specific conclusions stemming from the individual steps of our analysis.
+  The different denoising/clustering methods differ mostly in their identification of sequences that are in lower abundances.
+  Hence, they do not have much of an impact on the inferred co-occurrence networks when the sequences of low abundance are removed (Figure~\ref{fig:figure_s1}).
+  Comparison of inferred and expected reference sequences and their abundances in mock community datasets has allowed us to identify \ac{dada2} as the method which best recapitulates the expected sequence composition.
+  For the chimera checking module, we suggest using the remove bimera method since it was developed in conjunction with \ac{dada2} and its performance does not significantly differ from uchime-denovo.
   For the current work we have decided to focus on the tools most widely used at the time of the analysis.
-  Some tools there were not as widely used (e.g.
+  Some tools which were not as widely used (e.g.
 dbOTU3~\cite{Olesen2017}) as well as older popular methods like mothur~\cite{Schloss2009} have not been included in the study, but could be added into the pipelines in future updated analyses.
 
-  The choice of taxonomy database was found to be the most important factor in the inference of a microbial co-occurrence network, contributing $\sim65.4\%$ of the total variance.
+  The choice of taxonomy database was found to be the most important factor in the inference of microbial co-occurrence networks, contributing $\sim65.4\%$ of the total variance.
   The frequent changes in the taxonomy nomenclature coupled with the frequency of updates to the various 16S reference databases create inherent differences \cite{Balvociute2017} in taxonomy hierarchies in these databases.
-  Our analysis revealed that no particular reference database performs better than the others across all scenarios.
-We suggest that that choice of the database should be made based on possible reported or inferred biases in the representation of given biomes in a specific databases \cite{Balvociute2017}.
-  The default reference database in the pipeline is the \ac{gg} reference database along with the NaiveBayes classifier as the query tool.
-  The reason for our choice stem from the popularity of the \ac{gg} database in taxonomic studies, which would enable easy comparison across datasets.
-  We also enable users to use custom databases \cite{Ritari2015} with the blast+ and naive bayes classifiers that are incorporated into the pipeline (from \ac{qiime2}).
+  Our analysis revealed that no particular reference database performs better than the others across the different mock dataset benchmarks.
+  We suggest that that choice of the database should be made based on possible reported or inferred biases in the representation of given biomes in a specific databases~\cite{Balvociute2017}.
+  The default reference database in the pipeline is the \ac{gg} reference database along with the ``Naive Bayes'' classifier as the query tool.
+  The reason for our choice stems from the popularity of the \ac{gg} database in taxonomic studies, which would enable easy comparison across datasets.
+  We also enable users to use custom databases~\cite{Ritari2015,iiRESCRIPtReproducibleSequence2021} with the BLAST and Naive Bayes classifiers that are incorporated into the pipeline (from \ac{qiime2}).
 
   The \ac{op} step of the pipeline is second in its contribution to total network variance.
-  This can be attributed to the large number of nodes that are added to the final networks when filtering is turned off.
+  This can be attributed to the large number of nodes that are added to the final networks when the filtering is turned off.
   Additionally, a very large number of nodes also decreases the accuracy of the network inference algorithms for the same sample size~\cite{peschelNetCoMiNetworkConstruction2020} and increases the computational complexity~\cite{tackmannRapidInferenceDirect2019}.
-  We observe that filtering out taxa that are present in low abundances in all samples decrease the proportion of taxa in common between taxonomy tables generated using different reference databases, but not significantly (Figure~\cite{fig:figure_s5}).
-  However, we do observe that the reduction in the number of taxa leads to better agreement in the networks inferred through different methods (Figure~\ref{fig:figure_s1}).
+  We observe that filtering out taxa that are present in low abundances in all samples increases the proportion of taxa in common between taxonomy tables generated using different reference databases (Figure~\ref{fig:figure_s5}), providing another reason for filtering.
+  We also observe that the reduction in the number of taxa leads to better agreement in the networks inferred through different methods (Figure~\ref{fig:figure_s1}).
   Moreover, filtering is necessary in order to increase the power in tests of significance when the number of taxa is much greater than the number of samples.
@@
   The networks generated by different network inference methods show considerable differences in edge-density and connectivity.
-  One reason for this is the underlying assumptions regarding sparsity, distribution and compositionality that the algorithms make.
+  One reason for this is the underlying assumptions regarding sparsity, distribution and compositionality, that the algorithms make.
   The consensus network created by merging the networks inferred from the different network inference methods enables the creation of a network whose links have evidence based on multiple inference algorithms.
   We have developed two consensus algorithms to support this effort, the simple voting method and the scaled-sum method.
-  Ultimately, we find that the scaled-sum method along with pvalue merging (applied to the correlation-based methods) performs the best on synthetic datasets, therefore this is the default for the \ac{ni} step of the pipeline.
+  Ultimately, we find that the scaled-sum method along with p-value merging (applied to the correlation-based methods) performs the best on synthetic datasets, therefore this is the default for the \ac{ni} step of the pipeline.
+  The consensus algorithm not only leads to the creation of a network that has a higher number of true positives, it also decreases the variability in the networks due to choosing a different tool or parameter in any of the previous steps of the pipeline (Figure~\ref{fig:figure7}B).
 
-  Exploring the effects of these combinations of methods on the resultant networks is difficult and inconvenient since different tools differ in their input and output formats and require inter-converting between the various formats.
-  The pipeline facilitates this comparative exploration by providing a variety of modules for inter-conversion between various formats, and by allowing easy incorporation of new tools as modules.
-  Since there exists no ``best method'' for the various steps of 16S data analysis, especially for microbial network inference, \ac{micone} is intended to help researchers to identify the methods and algorithms that are most suitable for their datasets in a easy-to-use and reproducible manner.
+  Exploring the effects of these combinations of methods on the resultant networks is difficult and inconvenient since different tools differ in their input and output formats and require interconversions between the various formats.
+  The pipeline facilitates this comparative exploration by providing a variety of modules for interconversion between various formats, and by allowing for easy incorporation of new tools as modules.
+  Since there exists no ``best method'' for the various steps of 16S data analysis, especially for microbial network inference, \ac{micone} is intended to help researchers to identify the methods and algorithms that are most suitable for their datasets in an easy-to-use and reproducible manner.
   We envision that \ac{micone}, and the underlying tools and databases that help process amplicon sequencing data into co-occurrence networks, will be increasingly useful towards building large comparative analyses across studies.
   By having a unified transparent tool to compute networks, it will be possible to reprocess available 16S datasets to obtain networks that are directly comparable to each other.
-  Furthermore, even in the analysis of published networks across studies and processing methods, \ac{micone} could help understand underlying biases of each network, which could in turn be taken into account upon making cross-study comparisons.
+  Furthermore, even in the analysis of published networks across studies and processing methods, \ac{micone} could help understand underlying biases of each network, which could in turn be taken into account when making cross-study comparisons.
 
@@
 
-  The current analyses include one fecal microbiome transplant dataset with healthy and autistic samples, three mock community datasets and several datasets generated by two synthetic interaction methods.
+  The current analyses make use of one fecal microbiome transplant dataset with healthy and autistic samples, three mock community datasets and several datasets generated by two synthetic interaction methods.
   The analyses presented in this study are specific to the datasets utilized, and due to the lack of a gold standard reference dataset for microbial interaction data, it is not possible to generalize these results to datasets with widely different microbiome profiles.
@@
 
-  Although the pipeline currently supports changing of parameter values for the most commonly used parameters for several tools, we only use the those default parameter values that are recommended by the tool creators.
+  Although the pipeline currently supports changing of parameter values for the most commonly used parameters for several tools, we only use those default parameter values that are recommended by the tool creators.
   The exploration of the effects of parameter values on the inferred association networks might help optimize these values for better network inference, but these are not part of this current study.
-  Additionally, despite incorporating a wide range of tools and algorithsm into the \ac{micone} pipeline, there are several other tools that were omitted due time and computational constraints.
+  Additionally, despite incorporating a wide range of tools and algorithms into the \ac{micone} pipeline, there were several other tools that were omitted due time and computational constraints.
   However, the modular nature of the \ac{micone} pipeline enables easy addition of new tools and methods at any step of the workflow.
@@
 
-  % TODO: Health and disease
-  In summary, we believe that \ac{micone} is a useful tool for the exploration of microbial networks generated from 16S data, allowing for rapid, configurable and reproducible inference of microbial networks and further formulation of hypotheses about the role of these interactions on community composition and stability.
+  % TODO: Talk health and disease or something more specific?
+  In summary, we believe that \ac{micone} is a useful tool for the exploration of microbial networks generated from 16S data.
+  It allows for rapid, configurable and reproducible inference of microbial networks and further for the formulation of hypotheses about the role of these interactions on community composition and stability.

### 2022-05-07 sections/results.tex
Change: Update the tables and figures for draft v2

@@
   Figure~\ref{fig:figure3}B on the other hand shows an increase in dissimilarity between each pair of methods suggesting that the methods might differ in the assignment of sequences lower in abundance.
-  In order to verify this claim, for each of these methods we filtered out only the top 1000 representative sequences and recalculated the distance metrics for each method (Figure \ref{fig:figureS3}).
+  In order to verify this claim, for each of these methods we filtered out only the top 1000 representative sequences and recalculated the distance metrics for each method (Figure \ref{fig:figure_s3}).
   We observe that the both the weighted and unweighted UniFrac distances are increased implying that the top representative sequences generated by the different methods are not as similar to each other.
@@
   Despite this partial degree of assessment, we wish to suggest a combination of tools and parameters that produce networks that are derived from the combination of tools which performed best on the mock communities, and displayed the highest robustness to switching to alternative methods.
-  These tools and parameters are chosen as the defaults for the pipeline and are given in Table~\ref{tab:default_options}.
+  These tools and parameters are chosen as the defaults for the pipeline and are given in Table~\ref{tab:micone_tools}.
 
@@
   Figure~\ref{fig:figure7}C shows the co-occurrence networks inferred for the healthy subjects (control) and autistic subjects in the fecal microbiome transplant study~\cite{Kang2017}.
-  These consensus networks were generated using the default tools and parameters from Table~\ref{tab:default_options}.
+  These consensus networks were generated using the default tools and parameters from Table~\ref{tab:micone_tools}.

### 2022-05-07 sections/tables_figures.tex
Change: Update the Results for draft v2
Change: Update the tables and figures for draft v2

@@
 
-  \begin{table}[h]
+  \begin{table}[H]
     \centering
@@
                                                 &  & De novo & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
-                                                &  & Dada2 & \cite{Callahan2016} \\
+                                                &  & \rowcolor{lightgray} Dada2 & \cite{Callahan2016} \\
                                                 &  & Deblur & \cite{Amir2017,bolyenReproducibleInteractiveScalable2019} \\
@@
                                                 & \multirow{2}{*}{Chimera checking} & Uchime-denovo & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
-                                                & & Remove bimera & \cite{Callahan2016} \\
+                                                & & \rowcolor{lightgray} Remove bimera & \cite{Callahan2016} \\
                                                 \hline
       \multirow{5}{*}{Taxonomy assignment} &  \multirow{2}{*}{Query tool} & Blast & \cite{camachoBLASTArchitectureApplications2009,bokulichOptimizingTaxonomicClassification2018} \\
-                                           &  & Naive bayes classifier & \cite{bokulichOptimizingTaxonomicClassification2018} \\
+                                           &  & \rowcolor{lightgray} Naive bayes classifier & \cite{bokulichOptimizingTaxonomicClassification2018} \\
                                            \cline{2-4}
-                                           & \multirow{3}{*}{Database} & Greengenes & \cite{DeSantis2006} \\
-                                           & & SILVA & \cite{Quast2012} \\
-                                           & & NCBI RefSeq & \cite{Sayers2009} \\
+                                           & \multirow{3}{*}{Database} & \rowcolor{lightgray} Greengenes 13\_8 & \cite{DeSantis2006} \\
+                                           & & SILVA 138 & \cite{Quast2012} \\
+                                           & & NCBI RefSeq (Oct 2021) & \cite{Sayers2009} \\
       \hline
@@
                                       & & Observation sum threshold & 10             \\ \cline{2-4}
-                                      & \multirow{3}{*}{Filter(on)}  & Prevalence threshold      & 0.05           \\
-                                      & & Abundance threshold       & 0.01           \\
-                                      & & Observation sum threshold & 100            \\
+                                      & \multirow{3}{*}{Filter(on)}  & \rowcolor{lightgray} Prevalence threshold      & 0.05           \\
+                                      & & \rowcolor{lightgray} Abundance threshold       & 0.01           \\
+                                      & & \rowcolor{lightgray} Observation sum threshold & 100            \\
       \hline
-      \multirow{12}{*}{Network Inference} & \multirow{2}{*}{Bootstrapping}& \texttt{fastspar\_bootstraps} v1.0 & \cite{Watts2018} \\
-                                          & & \texttt{fastspar\_pvalues} v1.0 & \cite{Watts2018} \\
+      \multirow{12}{*}{Network Inference} & \multirow{2}{*}{Bootstrapping}& fastspar\_bootstraps v1.0 & \cite{Watts2018} \\
+                                          & & fastspar\_pvalues v1.0 & \cite{Watts2018} \\
                                           \cline{2-4}
@@
                                           & & Spearman & - \\
-                                          & & propr v2.1.2 & \cite{quinnProprRpackageIdentifying2017} \\
+                                          & & propr v2.1.2 & \cite{quinnProprRpackageIdentifying2017} \\ \cline{2-4}
+                                          & \multirow{2}{*}{Consensus algorithm} & \rowcolor{lightgray} scaled-sum & 0.667 \\
+                                          & & simple voting & 0.667 \\
       \hline
     \end{tabular}
-    \caption{Tools used in the \ac{micone} pipeline}
+    \caption{
+      \textbf{Tools used in the \ac{micone} pipeline}.
+      The tools highlighted in gray are the defaults for the pipeline and are recommended based on the benchmarks with the mock and synthetic datasets.
+      The consensus algorithm in the \ac{ni} step incorporates all the modules (bootstrapping, direct association and correlation-based) in that step to generate the consensus network.
+    }
     \label{tab:micone_tools}
-  \end{table}
-
-  \begin{table}[h]
-    \centering
-    \small
-    \begin{tabular}{|c|c|c|}
-      \hline
-      \textbf{Workflow step} & \textbf{Default tool} \\
-      \hline
-      Denoising and Clustering & Dada2 and remove bimera \\
-      Taxonomy Assignment & \ac{gg} with NaiveBayes classifier \\
-      OTU Processing & Thresholds on prevalence and abundance \\
-      Network Inference & scaled-sum consensus method \\
-      \hline
-    \end{tabular}
-    \caption{Default tools and parameters of the \ac{micone} pipeline}
-    \label{tab:default_options}
   \end{table}
@@
 
-  \begin{figure}[h]
+  \begin{figure}[H]
     \centering
@@
   \end{figure}
-  \begin{figure}[ht!]
+  \begin{figure}[H]
     \centering
@@
       \textbf{The workflow of the \ac{micone} pipeline}.
-      The steps can be grouped into five major groups: \textbf{(SP)} \textbf{S}equence \textbf{P}rocessig, \textbf{(DC)} \textbf{D}enoising and \textbf{C}lustering, \textbf{(TA)} \textbf{T}axonomy \textbf{A}ssignment, \textbf{(OP)} \textbf{O}TU or ESV \textbf{P}rocessing, and \textbf{(NI)} \textbf{N}etwork \textbf{I}nference.
+      The steps of the workflow can be broken down into five major groups: \textbf{(SP)} \textbf{S}equence \textbf{P}rocessig, \textbf{(DC)} \textbf{D}enoising and \textbf{C}lustering, \textbf{(TA)} \textbf{T}axonomy \textbf{A}ssignment, \textbf{(OP)} \textbf{O}TU and ESV \textbf{P}rocessing, and \textbf{(NI)} \textbf{N}etwork \textbf{I}nference.
       Each step incorporates several processes, each of which in turn have several alternate algorithms for the same task (indicated by the text to the right of the blue boxes).
       The text along the arrows describes the data that is being passed from one step to another.
-      The final output of the pipeline are consensus networks generated from the co-occurrence networks.
-      For details on each process and data types, see Methods.
+      The final output of the pipeline is the consensus network generated from the inferred co-occurrence networks.
+      For details on each process and the data outputs, see Methods.
     }
@@
 
-  \begin{figure}
+  % TODO: Add equations for the linear model to the left of the pie chart
+  \begin{figure}[H]
     \centering
-    \includegraphics[width=1.0\linewidth]{figure2.pdf}
+    \includegraphics[width=\textwidth]{figure2.pdf}
   \end{figure}
-  \begin{figure}
+  \begin{figure}[H]
     \centering
@@
       \textbf{The choice of reference database contributes to the most variance in the networks}.
-      \textbf{(A)} The total relative variance in the networks contributed by the DC, CC (chimera checking), TA and OP steps of the pipeline (right) and the linear model used to calculate the relative variance (left), see the Methods section for details.
-      The analyses was performed on the healthy stool samples in a fecal microbiome transplant dataset~\ref{Kang2017}.
-      The taxonomy database contributes most to the variance between the networks (65.4\%) followed by filtering of the counts matrix (26.8\%) in the OP step.
+      \textbf{(A)} The percentage of variance in the networks contributed by the DC, CC (chimera checking), TA and OP steps of the pipeline (right) and the linear model used to calculate the variance (left), see the Methods section for details.
+      The data used for the analysis were the healthy stool samples from the fecal microbiome transplant dataset~\ref{Kang2017}.
+      The taxonomy database contributes most to the variance between the networks (65.4\%) followed by the filtering of the counts matrix (26.8\%) in the OP step.
+      The variation due to the DC and CC steps are comparatively not as significant.
       \textbf{(B)} All combinations of inferred networks are shown as points on a PCA plot.
@@
       The color of the points corresponds to the reference taxonomy database used to reconstruct those networks (left), and the network inference algorithm used to infer the networks (right).
-      % TODO: The points aren't really separated
-      The points on the PCA plot can be separated based on the TA step and the differences due to the DC and CC steps are comparatively not as significant.
+      The points on the PCA plot can be grouped based on the TA step, but at the NI step only some methods create high variability in the inferred networks.
     }
@@
 
-  \begin{figure}
+  \begin{figure}[H]
     \centering
@@
   \end{figure}
-  \begin{figure}
+  \begin{figure}[H]
     \centering
@@
       \textbf{(B)} The relatively larger average unweighted UniFrac distance indicates that methods differ in their identification of sequences that are lower in abundance.
-      \textbf{(C, D)} The distributions of the average weighted and unweighted UniFrac distance between the expected sequence profile and the calculated sequence profile in the mock datasets.
+      The number of \ac{otu}s or \ac{esv}s generated by the respective methods are provided in the parenthesis next to the names.
+      \textbf{(C, D)} The distributions of the average weighted and unweighted UniFrac distance between the expected sequence profile and the predicted sequence profile in the mock datasets.
       The distributions of the average weighted UniFrac distance show that de novo (DN) and open reference (OR) were the best performing methods in most of the datasets, while they are worst performing methods under the unweighted UniFrac metric.
-      The above average performance of dada2 (D2) under both distance metrics combined with its approach of identifying \ac{esv}s using de novo methods, prompts us to propose it as the default method for the DC step.
+      The good performance of dada2 (D2) under both distance metrics combined with its approach of identifying \ac{esv}s using de novo methods, prompts us to propose it as the default method for the DC step.
     }
@@
       \textbf{Taxonomic reference databases vary widely in terms of their taxonomy assignments beyond the Order level.}
-      \textbf{(A)} The assignments of the top 50 representative sequences to their respective taxonomies using the three different reference databases show how the same sequences are assigned to different Genus under the different databases.
+      \textbf{(A)} The assignments of the top 50 representative sequences to their respective taxonomies using the three different reference databases.
+      This result clearly describes how the same sequences are assigned to different Genus under the different databases.
       A significant portion of the representative sequences are assigned to an ``unknown'' Genus in two of three databases (\ac{gg} and \ac{ncbi}).
       \textbf{(B)} The number of representative sequences assigned to the same taxonomic label when using different reference databases, shown for the top 100 sequences.
-      The number of mismatches are fewer at higher taxonomic levels but even at the Order level there exists greater than 51\% of mismatches, demonstrating the poor agreement in taxonomic labels assigned by the different databases.
+      The number of mismatches are fewer at higher taxonomic levels, but, even at the Order level there exists greater than 51\% of mismatches, demonstrating the poor agreement in taxonomic labels assigned by the different databases.
       \textbf{(C)} The Bray-Curtis dissimilarity between the expected taxonomy profile and calculated taxonomy profile in the mock datasets shows that there is no singular best choice of database for every dataset, as all the databases have similar performance.
-      However, we recommend the \ac{gg} database due to its popularity and if the goal is to compare against existing networks in literature.
+      However, we still recommend the \ac{gg} database and the Naive Bayes classifier as the defaults for the \ac{micone} pipeline.
+      Due to the popularity of the \ac{gg} database, comparison against existing networks in literature will be easier since many published datasets make use of this database in their analyses.
     }
@@
       \textbf{Networks generated using different network inference methods show notable differences both in terms of edge-density and connectivity}.
-      \textbf{(A)} The nine different networks generated by the different network inference methods (excluding \ac{mldm}) are very dissimilar.
-      The green links are positive associations and the orange links are negative associations.
+      \textbf{(A)} The nine different networks (excluding \ac{mldm}) generated by the different network inference methods are very dissimilar.
+      The green links are positive associations and the orange links represent negative associations.
       A threshold of 0.3 was set for the correlation-based methods (\ac{sparcc}, Propr, Spearman and Pearson) and a threshold of 0.01 was set for the other direct association methods.
-      \textbf{(B)} The node overlap Upset plot indicates that all the networks have a large number of common nodes involved in connections (33 out of 68).
+      The correlation-based methods in general produce graph with higher edge-densities.
+      \textbf{(B)} The node overlap Upset plot indicates that all the networks have a large proportion of common nodes involved in connections (33 out of 68).
       Whereas, \textbf{(C)} the edge overlap Upset plot shows that a very small fraction of these connections are actually shared (8 out of 202).
@@
       \textbf{The precision and sensitivity of the inferred networks on the NorTA synthetic interaction data.}
-      The ``I'' prefix indicates the individual network inference methods and the ``X'' prefix denotes the different consensus based methods.
+      The ``I'' prefix indicates the individual network inference methods and the ``X'' prefix denotes the consensus based methods that are a part of \ac{micone}.
       The different consensus based methods used are: pvalue merging (PM), scaled-sum (SS) and simple voting (SV) method.
@@
       The simple voting method when using presence of edges in all inferred networks as a requirement (parameter value 1.000), also outperforms \ac{spieceasi} on average precision (0.969).
+      Although \ac{spieceasi} has a higher performance on sensitivity, if the goal of network inference is to obtain the list of associations that have a high probability of existing in the real microbial community, then the consensus methods perform better.
       Therefore, we recommend the scaled-sum consensus method as the default tool for network inference.
@@
     \caption{
-      \textbf{The consensus algorithm drastically reduces variability in the networks and increases robustness of inferred networks}
+      \textbf{The consensus algorithm drastically reduces variability in the networks due to upstream data processing methods and increases robustness of inferred networks}
       \textbf{(A)} The network constructed using the default pipeline parameters (DC=\ac{dada2}, CC=remove bimera, TA=\ac{gg}, OP=Filter(on), NI=scaled-sum consensus) is compared with networks generated when one of the steps use a different tool.
-      The common connections (common with the default network) are in black, connections unique to the network are colored purple and connections in the default network but not present in the current network are gray.
+      The common connections (common with the default network) are in black, connections unique to the network are colored purple and connections in the default network, but not present in the current network are gray.
       Changing the NI, TA and OP steps leads to the creation of the most number of new nodes and edges.
-      \textbf{(B)} The boxplot showing the L1 distance between the networks generated by changing one step of the default pipeline and the network generated using the default parameters.
+      \textbf{(B)} The boxplot showing the L1 distance between the default network and the networks generated by changing one step of the default pipeline.
       The high value of L1 distances for the network inference step as compared to the other steps demonstrates how the consensus method is able to reduce the variability in the networks in the other steps of the pipeline.
+      This also shows that even though many new nodes were added in (A) when the TA and OP steps were changed, the L1 distances remain fairly small.
       \textbf{(C)} The networks generated for the control (left) and autistic (right) samples in the fecal microbiome transplant dataset.

### 2022-05-07 scripts/figure3/figure3_plot.R

@@ -73,8 +73,8 @@ unweighted_tidy <- tidy_up_data_ab(unweighted)
 weighted <- read.csv(weighted_unifrac_csv, sep = ",", header = TRUE)
 weighted_tidy <- tidy_up_data_ab(weighted)
 
-weighted_plot <- plot_heatmap(weighted_tidy, "weighted unifrac")
-unweighted_plot <- plot_heatmap(unweighted_tidy, "unweighted unifrac")
+weighted_plot <- plot_heatmap(weighted_tidy, "Weighted UniFrac")
+unweighted_plot <- plot_heatmap(unweighted_tidy, "UnWeighted UniFrac")
 
 heatmap_plot <- ggarrange(
     weighted_plot, unweighted_plot,
@@ -106,8 +106,8 @@ make_dot_plot <- function(unifrac_data, title) {
             y = "unifrac",
             color = "dataset",
             title = title,
-            xlab = "denoising/clustering method",
-            ylab = "unifrac distance",
+            xlab = "Denoising/Clustering method",
+            ylab = "UniFrac distance",
             ylim = c(0, 1),
             size = 3,
             add = "jitter",
@@ -132,7 +132,7 @@ mock12_weighted_tidy <- tidy_up_data_cd(mock12_weighted, "mock12")
 mock16_weighted_tidy <- tidy_up_data_cd(mock16_weighted, "mock16")
 
 mock_weighted_tidy <- rbind(mock4_weighted_tidy, mock12_weighted_tidy, mock16_weighted_tidy)
-mock_weighted_plot <- make_dot_plot(mock_weighted_tidy, "weighted unifrac")
+mock_weighted_plot <- make_dot_plot(mock_weighted_tidy, "Weighted UniFrac")
 
 # Unweighted
 mock4_unweighted <- read.csv(mock4_unweighted_csv, sep = ",", header = TRUE)
@@ -144,7 +144,7 @@ mock12_unweighted_tidy <- tidy_up_data_cd(mock12_unweighted, "mock12")
 mock16_unweighted_tidy <- tidy_up_data_cd(mock16_unweighted, "mock16")
 
 mock_unweighted_tidy <- rbind(mock4_unweighted_tidy, mock12_unweighted_tidy, mock16_unweighted_tidy)
-mock_unweighted_plot <- make_dot_plot(mock_unweighted_tidy, "unweighted unifrac")
[…]

### 2022-05-07 scripts/figure4/figure4_plot.R

@@ -187,5 +187,5 @@ ggsave(output_file_c, width = 11, height = 5)
 
 ################################################################################
 # Combine figures
-ggarrange(final_plot_a, combined_plot_b, dot_plot_facet_c, nrow = 3, ncol = 1)
+ggarrange(final_plot_a, combined_plot_b, dot_plot_facet_c, nrow = 3, ncol = 1, heights = c(2, 1, 1))
 ggsave(output_file, width = 11, height = 16)

### 2022-05-07 scripts/figure5/figure5_plot.R

@@ -182,4 +182,4 @@ ggsave(output_file_bc, width = 11, height = 5)
 
 
 final_plot <- ggarrange(a_plot, bc_plot, ncol = 1, heights = c(1.3, 1.0))
-ggsave(output_file, final_plot, width = 11, height = 12)
+ggsave(output_file, final_plot, width = 11, height = 15)

### 2022-05-07 scripts/figure7/figure7_plot.R

@@ -127,6 +127,8 @@ l1_data[l1_data == "closed_reference(gg_97)"] <- "CR"
 l1_data[l1_data == "de_novo"] <- "DN"
 l1_data[l1_data == "dada2"] <- "D2"
 l1_data[l1_data == "deblur"] <- "DB"
+l1_data[l1_data == "naive_bayes(silva_138_99)"] <- "NaiveBayes(SILVA)"
+l1_data[l1_data == "blast(ncbi)"] <- "BLAST(NCBI)"
 l1_data[l1_data == "normalize_filter(on)"] <- "Filter(on)"
 l1_data[l1_data == "normalize_filter(off)"] <- "Filter(off)"
 
@@ -213,5 +215,5 @@ c_plot <- annotate_figure(c_plot, fig.lab="C", fig.lab.pos="top.left", fig.lab.s
 
 
 final_plot <- ggarrange(a_plot, b_plot, c_plot, ncol = 1)
-ggsave(output_file, width = 11, height = 12)
+ggsave(output_file, width = 11, height = 14)
 

### 2022-05-13 sections/discussion.tex
Add: Information about biological benchmark data in Discussion

@@
   Incorporating datasets from a broad spectrum of biomes with varying microbial distributions into \ac{micone} will likely increase the robustness and generalizability of the results from these analyses.
+  Additionally, the utilization of literature-based interactions~\cite{lima-mendezDeterminantsCommunityStructure2015a} or biological benchmark interaction data~\cite{sungGlobalMetabolicInteraction2017a} as additional sources for benchmarking
   Furthermore, expanding the scope of the \ac{micone} pipeline to also include metagenomics sequencing data will enable an even broader access to a diverse range of datasets as well as domains of life that can be incorporated into the analysis.

### 2022-05-13 sections/results.tex
Add: Information about biological benchmark data in Discussion

@@
   % TODO: Show proof??
-  The change of taxonomy database and the filtering in the \ac{OP} step when turned off also adds a large number of new nodes, but the edges between the common nodes (with the default network) are fairly conserved.
+  The change of taxonomy database and the filtering in the \ac{op} step when turned off also adds a large number of new nodes, but the edges between the common nodes (with the default network) are fairly conserved.
 

### 2022-05-13 sections/discussion.tex
Change: Updates based on Reviewer 2 comments

@@
   Incorporating datasets from a broad spectrum of biomes with varying microbial distributions into \ac{micone} will likely increase the robustness and generalizability of the results from these analyses.
+  The network analyses in this study are primarily at the Genus level, wherein the lowest resolution of a node is a Genus and if an entity cannot be resolved to the Genus level and next lowest taxonomic level is used (e.g.
+Family).
+  As a consequence, two entities belonging to the same lineage where one entity is resolved to the Genus level and another is resolved to the Family level are treated as two different nodes in the network.
+  Therefore, the development of a metric of overlap to compare nodes with shared lineages within and across networks could provide more biologically and phylogentically relevant comparisons.
+  Further analyses could also include networks at the Species or other taxonomic levels.
+
   Additionally, the utilization of literature-based interactions~\cite{lima-mendezDeterminantsCommunityStructure2015a} or biological benchmark interaction data~\cite{sungGlobalMetabolicInteraction2017a} as additional sources for benchmarking

### 2022-05-13 sections/results.tex
Change: Updates based on Reviewer 2 comments

@@
   Based on their slightly better performance on the mock datasets, their \textit{de novo} error correcting nature and other previous studies~\cite{Nearing2018}, \ac{dada2} and Deblur appear to be the most reliable.
+  This is because the open-reference and de novo clustering methods return a much larger number of \ac{otu}s compared to the other pipelines and this would affect the accuracy of the network inference step if filtering is not performed.
   Overall, since \ac{dada2} as compared to Deblur, displays better performance on all the mock datasets and under the weighted UniFrac metric, we have decided to recommend this tool for the DC step of the pipeline.
+  However, if comparison across studies that have sequenced different 16S regions is required, closed-reference and open-reference might be a better option.
 

### 2022-05-13 sections/tables_figures.tex
Change: Updates based on Reviewer 2 comments

@@
       \textbf{(A)} The percentage of variance in the networks contributed by the DC, CC (chimera checking), TA and OP steps of the pipeline (right) and the linear model used to calculate the variance (left), see the Methods section for details.
-      The data used for the analysis were the healthy stool samples from the fecal microbiome transplant dataset~\ref{Kang2017}.
+      The data used for the analysis were the healthy stool samples from the fecal microbiome transplant dataset~\cite{Kang2017}.
       The taxonomy database contributes most to the variance between the networks (65.4\%) followed by the filtering of the counts matrix (26.8\%) in the OP step.
@@
       The number of \ac{otu}s or \ac{esv}s generated by the respective methods are provided in the parenthesis next to the names.
+      The data used for the analysis in \textbf{(A, B)} were all the samples (healthy and autistic) from the fecal microbiome transplant dataset.
       \textbf{(C, D)} The distributions of the average weighted and unweighted UniFrac distance between the expected sequence profile and the predicted sequence profile in the mock datasets.
@@
       The good performance of dada2 (D2) under both distance metrics combined with its approach of identifying \ac{esv}s using de novo methods, prompts us to propose it as the default method for the DC step.
+      The data used for the analysis in \textbf{(C, D)} were the mock datasets from mockrobiota~\cite{Bokulich2016}.
     }
@@
       The number of mismatches are fewer at higher taxonomic levels, but, even at the Order level there exists greater than 51\% of mismatches, demonstrating the poor agreement in taxonomic labels assigned by the different databases.
+      The data used for the analysis in \textbf{(A, B)} were all the samples (healthy and autistic) from the fecal microbiome transplant dataset.
       \textbf{(C)} The Bray-Curtis dissimilarity between the expected taxonomy profile and calculated taxonomy profile in the mock datasets shows that there is no singular best choice of database for every dataset, as all the databases have similar performance.
@@
       Due to the popularity of the \ac{gg} database, comparison against existing networks in literature will be easier since many published datasets make use of this database in their analyses.
+      The datasets used for the analysis in \textbf{(C)} were the mock datasets from mockrobiota.
     }
@@
       Whereas, \textbf{(C)} the edge overlap Upset plot shows that a very small fraction of these connections are actually shared (8 out of 202).
+      The data used in this analysis were the healthy stool samples from the fecal microbiome transplant dataset.
     }
@@
       These networks were generated using the default tools and parameters recommended by the \ac{micone} pipeline.
+      The data used in this analysis were the healthy stool samples from the fecal microbiome transplant dataset.
     }

### 2022-05-16 sections/discussion.tex
Change: Updates based on Reviewer 2 minor comments

@@
   Our analysis revealed that no particular reference database performs better than the others across the different mock dataset benchmarks.
-  We suggest that that choice of the database should be made based on possible reported or inferred biases in the representation of given biomes in a specific databases~\cite{Balvociute2017}.
   The default reference database in the pipeline is the \ac{gg} reference database along with the ``Naive Bayes'' classifier as the query tool.
   The reason for our choice stems from the popularity of the \ac{gg} database in taxonomic studies, which would enable easy comparison across datasets.
-  We also enable users to use custom databases~\cite{Ritari2015,iiRESCRIPtReproducibleSequence2021} with the BLAST and Naive Bayes classifiers that are incorporated into the pipeline (from \ac{qiime2}).
+  % TODO: Add ref: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6440677/
+  However, for newer studies we recommend using SILVA database because of its size and taxonomic comprehensiveness~\cite{iiRESCRIPtReproducibleSequence2021} and since \ac{gg} has not been updated since 2013.
+  Additionally, we also enable users to use custom databases~\cite{Ritari2015,iiRESCRIPtReproducibleSequence2021} with the BLAST and Naive Bayes classifiers that are incorporated into the pipeline (from \ac{qiime2}).
+  % TODO: Add ref: https://pubmed.ncbi.nlm.nih.gov/32106809/
+  We suggest that that choice of the database should be made based on possible reported or inferred biases in the representation of given biomes in a specific databases~\cite{Balvociute2017,iiRESCRIPtReproducibleSequence2021}.
 

### 2022-05-16 sections/results.tex
Change: Updates based on Reviewer 2 minor comments

@@
   The systematic analyses performed in the previous sections clearly show that the choice of tools and parameters can have a big impact on the final co-occurrence network.
-  For some of these choices (e.g. \ac{dada2} vs. Deblur), there is no clear metric to establish the best protocol.
+  For some of these choices (e.g. for the \ac{ta} step), there is no clear metric to establish the best protocol.
   For other choices, the mock communities and synthetic data provide an opportunity to select the combinations of tools that yield more accurate and robust results.
@@
   This reinforces the importance of the consensus network in increasing the robustness of the inferred networks to different processing methods.
+
   Figure~\ref{fig:figure7}C shows the co-occurrence networks inferred for the healthy subjects (control) and autistic subjects in the fecal microbiome transplant study~\cite{Kang2017}.
+  This figure demonstrates a typical use case of comparative analysis of networks using the \ac{micone} pipeline.
   These consensus networks were generated using the default tools and parameters from Table~\ref{tab:micone_tools}.
+  As a consequence of using the consensus network algorithm, the final co-occurrence networks are sparse and can be visually compared and examined.
+  % TODO: How much detail should we go into?
+  We observe 10 unique links in the network for control samples and 3 unique links in the network for autistic subjects.
+  These unique connections can serve as potential starting points for literature surveys and further for experimental confirmations.

### 2022-05-18 sections/results.tex
Change: Edit the supplementary section for draft v2

@@
   In other words, each point corresponds to a network generated using different combinations of tools, and captures how much the final network is affected by such a choice.
-  The importance of the TA step which was implied in Figure \ref{fig:figure2}A is also reflected here, since most of the networks can be clearly separated based on the database used (Figure \ref{fig:figure2}B).
+  The importance of the TA step which was implied in Figure \ref{fig:figure2}A is also reflected here, since most of the networks can be separated based on the database used (Figure \ref{fig:figure2}B).
   However, it is also evident that some networks, especially those inferred using the direct association based network inference methods, are much closer in the PCA plot regardless of the reference database used.
@@
   The Deblur method returns a very small number of representative sequences (2388) and this could account for the reason for the high dissimilarity with the other methods as well as irregular performance on the mock data.
-  There is no method that clearly outperforms the rest in all datasets.
+  There is no method that outperforms the rest in all datasets.
   Based on their slightly better performance on the mock datasets, their \textit{de novo} error correcting nature and other previous studies~\cite{Nearing2018}, \ac{dada2} and Deblur appear to be the most reliable.
@@
   In order to obtain a quantitative picture of the differences between the inferred networks, we analyzed the distribution of common nodes and edges (Figure~\ref{fig:figure5} B and C) using UpSet plots~\cite{Lex}.
-  Only \ac{harmonies}, \ac{cozine},\ac{spieceasi}, \ac{spring}, propr, and \ac{sparcc} are used in the comparison since Pearson and Spearman add a large number of spurious edges as they are not intended for compositional datasets).
+  Only \ac{harmonies}, \ac{cozine}, \ac{spieceasi}, \ac{spring}, propr, and \ac{sparcc} are used in the comparison since Pearson and Spearman add a large number of spurious edges as they are not intended for compositional datasets).
   The results for the node intersections show that the networks have a large proportion of nodes in common ($33$ out of $68$ total unique nodes) and no network possesses any unique node.
@@
 
-  The systematic analyses performed in the previous sections clearly show that the choice of tools and parameters can have a big impact on the final co-occurrence network.
-  For some of these choices (e.g. for the \ac{ta} step), there is no clear metric to establish the best protocol.
+  The systematic analyses performed in the previous sections illustrate that the choice of tools and parameters can have a big impact on the final co-occurrence network.
+  For some of these choices (e.g. for the \ac{op} step), there is no clear metric to establish the best protocol.
   For other choices, the mock communities and synthetic data provide an opportunity to select the combinations of tools that yield more accurate and robust results.

### 2022-05-18 sections/supplementary.tex
Change: Update the tables and figures for draft v2
Change: Edit the supplementary section for draft v2

@@
 
+  \subsection*{Figures}
+
+    \begin{figure}[H]
+      \centering
+      \includegraphics[width=1.0\linewidth]{figure_s1.pdf}
+    \end{figure}
+    \begin{figure}[H]
+      \centering
+        \caption{
+          \textbf{The networks which were filtered during the \ac{op} step are less variable}
+          All combinations of inferred networks are shown as points on a PCA plot and shows the variation in the inferred networks in the first two principal axes (similar to Figure~\ref{fig:figure2}B).
+          Each point on the PCA plot represents a network inferred using different combinations of tools and parameters that are available in the \ac{micone} pipeline.
+          The points are colored by the tools or parameters used in the \ac{dc} step (A) and the \ac{op} step (B).
+          The \ac{dc} step does not seem to have any correlation with the similarity of networks on the PCA plot, but all the networks with filtering at the \ac{op} step are similar to each other.
+          This further confirms that the variability in the networks decreases upon filtering out the taxonomic entities at low abundance.
+        }
+      \label{fig:figure_s1}
+    \end{figure}
+    \FloatBarrier
+    \newpage
+
+    % TODO: Verify this about t-SNE
+    \begin{figure}[H]
+      \centering
+      \includegraphics[width=1.0\linewidth]{figure_s2.pdf}
+    \end{figure}
+    \begin{figure}[H]
+      \centering
+        \caption{
+          \textbf{The t-SNE plot of all the inferred networks clusters the networks based on the taxonomy reference database used}.
+          Each point on the t-SNE plot represents a network inferred using different combinations of tools and parameters that are available in the \ac{micone} pipeline.
+          The points are colored by the tools and parameters used in \ac{dc} step (A), \ac{ta} step (B), \ac{op} step (C) and \ac{ni} step (D).
+          The separation of the points based on taxonomy reference database shows that the points might cluster based on reference database in high-dimensional space.
+        }
+      \label{fig:figure_s2}
+    \end{figure}
+    \FloatBarrier
+    \newpage
+
+    \begin{figure}[H]
+      \centering
+      \includegraphics[width=1.0\linewidth]{figure_s3.pdf}
+    \end{figure}
+    \begin{figure}[H]
+      \centering
+        \caption{
+          \textbf{The UniFrac distance between the 1000 most abundant representative sequences is higher than that when all sequences are considered}.
+          Each value is the average UniFrac distance between the reference sequences generated by the various methods in the \ac{dc} step (similar to Figure~\ref{fig:figure3}).
+          There is an increase in both weighted and unweighted UniFrac distances compared to when all the representative sequences are considered.
+          This shows that the 1000 most abundant representative sequences generated by the methods are not as similar to each other.
+          And since the weighted UniFrac is much smaller than the unweighted UniFrac distance, we can conclude that those reference sequences that are present in the middle of the abundance distribution are dissimilar.
+        }
+      \label{fig:figure_s3}
+    \end{figure}
+    \FloatBarrier
+    \newpage
+
+    \begin{figure}[H]
+      \centering
+      \includegraphics[width=1.0\linewidth]{figure_s4.pdf}
+    \end{figure}
+    \begin{figure}[H]
+      \centering
+        \caption{
+          \textbf{The weighted and unweighted UniFrac distances between representative sequences generated using remove bimera and uchime for each denoising method are low.}
+          The two chimera checking methods, uchime and remove bimera, produce similar outputs.
+          With the exception of de novo and open reference under the unweighted UniFrac metric, all the other methods have low dissimilarity.
+          This is especially true for the \ac{dada2} and Deblur methods which are the recommended denoising methods in the \ac{micone} pipeline.
+          Therefore, remove bimera is recommended as the default chimera method if one is using \ac{dada2} and uchime-denovo when one is using Deblur, since these methods were developed for these respective algorithms (\ac{qiime2} uses uchime-denovo in the Deblur workflow).
+        }
+      \label{fig:figure_s4}
+    \end{figure}
+    \FloatBarrier
+    \newpage
+
+    \begin{figure}[H]
+      \centering
+      \includegraphics[width=1.0\linewidth]{figure_s5.pdf}
+    \end{figure}
+    \begin{figure}[H]
+      \centering
+        \caption{
+          \textbf{The pairwise comparison of assignments generated using different databases for all representative sequences has a higher proportion of mismatches}.
+          The comparison made here is similar to Figure~\ref{fig:figure4}B, but instead of the top 100 taxonomic entities (by abundance), all the assignments from one database are matched with those from the other two databases.
+          Higher percentage of mismatches implies that the matching of the taxonomies in the more abundant sequences (top 100) are more consistent.
+        }
+      \label{fig:figure_s5}
+    \end{figure}
+    \FloatBarrier
+    \newpage
+
+  \begin{figure}[H]
+    \centering
+    \includegraphics[width=1.0\linewidth]{figure_s6.pdf}
+  \end{figure}
+  \begin{figure}[H]
+    \centering
+      \caption{
+        \textbf{The precision and sensitivity of the inferred networks on the NorTA synthetic interaction data when Pearson and Spearman are included in the consensus calculation}.
+        The ``I'' prefix indicates the individual network inference methods and the ``X'' prefix denotes the different consensus based methods.
+        The different consensus based methods used are: pvalue merging (PM), scaled-sum (SS) and simple voting (SV) method.
+        With the addition of Pearson and Spearman in the consensus calculation we observe a decrease in the average precision scores of both the consensus algorithms (compared to Figure~\ref{fig:figure6}.
+        The overall best precision was still consistently obtained by the scaled-sum method (0.985, 0.963 and 1.000), but the performance was not stable to an increase in the parameter value.
+        The simple voting method when using presence of edges in all inferred networks as a requirement (parameter value 1.000), again outperforms \ac{spieceasi} on average precision (0.961).
+      }
+    \label{fig:figure_s6}
+  \end{figure}
+  \FloatBarrier
+  \newpage
+
+  \begin{figure}[H]
+    \centering
+    \includegraphics[width=1.0\linewidth]{figure_s7.pdf}
+  \end{figure}
+  \begin{figure}[H]
+    \centering
+      \caption{
+        \textbf{The precision and sensitivity of the inferred networks on the seqtime synthetic interaction data.}
+        The ``I'' prefix indicates the individual network inference methods and the ``X'' prefix denotes the different consensus based methods.
+        The different consensus based methods used are: pvalue merging (PM), scaled-sum (SS) and simple voting (SV) method.
+        Pearson and Spearman methods are not used in the calculation of the consensus.
+        Among all the independent network inference methods, \ac{spieceasi} has the best average precision (0.624), but the overall best precision was consistently obtained by the scaled-sum method (0.688, 0.820 and 1.000).
+        The simple voting method when using presence of edges in all inferred networks as a requirement (parameter value 1.000), also outperforms \ac{spieceasi} on average precision (0.692).
+        These results show that the scaled-sum method is not only much better suited for inferring robust and accurate interactions from count data regardless of the distributions and topologies.
+        But, it is also capable of accurately extracting real interactions (gLV interactions) from count data.
+      }
+    \label{fig:figure_s7}
+  \end{figure}
+  \FloatBarrier
+  \newpage
+
   \subsection*{Processing the FMT data}
@@
     \subsubsection*{Data download and pre-processing}
-    The main dataset used in this study contained stool samples (healthy and autistic individuals) from a fecal microbiome transplant dataset~\cite{Kang2017}.
-    The data containing the 16S sequencing reads (V4 region) was downloaded from Qiita~\cite{qiita} (study id: 10532).
-    Only runs 2, 3, and 4 were used for the subsequent analysis because these runs were paired-end sequencing data whereas run 1 was single-end.
-    The sample metadata was updated so as to contain only bmi, sex, height, weight and experimental group.
-    This was necessary as some of the network inference algorithms (like \ac{mldm}) required information about envioronmental heterogeneity.
-    However, these results were not included in the current analyses.
+    The main biological dataset used in this study was the collection of 16S rRNA sequencing reads from stool samples (healthy and autistic individuals) for a fecal microbiome transplant study~\cite{Kang2017}.
+    The data containing the 16S sequencing reads (V4 region) was downloaded from Qiita~\cite{qiita} (study ID: 10532).
+    Only runs 2, 3, and 4 were used for the subsequent analysis as these runs consisted of paired-end sequencing data and run 1 contained single-end data.
+    The sample metadata was updated to contain only BMI, sex, height, weight and experimental group.
+    This was necessary as two of the network inference algorithms (\ac{mldm} and FlashWeave) required information about environmental heterogeneity.
+    However, these environmental correlations were not included in the current analyses.
 
@@
     % TODO: Store pipeline results
-    The results of the pipeline execution are stored on Zenodo[REF] at ...
+    The results of the pipeline execution for reproducing the analyses in the manuscript are stored on Zenodo[REF] at ...
 
@@
     The mock datasets, mock4, mock12 and mock16 used for this study, were obtained from mockrobiota~\cite{Bokulich2016}.
-    Mock 4 is a mock community composed of 21 bacterial strains represented in equal abundances in two replicate samples, and the same strains represented in uneven abundances in two replicate samples.
+    Mock 4 is a mock community composed of 21 bacterial strains represented in equal abundances in two replicate samples, and the same strains represented in uneven abundances in two other replicate samples.
     Mock 12 is composed of 27 bacterial strains containing closely related taxa, the members of which were chosen in part for their well-separated 16S rRNA gene sequences.
@@
     % TODO: Store pipeline results
-    The results of the pipeline execution are stored on Zenodo[REF] at ...
+    The results of the pipeline execution for reproducing the analyses in the manuscript are stored on Zenodo at ...
 
@@
     % TODO: Store pipeline results
-    The results of the pipeline execution are stored on Zenodo[REF] at ...
-
-  \subsection*{Effects of the \ac{dc} and \ac{op} steps on network variability}
-
-    Figure~\ref{fig:figure_s1} describes the variation in the inferred networks in the first two principal axes and is similar to Figure~\ref{fig:figure2}B.
-    The \ac{dc} step does not seem to have any correlation with the similarity of networks on the PCA plot, but all the networks with filtering at the \ac{op} step are similar to each other.
-    This further confirms that the variability in the networks decreases upon filtering out the taxonomic entities at low abundance.
-
-    \begin{figure}[H]
-      \centering
-      \includegraphics[width=1.0\linewidth]{figure_s1.pdf}
-    \end{figure}
-    \begin{figure}[H]
-      \centering
-        \caption{
-          \textbf{The variation in the inferred networks in the first two principal axes}
-          All combinations of inferred networks are shown as points on a PCA plot.
-          Each point on the PCA plot represents a network inferred using different combinations of tools and parameters that are available in the \ac{micone} pipeline.
-          The points are colored by the tools or parameters used in the \ac{dc} step (A) and the \ac{op} step (B).
-        }
-      \label{fig:figure_s1}
-    \end{figure}
-    \FloatBarrier
-    \newpage
-
-    % TODO: Verify this about t-SNE
-    \begin{figure}[H]
-      \centering
-      \includegraphics[width=1.0\linewidth]{figure_s2.pdf}
-    \end{figure}
-    \begin{figure}[H]
-      \centering
-        \caption{
-          \textbf{The t-SNE plot of the networks inferred by various methods}.
-          The points are colored by the tools and parameters used in \ac{dc} step (A), \ac{ta} step (B), \ac{op} step (C) and \ac{ni} step (D).
-          The separation of the points based on taxonomy reference database shows that the points might cluster based on reference database in high-dimensional space.
-        }
-      \label{fig:figure_s2}
-    \end{figure}
-    \FloatBarrier
-    \newpage
-
-  \subsection*{Similarity of high abundance reference sequences}
-
-    Figure~\ref{fig:figure_s3} is similar to Figure~\ref{fig:figure3} shows the average UniFrac distance between reference sequences generated by the various methods in the \ac{dc} step.
-    In Figure~\ref{fig:figure_s3} however, we use only the top 1000 representative sequences (by abundance) for this calculation.
-    We observe that both the weighted and unweighted UniFrac distances are increased compared to the figure with the all the representative sequences.
-    This shows that the 1000 most abundant representative sequences generated by the methods are not similar to each other.
-    And since the weighted UniFrac is much smaller than the unweighted UniFrac distance, we can conclude that those reference sequences that are present in the middle of the abundance distribution are dissimilar.
-
-    \begin{figure}[H]
-      \centering
-      \includegraphics[width=1.0\linewidth]{figure_s3.pdf}
-    \end{figure}
-    \begin{figure}[H]
-      \centering
-        \caption{
-          \textbf{The UniFrac distance between the 1000 most abundant representative sequences generated by various method in the \ac{dc} step}
-        }
-      \label{fig:figure_s3}
-    \end{figure}
-    \FloatBarrier
-    \newpage
-
-    We find that the two chimera checking methods, uchime and remove bimera, produce similar outputs.
-    This can be verified by comparing the output (representative sequence distributions) of the two chimera checking methods for each denoising method (Figure~\ref{fig:figure_s4}).
-    We find that with the exception of de novo and open-reference under the unweighted UniFrac metric, all the other methods have low dissimilarity.
-    This is especially true for the \ac{dada2} and Deblur methods which are the recommended denoising methods in the \ac{micone} pipeline.
-    Therefore, we recommended remove bimera as the default chimera method if one is using \ac{dada2} and uchime-denovo when one is using Deblur, since these methods were developed for these respective algorithms (\ac{qiime2} uses uchime-denovo in the Deblur workflow).
-
-    \begin{figure}[H]
-      \centering
-      \includegraphics[width=1.0\linewidth]{figure_s4.pdf}
-    \end{figure}
-    \begin{figure}[H]
-      \centering
-        \caption{
-          \textbf{The weighted and unweighted UniFrac distance between representative sequences generated using remove bimera and uchime for each denoising method.}
-        }
-      \label{fig:figure_s4}
-    \end{figure}
-    \FloatBarrier
-    \newpage
-
-  \subsection*{Mismatches in all assignments at the \ac{ta} step}
-
-    Figure~\ref{fig:figure_s5} is similar to Figure~\ref{fig:figure4}B, but instead of the matching only the top 100 taxonomic entities (by abundance), we match all the assignments from one database with those from the other two databases.
-    We observe that the percentage of mismatches are indeed higher when considering all the assignments, implying that matching of the taxonomies in the more abundant sequences are more consistent.
-
-    \begin{figure}[H]
-      \centering
-      \includegraphics[width=1.0\linewidth]{figure_s5.pdf}
-    \end{figure}
-    \begin{figure}[H]
-      \centering
-        \caption{
-          \textbf{The pairwise comparison of assignments generated using different databases for each representative sequence}
-        }
-      \label{fig:figure_s5}
-    \end{figure}
-    \FloatBarrier
-    \newpage
+    The results of the pipeline execution for reproducing the analyses in the manuscript are stored on Zenodo at ...
 
@@
   In Table~\ref{tab:network_metrics} we show various global network metrics calculated for each tool in the pipeline.
-  All the networks that make use of a particular tool are grouped together and the average metrics are calculated for each group.
-  All the metrics were calculated using the \texttt{networkx} Python package~\cite{hagbergExploringNetworkStructure2008}.
+  All the networks that make use of a particular tool are grouped together, and the following average metrics are calculated for each group:
   \begin{enumerate}
     \item The average shortest path length describes the average of all the shortest paths in the graph.
-No number is reported if the graph is not connected, thereforce, none of the networks that make use of \ac{harmonies}, \ac{cozine}, \ac{spring}, \ac{spieceasi} and Pearson are connected.
+No number is reported if the graph is not connected, therefore, the results indicate that none of the networks that make use of \ac{harmonies}, \ac{cozine}, \ac{spring}, \ac{spieceasi} and Pearson are connected.
     \item The average clustering is the average clustering coefficient of the graph.
@@
 We can observe that the networks that use correlation-based methods have the highest values while the direct-association based methods have the lowest.
-    \item The number of connected components are the highest for the direct-association based methods and the lowest for the correlation-based methods.
+    \item The number of connected components is the highest for the direct-association based methods and the lowest for the correlation-based methods.
 In the case of propr, all the networks have only one giant component.
-    \item Modularity metric is the average modularity over all partitions in a graph calculated using a label propagation algorithm.
-Positive values imply that there are more edges between vertices of the same type than we would expect by chance, and negative means that there are less.
+    \item The modularity metric is the modularity over all partitions in a graph calculated using a label propagation algorithm~\cite{cordascoCommunityDetectionSemisynchronous2010}.
+Positive values imply that there are more edges between vertices of the same type than we would expect by chance, and negative implies that there are less.
 The networks inferred by \ac{mldm} report very few edges, and skew the average modularity scores.
+This could also be an artifact of incomplete converge of the \ac{mldm} algorithm for some combinations.
     \item Node connectivity refers to the minimum number of nodes that must be removed from the graph to make it disconnected.
@@
   \end{enumerate}
-
-  \begin{landscape}
-  \begin{table}[H]
-    \centering
-    \small
-    \begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{2cm}|p{1.5cm}|p{1.5cm}|p{1.5cm}|p{1.5cm}|p{1.5cm}|p{1.5cm}|p{1.5cm}|}
-  \hline
-  DC & CC            & TA                & OP          & NI         & Average shortest path length & Average clustering & No.
-of connected components & Modularity & Node connectivity & Degree assortativity coefficient \\ \hline
-  DB & -             & -                 & -           & -          & 2.066                        & 0.265              & 27.088                      & 0.387      & 1.868             & 0.070                            \\
-  OR & -             & -                 & -           & -          & 1.904                        & 0.273              & 24.868                      & -17.060    & 1.877             & 0.099                            \\
-  D2 & -             & -                 & -           & -          & 2.263                        & 0.267              & 28.228                      & 0.326      & 1.789             & 0.078                            \\
-  DN & -             & -                 & -           & -          & 2.065                        & 0.272              & 25.132                      & 0.302      & 1.754             & 0.093                            \\
-  CR & -             & -                 & -           & -          & 1.937                        & 0.276              & 24.289                      & -1.865     & 2.079             & 0.097                            \\
-  -  & remove bimera & -                 & -           & -          & 2.070                        & 0.272              & 25.968                      & -7.439     & 1.863             & 0.086                            \\
-  -  & uchime        & -                 & -           & -          & 2.066                        & 0.269              & 25.874                      & 0.275      & 1.884             & 0.089                            \\
-  -  & -             & NCBI              & -           & -          & 2.218                        & 0.262              & 22.784                      & -3.683     & 1.495             & 0.100                            \\
-  -  & -             & SILVA             & -           & -          & 1.872                        & 0.275              & 30.716                      & 0.415      & 2.674             & 0.074                            \\
-  -  & -             & GG                & -           & -          & 2.086                        & 0.274              & 24.263                      & -7.478     & 1.453             & 0.088                            \\
-  -  & -             & -                 & On          & -          & 1.975                        & 0.252              & 18.493                      & -7.194     & 1.880             & 0.073                            \\
-  -  & -             & -                 & Off         & -          & 2.193                        & 0.291              & 34.174                      & 0.431      & 1.867             & 0.104                            \\
-  -  & -             & -                 & -           & propr      & 1.650                        & 0.528              & 1.000                       & 0.000      & 10.583            & 0.036                            \\
-  -  & -             & -                 & -           & harmonies  &                              & 0.108              & 82.633                      & 0.757      & 0.000             & 0.340                            \\
-  -  & -             & -                 & -           & flashweave & 3.878                        & 0.079              & 1.800                       & 0.458      & 0.383             & -0.032                           \\
-  -  & -             & -                 & -           & cozine     &                              & 0.090              & 82.167                      & 0.694      & 0.000             & -0.072                           \\
-  -  & -             & -                 & -           & spring     &                              & 0.092              & 11.767                      & 0.566      & 0.000             & -0.035                           \\
-  -  & -             & -                 & -           & sparcc     & 1.747                        & 0.525              & 1.033                       & 0.009      & 5.500             & -0.011                           \\
-  -  & -             & -                 & -           & spieceasi  &                              & 0.076              & 39.383                      & 0.717      & 0.000             & -0.031                           \\
-  -  & -             & -                 & -           & pearson    &                              & 0.457              & 23.883                      & 0.551      & 0.000             & 0.478                            \\
-  -  & -             & -                 & -           & spearman   & 1.875                        & 0.551              & 1.450                       & 0.044      & 1.117             & 0.170                            \\
-  -  & -             & -                 & -           & mldm       & 2.760                        & 0.130              & 2.267                       & -75.648    & 0.433             & -0.028                           \\ \hline
-  \end{tabular}
-  \caption{
-    \textbf{Table of global network metrics for all combinations of networks.}
-    The metrics are averaged for each tool in every step of the pipeline, implying that all combinations of networks that use a particular tool are grouped and averaged and their metrics are reported.
-  }
-  \label{tab:network_metrics}
-  \end{table}
-  \end{landscape}
+  All the metrics were calculated using the \texttt{networkx} Python package~\cite{hagbergExploringNetworkStructure2008}.
 
@@
 
-  Fisher~\cite{fisher_224a_1948} proposed that for $k$ independent p-values, each generated by $k$ different methods and denoted by $\bar{P}^i$, the following will hold true for the statistic $\Psi$:
+  Fisher~\cite{fisher_224a_1948} proposed that for $k$ independent p-values, each generated by $k$ different methods and denoted by $\bar{P}^i$ (notations are same as used in the "Consensus network and p-value merging" subsection of the Methods), the following will hold true for the statistic $\Psi$:
   \begin{equation*}
@@
   \begin{equation}
-    \begin{aligned}
+    \begin{aligned*}
         & \hat{P}_j = 1.0 - \Phi_{2f}\left( \psi / c \right) \\
         \text{where},~ &\psi = -2 \sum_{i=1}^k \log(\bar{P}^i_j) ~~~\text{and}~~~ \Phi_{2f} = \mathrm{CDF}\left( \chi^2_{2f} \right)
-    \end{aligned}
+    \end{aligned*}
     \label{eqn:suppl_pvalue-combined}
@@
 
-  The p-value merging and consensus method in \ac{micone} (refer Documentation) uses Equation~\ref{eqn:suppl_covariance-pvalues} to estimate the covariance of the p-values and Equation~\ref{eqn:suppl_pvalue-combined} to merge the p-values (obtained from bootstrapping) from the different correlation methods.
+  The p-value merging and consensus method in \ac{micone} (refer Methods) uses Equation~\ref{eqn:suppl_covariance-pvalues} to estimate the covariance of the p-values and Equation~\ref{eqn:suppl_pvalue-combined} to merge the p-values (obtained from bootstrapping) from the different correlation methods.
   Note that we do not use Pearson and Spearman methods in the p-value merging step and these algorithms are only used for demonstration and comparison.
-  The combined p-values are used to threshold for significance during the consensus network step.
-
-  \subsection*{Performance of alternate consensus network on NorTA data}
-
-    In Figure~\ref{fig:figure_s6} we calculate the consensus network by also taking into account the networks inferred by the Pearson and Spearman methods.
-    We see that the consensus networks in this case do not perform as well as those networks that did not incorporate these two algorithms.
-    However, we still observe that the consensus network at some parameter values still outperforms the best network inference method (\ac{spieceasi}).
-
-    % TODO: Rerun this figure
-    \begin{figure}[H]
-      \centering
-      \includegraphics[width=1.0\linewidth]{figure_s6.pdf}
-    \end{figure}
-    \begin{figure}[H]
-      \centering
-        \caption{
-          \textbf{The precision and sensitivity of the inferred networks on the NorTA synthetic interaction data with Pearson and Spearman added to the consensus.}
-            The ``I'' prefix indicates the individual network inference methods and the ``X'' prefix denotes the different consensus based methods.
-            The different consensus based methods used are: pvalue merging (PM), scaled-sum (SS) and simple voting (SV) method.
-            With the addition of Pearson and Spearman in the consensus calculation we observe a decrease in the average precision scores of both the consensus algorithms.
-            The overall best precision was still consistently obtained by the scaled-sum method (0.985, 0.963 and 1.000), but the performance was not stable to changes in parameter value.
-            The simple voting method when using presence of edges in all inferred networks as a requirement (parameter value 1.000), again outperforms \ac{spieceasi} on average precision (0.961).
-        }
-      \label{fig:figure_s6}
-    \end{figure}
-    \FloatBarrier
-    \newpage
-
-  \subsection*{Performance of network inference algorithms on seqtime data}
-
-    In Figure~\ref{fig:figure_s7} we perform the same analysis we did with the NorTA data using the seqtime data.
-    We observe that the best performing network inference algorithm is \ac{spieceasi}, but the consensus algorithms, both the scaled-sum and simple voting methods outperform it in terms of precision.
-    These results show that the scaled-sum method is much better suited for inferring robust and accurate interactions from count data regardless of the distributions, topologies and number of samples in the data.
-
-    \begin{figure}[H]
-      \centering
-      \includegraphics[width=1.0\linewidth]{figure_s7.pdf}
-    \end{figure}
-    \begin{figure}[H]
-      \centering
-        \caption{
-          \textbf{The precision and sensitivity of the inferred networks on the seqtime synthetic interaction data.}
-          The ``I'' prefix indicates the individual network inference methods and the ``X'' prefix denotes the different consensus based methods.
-          The different consensus based methods used are: pvalue merging (PM), scaled-sum (SS) and simple voting (SV) method.
-          Pearson and Spearman methods are not used in the calculation of the consensus.
-          Among all the independent network inference methods, \ac{spieceasi} has the best average precision (0.624), but the overall best precision was consistently obtained by the scaled-sum method (0.688, 0.820 and 1.000).
-          The simple voting method when using presence of edges in all inferred networks as a requirement (parameter value 1.000), also outperforms \ac{spieceasi} on average precision (0.692).
-        }
-      \label{fig:figure_s7}
-    \end{figure}
-    \FloatBarrier
-    \newpage
+  The combined p-values are used to threshold for significance the correlation-based networks during the consensus network step.
 
@@
     The default format \ac{micone} uses for storing the network files is the JSON (JavaScript Object Notation) format.
-    The custom JSON schema we have designed is able to store all network related information from nodes and links to metadata related to links and datasets.
+    The custom JSON schema we have designed is able to store all network related information pertaining to nodes, links and the metadata related to the links and datasets.
     Additionally, \ac{micone} also supports exporting of networks into a variety of other formats such as edge lists, .gml and cytoscape formats.
-    Since, we make use of \texttt{networkx}~\cite{hagbergExploringNetworkStructure2008} for the export functionality, networks can be exported to all supported formats but the corresponding metadata might not be exported appropriately.
+    Since, we make use of \texttt{networkx}~\cite{hagbergExploringNetworkStructure2008} for the export functionality, networks can be exported to all formats supported by the package.
+    However, not all the corresponding metadata will be exported appropriately, since most formats do not support this additional metadata.
     The details of the format and information about importing/exporting it and other network formats can be found in the \ac{micone} documentation.

### 2022-08-25 sections/abstract.tex
Change: Update abstract and results

@@
 \section*{Importance}
-  To understand and control the mechanisms that determine the structure and function of microbial communities, it is important to map the interrelationships between its constituent microbial species.
+  Mapping the interrelationships between the species in a microbial community is essential to understand and control their structure and function.
   The surge in the high-throughput sequencing of microbial communities has led to the creation of thousands of datasets containing information about microbial abundances.
-  These abundances can be transformed into networks of co-occurrences across multiple samples, providing a glimpse into the structure of microbiomes.
+  These abundances can be transformed into co-occurrence networks, providing a glimpse into the associations within microbiomes.
   However, processing these datasets to obtain co-occurrence information relies on several complex steps, each of which involves numerous choices of tools and corresponding parameters.
   These multiple options pose questions about the accuracy and uniqueness of the inferred networks.
-  In this study, we address this workflow and provide a systematic analysis of how these choices of tools and parameters affect the final network, and on how to select those that are most appropriate for a particular dataset.
+  In this study, we address this workflow and provide a systematic analysis of how these choices of tools affect the final network, and guidelines on appropriate tool selection for a particular dataset.
   We also develop a consensus network algorithm that helps generate more robust and accurate occurrence networks based on benchmarks synthetic datasets.

### 2022-08-25 sections/coverpage.tex
Change: Update abstract and results

--- 2020-05-29 sections/coverpage.tex
@@
   \vspace{1.5 cm}
-  Dileep Kishore\textsuperscript{1},
-  Gabriel Birzu\textsuperscript{2},
-  Zhenjun Hu\textsuperscript{1},
-  Charles DeLisi\textsuperscript{1,3},
-  Kirill Korolev\textsuperscript{$\dagger$1,2},\\
-  Daniel Segr\`{e}\textsuperscript{$\dagger$1,3,4}\\
+  Dileep Kishore\textsuperscript{a,b},
+  Gabriel Birzu\textsuperscript{c,f},
+  Zhenjun Hu\textsuperscript{a},
+  Charles DeLisi\textsuperscript{a,c},
+  Kirill S. Korolev\textsuperscript{a,b,c}\#,\\
+  Daniel Segr\`{e}\textsuperscript{a,b,d,e}\#\\
   \vspace{1cm}
-  \textsuperscript{1}Bioinformatics Program, Boston University, Boston, Massachusetts, USA\\
-  \textsuperscript{2}Department of Physics, Boston University, Boston, Massachusetts, USA\\
-  \textsuperscript{3}Department of Biomedical Engineering, Boston University, Boston, Massachusetts, USA\\
-  \textsuperscript{4}Department of Biology, Boston University, Boston, Massachusetts, USA\\
-  \textsuperscript{$\dagger$}Correspondence should be sent to \href{mailto:korolev@bu.edu}{korolev@bu.edu} or \href{mailto:dsegre@bu.edu}{dsegre@bu.edu}\\
+  \textsuperscript{a}Bioinformatics Program, Boston University, Boston, Massachusetts, USA\\
+  \textsuperscript{b}Biological Design Center, Boston University, Boston, Massachusetts, USA\\
+  \textsuperscript{c}Department of Physics, Boston University, Boston, Massachusetts, USA\\
+  \textsuperscript{d}Department of Biomedical Engineering, Boston University, Boston, Massachusetts, USA\\
+  \textsuperscript{e}Department of Biology, Boston University, Boston, Massachusetts, USA\\
+  \textsuperscript{f}Department of Applied Physics, Stanford University, Stanford, California, USA\\
+  \textsuperscript{$\#$}Correspondence should be sent to \href{mailto:korolev@bu.edu}{korolev@bu.edu} or \href{mailto:dsegre@bu.edu}{dsegre@bu.edu}\\
+
+  \vspace{2cm}
+  % TODO: Add word count
+  Abstract word count: 244\\
+  Text word count: \\
 

### 2022-08-29 sections/discussion.tex
Change: Update Results section for draft v3
Also, changes the figure order (code remains unchanged)

@@
 
-  \subsection*{Why \ac{micone}?}
+  % \subsection*{Why \ac{micone}?}
 
-  Co-occurrence associations in microbial communities help identify important interactions that drive microbial community structure and organization.
+  Co-occurrence associations in microbial communities help identify important interdependencies that may drive microbial community structure and organization.
   A myriad of tools and methods have been developed for different parts of the workflow for inference of co-occurrence networks from 16S rRNA data.
-  Our analysis shows that networks generated using different combinations of tools and approaches are significantly different from each other, highlighting the need for a clear assessment of the source of variability and of tools that provide the most robust and accurate results.
-  Our newly developed integrated software for the inference of co-occurrence networks from 16S rRNA data, \ac{micone}, is a freely customizable and user-friendly pipeline that allows users to easily test combinations of tools and to compare networks generated by multiple possible choices.
+  Our analysis shows that networks generated using different combinations of tools and approaches are significantly different from each other, highlighting the need for a clear assessment of the source of variability and for tools that provide the most robust and accurate results.
+  Our newly developed integrated software for the inference of co-occurrence networks from 16S rRNA data, \ac{micone}, is a freely customizable and user-friendly pipeline that allows users to compare networks generated by multiple possible combinations of tools and parameters.
   Importantly, in addition to revisiting the test cases presented in this work, users will be able to explore the effect of various tool combinations on their own datasets of interest.
-  The \ac{micone} pipeline is built in a modular fashion; its plug-and-play architecture will make it possible for users to add new tools and steps, either using existing packages that were not examined in the present work or those developed in the future.
-  The easy-to-use nature, the configurability and modular nature of the \ac{micone} package enables users to start and end the pipeline at any point in the workflow, enabling users to run parts of the pipeline in isolation.
-  This would allow for the user to bring in their own data at any step of the pipeline and explore the effects of the desired downstream data processing tools.
+  The \ac{micone} pipeline is built in a modular fashion; its plug-and-play architecture enables users to add new tools and steps, either using existing packages that were not examined in the present work or those developed in the future.
+  The easy-to-use nature, the configurability and modular nature of the \ac{micone} package enables users to start and end the pipeline at any point in the workflow, and to run parts of the pipeline in isolation.
   The \ac{micone} Python package provides functions and methods to perform detailed analysis of the count matrices and the co-occurrence networks.
-  Additionally, the inferred networks by default are exported to a custom JSON format (see Supplementary), but are easily exported to Cytoscape~\cite{shannonCytoscapeSoftwareEnvironment2003}, GML~\cite{himsoltGMLPortableGraph2010}, and many other popular formats via the \ac{micone} Python package.
+  Additionally, the inferred networks are exported to a custom JSON format (see Supplementary), but can also be exported to Cytoscape~\cite{shannonCytoscapeSoftwareEnvironment2003}, GML~\cite{himsoltGMLPortableGraph2010}, and many other popular formats via the \ac{micone} Python package.
 
-  There have been several tools/workflows such as \ac{qiime2}~\cite{bolyenReproducibleInteractiveScalable2019} and NetCoMi~\cite{peschelNetCoMiNetworkConstruction2020} that have been developed for analysis of 16S sequencing data and to infer networks from these datasets.
-  However, there has been no single tool that completely integrates the process of inferring microbial interaction networks from 16S sequencing reads.
-  \ac{micone} is unique as it offers this functionality packaged in a simple workflow that can be run locally, on the compute cluster or the cloud.
-  For an end user who is interested in identifying the microbial interactions present in a specific sequencing dataset, \ac{micone} offers a fast and reproducible way to compare and test novel hypotheses related to interactions.
+  While several tools/workflows such as \ac{qiime2}~\cite{bolyenReproducibleInteractiveScalable2019} and NetCoMi~\cite{peschelNetCoMiNetworkConstruction2020} can be used to generate co-occurrence networks from 16S sequencing data, no single tool exist that integrates the complete process of inferring microbial interaction networks from 16S sequencing reads.
+  \ac{micone} is unique as it offers this functionality packaged in a workflow that can be run locally, on the compute cluster or the cloud.
+  For an end user who is interested in identifying the microbial interdependencies present in a specific sequencing dataset, \ac{micone} offers a fast and reproducible way to compare and test novel hypotheses.
 
-  \subsection*{The default pipeline and recommended tools}
+  % \subsection*{The default pipeline and recommended tools}
 
@@
   Additionally, it is worth pointing out some additional more specific conclusions stemming from the individual steps of our analysis.
-  The different denoising/clustering methods differ mostly in their identification of sequences that are in lower abundances.
+  The different denoising/clustering methods differ mostly in their identification of sequences that are in low abundances.
   Hence, they do not have much of an impact on the inferred co-occurrence networks when the sequences of low abundance are removed (Figure~\ref{fig:figure_s1}).
@@
 
-  The choice of taxonomy database was found to be the most important factor in the inference of microbial co-occurrence networks, contributing $\sim65.4\%$ of the total variance.
+  The choice of taxonomy database was found to be the most important factor in the inference of microbial co-occurrence networks, contributing $65.4\%$ of the total variance.
   The frequent changes in the taxonomy nomenclature coupled with the frequency of updates to the various 16S reference databases create inherent differences \cite{Balvociute2017} in taxonomy hierarchies in these databases.
@@
   The default reference database in the pipeline is the \ac{gg} reference database along with the ``Naive Bayes'' classifier as the query tool.
-  The reason for our choice stems from the popularity of the \ac{gg} database in taxonomic studies, which would enable easy comparison across datasets.
+  The reason for our choice stems from the popularity of the \ac{gg} database [https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6440677/] in taxonomic studies, which would enable easy comparison across datasets.
   % TODO: Add ref: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6440677/
@@
   The consensus algorithm not only leads to the creation of a network that has a higher number of true positives, it also decreases the variability in the networks due to choosing a different tool or parameter in any of the previous steps of the pipeline (Figure~\ref{fig:figure7}B).
+  This provides the means to obtain a short list of associations, each of which, have a high probability of being present in the real interaction network
+  This small set of associations could be further tested and validated experimentally to verify their plausibility.
 
-  Exploring the effects of these combinations of methods on the resultant networks is difficult and inconvenient since different tools differ in their input and output formats and require interconversions between the various formats.
-  The pipeline facilitates this comparative exploration by providing a variety of modules for interconversion between various formats, and by allowing for easy incorporation of new tools as modules.
-  Since there exists no ``best method'' for the various steps of 16S data analysis, especially for microbial network inference, \ac{micone} is intended to help researchers to identify the methods and algorithms that are most suitable for their datasets in an easy-to-use and reproducible manner.
-  We envision that \ac{micone}, and the underlying tools and databases that help process amplicon sequencing data into co-occurrence networks, will be increasingly useful towards building large comparative analyses across studies.
-  By having a unified transparent tool to compute networks, it will be possible to reprocess available 16S datasets to obtain networks that are directly comparable to each other.
-  Furthermore, even in the analysis of published networks across studies and processing methods, \ac{micone} could help understand underlying biases of each network, which could in turn be taken into account when making cross-study comparisons.
-
-  \subsection*{Future directions}
+  % \subsection*{Future directions}
 
@@
   Incorporating datasets from a broad spectrum of biomes with varying microbial distributions into \ac{micone} will likely increase the robustness and generalizability of the results from these analyses.
+
   The network analyses in this study are primarily at the Genus level, wherein the lowest resolution of a node is a Genus and if an entity cannot be resolved to the Genus level and next lowest taxonomic level is used (e.g.
@@
 
-  Additionally, the utilization of literature-based interactions~\cite{lima-mendezDeterminantsCommunityStructure2015a} or biological benchmark interaction data~\cite{sungGlobalMetabolicInteraction2017a} as additional sources for benchmarking
+  Additionally, the utilization of literature-based interactions~\cite{lima-mendezDeterminantsCommunityStructure2015a} or biological benchmark interaction data~\cite{sungGlobalMetabolicInteraction2017a} as additional sources for benchmarking.
   Furthermore, expanding the scope of the \ac{micone} pipeline to also include metagenomics sequencing data will enable an even broader access to a diverse range of datasets as well as domains of life that can be incorporated into the analysis.
@@
   However, the modular nature of the \ac{micone} pipeline enables easy addition of new tools and methods at any step of the workflow.
+  Since there exists no ``best method'' for the various steps of 16S data analysis, especially for microbial network inference, \ac{micone} is intended to help researchers to identify the methods and algorithms that are most suitable for their datasets in an easy-to-use and reproducible manner.
 
-  % TODO: Should we move the last paragraph of the previous section here?
-
-  % TODO: Talk health and disease or something more specific?
+  We envision that \ac{micone}, and the underlying tools and databases that help process amplicon sequencing data into co-occurrence networks, will be increasingly useful towards building large comparative analyses across studies.
+  Furthermore, in the re-analysis of published networks across studies and processing methods, \ac{micone} could help understand underlying biases imposed on networks due to the selected methods.
   In summary, we believe that \ac{micone} is a useful tool for the exploration of microbial networks generated from 16S data.
   It allows for rapid, configurable and reproducible inference of microbial networks and further for the formulation of hypotheses about the role of these interactions on community composition and stability.
+  Additionally, our pipeline is coupled to the MIND database and web interface for the organization and visualization of microbial networks, enabling easier comparisons across studies and datasets.

### 2022-08-29 sections/introduction.tex
Change: Update Results section for draft v3
Also, changes the figure order (code remains unchanged)

@@
   With the advancement in DNA sequencing technologies \cite{Narihiro2017} and data processing methods,  more information can be extracted from these microbial community samples than ever before.
-  In particular, high-throughput sequencing, including metagenomic sequencing and sequencing of 16S rRNA gene amplicons of microbial communities, has the potential to help detect, identify and quantify a large portion of the constitutive microorganisms of a microbiome \cite{Jovel2016,Lloyd-Price2016}.
+  In particular, high-throughput sequencing, including metagenomic sequencing and sequencing of 16S rRNA gene amplicons (hereafter referred to as 16S data) of microbial communities, has the potential to help detect, identify and quantify a large portion of the constitutive microorganisms of a microbiome \cite{Jovel2016,Lloyd-Price2016}.
   These advances have led to large-scale data collection efforts involving environmental (\acl{emp}) \cite{Thompson2017}, marine (Tara Oceans Project) \cite{Zhang2015} and human-associated microbiota (Human Microbiome Project) \cite{HumanMicrobiomeProjectConsortium2012}.
@@
   % TODO: Link or describe MIND here (?)
-  Our pipeline is coupled to an online integrative tool for the organization, visualization and analysis of inter-microbial networks.
+ Our pipeline is coupled to an online integrative tool for the organization, visualization and analysis of inter-microbial networks [MIND REF], called \ac{mind}, that is available at [link].
   In addition to making this tool freely available, we completed a systematic comparative analysis to determine which steps of the pipeline have the largest influence on the final network, and what choice seems to have the best agreement with the tested mock and synthetic datasets.

### 2022-08-29 sections/methods.tex
Change: Edit the supplementary section for draft v2
Change: Update Results section for draft v3
Also, changes the figure order (code remains unchanged)

@@
     Sequences from both control (212 samples including neurotypical and donors) and autistic (126 samples) patients were included in the analyses, but the groups were analyzed independently in the pipeline after the \ac{op} step.
+    All the network analyses in the study, unless explicitly mentioned, are performed indeed on the healthy and autistic samples in the fecal microbiome transplant study.
+    However, all other analysis (preceding the NI step) are performed together on all the samples in the fecal microbiome transplant study.
     The mock community 16S datasets are sequencing data obtained for artificially assembled collections of species in known proportions.
@@
   The \ac{micone} Python API provides several biom (\ac{otu} table) and network related functions and methods, enabling detailed comparison of counts tables and inferred networks if desired.
+  Exploring the effects of these combinations of methods on the resultant networks is difficult and inconvenient since different tools differ in their input and output formats and require interconversions between the various formats.
+  The pipeline facilitates this comparative exploration by providing a variety of modules for interconversion between various formats, and by allowing for easy incorporation of new tools as modules.
   It also contains helper functions that can help in parsing taxonomies and communicate with the \ac{ncbi} taxonomy database to query taxonomy by name or taxonomic IDs.
@@
   The pipeline incorporates \ac{gg} 13\_8~\cite{DeSantis2006}, SILVA 138~\cite{Quast2012} and the \ac{ncbi} (16S RefSeq as of Oct 2021)~\cite{Sayers2009} databases for taxonomy assignment.
-  These databases were downloaded and built using the RESCRIPt qiime2 plugin~\cite{iiRESCRIPtReproducibleSequence2021}.
+  SILVA and \ac{gg} are two popular 16S databases used for taxonomy identification and the \ac{ncbi} RefSeq nucleotide database contains 16S rRNA sequences as a part of two BioProjects - 33175 and 33317.
+  The three databases vastly differ in terms of their last update status - \ac{gg} was last updated on May 2013, SILVA was last updated on August 2020 at the time of writing and \ac{ncbi} is updated regularly as new sequences are curated.
+  These databases were downloaded and built using the RESCRIPt \ac{qiime2} plugin~\cite{iiRESCRIPtReproducibleSequence2021}.
   The Naive Bayes classifier and the \ac{ncbi} blast used as the query tools in this study were from the \ac{qiime2} package and the parameters used were the defaults of the package.
   The consensus algorithm used is the default method used by the classifiers in \ac{qiime2}.
+  The assignment between SILVA and the other two databases were originally significantly different ($40\%$ mismatch) even at the Phylum level.
+  However, this was fixed by making minor adjustments to the taxonomic names, such as changing Bacteroidota to Bacteroidetes in the SILVA Phylum assignments.
+  The full list of changes can be found in \texttt{figure4ab_data.py} in the data and scripts  repository.
   The list of tools used in this step, along with their modules and references are provided in Table~\ref{tab:micone_tools}.
@@
   These associations can be represented as a network, with nodes representing taxonomies of the microorganisms and edges representing the associations between them.
+  A threshold of 0.3 was set for the correlation-based methods (\ac{sparcc}, Propr, Spearman and Pearson) and a threshold of 0.01 was set for the other direct association methods.
+  After filtering out edges based on the threshold, the isolated nodes are removed.
+
   A null model is created by re-sampling and bootstrapping the counts matrix and recalculating the correlations.
@@
 
-  \subsection*{Consensus Network and p-value merging}
+  \subsection*{Consensus network and p-value merging}
 
@@
   \vspace{-5mm}
-  This section defines notation for the network variability analysis performed for Figure~\ref{fig:figure2}.
+  This section defines notation for the network variability analysis performed for Figure~\ref{fig:figure6}.
 
@@
 
-   The variance contributed by each step of the pipeline is calculated for every component in $\mathbf{C}$ matrix through ANOVA using the Python statsmodels~\cite{seaboldStatsmodelsEconometricStatistical2010} package and is shown in Figure~\ref{fig:figure2}A.
+   The variance contributed by each step of the pipeline is calculated for every component in $\mathbf{C}$ matrix through ANOVA using the Python statsmodels~\cite{seaboldStatsmodelsEconometricStatistical2010} package and is shown in Figure~\ref{fig:figure6}A.
   The total variance for the network is calculated by adding the variances for each connection.
-  The merged network table $\mathbf{\bar{N}}$ is used as the input to the PCA analysis to generate Figure~\ref{fig:figure2}B.
+  The merged network table $\mathbf{\bar{N}}$ is used as the input to the PCA analysis to generate Figure~\ref{fig:figure6}B.
 
@@
   The interaction matrix was used as the source of expected (true) interactions and the associations predicted using \ac{micone} were the source of predicted interactions.
-  Finally, for each dataset we evaluated the precision and sensitivity of the associations predicted by the individual network inference methods as well as the consensus (Figure \ref{fig:figure6}).
+  Finally, for each dataset we evaluated the precision and sensitivity of the associations predicted by the individual network inference methods as well as the consensus (Figure \ref{fig:figure5}).
 
@@
   \vspace{-5mm}
-  In order to compare the representative sequences generated by the various methods in the DC step, we employ both the weighted~\cite{Lozupone2007} (Figure~\ref{fig:figure3}A) and unweighted UniFrac method~\cite{Lozupone2005} (Figure~\ref{fig:figure3}B).
+  In order to compare the representative sequences generated by the various methods in the DC step, we employ both the weighted~\cite{Lozupone2007} (Figure~\ref{fig:figure2}A) and unweighted UniFrac method~\cite{Lozupone2005} (Figure~\ref{fig:figure2}B).
   The UniFrac distance metric (unique fraction metric) is a beta-diversity measure that computes the distance between two sets of taxa as the fraction of the branch length of the tree that leads to descendants from either one environment or the other, but not both~\cite{Lozupone2005}
   The weighted UniFrac distance metric takes into account the abundances of the representative sequences when calculating shared and unshared branch lengths, whereas the unweighted UniFrac distance metric does not and hence gives equal weights to each sequence.
-  In Figure~\ref{fig:figure3} the distances between methods are the average distance calculated between the reference sequence distribution for a pair of methods generated for every corresponding sample in the dataset.
+  In Figure~\ref{fig:figure2} the distances between methods are the average distance calculated between the reference sequence distribution for a pair of methods generated for every corresponding sample in the dataset.
   All UniFrac calculations were performed using the \texttt{scikit-bio}~\cite{thescikit-biodevelopmentteamScikitbioBioinformaticsLibrary2022} v0.5.6 Python package.
@@
   \vspace{-5mm}
-  In Figure~\ref{fig:figure4}C we use the Bray-Curtis distance metric to calculate the distance between the predicted (using the taxonomy databases in the TA step) and expected taxonomic distribution.
+  In Figure~\ref{fig:figure3}C we use the Bray-Curtis distance metric to calculate the distance between the predicted (using the taxonomy databases in the TA step) and expected taxonomic distribution.
   The Bray-Curtis distance is used to quantify the compositional dissimilarity between two different taxonomic distributions.
@@
   \vspace{-5mm}
-  In Figure~\ref{fig:figure6} we calculate the accuracy and robustness of the inferred association networks (using the various network inference algorithms and the consensus methods) against the original interaction used to create the taxonomic distribution.
+  In Figure~\ref{fig:figure5} we calculate the accuracy and robustness of the inferred association networks (using the various network inference algorithms and the consensus methods) against the original interaction used to create the taxonomic distribution.
   We use precision and sensitivity to calculate the accuracy of the predictions.

### 2022-08-29 sections/results.tex
Change: Update abstract and results
Change: Update Results section for draft v3
Also, changes the figure order (code remains unchanged)

@@
 
-  We developed \ac{micone}, a flexible and modular pipeline for 16S amplicon sequencing rRNA data (hereafter mentioned simply as 16S data) analysis.
-  It incorporates various popular, publicly available tools as well as custom Python modules and scripts to facilitate inference of co-occurrence networks from 16S data (see Methods).
-  Using \ac{micone} one can obtain co-occurrence networks by applying to 16S data (or to already processed taxonomic count matrices) any combinations of the supported tools.
+  We developed \ac{micone}, a flexible and modular pipeline for the inference of occurrence networks from 16S data.
+  It incorporates various popular, publicly available tools as well as custom Python modules for 16S data analysis and network inference (see Methods).
+  Using \ac{micone} one can obtain co-occurrence networks by applying to 16S data (or to already processed taxonomic count matrices) any combination of the supported tools.
   The effects of changing any intermediate step can be monitored and evaluated in terms of its final network outcome, as well as on any of the intermediate metrics and data outputs.
@@
   A centralized configuration file contains all the specifications regarding the modules used in the pipeline, and can be modified by the user to choose the desired set of tools.
-  The default tools and parameters for the configuration file can be found in the \ac{micone} documentation.
+  The default tools and parameters for the configuration file can be found in the \ac{micone} documentation (\href{https://micone.readthedocs.io}{https://micone.readthedocs.io}).
   In the current study, we performed a systematic analysis of each step of the pipeline to estimate how much the final co-occurrence network depends on the possible choices at each step.
@@
   Our analysis involves two types of data: The first type consists of sets of 16S sequencing data from real communities sampled from human stool microbiomes from a fecal microbiome transplant study of autism~\cite{Kang2017}.
+  This dataset was chosen because the sequences were easily accessible on Qiita~\cite{qiita} and optimally pre-processed according to the \ac{emp}~\cite{Thompson2017} protocol and can be used directly as input to the \ac{micone} pipeline.
   The second are datasets synthetically or artificially created for the specific goal of helping evaluate computational analysis tools.
-  In particular, in order to objectively compare, to the extent possible, how well each step in \ac{micone} best captures the underlying data, we use both mock data (labelled mock4, mock12 and mock16) from mockrobiota~\cite{Bokulich2016} as well as, synthetically networks generated using the NorTA~\cite{Kurtz2015} and seqtime~\cite{Rottjers2018} methods.
+  In particular, in order to objectively compare, to the extent possible, how well each step in \ac{micone} best captures the underlying data, we use both mock data (labelled mock4, mock12 and mock16) from mockrobiota~\cite{Bokulich2016} as well as, synthetic networks generated using the NorTA~\cite{Kurtz2015} and seqtime~\cite{Rottjers2018} methods.
   Detailed information on the mock communities and the settings used to generate the synthetic data are provided in the Methods section.
@@
 
-  \subsection*{The choice of reference database has the biggest impact on inferred networks}
-
-  In order to analyze the effect of different processing methods on the inferred co-occurrence networks, we generated co-occurrence networks using all possible combinations of methods and estimated the variability in the networks due to each choice (Figure \ref{fig:figure2}A).
-  All the network analysis in the study, unless explicitly mentioned, are performed only on the healthy samples in the fecal microbiome transplant study.
-  However, all other analysis are performed on all the samples in the fecal microbiome transplant study.
-  The networks are filtered on both association strength (0.1) and pvalue (0.05, only correlation methods) before the networks are subjected to the analyses.
-  The effects of various steps on the final co-occurrence network is estimated by building a linear model of the edges of the network as a function the various steps in the pipeline workflow (see Methods).
-  Figure \ref{fig:figure2}A, shows the percentage of total variation among the co-occurrence networks due to the different steps of the pipeline.
-  The \ac{ta} step, more specifically, the choice of 16S reference database contributes the most ($\sim65.4\%$) to the variation in the networks, followed by the \ac{op} step ($\sim26.8\%$).
-  This result highlights the importance of the taxonomy assignment step in the 16S data analysis workflow, implying that a change in the reference database will result in drastically different inferred networks.
-  The reason for this seems to derive from the differential labeling of constitutive representative sequences into taxonomic entities (Figure~\ref{fig:figure4} and Figure~\ref{fig:figure_s5}), which drastically alter the nodes and hence the underlying network topology.
-  In fact the variability induced by taxonomy assignment is much more significant than that due to the variability introduced based on how the reference sequences themselves are identified (in the DC step).
-  The residual variation (Figure \ref{fig:figure2}A) can be seen as an artifact that arises when multiple steps are changed at the same time, however this is negligible.
-
-  Figure \ref{fig:figure2}B is the PCA plot of all the generated co-occurrence networks, colored by the databases in the TA step (left) and the network inference algorithms in the NI step (right).
-  In other words, each point corresponds to a network generated using different combinations of tools, and captures how much the final network is affected by such a choice.
-  The importance of the TA step which was implied in Figure \ref{fig:figure2}A is also reflected here, since most of the networks can be separated based on the database used (Figure \ref{fig:figure2}B).
-  However, it is also evident that some networks, especially those inferred using the direct association based network inference methods, are much closer in the PCA plot regardless of the reference database used.
-  It can also be seen that the variation (on the first two principal axes) between the networks decreases when the low abundance \ac{otu}s are removed from the network (Figure~\ref{fig:figure_s1}).
-  These results suggest that the most important criterion for accurate comparative analysis of co-occurrence networks is the taxonomy reference database followed by the level of filtering of the taxonomy tables prior to network inference.
-
-  \FloatBarrier
-
-  \subsection*{Denoising and clustering methods differ in their identification of less common reference sequences}
+  \subsection*{DC: Denoising and clustering methods differ in their identification of less common reference sequences}
 
   The \ac{dc} step is commonly carried out to generate representative sequences from the demultiplexed and trimmed 16S sequencing data to obtain the \ac{otu}/\ac{esv} tables (counts of representative sequences for each sample).
-  In order to compare the representative sequence count tables generated by different tools, we processed the same 16S sequencing reads (samples from a fecal microbiome transplant study~\cite{Kang2017}) using 5 different methods: open-reference clustering, closed-reference clustering, denovo clustering, \ac{dada2}~\cite{Callahan2016} and Deblur~\cite{Amir2017}.
-  The first three methods are from the vsearch plugin and Deblur is from the deblur plugin of \ac{qiime2} v2021.8.0~\cite{bolyenReproducibleInteractiveScalable2019}.
-  The \ac{dada2} method is from the \ac{dada2} R package.
+  In order to compare the representative sequence count tables generated by different tools, we processed the same 16S sequencing reads (samples from a fecal microbiome transplant study~\cite{Kang2017}) using 5 different methods: open-reference clustering, closed-reference clustering, de novo clustering, \ac{dada2}~\cite{Callahan2016} and Deblur~\cite{Amir2017}.
+  The first three methods are from the vsearch plugin from \ac{qiime2}~\cite{bolyenReproducibleInteractiveScalable2019}.
   The closed and open reference methods in this analysis use the \ac{gg} database for reference sequence alignment.
 
-  To compare the representative sequences generated by these methods we employ both the weighted~\cite{Lozupone2007} (Figure~\ref{fig:figure3}A) and unweighted UniFrac method~\cite{Lozupone2005} (Figure~\ref{fig:figure3}B).
+  To compare the representative sequences generated by these methods we employ both the weighted~\cite{Lozupone2007} (Figure~\ref{fig:figure2}A) and unweighted UniFrac method~\cite{Lozupone2005} (Figure~\ref{fig:figure2}B).
   In both the comparisons, we set a count threshold of 10, such that if the count of the representative sequences in a particular sample is less than the threshold, it is omitted from the analysis.
@@
   The weighted UniFrac distance metric takes into account the counts of the representative sequences, whereas the unweighted UniFrac distance metric does not and hence gives equal weights to each sequence.
-  From Figure~\ref{fig:figure3}A one can see that the representative sequences generated by the different methods, with the exception of Deblur, are similar to each other when weighted by their abundance.
-  Figure~\ref{fig:figure3}B on the other hand shows an increase in dissimilarity between each pair of methods suggesting that the methods might differ in the assignment of sequences lower in abundance.
-  In order to verify this claim, for each of these methods we filtered out only the top 1000 representative sequences and recalculated the distance metrics for each method (Figure \ref{fig:figure_s3}).
+  From Figure~\ref{fig:figure2}A one can see that the representative sequences generated by the different methods, with the exception of Deblur, are similar to each other when weighted by their abundance.
+  Figure~\ref{fig:figure2}B on the other hand shows an increase in dissimilarity between each pair of methods suggesting that the methods might differ in the assignment of sequences lower in abundance.
+  In order to verify this claim, for each of these methods we selected the top 1000 representative sequences for each method and recalculated the distance metrics (Figure \ref{fig:figure_s3}).
   We observe that the both the weighted and unweighted UniFrac distances are increased implying that the top representative sequences generated by the different methods are not as similar to each other.
-  % TODO: Can we make this conclusion?
-  This leads us to conclude that the representative sequences in the middle range of the abundance distribution are those that must be the most similar between the methods.
+  Therefore, since the weighted UniFrac distances are lower than the unweighted distances, we conclude that the representative sequences in the middle range of the abundance distribution are those that must be the most similar between the methods.
 
-  These comparisons only elucidate the pairwise similarity or dissimilarity of a pair of methods.
+  These comparisons only elucidate the similarity or dissimilarity of a pair of methods.
   In order to determine the tool that most accurately recapitulates the reference sequences in the samples, we used the information about the known 16S sequences from the mock datasets (mock4, mock12, and mock16).
-  In particular, we used the pipeline to process mock community datasets using each of the possible methods included for this step.
+  In particular, we used the pipeline to process these datasets using each of the possible methods included for this step.
   Subsequently, we compared the predicted representative sequences with the expected representative sequences and their distribution.
-  The results (Figure~\ref{fig:figure3}C and D) show that, for the mock datasets, the predicted sequence distributions are substantially different from the expected sequence distribution.
-  This can be seen in the large values for both unweighted and weighted UniFrac distance values when compared to Figure~\ref{fig:figure3}A and B.
-  This result is more exaggerated in the case of the unweighted UniFrac metric, where some processed mock datasets show a very high deviation from the expected sequences.
+  The results (Figure~\ref{fig:figure2}C and D) show that, the predicted sequence distributions are considerably different from the expected sequence distribution.
+  This can be seen in the large values for both unweighted and weighted UniFrac distance values when compared to Figure~\ref{fig:figure2}A and B.
   These high deviations are primarily in one or two of the three datasets that were analyzed and show that the datasets themselves play a big role in the performance of these methods.
+  % TODO: How to condense this? Move to supplementary?
   Open-reference and de novo clustering methods perform the best under the weighted UniFrac metric and the worst (marginally) under the unweighted UniFrac metric.
@@
   There is no method that outperforms the rest in all datasets.
-  Based on their slightly better performance on the mock datasets, their \textit{de novo} error correcting nature and other previous studies~\cite{Nearing2018}, \ac{dada2} and Deblur appear to be the most reliable.
+  Based on their slightly better performance on the mock datasets, their de novo error correcting nature and previous independent evaluation~\cite{Nearing2018}, \ac{dada2} and Deblur appear to be the most reliable.
   This is because the open-reference and de novo clustering methods return a much larger number of \ac{otu}s compared to the other pipelines and this would affect the accuracy of the network inference step if filtering is not performed.
-  Overall, since \ac{dada2} as compared to Deblur, displays better performance on all the mock datasets and under the weighted UniFrac metric, we have decided to recommend this tool for the DC step of the pipeline.
+  Overall, since \ac{dada2} as compared to Deblur, displays better performance on all the mock datasets and under the weighted UniFrac metric, we this tool as the default for the DC step of the pipeline.
   However, if comparison across studies that have sequenced different 16S regions is required, closed-reference and open-reference might be a better option.
@@
   The \ac{micone} pipeline supports two different chimera checking methods, uchime-denovo and remove bimera.
-  The uchime-denovo method is from \ac{qiime2}, whereas the remove bimera method is from the \ac{dada2} R package.
+  The ``uchime-denovo'' method is from \ac{qiime2}~\cite{bolyenReproducibleInteractiveScalable2019}, whereas the ``remove bimera'' method is from the \ac{dada2} R package~\cite{Callahan2016}.
   We did not notice any significant difference between the two chimera checking methods (Figure~\ref{fig:figure_s4}), implying that they identify and remove mostly the same set of sequences as chimeras.
   Since the remove bimera method was originally developed in conjunction with dada2 we use this method as the default for the chimera checking module.
+  The DC step results in a set of unique sequences, these will be referred to as representative sequences in the subsequent steps.
 
@@
 
-  \subsection*{Taxonomy databases vary widely in taxonomy hierarchy and update frequency}
+  \subsection*{TA: Taxonomy databases vary widely in taxonomy assignments beyond Order level}
 
   Taxonomy databases are used to assign taxonomic identities to the representative sequences obtained after the DC step.
-  In order to compare the assigned taxonomies from different databases, we use the same reference sequences and assign taxonomies to them using different taxonomy reference databases.
-  The three 16S taxonomic reference databases used in this study are SILVA~\cite{Quast2012}, \ac{gg}~\cite{DeSantis2006} and \ac{ncbi} RefSeq~\cite{Sayers2009}.
-  SILVA and \ac{gg} are two popular 16S databases used for taxonomy identification and the \ac{ncbi} RefSeq nucleotide database contains 16S rRNA sequences as a part of two BioProjects - 33175 and 33317.
-  The three databases vastly differ in terms of their last update status - \ac{gg} was last updated on May 2013, SILVA was last updated on August 2020 at the time of writing and \ac{ncbi} is updated regularly as new sequences are curated.
-  Since updates to taxonomic classifications are frequent, these databases vary significantly in terms of taxonomy hierarchies including species names and phylogenetic relationships~\cite{Balvociute2017}.
-  Assignment using a particular database also requires a query tool, we use the ``Naive Bayes'' classifier from \ac{qiime2} for the \ac{gg} and SILVA databases and the ``BLAST'' tool (also included as a \ac{qiime2} plugin) for the \ac{ncbi} database.
+  The three 16S taxonomic reference databases used in this study are SILVA~\cite{Quast2012}, \ac{gg}~\cite{DeSantis2006} and \ac{ncbi} RefSeq~\cite{Sayers2009} (see Methods).
+  These databases vary significantly in terms of taxonomy hierarchies including species names and phylogenetic relationships~\cite{Balvociute2017}.
+  Assignment using a particular database also requires a query tool.
+  We use the ``Naive Bayes'' classifier from \ac{qiime2} for the \ac{gg} and SILVA databases and the ``BLAST'' tool (also included as a \ac{qiime2} plugin) for the \ac{ncbi} database.
   These tools have been well quantified and optimized~\cite{bokulichOptimizingTaxonomicClassification2018}, hence, we make use of the default parameters in our analyses.
 
-  The representative sequences obtained from the \ac{dada2} method (with remove bimera for chimera checking) in the DC step were used for taxonomic assignment using the three reference databases.
-  Figure~\ref{fig:figure4}A depicts a flow diagram (alluvial plot) that shows how the top 50 representative sequences (sorted by abundance) are assigned a Genus according to the three databases.
-  We observe that the top 50 representative sequences are assigned to 17 Genus in \ac{gg}, 23 in SILVA and 17 in \ac{ncbi}, however the sequences that are assigned to a particular Genus vary significantly based on the taxonomy database.
+  The representative sequences obtained using the default settings of the DC step were used for taxonomic assignment using the three reference databases.
+  Figure~\ref{fig:figure3}A depicts a flow diagram (alluvial plot) that shows how the top 50 representative sequences (sorted by abundance) are assigned a Genus according to the three databases.
+  We observe that the top 50 representative sequences are assigned to 17 Genera in \ac{gg}, 23 in SILVA and 17 in \ac{ncbi}, however the sequences that are assigned to a particular Genus vary significantly based on the taxonomy database.
   We observe that not only does the assigned Genus composition vary significantly, but the percentage of unassigned representative sequences (pink) also differ.
-  The large percentage of unassigned sequences indicate that even among the most abundant representative sequences a large fraction is assigned an "unknown" Genus in two of the three databases (\ac{ncbi} and \ac{gg}).
-  A representative sequence might be assigned an "unknown" Genus for one of two reasons: the first is if the taxonomy identifier associated with the sequence in the database did not contain a Genus; the second (more likely) reason is that the database contains multiple sequences that are very similar to the query (representative) sequence and the consensus algorithm (from \ac{qiime2}) is unable to assign one particular Genus at the required confidence.
-  Additionally, we observe that other abundant Genus like Bacteroides and Escherichia are also assigned to different representative sequences (indicated by the different proportions in the alluvial plot).
-  Furthermore, the most abundant Genus (among the top 50 representative sequences) in the \ac{ncbi} taxonomy table was \textit{Escherichia}, but these sequences were assigned to "unknown" in \ac{gg} and in the SILVA taxonomy table they were assigned to \textit{Escherichia-Shigella}.
-  Although these are minor differences (in the case of \ac{ncbi} and SILVA), when comparing a large number of taxonomy composition tables these problems are hard to diagnose.
+  The large percentage of unassigned sequences are because a large fraction of the representative sequences are assigned to an "unknown" Genus during assignment process.
+  This is especially evident in two of the three databases, \ac{ncbi} and \ac{gg}.
+  A representative sequence might be assigned an "unknown" Genus for one of two reasons: the first is if the taxonomy identifier associated with the sequence in the database did not contain a given Genus; the second, more likely, reason is that the database contains multiple sequences that are very similar to the query (representative) sequence and the consensus algorithm (from \ac{qiime2}) is unable to assign one particular Genus at the required confidence.
+  Additionally, we observe that other abundant Genera like Bacteroides and Escherichia are also assigned to different representative sequences, as indicated by the different proportions in the alluvial plot.
+  These changes together result in taxonomic assignments generated from each database leading to drastically different taxonomic abundance distributions.
 
-  After assigning all the representative sequences to taxonomies we perform a pairwise comparison of the similarity between the top 100 assignments (by abundance) from different databases at every taxonomic level (Figure~\ref{fig:figure4}B).
-  The assignments beyond Family level (Family, Genus and Species) are very dissimilar with $<45\%$ similarity between any pair of databases.
-  The matches and mismatches in the assignments from all three databases are comparable, with \ac{ncbi} and SILVA being the most similar across most of the taxonomic levels, followed closed by \ac{gg} vs.
-SILVA.
-  However, the number of mismatches for taxonomy assignments beyond the Genus level (Genus, Species) are very high ($>67\%$), implying that the taxonomy assignments from each reference database are fairly unique and are largely responsible for the differences observed in the co-occurrence networks generated from different taxonomy databases.
-  The assignment between SILVA and the other two databases were originally significantly different ($40\%$ mismatch) even at the Phylum level, indicating that 40 of the top 100 assignments at the Phylum level in the SILVA database are different in SILVA.
-  However, this was fixed by making minor adjustments to the taxonomic names, such as changing Bacteroidota to Bacteroidetes in the SILVA Phylum assignments.
+  After assigning all the representative sequences to taxonomies we perform a pairwise comparison of the similarity between the top 100 assignments (by abundance) from different databases at every taxonomic level (Figure~\ref{fig:figure3}B).
+  The comparisons of the assignments beyond Family level (Family, Genus and Species) show less than $45\%$ similarity between any pair of databases.
+  However, the number of mismatches for taxonomy assignments beyond the Genus level (Genus, Species) are greater than $67\%$, implying that the taxonomy assignments from each reference database are fairly unique and are largely responsible for the differences observed in the co-occurrence networks generated from different taxonomy databases.
   The comparison of all assigned genera (Figure~\ref{fig:figure_s5}), instead of just the top 100, contains a higher percentage of mismatches.
@@
   As in the previous section, the comparisons so far only indicate similarity or dissimilarity between methods.
-  In order to obtain an absolute measure of accuracy of the taxonomic assignments we use the expected reference sequences from the mock datasets as the query sequences for the databases and the expected taxonomic composition as the standard to compare against (Figure~\ref{fig:figure4}C).
-  Here, we observe that none of the databases perform better than the others in absolute terms and that the dissimilarity with the expected compositon is high, indicating that all the databases perform poorly when trying to recapture the expected taxonomic composition.
+  In order to obtain an absolute measure of accuracy of the taxonomic assignments we use the expected reference sequences from the mock datasets as the query sequences for the databases and the expected taxonomic composition as the standard to compare against.
+  We use the Bray-Curtis distance metric~\cite{virtanenSciPyFundamentalAlgorithms2020} to calculate the distance between the predicted and expected taxonomic distribution (Figure~\ref{fig:figure3}C).
+  Here, we observe that none of the databases perform better than the others in absolute terms and that the dissimilarity with the expected composition is high ($>0.5$ for Family and Genus and $>0.9$ for Species), indicating that all the databases perform poorly when trying to recapture the expected taxonomic composition.
+  % TODO: What are the implications?
 
-  Given that no database performs better than others against mock datasets, the choice of which database to use should be driven by other reasons.
+  Since no database performs better than others against mock datasets, the choice of which database to use could be driven by other reasons.
   One user-specific way to choose, would be based on the known representation of taxa for the microbiome of interest (see Discussion).
   Another reason could be the frequency of updates and the potential for future growth.
-  Since the \ac{gg} database is more widely used in the field~\cite{Balvociute2017}, to enable easier comparison with published networks in literature, we set \ac{gg} as the \ac{micone} standard for taxonomy assignment.
+  Since the \ac{gg} database is used frequently as a 16S reference database in studies~\cite{Balvociute2017}, to enable easier comparison with published networks in literature, we set \ac{gg} as the default for taxonomy assignment.
   On the other hand, since the degree of similarity of assignments between \ac{ncbi} and \ac{gg} is fairly high (Figure~\ref{fig:figure_s5}), and due to its regular revision and maintenance the \ac{ncbi} database is also a good choice for the user.
-  % TODO: Do we need more refs to support our argument?
+  The TA step results in a taxonomic counts table that is used as input to the subsequent steps of the pipeline.
+  The count tables at different levels can be obtained by aggregation, for example, Genus count tables were obtained by summing up the counts of the lower taxonomy levels (Species and \ac{otu}) that map to the same higher taxonomy level name.
 
@@
 
-  \subsection*{Networks generated using different network inference methods show notable difference in edge-density and connectivity}
+  \subsection*{NI: Networks generated using different network inference methods show notable difference in edge-density and connectivity}
 
-  The ten different network inference methods used in this study are: \ac{spieceasi}~\cite{Kurtz2015}, FlashWeave~\cite{tackmannRapidInferenceDirect2019}, \ac{cozine}~\cite{haCompositionalZeroinflatedNetwork2020a}, \ac{harmonies}~\cite{jiangHARMONIESHybridApproach2020}, \ac{spring}~\cite{yoonMicrobialNetworksSPRING2019}, \ac{mldm}~\cite{Yang2017}, \ac{sparcc}~\cite{Friedman2012,Watts2018}, propr~\cite{quinnProprRpackageIdentifying2017}, Spearman and Pearson.
+  The ten different network inference methods used in this study are: \ac{spieceasi}~\cite{Kurtz2015}, FlashWeave~\cite{tackmannRapidInferenceDirect2019}, \ac{cozine}~\cite{haCompositionalZeroinflatedNetwork2020a}, \ac{harmonies}~\cite{jiangHARMONIESHybridApproach2020}, \ac{spring}~\cite{yoonMicrobialNetworksSPRING2019}, \ac{mldm}~\cite{Yang2017}, \ac{sparcc}~\cite{Friedman2012,Watts2018}, propr~\cite{quinnProprRpackageIdentifying2017}, Spearman and Pearson correlation methods.
   These network inference methods fall into two groups, the first set of methods (Pearson, Spearman, \ac{sparcc}, and propr) infer pairwise correlations while the second set infer direct associations (\ac{spieceasi}, FlashWeave, \ac{cozine}, \ac{harmonies}, \ac{spring}, and \ac{mldm}).
-  Pairwise correlation methods involve calculation of the correlation coefficient between every pair of node (taxonomic entity like Genus) leading to the detection of spurious indirect connections.
+  Pairwise correlation methods involve calculation of the correlation coefficient between each pair of nodes (taxonomic entity like Genera) leading to the inclusion of spurious indirect connections.
   On the other hand, direct association methods use conditional independence to avoid the detection of correlated but indirectly connected taxonomic entities~\cite{Kurtz2015,Menon2018}.
 
-  For the analysis presented in this section, we used the filtered (filtering performed during the \ac{op} step) Genus level counts table obtained using the \ac{gg} (dada2 and remove bimera methods were used in the DC step) reference database from the \ac{ta} step as the input for the network inference algorithms.
-  The Genus count tables were obtained by summing up the counts of the lower taxonomy levels (Species and \ac{otu}) that map to the same higher taxonomy level name.
-  Figure~\ref{fig:figure5}A shows the networks inferred from this dataset using the different inference algorithms.
+  For the analysis presented in this section, we used the filtered (performed during the \ac{op} step, refer Methods) Genus level counts table obtained using the default settings in the DC and TA steps as the input for the network inference algorithms.
+  Figure~\ref{fig:figure4}A shows the networks inferred from this dataset using the different inference algorithms.
   The different networks differ vastly in their edge-density and connectivity; even the edges in common to these networks can have their signs inverted.
-  Note, however, that some of these comparisons depend on the threshold that has to be applied to the networks (currently 0.3 for the correlation based methods, based on~\cite{Friedman2012} and 0.01 for the direct association methods).
-  \ac{mldm} is not shown in the comparisons because the algorithm failed to converge for most of the network combinations, hence, it is only used in the network variation analysis.
 
-  In order to obtain a quantitative picture of the differences between the inferred networks, we analyzed the distribution of common nodes and edges (Figure~\ref{fig:figure5} B and C) using UpSet plots~\cite{Lex}.
-  Only \ac{harmonies}, \ac{cozine}, \ac{spieceasi}, \ac{spring}, propr, and \ac{sparcc} are used in the comparison since Pearson and Spearman add a large number of spurious edges as they are not intended for compositional datasets).
-  The results for the node intersections show that the networks have a large proportion of nodes in common ($33$ out of $68$ total unique nodes) and no network possesses any unique node.
+  In order to obtain a quantitative picture of the differences between the inferred networks, we analyzed the distribution of common nodes and edges (Figure~\ref{fig:figure4} B and C) using UpSet plots~\cite{Lex}.
+  Only \ac{harmonies}, \ac{cozine}, \ac{spieceasi}, \ac{spring}, propr, and \ac{sparcc} are used in the comparison since Pearson and Spearman add a large number of spurious edges as they are not intended for compositional datasets.
+  The results for the node intersections show that the networks have $33$ out of $68$ total unique nodes in common and no network possesses an unique node.
   The edge intersections in contrast show that only $8$ edges (out of $202$ total unique edges) are in common between all the methods and each network has a large number of unique edges.
   These results indicate that there is a substantial rewiring of connections in the inferred networks.
-  Hence, in order to determine the interactions that are most likely to exist in the dataset we advocate for the development of a consensus algorithm.
+  Hence, in order to determine the interactions that are most likely to exist in the dataset we advocate for the development of a consensus algorithm (refer to next section).
 
@@
 
-  \subsection*{The consensus network algorithms}
+  \subsection*{NI: The consensus network algorithms}
 
-  We observe a high degree of rewiring in the inferred association networks generated by the different methods (Figure \ref{fig:figure5}C), implying that there is a lack of agreement on the associations that should be a part of the final network.
+  We observe a high degree of rewiring in the inferred association networks generated by the different methods (Figure \ref{fig:figure4}C), implying that there is a lack of agreement on the associations that should be a part of the final network.
   Therefore, we have developed a consensus method that would take into consideration the evidence offered by each network inference algorithm and generate a consensus network that would contain the most probable interactions for the input data.
   We have developed two algorithms for the generation of the consensus network, the simple voting and the scaled-sum methods (see Methods).
-  First, for the correlation-based methods we merge the p-values using the Brown's p-value merging method~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016,faustCoNetAppInference2016} (see Methods).
+  Both approaches combine network inferred from both correlation-based and direct association methods.
+  First, for the correlation-based methods we calculate p-values using null models and then merge the p-values using the Brown's p-value merging method~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016,faustCoNetAppInference2016} (see Methods).
   Second, we filter all the inferred networks based on an association strength threshold of 0.1 and a p-value cutoff of 0.05.
   Finally, we apply the consensus algorithms we have developed on these filtered networks.
-  The simple voting method does a consensus based on whether the number of inferred networks that have an edge between a pair of nodes is greater than a threshold.
-  The scaled-sum method does a consensus based on whether the normalized sum of an edge between a pair of nodes from all inferred networks is greater than a threshold.
+  % The simple voting method does a consensus based on whether the number of inferred networks that have an edge between a pair of nodes is greater than a threshold.
+  % The scaled-sum method does a consensus based on whether the normalized sum of an edge between a pair of nodes from all inferred networks is greater than a threshold.
   The default algorithm for the pipeline is the scaled-sum method, since this method takes into account the weights of the associations in the determination of the final consensus.
@@
 
-  Similar to the previous steps of the pipeline, where were we evaluated the performance of methods on mock datasets, we can evaluate the network inference algorithms and the final consensus network using synthetic interaction data.
+  Similar to what was done for the previous steps of the pipeline, we evaluated the network inference algorithms and the final consensus network using synthetic interaction data.
   For this purpose, we generated synthetic interaction data using the ``NorTA''~\cite{Kurtz2015} and ``seqtime''~\cite{faustSignaturesEcologicalProcesses2018} methods (see Methods), which have been previously used for benchmarking network inference methods.
@@
   This counts table is used as the input to the \ac{micone} pipeline to generate predicted associations.
-  The interaction network used to generate the counts table is used as the source of true interactions to calculate the precision and sensitivity for each network inference algorithm for the NorTA datasets (Figure \ref{fig:figure6}).
-  Each point in the scatter plot represents a network with different input network topology and OTU abundance distributions.
+  The interaction network used to generate the counts table is used as the source of true interactions to calculate the precision and sensitivity for each network inference algorithm for the NorTA datasets (Figure \ref{fig:figure5}).
   The goal of the consensus network algorithms is to increase the precision of the resultant consensus network by removing edges that are either not present in most of the inference methods or whose association strength is low.
-  From Figure \ref{fig:figure6} we can confirm that the consensus algorithm, especially the scaled-sum method, does indeed produce networks with more precision, but as a consequence the sensitivity of the networks is reduced.
-  Overall, we see that the scaled-sum method for a parameter value of 1.0 performs the best in terms of average precision ($1.000$).
-  However, the scaled-sum method for a parameter value of 0.667 has the second-highest average precision ($0.985$) and does not underperform in terms of sensitivity.
+  In Figure \ref{fig:figure5} we show that the consensus algorithm, especially the scaled-sum method, captures true associations with more precision, but as a consequence the sensitivity of the predictions is reduced.
+  Overall, we see that the scaled-sum method for $p=1.0$ performs the best in terms of average precision ($1.000$).
+  However, the scaled-sum method for $p=0.667$ has the second-highest average precision ($0.985$) and does not underperform in terms of sensitivity.
   Among the individual inference methods, we observe that \ac{spieceasi} has the best average precision ($0.944$)
   Therefore, the consensus networks provide the means to obtain a short list of associations that would have a high probability of being present in the real interaction network.
-  This small set of associations could be further tested and validated experimentally to verify their plausibility.
   Furthermore, we also find that the consensus algorithm performs drastically better in terms of precision (0.820 for scaled-sum method vs.
@@
   The precision and sensitivity graphs on the seqtime data can be found in the Supplementary section (Figure~\ref{fig:figure_s7}).
+
+  \FloatBarrier
+
+  \subsection*{Overall assessment of impact on inferred networks}
+
+  In order to analyze the effect of different processing methods on the inferred co-occurrence networks, we generated these networks using all possible combinations of methods and estimated the variability in the networks due to each choice (Figure \ref{fig:figure6}A).
+  The networks used in this analysis are those directly inferred from the correlation-based and direct inference methods and to which the consensus algorithm is yet to be applied.
+  The effects of various steps on the final co-occurrence network are estimated by building a linear model of the edges of the network as a function the various steps in the pipeline workflow (see Methods).
+  Figure \ref{fig:figure6}A, shows the percentage of total variation among the co-occurrence networks due to the different steps of the pipeline.
+  The \ac{ta} step or more specifically the choice of 16S reference database, contributes the most ($\sim65.4\%$) to the variation in the networks, followed by the \ac{op} step ($\sim26.8\%$).
+  This result highlights the importance of the taxonomy assignment step in the 16S data analysis workflow, implying that a change in the reference database will result in drastically different inferred networks.
+  The reason for this seems to derive from the differential labeling of constitutive representative sequences into taxonomic entities (Figure~\ref{fig:figure3} and Figure~\ref{fig:figure_s5}), which drastically alter the nodes and hence the underlying network topology.
+  In fact the variability induced by taxonomy assignment is much more significant than that due to the variability introduced based on how the reference sequences themselves are identified (in the DC step).
+  The fraction labeled as the residual is an artifact that arises when multiple steps are changed at the same time, however this can be ignored as the fraction is negligible.
+
+  The effects of the different steps of the pipeline on the inferred networks can be best visualized through dimensionality reduction.
+  Figure \ref{fig:figure6}B is the PCA plot of all the generated co-occurrence networks, colored by the databases in the TA step (left) and the network inference algorithms in the NI step (right).
+  In other words, each point corresponds to a network generated using different combinations of tools, and captures how much the final network is affected by such a choice.
+  The importance of the TA step, shown in Figure \ref{fig:figure6}A, is also reflected here since most of the networks can be separated based on the database used (Figure \ref{fig:figure6}B).
+  However, it is also evident that some networks, especially those inferred using the direct association based network inference methods, are much closer in the PCA plot regardless of the reference database used.
+  Additionally, the variation (on the first two principal axes) between the networks decreases when the low abundance \ac{otu}s are removed from the network (Figure~\ref{fig:figure_s1}).
+  These results suggest that the most important criterion for accurate comparative analysis of co-occurrence networks is the taxonomy reference database followed by the level of filtering of the taxonomy tables prior to network inference.
 
@@
   For some of these choices (e.g. for the \ac{op} step), there is no clear metric to establish the best protocol.
-  For other choices, the mock communities and synthetic data provide an opportunity to select the combinations of tools that yield more accurate and robust results.
-  Despite this partial degree of assessment, we wish to suggest a combination of tools and parameters that produce networks that are derived from the combination of tools which performed best on the mock communities, and displayed the highest robustness to switching to alternative methods.
+  For other choices, the mock communities and synthetic data provide an opportunity to select the combinations of tools that yield more accurate and robust results, though those may depends on assumptions and details of the mock data.
+  Despite this partial degree of assessment, we wish to suggest a combination of tools and parameters that led to the best performance based on the mock communities, and displayed the highest robustness to switching to alternative methods.
   These tools and parameters are chosen as the defaults for the pipeline and are given in Table~\ref{tab:micone_tools}.
 
-  The recommended tool for the \ac{dc} step (\ac{dada2}) was chosen based on its accuracy in recapitulating the reference sequences in mock communities.
-  For the chimera checking module of the \ac{dc} step, since both uchime-denovo and remove bimera perform similarly, we suggest using remove bimera since it was developed specifically for \ac{dada2} and should provide the best performance in conjunction with that denoising method.
-  The choice of the taxonomy reference database in the \ac{ta} step is dictated largely by the species expected to be present in the sample as well the database used in similar studies, if comparison is a goal.
-  Nevertheless, we suggest \ac{gg} database along with the Naive Bayes classifier as the query tool since the database is used most often in taxonomic studies.
-  The filtering thresholds at the \ac{op} step are set at values described in Table~\ref{tab:micone_tools} in order to prevent the addition of large numbers of nodes to networks and the subsequent decrease in the performance of network inference algorithms.
-  For the \ac{op} step, we recommend that the filtering be turned on and used at the default settings, unless a taxonomy entity of interest might be filtered out at the default settings.
-  This is necessary since a large number of rows in the counts matrix would reduce the accuracy of the inference methods and increase the computational time.
-  Finally, for the \ac{ni} step, we recommend the scaled-sum consensus method along with p-value merging (for the correlation-based methods) to generate a final consensus network for the input data.
-
-  Figure~\ref{fig:figure7}A show the default network compared against networks generated by altering one of the steps of the pipeline from the default.
+  Figure~\ref{fig:figure7}A shows the comparison of the default network against networks generated by altering one of the steps of the pipeline from the default.
   This result suggests that the most significant changes (many new nodes and edges are added) occur when the \ac{ta} step, \ac{op} step and the \ac{ni} step are changed.
@@
 
-  % TODO: Say that this shows how important the consensus network is in the whole pipeline
-  The low L1 distances show us that not only does the consensus help reconcile the variability in the edges due to multiple network inference algorithms, it also helps reconcile the variability to the other steps of the pipeline as well.
-  This result is different from that observed in Figure~\ref{fig:figure2} since in this particular analysis all the steps other than the \ac{ni} step have the consensus algorithm applied to them, which drastically reduces the edges in the graph that have lower probabilities.
+  The low L1 distances show us that not only does the consensus help reconcile the variability in the edges due to multiple network inference algorithms, it also helps reduce the variability due to the other steps of the pipeline as well.
+  This result is different from that observed in Figure~\ref{fig:figure6} since in this particular analysis all the steps other than the \ac{ni} step have the consensus algorithm applied to them, which drastically reduces the edges in the graph that have lower probabilities.
   This reinforces the importance of the consensus network in increasing the robustness of the inferred networks to different processing methods.
@@
   As a consequence of using the consensus network algorithm, the final co-occurrence networks are sparse and can be visually compared and examined.
-  % TODO: How much detail should we go into?
   We observe 10 unique links in the network for control samples and 3 unique links in the network for autistic subjects.

### 2022-08-29 sections/supplementary.tex
Change: Update Results section for draft v3
Also, changes the figure order (code remains unchanged)

@@
           \textbf{The networks which were filtered during the \ac{op} step are less variable}
-          All combinations of inferred networks are shown as points on a PCA plot and shows the variation in the inferred networks in the first two principal axes (similar to Figure~\ref{fig:figure2}B).
+          All combinations of inferred networks are shown as points on a PCA plot and shows the variation in the inferred networks in the first two principal axes (similar to Figure~\ref{fig:figure6}B).
           Each point on the PCA plot represents a network inferred using different combinations of tools and parameters that are available in the \ac{micone} pipeline.
@@
           \textbf{The UniFrac distance between the 1000 most abundant representative sequences is higher than that when all sequences are considered}.
-          Each value is the average UniFrac distance between the reference sequences generated by the various methods in the \ac{dc} step (similar to Figure~\ref{fig:figure3}).
+          Each value is the average UniFrac distance between the reference sequences generated by the various methods in the \ac{dc} step (similar to Figure~\ref{fig:figure2}).
           There is an increase in both weighted and unweighted UniFrac distances compared to when all the representative sequences are considered.
@@
           \textbf{The pairwise comparison of assignments generated using different databases for all representative sequences has a higher proportion of mismatches}.
-          The comparison made here is similar to Figure~\ref{fig:figure4}B, but instead of the top 100 taxonomic entities (by abundance), all the assignments from one database are matched with those from the other two databases.
+          The comparison made here is similar to Figure~\ref{fig:figure3}B, but instead of the top 100 taxonomic entities (by abundance), all the assignments from one database are matched with those from the other two databases.
           Higher percentage of mismatches implies that the matching of the taxonomies in the more abundant sequences (top 100) are more consistent.
@@
         The different consensus based methods used are: pvalue merging (PM), scaled-sum (SS) and simple voting (SV) method.
-        With the addition of Pearson and Spearman in the consensus calculation we observe a decrease in the average precision scores of both the consensus algorithms (compared to Figure~\ref{fig:figure6}.
+        With the addition of Pearson and Spearman in the consensus calculation we observe a decrease in the average precision scores of both the consensus algorithms (compared to Figure~\ref{fig:figure5}.
         The overall best precision was still consistently obtained by the scaled-sum method (0.985, 0.963 and 1.000), but the performance was not stable to an increase in the parameter value.

### 2022-08-29 sections/tables_figures.tex
Change: Edit the supplementary section for draft v2
Change: Update Results section for draft v3
Also, changes the figure order (code remains unchanged)

@@
   \newpage
-
-  \begin{figure}[H]
-    \centering
-    \includegraphics[width=0.74\linewidth]{figure1.pdf}
-  \end{figure}
 \begin{figure}[H]
@@
 
-  \FloatBarrier
-  \newpage
-
-  % TODO: Add equations for the linear model to the left of the pie chart
   \begin{figure}[H]
     \centering
-    \includegraphics[width=\textwidth]{figure2.pdf}
+    \caption{
+      \textbf{The representative sequences generated by the different denoising/clustering methods are very similar but differ in the sequences that are lower in abundance.}
+      \textbf{(A)} The average weighted UniFrac distance between the representative sequences shows that the representative sequences and their compositions are fairly identical between the methods (with the exception of Deblur (DB)).
+      \textbf{(B)} The relatively larger average unweighted UniFrac distance indicates that methods differ in their identification of sequences that are lower in abundance.
+      The number of \ac{otu}s or \ac{esv}s generated by the respective methods are provided in the parenthesis next to the names.
+      The data used for the analysis in \textbf{(A, B)} were all the samples (healthy and autistic) from the fecal microbiome transplant dataset.
+      \textbf{(C, D)} The distributions of the average weighted and unweighted UniFrac distance between the expected sequence profile and the predicted sequence profile in the mock datasets.
+      The distributions of the average weighted UniFrac distance show that de novo (DN) and open reference (OR) were the best performing methods in most of the datasets, while they are worst performing methods under the unweighted UniFrac metric.
+      The good performance of dada2 (D2) under both distance metrics combined with its approach of identifying \ac{esv}s using de novo methods, prompts us to propose it as the default method for the DC step.
+      The data used for the analysis in \textbf{(C, D)} were the mock datasets from mockrobiota~\cite{Bokulich2016}.
+    }
+    \label{fig:figure2}
   \end{figure}
+
+  \begin{figure}[H]
+    \centering
+    \caption{
+      \textbf{Taxonomic reference databases vary widely in terms of their taxonomy assignments beyond the Order level.}
+      \textbf{(A)} The assignments of the top 50 representative sequences to their respective taxonomies using the three different reference databases.
+      This result illustrates how the same sequences are assigned to different Genus under the different databases.
+      A significant portion of the representative sequences are assigned to an ``unknown'' Genus in two of three databases (\ac{gg} and \ac{ncbi}).
+      \textbf{(B)} The number of representative sequences assigned to the same taxonomic label when using different reference databases, shown for the top 100 sequences.
+      The number of mismatches are fewer at higher taxonomic levels, but, even at the Order level there exists greater than 51\% of mismatches, demonstrating the poor agreement in taxonomic labels assigned by the different databases.
+      The data used for the analysis in \textbf{(A, B)} were all the samples (healthy and autistic) from the fecal microbiome transplant dataset.
+      \textbf{(C)} The Bray-Curtis dissimilarity between the expected taxonomy profile and calculated taxonomy profile in the mock datasets shows that there is no singular best choice of database for every dataset, as all the databases have similar performance.
+      However, we still recommend the \ac{gg} database and the Naive Bayes classifier as the defaults for the \ac{micone} pipeline.
+      Due to the popularity of the \ac{gg} database, comparison against existing networks in literature will be easier since many published datasets make use of this database in their analyses.
+      The datasets used for the analysis in \textbf{(C)} were the mock datasets from mockrobiota.
+    }
+    \label{fig:figure3}
+  \end{figure}
+
+  \begin{figure}[H]
+    \centering
+    \caption{
+      \textbf{Networks generated using different network inference methods show notable differences both in terms of edge-density and connectivity}.
+      \textbf{(A)} The nine different networks (excluding \ac{mldm}) generated by the different network inference methods are very dissimilar.
+      The green links are positive associations and the orange links represent negative associations.
+      A threshold of 0.3 was set for the correlation-based methods (\ac{sparcc}, Propr, Spearman and Pearson) and a threshold of 0.01 was set for the other direct association methods.
+      The correlation-based methods in general produce graph with higher edge-densities.
+      \textbf{(B)} The node overlap Upset plot indicates that all the networks have a large proportion of common nodes involved in connections (33 out of 68).
+      Whereas, \textbf{(C)} the edge overlap Upset plot shows that a very small fraction of these connections are actually shared (8 out of 202).
+      The data used in this analysis were the healthy stool samples from the fecal microbiome transplant dataset.
+      \ac{mldm} is not shown in the comparisons because the algorithm failed to converge for most of the network combinations.
+    }
+    \label{fig:figure4}
+  \end{figure}
+
+  \begin{figure}[H]
+    \centering
+    \caption{
+      \textbf{The precision and sensitivity of the inferred networks on the NorTA synthetic interaction data.}
+      The ``I'' prefix indicates the individual network inference methods and the ``X'' prefix denotes the consensus based methods that are a part of \ac{micone}.
+      The different consensus based methods used are: pvalue merging (PM), scaled-sum (SS) and simple voting (SV) method.
+      Pearson and Spearman methods are not used in the calculation of the consensus.
+      Among all the independent network inference methods, \ac{spieceasi} has the best average precision (0.944), but the overall best precision was consistently obtained by the scaled-sum method (0.956, 0.985 and 1.000).
+      The simple voting method when using presence of edges in all inferred networks as a requirement (parameter value 1.000), also outperforms \ac{spieceasi} on average precision (0.969).
+      Although \ac{spieceasi} has a higher performance on sensitivity, if the goal of network inference is to obtain the list of associations that have a high probability of existing in the real microbial community, then the consensus methods perform better.
+      Therefore, we recommend the scaled-sum consensus method as the default tool for network inference.
+    }
+    \label{fig:figure5}
+  \end{figure}
+
   \begin{figure}[H]
@@
     }
-    \label{fig:figure2}
-  \end{figure}
-  \FloatBarrier
-  \newpage
-
-  \begin{figure}[H]
-    \centering
-    \includegraphics[width=\textwidth]{figure3.pdf}
-  \end{figure}
-  \begin{figure}[H]
-    \centering
-    \caption{
-      \textbf{The representative sequences generated by the different denoising/clustering methods are very similar but differ in the sequences that are lower in abundance.}
-      \textbf{(A)} The average weighted UniFrac distance between the representative sequences shows that the representative sequences and their compositions are fairly identical between the methods (with the exception of Deblur (DB)).
-      \textbf{(B)} The relatively larger average unweighted UniFrac distance indicates that methods differ in their identification of sequences that are lower in abundance.
-      The number of \ac{otu}s or \ac{esv}s generated by the respective methods are provided in the parenthesis next to the names.
-      The data used for the analysis in \textbf{(A, B)} were all the samples (healthy and autistic) from the fecal microbiome transplant dataset.
-      \textbf{(C, D)} The distributions of the average weighted and unweighted UniFrac distance between the expected sequence profile and the predicted sequence profile in the mock datasets.
-      The distributions of the average weighted UniFrac distance show that de novo (DN) and open reference (OR) were the best performing methods in most of the datasets, while they are worst performing methods under the unweighted UniFrac metric.
-      The good performance of dada2 (D2) under both distance metrics combined with its approach of identifying \ac{esv}s using de novo methods, prompts us to propose it as the default method for the DC step.
-      The data used for the analysis in \textbf{(C, D)} were the mock datasets from mockrobiota~\cite{Bokulich2016}.
-    }
-    \label{fig:figure3}
-  \end{figure}
-  \FloatBarrier
-  \newpage
-
-  \begin{figure}[H]
-    \centering
-    \includegraphics[width=\textwidth,height=1.2\textwidth]{figure4.pdf}
-  \end{figure}
-  \begin{figure}[H]
-    \centering
-    \caption{
-      \textbf{Taxonomic reference databases vary widely in terms of their taxonomy assignments beyond the Order level.}
-      \textbf{(A)} The assignments of the top 50 representative sequences to their respective taxonomies using the three different reference databases.
-      This result clearly describes how the same sequences are assigned to different Genus under the different databases.
-      A significant portion of the representative sequences are assigned to an ``unknown'' Genus in two of three databases (\ac{gg} and \ac{ncbi}).
-      \textbf{(B)} The number of representative sequences assigned to the same taxonomic label when using different reference databases, shown for the top 100 sequences.
-      The number of mismatches are fewer at higher taxonomic levels, but, even at the Order level there exists greater than 51\% of mismatches, demonstrating the poor agreement in taxonomic labels assigned by the different databases.
-      The data used for the analysis in \textbf{(A, B)} were all the samples (healthy and autistic) from the fecal microbiome transplant dataset.
-      \textbf{(C)} The Bray-Curtis dissimilarity between the expected taxonomy profile and calculated taxonomy profile in the mock datasets shows that there is no singular best choice of database for every dataset, as all the databases have similar performance.
-      However, we still recommend the \ac{gg} database and the Naive Bayes classifier as the defaults for the \ac{micone} pipeline.
-      Due to the popularity of the \ac{gg} database, comparison against existing networks in literature will be easier since many published datasets make use of this database in their analyses.
-      The datasets used for the analysis in \textbf{(C)} were the mock datasets from mockrobiota.
-    }
-    \label{fig:figure4}
-  \end{figure}
-  \FloatBarrier
-  \newpage
-
-  \begin{figure}[H]
-    \centering
-    \includegraphics[width=\textwidth]{figure5.pdf}
-  \end{figure}
-  \begin{figure}[H]
-    \centering
-    \caption{
-      \textbf{Networks generated using different network inference methods show notable differences both in terms of edge-density and connectivity}.
-      \textbf{(A)} The nine different networks (excluding \ac{mldm}) generated by the different network inference methods are very dissimilar.
-      The green links are positive associations and the orange links represent negative associations.
-      A threshold of 0.3 was set for the correlation-based methods (\ac{sparcc}, Propr, Spearman and Pearson) and a threshold of 0.01 was set for the other direct association methods.
-      The correlation-based methods in general produce graph with higher edge-densities.
-      \textbf{(B)} The node overlap Upset plot indicates that all the networks have a large proportion of common nodes involved in connections (33 out of 68).
-      Whereas, \textbf{(C)} the edge overlap Upset plot shows that a very small fraction of these connections are actually shared (8 out of 202).
-      The data used in this analysis were the healthy stool samples from the fecal microbiome transplant dataset.
-    }
-    \label{fig:figure5}
-  \end{figure}
-  \FloatBarrier
-  \newpage
-
-  \begin{figure}[H]
-    \centering
-    \includegraphics[width=1.0\linewidth]{figure6.pdf}
-    \caption{
-      \textbf{The precision and sensitivity of the inferred networks on the NorTA synthetic interaction data.}
-      The ``I'' prefix indicates the individual network inference methods and the ``X'' prefix denotes the consensus based methods that are a part of \ac{micone}.
-      The different consensus based methods used are: pvalue merging (PM), scaled-sum (SS) and simple voting (SV) method.
-      Pearson and Spearman methods are not used in the calculation of the consensus.
-      Among all the independent network inference methods, \ac{spieceasi} has the best average precision (0.944), but the overall best precision was consistently obtained by the scaled-sum method (0.956, 0.985 and 1.000).
-      The simple voting method when using presence of edges in all inferred networks as a requirement (parameter value 1.000), also outperforms \ac{spieceasi} on average precision (0.969).
-      Although \ac{spieceasi} has a higher performance on sensitivity, if the goal of network inference is to obtain the list of associations that have a high probability of existing in the real microbial community, then the consensus methods perform better.
-      Therefore, we recommend the scaled-sum consensus method as the default tool for network inference.
-    }
     \label{fig:figure6}
   \end{figure}
-  \FloatBarrier
-  \newpage
 
-  % TODO: Verify the validity of Figure 7A and 7B again
-  % Are they both using consensus?
-  \begin{figure}[H]
-    \centering
-    \includegraphics[width=1.0\linewidth]{figure7.pdf}
-  \end{figure}
   \begin{figure}[H]
@@
   \end{figure}
+
+  \FloatBarrier
+  \newpage
+
+  \begin{figure}[H]
+    \centering
+    \includegraphics[width=0.74\linewidth]{figure1.pdf}
+  \end{figure}
+
+  \FloatBarrier
+  \newpage
+
+  \begin{figure}[H]
+    \centering
+    \includegraphics[width=\textwidth]{figure2.pdf}
+  \end{figure}
+
+  \FloatBarrier
+  \newpage
+
+  \begin{figure}[H]
+    \centering
+    \includegraphics[width=\textwidth]{figure3.pdf}
+  \end{figure}
+
+  \FloatBarrier
+  \newpage
+
+  \begin{figure}[H]
+    \centering
+    \includegraphics[width=\textwidth]{figure4.pdf}
+  \end{figure}
+
+  \FloatBarrier
+  \newpage
+
+  \begin{figure}[H]
+    \centering
+    \includegraphics[width=1.0\linewidth]{figure5.pdf}
+  \end{figure}
+
+  \FloatBarrier
+  \newpage
+
+  % TODO: Add equations for the linear model to the left of the pie chart
+  \begin{figure}[H]
+    \centering
+    \includegraphics[width=\textwidth]{figure6.pdf}
+  \end{figure}
+
+  \FloatBarrier
+  \newpage
+
+  % TODO: Verify the validity of Figure 7A and 7B again
+  % Are they both using consensus?
+  \begin{figure}[H]
+    \centering
+    \includegraphics[width=1.0\linewidth]{figure7.pdf}
+  \end{figure}

### 2022-08-29 sections/abstract.tex
Change: Update Abstract and Intro for draft v3

@@
     Through this process, we map how different choices of algorithms and parameters affect the co-occurrence network and estimate steps that contribute most significantly to the variance.
-    We further determine the tools and parameters that generate the most accurate and robust co-occurrence networks and develop consensus network algorithms based on benchmarks with mock and synthetic datasets.
-    Ultimately, we develop a standardized pipeline (available at \href{https://github.com/segrelab/MiCoNE}{https://github.com/segrelab/MiCoNE}) that follows these default tools and parameters, but that can also help explore the outcome of any other combination of choices.
+    We further determine the tools and parameters that generate robust co-occurrence networks and develop consensus network algorithms based on benchmarks with mock and synthetic datasets.
+    The Microbial Co-occurrence Network Explorer or \acs{micone} (available at \href{https://github.com/segrelab/MiCoNE}{https://github.com/segrelab/MiCoNE}), follows these default tools and parameters, and can help explore the outcome of these combinations of choices on the inferred networks.
     We envisage that this pipeline could be used for integrating multiple data-sets, and for generating comparative analyses and consensus networks that can help understand microbial community assembly in different biomes.
@@
 \section*{Importance}
-  Mapping the interrelationships between the species in a microbial community is essential to understand and control their structure and function.
+  Mapping the interrelationships between different species in a microbial community is essential to understand and control their structure and function.
   The surge in the high-throughput sequencing of microbial communities has led to the creation of thousands of datasets containing information about microbial abundances.
@@
   However, processing these datasets to obtain co-occurrence information relies on several complex steps, each of which involves numerous choices of tools and corresponding parameters.
-  These multiple options pose questions about the accuracy and uniqueness of the inferred networks.
+  These multiple options pose questions about the robustness and uniqueness of the inferred networks.
   In this study, we address this workflow and provide a systematic analysis of how these choices of tools affect the final network, and guidelines on appropriate tool selection for a particular dataset.
-  We also develop a consensus network algorithm that helps generate more robust and accurate occurrence networks based on benchmarks synthetic datasets.
+  We also develop a consensus network algorithm that helps generate more robust co-occurrence networks based on benchmarks synthetic datasets.

### 2022-08-29 sections/introduction.tex
Change: Update Abstract and Intro for draft v3

@@
  Direct high-throughput measurement of interactions, e.g.
-through co-culture micro-droplet experiments \cite{Hsu2019,Jian2020}, or spatial visualization of natural communities \cite{Wilbert2020} is possible, but it requires specific technological capabilities, and has yet to be extensively used.
+through co-culture micro-droplet experiments~\cite{Hsu2019,Jian2020}, observation of direct contact between cells~\cite{konovalovaCloseEncountersContactdependent2011}, or spatial visualization of natural communities \cite{shiHighlyMultiplexedSpatial2020,hartmannQuantitativeImageAnalysis2021,Wilbert2020} is possible, but it requires specific technological capabilities.
  However, sequencing data across multiple samples, which is more readily available, can be used for estimating co-occurrence relationships between taxa.
@@
 Popular methods for different phases of the analysis of 16S data include tools for: (i) quality checking and trimming the sequencing reads; (ii) denoising and clustering the trimmed reads \cite{Caporaso2010,Callahan2016}; (iii) assigning taxonomy to the denoised reads \cite{DeSantis2006,Quast2012}; (iv) processing and transforming the taxonomy count matrices \cite{Weiss2015}; and (v) inferring the co-occurrence network \cite{Cougoul2019,Kurtz2015}.
-Different specific algorithms are often aggregated into popular platforms (like MG-RAST~\cite{Keegan2016}, Qiita~\cite{qiita}) and packages (such as QIIME2~\cite{bolyenReproducibleInteractiveScalable2019}) that provide pipelines for 16S data analysis.
+Different specific algorithms are often aggregated into popular online platforms (like MG-RAST~\cite{Keegan2016}, Qiita~\cite{qiita}) and software packages (such as QIIME2~\cite{bolyenReproducibleInteractiveScalable2019}) that provide pipelines for 16S data analysis.
 The different methods and tools developed to solve issues arising in 16S analysis can lead to vastly different inferences of community compositions and co-occurrence networks \cite{Golob2017,Weiss2016}, making it difficult to reliably compare networks across different publications and studies.
@@
  In this study, we present a standardized 16S data analysis pipeline called \ac{micone} that produces robust and reproducible co-occurrence networks from community 16S sequence data, and allow users to interactively explore how the network would change upon using different alternative tools and parameters at each step.
-  % TODO: Link or describe MIND here (?)
- Our pipeline is coupled to an online integrative tool for the organization, visualization and analysis of inter-microbial networks [MIND REF], called \ac{mind}, that is available at [link].
+ Our pipeline is coupled to an online integrative tool for the organization, visualization and analysis of inter-microbial networks called \ac{mind}~\cite{huResourceComparisonIntegration2022}, that is available at \href{http://microbialnet.org/}{http://microbialnet.org/}.
   In addition to making this tool freely available, we completed a systematic comparative analysis to determine which steps of the pipeline have the largest influence on the final network, and what choice seems to have the best agreement with the tested mock and synthetic datasets.

### 2022-08-29 sections/discussion.tex
Change: Update Discussion for draft v3

@@
 
-  % \subsection*{Why \ac{micone}?}
+  \subsection*{Why \ac{micone}?}
 
-  Co-occurrence associations in microbial communities help identify important interdependencies that may drive microbial community structure and organization.
+  Co-occurrence associations in microbial communities help identify important interdependencies that may drive or reflect microbial community structure and organization.
   A myriad of tools and methods have been developed for different parts of the workflow for inference of co-occurrence networks from 16S rRNA data.
-  Our analysis shows that networks generated using different combinations of tools and approaches are significantly different from each other, highlighting the need for a clear assessment of the source of variability and for tools that provide the most robust and accurate results.
+  Our analysis shows that networks generated using different combinations of tools and approaches can be significantly different from each other, highlighting the need for a clear assessment of the source of variability and for tools that provide the most robust and accurate results.
   Our newly developed integrated software for the inference of co-occurrence networks from 16S rRNA data, \ac{micone}, is a freely customizable and user-friendly pipeline that allows users to compare networks generated by multiple possible combinations of tools and parameters.
@@
   The \ac{micone} pipeline is built in a modular fashion; its plug-and-play architecture enables users to add new tools and steps, either using existing packages that were not examined in the present work or those developed in the future.
-  The easy-to-use nature, the configurability and modular nature of the \ac{micone} package enables users to start and end the pipeline at any point in the workflow, and to run parts of the pipeline in isolation.
   The \ac{micone} Python package provides functions and methods to perform detailed analysis of the count matrices and the co-occurrence networks.
-  Additionally, the inferred networks are exported to a custom JSON format (see Supplementary), but can also be exported to Cytoscape~\cite{shannonCytoscapeSoftwareEnvironment2003}, GML~\cite{himsoltGMLPortableGraph2010}, and many other popular formats via the \ac{micone} Python package.
+  The inferred networks are exported to a custom JSON format (see Supplementary) by default, but can also be exported to Cytoscape~\cite{shannonCytoscapeSoftwareEnvironment2003}, GML~\cite{himsoltGMLPortableGraph2010}, and many other popular formats via the \ac{micone} Python package.
 
@@
   \ac{micone} is unique as it offers this functionality packaged in a workflow that can be run locally, on the compute cluster or the cloud.
-  For an end user who is interested in identifying the microbial interdependencies present in a specific sequencing dataset, \ac{micone} offers a fast and reproducible way to compare and test novel hypotheses.
 
-  % \subsection*{The default pipeline and recommended tools}
+  \subsection*{The default pipeline and recommended tools}
 
-  The main outcome of this work is two-fold.
-  On one hand, we transparently reveal the dependence of co-occurrence networks on tool and parameter choices, and built a tool for this purpose, enabling to more rigorously assess and compare existing networks.
-  On the other hand, we take advantage of our spectrum of computational options and the availability of mock and synthetic datasets, to suggest a default standard setting, and a consensus approach, that is more likely to yield networks that are robust across multiple tool/parameter choices.
-  An important caveat related to this last point is the fact that due to the lack of a universal standard for microbial interaction data, our conclusions are based on the specific datasets used in our analysis.
+  Through \ac{micone}, in addition to transparently revealing the dependence of co-occurrence networks on tool and parameter choices, we take advantage of our spectrum of computational options and the availability of mock and synthetic datasets, to suggest a default standard setting.
+  Additionally, we develop a consensus approach, that can reliably generate networks that are robust across multiple tool/parameter choices.
+  An important caveat related to these results is the fact that due to the lack of a universal standard for microbial interaction data, our conclusions are based on the specific datasets used in our analysis.
   While our datasets comprise several mock and synthetic datasets that cover a fairly diverse range of abundance distributions and network topologies, datasets that have drastically different distributions may require a re-assessment of the best settings through our pipeline.
 
-  Additionally, it is worth pointing out some additional more specific conclusions stemming from the individual steps of our analysis.
-  The different denoising/clustering methods differ mostly in their identification of sequences that are in low abundances.
-  Hence, they do not have much of an impact on the inferred co-occurrence networks when the sequences of low abundance are removed (Figure~\ref{fig:figure_s1}).
-  Comparison of inferred and expected reference sequences and their abundances in mock community datasets has allowed us to identify \ac{dada2} as the method which best recapitulates the expected sequence composition.
-  For the chimera checking module, we suggest using the remove bimera method since it was developed in conjunction with \ac{dada2} and its performance does not significantly differ from uchime-denovo.
-  For the current work we have decided to focus on the tools most widely used at the time of the analysis.
-  Some tools which were not as widely used (e.g.
-dbOTU3~\cite{Olesen2017}) as well as older popular methods like mothur~\cite{Schloss2009} have not been included in the study, but could be added into the pipelines in future updated analyses.
-
-  The choice of taxonomy database was found to be the most important factor in the inference of microbial co-occurrence networks, contributing $65.4\%$ of the total variance.
-  The frequent changes in the taxonomy nomenclature coupled with the frequency of updates to the various 16S reference databases create inherent differences \cite{Balvociute2017} in taxonomy hierarchies in these databases.
-  Our analysis revealed that no particular reference database performs better than the others across the different mock dataset benchmarks.
-  The default reference database in the pipeline is the \ac{gg} reference database along with the ``Naive Bayes'' classifier as the query tool.
-  The reason for our choice stems from the popularity of the \ac{gg} database [https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6440677/] in taxonomic studies, which would enable easy comparison across datasets.
-  % TODO: Add ref: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6440677/
-  However, for newer studies we recommend using SILVA database because of its size and taxonomic comprehensiveness~\cite{iiRESCRIPtReproducibleSequence2021} and since \ac{gg} has not been updated since 2013.
-  Additionally, we also enable users to use custom databases~\cite{Ritari2015,iiRESCRIPtReproducibleSequence2021} with the BLAST and Naive Bayes classifiers that are incorporated into the pipeline (from \ac{qiime2}).
-  % TODO: Add ref: https://pubmed.ncbi.nlm.nih.gov/32106809/
-  We suggest that that choice of the database should be made based on possible reported or inferred biases in the representation of given biomes in a specific databases~\cite{Balvociute2017,iiRESCRIPtReproducibleSequence2021}.
-
-  The \ac{op} step of the pipeline is second in its contribution to total network variance.
-  This can be attributed to the large number of nodes that are added to the final networks when the filtering is turned off.
-  Additionally, a very large number of nodes also decreases the accuracy of the network inference algorithms for the same sample size~\cite{peschelNetCoMiNetworkConstruction2020} and increases the computational complexity~\cite{tackmannRapidInferenceDirect2019}.
-  We observe that filtering out taxa that are present in low abundances in all samples increases the proportion of taxa in common between taxonomy tables generated using different reference databases (Figure~\ref{fig:figure_s5}), providing another reason for filtering.
-  We also observe that the reduction in the number of taxa leads to better agreement in the networks inferred through different methods (Figure~\ref{fig:figure_s1}).
-  Moreover, filtering is necessary in order to increase the power in tests of significance when the number of taxa is much greater than the number of samples.
-
+  The extending discussion behind our choices and selection of default settings for the DC, TA and OP steps of the pipeline can be found in the supplementary discussion section (see Supplementary).
   The networks generated by different network inference methods show considerable differences in edge-density and connectivity.
@@
 
-  % \subsection*{Future directions}
+  \subsection*{Future directions}
 
+  Future work building upon our current results could enhance the network inference process in multiple ways.
   The current analyses make use of one fecal microbiome transplant dataset with healthy and autistic samples, three mock community datasets and several datasets generated by two synthetic interaction methods.
-  The analyses presented in this study are specific to the datasets utilized, and due to the lack of a gold standard reference dataset for microbial interaction data, it is not possible to generalize these results to datasets with widely different microbiome profiles.
   Incorporating datasets from a broad spectrum of biomes with varying microbial distributions into \ac{micone} will likely increase the robustness and generalizability of the results from these analyses.
@@
   Therefore, the development of a metric of overlap to compare nodes with shared lineages within and across networks could provide more biologically and phylogentically relevant comparisons.
-  Further analyses could also include networks at the Species or other taxonomic levels.
 
-  Additionally, the utilization of literature-based interactions~\cite{lima-mendezDeterminantsCommunityStructure2015a} or biological benchmark interaction data~\cite{sungGlobalMetabolicInteraction2017a} as additional sources for benchmarking.
-  Furthermore, expanding the scope of the \ac{micone} pipeline to also include metagenomics sequencing data will enable an even broader access to a diverse range of datasets as well as domains of life that can be incorporated into the analysis.
+  Futher benchmarking of co-occurrence networks could be pursued through the utilization of literature-based interactions~\cite{lima-mendezDeterminantsCommunityStructure2015a} or biological benchmark interaction data~\cite{sungGlobalMetabolicInteraction2017a}.
+  Furthermore, \ac{micone} could be extended to enable the processing of metagenomics sequencing data, facilitating the analysis of a much larger and diverse range of datasets and domains of life.
 
-  Although the pipeline currently supports changing of parameter values for the most commonly used parameters for several tools, we only use those default parameter values that are recommended by the tool creators.
-  The exploration of the effects of parameter values on the inferred association networks might help optimize these values for better network inference, but these are not part of this current study.
-  Additionally, despite incorporating a wide range of tools and algorithms into the \ac{micone} pipeline, there were several other tools that were omitted due time and computational constraints.
-  However, the modular nature of the \ac{micone} pipeline enables easy addition of new tools and methods at any step of the workflow.
-  Since there exists no ``best method'' for the various steps of 16S data analysis, especially for microbial network inference, \ac{micone} is intended to help researchers to identify the methods and algorithms that are most suitable for their datasets in an easy-to-use and reproducible manner.
+  Although in the current analysis, we only use default parameter values recommended by the tool creators, the \ac{micone} pipeline could be used in the future to explore any combinations of parameters and to optimize these values for improved network inference.
+  Overall, it is important to keep in mind that there likely is no ``best method'' for the various steps of 16S data analysis, especially for the network inference step.
+  Therefore, \ac{micone} is intended to help researchers to identify the methods and algorithms that are most suitable for their datasets in an easy-to-use and reproducible manner.
 
-  We envision that \ac{micone}, and the underlying tools and databases that help process amplicon sequencing data into co-occurrence networks, will be increasingly useful towards building large comparative analyses across studies.
-  Furthermore, in the re-analysis of published networks across studies and processing methods, \ac{micone} could help understand underlying biases imposed on networks due to the selected methods.
-  In summary, we believe that \ac{micone} is a useful tool for the exploration of microbial networks generated from 16S data.
+  We envision that \ac{micone}, and its underlying tools and databases, will be increasingly useful towards building large comparative analyses across studies.
   It allows for rapid, configurable and reproducible inference of microbial networks and further for the formulation of hypotheses about the role of these interactions on community composition and stability.
-  Additionally, our pipeline is coupled to the MIND database and web interface for the organization and visualization of microbial networks, enabling easier comparisons across studies and datasets.
+  These comparative analyses will require coupled network analysis and visualization tools (such as \ac{mind}~\cite{huResourceComparisonIntegration2022}), and will require a more systematic access to datasets, shared in accordance to FAIR standards~\cite{pachecoFAIRRepresentationsMicrobial2022}.

### 2022-08-29 sections/methods.tex
Change: Update Discussion for draft v3

@@
   The inputs to the pipeline by default are raw untrimmed 16S rRNA sequence reads, but the software can be alternatively configured to use trimmed sequences, \ac{otu} tables and other types of intermediate data (see documentation).
+  The easy-to-use nature, the configurability and modular nature of the \ac{micone} package enables users to start and end the pipeline at any point in the workflow, and to run parts of the pipeline in isolation.
   The pipeline supports both paired end and single end reads, and additionally supports independently processing reads from multiple runs and merging the OTU tables in the DC step.
@@
   However, this was fixed by making minor adjustments to the taxonomic names, such as changing Bacteroidota to Bacteroidetes in the SILVA Phylum assignments.
-  The full list of changes can be found in \texttt{figure4ab_data.py} in the data and scripts  repository.
+  The full list of changes can be found in \texttt{figure4ab\_data.py} in the data and scripts  repository.
   The list of tools used in this step, along with their modules and references are provided in Table~\ref{tab:micone_tools}.

### 2022-08-29 sections/supplementary.tex
Change: Update Discussion for draft v3

@@
     The details of the format and information about importing/exporting it and other network formats can be found in the \ac{micone} documentation.
+
+  \subsection*{Supplementary discussion}
+
+  Additionally, it is worth pointing out some additional more specific conclusions stemming from the individual steps of our analysis.
+  The different denoising/clustering methods differ mostly in their identification of sequences that are in low abundances.
+  Hence, they do not have much of an impact on the inferred co-occurrence networks when the sequences of low abundance are removed (Figure~\ref{fig:figure_s1}).
+  Comparison of inferred and expected reference sequences and their abundances in mock community datasets has allowed us to identify \ac{dada2} as the method which best recapitulates the expected sequence composition.
+  For the chimera checking module, we suggest using the remove bimera method since it was developed in conjunction with \ac{dada2} and its performance does not significantly differ from uchime-denovo.
+  For the current work we have decided to focus on the tools most widely used at the time of the analysis.
+  Some tools which were not as widely used (e.g.
+dbOTU3~\cite{Olesen2017}) as well as older popular methods like mothur~\cite{Schloss2009} have not been included in the study, but could be added into the pipelines in future updated analyses.
+
+  The choice of taxonomy database was found to be the most important factor in the inference of microbial co-occurrence networks, contributing $65.4\%$ of the total variance.
+  The frequent changes in the taxonomy nomenclature coupled with the frequency of updates to the various 16S reference databases create inherent differences \cite{Balvociute2017} in taxonomy hierarchies in these databases.
+  Our analysis revealed that no particular reference database performs better than the others across the different mock dataset benchmarks.
+  The default reference database in the pipeline is the \ac{gg} reference database along with the ``Naive Bayes'' classifier as the query tool.
+  The reason for our choice stems from the popularity of the \ac{gg} database~\cite{parkEvaluation16SRRNA2018} in taxonomic studies, which would enable easy comparison across datasets.
+  However, for newer studies we recommend using SILVA database because of its size and taxonomic comprehensiveness~\cite{iiRESCRIPtReproducibleSequence2021} and since \ac{gg} has not been updated since 2013.
+  Additionally, we also enable users to use custom databases~\cite{Ritari2015,iiRESCRIPtReproducibleSequence2021} with the BLAST and Naive Bayes classifiers that are incorporated into the pipeline (from \ac{qiime2}).
+  We suggest that that choice of the database should be made based on possible reported or inferred biases in the representation of given biomes in a specific databases~\cite{Balvociute2017,iiRESCRIPtReproducibleSequence2021}, as choosing taxon-specific databases have also been observed to compromise classification~\cite{rmarcelinoUseTaxonspecificReference2020}.
+
+  The \ac{op} step of the pipeline is second in its contribution to total network variance.
+  This can be attributed to the large number of nodes that are added to the final networks when the filtering is turned off.
+  Additionally, a very large number of nodes also decreases the accuracy of the network inference algorithms for the same sample size~\cite{peschelNetCoMiNetworkConstruction2020} and increases the computational complexity~\cite{tackmannRapidInferenceDirect2019}.
+  We observe that filtering out taxa that are present in low abundances in all samples increases the proportion of taxa in common between taxonomy tables generated using different reference databases (Figure~\ref{fig:figure_s5}), providing another reason for filtering.
+  We also observe that the reduction in the number of taxa leads to better agreement in the networks inferred through different methods (Figure~\ref{fig:figure_s1}).
+  Moreover, filtering is necessary in order to increase the power in tests of significance when the number of taxa is much greater than the number of samples.

### 2022-08-29 sections/methods.tex
Change: Update Methods for draft v3

@@
   The current analysis used stool samples from a fecal microbiome transplant study of autism~\cite{Kang2017} as biological datasets.
-    The study was composed of multiple sequencing runs and those runs that contained paired end reads (run 2 (10M reads), run 3 (750K reads) and run 4 (16M reads)), were downloaded from Qiita~\cite{qiita} (study ID 10532) and used as input sequences for the \ac{micone} pipeline.
+  The study was composed of multiple sequencing runs.
+  The runs that contained paired end reads (run 2 (10M reads), run 3 (750K reads) and run 4 (16M reads)), were downloaded from Qiita~\cite{qiita} (study ID 10532) and used as input sequences for the \ac{micone} pipeline.
   Sequences from both control (212 samples including neurotypical and donors) and autistic (126 samples) patients were included in the analyses, but the groups were analyzed independently in the pipeline after the \ac{op} step.
@@
   However, all other analysis (preceding the NI step) are performed together on all the samples in the fecal microbiome transplant study.
-    The mock community 16S datasets are sequencing data obtained for artificially assembled collections of species in known proportions.
+  The mock community 16S datasets are experimental sequencing data obtained for artificially assembled collections of species in known proportions.
   The mock datasets used for this study, obtained from mockrobiota~\cite{Bokulich2016}, are labelled mock4, mock12 and mock16.
@@
 
-  The \ac{micone} pipeline provides both a Python API together with a command-line interface and only uses a single configuration file (nextflow.config) to contain the configuration parameters.
-  The \ac{micone} Python API provides several biom (\ac{otu} table) and network related functions and methods, enabling detailed comparison of counts tables and inferred networks if desired.
+  The \ac{micone} pipeline provides both a Python API together with a command-line interface and only uses a single configuration file (\texttt{nextflow.config}) to encode the configuration parameters.
+  The \ac{micone} Python API provides several \ac{otu} table and network related functions and methods, enabling detailed comparison of counts tables and inferred networks if desired.
   Exploring the effects of these combinations of methods on the resultant networks is difficult and inconvenient since different tools differ in their input and output formats and require interconversions between the various formats.
@@
   It also contains helper functions that can help in parsing taxonomies and communicate with the \ac{ncbi} taxonomy database to query taxonomy by name or taxonomic IDs.
-  The configuration file along with the run file (main.nf) lists the inputs, output and the steps to be performed during runtime, along with the parameters to be used (if different from defaults) for the various steps.
+  The configuration file along with the run file (\texttt{main.nf}) lists the inputs, output and the steps to be performed during runtime, along with the parameters to be used (if different from defaults) for the various steps.
   Since the entire pipeline run is stored in the form of a text file (the configuration file), subsequent runs are highly reproducible and changes can be easily tracked using version control.
@@
   It consists of the demultiplexing and trimming processes.
+  The demultiplexing process deals with separating the multiplexed sequences into individual samples based on barcodes.
   The trimming process handles the quality control steps such as trimming adapters and trimming low-quality nucleotide stretches from the sequences.
@@
   Filtering, is performed to remove samples or features (\ac{otu}s or \ac{esv}s) from the counts matrix that are sparse.
-  By default, when the OP module is ``on'', the samples are filtered out if the total reads in a sample are less than 500 and features are filtered out if the abundance is less than 1\%, prevalence is less than 5\% and count sum across all the samples is less than 100.
+  By default, when the OP module is ``on'', the samples are filtered out if the total reads in a sample are less than 500 and features are filtered out if the relative abundance is less than 1\%, prevalence (percentage of samples containing feature) is less than 5\% and count sum across all the samples is less than 100.
   When the OP module is ``off'', the filtering is still performed but only to remove those features that are very low in presence.
   The parameters used are given in Table~\ref{tab:micone_tools}.
-  The forking operation splits the count matrix into multiple matrices based on sample metadata column.
-  The group operation transforms the \ac{otu} or \ac{esv} count matrix into a taxonomic count matrix at the requested level by adding up counts that map to the same taxonomy.
+  The forking operation splits the count matrix into multiple matrices based on sample metadata column, this is useful for example to compare case vs. control.
+  The group operation transforms the \ac{otu} or \ac{esv} count matrix into a taxonomic count matrix at the requested level by adding up counts that map to the same taxonomy and is carried out at the end of the OP step.
   Finally, transformations are performed in order to correct for and overcome the compositional bias that is inherent in the counts matrix (in the analysis performed in the study these are handled by the network inference algorithm).
@@
 
-  A null model is created by re-sampling and bootstrapping the counts matrix and recalculating the correlations.
+  A null model is created by re-sampling and bootstrapping the counts matrix and recalculating the correlations (see next section for details on network analysis and statistics).
   These bootstrapped association matrices are used to calculate the significance of the inferred correlations by calculating the p-values against this null model~\cite{Watts2018}.
@@
 
-  $w$, the number of co-occurrence networks to be integrated into the consensus network (corresponding to the number of the different approaches used to infer co-occurrence networks excluding Spearman and Pearson)
+  $w$, the number of co-occurrence networks to be integrated into the consensus network (corresponding to the number of the different approaches used to infer co-occurrence networks excluding Spearman and Pearson, $k$)
 
@@
   \vspace{-5mm}
-  For all correlation-based methods $k \subset w$, $1000$ permutations of the original \ac{otu} counts data were generated~\cite{Watts2018}.
+  For all correlation-based methods $k < w$, $1000$ permutations of the original \ac{otu} counts data were generated~\cite{Watts2018}.
   All the direct association based methods used in the study have their own regularization methods built-in and hence do not need to undergo bootstrapping.
@@
   The simple voting method performs a voting-based consensus to determine whether an edge will exist between a given node-pair in the final consensus network~\cite{bustinceFuzzySetsTheir2008,tsarevApplicationMajorityVoting2018}.
-  For each pair of nodes, we determine the number of network inference methods that report an edge $j=l$ between them, i.e. $\bar{N}^{i}_{j=l}, \forall i \in \{1,\dots,w\}$.
+  For each pair of nodes, we determine the number of network inference methods that report an edge $j$ between them, i.e. $\bar{N}^{i}_{j}, \forall i \in \{1,\dots,w\}$.
   Each node-pair will have an edge in the final consensus network if the number of reported edges is larger than a user-defined threshold (Equation~\ref{eqn:simple-voting2}).
@@
  For each edge $j$, we obtain $M_j$ which represents the number of networks in which edge $j$ is reported.
+ Formally, $M_j$ is calculated as the following function:
   \begin{equation}
@@
 
- The edge $j$ appears in the final consensus network if the number of reported edges is greater than the threshold.
+ The edge $j$ is selected to be present in the final consensus network if the number of networks in which $j$ appears is greater than a threshold, i.e:
   \begin{equation}
@@
   \begin{multline}
-  C_j = &\sum_{k=1}^5 \left( \alpha^{DC(k)}_j.\delta^{DC(k)}_j \right) +
-         \sum_{k=1}^2 \left( \alpha^{CC(k)}_j.\delta^{CC(k)}_j \right) +
-         \sum_{k=1}^3 \left( \alpha^{TA(k)}_j.\delta^{TA(k)}_j \right) +
-         \sum_{k=1}^2 \left( \alpha^{OP(k)}_j.\delta^{OP(k)}_j \right) + \\
-         &\sum_{k=1}^{10} \left( \alpha^{NI(k)}_j.\delta^{NI(k)}_j \right) +
+  C_j = &\sum_{i=1}^5 \left( \alpha^{DC(i)}_j.\delta^{DC(i)}_j \right) +
+         \sum_{i=1}^2 \left( \alpha^{CC(i)}_j.\delta^{CC(i)}_j \right) +
+         \sum_{i=1}^3 \left( \alpha^{TA(i)}_j.\delta^{TA(i)}_j \right) +
+         \sum_{i=1}^2 \left( \alpha^{OP(i)}_j.\delta^{OP(i)}_j \right) + \\
+         &\sum_{i=1}^{10} \left( \alpha^{NI(i)}_j.\delta^{NI(i)}_j \right) +
          \epsilon_j
@@
    \begin{enumerate}
-     \item $DC(k) \subset$ \{CR, OR, DN, D2, DB\}
-     \item $CC(k) \subset$ \{remove bimera, uchime-denovo\}
-     \item $TA(k) \subset$ \{NaiveBayes(GG), NaiveBayes(SILVA), BLAST(NCBI)\}
-     \item $OP(k) \subset$ \{Filter(on), Filter(off)\}
-     \item $NI(k) \subset$ \{\ac{sparcc}, propr, Spearman, Pearson, \ac{spieceasi}, \ac{cozine}, \ac{harmonies}, \ac{spring}, \ac{mldm}, FlashWeave\}
+     \item $DC(i) \in$ \{CR, OR, DN, D2, DB\}
+     \item $CC(i) \in$ \{remove bimera, uchime-denovo\}
+     \item $TA(i) \in$ \{NaiveBayes(GG), NaiveBayes(SILVA), BLAST(NCBI)\}
+     \item $OP(i) \in$ \{Filter(on), Filter(off)\}
+     \item $NI(i) \in$ \{\ac{sparcc}, propr, Spearman, Pearson, \ac{spieceasi}, \ac{cozine}, \ac{harmonies}, \ac{spring}, \ac{mldm}, FlashWeave\}
    \end{enumerate}
@@
   In order to compare the representative sequences generated by the various methods in the DC step, we employ both the weighted~\cite{Lozupone2007} (Figure~\ref{fig:figure2}A) and unweighted UniFrac method~\cite{Lozupone2005} (Figure~\ref{fig:figure2}B).
-  The UniFrac distance metric (unique fraction metric) is a beta-diversity measure that computes the distance between two sets of taxa as the fraction of the branch length of the tree that leads to descendants from either one environment or the other, but not both~\cite{Lozupone2005}
+  The UniFrac distance metric (unique fraction metric) is a beta-diversity measure that computes the distance between two sets of taxa as the fraction of the branch length of the tree that leads to descendants from either one environment or the other, but not both~\cite{Lozupone2005}.
   The weighted UniFrac distance metric takes into account the abundances of the representative sequences when calculating shared and unshared branch lengths, whereas the unweighted UniFrac distance metric does not and hence gives equal weights to each sequence.
@@
   In Figure~\ref{fig:figure3}C we use the Bray-Curtis distance metric to calculate the distance between the predicted (using the taxonomy databases in the TA step) and expected taxonomic distribution.
-  The Bray-Curtis distance is used to quantify the compositional dissimilarity between two different taxonomic distributions.
+  The Bray-Curtis distance is used to quantify the compositional dissimilarity between two different taxonomic distributions defined by vectors $u$ and $v$.
   It is defined as:
@@
   \vspace{-5mm}
-  In Figure~\ref{fig:figure5} we calculate the accuracy and robustness of the inferred association networks (using the various network inference algorithms and the consensus methods) against the original interaction used to create the taxonomic distribution.
+  In Figure~\ref{fig:figure5} we calculate the accuracy and robustness of the inferred association networks (using the various network inference algorithms and the consensus methods) against the original interaction network used to create the taxonomic distribution.
   We use precision and sensitivity to calculate the accuracy of the predictions.

### 2022-08-30 scripts/figure4/figure4_plot.R
Fix: Changes based on reviewer comments

@@ -187,5 +187,5 @@ ggsave(output_file_c, width = 11, height = 5)
 
 ################################################################################
 # Combine figures
-ggarrange(final_plot_a, combined_plot_b, dot_plot_facet_c, nrow = 3, ncol = 1, heights = c(2, 1, 1))
+ggarrange(final_plot_a, combined_plot_b, dot_plot_facet_c, nrow = 3, ncol = 1, heights = c(2, 1, 1), labels = c("A", "B", "C"))
 ggsave(output_file, width = 11, height = 16)

### 2022-09-01 sections/discussion.tex
Change: Update manuscript based on reviewer comments

@@
 
-  Futher benchmarking of co-occurrence networks could be pursued through the utilization of literature-based interactions~\cite{lima-mendezDeterminantsCommunityStructure2015a} or biological benchmark interaction data~\cite{sungGlobalMetabolicInteraction2017a}.
-  Furthermore, \ac{micone} could be extended to enable the processing of metagenomics sequencing data, facilitating the analysis of a much larger and diverse range of datasets and domains of life.
+  Further, benchmarking of co-occurrence networks could also be pursued through the utilization of literature-based interactions~\cite{lima-mendezDeterminantsCommunityStructure2015a} or biological benchmark interaction data~\cite{sungGlobalMetabolicInteraction2017a}.
+  However, direct comparisons between co-occurrence networks and directly measured interactions are difficult to interpret and highly debated~\cite{hiranoDifficultyInferringMicrobial2019,gobernaCautionaryNotesUse2022}.
+  Additionally, \ac{micone} could be extended to enable the processing of metagenomics sequencing data, facilitating the analysis of a much larger and diverse range of datasets and domains of life.
 

### 2022-09-01 sections/methods.tex
Fix: Changes based on reviewer comments
Change: Update manuscript based on reviewer comments

@@
   The module also supports rarefaction, which is a normalization technique used to overcome the bias that might arise due to variable sampling depth in different samples.
+  However, due to the potential biases and false positives~\cite{mcmurdieWasteNotWant2014,chaoCoveragebasedRarefactionExtrapolation2012} that might arise during the process, the rarefaction module is disabled by default and can be enabled in the configuration if needed.
   This is performed either by sub-sampling or by normalization of the matrix to the lowest sampling depth~\cite{Weiss2015}.
@@
   \vspace{-5mm}
-  For all correlation-based methods $k < w$, $1000$ permutations of the original \ac{otu} counts data were generated~\cite{Watts2018}.
+  For all correlation-based methods $k \le w$, $1000$ permutations of the original \ac{otu} counts data were generated~\cite{Watts2018}.
   All the direct association based methods used in the study have their own regularization methods built-in and hence do not need to undergo bootstrapping.
@@
 
-  \subsection*{Network Variability}
+  \subsection*{Network variability}
 

### 2022-09-01 sections/results.tex
Change: Update manuscript based on reviewer comments

@@
   The TA step results in a taxonomic counts table that is used as input to the subsequent steps of the pipeline.
-  The count tables at different levels can be obtained by aggregation, for example, Genus count tables were obtained by summing up the counts of the lower taxonomy levels (Species and \ac{otu}) that map to the same higher taxonomy level name.
+  The count tables at different levels can be obtained through aggregation, for example, Genus count tables were obtained by summing up the counts of the lower taxonomy levels (Species and \ac{otu}) that map to the same higher taxonomy level name.
 
@@
   Similar to what was done for the previous steps of the pipeline, we evaluated the network inference algorithms and the final consensus network using synthetic interaction data.
-  For this purpose, we generated synthetic interaction data using the ``NorTA''~\cite{Kurtz2015} and ``seqtime''~\cite{faustSignaturesEcologicalProcesses2018} methods (see Methods), which have been previously used for benchmarking network inference methods.
+  For this purpose, we generated synthetic interaction data using the ``NorTA''~\cite{Kurtz2015} and ``seqtime''~\cite{faustSignaturesEcologicalProcesses2018} methods (see Methods), which have been previously used for benchmarking network inference methods~\cite{Kurtz2015,Weiss2016}.
   For each method, an \ac{otu} counts table is generated based on the selected interaction network topology and properties.

### 2022-09-01 sections/supplementary.tex
Fix: Changes based on reviewer comments
Change: Update manuscript based on reviewer comments

@@
 \section*{Supplementary}
+
+  {\color{red} Tables S1-S5 will be moved to GitHub}
 
@@
   However, for newer studies we recommend using SILVA database because of its size and taxonomic comprehensiveness~\cite{iiRESCRIPtReproducibleSequence2021} and since \ac{gg} has not been updated since 2013.
-  Additionally, we also enable users to use custom databases~\cite{Ritari2015,iiRESCRIPtReproducibleSequence2021} with the BLAST and Naive Bayes classifiers that are incorporated into the pipeline (from \ac{qiime2}).
+  Additionally, a particular database might be more appropriate than the rest based on specific requirements.
+  For example, in order to generate a dataset that is compatible with the \ac{mind} platform~\cite{huResourceComparisonIntegration2022} \ac{ncbi} is the most appropriate choice as it guarantees compatibility of taxonomic hierarchy and therefore comparability with other datasets.
+  Furthermore, we also enable users to use custom databases~\cite{Ritari2015,iiRESCRIPtReproducibleSequence2021} with the BLAST and Naive Bayes classifiers that are incorporated into the pipeline (from \ac{qiime2}).
   We suggest that that choice of the database should be made based on possible reported or inferred biases in the representation of given biomes in a specific databases~\cite{Balvociute2017,iiRESCRIPtReproducibleSequence2021}, as choosing taxon-specific databases have also been observed to compromise classification~\cite{rmarcelinoUseTaxonspecificReference2020}.

### 2022-09-05 scripts/pdf2eps.sh
Add: Script to convert pdf to eps

@@ -1,0 +1,10 @@
+#!/bin/sh
+# $Id: pdf2eps,v 0.01 2005/10/28 00:55:46 Herbert Voss Exp $
+# Convert PDF to encapsulated PostScript.
+# usage:
+# pdf2eps <page number> <pdf file without ext>
+
+pdfcrop $2.pdf
+pdftops -f $1 -l $1 -eps "$2-crop.pdf"
+rm "$2-crop.pdf"
+mv "$2-crop.eps" $2.eps

### 2022-09-08 sections/discussion.tex
Update on Overleaf.

@@
 
-  Co-occurrence associations in microbial communities help identify important interdependencies that may drive or reflect microbial community structure and organization.
   A myriad of tools and methods have been developed for different parts of the workflow for inference of co-occurrence networks from 16S rRNA data.
@@
   The \ac{micone} pipeline is built in a modular fashion; its plug-and-play architecture enables users to add new tools and steps, either using existing packages that were not examined in the present work or those developed in the future.
-  The \ac{micone} Python package provides functions and methods to perform detailed analysis of the count matrices and the co-occurrence networks.
+  The \ac{micone} Python package provides functions and methods to perform a detailed analysis of the count matrices and the co-occurrence networks.
   The inferred networks are exported to a custom JSON format (see Supplementary) by default, but can also be exported to Cytoscape~\cite{shannonCytoscapeSoftwareEnvironment2003}, GML~\cite{himsoltGMLPortableGraph2010}, and many other popular formats via the \ac{micone} Python package.
@@
   While several tools/workflows such as \ac{qiime2}~\cite{bolyenReproducibleInteractiveScalable2019} and NetCoMi~\cite{peschelNetCoMiNetworkConstruction2020} can be used to generate co-occurrence networks from 16S sequencing data, no single tool exist that integrates the complete process of inferring microbial interaction networks from 16S sequencing reads.
-  \ac{micone} is unique as it offers this functionality packaged in a workflow that can be run locally, on the compute cluster or the cloud.
+  \ac{micone} is unique as it offers this functionality packaged in a workflow that can be run locally, on the compute cluster, or in the cloud.
 
@@
 
-  Through \ac{micone}, in addition to transparently revealing the dependence of co-occurrence networks on tool and parameter choices, we take advantage of our spectrum of computational options and the availability of mock and synthetic datasets, to suggest a default standard setting.
+  Through \ac{micone}, in addition to transparently revealing the dependence of co-occurrence networks on tool and parameter choices (see Discussion in Supplementary Text for details on the DC, TA and OP steps), we take advantage of our spectrum of computational options and the availability of mock and synthetic datasets, to suggest a default standard setting.
   Additionally, we develop a consensus approach, that can reliably generate networks that are robust across multiple tool/parameter choices.
   An important caveat related to these results is the fact that due to the lack of a universal standard for microbial interaction data, our conclusions are based on the specific datasets used in our analysis.
-  While our datasets comprise several mock and synthetic datasets that cover a fairly diverse range of abundance distributions and network topologies, datasets that have drastically different distributions may require a re-assessment of the best settings through our pipeline.
+  While our analysis is based on several mock and synthetic datasets that cover a diverse range of abundance distributions and network topologies, datasets that have drastically different distributions may require a re-assessment of the best settings.
 
-  The extending discussion behind our choices and selection of default settings for the DC, TA and OP steps of the pipeline can be found in the supplementary discussion section (see Supplementary).
-  The networks generated by different network inference methods show considerable differences in edge-density and connectivity.
-  One reason for this is the underlying assumptions regarding sparsity, distribution and compositionality, that the algorithms make.
-  The consensus network created by merging the networks inferred from the different network inference methods enables the creation of a network whose links have evidence based on multiple inference algorithms.
-  We have developed two consensus algorithms to support this effort, the simple voting method and the scaled-sum method.
-  Ultimately, we find that the scaled-sum method along with p-value merging (applied to the correlation-based methods) performs the best on synthetic datasets, therefore this is the default for the \ac{ni} step of the pipeline.
-  The consensus algorithm not only leads to the creation of a network that has a higher number of true positives, it also decreases the variability in the networks due to choosing a different tool or parameter in any of the previous steps of the pipeline (Figure~\ref{fig:figure7}B).
-  This provides the means to obtain a short list of associations, each of which, have a high probability of being present in the real interaction network
-  This small set of associations could be further tested and validated experimentally to verify their plausibility.
+  The networks generated by different network inference methods show considerable differences in edge-density and connectivity, partially due to the underlying assumptions regarding sparsity, distribution and compositionality.
+  To address this issue, we have developed two consensus algorithms (simple voting and scaled-sum method) that generate networks whose links have evidence based on multiple inference algorithms.
+
+ We find that the scaled-sum method (along with p-value merging for correlation-based methods) performs the best on synthetic datasets, and is therefore chosen as default for the \ac{ni} step of the pipeline.
+Notably, the consensus network displays a higher number of true positives, while also decreasing the overall variability across the whole pipeline (Figure~\ref{fig:figure7}B).
+The concise list of robust associations revealed by the consensus algorithm represents a valuable set for experimental validation follow-up.
 
@@
   Future work building upon our current results could enhance the network inference process in multiple ways.
-  The current analyses make use of one fecal microbiome transplant dataset with healthy and autistic samples, three mock community datasets and several datasets generated by two synthetic interaction methods.
+  The current analyses make use of one fecal microbiome transplant dataset with healthy and autistic samples, three mock community datasets, and several datasets generated by two synthetic interaction methods.
   Incorporating datasets from a broad spectrum of biomes with varying microbial distributions into \ac{micone} will likely increase the robustness and generalizability of the results from these analyses.
 
-  The network analyses in this study are primarily at the Genus level, wherein the lowest resolution of a node is a Genus and if an entity cannot be resolved to the Genus level and next lowest taxonomic level is used (e.g.
+  The network analyses in this study are primarily at the Genus level, wherein the lowest resolution of a node is a Genus and if an entity cannot be resolved to the Genus level, the next lowest taxonomic level is used (e.g.
 Family).
   As a consequence, two entities belonging to the same lineage where one entity is resolved to the Genus level and another is resolved to the Family level are treated as two different nodes in the network.
-  Therefore, the development of a metric of overlap to compare nodes with shared lineages within and across networks could provide more biologically and phylogentically relevant comparisons.
+  Therefore, the development of a metric of overlap to compare nodes with shared lineages within and across networks could provide more biologically and phylogenetically relevant comparisons.
 
+  Although direct comparisons between co-occurrence networks and directly measured interactions are difficult to interpret and highly debated~\cite{hiranoDifficultyInferringMicrobial2019,gobernaCautionaryNotesUse2022}.
   Further, benchmarking of co-occurrence networks could also be pursued through the utilization of literature-based interactions~\cite{lima-mendezDeterminantsCommunityStructure2015a} or biological benchmark interaction data~\cite{sungGlobalMetabolicInteraction2017a}.
-  However, direct comparisons between co-occurrence networks and directly measured interactions are difficult to interpret and highly debated~\cite{hiranoDifficultyInferringMicrobial2019,gobernaCautionaryNotesUse2022}.
   Additionally, \ac{micone} could be extended to enable the processing of metagenomics sequencing data, facilitating the analysis of a much larger and diverse range of datasets and domains of life.
@@
   Although in the current analysis, we only use default parameter values recommended by the tool creators, the \ac{micone} pipeline could be used in the future to explore any combinations of parameters and to optimize these values for improved network inference.
-  Overall, it is important to keep in mind that there likely is no ``best method'' for the various steps of 16S data analysis, especially for the network inference step.
+  Overall, it is important to keep in mind that there likely is no ``best method'' for the various steps of 16S data analysis.
   Therefore, \ac{micone} is intended to help researchers to identify the methods and algorithms that are most suitable for their datasets in an easy-to-use and reproducible manner.
 
-  We envision that \ac{micone}, and its underlying tools and databases, will be increasingly useful towards building large comparative analyses across studies.
-  It allows for rapid, configurable and reproducible inference of microbial networks and further for the formulation of hypotheses about the role of these interactions on community composition and stability.
-  These comparative analyses will require coupled network analysis and visualization tools (such as \ac{mind}~\cite{huResourceComparisonIntegration2022}), and will require a more systematic access to datasets, shared in accordance to FAIR standards~\cite{pachecoFAIRRepresentationsMicrobial2022}.
+  We envision that \ac{micone}, and its underlying tools and databases, will be increasingly useful for building large comparative analyses across studies.
+  It allows for rapid, configurable, and reproducible inference of microbial networks and further for the formulation of hypotheses about the role of these interactions on community composition and stability.
+  These comparative analyses will require coupled network analysis and visualization tools (such as \ac{mind}~\cite{huResourceComparisonIntegration2022}) and will require more systematic access to datasets, shared in accordance with FAIR standards~\cite{pachecoFAIRRepresentationsMicrobial2022}.

### 2022-09-08 sections/introduction.tex
Update on Overleaf.

@@
 
-  Microbial communities are ubiquitous and play an important role in marine and terrestrial environments, urban ecosystems, and human health \cite{Ghoul2016,Thompson2017}.
-  These microbial communities, or microbiomes, often comprise several hundreds of different microbial strains interacting with each other and their environment, often through intricate metabolic and signaling relationships.
+  Microbial communities are ubiquitous and play an important role in marine and terrestrial environments, urban ecosystems, and human health \cite{Ghoul2016,Thompson2017} [MORE REFS].
+  These microbial communities, or microbiomes, often comprise several hundreds of different microbial strains interacting with each other and their environment, often through intricate metabolic and signaling relationships [REFS].
   Understanding how these interconnections shape community structure and functionalities is a fundamental challenge in microbial ecology, and has applications in the study of microbial ecosystems across different biomes.
@@
   In particular, high-throughput sequencing, including metagenomic sequencing and sequencing of 16S rRNA gene amplicons (hereafter referred to as 16S data) of microbial communities, has the potential to help detect, identify and quantify a large portion of the constitutive microorganisms of a microbiome \cite{Jovel2016,Lloyd-Price2016}.
-  These advances have led to large-scale data collection efforts involving environmental (\acl{emp}) \cite{Thompson2017}, marine (Tara Oceans Project) \cite{Zhang2015} and human-associated microbiota (Human Microbiome Project) \cite{HumanMicrobiomeProjectConsortium2012}.
+  These advances have led to large-scale data collection efforts involving terrestrial (\acl{emp}) \cite{Thompson2017}, marine (Tara Oceans Project) \cite{Zhang2015} and human-associated microbiota (Human Microbiome Project) \cite{HumanMicrobiomeProjectConsortium2012}.
+[3 REFS for each]
 
- This wealth of information on the composition and functions of a community at different times and under different environmental conditions has the potential to help us understand how communities assemble and operate.
- A powerful tool for translating microbiome composition data into knowledge is the construction of possible inter-dependence networks across species.
- The importance of these networks of relationships is two-fold: first, such networks can serve as maps that help identify hubs of keystone species \cite{Menon2018,Rottjers2018}, and rewiring of interactions that occur as a consequence of environmental perturbations or underlying host conditions \cite{Gilbert2016}; second, networks of interdependencies can serve as a key bridge towards building mechanistic models of microbial communities, greatly enhancing our capacity to understand and control them.
- For example, multiple studies have shown the importance of specific microbial interactions in the healthy microbiome \cite{Lloyd-Price2016} and others have shown how changes in these interactions can lead to dysbiosis \cite{Wang2017,Gilbert2016,Belizario2015}.
- In the context of terrestrial bio-geochemistry, co-occurrence networks have been proposed as a valuable approach towards reconstructing the processes leading to microbiome assembly \cite{Fierer2017}, and understanding the response of microbial communities to environmental perturbations \cite{Jiao2019}.
+ This wealth of information has the potential to help us understand how communities assemble and operate.
+In particular, a powerful tool for translating microbiome composition data into knowledge is the construction of association (or co-occurrence, or correlation) networks, in which  microbial taxa are represented by nodes, and frequent co-occurrences (or negative correlations) across datasets are encoded in edges between nodes.
+ While the relationship between directly measured interactions [REFS] and statistically inferred co-occurrence is still poorly understood \cite{Zuniga2017}, a significant amount of effort has gone into estimating correlations from large microbiome sequence datasets.
 
- Direct high-throughput measurement of interactions, e.g.
-through co-culture micro-droplet experiments~\cite{Hsu2019,Jian2020}, observation of direct contact between cells~\cite{konovalovaCloseEncountersContactdependent2011}, or spatial visualization of natural communities \cite{shiHighlyMultiplexedSpatial2020,hartmannQuantitativeImageAnalysis2021,Wilbert2020} is possible, but it requires specific technological capabilities.
- However, sequencing data across multiple samples, which is more readily available, can be used for estimating co-occurrence relationships between taxa.
- While the relationship between directly measured interactions and statistically inferred co-occurrence is still poorly understood \cite{Zuniga2017}, a significant amount of effort has gone into estimating correlations from large microbiome sequence datasets.
- Co-occurrence networks have microbial taxa as nodes, and edges that represent the frequent co-occurrences (or negative correlations) across different datasets.
+ The importance of these networks is two-fold: first, they can serve as maps that help identify hubs of keystone species \cite{Menon2018,Rottjers2018}, and the community response to environmental perturbations or underlying host conditions \cite{Gilbert2016}; second, they can serve as a bridge towards building mechanistic models of microbial communities, greatly enhancing our capacity to understand and control them.
+ For example, multiple studies have shown the importance of specific microbial  associations in the healthy microbiome \cite{Lloyd-Price2016} and the rise of dysbiosis \cite{Wang2017,Gilbert2016,Belizario2015}.
+ In the context of terrestrial bio-geochemistry, co-occurrence networks were shown to help understand microbiome assembly \cite{Fierer2017}, and  the response of microbial communities to environmental perturbations \cite{Jiao2019}.
 
 One of the most frequently used avenues for inferring co-occurrence networks is the parsing and analysis of 16S sequencing data \cite{Rottjers2018,Friedman2012}.
-Numerous software tools and pipelines have been developed to analyze 16S sequencing data, often focused on addressing the many known limitations of this methodology, including resolution, sequencing depth, compositional nature, sequencing errors and copy number variations \cite{Bharti2019,Pollock2018}.
+Numerous software tools and pipelines have been developed to analyze 16S sequencing data, with strong emphasis on the known limitations of this methodology, including resolution, sequencing depth, compositional nature, sequencing errors and copy number variations \cite{Bharti2019,Pollock2018}.
 Popular methods for different phases of the analysis of 16S data include tools for: (i) quality checking and trimming the sequencing reads; (ii) denoising and clustering the trimmed reads \cite{Caporaso2010,Callahan2016}; (iii) assigning taxonomy to the denoised reads \cite{DeSantis2006,Quast2012}; (iv) processing and transforming the taxonomy count matrices \cite{Weiss2015}; and (v) inferring the co-occurrence network \cite{Cougoul2019,Kurtz2015}.
-Different specific algorithms are often aggregated into popular online platforms (like MG-RAST~\cite{Keegan2016}, Qiita~\cite{qiita}) and software packages (such as QIIME2~\cite{bolyenReproducibleInteractiveScalable2019}) that provide pipelines for 16S data analysis.
+Different specific algorithms are often aggregated into popular online platforms (like MG-RAST~\cite{Keegan2016}, Qiita~\cite{qiita}) and software packages (such as QIIME2~\cite{bolyenReproducibleInteractiveScalable2019}).
 The different methods and tools developed to solve issues arising in 16S analysis can lead to vastly different inferences of community compositions and co-occurrence networks \cite{Golob2017,Weiss2016}, making it difficult to reliably compare networks across different publications and studies.
- This is partially due to the fact that existing platforms are typically focused on \ac{otu} generation and not on the effects of upstream statistical methods on the inferred co-occurrence networks.
+ This is partially due to the fact that existing platforms are typically focused on \ac{otu} or \ac{esv} generation and not on the effects of upstream statistical methods on the inferred co-occurrence networks.
  Furthermore, no organized framework currently exist to systematically analyze and compare existing components of the data analysis from amplicons to networks.
-  More broadly, given the lack of comprehensive comparisons between directly observed microbial interactions (e.g.
-from co-culture experiments) and co-occurrence networks, there is no straightforward way to determine which set of tools or methods generate the most accurate networks.
 
@@
  Our pipeline is coupled to an online integrative tool for the organization, visualization and analysis of inter-microbial networks called \ac{mind}~\cite{huResourceComparisonIntegration2022}, that is available at \href{http://microbialnet.org/}{http://microbialnet.org/}.
-  In addition to making this tool freely available, we completed a systematic comparative analysis to determine which steps of the pipeline have the largest influence on the final network, and what choice seems to have the best agreement with the tested mock and synthetic datasets.
-  We believe that these steps will ensure better reproducibility and easier comparison of co-occurrence networks across datasets.
+Through a systematic comparative analysis we determined which steps of the \ac{micone} pipeline have the largest influence on the final network, and what choice seems to have the best agreement with the tested mock and synthetic datasets.
+These steps will ensure better reproducibility and easier comparison of co-occurrence networks across datasets.
 We expect that our tool will also be useful for benchmarking future alternative methods, and for ensuring a transparent evaluation of the possible biases introduced by the use of specific tools.

### 2022-09-08 sections/methods.tex
Fix: Add reference to Table 1 to SP section of Methods
Update on Overleaf.

@@
   The current analysis used stool samples from a fecal microbiome transplant study of autism~\cite{Kang2017} as biological datasets.
+  This dataset was chosen because the sequences were easily accessible on Qiita~\cite{qiita} and optimally pre-processed according to the \ac{emp}~\cite{Thompson2017} protocol and can be used directly as input to the \ac{micone} pipeline.
   The study was composed of multiple sequencing runs.
-  The runs that contained paired end reads (run 2 (10M reads), run 3 (750K reads) and run 4 (16M reads)), were downloaded from Qiita~\cite{qiita} (study ID 10532) and used as input sequences for the \ac{micone} pipeline.
+  The runs that contained paired-end reads (run 2 (10M reads), run 3 (750K reads) and run 4 (16M reads)), were downloaded from Qiita~\cite{qiita} (study ID 10532) and used as input sequences for the \ac{micone} pipeline.
   Sequences from both control (212 samples including neurotypical and donors) and autistic (126 samples) patients were included in the analyses, but the groups were analyzed independently in the pipeline after the \ac{op} step.
   All the network analyses in the study, unless explicitly mentioned, are performed indeed on the healthy and autistic samples in the fecal microbiome transplant study.
-  However, all other analysis (preceding the NI step) are performed together on all the samples in the fecal microbiome transplant study.
+  However, all other analyses (preceding the NI step) are performed together on all the samples in the fecal microbiome transplant study.
   The mock community 16S datasets are experimental sequencing data obtained for artificially assembled collections of species in known proportions.
-  The mock datasets used for this study, obtained from mockrobiota~\cite{Bokulich2016}, are labelled mock4, mock12 and mock16.
+  The mock datasets used for this study, obtained from mockrobiota~\cite{Bokulich2016}, are labeled mock4, mock12, and mock16.
   The mock4 community is composed of 21 bacterial strains.
   Two replicate samples from mock4 contain all species in equal abundances, and two additional replicate samples contain the same species in unequal abundances.
-  The mock12 community is composed of 27 bacterial strains that include closely related taxa with some pairs having only one to two nucleotide difference from one another.
-  The mock16 community is composed of 49 bacteria and 10 Archaea, all represented in equal amount.
+  The mock12 community is composed of 27 bacterial strains that include closely related taxa with some pairs having only one to two nucleotide differences from one another.
+  The mock16 community is composed of 49 bacteria and 10 Archaea, all represented in equal amounts.
 
@@
   The inputs to the pipeline by default are raw untrimmed 16S rRNA sequence reads, but the software can be alternatively configured to use trimmed sequences, \ac{otu} tables and other types of intermediate data (see documentation).
-  The easy-to-use nature, the configurability and modular nature of the \ac{micone} package enables users to start and end the pipeline at any point in the workflow, and to run parts of the pipeline in isolation.
-  The pipeline supports both paired end and single end reads, and additionally supports independently processing reads from multiple runs and merging the OTU tables in the DC step.
+  The configuration and modular nature of the \ac{micone} package enables users to start and end the pipeline at any point in the workflow, and to run parts of the pipeline in isolation.
+  The pipeline supports both paired-end and single-end reads, and additionally supports independently processing reads from multiple runs and merging the OTU tables in the DC step.
   The final output of the pipeline is the inferred network of co-occurrence relationships among the microbes present in the samples.
@@
   The \ac{micone} pipeline provides both a Python API together with a command-line interface and only uses a single configuration file (\texttt{nextflow.config}) to encode the configuration parameters.
-  The \ac{micone} Python API provides several \ac{otu} table and network related functions and methods, enabling detailed comparison of counts tables and inferred networks if desired.
+  The \ac{micone} Python API provides several \ac{otu} table and network-related functions and methods, enabling detailed comparison of counts tables and inferred networks if desired.
   Exploring the effects of these combinations of methods on the resultant networks is difficult and inconvenient since different tools differ in their input and output formats and require interconversions between the various formats.
@@
   It also contains helper functions that can help in parsing taxonomies and communicate with the \ac{ncbi} taxonomy database to query taxonomy by name or taxonomic IDs.
-  The configuration file along with the run file (\texttt{main.nf}) lists the inputs, output and the steps to be performed during runtime, along with the parameters to be used (if different from defaults) for the various steps.
+  The configuration file along with the run file (\texttt{main.nf}) lists the inputs, output, and the steps to be performed during runtime, along with the parameters to be used (if different from defaults) for the various steps.
   Since the entire pipeline run is stored in the form of a text file (the configuration file), subsequent runs are highly reproducible and changes can be easily tracked using version control.
-  It makes use the nextflow workflow manager~\cite{Tommaso2015} under the hood, making it readily usable on the local machine, cluster or cloud with minimal configuration change.
+  It makes use of the nextflow workflow manager~\cite{Tommaso2015} under the hood, making it readily usable on the local machine, cluster, or cloud with minimal configuration change.
   It also allows for automatic parallelization of all possible processes, both within and across samples.
@@
   The various tools used for the processes were adapted from \ac{qiime2} v2021.8.0~\cite{bolyenReproducibleInteractiveScalable2019}.
+  The list of tools used in this step, along with their modules and references are provided in Table~\ref{tab:micone_tools}.
 
@@
   The list of tools used in this step, along with their modules and references are provided in Table~\ref{tab:micone_tools}.
+  In all UniFrac comparisons, we set a count threshold of 10, such that if the count of the representative sequences in a particular sample is less than the threshold, it is omitted from the analysis.
 
@@
   These bootstrapped association matrices are used to calculate the significance of the inferred correlations by calculating the p-values against this null model~\cite{Watts2018}.
-  The pipeline includes 4 methods for pairwise correlation metrics, and 6 methods for direct association metrics (refer Table~\ref{tab:micone_tools}).
-  The Brown's p-value merging method~\cite{brown_400_1975} is used for combining p-values from the pairwise correlations methods to obtain a consensus p-value, which can be used to filter for significance.
+  The pipeline includes 4 methods for pairwise correlation metrics, and 6 methods for direct association metrics (refer to Table~\ref{tab:micone_tools}).
+  Pairwise correlation methods involve the calculation of the correlation coefficient between each pair of nodes (taxonomic entity like Genera) leading to the inclusion of spurious indirect connections.
+  On the other hand, direct association methods use conditional independence to avoid the detection of correlated but indirectly connected taxonomic entities~\cite{Kurtz2015,Menon2018}.
+  Brown's p-value merging method~\cite{brown_400_1975} is used for combining p-values from the pairwise correlations methods to obtain a consensus p-value, which can be used to filter for significance.
   The bootstrapping and p-value calculations are only performed on the correlations-based methods.
   In the final module of this step, the consensus algorithms are used to create the final consensus network using associations from all the network inference methods (except Pearson and Spearman, by default).
-  The output of this step are co-occurrence association networks encoded in the JSON format (refer Supplementary section) and which can also be exportable to a variety of network formats.
+  The outputs of this step are co-occurrence association networks encoded in the JSON format (refer to Supplementary section) and which can also be exportable to a variety of network formats.
   The list of tools used in this step, along with their modules and references are provided in Table~\ref{tab:micone_tools}.
@@
   For all correlation-based methods $k \le w$, $1000$ permutations of the original \ac{otu} counts data were generated~\cite{Watts2018}.
-  All the direct association based methods used in the study have their own regularization methods built-in and hence do not need to undergo bootstrapping.
+  All the direct association-based methods used in the study have their own regularization methods built in and hence do not need to undergo bootstrapping.
   The correlations in the permuted \ac{otu} tables are recalculated using the different correlation-based algorithms.
@@
     If a given pair of nodes is inferred to have edges that satisfy the above relationships, all corresponding edges from the $w$ networks will be returned by the algorithm, as a multigraph.
+    Based on this approach, \ac{micone} reports as the default output, the consensus network where each edge is annotated with weights (correlations for the correlation-based methods and direct associations for the other methods) from all the methods used in the consensus algorithm.
 

### 2022-09-08 sections/results.tex
Update on Overleaf.

@@
   It incorporates various popular, publicly available tools as well as custom Python modules for 16S data analysis and network inference (see Methods).
-  Using \ac{micone} one can obtain co-occurrence networks by applying to 16S data (or to already processed taxonomic count matrices) any combination of the supported tools.
-  The effects of changing any intermediate step can be monitored and evaluated in terms of its final network outcome, as well as on any of the intermediate metrics and data outputs.
   The main workflow of the \ac{micone} pipeline is shown in Figure~\ref{fig:figure1}.
-  The different steps that are a part of the co-occurrence network inference workflow can be grouped into five major modules; (i) the sequence processing (SP) step, which handles the demultiplexing, quality control and trimming of the raw 16S sequencing data into demultiplexed and trimmed sequences; (ii) the denoising and clustering (DC) step, which handles denoising of the demultiplexed and trimmed sequences into representative sequences; (iii) the taxonomy assignment (TA) step that assigns taxonomic labels to the representative sequences; (iv) the \ac{otu} processing (OP) step that filters and transforms the taxonomy abundance table; and finally (v) the network inferences (NI) step which infers the microbial co-occurrence network.
+  The different steps that are a part of the \ac{micone} co-occurrence network inference workflow can be grouped into five major modules; (i) sequence processing (SP) (ii) denoising and clustering (DC) (iii) taxonomy assignment (TA) (iv) \ac{otu} processing (OP) (v) network inferences (NI).
   Each process in the pipeline supports alternate tools for performing the same task (see Methods and Figure~\ref{fig:figure1}).
-  A centralized configuration file contains all the specifications regarding the modules used in the pipeline, and can be modified by the user to choose the desired set of tools.
-  The default tools and parameters for the configuration file can be found in the \ac{micone} documentation (\href{https://micone.readthedocs.io}{https://micone.readthedocs.io}).
+  The effects of changing any intermediate step of the pipeline can be monitored and evaluated in terms of its final network outcome, as well as on any of the intermediate metrics and data outputs.
+  A centralized configuration file contains all the specifications regarding the modules used in the pipeline and can be modified by the user to choose the desired set of tools.
+  The default tools and parameters for the configuration file can be found in Tables S2-S6 at \href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper}.
   In the current study, we performed a systematic analysis of each step of the pipeline to estimate how much the final co-occurrence network depends on the possible choices at each step.
@@
   Our analysis involves two types of data: The first type consists of sets of 16S sequencing data from real communities sampled from human stool microbiomes from a fecal microbiome transplant study of autism~\cite{Kang2017}.
-  This dataset was chosen because the sequences were easily accessible on Qiita~\cite{qiita} and optimally pre-processed according to the \ac{emp}~\cite{Thompson2017} protocol and can be used directly as input to the \ac{micone} pipeline.
-  The second are datasets synthetically or artificially created for the specific goal of helping evaluate computational analysis tools.
-  In particular, in order to objectively compare, to the extent possible, how well each step in \ac{micone} best captures the underlying data, we use both mock data (labelled mock4, mock12 and mock16) from mockrobiota~\cite{Bokulich2016} as well as, synthetic networks generated using the NorTA~\cite{Kurtz2015} and seqtime~\cite{Rottjers2018} methods.
+  The second type are datasets synthetically or artificially created for the specific goal of helping evaluate computational analysis tools.
+  In particular, in order to objectively compare, to the extent possible, how well each step in \ac{micone} best captures the underlying data, we use both mock data (labeled mock4, mock12, and mock16) from mockrobiota~\cite{Bokulich2016} as well as, synthetic networks generated using the NorTA~\cite{Kurtz2015} and seqtime~\cite{Rottjers2018} methods.
   Detailed information on the mock communities and the settings used to generate the synthetic data are provided in the Methods section.
@@
 
-  The \ac{dc} step is commonly carried out to generate representative sequences from the demultiplexed and trimmed 16S sequencing data to obtain the \ac{otu}/\ac{esv} tables (counts of representative sequences for each sample).
-  In order to compare the representative sequence count tables generated by different tools, we processed the same 16S sequencing reads (samples from a fecal microbiome transplant study~\cite{Kang2017}) using 5 different methods: open-reference clustering, closed-reference clustering, de novo clustering, \ac{dada2}~\cite{Callahan2016} and Deblur~\cite{Amir2017}.
+  The \ac{dc} step is commonly carried out to generate representative sequences from the demultiplexed and trimmed 16S sequencing data to obtain the \acs{otu}/\acs{esv} tables (counts of representative sequences for each sample).
+  In order to compare the representative sequence count tables generated by different tools, we processed the same 16S sequencing reads (samples from the FMT study~\cite{Kang2017}) using 5 different methods: open-reference clustering, closed-reference clustering, de novo clustering, \ac{dada2}~\cite{Callahan2016} and Deblur~\cite{Amir2017}.
   The first three methods are from the vsearch plugin from \ac{qiime2}~\cite{bolyenReproducibleInteractiveScalable2019}.
@@
   To compare the representative sequences generated by these methods we employ both the weighted~\cite{Lozupone2007} (Figure~\ref{fig:figure2}A) and unweighted UniFrac method~\cite{Lozupone2005} (Figure~\ref{fig:figure2}B).
-  In both the comparisons, we set a count threshold of 10, such that if the count of the representative sequences in a particular sample is less than the threshold, it is omitted from the analysis.
+  In both comparisons, we set a count threshold of 10, such that if the count of the representative sequences in a particular sample is less than the threshold, it is omitted from the analysis.
   The values shown in the figures correspond to the mean UniFrac values across all samples in the study.
@@
   Figure~\ref{fig:figure2}B on the other hand shows an increase in dissimilarity between each pair of methods suggesting that the methods might differ in the assignment of sequences lower in abundance.
-  In order to verify this claim, for each of these methods we selected the top 1000 representative sequences for each method and recalculated the distance metrics (Figure \ref{fig:figure_s3}).
-  We observe that the both the weighted and unweighted UniFrac distances are increased implying that the top representative sequences generated by the different methods are not as similar to each other.
+  In order to verify this claim, for each of these methods, we selected the top 1000 representative sequences for each method and recalculated the distance metrics (Figure \ref{fig:figure_s3}).
+  We observe that both the weighted and unweighted UniFrac distances are increased, implying that the top representative sequences generated by the different methods are not as similar to each other.
   Therefore, since the weighted UniFrac distances are lower than the unweighted distances, we conclude that the representative sequences in the middle range of the abundance distribution are those that must be the most similar between the methods.
@@
   In order to determine the tool that most accurately recapitulates the reference sequences in the samples, we used the information about the known 16S sequences from the mock datasets (mock4, mock12, and mock16).
-  In particular, we used the pipeline to process these datasets using each of the possible methods included for this step.
+  In particular, we used the pipeline to process these datasets using each of the possible methods included in this step.
   Subsequently, we compared the predicted representative sequences with the expected representative sequences and their distribution.
-  The results (Figure~\ref{fig:figure2}C and D) show that, the predicted sequence distributions are considerably different from the expected sequence distribution.
-  This can be seen in the large values for both unweighted and weighted UniFrac distance values when compared to Figure~\ref{fig:figure2}A and B.
+  The results (Figure~\ref{fig:figure2}C and D) show that the predicted sequence distributions are considerably different from the expected sequence distribution.
   These high deviations are primarily in one or two of the three datasets that were analyzed and show that the datasets themselves play a big role in the performance of these methods.
-  % TODO: How to condense this? Move to supplementary?
-  Open-reference and de novo clustering methods perform the best under the weighted UniFrac metric and the worst (marginally) under the unweighted UniFrac metric.
-  This result can be attributed to the large number of low abundance representative sequences that are generated by these methods.
-  Deblur performs poorly under weighted Unifrac and although its performance on the mock4 dataset is the best under unweighted UniFrac, its performance on the other datasets is average.
-  The Deblur method returns a very small number of representative sequences (2388) and this could account for the reason for the high dissimilarity with the other methods as well as irregular performance on the mock data.
-  There is no method that outperforms the rest in all datasets.
+  There is no method that outperforms the rest in all datasets (see Supplementary for an extended discussion).
   Based on their slightly better performance on the mock datasets, their de novo error correcting nature and previous independent evaluation~\cite{Nearing2018}, \ac{dada2} and Deblur appear to be the most reliable.
   This is because the open-reference and de novo clustering methods return a much larger number of \ac{otu}s compared to the other pipelines and this would affect the accuracy of the network inference step if filtering is not performed.
-  Overall, since \ac{dada2} as compared to Deblur, displays better performance on all the mock datasets and under the weighted UniFrac metric, we this tool as the default for the DC step of the pipeline.
+  Overall, since \ac{dada2} as compared to Deblur, displays better performance on all the mock datasets on the weighted UniFrac metric, we set this tool as the default for the DC step of the pipeline.
   However, if comparison across studies that have sequenced different 16S regions is required, closed-reference and open-reference might be a better option.
@@
   After the denoising, the sequences are subject to chimera checking.
-  The \ac{micone} pipeline supports two different chimera checking methods, uchime-denovo and remove bimera.
-  The ``uchime-denovo'' method is from \ac{qiime2}~\cite{bolyenReproducibleInteractiveScalable2019}, whereas the ``remove bimera'' method is from the \ac{dada2} R package~\cite{Callahan2016}.
+  The \ac{micone} pipeline supports two different chimera checking methods, ``uchime-denovo"~\cite{bolyenReproducibleInteractiveScalable2019}, and ``remove bimera"~\cite{Callahan2016}.
   We did not notice any significant difference between the two chimera checking methods (Figure~\ref{fig:figure_s4}), implying that they identify and remove mostly the same set of sequences as chimeras.
@@
   Taxonomy databases are used to assign taxonomic identities to the representative sequences obtained after the DC step.
-  The three 16S taxonomic reference databases used in this study are SILVA~\cite{Quast2012}, \ac{gg}~\cite{DeSantis2006} and \ac{ncbi} RefSeq~\cite{Sayers2009} (see Methods).
+  The three 16S taxonomic reference databases used in this study are SILVA~\cite{Quast2012}, \acs{gg}~\cite{DeSantis2006} and \acs{ncbi} RefSeq~\cite{Sayers2009} (see Methods).
   These databases vary significantly in terms of taxonomy hierarchies including species names and phylogenetic relationships~\cite{Balvociute2017}.
@@
   Figure~\ref{fig:figure3}A depicts a flow diagram (alluvial plot) that shows how the top 50 representative sequences (sorted by abundance) are assigned a Genus according to the three databases.
-  We observe that the top 50 representative sequences are assigned to 17 Genera in \ac{gg}, 23 in SILVA and 17 in \ac{ncbi}, however the sequences that are assigned to a particular Genus vary significantly based on the taxonomy database.
-  We observe that not only does the assigned Genus composition vary significantly, but the percentage of unassigned representative sequences (pink) also differ.
-  The large percentage of unassigned sequences are because a large fraction of the representative sequences are assigned to an "unknown" Genus during assignment process.
-  This is especially evident in two of the three databases, \ac{ncbi} and \ac{gg}.
-  A representative sequence might be assigned an "unknown" Genus for one of two reasons: the first is if the taxonomy identifier associated with the sequence in the database did not contain a given Genus; the second, more likely, reason is that the database contains multiple sequences that are very similar to the query (representative) sequence and the consensus algorithm (from \ac{qiime2}) is unable to assign one particular Genus at the required confidence.
-  Additionally, we observe that other abundant Genera like Bacteroides and Escherichia are also assigned to different representative sequences, as indicated by the different proportions in the alluvial plot.
+  We observe that the top 50 representative sequences are assigned to 17 Genera in \ac{gg}, 23 in SILVA, and 17 in \ac{ncbi}, however, the sequences that are assigned to a particular Genus vary significantly based on the taxonomy database.
+  We observe that not only does the assigned Genus composition vary significantly, but the percentage of unassigned representative sequences (pink) also differs.
+  The large percentage of unassigned sequences is because a large fraction of the representative sequences is assigned to an "unknown" Genus during the assignment process.
+  A representative sequence might be assigned an "unknown" Genus for one of two reasons: the first is if the taxonomy identifier associated with the sequence in the database did not contain a given Genus; the second, more likely reason, is that the database contains multiple sequences that are very similar to the query (representative) sequence and the consensus algorithm (from \ac{qiime2}) is unable to assign one particular Genus at the required confidence.
+  Additionally, we observe that other abundant Genera like Bacteroides and Escherichia are also assigned to different representative sequences.
   These changes together result in taxonomic assignments generated from each database leading to drastically different taxonomic abundance distributions.
@@
   After assigning all the representative sequences to taxonomies we perform a pairwise comparison of the similarity between the top 100 assignments (by abundance) from different databases at every taxonomic level (Figure~\ref{fig:figure3}B).
-  The comparisons of the assignments beyond Family level (Family, Genus and Species) show less than $45\%$ similarity between any pair of databases.
-  However, the number of mismatches for taxonomy assignments beyond the Genus level (Genus, Species) are greater than $67\%$, implying that the taxonomy assignments from each reference database are fairly unique and are largely responsible for the differences observed in the co-occurrence networks generated from different taxonomy databases.
+  The comparisons of the assignments beyond the Family level (Family, Genus, and Species) show less than $45\%$ similarity between any pair of databases.
+  This implies that the taxonomy assignments from each reference database are fairly unique.
   The comparison of all assigned genera (Figure~\ref{fig:figure_s5}), instead of just the top 100, contains a higher percentage of mismatches.
-  This suggests that the most abundant sequences are the ones that are consistently matched to the same taxonomies in the different reference databases, at least for the dataset tested in the current analysis.
+  This suggests that many of the most abundant sequences are consistently matched to the same taxonomies in the different reference databases, at least for the dataset tested in the current analysis.
 
-  As in the previous section, the comparisons so far only indicate similarity or dissimilarity between methods.
-  In order to obtain an absolute measure of accuracy of the taxonomic assignments we use the expected reference sequences from the mock datasets as the query sequences for the databases and the expected taxonomic composition as the standard to compare against.
-  We use the Bray-Curtis distance metric~\cite{virtanenSciPyFundamentalAlgorithms2020} to calculate the distance between the predicted and expected taxonomic distribution (Figure~\ref{fig:figure3}C).
+  In order to obtain an absolute measure of the accuracy of the taxonomic assignments, we used the expected reference sequences from the mock datasets as the query sequences for the databases and the expected taxonomic composition as the standard to compare against.
+  We used the Bray-Curtis distance metric~\cite{virtanenSciPyFundamentalAlgorithms2020} to calculate the distance between the predicted and expected taxonomic distribution (Figure~\ref{fig:figure3}C).
   Here, we observe that none of the databases perform better than the others in absolute terms and that the dissimilarity with the expected composition is high ($>0.5$ for Family and Genus and $>0.9$ for Species), indicating that all the databases perform poorly when trying to recapture the expected taxonomic composition.
@@
 
-  Since no database performs better than others against mock datasets, the choice of which database to use could be driven by other reasons.
-  One user-specific way to choose, would be based on the known representation of taxa for the microbiome of interest (see Discussion).
-  Another reason could be the frequency of updates and the potential for future growth.
+  Since no database performs better than others against mock datasets, the choice of which database to use could be driven by other reasons (see Discussion).
+  One reason to choose a particular database could be the frequency of updates and the potential for future growth.
   Since the \ac{gg} database is used frequently as a 16S reference database in studies~\cite{Balvociute2017}, to enable easier comparison with published networks in literature, we set \ac{gg} as the default for taxonomy assignment.
-  On the other hand, since the degree of similarity of assignments between \ac{ncbi} and \ac{gg} is fairly high (Figure~\ref{fig:figure_s5}), and due to its regular revision and maintenance the \ac{ncbi} database is also a good choice for the user.
+  On the other hand, since the degree of similarity of assignments between \ac{ncbi} and \ac{gg} is fairly high (Figure~\ref{fig:figure_s5}), and due to its regular revision and maintenance, the \ac{ncbi} database is also a good choice for the user.
   The TA step results in a taxonomic counts table that is used as input to the subsequent steps of the pipeline.
-  The count tables at different levels can be obtained through aggregation, for example, Genus count tables were obtained by summing up the counts of the lower taxonomy levels (Species and \ac{otu}) that map to the same higher taxonomy level name.
+  The count tables at different levels can be obtained through aggregation, for example, Genus count tables were obtained by summing up the counts of the lower taxonomy levels (Species and \ac{otu}) that map to the same higher taxonomy level entity.
 
@@
 
-  \subsection*{NI: Networks generated using different network inference methods show notable difference in edge-density and connectivity}
+  \subsection*{NI: Networks generated using different network inference methods show a notable difference in edge-density and connectivity}
 
-  The ten different network inference methods used in this study are: \ac{spieceasi}~\cite{Kurtz2015}, FlashWeave~\cite{tackmannRapidInferenceDirect2019}, \ac{cozine}~\cite{haCompositionalZeroinflatedNetwork2020a}, \ac{harmonies}~\cite{jiangHARMONIESHybridApproach2020}, \ac{spring}~\cite{yoonMicrobialNetworksSPRING2019}, \ac{mldm}~\cite{Yang2017}, \ac{sparcc}~\cite{Friedman2012,Watts2018}, propr~\cite{quinnProprRpackageIdentifying2017}, Spearman and Pearson correlation methods.
-  These network inference methods fall into two groups, the first set of methods (Pearson, Spearman, \ac{sparcc}, and propr) infer pairwise correlations while the second set infer direct associations (\ac{spieceasi}, FlashWeave, \ac{cozine}, \ac{harmonies}, \ac{spring}, and \ac{mldm}).
-  Pairwise correlation methods involve calculation of the correlation coefficient between each pair of nodes (taxonomic entity like Genera) leading to the inclusion of spurious indirect connections.
-  On the other hand, direct association methods use conditional independence to avoid the detection of correlated but indirectly connected taxonomic entities~\cite{Kurtz2015,Menon2018}.
+  The ten different network inference methods used in this study are: \acs{spieceasi}~\cite{Kurtz2015}, FlashWeave~\cite{tackmannRapidInferenceDirect2019}, \acs{cozine}~\cite{haCompositionalZeroinflatedNetwork2020a}, \acs{harmonies}~\cite{jiangHARMONIESHybridApproach2020}, \acs{spring}~\cite{yoonMicrobialNetworksSPRING2019}, \acs{mldm}~\cite{Yang2017}, \acs{sparcc}~\cite{Friedman2012,Watts2018}, propr~\cite{quinnProprRpackageIdentifying2017}, Spearman and Pearson correlation methods.
+  These network inference methods fall into two groups, the first set of methods (Pearson, Spearman, \acs{sparcc}, and propr) infer pairwise correlations while the second set infer direct associations (\acs{spieceasi}, FlashWeave, \acs{cozine}, \acs{harmonies}, \acs{spring}, and \acs{mldm}).
 
-  For the analysis presented in this section, we used the filtered (performed during the \ac{op} step, refer Methods) Genus level counts table obtained using the default settings in the DC and TA steps as the input for the network inference algorithms.
+  For the analysis presented in this section, we used the filtered (performed during the \ac{op} step, refer to Methods) Genus level counts table obtained using the default settings in the DC, TA, and OP steps as the input for the network inference algorithms.
   Figure~\ref{fig:figure4}A shows the networks inferred from this dataset using the different inference algorithms.
@@
   In order to obtain a quantitative picture of the differences between the inferred networks, we analyzed the distribution of common nodes and edges (Figure~\ref{fig:figure4} B and C) using UpSet plots~\cite{Lex}.
-  Only \ac{harmonies}, \ac{cozine}, \ac{spieceasi}, \ac{spring}, propr, and \ac{sparcc} are used in the comparison since Pearson and Spearman add a large number of spurious edges as they are not intended for compositional datasets.
-  The results for the node intersections show that the networks have $33$ out of $68$ total unique nodes in common and no network possesses an unique node.
+  Only \acs{harmonies}, \acs{cozine}, \acs{spieceasi}, \acs{spring}, propr, and \acs{sparcc} are used in the comparison since Pearson and Spearman add a large number of spurious edges as they are not intended for compositional datasets.
+  The results for the node intersections show that the networks have $33$ out of $68$ total unique nodes in common and no network possesses a unique node.
   The edge intersections in contrast show that only $8$ edges (out of $202$ total unique edges) are in common between all the methods and each network has a large number of unique edges.
@@
   Therefore, we have developed a consensus method that would take into consideration the evidence offered by each network inference algorithm and generate a consensus network that would contain the most probable interactions for the input data.
-  We have developed two algorithms for the generation of the consensus network, the simple voting and the scaled-sum methods (see Methods).
-  Both approaches combine network inferred from both correlation-based and direct association methods.
-  First, for the correlation-based methods we calculate p-values using null models and then merge the p-values using the Brown's p-value merging method~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016,faustCoNetAppInference2016} (see Methods).
+  We have developed two algorithms for the generation of the consensus network, simple voting and the scaled-sum methods (see Methods).
+  Both approaches combine networks inferred from both correlation-based and direct association methods.
+  First, for the correlation-based methods, we calculate p-values using null models and then merge the p-values using Brown's p-value merging method~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016,faustCoNetAppInference2016} (see Methods).
   Second, we filter all the inferred networks based on an association strength threshold of 0.1 and a p-value cutoff of 0.05.
   Finally, we apply the consensus algorithms we have developed on these filtered networks.
-  % The simple voting method does a consensus based on whether the number of inferred networks that have an edge between a pair of nodes is greater than a threshold.
-  % The scaled-sum method does a consensus based on whether the normalized sum of an edge between a pair of nodes from all inferred networks is greater than a threshold.
-  The default algorithm for the pipeline is the scaled-sum method, since this method takes into account the weights of the associations in the determination of the final consensus.
-  By default, of the correlation-based methods only \ac{sparcc} and propr are included in the calculation of consensus network, for consensus networks that include Spearman and Pearson see Figure~\ref{fig:figure_s6} in the Supplementary section.
-  Additionally, the pipeline also allows for a different subset of methods to be used in the calculation of the consensus if it is required.
-  Based on this approach, \ac{micone} reports as default output, the consensus network, where is each edge is annotated with weights (correlations for the correlation-based methods and direct associations for the other methods) from all the methods used in the consensus algorithm.
+  The default algorithm for the pipeline is the scaled-sum method since this method takes into account the weights of the associations in the determination of the final consensus.
+  By default, of the correlation-based methods, only \acs{sparcc} and propr are included in the calculation of the consensus network.
+  Additionally, the pipeline also allows for the selection of the subset of methods to be used in the calculation of the consensus (for consensus networks that include Spearman and Pearson, see Figure~\ref{fig:figure_s6} in the Supplementary section).
 
@@
   This counts table is used as the input to the \ac{micone} pipeline to generate predicted associations.
-  The interaction network used to generate the counts table is used as the source of true interactions to calculate the precision and sensitivity for each network inference algorithm for the NorTA datasets (Figure \ref{fig:figure5}).
+  The interaction network used to generate the counts table is used as the source of true interactions to calculate the precision and sensitivity for each network inference algorithm (Figure \ref{fig:figure5}).
   The goal of the consensus network algorithms is to increase the precision of the resultant consensus network by removing edges that are either not present in most of the inference methods or whose association strength is low.
-  In Figure \ref{fig:figure5} we show that the consensus algorithm, especially the scaled-sum method, captures true associations with more precision, but as a consequence the sensitivity of the predictions is reduced.
+  In Figure \ref{fig:figure5} we show that the consensus algorithm, especially the scaled-sum method, captures true associations with more precision, but as a consequence, the sensitivity of the predictions is reduced.
+  Among the individual inference methods, we observe that \acs{spieceasi} has the best average precision ($0.944$)
   Overall, we see that the scaled-sum method for $p=1.0$ performs the best in terms of average precision ($1.000$).
   However, the scaled-sum method for $p=0.667$ has the second-highest average precision ($0.985$) and does not underperform in terms of sensitivity.
-  Among the individual inference methods, we observe that \ac{spieceasi} has the best average precision ($0.944$)
   Therefore, the consensus networks provide the means to obtain a short list of associations that would have a high probability of being present in the real interaction network.
   Furthermore, we also find that the consensus algorithm performs drastically better in terms of precision (0.820 for scaled-sum method vs.
-0.624 for \ac{spieceasi}) in recapturing the original interactions in the ``seqtime'' data.
+0.624 for \acs{spieceasi}) in recapturing the original interactions in the ``seqtime'' data.
   The precision and sensitivity graphs on the seqtime data can be found in the Supplementary section (Figure~\ref{fig:figure_s7}).
@@
 
-  \subsection*{Overall assessment of impact on inferred networks}
+  \subsection*{Overall assessment of the impact on inferred networks}
 
-  In order to analyze the effect of different processing methods on the inferred co-occurrence networks, we generated these networks using all possible combinations of methods and estimated the variability in the networks due to each choice (Figure \ref{fig:figure6}A).
+  In order to analyze the effect of different processing methods on the inferred co-occurrence networks, we generated networks using all possible combinations of methods and estimated the variability due to each choice (Figure \ref{fig:figure6}A).
   The networks used in this analysis are those directly inferred from the correlation-based and direct inference methods and to which the consensus algorithm is yet to be applied.
-  The effects of various steps on the final co-occurrence network are estimated by building a linear model of the edges of the network as a function the various steps in the pipeline workflow (see Methods).
+  The effects of various steps on the final co-occurrence network are estimated by building a linear model of the edges of the network as a function of the various steps in the pipeline workflow (see Methods).
   Figure \ref{fig:figure6}A, shows the percentage of total variation among the co-occurrence networks due to the different steps of the pipeline.
@@
   The reason for this seems to derive from the differential labeling of constitutive representative sequences into taxonomic entities (Figure~\ref{fig:figure3} and Figure~\ref{fig:figure_s5}), which drastically alter the nodes and hence the underlying network topology.
-  In fact the variability induced by taxonomy assignment is much more significant than that due to the variability introduced based on how the reference sequences themselves are identified (in the DC step).
-  The fraction labeled as the residual is an artifact that arises when multiple steps are changed at the same time, however this can be ignored as the fraction is negligible.
+  The fraction labeled as the residual is an artifact that arises when multiple steps are changed at the same time, however, this can be ignored as the fraction is negligible.
 
@@
   Figure \ref{fig:figure6}B is the PCA plot of all the generated co-occurrence networks, colored by the databases in the TA step (left) and the network inference algorithms in the NI step (right).
-  In other words, each point corresponds to a network generated using different combinations of tools, and captures how much the final network is affected by such a choice.
+  In other words, each point corresponds to a network generated using different combinations of tools.
   The importance of the TA step, shown in Figure \ref{fig:figure6}A, is also reflected here since most of the networks can be separated based on the database used (Figure \ref{fig:figure6}B).
-  However, it is also evident that some networks, especially those inferred using the direct association based network inference methods, are much closer in the PCA plot regardless of the reference database used.
+  However, it is also evident that some networks, especially those inferred using the direct association network inference methods, are much closer in the PCA plot regardless of the reference database used.
   Additionally, the variation (on the first two principal axes) between the networks decreases when the low abundance \ac{otu}s are removed from the network (Figure~\ref{fig:figure_s1}).
@@
   The systematic analyses performed in the previous sections illustrate that the choice of tools and parameters can have a big impact on the final co-occurrence network.
-  For some of these choices (e.g. for the \ac{op} step), there is no clear metric to establish the best protocol.
-  For other choices, the mock communities and synthetic data provide an opportunity to select the combinations of tools that yield more accurate and robust results, though those may depends on assumptions and details of the mock data.
-  Despite this partial degree of assessment, we wish to suggest a combination of tools and parameters that led to the best performance based on the mock communities, and displayed the highest robustness to switching to alternative methods.
+  The mock communities and synthetic data provide an opportunity to select the combinations of tools that yield more accurate and robust results.
+  However, our analysis could not yield the best protocol for some of the steps of the workflow (e.g. for the \ac{ta} step).
+  Despite this partial degree of assessment, we wish to suggest a combination of tools and parameters that led to the best performance based on the mock and synthetic data and which displayed the highest robustness to switching to alternative methods.
   These tools and parameters are chosen as the defaults for the pipeline and are given in Table~\ref{tab:micone_tools}.
@@
   Figure~\ref{fig:figure7}A shows the comparison of the default network against networks generated by altering one of the steps of the pipeline from the default.
-  This result suggests that the most significant changes (many new nodes and edges are added) occur when the \ac{ta} step, \ac{op} step and the \ac{ni} step are changed.
-  However, the L1 distance of networks generated by altering one of the steps of the pipeline from the default against the default network (Figure~\ref{fig:figure7}B) shows that the biggest deviations from the default network occur when the \ac{ni} step is changed.
-  These results indicate that the biggest rewiring in networks occur when the network inference algorithm is changed and that the changes in the others steps do not create as significant of a difference.
-  % TODO: Show proof??
-  The change of taxonomy database and the filtering in the \ac{op} step when turned off also adds a large number of new nodes, but the edges between the common nodes (with the default network) are fairly conserved.
+  This result suggests that the most significant changes (many new nodes and edges are added) occur when the \ac{ta} step, \ac{op} step, and the \ac{ni} step are changed.
+  However, the L1 distance of networks generated by altering one of the steps of the pipeline from the default against the default network (Figure~\ref{fig:figure7}B) shows that the biggest deviations from the default network occur when the \ac{ni} step is changed from the consensus algorithm to one of the individual algorithms.
+  These results indicate that the biggest rewiring in networks occurs when the network inference algorithm is changed and that the changes in the other steps do not create as significant of a difference.
+  The change of the taxonomy database and the filtering in the \ac{op} step when turned off also adds a large number of new nodes, but the edges between the common nodes (with the default network) are fairly conserved.
 
-  The low L1 distances show us that not only does the consensus help reconcile the variability in the edges due to multiple network inference algorithms, it also helps reduce the variability due to the other steps of the pipeline as well.
+  The low L1 distances show us that not only does the consensus help reconcile the variability in the edges due to multiple network inference algorithms, but it also helps reduce the variability due to the other steps of the pipeline as well.
   This result is different from that observed in Figure~\ref{fig:figure6} since in this particular analysis all the steps other than the \ac{ni} step have the consensus algorithm applied to them, which drastically reduces the edges in the graph that have lower probabilities.
-  This reinforces the importance of the consensus network in increasing the robustness of the inferred networks to different processing methods.
+  This reinforces the importance of the consensus network in increasing the robustness of the inferred networks to change in processing methods.
 
-  Figure~\ref{fig:figure7}C shows the co-occurrence networks inferred for the healthy subjects (control) and autistic subjects in the fecal microbiome transplant study~\cite{Kang2017}.
+  Figure~\ref{fig:figure7}C shows the co-occurrence networks inferred for the healthy subjects (control) and autistic subjects in the fecal microbiome transplant study~\cite{Kang2017} (using the default tools and parameters from Table~\ref{tab:micone_tools}).
   This figure demonstrates a typical use case of comparative analysis of networks using the \ac{micone} pipeline.
-  These consensus networks were generated using the default tools and parameters from Table~\ref{tab:micone_tools}.
   As a consequence of using the consensus network algorithm, the final co-occurrence networks are sparse and can be visually compared and examined.
   We observe 10 unique links in the network for control samples and 3 unique links in the network for autistic subjects.
-  These unique connections can serve as potential starting points for literature surveys and further for experimental confirmations.
+  These unique connections can serve as potential starting points for literature surveys and further experimental confirmations.

### 2022-09-08 sections/supplementary.tex
Update on Overleaf.

@@
 
+    \subsubsection*{Interpretation of Unifrac results in the DC step}
+    Open-reference and de novo clustering methods perform the best under the weighted UniFrac metric and the worst (marginally) under the unweighted UniFrac metric.
+    This result can be attributed to the large number of low abundance representative sequences that are generated by these methods.
+    Deblur performs poorly under weighted Unifrac and although its performance on the mock4 dataset is the best under unweighted UniFrac, its performance on the other datasets is average.
+    The Deblur method returns a very small number of representative sequences (2388) and this could account for the reason for the high dissimilarity with the other methods as well as irregular performance on the mock data.
+
   \subsection*{Synthetic interaction data}

### 2022-09-08 sections/abstract.tex
Update on Overleaf.

@@
     These snapshots, when collected from multiple samples, have the potential to reveal which microbes co-occur, providing a glimpse into the network of associations in these communities.
-    The inference of networks from 16S data is prone to statistical artifacts.
-    There are many tools for performing each step of the 16S analysis workflow, but the extent to which these steps affect the final network is still unclear.
+    The inference of networks from 16S data involves numerous steps, each requiring specific tools and parameter choices.The extent to which the steps in this workflow affect the final network is still unclear.
     In this study, we perform a meticulous analysis of each step of a pipeline that can convert 16S sequencing data into a network of microbial associations.

### 2022-09-13 sections/abstract.tex
Update on Overleaf.

@@
     These snapshots, when collected from multiple samples, have the potential to reveal which microbes co-occur, providing a glimpse into the network of associations in these communities.
-    The inference of networks from 16S data involves numerous steps, each requiring specific tools and parameter choices.The extent to which the steps in this workflow affect the final network is still unclear.
+    The inference of networks from 16S data involves numerous steps, each requiring specific tools and parameter choices.
+    The extent to which the steps in this workflow affect the final network is still unclear.
     In this study, we perform a meticulous analysis of each step of a pipeline that can convert 16S sequencing data into a network of microbial associations.
@@
     We further determine the tools and parameters that generate robust co-occurrence networks and develop consensus network algorithms based on benchmarks with mock and synthetic datasets.
-    The Microbial Co-occurrence Network Explorer or \acs{micone} (available at \href{https://github.com/segrelab/MiCoNE}{https://github.com/segrelab/MiCoNE}), follows these default tools and parameters, and can help explore the outcome of these combinations of choices on the inferred networks.
-    We envisage that this pipeline could be used for integrating multiple data-sets, and for generating comparative analyses and consensus networks that can help understand microbial community assembly in different biomes.
+    The Microbial Co-occurrence Network Explorer or \acs{micone} (available at \href{https://github.com/segrelab/MiCoNE}{https://github.com/segrelab/MiCoNE}), follows these default tools and parameters and can help explore the outcome of these combinations of choices on the inferred networks.
+    We envisage that this pipeline could be used for integrating multiple data sets, and for generating comparative analyses and consensus networks that can help understand microbial community assembly in different biomes.
   }
@@
 \section*{Importance}
-  Mapping the interrelationships between different species in a microbial community is essential to understand and control their structure and function.
+  Mapping the interrelationships between different species in a microbial community is essential to understanding and controlling their structure and function.
   The surge in the high-throughput sequencing of microbial communities has led to the creation of thousands of datasets containing information about microbial abundances.
@@
   In this study, we address this workflow and provide a systematic analysis of how these choices of tools affect the final network, and guidelines on appropriate tool selection for a particular dataset.
-  We also develop a consensus network algorithm that helps generate more robust co-occurrence networks based on benchmarks synthetic datasets.
+  We also develop a consensus network algorithm that helps generate more robust co-occurrence networks based on benchmark synthetic datasets.

### 2022-09-13 sections/supplementary.tex
Update on Overleaf.

@@
 
+% TODO: Update this section and link
     \subsubsection*{Interpretation of Unifrac results in the DC step}
+  In order to verify this claim, for each of these methods, we selected the top 1000 representative sequences for each method and recalculated the distance metrics (Figure \ref{fig:figure_s3}).
+  We observe that both the weighted and unweighted UniFrac distances are increased, implying that the top representative sequences generated by the different methods are not as similar to each other.
+  Therefore, since the weighted UniFrac distances are lower than the unweighted distances, we conclude that the representative sequences in the middle range of the abundance distribution are those that must be the most similar between the methods.
+
     Open-reference and de novo clustering methods perform the best under the weighted UniFrac metric and the worst (marginally) under the unweighted UniFrac metric.

### 2022-09-13 sections/tables_figures.tex
Change: Update manuscript based on reviewer comments
Change: Remove figures from manuscript
Update on Overleaf.

@@
                                           & \multirow{2}{*}{Consensus algorithm} & \rowcolor{lightgray} scaled-sum & 0.667 \\
-                                          & & simple voting & 0.667 \\
+                                          & & simple voting & 1.000 \\
       \hline
@@
 % Figures
+
   \FloatBarrier
   \newpage
+
+  \FloatBarrier
+  \newpage
+  % \begin{figure}[H]
+  %   \centering
+  %   \includegraphics[width=0.74\linewidth]{figure1.pdf}
+  % \end{figure}
   \begin{figure}[H]
@@
 
+  \FloatBarrier
+  \newpage
+  % \begin{figure}[H]
+  %   \centering
+  %   \includegraphics[width=\textwidth]{figure2.pdf}
+  % \end{figure}
   \begin{figure}[H]
@@
       \textbf{(C, D)} The distributions of the average weighted and unweighted UniFrac distance between the expected sequence profile and the predicted sequence profile in the mock datasets.
-      The distributions of the average weighted UniFrac distance show that de novo (DN) and open reference (OR) were the best performing methods in most of the datasets, while they are worst performing methods under the unweighted UniFrac metric.
+      The distributions of the average weighted UniFrac distance show that de novo (DN) and open reference (OR) were the best-performing methods in most of the datasets, while they are worst performing methods under the unweighted UniFrac metric.
       The good performance of dada2 (D2) under both distance metrics combined with its approach of identifying \ac{esv}s using de novo methods, prompts us to propose it as the default method for the DC step.
@@
 
+  \FloatBarrier
+  \newpage
+  % \begin{figure}[H]
+  %   \centering
+  %   \includegraphics[width=\textwidth]{figure3.pdf}
+  % \end{figure}
   \begin{figure}[H]
@@
       \textbf{(A)} The assignments of the top 50 representative sequences to their respective taxonomies using the three different reference databases.
-      This result illustrates how the same sequences are assigned to different Genus under the different databases.
+      This result illustrates how the same sequences are assigned to different Genus under different databases.
       A significant portion of the representative sequences are assigned to an ``unknown'' Genus in two of three databases (\ac{gg} and \ac{ncbi}).
@@
 
+  \FloatBarrier
+  \newpage
+  % \begin{figure}[H]
+  %   \centering
+  %   \includegraphics[width=\textwidth]{figure4.pdf}
+  % \end{figure}
   \begin{figure}[H]
@@
 
+  % TODO: Move caption to Figure S6
+  % Give reason for using those 4 individual methods
+  \FloatBarrier
+  \newpage
+  % \begin{figure}[H]
+  %   \centering
+  %   \includegraphics[width=1.0\linewidth]{figure5.pdf}
+  % \end{figure}
   \begin{figure}[H]
@@
 
+  \FloatBarrier
+  \newpage
+  % TODO: Add equations for the linear model to the left of the pie chart
+  % \begin{figure}[H]
+  %   \centering
+  %   \includegraphics[width=\textwidth]{figure6.pdf}
+  % \end{figure}
   \begin{figure}[H]
@@
       The variation due to the DC and CC steps are comparatively not as significant.
+      The fraction labeled as the residual is an artifact that arises when multiple steps are changed at the same time, however, this can be ignored as this fraction is negligible.
       \textbf{(B)} All combinations of inferred networks are shown as points on a PCA plot.
@@
 
+  \FloatBarrier
+  \newpage
+  % TODO: Verify the validity of Figure 7A and 7B again
+  % Are they both using consensus?
+  % \begin{figure}[H]
+  %   \centering
+  %   \includegraphics[width=1.0\linewidth]{figure7.pdf}
+  % \end{figure}
   \begin{figure}[H]
@@
   \end{figure}
-
-  \FloatBarrier
-  \newpage
-
-  \begin{figure}[H]
-    \centering
-    \includegraphics[width=0.74\linewidth]{figure1.pdf}
-  \end{figure}
-
-  \FloatBarrier
-  \newpage
-
-  \begin{figure}[H]
-    \centering
-    \includegraphics[width=\textwidth]{figure2.pdf}
-  \end{figure}
-
-  \FloatBarrier
-  \newpage
-
-  \begin{figure}[H]
-    \centering
-    \includegraphics[width=\textwidth]{figure3.pdf}
-  \end{figure}
-
-  \FloatBarrier
-  \newpage
-
-  \begin{figure}[H]
-    \centering
-    \includegraphics[width=\textwidth]{figure4.pdf}
-  \end{figure}
-
-  \FloatBarrier
-  \newpage
-
-  \begin{figure}[H]
-    \centering
-    \includegraphics[width=1.0\linewidth]{figure5.pdf}
-  \end{figure}
-
-  \FloatBarrier
-  \newpage
-
-  % TODO: Add equations for the linear model to the left of the pie chart
-  \begin{figure}[H]
-    \centering
-    \includegraphics[width=\textwidth]{figure6.pdf}
-  \end{figure}
-
-  \FloatBarrier
-  \newpage
-
-  % TODO: Verify the validity of Figure 7A and 7B again
-  % Are they both using consensus?
-  \begin{figure}[H]
-    \centering
-    \includegraphics[width=1.0\linewidth]{figure7.pdf}
-  \end{figure}

### 2022-09-13 sections/methods.tex

@@
   The Naive Bayes classifier and the \ac{ncbi} blast used as the query tools in this study were from the \ac{qiime2} package and the parameters used were the defaults of the package.
+  During the assignment, a representative sequence might be assigned an "unknown" Genus for one of two reasons: the first is if the taxonomy identifier associated with the sequence in the database did not contain a given Genus; the second, more likely reason, is that the database contains multiple sequences that are very similar to the query (representative) sequence and the consensus algorithm (from \ac{qiime2}) is unable to assign one particular Genus at the required confidence.
   The consensus algorithm used is the default method used by the classifiers in \ac{qiime2}.
@@
   \subsection*{Consensus network and p-value merging}
+  The consensus methods combine networks inferred from both correlation-based and direct association methods. 
+  First, for the correlation-based methods, we calculate p-values using null models and then merge the p-values using Brown's p-value merging method~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016,faustCoNetAppInference2016} (see Methods).
+  Second, we filter all the inferred networks based on an association strength threshold of 0.1 and a p-value cutoff of 0.05.
+  Finally, we apply the consensus algorithms we have developed on these filtered networks.
+  These steps are elaborated on in the subsequent sections.
 
@@
   For all correlation-based methods $k \le w$, $1000$ permutations of the original \ac{otu} counts data were generated~\cite{Watts2018}.
-  All the direct association-based methods used in the study have their own regularization methods built in and hence do not need to undergo bootstrapping.
+  All the direct association-based methods used in the study have their own regularization methods built-in and hence do not need to undergo bootstrapping.
   The correlations in the permuted \ac{otu} tables are recalculated using the different correlation-based algorithms.

### 2022-09-13 sections/results.tex

@@
 
-  We developed \ac{micone}, a flexible and modular pipeline for the inference of occurrence networks from 16S data.
-  It incorporates various popular, publicly available tools as well as custom Python modules for 16S data analysis and network inference (see Methods).
-  The main workflow of the \ac{micone} pipeline is shown in Figure~\ref{fig:figure1}.
-  The different steps that are a part of the \ac{micone} co-occurrence network inference workflow can be grouped into five major modules; (i) sequence processing (SP) (ii) denoising and clustering (DC) (iii) taxonomy assignment (TA) (iv) \ac{otu} processing (OP) (v) network inferences (NI).
-  Each process in the pipeline supports alternate tools for performing the same task (see Methods and Figure~\ref{fig:figure1}).
-  The effects of changing any intermediate step of the pipeline can be monitored and evaluated in terms of its final network outcome, as well as on any of the intermediate metrics and data outputs.
-  A centralized configuration file contains all the specifications regarding the modules used in the pipeline and can be modified by the user to choose the desired set of tools.
-  The default tools and parameters for the configuration file can be found in Tables S2-S6 at \href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper}.
-  In the current study, we performed a systematic analysis of each step of the pipeline to estimate how much the final co-occurrence network depends on the possible choices at each step.
-  We also evaluated numerous tool combinations to determine a set of recommended default options for the pipeline and provide the users with a set of guidelines to facilitate tool selection as appropriate for their data.
+  We developed \ac{micone}, a flexible and modular pipeline for the inference of co-occurrence networks from 16S data.
+  \ac{micone} incorporates various popular, publicly available tools as well as custom Python modules for 16S data analysis and network inference (see Methods).
+  The different steps that are a part of the \ac{micone} co-occurrence network inference workflow (Figure~\ref{fig:figure1}) can be grouped into five major modules; (i) sequence processing (SP); (ii) denoising and clustering (DC); (iii) taxonomy assignment (TA); (iv) \ac{otu} processing (OP); (v) network inferences (NI).
+  Each process in the pipeline can be implemented through  multiple tools (see Methods and Figure~\ref{fig:figure1}).
+  The effects of changing any intermediate step of the pipeline can be evaluated in terms of the final network outcome, as well as on any of the intermediate metrics and data outputs.
+  The choice of tools and parameters is encoded in a configuration file (with defaults shown in Tables S2-S6 at \href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper}).
+  Through a systematic analysis of tool combinations at each step of the pipeline, we estimated how much the final co-occurrence network depends on the possible choices at each step.
 
-  Our analysis involves two types of data: The first type consists of sets of 16S sequencing data from real communities sampled from human stool microbiomes from a fecal microbiome transplant study of autism~\cite{Kang2017}.
-  The second type are datasets synthetically or artificially created for the specific goal of helping evaluate computational analysis tools.
-  In particular, in order to objectively compare, to the extent possible, how well each step in \ac{micone} best captures the underlying data, we use both mock data (labeled mock4, mock12, and mock16) from mockrobiota~\cite{Bokulich2016} as well as, synthetic networks generated using the NorTA~\cite{Kurtz2015} and seqtime~\cite{Rottjers2018} methods.
-  Detailed information on the mock communities and the settings used to generate the synthetic data are provided in the Methods section.
+  Our analysis involves two types of data: The first type consists of sets of 16S sequencing data from samples of human stool microbiomes from a fecal microbiome transplant study of autism~\cite{Kang2017}.
+  The second type is a collection of datasets synthetically or artificially created for the specific goal of evaluating computational analysis tools.
+  In particular, in order to  compare how well each step in \ac{micone} best captures the underlying data, we use both mock data (labeled mock4, mock12, and mock16) from mockrobiota~\cite{Bokulich2016} and synthetic networks generated using the NorTA~\cite{Kurtz2015} and seqtime~\cite{Rottjers2018} approaches (See Methods).
 
@@
 
-  \subsection*{DC: Denoising and clustering methods differ in their identification of less common reference sequences}
+  \subsection*{DC: Denoising and clustering differ in their identification of sequences that are low in abundance}
 
-  The \ac{dc} step is commonly carried out to generate representative sequences from the demultiplexed and trimmed 16S sequencing data to obtain the \acs{otu}/\acs{esv} tables (counts of representative sequences for each sample).
-  In order to compare the representative sequence count tables generated by different tools, we processed the same 16S sequencing reads (samples from the FMT study~\cite{Kang2017}) using 5 different methods: open-reference clustering, closed-reference clustering, de novo clustering, \ac{dada2}~\cite{Callahan2016} and Deblur~\cite{Amir2017}.
+  The \ac{dc} step is commonly carried out to generate representative sequences (in the form of the \acs{otu}/\acs{esv} tables) from the demultiplexed and trimmed 16S sequencing data.
+  In order to compare the count tables generated by different tools, we processed the same 16S sequencing reads (from the FMT study~\cite{Kang2017}) using 5 different methods: open-reference clustering, closed-reference clustering, de novo clustering, \ac{dada2}~\cite{Callahan2016} and Deblur~\cite{Amir2017}.
   The first three methods are from the vsearch plugin from \ac{qiime2}~\cite{bolyenReproducibleInteractiveScalable2019}.
-  The closed and open reference methods in this analysis use the \ac{gg} database for reference sequence alignment.
+  The closed and open reference methods in this analysis use the \ac{gg}~\cite{DeSantis2006} database for reference sequence alignment.
 
-  To compare the representative sequences generated by these methods we employ both the weighted~\cite{Lozupone2007} (Figure~\ref{fig:figure2}A) and unweighted UniFrac method~\cite{Lozupone2005} (Figure~\ref{fig:figure2}B).
-  In both comparisons, we set a count threshold of 10, such that if the count of the representative sequences in a particular sample is less than the threshold, it is omitted from the analysis.
-  The values shown in the figures correspond to the mean UniFrac values across all samples in the study.
-  The weighted UniFrac distance metric takes into account the counts of the representative sequences, whereas the unweighted UniFrac distance metric does not and hence gives equal weights to each sequence.
-  From Figure~\ref{fig:figure2}A one can see that the representative sequences generated by the different methods, with the exception of Deblur, are similar to each other when weighted by their abundance.
-  Figure~\ref{fig:figure2}B on the other hand shows an increase in dissimilarity between each pair of methods suggesting that the methods might differ in the assignment of sequences lower in abundance.
-  In order to verify this claim, for each of these methods, we selected the top 1000 representative sequences for each method and recalculated the distance metrics (Figure \ref{fig:figure_s3}).
-  We observe that both the weighted and unweighted UniFrac distances are increased, implying that the top representative sequences generated by the different methods are not as similar to each other.
-  Therefore, since the weighted UniFrac distances are lower than the unweighted distances, we conclude that the representative sequences in the middle range of the abundance distribution are those that must be the most similar between the methods.
+  A comparison of the different approaches is visualized in the form of mean UniFrac values across all samples (Figure~\ref{fig:figure2}). 
+  The analysis was performed using both the weighted UniFrac~\cite{Lozupone2007} (Figure~\ref{fig:figure2}A) distance metric, which takes into account the counts of the representative sequences, and the unweighted UniFrac~\cite{Lozupone2005} (Figure~\ref{fig:figure2}B) distance metric, which gives equal weights to each sequence.
 
-  These comparisons only elucidate the similarity or dissimilarity of a pair of methods.
-  In order to determine the tool that most accurately recapitulates the reference sequences in the samples, we used the information about the known 16S sequences from the mock datasets (mock4, mock12, and mock16).
-  In particular, we used the pipeline to process these datasets using each of the possible methods included in this step.
-  Subsequently, we compared the predicted representative sequences with the expected representative sequences and their distribution.
-  The results (Figure~\ref{fig:figure2}C and D) show that the predicted sequence distributions are considerably different from the expected sequence distribution.
-  These high deviations are primarily in one or two of the three datasets that were analyzed and show that the datasets themselves play a big role in the performance of these methods.
+  The first main message emerging from this analysis is that the representative sequences generated by the different methods, with the exception of Deblur, are similar to each other when weighted by their abundance (Figure~\ref{fig:figure2}A).
+  A second message is that the different methods differ mainly in the assignment of sequences of lower abundance.
+This can be inferred from the fact that the unweighted comparison (Figure~\ref{fig:figure2}B) shows an increase in dissimilarity between each pair of methods (see additional details in Supplementary and Figure \ref{fig:figure_s3}).
+
+  These comparisons only elucidate the similarity of a pair of methods.
+  In order to determine which tool most accurately recapitulates the reference sequences in the samples, we applied the same pipeline step to process the mock datasets (mock4, mock12, and mock16) and compared the predicted representative sequences with the true sequences and their distribution.
+  The results (Figure~\ref{fig:figure2}C and D) show that the predicted sequence distributions are overall considerably different from the expected ones.
+  The significant variation across datasets indicates that the datasets themselves play a big role in method performance.
   There is no method that outperforms the rest in all datasets (see Supplementary for an extended discussion).
@@
   The \ac{micone} pipeline supports two different chimera checking methods, ``uchime-denovo"~\cite{bolyenReproducibleInteractiveScalable2019}, and ``remove bimera"~\cite{Callahan2016}.
-  We did not notice any significant difference between the two chimera checking methods (Figure~\ref{fig:figure_s4}), implying that they identify and remove mostly the same set of sequences as chimeras.
+  We did not notice any significant difference between the two methods (Figure~\ref{fig:figure_s4}), implying that they identify and remove mostly the same set of sequences as chimeras.
   Since the remove bimera method was originally developed in conjunction with dada2 we use this method as the default for the chimera checking module.
-  The DC step results in a set of unique sequences, these will be referred to as representative sequences in the subsequent steps.
+  The DC step thus results in a reduced set of unique sequences, which will be referred to as representative sequences in the subsequent steps.
 
@@
   The three 16S taxonomic reference databases used in this study are SILVA~\cite{Quast2012}, \acs{gg}~\cite{DeSantis2006} and \acs{ncbi} RefSeq~\cite{Sayers2009} (see Methods).
-  These databases vary significantly in terms of taxonomy hierarchies including species names and phylogenetic relationships~\cite{Balvociute2017}.
+  These databases vary significantly in terms of taxonomy hierarchies, including species names and phylogenetic relationships~\cite{Balvociute2017}.
   Assignment using a particular database also requires a query tool.
-  We use the ``Naive Bayes'' classifier from \ac{qiime2} for the \ac{gg} and SILVA databases and the ``BLAST'' tool (also included as a \ac{qiime2} plugin) for the \ac{ncbi} database.
+  We use the ``Naive Bayes'' classifier from \ac{qiime2} for the \ac{gg} and SILVA databases and the ``BLAST'' tool (included as a \ac{qiime2} plugin) for the \ac{ncbi} database.
   These tools have been well quantified and optimized~\cite{bokulichOptimizingTaxonomicClassification2018}, hence, we make use of the default parameters in our analyses.
@@
   The representative sequences obtained using the default settings of the DC step were used for taxonomic assignment using the three reference databases.
-  Figure~\ref{fig:figure3}A depicts a flow diagram (alluvial plot) that shows how the top 50 representative sequences (sorted by abundance) are assigned a Genus according to the three databases.
-  We observe that the top 50 representative sequences are assigned to 17 Genera in \ac{gg}, 23 in SILVA, and 17 in \ac{ncbi}, however, the sequences that are assigned to a particular Genus vary significantly based on the taxonomy database.
-  We observe that not only does the assigned Genus composition vary significantly, but the percentage of unassigned representative sequences (pink) also differs.
-  The large percentage of unassigned sequences is because a large fraction of the representative sequences is assigned to an "unknown" Genus during the assignment process.
-  A representative sequence might be assigned an "unknown" Genus for one of two reasons: the first is if the taxonomy identifier associated with the sequence in the database did not contain a given Genus; the second, more likely reason, is that the database contains multiple sequences that are very similar to the query (representative) sequence and the consensus algorithm (from \ac{qiime2}) is unable to assign one particular Genus at the required confidence.
-  Additionally, we observe that other abundant Genera like Bacteroides and Escherichia are also assigned to different representative sequences.
-  These changes together result in taxonomic assignments generated from each database leading to drastically different taxonomic abundance distributions.
+  Figure~\ref{fig:figure3}A depicts a flow diagram that shows how the top 50 representative sequences (sorted by abundance) are assigned a Genus according to the three databases.
+  The different databases lead to assignments that qualitatively display similar distributions.
+However, the assigned Genus compositions also display clear differences, as does the percentage of unassigned representative sequences (pink).
+  Some of the differences in Genus composition have a clear explanation, for example, abundant Genera like Bacteroides and Escherichia are assigned to different representative sequences.
+  The large percentage of unassigned sequences is due to the large fraction of the representative sequences assigned to an "unknown" Genus during the assignment process (see Methods).
 
-  After assigning all the representative sequences to taxonomies we perform a pairwise comparison of the similarity between the top 100 assignments (by abundance) from different databases at every taxonomic level (Figure~\ref{fig:figure3}B).
+  After assigning all the representative sequences to taxonomies we performed a pairwise comparison of the similarity between the top 100 assignments (by abundance) from different databases at every taxonomic level (Figure~\ref{fig:figure3}B).
   The comparisons of the assignments beyond the Family level (Family, Genus, and Species) show less than $45\%$ similarity between any pair of databases.
@@
 
-  In order to obtain an absolute measure of the accuracy of the taxonomic assignments, we used the expected reference sequences from the mock datasets as the query sequences for the databases and the expected taxonomic composition as the standard to compare against.
+  In order to obtain an absolute measure of the accuracy of the taxonomic assignments, we used the representative sequences from the DC step for mock datasets as the query sequences and the expected taxonomic composition as the standard to compare against.
   We used the Bray-Curtis distance metric~\cite{virtanenSciPyFundamentalAlgorithms2020} to calculate the distance between the predicted and expected taxonomic distribution (Figure~\ref{fig:figure3}C).
-  Here, we observe that none of the databases perform better than the others in absolute terms and that the dissimilarity with the expected composition is high ($>0.5$ for Family and Genus and $>0.9$ for Species), indicating that all the databases perform poorly when trying to recapture the expected taxonomic composition.
-  % TODO: What are the implications?
+  We find that none of the databases perform better than the others in absolute terms and that the dissimilarity with the expected composition is high ($>0.5$ for Family and Genus and $>0.9$ for Species), indicating that all the databases have some limitations when trying to recapture the expected taxonomic composition.
 
@@
   One reason to choose a particular database could be the frequency of updates and the potential for future growth.
-  Since the \ac{gg} database is used frequently as a 16S reference database in studies~\cite{Balvociute2017}, to enable easier comparison with published networks in literature, we set \ac{gg} as the default for taxonomy assignment.
-  On the other hand, since the degree of similarity of assignments between \ac{ncbi} and \ac{gg} is fairly high (Figure~\ref{fig:figure_s5}), and due to its regular revision and maintenance, the \ac{ncbi} database is also a good choice for the user.
+  Both \ac{gg}, due to its frequent use in the literature~\cite{Balvociute2017}, and \ac{ncbi}, due to its regular revision and maintenance, could be good choices for taxonomy assignment, with few differences observed between the two in our analysis  (Figure~\ref{fig:figure_s5}).
+  In our default pipeline, we choose \ac{gg} as the default method.
+
   The TA step results in a taxonomic counts table that is used as input to the subsequent steps of the pipeline.
-  The count tables at different levels can be obtained through aggregation, for example, Genus count tables were obtained by summing up the counts of the lower taxonomy levels (Species and \ac{otu}) that map to the same higher taxonomy level entity.
+  Note that the count tables at different levels can be obtained through aggregation: for example, Genus count tables were obtained by summing up the counts of the lower taxonomy levels (Species and \ac{otu}) that map to the same higher taxonomy level entity.
 
@@
 
-  \subsection*{NI: Networks generated using different network inference methods show a notable difference in edge-density and connectivity}
+  \subsection*{NI: Different network inference methods drastically affect edge-density and connectivity}
 
-  The ten different network inference methods used in this study are: \acs{spieceasi}~\cite{Kurtz2015}, FlashWeave~\cite{tackmannRapidInferenceDirect2019}, \acs{cozine}~\cite{haCompositionalZeroinflatedNetwork2020a}, \acs{harmonies}~\cite{jiangHARMONIESHybridApproach2020}, \acs{spring}~\cite{yoonMicrobialNetworksSPRING2019}, \acs{mldm}~\cite{Yang2017}, \acs{sparcc}~\cite{Friedman2012,Watts2018}, propr~\cite{quinnProprRpackageIdentifying2017}, Spearman and Pearson correlation methods.
-  These network inference methods fall into two groups, the first set of methods (Pearson, Spearman, \acs{sparcc}, and propr) infer pairwise correlations while the second set infer direct associations (\acs{spieceasi}, FlashWeave, \acs{cozine}, \acs{harmonies}, \acs{spring}, and \acs{mldm}).
+   The ten network inference methods we used in this step fall into two groups: the first set of methods (Pearson, Spearman, \acs{sparcc}~\cite{Friedman2012,Watts2018}, and propr~\cite{quinnProprRpackageIdentifying2017}) infer pairwise correlations while the second set infer direct associations (\acs{spieceasi}~\cite{Kurtz2015}, FlashWeave~\cite{tackmannRapidInferenceDirect2019}, \acs{cozine}~\cite{haCompositionalZeroinflatedNetwork2020a}, \acs{harmonies}~\cite{jiangHARMONIESHybridApproach2020}, \acs{spring}~\cite{yoonMicrobialNetworksSPRING2019}, and \acs{mldm}~\cite{Yang2017}).
+   Note that while Pearson and Spearman methods are included in the pipeline for completeness, they tend to generate a large number of spurious edges as they are not intended for compositional datasets.
+   Thus they are not included in subsequent quantitative analyses.
 
-  For the analysis presented in this section, we used the filtered (performed during the \ac{op} step, refer to Methods) Genus level counts table obtained using the default settings in the DC, TA, and OP steps as the input for the network inference algorithms.
-  Figure~\ref{fig:figure4}A shows the networks inferred from this dataset using the different inference algorithms.
-  The different networks differ vastly in their edge-density and connectivity; even the edges in common to these networks can have their signs inverted.
+  Filtered (see \ac{op} step in Methods) genus-level counts table obtained using the default settings in the previous steps were used as input for the different network inference algorithms (Figure~\ref{fig:figure4}A).
+  Even from a visual inspection, one can see that the different networks differ vastly in their edge-density and connectivity, with common edges often displaying inverted signs.
 
-  In order to obtain a quantitative picture of the differences between the inferred networks, we analyzed the distribution of common nodes and edges (Figure~\ref{fig:figure4} B and C) using UpSet plots~\cite{Lex}.
-  Only \acs{harmonies}, \acs{cozine}, \acs{spieceasi}, \acs{spring}, propr, and \acs{sparcc} are used in the comparison since Pearson and Spearman add a large number of spurious edges as they are not intended for compositional datasets.
-  The results for the node intersections show that the networks have $33$ out of $68$ total unique nodes in common and no network possesses a unique node.
-  The edge intersections in contrast show that only $8$ edges (out of $202$ total unique edges) are in common between all the methods and each network has a large number of unique edges.
-  These results indicate that there is a substantial rewiring of connections in the inferred networks.
-  Hence, in order to determine the interactions that are most likely to exist in the dataset we advocate for the development of a consensus algorithm (refer to next section).
+  In order to quantify the differences between the networks, we analyzed the distribution of common nodes and edges (Figure~\ref{fig:figure4} B and C) using UpSet plots~\cite{Lex}.
+  The node intersection analysis shows that the networks have $33$ out of $68$ total unique nodes in common and that no network possesses a unique node.
+  Edge intersections in contrast show that only $8$ edges (out of $202$ total unique edges) are in common between all the methods and each network has a large number of unique edges.
+  These results indicate that there is a substantial rewiring of connections in different inferred networks, prompting us to identify associations robust across methods, through consensus algorithms.
 
@@
 
-  We observe a high degree of rewiring in the inferred association networks generated by the different methods (Figure \ref{fig:figure4}C), implying that there is a lack of agreement on the associations that should be a part of the final network.
-  Therefore, we have developed a consensus method that would take into consideration the evidence offered by each network inference algorithm and generate a consensus network that would contain the most probable interactions for the input data.
-  We have developed two algorithms for the generation of the consensus network, simple voting and the scaled-sum methods (see Methods).
-  Both approaches combine networks inferred from both correlation-based and direct association methods.
-  First, for the correlation-based methods, we calculate p-values using null models and then merge the p-values using Brown's p-value merging method~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016,faustCoNetAppInference2016} (see Methods).
-  Second, we filter all the inferred networks based on an association strength threshold of 0.1 and a p-value cutoff of 0.05.
-  Finally, we apply the consensus algorithms we have developed on these filtered networks.
-  The default algorithm for the pipeline is the scaled-sum method since this method takes into account the weights of the associations in the determination of the final consensus.
-  By default, of the correlation-based methods, only \acs{sparcc} and propr are included in the calculation of the consensus network.
-  Additionally, the pipeline also allows for the selection of the subset of methods to be used in the calculation of the consensus (for consensus networks that include Spearman and Pearson, see Figure~\ref{fig:figure_s6} in the Supplementary section).
+ Inspired by previous approaches~\cite{bustinceFuzzySetsTheir2008,tsarevApplicationMajorityVoting2018}, we developed two methods that take into consideration the evidence offered by each network inference algorithm and generate a consensus network that contains the most probable associations.
 
-  Similar to what was done for the previous steps of the pipeline, we evaluated the network inference algorithms and the final consensus network using synthetic interaction data.
-  For this purpose, we generated synthetic interaction data using the ``NorTA''~\cite{Kurtz2015} and ``seqtime''~\cite{faustSignaturesEcologicalProcesses2018} methods (see Methods), which have been previously used for benchmarking network inference methods~\cite{Kurtz2015,Weiss2016}.
+  Both of our approaches - simple voting (SV) and scaled-sum (SS) - combine appropriately filtered networks inferred from correlation-based and direct association methods (see Methods, section "Consensus network and p-value merging").
+  We chose the scaled-sum method as the pipeline default since this method takes into account the weights of the associations in the determination of the final consensus.
+  The pipeline allows for any choice of a subset of methods for the consensus calculation. Currently, by default, all direct methods are used, together with \acs{sparcc} and propr.
+
+  Similar to what was done for the previous steps of the pipeline, and in analogy with previous estimations of network inference accuracy~\cite{Kurtz2015,Weiss2016}, we evaluated the network inference algorithms and the final consensus network using synthetic interaction data.
+  For this purpose, we generated synthetic interaction data using the ``NorTA''~\cite{Kurtz2015} and ``seqtime''~\cite{faustSignaturesEcologicalProcesses2018} methods (see Methods).
   For each method, an \ac{otu} counts table is generated based on the selected interaction network topology and properties.
   This counts table is used as the input to the \ac{micone} pipeline to generate predicted associations.
-  The interaction network used to generate the counts table is used as the source of true interactions to calculate the precision and sensitivity for each network inference algorithm (Figure \ref{fig:figure5}).
-  The goal of the consensus network algorithms is to increase the precision of the resultant consensus network by removing edges that are either not present in most of the inference methods or whose association strength is low.
-  In Figure \ref{fig:figure5} we show that the consensus algorithm, especially the scaled-sum method, captures true associations with more precision, but as a consequence, the sensitivity of the predictions is reduced.
-  Among the individual inference methods, we observe that \acs{spieceasi} has the best average precision ($0.944$)
-  Overall, we see that the scaled-sum method for $p=1.0$ performs the best in terms of average precision ($1.000$).
-  However, the scaled-sum method for $p=0.667$ has the second-highest average precision ($0.985$) and does not underperform in terms of sensitivity.
-  Therefore, the consensus networks provide the means to obtain a short list of associations that would have a high probability of being present in the real interaction network.
-  Furthermore, we also find that the consensus algorithm performs drastically better in terms of precision (0.820 for scaled-sum method vs.
-0.624 for \acs{spieceasi}) in recapturing the original interactions in the ``seqtime'' data.
-  The precision and sensitivity graphs on the seqtime data can be found in the Supplementary section (Figure~\ref{fig:figure_s7}).
+  The interaction network used to generate the counts table is used as the source of true interactions to calculate the precision (Figure \ref{fig:figure5}) and sensitivity (Figure \ref{fig:figure_s6} and \ref{fig:figure_s7}) for each network inference algorithm.
+  As shown in Figure \ref{fig:figure5} the consensus algorithm, especially the scaled-sum method, captures true associations with high precision (through the removal of edges that are either not present in most of the inference methods or whose association strength is low across methods).
+  Overall, the scaled-sum method for $p=1.0$ performs the best (precision = $1.000$ for both NorTA and seqtime). 
+  The scaled-sum method for $p=0.667$ (default option in the pipeline) has the second-highest average precision ($0.985$ with NorTA; $0.820$ with seqtime), without displaying significant reduction in sensitivity (Figure \ref{fig:figure_s6} and Figure~\ref{fig:figure_s7}).
+  Therefore, the consensus networks provide the means to obtain a short list of associations that would have a high probability of being present in the real association network.
 
@@
 
-  \subsection*{Overall assessment of the impact on inferred networks}
+  \subsection*{Impact of different pipeline steps on co-occurrence networks}
 
-  In order to analyze the effect of different processing methods on the inferred co-occurrence networks, we generated networks using all possible combinations of methods and estimated the variability due to each choice (Figure \ref{fig:figure6}A).
-  The networks used in this analysis are those directly inferred from the correlation-based and direct inference methods and to which the consensus algorithm is yet to be applied.
-  The effects of various steps on the final co-occurrence network are estimated by building a linear model of the edges of the network as a function of the various steps in the pipeline workflow (see Methods).
+  In order to analyze the effect of different processing methods on the inferred co-occurrence networks (before consensus estimation), we generated networks using all possible combinations of methods and quantified the variability due to each choice (Figure \ref{fig:figure6}A).
+ This was achieved by building a linear model of the edges of the network as a function of the various steps in the pipeline workflow (see Methods).
   Figure \ref{fig:figure6}A, shows the percentage of total variation among the co-occurrence networks due to the different steps of the pipeline.
-  The \ac{ta} step or more specifically the choice of 16S reference database, contributes the most ($\sim65.4\%$) to the variation in the networks, followed by the \ac{op} step ($\sim26.8\%$).
+  The \ac{ta} step, or more specifically the choice of 16S reference database, contributes the most ($\sim65.4\%$) to the variation in the networks, followed by the \ac{op} step ($\sim26.8\%$).
   This result highlights the importance of the taxonomy assignment step in the 16S data analysis workflow, implying that a change in the reference database will result in drastically different inferred networks.
-  The reason for this seems to derive from the differential labeling of constitutive representative sequences into taxonomic entities (Figure~\ref{fig:figure3} and Figure~\ref{fig:figure_s5}), which drastically alter the nodes and hence the underlying network topology.
-  The fraction labeled as the residual is an artifact that arises when multiple steps are changed at the same time, however, this can be ignored as the fraction is negligible.
+  This is likely due to the differential assignment of representative sequences to taxonomic entities (Figure~\ref{fig:figure3} and Figure~\ref{fig:figure_s5}), which drastically alter the nodes and hence the underlying network topology.
 
   The effects of the different steps of the pipeline on the inferred networks can be best visualized through dimensionality reduction.
-  Figure \ref{fig:figure6}B is the PCA plot of all the generated co-occurrence networks, colored by the databases in the TA step (left) and the network inference algorithms in the NI step (right).
-  In other words, each point corresponds to a network generated using different combinations of tools.
-  The importance of the TA step, shown in Figure \ref{fig:figure6}A, is also reflected here since most of the networks can be separated based on the database used (Figure \ref{fig:figure6}B).
-  However, it is also evident that some networks, especially those inferred using the direct association network inference methods, are much closer in the PCA plot regardless of the reference database used.
+  The PCA in Figure \ref{fig:figure6}B shows all the above networks, colored by the databases used in TA (left) and by the NI approach (right).
+  The major effect of the TA step choice, shown before in Figure \ref{fig:figure6}A, is also reflected in the PCA plot, where networks segregate based on the database used (Figure \ref{fig:figure6}B).
+  It is also evident that some networks, especially those inferred using the direct association network inference methods, are much closer in the PCA plot regardless of the reference database used.
   Additionally, the variation (on the first two principal axes) between the networks decreases when the low abundance \ac{otu}s are removed from the network (Figure~\ref{fig:figure_s1}).
-  These results suggest that the most important criterion for accurate comparative analysis of co-occurrence networks is the taxonomy reference database followed by the level of filtering of the taxonomy tables prior to network inference.
+  These results suggest that the most important criterion for accurate comparative analysis of co-occurrence networks is the taxonomy reference database followed by the level of filtering of the taxonomy tables.
 
@@
 
-  The systematic analyses performed in the previous sections illustrate that the choice of tools and parameters can have a big impact on the final co-occurrence network.
-  The mock communities and synthetic data provide an opportunity to select the combinations of tools that yield more accurate and robust results.
-  However, our analysis could not yield the best protocol for some of the steps of the workflow (e.g. for the \ac{ta} step).
-  Despite this partial degree of assessment, we wish to suggest a combination of tools and parameters that led to the best performance based on the mock and synthetic data and which displayed the highest robustness to switching to alternative methods.
-  These tools and parameters are chosen as the defaults for the pipeline and are given in Table~\ref{tab:micone_tools}.
+  The systematic analyses in the previous sections illustrate that the choice of tools and parameters can have a big impact on the final consensus co-occurrence network.
+  However, the mock communities and synthetic data provide an opportunity to select combinations of tools that yield accurate and robust results.
+  As highlighted in the above sections for individual steps, we propose a set of tools and parameters as the defaults for the pipeline (Table~\ref{tab:micone_tools}).
 
-  Figure~\ref{fig:figure7}A shows the comparison of the default network against networks generated by altering one of the steps of the pipeline from the default.
-  This result suggests that the most significant changes (many new nodes and edges are added) occur when the \ac{ta} step, \ac{op} step, and the \ac{ni} step are changed.
-  However, the L1 distance of networks generated by altering one of the steps of the pipeline from the default against the default network (Figure~\ref{fig:figure7}B) shows that the biggest deviations from the default network occur when the \ac{ni} step is changed from the consensus algorithm to one of the individual algorithms.
-  These results indicate that the biggest rewiring in networks occurs when the network inference algorithm is changed and that the changes in the other steps do not create as significant of a difference.
-  The change of the taxonomy database and the filtering in the \ac{op} step when turned off also adds a large number of new nodes, but the edges between the common nodes (with the default network) are fairly conserved.
+  Figure~\ref{fig:figure7}A shows the comparison of this default network against networks generated by altering one of the steps of the pipeline relative to the default.
+  This result suggests that the most significant changes occur when the \ac{ta} step, \ac{op} step, and the \ac{ni} step are changed.
+  However, the L1 distance (see Methods) between networks generated by altering one of the steps and the default network (Figure~\ref{fig:figure7}B) shows that the biggest deviations from the default network occur when the \ac{ni} step is changed from the consensus algorithm to one of the individual algorithms.
+  % TODO: Update this and the corresponding figure
+  XXXXXXXX Notably, despite the overall big effect of the TA database on the whole pipeline, this choice does not , but the edges between the common nodes (with the default network) are fairly conserved.
 

### 2022-09-14 sections/results.tex
Update on Overleaf.

@@
   \ac{micone} incorporates various popular, publicly available tools as well as custom Python modules for 16S data analysis and network inference (see Methods).
-  The different steps that are a part of the \ac{micone} co-occurrence network inference workflow (Figure~\ref{fig:figure1}) can be grouped into five major modules; (i) sequence processing (SP); (ii) denoising and clustering (DC); (iii) taxonomy assignment (TA); (iv) \ac{otu} processing (OP); (v) network inferences (NI).
+  The different steps that are a part of the \ac{micone} co-occurrence network inference workflow (Figure~\ref{fig:figure1}) can be grouped into five major modules; (i) \ac{sp}; (ii) \ac{dc}; (iii) \ac{ta}; (iv) \ac{op}; (v) \ac{ni}.
   Each process in the pipeline can be implemented through  multiple tools (see Methods and Figure~\ref{fig:figure1}).
@@
   The systematic analyses in the previous sections illustrate that the choice of tools and parameters can have a big impact on the final consensus co-occurrence network.
-  However, the mock communities and synthetic data provide an opportunity to select combinations of tools that yield accurate and robust results.
+  However, the mock communities and synthetic data provide an opportunity to select combinations of tools that yield the most accurate and robust results.
   As highlighted in the above sections for individual steps, we propose a set of tools and parameters as the defaults for the pipeline (Table~\ref{tab:micone_tools}).
 
-  Figure~\ref{fig:figure7}A shows the comparison of this default network against networks generated by altering one of the steps of the pipeline relative to the default.
-  This result suggests that the most significant changes occur when the \ac{ta} step, \ac{op} step, and the \ac{ni} step are changed.
-  However, the L1 distance (see Methods) between networks generated by altering one of the steps and the default network (Figure~\ref{fig:figure7}B) shows that the biggest deviations from the default network occur when the \ac{ni} step is changed from the consensus algorithm to one of the individual algorithms.
-  % TODO: Update this and the corresponding figure
-  XXXXXXXX Notably, despite the overall big effect of the TA database on the whole pipeline, this choice does not , but the edges between the common nodes (with the default network) are fairly conserved.
-
-  The low L1 distances show us that not only does the consensus help reconcile the variability in the edges due to multiple network inference algorithms, but it also helps reduce the variability due to the other steps of the pipeline as well.
-  This result is different from that observed in Figure~\ref{fig:figure6} since in this particular analysis all the steps other than the \ac{ni} step have the consensus algorithm applied to them, which drastically reduces the edges in the graph that have lower probabilities.
-  This reinforces the importance of the consensus network in increasing the robustness of the inferred networks to change in processing methods.
+  [REVISIT THIS PART]
+  Figure~\ref{fig:figure7} shows the comparison of this default network against networks generated by altering one of the steps of the pipeline relative to the default.
+  This result, both visually (Figure~\ref{fig:figure7}A), and quantitatively (Figure~\ref{fig:figure7}B)  suggests that the most significant changes occur when the \ac{ta} step and the \ac{op} changed.
+  Notably, however despite the overall big effect of the TA database on the whole pipeline, the edges between the common nodes (with the default network) are fairly conserved, suggesting that the pipeline, with the consensus network at the end, provides a robust avenue for inferring associations.
 

### 2022-09-14 sections/methods.tex
Fix: Figure 7 issues

@@
   Finally, a consensus method is used to identify the most probable match from the list of best matches.
-  The pipeline incorporates \ac{gg} 13\_8~\cite{DeSantis2006}, SILVA 138~\cite{Quast2012} and the \ac{ncbi} (16S RefSeq as of Oct 2021)~\cite{Sayers2009} databases for taxonomy assignment.
+  The pipeline incorporates \ac{gg} 13\_8~\cite{DeSantis2006} (99\% identity), SILVA 138~\cite{Quast2012} (99\% identity) and the \ac{ncbi} (16S RefSeq as of Oct 2021)~\cite{Sayers2009} databases for taxonomy assignment.
   SILVA and \ac{gg} are two popular 16S databases used for taxonomy identification and the \ac{ncbi} RefSeq nucleotide database contains 16S rRNA sequences as a part of two BioProjects - 33175 and 33317.

### 2022-09-14 scripts/figure7/figure7_data.py

@@ -224,10 +224,12 @@ def main(
         "CC": dict(),
         "TA": dict(),
         "OP": dict(),
-        "NI": dict(),
+        # "NI": dict(),
     }
     for step_folder in tqdm(list(folder.iterdir())):
         step_name = step_folder.stem
+        if step_name == "NI":
+            continue
         assert step_name in networks_dict
         for process_folder in step_folder.iterdir():
             process_name = process_folder.stem
@@ -268,7 +270,7 @@ def main(
         f"CC_{cc}": networks_dict["CC"][cc],
         f"TA_{ta}": networks_dict["TA"][ta],
         f"OP_{op}": networks_dict["OP"][op],
-        f"NI_{ni}": networks_dict["NI"][ni],
+        # f"NI_{ni}": networks_dict["NI"][ni],
         "default": networks_dict["default"]["default"],
     }
     nodes, edges = write_networks(

### 2022-09-14 scripts/figure7/figure7_plot.R

@@ -31,7 +31,7 @@ clustering_gml <- paste0(data_folder1, "DC_closed_reference(gg_97).gml")
 chimera_checking_gml <- paste0(data_folder1, "CC_uchime.gml")
 database_gml <- paste0(data_folder1, "TA_blast(ncbi).gml")
 otu_filtering_gml <- paste0(data_folder1, "OP_normalize_filter(off).gml")
-network_inference_gml <- paste0(data_folder1, "NI_sparcc.gml")
+# network_inference_gml <- paste0(data_folder1, "NI_sparcc.gml")
 l1_distance_csv <- paste0(data_folder1, "l1_distance_to_ref.csv")
 edit_distance_csv <- paste0(data_folder1, "edit_distance_to_ref.csv")
 combined_gml_2 <- paste0(data_folder2, "combined.gml")
@@ -106,11 +106,11 @@ clustering_plot <- plot_network(clustering_gml, combined_layout, 0.3, "DC=CR", p
 chimera_checking_plot <- plot_network(chimera_checking_gml, combined_layout, 0.3, "CC=uchime", palette, 4)
 database_plot <- plot_network(database_gml, combined_layout, 0.3, "TA=BLAST(NCBI)", palette, 2)
 otu_filtering_plot <- plot_network(otu_filtering_gml, combined_layout, 0.3, "OP=Filter(off)", palette, 4)
-network_inference_plot <- plot_network(network_inference_gml, combined_layout, 0.3, "NI=SparCC", palette, 5)
-legend_grob <- get_legend(network_inference_plot)
+legend_grob <- get_legend(otu_filtering_plot)
+# network_inference_plot <- plot_network(network_inference_gml, combined_layout, 0.3, "NI=SparCC", palette, 5)
 
 combined_plot <- ggarrange(
-    default_plot, database_plot, chimera_checking_plot, clustering_plot, otu_filtering_plot, network_inference_plot,
+    default_plot, database_plot, chimera_checking_plot, clustering_plot, otu_filtering_plot,
     ncol = 3, nrow = 2, common.legend = TRUE, legend = "right", legend.grob = legend_grob
 )
 a_plot <- annotate_figure(combined_plot, fig.lab = "A", fig.lab.pos = "top.left", fig.lab.size = 20)
@@ -164,10 +164,10 @@ plot_network_c <- function(network_file, combined_layout, interaction_threshold,
   graph_layout <- data.frame(temp_layout)
[…]

### 2022-09-16 scripts/figure7/figure7_data.py
Fix: Figure 7 filtering thresholds

@@ -267,13 +267,14 @@ def main(
                 network_group = NetworkGroup(networks_for_consensus)
             cids = list(range(len(network_group.contexts)))
             consensus_network = network_group.get_consensus_network(
-                cids, method="scaled_sum", parameter=0.5
+                cids, method="scaled_sum", parameter=0.333
             )
             # STEP3: Create a simple graph version of the consensus
             simple_consensus = consensus_network.to_network()
             simple_consensus.interaction_threshold = 0.1
-            simple_consensus.pvalue_threshold = 0.05
-            networks_dict[step_name][process_name] = simple_consensus.filter(True, True)
+            networks_dict[step_name][process_name] = simple_consensus.filter(
+                False, True
+            )
     # STEP4: Then, feed the above to the previous function
     output_path = pathlib.Path(output) / dataset
     output_path.mkdir(parents=True, exist_ok=True)

### 2022-09-16 scripts/figure7/figure7_plot.R

@@ -27,7 +27,7 @@ data_folder1 <- paste0(data_folder, dataset1, "/")
 data_folder2 <- paste0(data_folder, dataset2, "/")
 combined_gml <- paste0(data_folder1, "combined.gml")
 default_gml <- paste0(data_folder1, "default.gml")
-clustering_gml <- paste0(data_folder1, "DC_closed_reference(gg_97).gml")
+clustering_gml <- paste0(data_folder1, "DC_de_novo.gml")
 chimera_checking_gml <- paste0(data_folder1, "CC_uchime.gml")
 database_gml <- paste0(data_folder1, "TA_blast(ncbi).gml")
 otu_filtering_gml <- paste0(data_folder1, "OP_normalize_filter(off).gml")
@@ -64,24 +64,97 @@ plot_facet_network <- function(network_file, interaction_threshold) {
         filter(abs(weight) > interaction_threshold) %>%
         mutate(color = get_edgecolor(weight)) %>%
         mutate(layer = get_layer(title)) %>%
-        mutate(step = factor(title, levels=c("default", "DC=closed_reference(gg_97)", "CC=uchime", "TA=blast(ncbi)", "OP=normalize_filter(off)"))) %>%
+        mutate(step = factor(title, levels = c("default", "DC=de_novo", "CC=uchime", "TA=blast(ncbi)", "OP=normalize_filter(off)"))) %>%
+        mutate(weight = abs(weight)) %>%
         activate(nodes) %>%
-        mutate(isolated = node_is_isolated()) %>%
-        filter(isolated == FALSE)
+        filter(!node_is_isolated())
     graph_plot <- ggraph(graph = graph, layout = "nicely") +
         geom_edge_link(aes(color = color)) +
-        # geom_edge_link(aes(color = layer, edge_linetype = color)) +
         geom_node_point() +
-        facet_edges(~step, drop = FALSE) +
-    scale_edge_color_manual(values = c(negative="#d95f02", positive="#1b9e77")) +
-    scale_edge_linetype_manual(values = c(negative="solid", positive="solid")) +
-    scale_edge_alpha_manual(values = c(negative=0.6, positive=0.3)) +
-        # scale_edge_color_manual(values = c(foreground = foreground, background = "#c1c1c1", common = "black")) +
-        # scale_edge_linetype_manual(values = c(negative = "dashed", positive = "solid")) +
[…]

### 2022-09-16 scripts/figure7/figure7_process.sh

@@ -5,7 +5,7 @@ python -W ignore figure7_data.py \
   --folder "../../data/figure7/input/$DATASET" \
   --color_key_level "Family" \
   --multigraph "False" \
-  --dc "closed_reference(gg_97)" \
+  --dc "de_novo" \
   --cc "uchime" \
   --ta "blast(ncbi)" \
   --op "normalize_filter(off)" \
@@ -16,7 +16,7 @@ python -W ignore figure7_data.py \
   --folder "../../data/figure7/input/$DATASET" \
   --color_key_level "Family" \
   --multigraph "False" \
-  --dc "closed_reference(gg_97)" \
+  --dc "de_novo" \
   --cc "uchime" \
   --ta "blast(ncbi)" \
   --op "normalize_filter(off)" \

### 2022-09-17 sections/methods.tex
Change: Use L1 distance for Figure 7 and update figures

@@
   In all UniFrac comparisons, we set a count threshold of 10, such that if the count of the representative sequences in a particular sample is less than the threshold, it is omitted from the analysis.
+  The expected sequences from the mock communities are trimmed to the V4 region before being subject to UniFrac analyses.
 
@@
   $$\frac{|A_{x} - A_{y}|}{2}$$
-  where, $A$ is the adjacency matrix of the network.
+  where, $A$ is the adjacency matrix of the network where edges with weights greater than $0.1$ are assigned a value of $+1$ and those lesser than $-0.1$ are assigned a value of $-1$.
 

### 2022-09-17 scripts/figure7/figure7_plot.R

@@ -94,20 +94,20 @@ ggsave(output_file_a, width = 11, height = 5.5)
 #########################################
 # b
 #########################################
-edit_data <- read.csv(edit_distance_csv, header = TRUE, sep = ",")
-edit_data[edit_data == "de_novo"] <- "OR"
-edit_data[edit_data == "closed_reference(gg_97)"] <- "CR"
-edit_data[edit_data == "de_novo"] <- "DN"
-edit_data[edit_data == "dada2"] <- "D2"
-edit_data[edit_data == "deblur"] <- "DB"
-edit_data[edit_data == "naive_bayes(silva_138_99)"] <- "NaiveBayes(SILVA)"
-edit_data[edit_data == "blast(ncbi)"] <- "BLAST(NCBI)"
-edit_data[edit_data == "normalize_filter(on)"] <- "Filter(on)"
-edit_data[edit_data == "normalize_filter(off)"] <- "Filter(off)"
+l1_data <- read.csv(l1_distance_csv, header = TRUE, sep = ",")
+l1_data[l1_data == "de_novo"] <- "OR"
+l1_data[l1_data == "closed_reference(gg_97)"] <- "CR"
+l1_data[l1_data == "de_novo"] <- "DN"
+l1_data[l1_data == "dada2"] <- "D2"
+l1_data[l1_data == "deblur"] <- "DB"
+l1_data[l1_data == "naive_bayes(silva_138_99)"] <- "NaiveBayes(SILVA)"
+l1_data[l1_data == "blast(ncbi)"] <- "BLAST(NCBI)"
+l1_data[l1_data == "normalize_filter(on)"] <- "Filter(on)"
+l1_data[l1_data == "normalize_filter(off)"] <- "Filter(off)"
 
 box_plot <- ggdotplot(
-    edit_data,
-    x = "step", y = "edit_distance", label = "process", repel = TRUE, order = c("DC", "CC", "TA", "OP"), fill = "step", add = "jitter", palette = "Set2", ylab = "edit distance", xlab = "Pipeline step"
+    l1_data,
+    x = "step", y = "l1_distance", label = "process", repel = TRUE, order = c("DC", "CC", "TA", "OP"), fill = "step", add = "jitter", palette = "Set2", ylab = "L1 distance", xlab = "Pipeline step"
 ) +
     theme(
         text = element_text(size = 18),

### 2022-09-27 tables/figure_s8_caption.tex
Fix: Thresholds and alpha in Figure 4

--- 2022-09-01 sections/supplementary.tex
@@
-%!TEX root = ../main.tex
-
-\newpage
-\section*{Supplementary}
-
-  {\color{red} Tables S1-S5 will be moved to GitHub}
-
-  \renewcommand{\thefigure}{S\arabic{figure}}
-  \setcounter{figure}{0}
-
-  \renewcommand{\thetable}{S\arabic{table}}
-  \setcounter{table}{0}
-
-  \input{sections/supplementary_table.tex}
-
-  \subsection*{Figures}
-
-    \begin{figure}[H]
-      \centering
-      \includegraphics[width=1.0\linewidth]{figure_s1.pdf}
-    \end{figure}
-    \begin{figure}[H]
-      \centering
-        \caption{
-          \textbf{The networks which were filtered during the \ac{op} step are less variable}
-          All combinations of inferred networks are shown as points on a PCA plot and shows the variation in the inferred networks in the first two principal axes (similar to Figure~\ref{fig:figure6}B).
-          Each point on the PCA plot represents a network inferred using different combinations of tools and parameters that are available in the \ac{micone} pipeline.
-          The points are colored by the tools or parameters used in the \ac{dc} step (A) and the \ac{op} step (B).
-          The \ac{dc} step does not seem to have any correlation with the similarity of networks on the PCA plot, but all the networks with filtering at the \ac{op} step are similar to each other.
-          This further confirms that the variability in the networks decreases upon filtering out the taxonomic entities at low abundance.
-        }
-      \label{fig:figure_s1}
-    \end{figure}
-    \FloatBarrier
-    \newpage
-
-    % TODO: Verify this about t-SNE
-    \begin{figure}[H]
-      \centering
-      \includegraphics[width=1.0\linewidth]{figure_s2.pdf}
-    \end{figure}
-    \begin{figure}[H]
-      \centering
-        \caption{
-          \textbf{The t-SNE plot of all the inferred networks clusters the networks based on the taxonomy reference database used}.
-          Each point on the t-SNE plot represents a network inferred using different combinations of tools and parameters that are available in the \ac{micone} pipeline.
-          The points are colored by the tools and parameters used in \ac{dc} step (A), \ac{ta} step (B), \ac{op} step (C) and \ac{ni} step (D).
-          The separation of the points based on taxonomy reference database shows that the points might cluster based on reference database in high-dimensional space.
-        }
-      \label{fig:figure_s2}
-    \end{figure}
-    \FloatBarrier
-    \newpage
-
-    \begin{figure}[H]
-      \centering
-      \includegraphics[width=1.0\linewidth]{figure_s3.pdf}
-    \end{figure}
-    \begin{figure}[H]
-      \centering
-        \caption{
-          \textbf{The UniFrac distance between the 1000 most abundant representative sequences is higher than that when all sequences are considered}.
-          Each value is the average UniFrac distance between the reference sequences generated by the various methods in the \ac{dc} step (similar to Figure~\ref{fig:figure2}).
-          There is an increase in both weighted and unweighted UniFrac distances compared to when all the representative sequences are considered.
-          This shows that the 1000 most abundant representative sequences generated by the methods are not as similar to each other.
-          And since the weighted UniFrac is much smaller than the unweighted UniFrac distance, we can conclude that those reference sequences that are present in the middle of the abundance distribution are dissimilar.
-        }
-      \label{fig:figure_s3}
-    \end{figure}
-    \FloatBarrier
-    \newpage
-
-    \begin{figure}[H]
-      \centering
-      \includegraphics[width=1.0\linewidth]{figure_s4.pdf}
-    \end{figure}
-    \begin{figure}[H]
-      \centering
-        \caption{
-          \textbf{The weighted and unweighted UniFrac distances between representative sequences generated using remove bimera and uchime for each denoising method are low.}
-          The two chimera checking methods, uchime and remove bimera, produce similar outputs.
-          With the exception of de novo and open reference under the unweighted UniFrac metric, all the other methods have low dissimilarity.
-          This is especially true for the \ac{dada2} and Deblur methods which are the recommended denoising methods in the \ac{micone} pipeline.
-          Therefore, remove bimera is recommended as the default chimera method if one is using \ac{dada2} and uchime-denovo when one is using Deblur, since these methods were developed for these respective algorithms (\ac{qiime2} uses uchime-denovo in the Deblur workflow).
-        }
-      \label{fig:figure_s4}
-    \end{figure}
-    \FloatBarrier
-    \newpage
-
-    \begin{figure}[H]
-      \centering
-      \includegraphics[width=1.0\linewidth]{figure_s5.pdf}
-    \end{figure}
-    \begin{figure}[H]
-      \centering
-        \caption{
-          \textbf{The pairwise comparison of assignments generated using different databases for all representative sequences has a higher proportion of mismatches}.
-          The comparison made here is similar to Figure~\ref{fig:figure3}B, but instead of the top 100 taxonomic entities (by abundance), all the assignments from one database are matched with those from the other two databases.
-          Higher percentage of mismatches implies that the matching of the taxonomies in the more abundant sequences (top 100) are more consistent.
-        }
-      \label{fig:figure_s5}
-    \end{figure}
-    \FloatBarrier
-    \newpage
-
-  \begin{figure}[H]
-    \centering
-    \includegraphics[width=1.0\linewidth]{figure_s6.pdf}
-  \end{figure}
-  \begin{figure}[H]
-    \centering
 \caption{
@@
     The different consensus based methods used are: pvalue merging (PM), scaled-sum (SS) and simple voting (SV) method.
-        With the addition of Pearson and Spearman in the consensus calculation we observe a decrease in the average precision scores of both the consensus algorithms (compared to Figure~\ref{fig:figure5}.
+    With the addition of Pearson and Spearman in the consensus calculation we observe a decrease in the average precision scores of both the consensus algorithms (compared to Figure~\ref{fig:figure_s6}.
     The overall best precision was still consistently obtained by the scaled-sum method (0.985, 0.963 and 1.000), but the performance was not stable to an increase in the parameter value.
@@
 }
-    \label{fig:figure_s6}
-  \end{figure}
-  \FloatBarrier
-  \newpage
-
-  \begin{figure}[H]
-    \centering
-    \includegraphics[width=1.0\linewidth]{figure_s7.pdf}
-  \end{figure}
-  \begin{figure}[H]
-    \centering
-      \caption{
-        \textbf{The precision and sensitivity of the inferred networks on the seqtime synthetic interaction data.}
-        The ``I'' prefix indicates the individual network inference methods and the ``X'' prefix denotes the different consensus based methods.
-        The different consensus based methods used are: pvalue merging (PM), scaled-sum (SS) and simple voting (SV) method.
-        Pearson and Spearman methods are not used in the calculation of the consensus.
-        Among all the independent network inference methods, \ac{spieceasi} has the best average precision (0.624), but the overall best precision was consistently obtained by the scaled-sum method (0.688, 0.820 and 1.000).
-        The simple voting method when using presence of edges in all inferred networks as a requirement (parameter value 1.000), also outperforms \ac{spieceasi} on average precision (0.692).
-        These results show that the scaled-sum method is not only much better suited for inferring robust and accurate interactions from count data regardless of the distributions and topologies.
-        But, it is also capable of accurately extracting real interactions (gLV interactions) from count data.
-      }
-    \label{fig:figure_s7}
-  \end{figure}
-  \FloatBarrier
-  \newpage
-
-  \subsection*{Processing the FMT data}
-
-    \subsubsection*{Data download and pre-processing}
-    The main biological dataset used in this study was the collection of 16S rRNA sequencing reads from stool samples (healthy and autistic individuals) for a fecal microbiome transplant study~\cite{Kang2017}.
-    The data containing the 16S sequencing reads (V4 region) was downloaded from Qiita~\cite{qiita} (study ID: 10532).
-    Only runs 2, 3, and 4 were used for the subsequent analysis as these runs consisted of paired-end sequencing data and run 1 contained single-end data.
-    The sample metadata was updated to contain only BMI, sex, height, weight and experimental group.
-    This was necessary as two of the network inference algorithms (\ac{mldm} and FlashWeave) required information about environmental heterogeneity.
-    However, these environmental correlations were not included in the current analyses.
-
-    \subsubsection*{Processing using the \ac{micone} pipeline}
-    The data was then processed using the \ac{micone} pipeline starting at the \ac{sp} step and ending at the \ac{ni} step with the consensus algorithm.
-    The configuration files (main.nf and nextflow.config) used to run the \ac{micone} pipeline as well the details of the pipeline execution (dag, report, timeline and trace) are in the "runs/FMT" directory of the data and scripts repository (\href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper})
-    % TODO: Store pipeline results
-    The results of the pipeline execution for reproducing the analyses in the manuscript are stored on Zenodo[REF] at ...
-
-  \subsection*{Processing the mock data}
-
-    \subsubsection*{Data download and pre-processing}
-    The mock datasets, mock4, mock12 and mock16 used for this study, were obtained from mockrobiota~\cite{Bokulich2016}.
-    Mock 4 is a mock community composed of 21 bacterial strains represented in equal abundances in two replicate samples, and the same strains represented in uneven abundances in two other replicate samples.
-    Mock 12 is composed of 27 bacterial strains containing closely related taxa, the members of which were chosen in part for their well-separated 16S rRNA gene sequences.
-Some pairs of strains differ by as little as one nucleotide, but all the strains are distinguishable over the sequenced region of the 16S rRNA gene.
-    Mock 16 is a mock community composed of even amounts of purified genomic DNA from 49 bacteria and 10 archaea.
-    The datasets did not require any preprocessing and could be directly used as input to the pipeline
-
-    \subsubsection*{Processing using the \ac{micone} pipeline}
-    The data was processed using the \ac{micone} pipeline starting at the \ac{sp} step and ending at the \ac{op} step with the filtered taxonomic tables as the final output.
-    The configuration files (main.nf and nextflow.config) used to run the \ac{micone} pipeline as well the details of the pipeline execution (dag, report, timeline and trace) are in the "runs/mock*" directory of the data and scripts repository (\href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper})
-    % TODO: Store pipeline results
-    The results of the pipeline execution for reproducing the analyses in the manuscript are stored on Zenodo at ...
-
-  \subsection*{Synthetic interaction data}
-
-    \subsubsection*{Data generation}
-    The synthetic interaction data for the study were generated using two methods.
-    The first method, ``seqtime''~\cite{faustSignaturesEcologicalProcesses2018} utilized generalized Lotka-Volterra (gLV) equations to model the microbial community dynamics and made use of the Klemm–Eguı́luz algorithm to generate clique-based interaction networks~\cite{Rottjers2018}.
-    We used the seqtime R package to simulate communities with different numbers of species and samples (see Methods for details).
-    The second method, ``NorTA'' used the Normal to Anything (NorTA) approach coupled with a given interaction network topology to generate the abundance distribution of the microbial community~\cite{Kurtz2015}.
-    We used the spieceasi R package to simulate communities with different abundance distributions and network topologies (see Methods for details).
-    The scripts to generate these datasets can be found in the synthetic data and scripts repository (\href{https://github.com/segrelab/MiCoNE-synthetic-data}{https://github.com/segrelab/MiCoNE-synthetic-data})
-
-    \subsubsection*{Processing using the \ac{micone} pipeline}
-    The data was processed using the \ac{micone} pipeline using only the \ac{ni} step with the consensus networks as the final output.
-    The configuration files (main.nf and nextflow.config) used to run the \ac{micone} pipeline as well the details of the pipeline execution (dag, report, timeline and trace) are in the "runs/norta" and "runs/seqtime" directories of the data and scripts repository (\href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper})
-    % TODO: Store pipeline results
-    The results of the pipeline execution for reproducing the analyses in the manuscript are stored on Zenodo at ...
-
-  \subsection*{Network metrics}
-
-  In Table~\ref{tab:network_metrics} we show various global network metrics calculated for each tool in the pipeline.
-  All the networks that make use of a particular tool are grouped together, and the following average metrics are calculated for each group:
-  \begin{enumerate}
-    \item The average shortest path length describes the average of all the shortest paths in the graph.
-No number is reported if the graph is not connected, therefore, the results indicate that none of the networks that make use of \ac{harmonies}, \ac{cozine}, \ac{spring}, \ac{spieceasi} and Pearson are connected.
-    \item The average clustering is the average clustering coefficient of the graph.
-The closer the value is to 1.0, the more densely connected is the graph.
-We can observe that the networks that use correlation-based methods have the highest values while the direct-association based methods have the lowest.
-    \item The number of connected components is the highest for the direct-association based methods and the lowest for the correlation-based methods.
-In the case of propr, all the networks have only one giant component.
-    \item The modularity metric is the modularity over all partitions in a graph calculated using a label propagation algorithm~\cite{cordascoCommunityDetectionSemisynchronous2010}.
-Positive values imply that there are more edges between vertices of the same type than we would expect by chance, and negative implies that there are less.
-The networks inferred by \ac{mldm} report very few edges, and skew the average modularity scores.
-This could also be an artifact of incomplete converge of the \ac{mldm} algorithm for some combinations.
-    \item Node connectivity refers to the minimum number of nodes that must be removed from the graph to make it disconnected.
-We observe that only the networks generated using propr have a high value since most of these networks are connected.
-    \item Degree assortativity coefficient measures the similarity of connections in the graph with respect to the node degree.
-Again we observe that the direct-association based methods have negative degree assortativity, meaning that there are many hubs in these networks.
-The correlation-based methods have positive values implying that in these networks nodes with similar degrees attach to one another.
-  \end{enumerate}
-  All the metrics were calculated using the \texttt{networkx} Python package~\cite{hagbergExploringNetworkStructure2008}.
-
-  \subsection*{p-value merging}
-
-  Fisher~\cite{fisher_224a_1948} proposed that for $k$ independent p-values, each generated by $k$ different methods and denoted by $\bar{P}^i$ (notations are same as used in the "Consensus network and p-value merging" subsection of the Methods), the following will hold true for the statistic $\Psi$:
-  \begin{equation*}
-    \begin{aligned}
-      \Psi &= \sum_{i=1}^k -2 \log \left( \bar{P}^i \right) \\
-        \Psi &\sim \chi^2_{2k}
-    \end{aligned}
-  \end{equation*}
-
-  Brown~\cite{brown_400_1975} extended Fisher's method to dependent p-values by using a re-scaled $\chi^2$ distribution:
-  \begin{equation*}
-    \Psi \sim c \chi^2_{2f}
-  \end{equation*}
-  where, $f$ is the degrees of freedom and $c$ is the scale factor and are given by:
-  \begin{equation*}
-    f = \frac{\mathrm{E}[\Psi]^2}{\mathrm{Var}[\Psi]} ~~~\text{and}~~~ c = \frac{\mathrm{Var}[\Psi]}{2\mathrm{E}[\Psi]} = \frac{k}{f}
-  \end{equation*}
-
-  Furthermore, Brown showed that $\mathrm{E}[\Psi]$ and $\mathrm{Var}[\Psi]$ can be calculated via a numerical integration:
-  \begin{equation*}
-    \mathrm{E}[\Psi] = 2k ~~~\text{and}~~~ \mathrm{Var}[\Psi] = 4k + 2\sum_{i<j} \mathrm{Cov}\left( -2\log(\bar{P}^i), -2\log(\bar{P}^j) \right)
-  \end{equation*}
-
-  Kost and McDermott~\cite{kost_combining_2002} further fit a third-order polynomial to approximate the covariance
-  \begin{equation}
-    \mathrm{Cov}\left( -2\log(\bar{P}^i), -2\log(\bar{P}^j) \right) \approx 3.263 \rho_{ij} + 0.710 \rho_{ij}^2 + 0.027 \rho_{ij}^3
-    \label{eqn:suppl_covariance-pvalues}
-  \end{equation}
-  where, $\rho_{ij}$ is the correlation between method $i$ and method $j$
-
-  The final combined p-value~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016} is then given by:
-  \begin{equation}
-    \begin{aligned*}
-        & \hat{P}_j = 1.0 - \Phi_{2f}\left( \psi / c \right) \\
-        \text{where},~ &\psi = -2 \sum_{i=1}^k \log(\bar{P}^i_j) ~~~\text{and}~~~ \Phi_{2f} = \mathrm{CDF}\left( \chi^2_{2f} \right)
-    \end{aligned*}
-    \label{eqn:suppl_pvalue-combined}
-  \end{equation}
-
-  The p-value merging and consensus method in \ac{micone} (refer Methods) uses Equation~\ref{eqn:suppl_covariance-pvalues} to estimate the covariance of the p-values and Equation~\ref{eqn:suppl_pvalue-combined} to merge the p-values (obtained from bootstrapping) from the different correlation methods.
-  Note that we do not use Pearson and Spearman methods in the p-value merging step and these algorithms are only used for demonstration and comparison.
-  The combined p-values are used to threshold for significance the correlation-based networks during the consensus network step.
-
-  % TODO: Should we mention the MIND database here?
-  \subsection*{The JSON network format and network exports}
-
-    The default format \ac{micone} uses for storing the network files is the JSON (JavaScript Object Notation) format.
-    The custom JSON schema we have designed is able to store all network related information pertaining to nodes, links and the metadata related to the links and datasets.
-    Additionally, \ac{micone} also supports exporting of networks into a variety of other formats such as edge lists, .gml and cytoscape formats.
-    Since, we make use of \texttt{networkx}~\cite{hagbergExploringNetworkStructure2008} for the export functionality, networks can be exported to all formats supported by the package.
-    However, not all the corresponding metadata will be exported appropriately, since most formats do not support this additional metadata.
-    The details of the format and information about importing/exporting it and other network formats can be found in the \ac{micone} documentation.
-
-  \subsection*{Supplementary discussion}
-
-  Additionally, it is worth pointing out some additional more specific conclusions stemming from the individual steps of our analysis.
-  The different denoising/clustering methods differ mostly in their identification of sequences that are in low abundances.
-  Hence, they do not have much of an impact on the inferred co-occurrence networks when the sequences of low abundance are removed (Figure~\ref{fig:figure_s1}).
-  Comparison of inferred and expected reference sequences and their abundances in mock community datasets has allowed us to identify \ac{dada2} as the method which best recapitulates the expected sequence composition.
-  For the chimera checking module, we suggest using the remove bimera method since it was developed in conjunction with \ac{dada2} and its performance does not significantly differ from uchime-denovo.
-  For the current work we have decided to focus on the tools most widely used at the time of the analysis.
-  Some tools which were not as widely used (e.g.
-dbOTU3~\cite{Olesen2017}) as well as older popular methods like mothur~\cite{Schloss2009} have not been included in the study, but could be added into the pipelines in future updated analyses.
-
-  The choice of taxonomy database was found to be the most important factor in the inference of microbial co-occurrence networks, contributing $65.4\%$ of the total variance.
-  The frequent changes in the taxonomy nomenclature coupled with the frequency of updates to the various 16S reference databases create inherent differences \cite{Balvociute2017} in taxonomy hierarchies in these databases.
-  Our analysis revealed that no particular reference database performs better than the others across the different mock dataset benchmarks.
-  The default reference database in the pipeline is the \ac{gg} reference database along with the ``Naive Bayes'' classifier as the query tool.
-  The reason for our choice stems from the popularity of the \ac{gg} database~\cite{parkEvaluation16SRRNA2018} in taxonomic studies, which would enable easy comparison across datasets.
-  However, for newer studies we recommend using SILVA database because of its size and taxonomic comprehensiveness~\cite{iiRESCRIPtReproducibleSequence2021} and since \ac{gg} has not been updated since 2013.
-  Additionally, a particular database might be more appropriate than the rest based on specific requirements.
-  For example, in order to generate a dataset that is compatible with the \ac{mind} platform~\cite{huResourceComparisonIntegration2022} \ac{ncbi} is the most appropriate choice as it guarantees compatibility of taxonomic hierarchy and therefore comparability with other datasets.
-  Furthermore, we also enable users to use custom databases~\cite{Ritari2015,iiRESCRIPtReproducibleSequence2021} with the BLAST and Naive Bayes classifiers that are incorporated into the pipeline (from \ac{qiime2}).
-  We suggest that that choice of the database should be made based on possible reported or inferred biases in the representation of given biomes in a specific databases~\cite{Balvociute2017,iiRESCRIPtReproducibleSequence2021}, as choosing taxon-specific databases have also been observed to compromise classification~\cite{rmarcelinoUseTaxonspecificReference2020}.
-
-  The \ac{op} step of the pipeline is second in its contribution to total network variance.
-  This can be attributed to the large number of nodes that are added to the final networks when the filtering is turned off.
-  Additionally, a very large number of nodes also decreases the accuracy of the network inference algorithms for the same sample size~\cite{peschelNetCoMiNetworkConstruction2020} and increases the computational complexity~\cite{tackmannRapidInferenceDirect2019}.
-  We observe that filtering out taxa that are present in low abundances in all samples increases the proportion of taxa in common between taxonomy tables generated using different reference databases (Figure~\ref{fig:figure_s5}), providing another reason for filtering.
-  We also observe that the reduction in the number of taxa leads to better agreement in the networks inferred through different methods (Figure~\ref{fig:figure_s1}).
-  Moreover, filtering is necessary in order to increase the power in tests of significance when the number of taxa is much greater than the number of samples.

### 2022-09-27 scripts/figure4/figure4_plot.R

@@ -36,6 +36,8 @@ output_file <- paste0(output_folder, "figure4.pdf")
 output_file_a <- paste0(output_folder, "figure4a.pdf")
 output_file_bc <- paste0(output_folder, "figure4bc.pdf")
 
+corr_thres <- 0.3
+direct_thres <- 0.01
 
 #########################################
 # a
@@ -55,6 +57,7 @@ plot_network <- function(network_file, combined_layout, interaction_threshold, t
         activate(edges) %>%
         filter(abs(weight) > interaction_threshold) %>%
         mutate(color = get_edgecolor(weight)) %>%
+        mutate(abs_weight = abs(weight)) %>%
         activate(nodes) %>%
         mutate(isolated = node_is_isolated()) %>%
         filter(isolated == FALSE)
@@ -69,7 +72,7 @@ plot_network <- function(network_file, combined_layout, interaction_threshold, t
         geom_node_point() +
         scale_edge_color_manual(values = c(negative = "#d95f02", positive = "#1b9e77")) +
         scale_edge_linetype_manual(values = c(negative = "solid", positive = "solid")) +
-        scale_edge_alpha_manual(values = c(negative = 0.6, positive = 0.3)) +
+        scale_edge_alpha_manual(values = c(negative = 0.5, positive = 0.5)) +
         labs(title = title) +
         coord_fixed() +
         theme_bw() +
@@ -92,17 +95,17 @@ combined_layout <- create_layout(graph = combined_graph, layout = "linear", circ
 
 palette <- brewer.pal(n = 10, name = "Dark2")
 
-flashweave_plot <- plot_network(flashweave_gml, combined_layout, 0.01, "flashweave", palette[[1]])
-# mldm_plot <- plot_network(mldm_gml, combined_layout, 0.01, "mldm", palette[[2]])
-spieceasi_plot <- plot_network(spieceasi_gml, combined_layout, 0.01, "spieceasi", palette[[2]])
-cozine_plot <- plot_network(cozine_gml, combined_layout, 0.01, "cozine", palette[[3]])
[…]

### 2022-09-27 sections/discussion.tex
Fix: Finalize figures for resubmission

@@
  We find that the scaled-sum method (along with p-value merging for correlation-based methods) performs the best on synthetic datasets, and is therefore chosen as default for the \ac{ni} step of the pipeline.
-Notably, the consensus network displays a higher number of true positives, while also decreasing the overall variability across the whole pipeline (Figure~\ref{fig:figure7}B).
-The concise list of robust associations revealed by the consensus algorithm represents a valuable set for experimental validation follow-up.
+ Notably, the consensus network displays a higher precision, and returns a concise list of robust associations which represent a valuable set for experimental validation follow-up.
 

### 2022-09-27 sections/methods.tex
Fix: Finalize figures for resubmission

--- 2022-09-14 sections/methods.tex
@@
   The Naive Bayes classifier and the \ac{ncbi} blast used as the query tools in this study were from the \ac{qiime2} package and the parameters used were the defaults of the package.
+  The consensus algorithm used is the default method used by the classifiers in \ac{qiime2}.
   During the assignment, a representative sequence might be assigned an "unknown" Genus for one of two reasons: the first is if the taxonomy identifier associated with the sequence in the database did not contain a given Genus; the second, more likely reason, is that the database contains multiple sequences that are very similar to the query (representative) sequence and the consensus algorithm (from \ac{qiime2}) is unable to assign one particular Genus at the required confidence.
-  The consensus algorithm used is the default method used by the classifiers in \ac{qiime2}.
   The assignment between SILVA and the other two databases were originally significantly different ($40\%$ mismatch) even at the Phylum level.
@@
   The interaction matrix was used as the source of expected (true) interactions and the associations predicted using \ac{micone} were the source of predicted interactions.
-  Finally, for each dataset we evaluated the precision and sensitivity of the associations predicted by the individual network inference methods as well as the consensus (Figure \ref{fig:figure5}).
+  Finally, for each dataset we evaluated the precision of the associations predicted by the individual network inference methods as well as the consensus (Figure \ref{fig:figure5}).
 
@@
 
-  In Figure~\ref{fig:figure7}B we use the L1 distance metric to measure to the distance between the network inferred using different tools at the same step of the workflow.
-  The L1 distance between networks $x$ and $y$ is calculated using the following formula:
-  $$\frac{|A_{x} - A_{y}|}{2}$$
-  where, $A$ is the adjacency matrix of the network.
-
   \subsection*{Code and Data Availability}

### 2022-09-27 sections/results.tex
Fix: Finalize figures for resubmission

@@
   This counts table is used as the input to the \ac{micone} pipeline to generate predicted associations.
-  The interaction network used to generate the counts table is used as the source of true interactions to calculate the precision (Figure \ref{fig:figure5}) and sensitivity (Figure \ref{fig:figure_s6} and \ref{fig:figure_s7}) for each network inference algorithm.
+  The interaction network used to generate the counts table is used as the source of true interactions to calculate the precision (Figure \ref{fig:figure5}) and sensitivity (Figure \ref{fig:figure_s6} and Figure \ref{fig:figure_s7}) for each network inference algorithm.
   As shown in Figure \ref{fig:figure5} the consensus algorithm, especially the scaled-sum method, captures true associations with high precision (through the removal of edges that are either not present in most of the inference methods or whose association strength is low across methods).
-  Overall, the scaled-sum method for $p=1.0$ performs the best (precision = $1.000$ for both NorTA and seqtime). 
-  The scaled-sum method for $p=0.667$ (default option in the pipeline) has the second-highest average precision ($0.985$ with NorTA; $0.820$ with seqtime), without displaying significant reduction in sensitivity (Figure \ref{fig:figure_s6} and Figure~\ref{fig:figure_s7}).
+  Overall, the scaled-sum method for $p=1.000$ performs the best (precision = $1.000$ for both NorTA and seqtime).
+  The scaled-sum method for $p=0.333$ (default option in the pipeline) shows a high precision ($0.956$ with NorTA; $0.688$ with seqtime), without displaying significant reduction in sensitivity (Figure~\ref{fig:figure_s6} and Figure~\ref{fig:figure_s7}).
+  However, if higher precision is required $p>0.5$ can be considered.
   Therefore, the consensus networks provide the means to obtain a short list of associations that would have a high probability of being present in the real association network.
@@
 
-  [REVISIT THIS PART]
-  Figure~\ref{fig:figure7} shows the comparison of this default network against networks generated by altering one of the steps of the pipeline relative to the default.
-  This result, both visually (Figure~\ref{fig:figure7}A), and quantitatively (Figure~\ref{fig:figure7}B)  suggests that the most significant changes occur when the \ac{ta} step and the \ac{op} changed.
-  Notably, however despite the overall big effect of the TA database on the whole pipeline, the edges between the common nodes (with the default network) are fairly conserved, suggesting that the pipeline, with the consensus network at the end, provides a robust avenue for inferring associations.
-
-  Figure~\ref{fig:figure7}C shows the co-occurrence networks inferred for the healthy subjects (control) and autistic subjects in the fecal microbiome transplant study~\cite{Kang2017} (using the default tools and parameters from Table~\ref{tab:micone_tools}).
+  Figure~\ref{fig:figure7} shows the co-occurrence networks inferred for the healthy subjects (control) and autistic subjects in the fecal microbiome transplant study~\cite{Kang2017} (using the default tools and parameters from Table~\ref{tab:micone_tools}).
   This figure demonstrates a typical use case of comparative analysis of networks using the \ac{micone} pipeline.
   As a consequence of using the consensus network algorithm, the final co-occurrence networks are sparse and can be visually compared and examined.
-  We observe 10 unique links in the network for control samples and 3 unique links in the network for autistic subjects.
+  % TODO: Statistical test for differences + Refer to original paper and create conclusions
+  We observed 20 unique links in the network for control samples, 12 unique links in the network for autistic subjects and 7 edges in common between both networks.
   These unique connections can serve as potential starting points for literature surveys and further experimental confirmations.
+
+  Figure~\ref{fig:figure_s7} shows a sensitivity analysis in which we compared the default network against networks generated by altering one of the steps of the pipeline relative to the default.
+  This result, both visually (Figure~\ref{fig:figure_s7}A), and quantitatively (Figure~\ref{fig:figure_s7}B)  suggests that the most significant changes occur when the \ac{op} or \ac{ta} steps are changed from the default value.

### 2022-09-27 sections/supplementary.tex
Fix: Finalize figures for resubmission

@@
   \input{sections/supplementary_table.tex}
-
-  \subsection*{Figures}
-
-    \begin{figure}[H]
-      \centering
-      \includegraphics[width=1.0\linewidth]{figure_s1.pdf}
-    \end{figure}
-    \begin{figure}[H]
-      \centering
-        \caption{
-          \textbf{The networks which were filtered during the \ac{op} step are less variable}
-          All combinations of inferred networks are shown as points on a PCA plot and shows the variation in the inferred networks in the first two principal axes (similar to Figure~\ref{fig:figure6}B).
-          Each point on the PCA plot represents a network inferred using different combinations of tools and parameters that are available in the \ac{micone} pipeline.
-          The points are colored by the tools or parameters used in the \ac{dc} step (A) and the \ac{op} step (B).
-          The \ac{dc} step does not seem to have any correlation with the similarity of networks on the PCA plot, but all the networks with filtering at the \ac{op} step are similar to each other.
-          This further confirms that the variability in the networks decreases upon filtering out the taxonomic entities at low abundance.
-        }
-      \label{fig:figure_s1}
-    \end{figure}
-    \FloatBarrier
-    \newpage
-
-    % TODO: Verify this about t-SNE
-    \begin{figure}[H]
-      \centering
-      \includegraphics[width=1.0\linewidth]{figure_s2.pdf}
-    \end{figure}
-    \begin{figure}[H]
-      \centering
-        \caption{
-          \textbf{The t-SNE plot of all the inferred networks clusters the networks based on the taxonomy reference database used}.
-          Each point on the t-SNE plot represents a network inferred using different combinations of tools and parameters that are available in the \ac{micone} pipeline.
-          The points are colored by the tools and parameters used in \ac{dc} step (A), \ac{ta} step (B), \ac{op} step (C) and \ac{ni} step (D).
-          The separation of the points based on taxonomy reference database shows that the points might cluster based on reference database in high-dimensional space.
-        }
-      \label{fig:figure_s2}
-    \end{figure}
-    \FloatBarrier
-    \newpage
-
-    \begin{figure}[H]
-      \centering
-      \includegraphics[width=1.0\linewidth]{figure_s3.pdf}
-    \end{figure}
-    \begin{figure}[H]
-      \centering
-        \caption{
-          \textbf{The UniFrac distance between the 1000 most abundant representative sequences is higher than that when all sequences are considered}.
-          Each value is the average UniFrac distance between the reference sequences generated by the various methods in the \ac{dc} step (similar to Figure~\ref{fig:figure2}).
-          There is an increase in both weighted and unweighted UniFrac distances compared to when all the representative sequences are considered.
-          This shows that the 1000 most abundant representative sequences generated by the methods are not as similar to each other.
-          And since the weighted UniFrac is much smaller than the unweighted UniFrac distance, we can conclude that those reference sequences that are present in the middle of the abundance distribution are dissimilar.
-        }
-      \label{fig:figure_s3}
-    \end{figure}
-    \FloatBarrier
-    \newpage
-
-    \begin{figure}[H]
-      \centering
-      \includegraphics[width=1.0\linewidth]{figure_s4.pdf}
-    \end{figure}
-    \begin{figure}[H]
-      \centering
-        \caption{
-          \textbf{The weighted and unweighted UniFrac distances between representative sequences generated using remove bimera and uchime for each denoising method are low.}
-          The two chimera checking methods, uchime and remove bimera, produce similar outputs.
-          With the exception of de novo and open reference under the unweighted UniFrac metric, all the other methods have low dissimilarity.
-          This is especially true for the \ac{dada2} and Deblur methods which are the recommended denoising methods in the \ac{micone} pipeline.
-          Therefore, remove bimera is recommended as the default chimera method if one is using \ac{dada2} and uchime-denovo when one is using Deblur, since these methods were developed for these respective algorithms (\ac{qiime2} uses uchime-denovo in the Deblur workflow).
-        }
-      \label{fig:figure_s4}
-    \end{figure}
-    \FloatBarrier
-    \newpage
-
-    \begin{figure}[H]
-      \centering
-      \includegraphics[width=1.0\linewidth]{figure_s5.pdf}
-    \end{figure}
-    \begin{figure}[H]
-      \centering
-        \caption{
-          \textbf{The pairwise comparison of assignments generated using different databases for all representative sequences has a higher proportion of mismatches}.
-          The comparison made here is similar to Figure~\ref{fig:figure3}B, but instead of the top 100 taxonomic entities (by abundance), all the assignments from one database are matched with those from the other two databases.
-          Higher percentage of mismatches implies that the matching of the taxonomies in the more abundant sequences (top 100) are more consistent.
-        }
-      \label{fig:figure_s5}
-    \end{figure}
-    \FloatBarrier
-    \newpage
-
-  \begin{figure}[H]
-    \centering
-    \includegraphics[width=1.0\linewidth]{figure_s6.pdf}
-  \end{figure}
-  \begin{figure}[H]
-    \centering
-      \caption{
-        \textbf{The precision and sensitivity of the inferred networks on the NorTA synthetic interaction data when Pearson and Spearman are included in the consensus calculation}.
-        The ``I'' prefix indicates the individual network inference methods and the ``X'' prefix denotes the different consensus based methods.
-        The different consensus based methods used are: pvalue merging (PM), scaled-sum (SS) and simple voting (SV) method.
-        With the addition of Pearson and Spearman in the consensus calculation we observe a decrease in the average precision scores of both the consensus algorithms (compared to Figure~\ref{fig:figure5}.
-        The overall best precision was still consistently obtained by the scaled-sum method (0.985, 0.963 and 1.000), but the performance was not stable to an increase in the parameter value.
-        The simple voting method when using presence of edges in all inferred networks as a requirement (parameter value 1.000), again outperforms \ac{spieceasi} on average precision (0.961).
-      }
-    \label{fig:figure_s6}
-  \end{figure}
-  \FloatBarrier
-  \newpage
-
-  \begin{figure}[H]
-    \centering
-    \includegraphics[width=1.0\linewidth]{figure_s7.pdf}
-  \end{figure}
-  \begin{figure}[H]
-    \centering
-      \caption{
-        \textbf{The precision and sensitivity of the inferred networks on the seqtime synthetic interaction data.}
-        The ``I'' prefix indicates the individual network inference methods and the ``X'' prefix denotes the different consensus based methods.
-        The different consensus based methods used are: pvalue merging (PM), scaled-sum (SS) and simple voting (SV) method.
-        Pearson and Spearman methods are not used in the calculation of the consensus.
-        Among all the independent network inference methods, \ac{spieceasi} has the best average precision (0.624), but the overall best precision was consistently obtained by the scaled-sum method (0.688, 0.820 and 1.000).
-        The simple voting method when using presence of edges in all inferred networks as a requirement (parameter value 1.000), also outperforms \ac{spieceasi} on average precision (0.692).
-        These results show that the scaled-sum method is not only much better suited for inferring robust and accurate interactions from count data regardless of the distributions and topologies.
-        But, it is also capable of accurately extracting real interactions (gLV interactions) from count data.
-      }
-    \label{fig:figure_s7}
-  \end{figure}
-  \FloatBarrier
-  \newpage
 

### 2022-09-27 sections/tables_figures.tex
Fix: Figure 7 filtering thresholds
Fix: Finalize figures for resubmission

@@
                                           & & propr v2.1.2 & \cite{quinnProprRpackageIdentifying2017} \\ \cline{2-4}
-                                          & \multirow{2}{*}{Consensus algorithm} & \rowcolor{lightgray} scaled-sum & 0.667 \\
+                                          & \multirow{2}{*}{Consensus algorithm} & \rowcolor{lightgray} scaled-sum & 0.333 \\
                                           & & simple voting & 1.000 \\
@@
 
-  \FloatBarrier
-  \newpage
-
-  \FloatBarrier
-  \newpage
+  % \FloatBarrier
+  % \newpage
   % \begin{figure}[H]
@@
 
-  \FloatBarrier
-  \newpage
+  % \FloatBarrier
+  % \newpage
   % \begin{figure}[H]
@@
       A significant portion of the representative sequences are assigned to an ``unknown'' Genus in two of three databases (\ac{gg} and \ac{ncbi}).
+      The number of assigned genera for each database are displayed at the top of each column.
       \textbf{(B)} The number of representative sequences assigned to the same taxonomic label when using different reference databases, shown for the top 100 sequences.
@@
 
-  % TODO: Move caption to Figure S6
-  % Give reason for using those 4 individual methods
-  \FloatBarrier
-  \newpage
+  % \FloatBarrier
+  % \newpage
   % \begin{figure}[H]
@@
     \caption{
-      \textbf{The precision and sensitivity of the inferred networks on the NorTA synthetic interaction data.}
-      The ``I'' prefix indicates the individual network inference methods and the ``X'' prefix denotes the consensus based methods that are a part of \ac{micone}.
-      The different consensus based methods used are: pvalue merging (PM), scaled-sum (SS) and simple voting (SV) method.
-      Pearson and Spearman methods are not used in the calculation of the consensus.
-      Among all the independent network inference methods, \ac{spieceasi} has the best average precision (0.944), but the overall best precision was consistently obtained by the scaled-sum method (0.956, 0.985 and 1.000).
-      The simple voting method when using presence of edges in all inferred networks as a requirement (parameter value 1.000), also outperforms \ac{spieceasi} on average precision (0.969).
-      Although \ac{spieceasi} has a higher performance on sensitivity, if the goal of network inference is to obtain the list of associations that have a high probability of existing in the real microbial community, then the consensus methods perform better.
-      Therefore, we recommend the scaled-sum consensus method as the default tool for network inference.
+      \textbf{The scaled-sum consensus method returns associations with high precision when applied to the synthetic benchmark datasets}.
+      Each point on the box plot shows the precision of the co-occurrence network generated by the individual and consensus methods using the ``NorTA'' and ``seqtime'' methods (refer to Methods).
+      The independent algorithms chosen for the comparison are the two best-performing correlation-based (propr, \ac{sparcc}) and direct association based (\ac{spieceasi}, FlashWeave) methods.
+      A weight threshold of 0.1 and a p-value threshold of 0.05 was applied to each network before the calculation of precision.
+      The different consensus based methods used are the scaled-sum (SS) and simple voting (SV) method.
+      The overall best precision was consistently obtained by the scaled-sum method for $p \geq 0.333$ on both datasets.
+      Among all the independent network inference methods, \ac{spieceasi} has the best average precision.
+      The simple voting method when using presence of edges in all inferred networks as a requirement ($p = 1.000$), also outperforms \ac{spieceasi} on average precision.
+      Therefore, we recommend the scaled-sum consensus method with $p = 0.333$ as the default tool for network inference, as this option provides a good balance of precision and sensitivity.
     }
@@
 
-  \FloatBarrier
-  \newpage
-  % TODO: Add equations for the linear model to the left of the pie chart
+  % \FloatBarrier
+  % \newpage
   % \begin{figure}[H]
@@
       The data used for the analysis were the healthy stool samples from the fecal microbiome transplant dataset~\cite{Kang2017}.
+      A weight threshold of 0.1 and a p-value threshold of 0.05 was applied to each network before the analysis.
       The taxonomy database contributes most to the variance between the networks (65.4\%) followed by the filtering of the counts matrix (26.8\%) in the OP step.
@@
       Each point on the PCA plot represents a network inferred using different combinations of tools and parameters that are available in the \ac{micone} pipeline.
-      The color of the points corresponds to the reference taxonomy database used to reconstruct those networks (left), and the network inference algorithm used to infer the networks (right).
+      The color of the points corresponds to the tools used at each step of the pipeline (DC, TA, OP and NI).
       The points on the PCA plot can be grouped based on the TA step, but at the NI step only some methods create high variability in the inferred networks.
+      The \ac{dc} step does not seem to have any correlation with the similarity of networks on the PCA plot, but all the networks with filtering at the \ac{op} step are similar to each other.
+      This further confirms that the variability in the networks decreases upon filtering out the taxonomic entities at low abundance.
     }
@@
 
-  \FloatBarrier
-  \newpage
-  % TODO: Verify the validity of Figure 7A and 7B again
-  % Are they both using consensus?
+  % \FloatBarrier
+  % \newpage
   % \begin{figure}[H]
@@
     \caption{
-      \textbf{The consensus algorithm drastically reduces variability in the networks due to upstream data processing methods and increases robustness of inferred networks}
-      \textbf{(A)} The network constructed using the default pipeline parameters (DC=\ac{dada2}, CC=remove bimera, TA=\ac{gg}, OP=Filter(on), NI=scaled-sum consensus) is compared with networks generated when one of the steps use a different tool.
-      The common connections (common with the default network) are in black, connections unique to the network are colored purple and connections in the default network, but not present in the current network are gray.
-      Changing the NI, TA and OP steps leads to the creation of the most number of new nodes and edges.
-      \textbf{(B)} The boxplot showing the L1 distance between the default network and the networks generated by changing one step of the default pipeline.
-      The high value of L1 distances for the network inference step as compared to the other steps demonstrates how the consensus method is able to reduce the variability in the networks in the other steps of the pipeline.
-      This also shows that even though many new nodes were added in (A) when the TA and OP steps were changed, the L1 distances remain fairly small.
-      \textbf{(C)} The networks generated for the control (left) and autistic (right) samples in the fecal microbiome transplant dataset.
-      These networks were generated using the default tools and parameters recommended by the \ac{micone} pipeline.
-      The data used in this analysis were the healthy stool samples from the fecal microbiome transplant dataset.
+      \textbf{Comparison of networks generated from autistic and control samples using the \ac{micone} pipeline}.
+      The networks for the control (left) and autistic (right) samples in the fecal microbiome transplant dataset were generated using the default tools and parameters recommended by the \ac{micone} pipeline.
+      There are 20 unique links in the network for control samples, 12 unique links in the network for autistic subjects and 7 edges in common between both networks.
     }
@@
   \end{figure}
+
+  \subsection*{Supplementary figures}
+
+    % TODO: Verify this about t-SNE
+    % \begin{figure}[H]
+    %   \centering
+    %   \includegraphics[width=1.0\linewidth]{figure_s1.pdf}
+    % \end{figure}
+    \begin{figure}[H]
+      \centering
+        \caption{
+          \textbf{The t-SNE plot of all the inferred networks clusters the networks based on the taxonomy reference database used}.
+          Each point on the t-SNE plot represents a network inferred using different combinations of tools and parameters that are available in the \ac{micone} pipeline.
+          The points are colored by the tools and parameters used in \ac{dc} step (A), \ac{ta} step (B), \ac{op} step (C) and \ac{ni} step (D).
+          The separation of the points based on taxonomy reference database shows that the points might cluster based on reference database in high-dimensional space.
+        }
+      \label{fig:figure_s1}
+    \end{figure}
+    % \FloatBarrier
+    % \newpage
+
+    % \begin{figure}[H]
+    %   \centering
+    %   \includegraphics[width=1.0\linewidth]{figure_s2.pdf}
+    % \end{figure}
+    \begin{figure}[H]
+      \centering
+        \caption{
+          \textbf{The UniFrac distance between the 1000 most abundant representative sequences is higher than that when all sequences are considered}.
+          Each value is the average UniFrac distance between the reference sequences generated by the various methods in the \ac{dc} step (similar to Figure~\ref{fig:figure2}).
+          There is an increase in both weighted and unweighted UniFrac distances compared to when all the representative sequences are considered.
+          This shows that the 1000 most abundant representative sequences generated by the methods are not as similar to each other.
+          And since the weighted UniFrac is much smaller than the unweighted UniFrac distance, we can conclude that those reference sequences that are present in the middle of the abundance distribution are dissimilar.
+        }
+      \label{fig:figure_s2}
+    \end{figure}
+    % \FloatBarrier
+    % \newpage
+
+    % \begin{figure}[H]
+    %   \centering
+    %   \includegraphics[width=1.0\linewidth]{figure_s3.pdf}
+    % \end{figure}
+    \begin{figure}[H]
+      \centering
+        \caption{
+          \textbf{The weighted and unweighted UniFrac distances between representative sequences generated using remove bimera and uchime for each denoising method are low.}
+          The two chimera checking methods, uchime and remove bimera, produce similar outputs.
+          With the exception of de novo and open reference under the unweighted UniFrac metric, all the other methods have low dissimilarity.
+          This is especially true for the \ac{dada2} and Deblur methods which are the recommended denoising methods in the \ac{micone} pipeline.
+          Therefore, remove bimera is recommended as the default chimera method if one is using \ac{dada2} and uchime-denovo when one is using Deblur, since these methods were developed for these respective algorithms (\ac{qiime2} uses uchime-denovo in the Deblur workflow).
+        }
+      \label{fig:figure_s3}
+    \end{figure}
+    % \FloatBarrier
+    % \newpage
+
+    % \begin{figure}[H]
+    %   \centering
+    %   \includegraphics[width=1.0\linewidth]{figure_s4.pdf}
+    % \end{figure}
+    \begin{figure}[H]
+      \centering
+        \caption{
+          \textbf{The pairwise comparison of assignments generated using different databases for all representative sequences has a higher proportion of mismatches}.
+          The comparison made here is similar to Figure~\ref{fig:figure3}B, but instead of the top 100 taxonomic entities (by abundance), all the assignments from one database are matched with those from the other two databases.
+          Higher percentage of mismatches implies that the matching of the taxonomies in the more abundant sequences (top 100) are more consistent.
+        }
+      \label{fig:figure_s4}
+    \end{figure}
+    % \FloatBarrier
+    % \newpage
+
+  % \begin{figure}[H]
+  %   \centering
+  %   \includegraphics[width=1.0\linewidth]{figure_s5.pdf}
+  % \end{figure}
+  \begin{figure}[H]
+    \centering
+      \caption{
+        \textbf{The precision and sensitivity of the inferred networks on the ``NorTA'' synthetic interaction data.}
+        The different consensus based methods used are: scaled-sum (SS) and simple voting (SV) method.
+        Pearson and Spearman methods are not used in the calculation of the consensus.
+        Among all the independent network inference methods, \ac{spieceasi} has the best average precision (0.944), but the overall best precision was consistently obtained by the scaled-sum method (0.956, 0.985 and 1.000).
+        The simple voting method when using presence of edges in all inferred networks as a requirement ($p = 1.000$), also outperforms \ac{spieceasi} on average precision (0.969).
+        Although \ac{spieceasi} has a higher performance on sensitivity, if the goal of network inference is to obtain the list of associations that have a high probability of existing in the real microbial community, then the consensus methods perform better.
+      }
+    \label{fig:figure_s5}
+  \end{figure}
+  % \FloatBarrier
+  % \newpage
+
+  % \begin{figure}[H]
+  %   \centering
+  %   \includegraphics[width=1.0\linewidth]{figure_s6.pdf}
+  % \end{figure}
+  \begin{figure}[H]
+    \centering
+      \caption{
+        \textbf{The precision and sensitivity of the inferred networks on the ``seqtime'' synthetic interaction data.}
+        The different consensus based methods used are: scaled-sum (SS) and simple voting (SV) method.
+        Pearson and Spearman methods are not used in the calculation of the consensus.
+        Among all the independent network inference methods, \ac{spieceasi} has the best average precision (0.624), but the overall best precision was consistently obtained by the scaled-sum method (0.688, 0.820 and 1.000).
+        The simple voting method when using presence of edges in all inferred networks as a requirement ($p = 1.000$), also outperforms \ac{spieceasi} on average precision (0.692).
+        These results show that the scaled-sum method is not only much better suited for inferring robust and accurate interactions from count data regardless of the distributions and topologies, but it is also capable of accurately extracting real associations (gLV interactions) from count data.
+      }
+    \label{fig:figure_s6}
+  \end{figure}
+  % \FloatBarrier
+  % \newpage
+
+  % \begin{figure}[H]
+  %   \centering
+  %   \includegraphics[width=1.0\linewidth]{figure_s7.pdf}
+  % \end{figure}
+  \begin{figure}[H]
+    \centering
+    \caption{
+      \textbf{Sensitivity analysis of the default settings of the \ac{micone} pipeline}.
+      \textbf{(A)} The network constructed using the default pipeline parameters (DC=\ac{dada2}, CC=remove bimera, TA=\ac{gg}, OP=Filter(on), NI=scaled-sum consensus) is compared with networks generated when one of the steps use a different tool.
+      The layout is created by fixing the positions of all the nodes from all networks and then drawing only the relevant edges.
+      The edges colored green are positive associations and those in red are negative associations.
+      We observe that changing the TA and OP steps leads to the creation of the most number of unique edges.
+      \textbf{(B)} The dot plot showing the fraction of nodes (left) and edges (right) in common between the default network and the networks generated by changing one step of the default pipeline.
+      The low value of the common fraction for TA and OP steps shows that these steps induce the biggest changes nodes and edges.
+      The NI step is not shown in this analysis because the consensus method uses edges from the individual network inference methods and a comparison would be biased.
+    }
+    \label{fig:figure_s7}
+  \end{figure}
+  % \FloatBarrier
+  % \newpage

### 2022-09-27 scripts/figure2/figure2_plot.R

@@ -57,6 +57,7 @@ plot_heatmap <- function(data, title) {
             labels = c("Similar", 0.25, 0.5, 0.75, "Dissimilar")
         ) +
         ggtitle(title) +
+        labs(fill="UniFrac") +
         theme_pubr() +
         theme(
             plot.title = element_text(hjust = 0.5),
@@ -74,7 +75,7 @@ weighted <- read.csv(weighted_unifrac_csv, sep = ",", header = TRUE)
 weighted_tidy <- tidy_up_data_ab(weighted)
 
 weighted_plot <- plot_heatmap(weighted_tidy, "Weighted UniFrac")
-unweighted_plot <- plot_heatmap(unweighted_tidy, "UnWeighted UniFrac")
+unweighted_plot <- plot_heatmap(unweighted_tidy, "Unweighted UniFrac")
 
 heatmap_plot <- ggarrange(
     weighted_plot, unweighted_plot,
@@ -112,6 +113,7 @@ make_dot_plot <- function(unifrac_data, title) {
             size = 3,
             add = "jitter",
         ) +
+        labs(fill="Dataset") +
             theme_pubr() +
             theme(
                 plot.title = element_text(hjust = 0.5),
@@ -144,7 +146,7 @@ mock12_unweighted_tidy <- tidy_up_data_cd(mock12_unweighted, "mock12")
 mock16_unweighted_tidy <- tidy_up_data_cd(mock16_unweighted, "mock16")
 
 mock_unweighted_tidy <- rbind(mock4_unweighted_tidy, mock12_unweighted_tidy, mock16_unweighted_tidy)
-mock_unweighted_plot <- make_dot_plot(mock_unweighted_tidy, "UnWeighted UniFrac")
+mock_unweighted_plot <- make_dot_plot(mock_unweighted_tidy, "Unweighted UniFrac")
 
 
 final_dot_plot <- ggarrange(

### 2022-09-27 scripts/figure3/figure3_plot.R

@@ -67,6 +67,7 @@ make_alluvial_plot2 <- function(db_data, title) {
         scale_fill_manual(values = palette) +
         geom_text_repel(aes(label = Genus), stat = "stratum", size = 3.5, direction = "y", nudge_x = 0.5) +
         ggtitle(title) +
+        xlab("Database") +
         theme_pubr() +
         theme(
             text = element_text(size = 15),
@@ -79,7 +80,8 @@ gg_n_genus <- length(unique(gg$Genus))
 silva_n_genus <- length(unique(silva$Genus))
 ncbi_n_genus <- length(unique(ncbi$Genus))
 
-combined_plot <- make_alluvial_plot2(combined, paste("NaiveBayes(GG)=", gg_n_genus, ", NaiveBayes(SILVA)=", silva_n_genus, ", BLAST(NCBI)=", ncbi_n_genus))
+# TODO: Fix this
+combined_plot <- make_alluvial_plot2(combined, paste("NaiveBayes(GG) =", gg_n_genus, ", NaiveBayes(SILVA) =", silva_n_genus, ", BLAST(NCBI) =", ncbi_n_genus))
 
 final_plot_a <- ggarrange(combined_plot, nrow = 1, ncol = 1, common.legend = FALSE)
 annotate_figure(final_plot_a, fig.lab = "A", fig.lab.pos = "top.left", fig.lab.size = 20)
@@ -114,6 +116,7 @@ make_bar_plot <- function(data, title) {
         palette = c("#00AFBB", "#FC4E07"),
         # palette="Paired"
     ) +
+        labs(fill="Assignments") +
         theme(
             text = element_text(size = 15),
             # plot.title = element_text(size = 10),
@@ -164,6 +167,7 @@ make_dot_plot <- function(braycurtis_data) {
             size = 3,
             add = "jitter",
         ) +
+        labs(fill="Dataset") +
             theme(
                 plot.title = element_text(hjust = 0.5),
                 text = element_text(size = 15),

### 2022-09-27 scripts/figure4/figure4_plot.R

@@ -56,7 +56,7 @@ plot_network <- function(network_file, combined_layout, interaction_threshold, t
     graph <- graph_raw %>%
         activate(edges) %>%
         filter(abs(weight) > interaction_threshold) %>%
-        mutate(color = get_edgecolor(weight)) %>%
+        mutate(interaction = get_edgecolor(weight)) %>%
         mutate(abs_weight = abs(weight)) %>%
         activate(nodes) %>%
         mutate(isolated = node_is_isolated()) %>%
@@ -67,7 +67,7 @@ plot_network <- function(network_file, combined_layout, interaction_threshold, t
     graph_layout$x <- combined_layout[match_inds, ]$x
     graph_layout$y <- combined_layout[match_inds, ]$y
     graph_plot <- ggraph(graph = graph, layout = "manual", x = graph_layout$x, y = graph_layout$y, circular = TRUE) +
-        geom_edge_arc(aes(color = color, edge_linetype = color, edge_alpha = color)) +
+        geom_edge_arc(aes(color = interaction, edge_linetype = interaction, edge_alpha = interaction)) +
         # geom_node_point(aes(color=factor(colorkey))) +
         geom_node_point() +
         scale_edge_color_manual(values = c(negative = "#d95f02", positive = "#1b9e77")) +

### 2022-09-27 scripts/figure5/figure5_plot.R

@@ -42,7 +42,7 @@ plot_boxplot <- function(data, title) {
         ) +
         xlab("Algorithm") +
         ylab("Precision") +
-        labs(title = title)
+        labs(title = title, fill = "NI method")
 }
 
 methods <- c("propr", "sparcc", "flashweave", "spieceasi", "SS[0.333]", "SS[0.667]", "SS[1.000]", "SV[0.333]", "SV[0.667]", "SV[1.000]")

### 2022-09-27 sections/abstract.tex
Change: Incorporate comments from Gabriel and add refs

@@
     \noindent
-    Microbes tend to organize into communities consisting of hundreds of species involved in complex interactions with each other.
+    Microbes commonly organize into communities consisting of hundreds of species involved in complex interactions with each other.
     16S ribosomal RNA (16S rRNA) amplicon profiling provides snapshots that reveal the phylogenies and abundance profiles of these microbial communities.
     These snapshots, when collected from multiple samples, have the potential to reveal which microbes co-occur, providing a glimpse into the network of associations in these communities.
-    The inference of networks from 16S data involves numerous steps, each requiring specific tools and parameter choices.
-    The extent to which the steps in this workflow affect the final network is still unclear.
+    However, the inference of networks from 16S data involves numerous steps, each requiring specific tools and parameter choices.
+    Moreover, the extent to which these steps affect the final network is still unclear.
     In this study, we perform a meticulous analysis of each step of a pipeline that can convert 16S sequencing data into a network of microbial associations.
@@
     The Microbial Co-occurrence Network Explorer or \acs{micone} (available at \href{https://github.com/segrelab/MiCoNE}{https://github.com/segrelab/MiCoNE}), follows these default tools and parameters and can help explore the outcome of these combinations of choices on the inferred networks.
-    We envisage that this pipeline could be used for integrating multiple data sets, and for generating comparative analyses and consensus networks that can help understand microbial community assembly in different biomes.
+    We envisage that this pipeline could be used for integrating multiple datasets, and for generating comparative analyses and consensus networks that can help understand microbial community assembly in different biomes.
   }
@@
 \keywords{Microbiome, 16S rRNA, Pipeline, Interaction, Denoising, Taxonomy, Network Inference, Correlations, Qiime2, Co-occurrence, Networks, Consensus algorithm}
+
+% \doublespacing
 

### 2022-09-27 sections/discussion.tex
Change: Incorporate comments from Gabriel and add refs

@@
   It allows for rapid, configurable, and reproducible inference of microbial networks and further for the formulation of hypotheses about the role of these interactions on community composition and stability.
-  These comparative analyses will require coupled network analysis and visualization tools (such as \ac{mind}~\cite{huResourceComparisonIntegration2022}) and will require more systematic access to datasets, shared in accordance with FAIR standards~\cite{pachecoFAIRRepresentationsMicrobial2022}.
+  These comparative analyses will require coupled network analysis and visualization tools (such as \ac{mind}~\cite{huResourceComparisonIntegration2022}) and need systematic access to datasets, shared in accordance with FAIR standards~\cite{pachecoFAIRRepresentationsMicrobial2022}.

### 2022-09-27 sections/introduction.tex
Change: Incorporate comments from Gabriel and add refs

@@
 
-  Microbial communities are ubiquitous and play an important role in marine and terrestrial environments, urban ecosystems, and human health \cite{Ghoul2016,Thompson2017} [MORE REFS].
-  These microbial communities, or microbiomes, often comprise several hundreds of different microbial strains interacting with each other and their environment, often through intricate metabolic and signaling relationships [REFS].
+Microbial communities are ubiquitous and play an important role in marine and terrestrial environments, urban ecosystems, and human health \cite{Zhang2015,Thompson2017,Ghoul2016,HumanMicrobiomeProjectConsortium2012,royo-llonchCompendium530Metagenomeassembled2021,tedersooFungalBiogeographyGlobal2014}.
+  These microbial communities, or microbiomes, often comprise several hundreds of different microbial strains interacting with each other and their environment, often through intricate metabolic and signaling relationships~\cite{lima-mendezDeterminantsCommunityStructure2015a,zelezniakMetabolicDependenciesDrive2015,coyteUnderstandingCompetitionCooperation2019,DSouza2018}.
   Understanding how these interconnections shape community structure and functionalities is a fundamental challenge in microbial ecology, and has applications in the study of microbial ecosystems across different biomes.
@@
   In particular, high-throughput sequencing, including metagenomic sequencing and sequencing of 16S rRNA gene amplicons (hereafter referred to as 16S data) of microbial communities, has the potential to help detect, identify and quantify a large portion of the constitutive microorganisms of a microbiome \cite{Jovel2016,Lloyd-Price2016}.
-  These advances have led to large-scale data collection efforts involving terrestrial (\acl{emp}) \cite{Thompson2017}, marine (Tara Oceans Project) \cite{Zhang2015} and human-associated microbiota (Human Microbiome Project) \cite{HumanMicrobiomeProjectConsortium2012}.
-[3 REFS for each]
+  These advances have led to large-scale data collection efforts involving terrestrial~\cite{Thompson2017,gilbertMeetingReportTerabase2010,tedersooFungalBiogeographyGlobal2014}, marine~\cite{Zhang2015,lima-mendezDeterminantsCommunityStructure2015a,royo-llonchCompendium530Metagenomeassembled2021} and human-associated microbiota~\cite{HumanMicrobiomeProjectConsortium2012,proctorIntegrativeHumanMicrobiome2019,Lloyd-Price2016}.
 
@@
  In particular, a powerful tool for translating microbiome composition data into knowledge is the construction of association (or co-occurrence, or correlation) networks, in which  microbial taxa are represented by nodes, and frequent co-occurrences (or negative correlations) across datasets are encoded in edges between nodes.
- While the relationship between directly measured interactions [REFS] and statistically inferred co-occurrence is still poorly understood \cite{Zuniga2017}, a significant amount of effort has gone into estimating correlations from large microbiome sequence datasets.
+ While the relationship between directly measured interactions~\cite{lubbeExometabolomicAnalysisCrossFeeding2017,Jian2020,Hsu2019} and statistically inferred co-occurrence is still poorly understood \cite{Zuniga2017,Rottjers2018}, a significant amount of effort has gone into estimating correlations from large microbiome sequence datasets~\cite{faustMicrobialInteractionsNetworks2012a,maEarthMicrobialCooccurrence2020a}.
 
  The importance of these networks is two-fold: first, they can serve as maps that help identify hubs of keystone species \cite{Menon2018,Rottjers2018}, and the community response to environmental perturbations or underlying host conditions \cite{Gilbert2016}; second, they can serve as a bridge towards building mechanistic models of microbial communities, greatly enhancing our capacity to understand and control them.
- For example, multiple studies have shown the importance of specific microbial  associations in the healthy microbiome \cite{Lloyd-Price2016} and the rise of dysbiosis \cite{Wang2017,Gilbert2016,Belizario2015}.
- In the context of terrestrial bio-geochemistry, co-occurrence networks were shown to help understand microbiome assembly \cite{Fierer2017}, and  the response of microbial communities to environmental perturbations \cite{Jiao2019}.
+ For example, multiple studies have shown the importance of specific microbial associations in the healthy microbiome \cite{Lloyd-Price2016,Wu2016,HumanMicrobiomeProjectConsortium2012} and the rise of dysbiosis \cite{Wang2017,Gilbert2016,Belizario2015}.
+ In the context of terrestrial biogeochemistry, co-occurrence networks were shown to help understand microbiome assembly \cite{Fierer2017}, and  the response of microbial communities to environmental perturbations \cite{Jiao2019}.
 

### 2022-09-27 sections/methods.tex
Change: Incorporate comments from Gabriel and add refs

@@
 
-  Note that we do not use Pearson and Spearman methods in the p-value merging step to determine the consensus network, these methods are only used for demonstration and comparison.
+  Note that we do not use Pearson and Spearman methods in the p-value merging step to determine the consensus network.
+  These methods are only used for demonstration and comparison.
   The combined p-values are used to threshold for significance right before the consensus algorithm is applied to the inferred networks.
@@
   The consensus algorithm was designed to increase the precision (number of true positives) at the end of the network inference step.
-  For this purpose, we developed two simple algorithms that combine the edges reported by the different network inference tools using a consensus based on a parameter $p$.
+  For this purpose, we developed two simple algorithms that combine the edges reported by the different network inference tools.
+  Both the algorithms make use of a user-defined parameter $p$ ($0 \leq p \leq 1$), in order to threshold the edges from the individual methods.
   The inputs to both the algorithms are the co-occurrence networks (association matrices) $\bar{N}^i$ (flattened version of $N^{i}$) generated by each method $i$ and the threshold parameter $p$.
@@
 
-  The simple voting method returns the union of the networks when $0 \leq p \leq \frac{1}{w}$ and will return the intersection when $(w - 1) \leq p \leq 1$.
-  In general, if $(n - 1) \leq p \leq \frac{n}{w}$, this algorithm will report an edge in the consensus network when at least $n$ network inference methods report this edge.
+  The simple voting method returns the union of the networks when $0 \leq p \leq \frac{1}{w}$ and will return the intersection when $\frac{(w - 1)}{w} \leq p \leq 1$.
+  In general, if $\frac{(n - 1)}{w} \leq p \leq \frac{n}{w}$, this algorithm will report an edge in the consensus network when at least $n$ network inference methods report this edge.
 

### 2022-09-27 sections/results.tex
Change: Incorporate comments from Gabriel and add refs

@@
 
-  Since no database performs better than others against mock datasets, the choice of which database to use could be driven by other reasons (see Discussion).
+  Since no database performs better than others against mock datasets, the choice of which database to use could be driven by other reasons (see Supplementary discussion).
   One reason to choose a particular database could be the frequency of updates and the potential for future growth.

### 2022-09-27 sections/supplementary.tex
Change: Incorporate comments from Gabriel and add refs

@@
 
-  Furthermore, Brown showed that $\mathrm{E}[\Psi]$ and $\mathrm{Var}[\Psi]$ can be calculated via a numerical integration:
+  We first calculate $E[\Psi]$ and $Var[\Psi]$ under the null hypothesis that the data are drawn from a multivariate Gaussian with some covariance matrix.
+  We then use these values to parametrize a $\chi^2$ distribution from which p-value corresponding to $\frac{\psi}{c}$ can be calculated.
+  Furthermore, Brown showed that $\mathrm{E}[\Psi]$ and $\mathrm{Var}[\Psi]$ can be calculated via a numerical approximation:
   \begin{equation*}
@@
 
-  Kost and McDermott~\cite{kost_combining_2002} further fit a third-order polynomial to approximate the covariance
+  The above formulation was improved by Kost and McDermott~\cite{kost_combining_2002} by further fitting a third-order polynomial to approximate the covariance
   \begin{equation}
@@
 
-  The final combined p-value~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016} is then given by:
+  Using $E[\Psi]$ and $Var[\Psi]$ one then fits a $\chi^2$ distribution with the parameters $c$ and $f$.
+  Note that since, in general, $f$ will not be an integer, this should be understood as a Gamma distribution with a shape parameter $f$, as mentioned by Brown~\cite{brown_400_1975}.
+  Using this, one calculates the test $\psi$ and computes the p-value from the CDF of the $\chi^2$ distribution, given in Equation~\ref{eqn:suppl_pvalue-combined}.
+  Therefore, the final combined p-value~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016} is then given by:
   \begin{equation}
@@
 
-  The p-value merging and consensus method in \ac{micone} (refer Methods) uses Equation~\ref{eqn:suppl_covariance-pvalues} to estimate the covariance of the p-values and Equation~\ref{eqn:suppl_pvalue-combined} to merge the p-values (obtained from bootstrapping) from the different correlation methods.
+  The p-value merging and consensus method in \ac{micone} (see Methods) uses Equation~\ref{eqn:suppl_covariance-pvalues} to estimate the covariance of the p-values and Equation~\ref{eqn:suppl_pvalue-combined} to merge the p-values (obtained from bootstrapping) from the different correlation methods.
   Note that we do not use Pearson and Spearman methods in the p-value merging step and these algorithms are only used for demonstration and comparison.

### 2022-09-27 sections/tables_figures.tex
Change: Incorporate comments from Gabriel and add refs

@@
       The final output of the pipeline is the consensus network generated from the inferred co-occurrence networks.
-      For details on each process and the data outputs, see Methods.
+      For details on each process and the outputs, see Methods.
     }
@@
     \caption{
-      \textbf{The representative sequences generated by the different denoising/clustering methods are very similar but differ in the sequences that are lower in abundance.}
+      \textbf{The representative sequences generated by the different DC methods differ in their identification of sequences that are low in abundance.}
       \textbf{(A)} The average weighted UniFrac distance between the representative sequences shows that the representative sequences and their compositions are fairly identical between the methods (with the exception of Deblur (DB)).
@@
       The number of \ac{otu}s or \ac{esv}s generated by the respective methods are provided in the parenthesis next to the names.
-      The data used for the analysis in \textbf{(A, B)} were all the samples (healthy and autistic) from the fecal microbiome transplant dataset.
+      The data used for the analysis in \textbf{(A, B)} were all the samples (healthy and autistic) from the fecal microbiome transplant (FMT) dataset~\cite{Kang2017}.
       \textbf{(C, D)} The distributions of the average weighted and unweighted UniFrac distance between the expected sequence profile and the predicted sequence profile in the mock datasets.
@@
     \caption{
-      \textbf{Taxonomic reference databases vary widely in terms of their taxonomy assignments beyond the Order level.}
+      \textbf{Taxonomic reference databases vary in terms of their taxonomy assignments beyond the Order level.}
       \textbf{(A)} The assignments of the top 50 representative sequences to their respective taxonomies using the three different reference databases.
@@
       The number of mismatches are fewer at higher taxonomic levels, but, even at the Order level there exists greater than 51\% of mismatches, demonstrating the poor agreement in taxonomic labels assigned by the different databases.
-      The data used for the analysis in \textbf{(A, B)} were all the samples (healthy and autistic) from the fecal microbiome transplant dataset.
+      The data used for the analysis in \textbf{(A, B)} were all the samples (healthy and autistic) from the FMT dataset.
       \textbf{(C)} The Bray-Curtis dissimilarity between the expected taxonomy profile and calculated taxonomy profile in the mock datasets shows that there is no singular best choice of database for every dataset, as all the databases have similar performance.
-      However, we still recommend the \ac{gg} database and the Naive Bayes classifier as the defaults for the \ac{micone} pipeline.
-      Due to the popularity of the \ac{gg} database, comparison against existing networks in literature will be easier since many published datasets make use of this database in their analyses.
+      The \ac{gg} database and the Naive Bayes classifier are chosen as the defaults for the TA step of \ac{micone}.
       The datasets used for the analysis in \textbf{(C)} were the mock datasets from mockrobiota.
@@
       The green links are positive associations and the orange links represent negative associations.
-      A threshold of 0.3 was set for the correlation-based methods (\ac{sparcc}, Propr, Spearman and Pearson) and a threshold of 0.01 was set for the other direct association methods.
+      A threshold of 0.3 was set for the correlation-based methods (sparcc, propr, spearman and pearson) and a threshold of 0.01 was set for the direct association methods (flashweave, spieceasi, cozine, harmonies, and spring).
       The correlation-based methods in general produce graph with higher edge-densities.
@@
       Whereas, \textbf{(C)} the edge overlap Upset plot shows that a very small fraction of these connections are actually shared (8 out of 202).
-      The data used in this analysis were the healthy stool samples from the fecal microbiome transplant dataset.
-      \ac{mldm} is not shown in the comparisons because the algorithm failed to converge for most of the network combinations.
+      The data used in this analysis were the healthy stool samples from the FMT dataset.
+      \ac{mldm} is not shown in the comparisons because the algorithm failed to converge for the particular network combination.
     }
@@
       \caption{
-      \textbf{The choice of reference database contributes to the most variance in the networks}.
-      \textbf{(A)} The percentage of variance in the networks contributed by the DC, CC (chimera checking), TA and OP steps of the pipeline (right) and the linear model used to calculate the variance (left), see the Methods section for details.
-      The data used for the analysis were the healthy stool samples from the fecal microbiome transplant dataset~\cite{Kang2017}.
+      \textbf{The choice of reference database contributes has the largest impact on network variance}.
+      \textbf{(A)} The percentage of variance in the networks contributed by the DC, CC (chimera checking), TA, OP and NI steps of the pipeline calculated using ANOVA on the linear model, see the Methods section for details.
+      The data used for the analysis were the healthy stool samples from the FMT dataset.
       A weight threshold of 0.1 and a p-value threshold of 0.05 was applied to each network before the analysis.
       The taxonomy database contributes most to the variance between the networks (65.4\%) followed by the filtering of the counts matrix (26.8\%) in the OP step.
-      The variation due to the DC and CC steps are comparatively not as significant.
+    The variation due to the DC and CC steps are much smaller in comparison (0.648\% and 0.003\% respectively).
       The fraction labeled as the residual is an artifact that arises when multiple steps are changed at the same time, however, this can be ignored as this fraction is negligible.
@@
       \textbf{Comparison of networks generated from autistic and control samples using the \ac{micone} pipeline}.
-      The networks for the control (left) and autistic (right) samples in the fecal microbiome transplant dataset were generated using the default tools and parameters recommended by the \ac{micone} pipeline.
+      The networks for the control (left) and autistic (right) samples in the FMT dataset were generated using the default tools and parameters recommended by the \ac{micone} pipeline.
       There are 20 unique links in the network for control samples, 12 unique links in the network for autistic subjects and 7 edges in common between both networks.
@@
 
-    % TODO: Verify this about t-SNE
     % \begin{figure}[H]
@@
           The points are colored by the tools and parameters used in \ac{dc} step (A), \ac{ta} step (B), \ac{op} step (C) and \ac{ni} step (D).
-          The separation of the points based on taxonomy reference database shows that the points might cluster based on reference database in high-dimensional space.
+          The separation of the points based on taxonomy reference database shows that the points cluster based on reference database in high-dimensional space.
         }

### 2022-09-28 sections/discussion.tex
Change: Updates to figure 7 text

@@
   Future work building upon our current results could enhance the network inference process in multiple ways.
-  The current analyses make use of one fecal microbiome transplant dataset with healthy and autistic samples, three mock community datasets, and several datasets generated by two synthetic interaction methods.
+  The current analyses make use of one fecal microbiome transplant dataset with healthy and ASD samples, three mock community datasets, and several datasets generated by two synthetic interaction methods.
   Incorporating datasets from a broad spectrum of biomes with varying microbial distributions into \ac{micone} will likely increase the robustness and generalizability of the results from these analyses.

### 2022-09-28 sections/methods.tex
Change: Updates to figure 7 text

@@
   The runs that contained paired-end reads (run 2 (10M reads), run 3 (750K reads) and run 4 (16M reads)), were downloaded from Qiita~\cite{qiita} (study ID 10532) and used as input sequences for the \ac{micone} pipeline.
-  Sequences from both control (212 samples including neurotypical and donors) and autistic (126 samples) patients were included in the analyses, but the groups were analyzed independently in the pipeline after the \ac{op} step.
-  All the network analyses in the study, unless explicitly mentioned, are performed indeed on the healthy and autistic samples in the fecal microbiome transplant study.
+  Sequences from both control (212 samples including neurotypical and donors) and autism spectrum disorder (ASD) (126 samples) patients were included in the analyses, but the groups were analyzed independently in the pipeline after the \ac{op} step.
+  All the network analyses in the study, unless explicitly mentioned, are performed indeed on the healthy and ASD samples in the fecal microbiome transplant study.
   However, all other analyses (preceding the NI step) are performed together on all the samples in the fecal microbiome transplant study.

### 2022-09-28 sections/results.tex
Change: Updates to figure 7 text

@@
 
-  Our analysis involves two types of data: The first type consists of sets of 16S sequencing data from samples of human stool microbiomes from a fecal microbiome transplant study of autism~\cite{Kang2017}.
+  Our analysis involves two types of data: The first type consists of sets of 16S sequencing data from samples of human stool microbiomes from a fecal microbiome transplant (FMT) study of autism~\cite{Kang2017}.
   The second type is a collection of datasets synthetically or artificially created for the specific goal of evaluating computational analysis tools.
@@
 
-  Figure~\ref{fig:figure7} shows the co-occurrence networks inferred for the healthy subjects (control) and autistic subjects in the fecal microbiome transplant study~\cite{Kang2017} (using the default tools and parameters from Table~\ref{tab:micone_tools}).
+  Figure~\ref{fig:figure7} shows the co-occurrence networks inferred for the healthy subjects (control) and subjects with autism specific disorder (ASD) in the fecal microbiome transplant study~\cite{Kang2017} (using the default tools and parameters from Table~\ref{tab:micone_tools}).
   This figure demonstrates a typical use case of comparative analysis of networks using the \ac{micone} pipeline.
   As a consequence of using the consensus network algorithm, the final co-occurrence networks are sparse and can be visually compared and examined.
+
   % TODO: Statistical test for differences + Refer to original paper and create conclusions
-  We observed 20 unique links in the network for control samples, 12 unique links in the network for autistic subjects and 7 edges in common between both networks.
+
+  Prevotella and Clostridiales are shown to have consistent evidence of association with ASD diagnostic status [REFS - Ho et al., 2020].
+  Notably, the abundance of these genera are lowered in cases with
+
+  We observed 20 unique links in the network for control samples, 12 unique links in the network for ASD subjects and 7 edges in common between both networks.
   These unique connections can serve as potential starting points for literature surveys and further experimental confirmations.

### 2022-09-28 sections/tables_figures.tex
Change: Updates to figure 7 text

@@
       The number of \ac{otu}s or \ac{esv}s generated by the respective methods are provided in the parenthesis next to the names.
-      The data used for the analysis in \textbf{(A, B)} were all the samples (healthy and autistic) from the fecal microbiome transplant (FMT) dataset~\cite{Kang2017}.
+      The data used for the analysis in \textbf{(A, B)} were all the samples from the fecal microbiome transplant (FMT) dataset~\cite{Kang2017}, containing both healthy subjects and subjects with autism spectrum disorder (ASD).
       \textbf{(C, D)} The distributions of the average weighted and unweighted UniFrac distance between the expected sequence profile and the predicted sequence profile in the mock datasets.
@@
       The number of mismatches are fewer at higher taxonomic levels, but, even at the Order level there exists greater than 51\% of mismatches, demonstrating the poor agreement in taxonomic labels assigned by the different databases.
-      The data used for the analysis in \textbf{(A, B)} were all the samples (healthy and autistic) from the FMT dataset.
+      The data used for the analysis in \textbf{(A, B)} were all the samples (healthy and ASD) from the FMT dataset.
       \textbf{(C)} The Bray-Curtis dissimilarity between the expected taxonomy profile and calculated taxonomy profile in the mock datasets shows that there is no singular best choice of database for every dataset, as all the databases have similar performance.
@@
     \caption{
-      \textbf{Comparison of networks generated from autistic and control samples using the \ac{micone} pipeline}.
-      The networks for the control (left) and autistic (right) samples in the FMT dataset were generated using the default tools and parameters recommended by the \ac{micone} pipeline.
-      There are 20 unique links in the network for control samples, 12 unique links in the network for autistic subjects and 7 edges in common between both networks.
+      \textbf{Comparison of networks generated from control and ASD samples using the \ac{micone} pipeline}.
+      The networks for the control (left) and ASD (right) samples in the FMT dataset were generated using the default tools and parameters recommended by the \ac{micone} pipeline.
+      There are 20 unique links in the network for control samples, 12 unique links in the network for ASD subjects and 7 edges in common between both networks.
     }

### 2022-09-28 sections/introduction.tex
Update on Overleaf.

@@
 
-Microbial communities are ubiquitous and play an important role in marine and terrestrial environments, urban ecosystems, and human health \cite{Zhang2015,Thompson2017,Ghoul2016,HumanMicrobiomeProjectConsortium2012,royo-llonchCompendium530Metagenomeassembled2021,tedersooFungalBiogeographyGlobal2014}.
-  These microbial communities, or microbiomes, often comprise several hundreds of different microbial strains interacting with each other and their environment, often through intricate metabolic and signaling relationships~\cite{lima-mendezDeterminantsCommunityStructure2015a,zelezniakMetabolicDependenciesDrive2015,coyteUnderstandingCompetitionCooperation2019,DSouza2018}.
+Microbial communities are ubiquitous and play an important role in marine and terrestrial environments, urban ecosystems, and human health \cite{lima-mendezDeterminantsCommunityStructure2015a,Thompson2017,Ghoul2016,HumanMicrobiomeProjectConsortium2012,royo-llonchCompendium530Metagenomeassembled2021,tedersooFungalBiogeographyGlobal2014}.
+  These microbial communities, or microbiomes, often comprise several hundreds of different microbial strains interacting with each other and their environment, often through intricate metabolic and signaling relationships~\cite{Zhang2015,zelezniakMetabolicDependenciesDrive2015,coyteUnderstandingCompetitionCooperation2019,DSouza2018}.
   Understanding how these interconnections shape community structure and functionalities is a fundamental challenge in microbial ecology, and has applications in the study of microbial ecosystems across different biomes.

### 2022-09-28 sections/introduction.tex
Change: Reference style and intro refs

@@
 Microbial communities are ubiquitous and play an important role in marine and terrestrial environments, urban ecosystems, and human health \cite{lima-mendezDeterminantsCommunityStructure2015a,Thompson2017,Ghoul2016,HumanMicrobiomeProjectConsortium2012,royo-llonchCompendium530Metagenomeassembled2021,tedersooFungalBiogeographyGlobal2014}.
-  These microbial communities, or microbiomes, often comprise several hundreds of different microbial strains interacting with each other and their environment, often through intricate metabolic and signaling relationships~\cite{Zhang2015,zelezniakMetabolicDependenciesDrive2015,coyteUnderstandingCompetitionCooperation2019,DSouza2018}.
+  These microbial communities, or microbiomes, often comprise several hundreds of different microbial strains interacting with each other and their environment, often through intricate metabolic and signaling relationships~\cite{zelezniakMetabolicDependenciesDrive2015,coyteUnderstandingCompetitionCooperation2019,DSouza2018}.
   Understanding how these interconnections shape community structure and functionalities is a fundamental challenge in microbial ecology, and has applications in the study of microbial ecosystems across different biomes.
+  % TODO: Change Narihiro2017 ref
   With the advancement in DNA sequencing technologies \cite{Narihiro2017} and data processing methods,  more information can be extracted from these microbial community samples than ever before.

### 2022-09-28 sections/results.tex
Add: Analysis for Figure 7 (text)

@@
 
-  % TODO: Statistical test for differences + Refer to original paper and create conclusions
-
-  Prevotella and Clostridiales are shown to have consistent evidence of association with ASD diagnostic status [REFS - Ho et al., 2020].
-  Notably, the abundance of these genera are lowered in cases with
-
+  The analysis of the rewiring of associations in the ASD samples with respect to the control provides a guide for identification of key genera that could be linked to dysbiosis.
   We observed 20 unique links in the network for control samples, 12 unique links in the network for ASD subjects and 7 edges in common between both networks.
-  These unique connections can serve as potential starting points for literature surveys and further experimental confirmations.
+  Although these unique connections do not imply real interactions, they can still serve as potential starting points for literature surveys and further experimental confirmations.
+  For example, the changed connectivity of Prevotella and Porphyromonas in our networks show that these genera could be useful candidates for further downstream verification.
+  Correspondingly, we have found previous studies that indicate that these genera are implicated in ASD~\cite{Kang2017,hoGutMicrobiotaChanges2020} and cognitive impairment~\cite{chiPorphyromonasGingivalisInducedCognitive2021}.
 

### 2022-09-28 sections/tables_figures.tex
Update on Overleaf.
Add: Analysis for Figure 7 (text)
Change: Move tables and fix headers
1. Tables now occupy their own pages at the end (without page numbers)
   we just have to split those pages
2. Updated the preamble such that there are no headers anymore

@@
     \centering
-    \small
-    \begin{tabular}{|c|c|c|c|}
-      \hline
-      \textbf{Workflow step} & \textbf{Module/Condition} & \textbf{Tool/Parameter} & \textbf{References/Value} \\
-      \hline
-      \multirow{7}{*}{Denoising and Clustering} & \multirow{5}{*}{Denoise and Cluster} & Closed reference & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
-                                                &  & Open reference & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
-                                                &  & De novo & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
-                                                &  & \rowcolor{lightgray} Dada2 & \cite{Callahan2016} \\
-                                                &  & Deblur & \cite{Amir2017,bolyenReproducibleInteractiveScalable2019} \\
-                                                \cline{2-4}
-                                                & \multirow{2}{*}{Chimera checking} & Uchime-denovo & \cite{rognesVSEARCHVersatileOpen2016,bolyenReproducibleInteractiveScalable2019} \\
-                                                & & \rowcolor{lightgray} Remove bimera & \cite{Callahan2016} \\
-                                                \hline
-      \multirow{5}{*}{Taxonomy assignment} &  \multirow{2}{*}{Query tool} & Blast & \cite{camachoBLASTArchitectureApplications2009,bokulichOptimizingTaxonomicClassification2018} \\
-                                           &  & \rowcolor{lightgray} Naive bayes classifier & \cite{bokulichOptimizingTaxonomicClassification2018} \\
-                                           \cline{2-4}
-                                           & \multirow{3}{*}{Database} & \rowcolor{lightgray} Greengenes 13\_8 & \cite{DeSantis2006} \\
-                                           & & SILVA 138 & \cite{Quast2012} \\
-                                           & & NCBI RefSeq (Oct 2021) & \cite{Sayers2009} \\
-      \hline
-      \multirow{6}{*}{OTU processing} & \multirow{3}{*}{Filter(off)} & Prevalence threshold      & 2 / n\_samples \\
-                                      & & Abundance threshold       & 0.001          \\
-                                      & & Observation sum threshold & 10             \\ \cline{2-4}
-                                      & \multirow{3}{*}{Filter(on)}  & \rowcolor{lightgray} Prevalence threshold      & 0.05           \\
-                                      & & \rowcolor{lightgray} Abundance threshold       & 0.01           \\
-                                      & & \rowcolor{lightgray} Observation sum threshold & 100            \\
-      \hline
-      \multirow{12}{*}{Network Inference} & \multirow{2}{*}{Bootstrapping}& fastspar\_bootstraps v1.0 & \cite{Watts2018} \\
-                                          & & fastspar\_pvalues v1.0 & \cite{Watts2018} \\
-                                          \cline{2-4}
-                                          & \multirow{6}{*}{Direct association} & \ac{spieceasi} v1.1.2 & \cite{Kurtz2015} \\
-                                          & & FlashWeave.jl v0.18.1 & \cite{tackmannRapidInferenceDirect2019} \\
-                                          & & \ac{cozine} v1.0 & \cite{haCompositionalZeroinflatedNetwork2020a} \\
-                                          & & \ac{harmonies} v1.0 & \cite{jiangHARMONIESHybridApproach2020} \\
-                                          & & \ac{spring} v1.0.4 & \cite{yoonMicrobialNetworksSPRING2019} \\
-                                          & & \ac{mldm} v1.1 & \cite{Yang2017} \\
-                                          \cline{2-4}
-                                          & \multirow{2}{*}{Correlation-based} & FastSpar (\ac{sparcc}) v1.0 & \cite{Watts2018} \\
-                                          & & Pearson & - \\
-                                          & & Spearman & - \\
-                                          & & propr v2.1.2 & \cite{quinnProprRpackageIdentifying2017} \\ \cline{2-4}
-                                          & \multirow{2}{*}{Consensus algorithm} & \rowcolor{lightgray} scaled-sum & 0.333 \\
-                                          & & simple voting & 1.000 \\
-      \hline
-    \end{tabular}
     \caption{
@@
       There are 20 unique links in the network for control samples, 12 unique links in the network for ASD subjects and 7 edges in common between both networks.
+      These connections can serve as potential starting points for further experimental validations or literature surveys.
     }
@@
 
-  \subsection*{Supplementary figures}
+  \FloatBarrier
+  \newpage
+  \subsection*{Supplementary Tables and Figures}
+
+  \renewcommand{\thefigure}{S\arabic{figure}}
+  \setcounter{figure}{0}
+
+  \renewcommand{\thetable}{S\arabic{table}}
+  \setcounter{table}{0}
+
+  \begin{table}[H]
+    \centering
+    \caption{
+      \textbf{Table of global network metrics for networks inferred from all possible combinations of tools}.
+      The metrics are averaged for each tool in every step of the pipeline, therefore, all combinations of networks that use a particular tool are grouped and averaged.
+    }
+    \label{tab:network_metrics}
+  \end{table}
 

### 2022-09-28 sections/supplementary.tex
Update on Overleaf.
Change: Move tables and fix headers
1. Tables now occupy their own pages at the end (without page numbers)
   we just have to split those pages
2. Updated the preamble such that there are no headers anymore
Change:Remove Zenodo links until published

@@
 \newpage
-\section*{Supplementary}
-
-  {\color{red} Tables S1-S5 will be moved to GitHub}
-
-  \renewcommand{\thefigure}{S\arabic{figure}}
-  \setcounter{figure}{0}
-
-  \renewcommand{\thetable}{S\arabic{table}}
-  \setcounter{table}{0}
-
-  \input{sections/supplementary_table.tex}
+\section*{Supplementary Text}
 
@@
     % TODO: Store pipeline results
-    The results of the pipeline execution for reproducing the analyses in the manuscript are stored on Zenodo[REF] at ...
+    The results of the pipeline execution for reproducing the analyses in the manuscript are stored on Zenodo.
 
@@
     % TODO: Store pipeline results
-    The results of the pipeline execution for reproducing the analyses in the manuscript are stored on Zenodo at ...
+    The results of the pipeline execution for reproducing the analyses in the manuscript are stored on Zenodo.
 
@@
   Moreover, filtering is necessary in order to increase the power in tests of significance when the number of taxa is much greater than the number of samples.
+
+  \pagenumbering{goggle}
+  \thispagestyle{empty}
+  \input{sections/table1.tex}
+  \thispagestyle{empty}
+  \input{sections/table_s1.tex}

### 2022-09-29 sections/introduction.tex
Change: Add and update references for the introduction

--- 2022-09-28 sections/introduction.tex
@@
 
-Microbial communities are ubiquitous and play an important role in marine and terrestrial environments, urban ecosystems, and human health \cite{lima-mendezDeterminantsCommunityStructure2015a,Thompson2017,Ghoul2016,HumanMicrobiomeProjectConsortium2012,royo-llonchCompendium530Metagenomeassembled2021,tedersooFungalBiogeographyGlobal2014}.
-  These microbial communities, or microbiomes, often comprise several hundreds of different microbial strains interacting with each other and their environment, often through intricate metabolic and signaling relationships~\cite{Zhang2015,zelezniakMetabolicDependenciesDrive2015,coyteUnderstandingCompetitionCooperation2019,DSouza2018}.
+Microbial communities are ubiquitous and play an important role in marine and terrestrial environments, urban ecosystems, and human health \cite{lima-mendezDeterminantsCommunityStructure2015a,Thompson2017,royo-llonchCompendium530Metagenomeassembled2021,tedersooFungalBiogeographyGlobal2014,dankoGlobalMetagenomicMap2021,mclellanMicrobiomeUrbanWaters2015,HumanMicrobiomeProjectConsortium2012}.
+  These microbial communities, or microbiomes, often comprise several hundreds of different microbial strains interacting with each other and their environment, often through intricate metabolic and signaling relationships~\cite{zelezniakMetabolicDependenciesDrive2015,Ghoul2016,coyteUnderstandingCompetitionCooperation2019,DSouza2018}.
   Understanding how these interconnections shape community structure and functionalities is a fundamental challenge in microbial ecology, and has applications in the study of microbial ecosystems across different biomes.
-  With the advancement in DNA sequencing technologies \cite{Narihiro2017} and data processing methods,  more information can be extracted from these microbial community samples than ever before.
-  In particular, high-throughput sequencing, including metagenomic sequencing and sequencing of 16S rRNA gene amplicons (hereafter referred to as 16S data) of microbial communities, has the potential to help detect, identify and quantify a large portion of the constitutive microorganisms of a microbiome \cite{Jovel2016,Lloyd-Price2016}.
-  These advances have led to large-scale data collection efforts involving terrestrial~\cite{Thompson2017,gilbertMeetingReportTerabase2010,tedersooFungalBiogeographyGlobal2014}, marine~\cite{Zhang2015,lima-mendezDeterminantsCommunityStructure2015a,royo-llonchCompendium530Metagenomeassembled2021} and human-associated microbiota~\cite{HumanMicrobiomeProjectConsortium2012,proctorIntegrativeHumanMicrobiome2019,Lloyd-Price2016}.
+  With the advancement in DNA sequencing technologies~\cite{huNextgenerationSequencingTechnologies2021,buermansNextGenerationSequencing2014,Narihiro2017},  more information can be extracted from these microbial community samples than ever before.
+  In particular, high-throughput sequencing, including metagenomic sequencing and sequencing of 16S rRNA gene amplicons (hereafter referred to as 16S data) of microbial communities, has the potential to help detect, identify and quantify a large portion of the constitutive microorganisms of a microbiome \cite{ju16SRRNAGene2015,Jovel2016,quinceShotgunMetagenomicsSampling2017,sedlarBioinformaticsStrategiesTaxonomy2017}.
+  These advances have led to large-scale data collection efforts involving terrestrial~\cite{Thompson2017,gilbertMeetingReportTerabase2010,tedersooFungalBiogeographyGlobal2014}, marine~\cite{lima-mendezDeterminantsCommunityStructure2015a,royo-llonchCompendium530Metagenomeassembled2021} and human-associated microbiota~\cite{HumanMicrobiomeProjectConsortium2012,proctorIntegrativeHumanMicrobiome2019,Lloyd-Price2016}.
 
@@
  In particular, a powerful tool for translating microbiome composition data into knowledge is the construction of association (or co-occurrence, or correlation) networks, in which  microbial taxa are represented by nodes, and frequent co-occurrences (or negative correlations) across datasets are encoded in edges between nodes.
- While the relationship between directly measured interactions~\cite{lubbeExometabolomicAnalysisCrossFeeding2017,Jian2020,Hsu2019} and statistically inferred co-occurrence is still poorly understood \cite{Zuniga2017,Rottjers2018}, a significant amount of effort has gone into estimating correlations from large microbiome sequence datasets~\cite{faustMicrobialInteractionsNetworks2012a,maEarthMicrobialCooccurrence2020a}.
+ While the relationship between directly measured interactions~\cite{lubbeExometabolomicAnalysisCrossFeeding2017,Jian2020,Hsu2019} and statistically inferred co-occurrence is still poorly understood \cite{Zuniga2017,Rottjers2018}, a significant amount of effort has gone into estimating correlations from large microbiome sequence datasets~\cite{faustMicrobialCooccurrenceRelationships2012,leeCrosskingdomCooccurrenceNetworks2022,faustMicrobialInteractionsNetworks2012a,maEarthMicrobialCooccurrence2020a}.
 
@@
 One of the most frequently used avenues for inferring co-occurrence networks is the parsing and analysis of 16S sequencing data \cite{Rottjers2018,Friedman2012}.
-Numerous software tools and pipelines have been developed to analyze 16S sequencing data, with strong emphasis on the known limitations of this methodology, including resolution, sequencing depth, compositional nature, sequencing errors and copy number variations \cite{Bharti2019,Pollock2018}.
-Popular methods for different phases of the analysis of 16S data include tools for: (i) quality checking and trimming the sequencing reads; (ii) denoising and clustering the trimmed reads \cite{Caporaso2010,Callahan2016}; (iii) assigning taxonomy to the denoised reads \cite{DeSantis2006,Quast2012}; (iv) processing and transforming the taxonomy count matrices \cite{Weiss2015}; and (v) inferring the co-occurrence network \cite{Cougoul2019,Kurtz2015}.
+Numerous software tools and pipelines have been developed to analyze 16S sequencing data, with strong emphasis on the known limitations of this methodology, including resolution, sequencing depth, compositional nature, sequencing errors and copy number variations \cite{Bharti2019,pollockMadnessMicrobiomeAttempting2018}.
+Popular methods for different phases of the analysis of 16S data include tools for: (i) quality checking and trimming the sequencing reads; (ii) denoising and clustering the trimmed reads \cite{Caporaso2010,Callahan2016,Amir2017}; (iii) assigning taxonomy to the denoised reads \cite{bokulichOptimizingTaxonomicClassification2018}; (iv) processing and transforming the taxonomy count matrices \cite{Weiss2015}; and (v) inferring the co-occurrence network \cite{Watts2018,Kurtz2015,tackmannRapidInferenceDirect2019}.
 Different specific algorithms are often aggregated into popular online platforms (like MG-RAST~\cite{Keegan2016}, Qiita~\cite{qiita}) and software packages (such as QIIME2~\cite{bolyenReproducibleInteractiveScalable2019}).

### 2022-09-29 sections/tables_figures.tex
Change: Update figure captions

@@
       \textbf{Tools used in the \ac{micone} pipeline}.
-      The tools highlighted in gray are the defaults for the pipeline and are recommended based on the benchmarks with the mock and synthetic datasets.
+      The tools highlighted in gray are the defaults for the pipeline that are recommended based on the benchmarks with the mock and synthetic datasets.
       The consensus algorithm in the \ac{ni} step incorporates all the modules (bootstrapping, direct association and correlation-based) in that step to generate the consensus network.
@@
       The steps of the workflow can be broken down into five major groups: \textbf{(SP)} \textbf{S}equence \textbf{P}rocessig, \textbf{(DC)} \textbf{D}enoising and \textbf{C}lustering, \textbf{(TA)} \textbf{T}axonomy \textbf{A}ssignment, \textbf{(OP)} \textbf{O}TU and ESV \textbf{P}rocessing, and \textbf{(NI)} \textbf{N}etwork \textbf{I}nference.
-      Each step incorporates several processes, each of which in turn have several alternate algorithms for the same task (indicated by the text to the right of the blue boxes).
-      The text along the arrows describes the data that is being passed from one step to another.
-      The final output of the pipeline is the consensus network generated from the inferred co-occurrence networks.
+      Each step incorporates several processes (blue boxes), each of which in turn have several alternate algorithms for the same task (indicated by the text to the right of the blue boxes).
+      Each arrow describes the data that is being passed from one step to another.
+      The input to the pipeline are 16S sequencing reads, and the final output is the consensus network generated from the inferred co-occurrence networks.
       For details on each process and the outputs, see Methods.
@@
     \caption{
-      \textbf{The representative sequences generated by the different DC methods differ in their identification of sequences that are low in abundance.}
-      \textbf{(A)} The average weighted UniFrac distance between the representative sequences shows that the representative sequences and their compositions are fairly identical between the methods (with the exception of Deblur (DB)).
+      \textbf{The representative sequences generated by the different denoising and clustering methods differ in their identification of sequences that are low in abundance}.
+      \textbf{(A)} The average weighted UniFrac distance between the representative sequences shows that the representative sequences and their compositions are fairly identical between the methods (with the exception of Deblur (DB) due to the low ESV count).
       \textbf{(B)} The relatively larger average unweighted UniFrac distance indicates that methods differ in their identification of sequences that are lower in abundance.
-      The number of \ac{otu}s or \ac{esv}s generated by the respective methods are provided in the parenthesis next to the names.
-      The data used for the analysis in \textbf{(A, B)} were all the samples from the fecal microbiome transplant (FMT) dataset~\cite{Kang2017}, containing both healthy subjects and subjects with autism spectrum disorder (ASD).
-      \textbf{(C, D)} The distributions of the average weighted and unweighted UniFrac distance between the expected sequence profile and the predicted sequence profile in the mock datasets.
+      The number of \ac{otu}s or \ac{esv}s generated by the respective methods are provided in the parenthesis next to their names.
+      The data used for the analysis in \textbf{(A, B)} were samples from the fecal microbiome transplant (FMT) dataset~\cite{Kang2017}, containing both healthy subjects and subjects with autism spectrum disorder (ASD).
+      \textbf{(C, D)} The distributions of the average weighted and unweighted UniFrac distance between the predicted sequence profile and the expected sequence profile in the mock datasets.
       The distributions of the average weighted UniFrac distance show that de novo (DN) and open reference (OR) were the best-performing methods in most of the datasets, while they are worst performing methods under the unweighted UniFrac metric.
       The good performance of dada2 (D2) under both distance metrics combined with its approach of identifying \ac{esv}s using de novo methods, prompts us to propose it as the default method for the DC step.
-      The data used for the analysis in \textbf{(C, D)} were the mock datasets from mockrobiota~\cite{Bokulich2016}.
+      The data used for the analysis in \textbf{(C, D)} were the mock4, mock12 and mock16 datasets from mockrobiota~\cite{Bokulich2016}.
     }
@@
     \caption{
-      \textbf{Taxonomic reference databases vary in terms of their taxonomy assignments beyond the Order level.}
-      \textbf{(A)} The assignments of the top 50 representative sequences to their respective taxonomies using the three different reference databases.
-      This result illustrates how the same sequences are assigned to different Genus under different databases.
+      \textbf{Taxonomic reference databases vary in terms of their taxonomy assignments below the Order level}.
+      \textbf{(A)} The taxonomic assignments of the top 50 representative sequences using the three different reference databases.
+      This result illustrates how the same sequences are assigned to different genera under different databases.
       A significant portion of the representative sequences are assigned to an ``unknown'' Genus in two of three databases (\ac{gg} and \ac{ncbi}).
       The number of assigned genera for each database are displayed at the top of each column.
-      \textbf{(B)} The number of representative sequences assigned to the same taxonomic label when using different reference databases, shown for the top 100 sequences.
-      The number of mismatches are fewer at higher taxonomic levels, but, even at the Order level there exists greater than 51\% of mismatches, demonstrating the poor agreement in taxonomic labels assigned by the different databases.
-      The data used for the analysis in \textbf{(A, B)} were all the samples (healthy and ASD) from the FMT dataset.
-      \textbf{(C)} The Bray-Curtis dissimilarity between the expected taxonomy profile and calculated taxonomy profile in the mock datasets shows that there is no singular best choice of database for every dataset, as all the databases have similar performance.
-      The \ac{gg} database and the Naive Bayes classifier are chosen as the defaults for the TA step of \ac{micone}.
+      \textbf{(B)} The number of representative sequences assigned to the same taxonomic label when using different reference databases (for the top 100 sequences).
+      The mismatches are fewer at higher taxonomic levels, but, even at the Order level there exists greater than 51\% of mismatches, demonstrating the poor agreement in taxonomic labels assigned by the different databases.
+      The data used for the analysis in \textbf{(A, B)} were samples (healthy and ASD) from the FMT dataset.
+      \textbf{(C)} The Bray-Curtis dissimilarity between the predicted taxonomy profile and expected taxonomy profile in the mock datasets shows that there is no singular best choice of database for every dataset, as all the databases have similar performance.
+      The \ac{gg} database and the Naive Bayes classifier are chosen as the defaults for the TA step of \ac{micone} due to their popularity.
       The datasets used for the analysis in \textbf{(C)} were the mock datasets from mockrobiota.
@@
     \caption{
-      \textbf{Networks generated using different network inference methods show notable differences both in terms of edge-density and connectivity}.
-      \textbf{(A)} The nine different networks (excluding \ac{mldm}) generated by the different network inference methods are very dissimilar.
+      \textbf{Networks generated using different network inference methods show notable differences in terms of edge-density and connectivity}.
+      \textbf{(A)} The nine different networks (excluding \ac{mldm}) generated by the different network inference methods.
+      The nodes for each network are arranged in the same positions in a circular layout and the differences in the connections can be directly visualized and compared.
       The green links are positive associations and the orange links represent negative associations.
+      The networks are dissimilar and vary widely in terms of connectivity with the correlation-based methods, in general, producing networks with higher edge-densities.
       A threshold of 0.3 was set for the correlation-based methods (sparcc, propr, spearman and pearson) and a threshold of 0.01 was set for the direct association methods (flashweave, spieceasi, cozine, harmonies, and spring).
-      The correlation-based methods in general produce graph with higher edge-densities.
       \textbf{(B)} The node overlap Upset plot indicates that all the networks have a large proportion of common nodes involved in connections (33 out of 68).
@@
       The data used in this analysis were the healthy stool samples from the FMT dataset.
-      \ac{mldm} is not shown in the comparisons because the algorithm failed to converge for the particular network combination.
+      \ac{mldm} is not shown in the comparisons because the algorithm failed to converge for the particular network combination used here (default setting of the \ac{micone} pipeline).
     }
@@
     \caption{
-      \textbf{The scaled-sum consensus method returns associations with high precision when applied to the synthetic benchmark datasets}.
+      \textbf{The associations generated by the scaled-sum consensus method show high precision in benchmarks using synthetic datasets}.
       Each point on the box plot shows the precision of the co-occurrence network generated by the individual and consensus methods using the ``NorTA'' and ``seqtime'' methods (refer to Methods).
-      The independent algorithms chosen for the comparison are the two best-performing correlation-based (propr, \ac{sparcc}) and direct association based (\ac{spieceasi}, FlashWeave) methods.
+      The independent algorithms chosen for the comparison are the two best-performing correlation-based (propr, sparcc) and direct association based (spieceasi, flashweave) methods.
+      The different consensus based methods used are the scaled-sum (SS) and simple voting (SV) method.
       A weight threshold of 0.1 and a p-value threshold of 0.05 was applied to each network before the calculation of precision.
-      The different consensus based methods used are the scaled-sum (SS) and simple voting (SV) method.
-      The overall best precision was consistently obtained by the scaled-sum method for $p \geq 0.333$ on both datasets.
-      Among all the independent network inference methods, \ac{spieceasi} has the best average precision.
-      The simple voting method when using presence of edges in all inferred networks as a requirement ($p = 1.000$), also outperforms \ac{spieceasi} on average precision.
-      Therefore, we recommend the scaled-sum consensus method with $p = 0.333$ as the default tool for network inference, as this option provides a good balance of precision and sensitivity.
+      The overall best precision was consistently obtained by the scaled-sum method for $p \geq 0.333$ on both benchmark datasets and among all the individual network inference methods, spieceasi shows the best average precision.
+      The simple voting method when using presence of edges in all inferred networks as a requirement ($p = 1.000$), also outperforms spieceasi on average precision.
+      Therefore, we recommend the scaled-sum consensus method with $p = 0.333$ as the default tool for network inference, since this option provides a good balance of precision and sensitivity.
     }
@@
       \caption{
-      \textbf{The choice of reference database contributes has the largest impact on network variance}.
-      \textbf{(A)} The percentage of variance in the networks contributed by the DC, CC (chimera checking), TA, OP and NI steps of the pipeline calculated using ANOVA on the linear model, see the Methods section for details.
+      \textbf{The choice of reference database has the largest impact on network variance}.
+      \textbf{(A)} The percentage of variance in the networks contributed by the \acf{dc}, CC (chimera checking), \acf{ta}, \acf{op} and \acf{ni} steps of the pipeline calculated using ANOVA on the linear model, see the Methods section for details.
       The data used for the analysis were the healthy stool samples from the FMT dataset.
@@
       The taxonomy database contributes most to the variance between the networks (65.4\%) followed by the filtering of the counts matrix (26.8\%) in the OP step.
-    The variation due to the DC and CC steps are much smaller in comparison (0.648\% and 0.003\% respectively).
+    The variation due to the NI, DC and CC steps are much smaller in comparison (6.553\%, 0.648\%, and 0.003\% respectively).
       The fraction labeled as the residual is an artifact that arises when multiple steps are changed at the same time, however, this can be ignored as this fraction is negligible.
-      \textbf{(B)} All combinations of inferred networks are shown as points on a PCA plot.
+      \textbf{(B)} All the inferred networks generated from various combinations of tools are shown as points on a PCA plot.
       Each point on the PCA plot represents a network inferred using different combinations of tools and parameters that are available in the \ac{micone} pipeline.
       The color of the points corresponds to the tools used at each step of the pipeline (DC, TA, OP and NI).
-      The points on the PCA plot can be grouped based on the TA step, but at the NI step only some methods create high variability in the inferred networks.
-      The \ac{dc} step does not seem to have any correlation with the similarity of networks on the PCA plot, but all the networks with filtering at the \ac{op} step are similar to each other.
-      This further confirms that the variability in the networks decreases upon filtering out the taxonomic entities at low abundance.
+      The points on the PCA plot can be grouped based on the TA step, but the extent of this grouping decreasing when the filtering is turned on in the OP step, confirming that the variability in the networks decreased upon filtering out the taxonomic entities at low abundance.
+      Some algorithms, especially the direct association methods, at the NI step can also be seen to generate networks that are less variable compared to the others.
+      The \ac{dc} step does not seem to have any correlation with the similarity of networks on the PCA plot.
     }
@@
     \caption{
-      \textbf{Comparison of networks generated from control and ASD samples using the \ac{micone} pipeline}.
-      The networks for the control (left) and ASD (right) samples in the FMT dataset were generated using the default tools and parameters recommended by the \ac{micone} pipeline.
-      There are 20 unique links in the network for control samples, 12 unique links in the network for ASD subjects and 7 edges in common between both networks.
-      These connections can serve as potential starting points for further experimental validations or literature surveys.
+      \textbf{Comparison of networks generated from the control and ASD samples of the FMT dataset using the \ac{micone} pipeline}.
+      The networks for the control (left) and ASD (right) samples were generated using the default tools and parameters recommended by the \ac{micone} pipeline as described in Table~\ref{tab:micone_tools}.
+      There are 22 unique links in the network for control samples, 12 unique links in the network for ASD subjects and 7 edges in common between both networks.
+      The changes in these connections can serve as potential starting points for further experimental validations or literature surveys.
     }
@@
       \textbf{Table of global network metrics for networks inferred from all possible combinations of tools}.
-      The metrics are averaged for each tool in every step of the pipeline, therefore, all combinations of networks that use a particular tool are grouped and averaged.
+      In each row one tool in a particular step is kept constant, and the metric is calculated for every possible combination of tools for the other steps of the pipeline.
+      Therefore, each row shows the grouped average metric for each tool in every step of the pipeline.
+      The different network inference methods show the most variation in the global network metrics.
     }
@@
           There is an increase in both weighted and unweighted UniFrac distances compared to when all the representative sequences are considered.
-          This shows that the 1000 most abundant representative sequences generated by the methods are not as similar to each other.
-          And since the weighted UniFrac is much smaller than the unweighted UniFrac distance, we can conclude that those reference sequences that are present in the middle of the abundance distribution are dissimilar.
+          This shows that the 1000 most abundant representative sequences generated by the methods are not similar to each other.
+          And since the weighted UniFrac is much smaller than the unweighted UniFrac distance, we can conclude that those reference sequences that are present in the middle of the abundance distribution (considering all sequences) are dissimilar.
         }
@@
           \textbf{The weighted and unweighted UniFrac distances between representative sequences generated using remove bimera and uchime for each denoising method are low.}
-          The two chimera checking methods, uchime and remove bimera, produce similar outputs.
-          With the exception of de novo and open reference under the unweighted UniFrac metric, all the other methods have low dissimilarity.
+          With the exception of de novo and open reference under the unweighted UniFrac metric, all the other methods have low dissimilarity, implying that the two chimera checking methods, uchime and remove bimera, produce similar outputs.
           This is especially true for the \ac{dada2} and Deblur methods which are the recommended denoising methods in the \ac{micone} pipeline.
@@
           The comparison made here is similar to Figure~\ref{fig:figure3}B, but instead of the top 100 taxonomic entities (by abundance), all the assignments from one database are matched with those from the other two databases.
-          Higher percentage of mismatches implies that the matching of the taxonomies in the more abundant sequences (top 100) are more consistent.
+          Higher percentage of mismatches implies that the assigned taxonomies in the more abundant sequences (top 100) match more consistently.
         }
@@
       \caption{
-        \textbf{The precision and sensitivity of the inferred networks on the ``NorTA'' synthetic interaction data.}
+        \textbf{The precision and sensitivity of the inferred networks on the ``NorTA'' synthetic interaction data}.
         The different consensus based methods used are: scaled-sum (SS) and simple voting (SV) method.
@@
         The simple voting method when using presence of edges in all inferred networks as a requirement ($p = 1.000$), also outperforms \ac{spieceasi} on average precision (0.969).
-        Although \ac{spieceasi} has a higher performance on sensitivity, if the goal of network inference is to obtain the list of associations that have a high probability of existing in the real microbial community, then the consensus methods perform better.
+        Although \ac{spieceasi} has a higher sensitivity, if the goal of network inference is to obtain the list of associations that have a high probability of existing in the real microbial community, then the consensus methods perform better.
       }
@@
         The simple voting method when using presence of edges in all inferred networks as a requirement ($p = 1.000$), also outperforms \ac{spieceasi} on average precision (0.692).
-        These results show that the scaled-sum method is not only much better suited for inferring robust and accurate interactions from count data regardless of the distributions and topologies, but it is also capable of accurately extracting real associations (gLV interactions) from count data.
+        These results show that the scaled-sum method is not only much better suited for inferring robust and accurate interactions from count data generated from network topologies (NorTA), but it is also capable of accurately extracting real associations from Lotka-Volterra simulations.
       }
@@
       \textbf{Sensitivity analysis of the default settings of the \ac{micone} pipeline}.
-      \textbf{(A)} The network constructed using the default pipeline parameters (DC=\ac{dada2}, CC=remove bimera, TA=\ac{gg}, OP=Filter(on), NI=scaled-sum consensus) is compared with networks generated when one of the steps use a different tool.
-      The layout is created by fixing the positions of all the nodes from all networks and then drawing only the relevant edges.
+      \textbf{(A)} The network constructed using the default pipeline parameters (DC=\ac{dada2}, CC=remove bimera, TA=\ac{gg}, OP=Filter(on), NI=scaled-sum consensus) is compared against networks generated when one of the steps uses a different tool.
+      The layout is created by fixing the positions of all the nodes from all networks and then drawing only the relevant edges, making the connections directly comparable.
       The edges colored green are positive associations and those in red are negative associations.
@@
       \textbf{(B)} The dot plot showing the fraction of nodes (left) and edges (right) in common between the default network and the networks generated by changing one step of the default pipeline.
-      The low value of the common fraction for TA and OP steps shows that these steps induce the biggest changes nodes and edges.
+      The low value of the common fraction for TA and OP steps shows that these steps induce the biggest changes in nodes and edges.
       The NI step is not shown in this analysis because the consensus method uses edges from the individual network inference methods and a comparison would be biased.

### 2022-09-29 sections/abstract.tex
Update on Overleaf.

@@
     16S ribosomal RNA (16S rRNA) amplicon profiling provides snapshots that reveal the phylogenies and abundance profiles of these microbial communities.
-    These snapshots, when collected from multiple samples, have the potential to reveal which microbes co-occur, providing a glimpse into the network of associations in these communities.
+    These snapshots, when collected from multiple samples, can reveal the co-occurrence of microbes, providing a glimpse into the network of associations in these communities.
     However, the inference of networks from 16S data involves numerous steps, each requiring specific tools and parameter choices.
@@
     In this study, we perform a meticulous analysis of each step of a pipeline that can convert 16S sequencing data into a network of microbial associations.
-    Through this process, we map how different choices of algorithms and parameters affect the co-occurrence network and estimate steps that contribute most significantly to the variance.
+    Through this process, we map how different choices of algorithms and parameters affect the co-occurrence network and identify the steps that contribute substantially to the variance.
     We further determine the tools and parameters that generate robust co-occurrence networks and develop consensus network algorithms based on benchmarks with mock and synthetic datasets.
     The Microbial Co-occurrence Network Explorer or \acs{micone} (available at \href{https://github.com/segrelab/MiCoNE}{https://github.com/segrelab/MiCoNE}), follows these default tools and parameters and can help explore the outcome of these combinations of choices on the inferred networks.
-    We envisage that this pipeline could be used for integrating multiple datasets, and for generating comparative analyses and consensus networks that can help understand microbial community assembly in different biomes.
+    We envisage that this pipeline could be used for integrating multiple datasets, and for generating comparative analyses and consensus networks that can guide our understanding of microbial community assembly in different biomes.
   }
@@
 % Insert keywords here
-\keywords{Microbiome, 16S rRNA, Pipeline, Interaction, Denoising, Taxonomy, Network Inference, Correlations, Qiime2, Co-occurrence, Networks, Consensus algorithm}
+\keywords{Microbiome, 16S rRNA, Interaction, Denoising, Taxonomy, Network Inference, Correlations, QIIME2, Co-occurrence, Networks, Consensus algorithm, Pipeline, nextflow}
 
@@
 \section*{Importance}
-  Mapping the interrelationships between different species in a microbial community is essential to understanding and controlling their structure and function.
+  Mapping the interrelationships between different species in a microbial community is important for understanding and controlling their structure and function.
   The surge in the high-throughput sequencing of microbial communities has led to the creation of thousands of datasets containing information about microbial abundances.

### 2022-09-29 sections/introduction.tex
Update on Overleaf.

@@
 Microbial communities are ubiquitous and play an important role in marine and terrestrial environments, urban ecosystems, and human health \cite{lima-mendezDeterminantsCommunityStructure2015a,Thompson2017,royo-llonchCompendium530Metagenomeassembled2021,tedersooFungalBiogeographyGlobal2014,dankoGlobalMetagenomicMap2021,mclellanMicrobiomeUrbanWaters2015,HumanMicrobiomeProjectConsortium2012}.
-  These microbial communities, or microbiomes, often comprise several hundreds of different microbial strains interacting with each other and their environment, often through intricate metabolic and signaling relationships~\cite{zelezniakMetabolicDependenciesDrive2015,Ghoul2016,coyteUnderstandingCompetitionCooperation2019,DSouza2018}.
-  Understanding how these interconnections shape community structure and functionalities is a fundamental challenge in microbial ecology, and has applications in the study of microbial ecosystems across different biomes.
+  These microbial communities, or microbiomes, often comprise several hundreds of different microbial strains interacting with each other and their environment, often through complex metabolic and signaling relationships~\cite{zelezniakMetabolicDependenciesDrive2015,Ghoul2016,coyteUnderstandingCompetitionCooperation2019,DSouza2018}.
+  Understanding how these interconnections shape community structure and function is a fundamental challenge in microbial ecology, and has applications in the study of microbial ecosystems across different biomes.
   With the advancement in DNA sequencing technologies~\cite{huNextgenerationSequencingTechnologies2021,buermansNextGenerationSequencing2014,Narihiro2017},  more information can be extracted from these microbial community samples than ever before.
-  In particular, high-throughput sequencing, including metagenomic sequencing and sequencing of 16S rRNA gene amplicons (hereafter referred to as 16S data) of microbial communities, has the potential to help detect, identify and quantify a large portion of the constitutive microorganisms of a microbiome \cite{ju16SRRNAGene2015,Jovel2016,quinceShotgunMetagenomicsSampling2017,sedlarBioinformaticsStrategiesTaxonomy2017}.
+  In particular, high-throughput sequencing, including metagenomic sequencing and sequencing of 16S rRNA gene amplicons (hereafter referred to as 16S data) of microbial communities, can help detect, identify and quantify a large portion of the constitutive microorganisms of a microbiome \cite{ju16SRRNAGene2015,Jovel2016,quinceShotgunMetagenomicsSampling2017,sedlarBioinformaticsStrategiesTaxonomy2017}.
   These advances have led to large-scale data collection efforts involving terrestrial~\cite{Thompson2017,gilbertMeetingReportTerabase2010,tedersooFungalBiogeographyGlobal2014}, marine~\cite{lima-mendezDeterminantsCommunityStructure2015a,royo-llonchCompendium530Metagenomeassembled2021} and human-associated microbiota~\cite{HumanMicrobiomeProjectConsortium2012,proctorIntegrativeHumanMicrobiome2019,Lloyd-Price2016}.
@@
  This wealth of information has the potential to help us understand how communities assemble and operate.
- In particular, a powerful tool for translating microbiome composition data into knowledge is the construction of association (or co-occurrence, or correlation) networks, in which  microbial taxa are represented by nodes, and frequent co-occurrences (or negative correlations) across datasets are encoded in edges between nodes.
+ In particular, a powerful tool for translating microbiome composition data into knowledge is the construction of association (co-occurrence or correlation) networks, in which  microbial taxa are represented by nodes, and frequent co-occurrences (or negative correlations) across datasets are encoded as edges between nodes.
  While the relationship between directly measured interactions~\cite{lubbeExometabolomicAnalysisCrossFeeding2017,Jian2020,Hsu2019} and statistically inferred co-occurrence is still poorly understood \cite{Zuniga2017,Rottjers2018}, a significant amount of effort has gone into estimating correlations from large microbiome sequence datasets~\cite{faustMicrobialCooccurrenceRelationships2012,leeCrosskingdomCooccurrenceNetworks2022,faustMicrobialInteractionsNetworks2012a,maEarthMicrobialCooccurrence2020a}.
@@
  The importance of these networks is two-fold: first, they can serve as maps that help identify hubs of keystone species \cite{Menon2018,Rottjers2018}, and the community response to environmental perturbations or underlying host conditions \cite{Gilbert2016}; second, they can serve as a bridge towards building mechanistic models of microbial communities, greatly enhancing our capacity to understand and control them.
- For example, multiple studies have shown the importance of specific microbial associations in the healthy microbiome \cite{Lloyd-Price2016,Wu2016,HumanMicrobiomeProjectConsortium2012} and the rise of dysbiosis \cite{Wang2017,Gilbert2016,Belizario2015}.
+ For example, multiple studies have shown the importance of specific microbial associations in the healthy microbiome \cite{Lloyd-Price2016,Wu2016,HumanMicrobiomeProjectConsortium2012} and their role in dysbiosis \cite{Wang2017,Gilbert2016,Belizario2015}.
  In the context of terrestrial biogeochemistry, co-occurrence networks were shown to help understand microbiome assembly \cite{Fierer2017}, and  the response of microbial communities to environmental perturbations \cite{Jiao2019}.
@@
 One of the most frequently used avenues for inferring co-occurrence networks is the parsing and analysis of 16S sequencing data \cite{Rottjers2018,Friedman2012}.
-Numerous software tools and pipelines have been developed to analyze 16S sequencing data, with strong emphasis on the known limitations of this methodology, including resolution, sequencing depth, compositional nature, sequencing errors and copy number variations \cite{Bharti2019,pollockMadnessMicrobiomeAttempting2018}.
+Numerous software tools and pipelines have been developed to analyze 16S sequencing data, with a strong emphasis on the known limitations of this method, including resolution, sequencing depth, compositional nature, sequencing errors, and copy number variations \cite{Bharti2019,pollockMadnessMicrobiomeAttempting2018}.
 Popular methods for different phases of the analysis of 16S data include tools for: (i) quality checking and trimming the sequencing reads; (ii) denoising and clustering the trimmed reads \cite{Caporaso2010,Callahan2016,Amir2017}; (iii) assigning taxonomy to the denoised reads \cite{bokulichOptimizingTaxonomicClassification2018}; (iv) processing and transforming the taxonomy count matrices \cite{Weiss2015}; and (v) inferring the co-occurrence network \cite{Watts2018,Kurtz2015,tackmannRapidInferenceDirect2019}.
-Different specific algorithms are often aggregated into popular online platforms (like MG-RAST~\cite{Keegan2016}, Qiita~\cite{qiita}) and software packages (such as QIIME2~\cite{bolyenReproducibleInteractiveScalable2019}).
-The different methods and tools developed to solve issues arising in 16S analysis can lead to vastly different inferences of community compositions and co-occurrence networks \cite{Golob2017,Weiss2016}, making it difficult to reliably compare networks across different publications and studies.
- This is partially due to the fact that existing platforms are typically focused on \ac{otu} or \ac{esv} generation and not on the effects of upstream statistical methods on the inferred co-occurrence networks.
- Furthermore, no organized framework currently exist to systematically analyze and compare existing components of the data analysis from amplicons to networks.
+Different specific algorithms are often aggregated into popular online platforms (like MG-RAST~\cite{Keegan2016}, Qiita~\cite{qiita}) and software packages (such as \ac{qiime2}~\cite{bolyenReproducibleInteractiveScalable2019}).
+The different methods and tools can lead to vastly different inferences of community compositions and co-occurrence networks \cite{Golob2017,Weiss2016}, making it difficult to reliably compare networks across different publications and studies.
+ This difference is partially due to the focus of existing platforms on \ac{otu} or \ac{esv} generation and not on the effects of upstream statistical methods on the inferred co-occurrence networks.
+ Furthermore, no organized framework currently exists that can systematically analyze and compare each step in the pipeline for processing amplicons into co-occurrence networks.
 
- In this study, we present a standardized 16S data analysis pipeline called \ac{micone} that produces robust and reproducible co-occurrence networks from community 16S sequence data, and allow users to interactively explore how the network would change upon using different alternative tools and parameters at each step.
- Our pipeline is coupled to an online integrative tool for the organization, visualization and analysis of inter-microbial networks called \ac{mind}~\cite{huResourceComparisonIntegration2022}, that is available at \href{http://microbialnet.org/}{http://microbialnet.org/}.
-Through a systematic comparative analysis we determined which steps of the \ac{micone} pipeline have the largest influence on the final network, and what choice seems to have the best agreement with the tested mock and synthetic datasets.
-These steps will ensure better reproducibility and easier comparison of co-occurrence networks across datasets.
+ In this study, we present a standardized 16S data analysis pipeline called \ac{micone} that produces robust and reproducible co-occurrence networks from 16S sequence data of microbial communities, and enable users to interactively explore how the network would change upon using different alternative tools and parameters at each step.
+ Our pipeline is coupled to an online integrative tool for the organization, visualization, and analysis of inter-microbial networks called \ac{mind}~\cite{huResourceComparisonIntegration2022}, which is available at \href{http://microbialnet.org/}{http://microbialnet.org/}.
+Through a systematic comparative analysis, we determine which steps of the \ac{micone} pipeline have the largest influence on the final network, and which choice seems to have the optimal agreement with the tested mock and synthetic datasets.
+These steps together with our default settings ensure better reproducibility and easier comparison of co-occurrence networks across datasets.
 We expect that our tool will also be useful for benchmarking future alternative methods, and for ensuring a transparent evaluation of the possible biases introduced by the use of specific tools.

### 2022-09-29 sections/results.tex
Change: Add and update references for the introduction
Update on Overleaf.

@@
   We developed \ac{micone}, a flexible and modular pipeline for the inference of co-occurrence networks from 16S data.
-  \ac{micone} incorporates various popular, publicly available tools as well as custom Python modules for 16S data analysis and network inference (see Methods).
-  The different steps that are a part of the \ac{micone} co-occurrence network inference workflow (Figure~\ref{fig:figure1}) can be grouped into five major modules; (i) \ac{sp}; (ii) \ac{dc}; (iii) \ac{ta}; (iv) \ac{op}; (v) \ac{ni}.
-  Each process in the pipeline can be implemented through  multiple tools (see Methods and Figure~\ref{fig:figure1}).
+  \ac{micone} incorporates various popular, publicly available tools as well as custom Python modules for 16S data analysis and network inference (Methods).
+  The different steps that are a part of the \ac{micone} co-occurrence network inference workflow (Figure~\ref{fig:figure1}) can be grouped into five major modules; (i) \ac{sp}; (ii) \ac{dc}; (iii) \ac{ta}; (iv) \ac{op}; and (v) \ac{ni}.
+  Each process in the pipeline is implemented through multiple tools (see Methods and Figure~\ref{fig:figure1}).
   The effects of changing any intermediate step of the pipeline can be evaluated in terms of the final network outcome, as well as on any of the intermediate metrics and data outputs.
-  The choice of tools and parameters is encoded in a configuration file (with defaults shown in Tables S2-S6 at \href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper}).
+  The choice of tools and parameters is encoded in a configuration file (with parameters as shown in Tables S2-S6 at \href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper}).
   Through a systematic analysis of tool combinations at each step of the pipeline, we estimated how much the final co-occurrence network depends on the possible choices at each step.
 
-  Our analysis involves two types of data: The first type consists of sets of 16S sequencing data from samples of human stool microbiomes from a fecal microbiome transplant (FMT) study of autism~\cite{Kang2017}.
-  The second type is a collection of datasets synthetically or artificially created for the specific goal of evaluating computational analysis tools.
-  In particular, in order to  compare how well each step in \ac{micone} best captures the underlying data, we use both mock data (labeled mock4, mock12, and mock16) from mockrobiota~\cite{Bokulich2016} and synthetic networks generated using the NorTA~\cite{Kurtz2015} and seqtime~\cite{Rottjers2018} approaches (See Methods).
+  Our analysis involved two types of data: The first type consisted of 16S sequencing data from samples of human stool microbiomes from a fecal microbiome transplant (FMT) study of autism~\cite{Kang2017}.
+  The second type was a collection of datasets synthetically or artificially created for the specific goal of evaluating computational analysis tools.
+  In particular, in order to benchmark each step in \ac{micone}, we used both mock data (labeled mock4, mock12, and mock16) from mockrobiota~\cite{Bokulich2016} and synthetic networks generated using the NorTA~\cite{Kurtz2015} and seqtime~\cite{Rottjers2018} approaches (See Methods).
 
@@
 
-  \subsection*{DC: Denoising and clustering differ in their identification of sequences that are low in abundance}
+  \subsection*{DC: Denoising and clustering methods differ in their identification of sequences that are low in abundance}
 
   The \ac{dc} step is commonly carried out to generate representative sequences (in the form of the \acs{otu}/\acs{esv} tables) from the demultiplexed and trimmed 16S sequencing data.
-  In order to compare the count tables generated by different tools, we processed the same 16S sequencing reads (from the FMT study~\cite{Kang2017}) using 5 different methods: open-reference clustering, closed-reference clustering, de novo clustering, \ac{dada2}~\cite{Callahan2016} and Deblur~\cite{Amir2017}.
+  In order to compare the count tables generated by different tools, we processed the 16S sequencing reads (from the FMT study~\cite{Kang2017}) using 5 different methods: open-reference clustering, closed-reference clustering, de novo clustering, \ac{dada2}~\cite{Callahan2016} and Deblur~\cite{Amir2017}.
   The first three methods are from the vsearch plugin from \ac{qiime2}~\cite{bolyenReproducibleInteractiveScalable2019}.
-  The closed and open reference methods in this analysis use the \ac{gg}~\cite{DeSantis2006} database for reference sequence alignment.
+  The closed and open reference methods in this analysis use the \acl{gg}~\cite{DeSantis2006} database for reference sequence alignment.
 
-  A comparison of the different approaches is visualized in the form of mean UniFrac values across all samples (Figure~\ref{fig:figure2}).
+  A comparison of the different methods was carried out by calculating the mean UniFrac distances across all samples (Figure~\ref{fig:figure2}).
   The analysis was performed using both the weighted UniFrac~\cite{Lozupone2007} (Figure~\ref{fig:figure2}A) distance metric, which takes into account the counts of the representative sequences, and the unweighted UniFrac~\cite{Lozupone2005} (Figure~\ref{fig:figure2}B) distance metric, which gives equal weights to each sequence.
@@
   A second message is that the different methods differ mainly in the assignment of sequences of lower abundance.
-This can be inferred from the fact that the unweighted comparison (Figure~\ref{fig:figure2}B) shows an increase in dissimilarity between each pair of methods (see additional details in Supplementary and Figure \ref{fig:figure_s3}).
+  This can be inferred from the unweighted comparison (Figure~\ref{fig:figure2}B) which shows an increase in dissimilarity between each pair of methods (see additional details in Supplementary and Figure \ref{fig:figure_s2}).
 
-  These comparisons only elucidate the similarity of a pair of methods.
-  In order to determine which tool most accurately recapitulates the reference sequences in the samples, we applied the same pipeline step to process the mock datasets (mock4, mock12, and mock16) and compared the predicted representative sequences with the true sequences and their distribution.
-  The results (Figure~\ref{fig:figure2}C and D) show that the predicted sequence distributions are overall considerably different from the expected ones.
-  The significant variation across datasets indicates that the datasets themselves play a big role in method performance.
+  These comparisons only elucidate the similarity between a pair of methods.
+  To determine which tool most accurately recapitulates the reference sequences in the samples, we applied the same pipeline step to process the mock datasets (mock4, mock12, and mock16) and compared the predicted representative sequences with the true sequences and their distribution.
+  The results (Figure~\ref{fig:figure2}C and \ref{fig:figure2}D) show that the predicted sequence distributions are overall different from the expected ones.
+  The variation across datasets indicates that the datasets themselves play a big role in method performance.
   There is no method that outperforms the rest in all datasets (see Supplementary for an extended discussion).
   Based on their slightly better performance on the mock datasets, their de novo error correcting nature and previous independent evaluation~\cite{Nearing2018}, \ac{dada2} and Deblur appear to be the most reliable.
-  This is because the open-reference and de novo clustering methods return a much larger number of \ac{otu}s compared to the other pipelines and this would affect the accuracy of the network inference step if filtering is not performed.
+  This is because the open-reference and de novo clustering methods return a much larger number of \ac{otu}s compared to the other pipelines and would affect the accuracy of the network inference step if stringent filtering is not performed.
   Overall, since \ac{dada2} as compared to Deblur, displays better performance on all the mock datasets on the weighted UniFrac metric, we set this tool as the default for the DC step of the pipeline.
@@
 
-  After the denoising, the sequences are subject to chimera checking.
+  After the denoising, the sequences are subject to Chimera Checking (CC).
   The \ac{micone} pipeline supports two different chimera checking methods, ``uchime-denovo"~\cite{bolyenReproducibleInteractiveScalable2019}, and ``remove bimera"~\cite{Callahan2016}.
-  We did not notice any significant difference between the two methods (Figure~\ref{fig:figure_s4}), implying that they identify and remove mostly the same set of sequences as chimeras.
-  Since the remove bimera method was originally developed in conjunction with dada2 we use this method as the default for the chimera checking module.
+  We did not notice any notable difference between the two methods (Figure~\ref{fig:figure_s3}), implying that they identify and remove mostly the same set of sequences as chimeras.
+  Since the remove bimera method was originally developed in conjunction with dada2 we use this method as the default.
   The DC step thus results in a reduced set of unique sequences, which will be referred to as representative sequences in the subsequent steps.
@@
   This implies that the taxonomy assignments from each reference database are fairly unique.
-  The comparison of all assigned genera (Figure~\ref{fig:figure_s5}), instead of just the top 100, contains a higher percentage of mismatches.
+  The comparison of all assigned genera (Figure~\ref{fig:figure_s4}), instead of just the top 100, contains a higher percentage of mismatches.
   This suggests that many of the most abundant sequences are consistently matched to the same taxonomies in the different reference databases, at least for the dataset tested in the current analysis.
@@
   One reason to choose a particular database could be the frequency of updates and the potential for future growth.
-  Both \ac{gg}, due to its frequent use in the literature~\cite{Balvociute2017}, and \ac{ncbi}, due to its regular revision and maintenance, could be good choices for taxonomy assignment, with few differences observed between the two in our analysis  (Figure~\ref{fig:figure_s5}).
+  Both \ac{gg}, due to its frequent use in the literature~\cite{Balvociute2017}, and \ac{ncbi}, due to its regular revision and maintenance, could be good choices for taxonomy assignment.
   In our default pipeline, we choose \ac{gg} as the default method.
@@
   This counts table is used as the input to the \ac{micone} pipeline to generate predicted associations.
-  The interaction network used to generate the counts table is used as the source of true interactions to calculate the precision (Figure \ref{fig:figure5}) and sensitivity (Figure \ref{fig:figure_s6} and Figure \ref{fig:figure_s7}) for each network inference algorithm.
+  The interaction network used to generate the counts table is used as the source of true interactions to calculate the precision (Figure \ref{fig:figure5}) and sensitivity (Figure \ref{fig:figure_s5} and Figure \ref{fig:figure_s6}) for each network inference algorithm.
   As shown in Figure \ref{fig:figure5} the consensus algorithm, especially the scaled-sum method, captures true associations with high precision (through the removal of edges that are either not present in most of the inference methods or whose association strength is low across methods).
   Overall, the scaled-sum method for $p=1.000$ performs the best (precision = $1.000$ for both NorTA and seqtime).
-  The scaled-sum method for $p=0.333$ (default option in the pipeline) shows a high precision ($0.956$ with NorTA; $0.688$ with seqtime), without displaying significant reduction in sensitivity (Figure~\ref{fig:figure_s6} and Figure~\ref{fig:figure_s7}).
+  The scaled-sum method for $p=0.333$ (default option in the pipeline) shows a high precision ($0.956$ with NorTA; $0.688$ with seqtime), without displaying significant reduction in sensitivity (Figure~\ref{fig:figure_s5} and Figure~\ref{fig:figure_s6}).
   However, if higher precision is required $p>0.5$ can be considered.
@@
   This result highlights the importance of the taxonomy assignment step in the 16S data analysis workflow, implying that a change in the reference database will result in drastically different inferred networks.
-  This is likely due to the differential assignment of representative sequences to taxonomic entities (Figure~\ref{fig:figure3} and Figure~\ref{fig:figure_s5}), which drastically alter the nodes and hence the underlying network topology.
+  This is likely due to the differential assignment of representative sequences to taxonomic entities (Figure~\ref{fig:figure3} and Figure~\ref{fig:figure_s4}), which drastically alter the nodes and hence the underlying network topology.
 
@@
   The PCA in Figure \ref{fig:figure6}B shows all the above networks, colored by the databases used in TA (left) and by the NI approach (right).
-  The major effect of the TA step choice, shown before in Figure \ref{fig:figure6}A, is also reflected in the PCA plot, where networks segregate based on the database used (Figure \ref{fig:figure6}B).
+  The major effect of the TA step choice, shown before in Figure \ref{fig:figure6}A, is also reflected in the PCA plot, where networks segregate based on the database used (Figure~\ref{fig:figure6}B and Figure~\ref{fig:figure_s1}).
   It is also evident that some networks, especially those inferred using the direct association network inference methods, are much closer in the PCA plot regardless of the reference database used.
-  Additionally, the variation (on the first two principal axes) between the networks decreases when the low abundance \ac{otu}s are removed from the network (Figure~\ref{fig:figure_s1}).
+  Additionally, the variation (on the first two principal axes) between the networks decreases when the low abundance \ac{otu}s are removed from the network.
   These results suggest that the most important criterion for accurate comparative analysis of co-occurrence networks is the taxonomy reference database followed by the level of filtering of the taxonomy tables.
@@
 
-  The analysis of the rewiring of associations in the ASD samples with respect to the control provides a guide for identification of key genera that could be linked to dysbiosis.
-  We observed 20 unique links in the network for control samples, 12 unique links in the network for ASD subjects and 7 edges in common between both networks.
+  The analysis of the rewiring of associations in the ASD samples with respect to the control provides a guide for the identification of key genera that could be linked to dysbiosis.
+  We observed 20 unique links in the network for control samples, 12 unique links in the network for ASD subjects, and 7 edges in common between both networks.
   Although these unique connections do not imply real interactions, they can still serve as potential starting points for literature surveys and further experimental confirmations.
@@
   Correspondingly, we have found previous studies that indicate that these genera are implicated in ASD~\cite{Kang2017,hoGutMicrobiotaChanges2020} and cognitive impairment~\cite{chiPorphyromonasGingivalisInducedCognitive2021}.
+  Additional visualization and comparison of networks can be performed using the \acl{mind}~\cite{huResourceComparisonIntegration2022}.
 

### 2022-09-29 sections/discussion.tex
Update on Overleaf.

@@
   A myriad of tools and methods have been developed for different parts of the workflow for inference of co-occurrence networks from 16S rRNA data.
-  Our analysis shows that networks generated using different combinations of tools and approaches can be significantly different from each other, highlighting the need for a clear assessment of the source of variability and for tools that provide the most robust and accurate results.
-  Our newly developed integrated software for the inference of co-occurrence networks from 16S rRNA data, \ac{micone}, is a freely customizable and user-friendly pipeline that allows users to compare networks generated by multiple possible combinations of tools and parameters.
+  Our analyses have shown that networks generated using different combinations of tools and approaches can be substantially different from each other, highlighting the need for a clear evaluation of the source of variability and for tools that provide the most robust and accurate results.
+  Our newly developed software, \ac{micone}, is a customizable pipeline for the inference of co-occurrence networks from 16S rRNA data that enables users to compare networks generated by multiple possible combinations of tools and parameters.
   Importantly, in addition to revisiting the test cases presented in this work, users will be able to explore the effect of various tool combinations on their own datasets of interest.
-  The \ac{micone} pipeline is built in a modular fashion; its plug-and-play architecture enables users to add new tools and steps, either using existing packages that were not examined in the present work or those developed in the future.
+  The \ac{micone} pipeline has been built in a modular fashion; its plug-and-play architecture enables users to add new tools and steps, either using existing packages that have not been examined in the present work or those developed in the future.
   The \ac{micone} Python package provides functions and methods to perform a detailed analysis of the count matrices and the co-occurrence networks.
-  The inferred networks are exported to a custom JSON format (see Supplementary) by default, but can also be exported to Cytoscape~\cite{shannonCytoscapeSoftwareEnvironment2003}, GML~\cite{himsoltGMLPortableGraph2010}, and many other popular formats via the \ac{micone} Python package.
+  The inferred networks are exported to a custom JSON format (see Supplementary) by default, but can also be exported to Cytoscape~\cite{shannonCytoscapeSoftwareEnvironment2003}, GML~\cite{himsoltGMLPortableGraph2010}, and many other popular formats via the Python package.
 
@@
 
-  Through \ac{micone}, in addition to transparently revealing the dependence of co-occurrence networks on tool and parameter choices (see Discussion in Supplementary Text for details on the DC, TA and OP steps), we take advantage of our spectrum of computational options and the availability of mock and synthetic datasets, to suggest a default standard setting.
-  Additionally, we develop a consensus approach, that can reliably generate networks that are robust across multiple tool/parameter choices.
-  An important caveat related to these results is the fact that due to the lack of a universal standard for microbial interaction data, our conclusions are based on the specific datasets used in our analysis.
+  Through \ac{micone}, in addition to transparently revealing the dependence of co-occurrence networks on tool and parameter choices (see Discussion in Supplementary Text for details on the DC, TA and OP steps), we have taken advantage of our spectrum of computational options and the availability of mock and synthetic datasets, to suggest a default standard setting.
+  Additionally, we have developed a consensus approach, that can reliably generate networks that are fairly robust across multiple tool choices.
+  An important caveat related to these results is that due to the lack of a universal standard for microbial interaction data, our conclusions are based on the specific datasets used in our analysis.
   While our analysis is based on several mock and synthetic datasets that cover a diverse range of abundance distributions and network topologies, datasets that have drastically different distributions may require a re-assessment of the best settings.
 
-  The networks generated by different network inference methods show considerable differences in edge-density and connectivity, partially due to the underlying assumptions regarding sparsity, distribution and compositionality.
+  The networks generated by different network inference methods show considerable differences in edge-density and connectivity, partially due to the underlying assumptions regarding sparsity, distribution, and compositionality.
   To address this issue, we have developed two consensus algorithms (simple voting and scaled-sum method) that generate networks whose links have evidence based on multiple inference algorithms.
 
- We find that the scaled-sum method (along with p-value merging for correlation-based methods) performs the best on synthetic datasets, and is therefore chosen as default for the \ac{ni} step of the pipeline.
- Notably, the consensus network displays a higher precision, and returns a concise list of robust associations which represent a valuable set for experimental validation follow-up.
+ We find that the scaled-sum method performs the best on synthetic datasets, and is therefore chosen as the default for the \ac{ni} step of the pipeline.
+ Notably, the consensus network displays a higher precision and returns a concise list of robust associations which represent a valuable set for experimental validation follow-up.
 
@@
 
-  The network analyses in this study are primarily at the Genus level, wherein the lowest resolution of a node is a Genus and if an entity cannot be resolved to the Genus level, the next lowest taxonomic level is used (e.g.
-Family).
+  The network analyses in this study are primarily at the Genus level, wherein the lowest resolution of a node is a Genus and if an entity cannot be resolved to the Genus level, the next lowest taxonomic level is used (for example, Family).
   As a consequence, two entities belonging to the same lineage where one entity is resolved to the Genus level and another is resolved to the Family level are treated as two different nodes in the network.
-  Therefore, the development of a metric of overlap to compare nodes with shared lineages within and across networks could provide more biologically and phylogenetically relevant comparisons.
+  Thus, the development of a metric of overlap to compare nodes with shared lineages within and across networks could enable more biologically and phylogenetically relevant comparisons.
 
   Although direct comparisons between co-occurrence networks and directly measured interactions are difficult to interpret and highly debated~\cite{hiranoDifficultyInferringMicrobial2019,gobernaCautionaryNotesUse2022}.
-  Further, benchmarking of co-occurrence networks could also be pursued through the utilization of literature-based interactions~\cite{lima-mendezDeterminantsCommunityStructure2015a} or biological benchmark interaction data~\cite{sungGlobalMetabolicInteraction2017a}.
+  Further, benchmarking of co-occurrence networks could also be pursued through the use of literature-based interactions~\cite{lima-mendezDeterminantsCommunityStructure2015a} or biological benchmark interaction data~\cite{sungGlobalMetabolicInteraction2017a}.
   Additionally, \ac{micone} could be extended to enable the processing of metagenomics sequencing data, facilitating the analysis of a much larger and diverse range of datasets and domains of life.
 
-  Although in the current analysis, we only use default parameter values recommended by the tool creators, the \ac{micone} pipeline could be used in the future to explore any combinations of parameters and to optimize these values for improved network inference.
-  Overall, it is important to keep in mind that there likely is no ``best method'' for the various steps of 16S data analysis.
-  Therefore, \ac{micone} is intended to help researchers to identify the methods and algorithms that are most suitable for their datasets in an easy-to-use and reproducible manner.
+  Although in the current analysis, we have only used default parameter values recommended by the tool creators, the \ac{micone} pipeline could be used in the future to explore any combinations of parameters and to optimize these values for improved network inference.
+  Overall, there likely is no ``best method'' for the various steps of 16S data analysis, and hence, \ac{micone} is intended to help researchers to identify the methods and algorithms that are most suitable for their datasets in an easy-to-use and reproducible manner.
 
   We envision that \ac{micone}, and its underlying tools and databases, will be increasingly useful for building large comparative analyses across studies.
-  It allows for rapid, configurable, and reproducible inference of microbial networks and further for the formulation of hypotheses about the role of these interactions on community composition and stability.
+  It enables rapid, configurable, and reproducible inference of microbial networks and furthers the formulation of hypotheses about the role of these interactions on community composition and stability.
   These comparative analyses will require coupled network analysis and visualization tools (such as \ac{mind}~\cite{huResourceComparisonIntegration2022}) and need systematic access to datasets, shared in accordance with FAIR standards~\cite{pachecoFAIRRepresentationsMicrobial2022}.

### 2022-09-29 sections/results.tex
Update on Overleaf.

@@
   The variation across datasets indicates that the datasets themselves play a big role in method performance.
-  There is no method that outperforms the rest in all datasets (see Supplementary for an extended discussion).
+  We note that there is no method that outperforms the rest in all datasets (see Supplementary for an extended discussion).
   Based on their slightly better performance on the mock datasets, their de novo error correcting nature and previous independent evaluation~\cite{Nearing2018}, \ac{dada2} and Deblur appear to be the most reliable.
@@
   Taxonomy databases are used to assign taxonomic identities to the representative sequences obtained after the DC step.
-  The three 16S taxonomic reference databases used in this study are SILVA~\cite{Quast2012}, \acs{gg}~\cite{DeSantis2006} and \acs{ncbi} RefSeq~\cite{Sayers2009} (see Methods).
-  These databases vary significantly in terms of taxonomy hierarchies, including species names and phylogenetic relationships~\cite{Balvociute2017}.
+  The three 16S taxonomic reference databases used in this study are SILVA~\cite{Quast2012}, \ac{gg}~\cite{DeSantis2006} and \ac{ncbi} RefSeq~\cite{Sayers2009} (Methods).
+  These databases vary substantially in terms of taxonomy hierarchies, including species names and phylogenetic relationships~\cite{Balvociute2017}.
   Assignment using a particular database also requires a query tool.
-  We use the ``Naive Bayes'' classifier from \ac{qiime2} for the \ac{gg} and SILVA databases and the ``BLAST'' tool (included as a \ac{qiime2} plugin) for the \ac{ncbi} database.
-  These tools have been well quantified and optimized~\cite{bokulichOptimizingTaxonomicClassification2018}, hence, we make use of the default parameters in our analyses.
+  We used the ``Naive Bayes'' classifier from \ac{qiime2} for the \ac{gg} and SILVA databases and the ``BLAST'' tool (included as a \ac{qiime2} plugin) for the \ac{ncbi} database.
+  These tools have been well quantified and optimized~\cite{bokulichOptimizingTaxonomicClassification2018}, hence, we made use of the default parameters in our analyses.
 
@@
   Some of the differences in Genus composition have a clear explanation, for example, abundant Genera like Bacteroides and Escherichia are assigned to different representative sequences.
-  The large percentage of unassigned sequences is due to the large fraction of the representative sequences assigned to an "unknown" Genus during the assignment process (see Methods).
+  The large percentage of unassigned sequences is due to the large fraction of the representative sequences assigned to an "unknown" Genus during the assignment process (Methods).
 
-  After assigning all the representative sequences to taxonomies we performed a pairwise comparison of the similarity between the top 100 assignments (by abundance) from different databases at every taxonomic level (Figure~\ref{fig:figure3}B).
-  The comparisons of the assignments beyond the Family level (Family, Genus, and Species) show less than $45\%$ similarity between any pair of databases.
+  After the assignment, we performed a pairwise comparison of the similarity between the top 100 assignments (by abundance) from different databases at every taxonomic level (Figure~\ref{fig:figure3}B).
+  The comparisons of the assignments below the Order level (Family, Genus, and Species) show less than $45\%$ similarity between any pair of databases.
   This implies that the taxonomy assignments from each reference database are fairly unique.
   The comparison of all assigned genera (Figure~\ref{fig:figure_s4}), instead of just the top 100, contains a higher percentage of mismatches.
-  This suggests that many of the most abundant sequences are consistently matched to the same taxonomies in the different reference databases, at least for the dataset tested in the current analysis.
+  This suggests that, comparatively, the most abundant sequences are more consistently matched to the same taxonomies, at least for the dataset tested in the current analysis.
 
-  In order to obtain an absolute measure of the accuracy of the taxonomic assignments, we used the representative sequences from the DC step for mock datasets as the query sequences and the expected taxonomic composition as the standard to compare against.
+  To obtain an absolute measure of the accuracy of the taxonomic assignments, we used the representative sequences from the DC step for mock datasets as the query sequences and the expected taxonomic composition as the standard to compare against.
   We used the Bray-Curtis distance metric~\cite{virtanenSciPyFundamentalAlgorithms2020} to calculate the distance between the predicted and expected taxonomic distribution (Figure~\ref{fig:figure3}C).
@@
   The TA step results in a taxonomic counts table that is used as input to the subsequent steps of the pipeline.
-  Note that the count tables at different levels can be obtained through aggregation: for example, Genus count tables were obtained by summing up the counts of the lower taxonomy levels (Species and \ac{otu}) that map to the same higher taxonomy level entity.
+  Note that the count tables at different levels can be obtained through aggregation; for example, Genus count tables were obtained by summing up the counts of the lower taxonomy levels (Species and \ac{otu}) that map to the same higher taxonomy level entity.
 
@@
 
-   The ten network inference methods we used in this step fall into two groups: the first set of methods (Pearson, Spearman, \acs{sparcc}~\cite{Friedman2012,Watts2018}, and propr~\cite{quinnProprRpackageIdentifying2017}) infer pairwise correlations while the second set infer direct associations (\acs{spieceasi}~\cite{Kurtz2015}, FlashWeave~\cite{tackmannRapidInferenceDirect2019}, \acs{cozine}~\cite{haCompositionalZeroinflatedNetwork2020a}, \acs{harmonies}~\cite{jiangHARMONIESHybridApproach2020}, \acs{spring}~\cite{yoonMicrobialNetworksSPRING2019}, and \acs{mldm}~\cite{Yang2017}).
+   The ten network inference methods we used in this step fall into two groups: the first set of methods (Pearson, Spearman, \acs{sparcc}~\cite{Friedman2012,Watts2018}, and propr~\cite{quinnProprRpackageIdentifying2017}) infer pairwise correlations while the second set (\acs{spieceasi}~\cite{Kurtz2015}, FlashWeave~\cite{tackmannRapidInferenceDirect2019}, \acs{cozine}~\cite{haCompositionalZeroinflatedNetwork2020a}, \acs{harmonies}~\cite{jiangHARMONIESHybridApproach2020}, \acs{spring}~\cite{yoonMicrobialNetworksSPRING2019}, and \acs{mldm}~\cite{Yang2017}) infer direct associations.
    Note that while Pearson and Spearman methods are included in the pipeline for completeness, they tend to generate a large number of spurious edges as they are not intended for compositional datasets.
-   Thus they are not included in subsequent quantitative analyses.
+   Thus, they are not included in subsequent quantitative analyses.
 
-  Filtered (see \ac{op} step in Methods) genus-level counts table obtained using the default settings in the previous steps were used as input for the different network inference algorithms (Figure~\ref{fig:figure4}A).
-  Even from a visual inspection, one can see that the different networks differ vastly in their edge-density and connectivity, with common edges often displaying inverted signs.
+  Filtered (see \ac{op} step in Methods) genus-level counts table obtained using the default settings in the previous steps were used as input for the different network inference algorithms (Figure~\ref{fig:figure4}).
+  Even from a visual inspection (Figure~\ref{fig:figure4}A), one can see that the different networks differ vastly in their edge-density and connectivity, with common edges often displaying inverted signs.
 
-  In order to quantify the differences between the networks, we analyzed the distribution of common nodes and edges (Figure~\ref{fig:figure4} B and C) using UpSet plots~\cite{Lex}.
+  To quantify the differences between the networks, we analyzed the distribution of common nodes and edges (Figure~\ref{fig:figure4} B and \ref{fig:figure4}C) using UpSet plots~\cite{Lex}.
   The node intersection analysis shows that the networks have $33$ out of $68$ total unique nodes in common and that no network possesses a unique node.
-  Edge intersections in contrast show that only $8$ edges (out of $202$ total unique edges) are in common between all the methods and each network has a large number of unique edges.
-  These results indicate that there is a substantial rewiring of connections in different inferred networks, prompting us to identify associations robust across methods, through consensus algorithms.
+  Edge intersections in contrast show that only $8$ edges (out of $202$ total unique edges) are in common between all the methods and each network has many unique edges.
+  These results showed a substantial rewiring of connections in different inferred networks and prompted us to identify associations robust across methods, through consensus algorithms.
 
@@
 
-  \subsection*{NI: The consensus network algorithms}
+  \subsection*{NI: The scaled-sum consensus method shows high precision on benchmark datasets}
 
- Inspired by previous approaches~\cite{bustinceFuzzySetsTheir2008,tsarevApplicationMajorityVoting2018}, we developed two methods that take into consideration the evidence offered by each network inference algorithm and generate a consensus network that contains the most probable associations.
+ Inspired by previous approaches~\cite{bustinceFuzzySetsTheir2008,tsarevApplicationMajorityVoting2018}, we developed two methods that take into consideration the evidence offered by each network inference algorithm and generate a consensus network that contains the common edges among the inferred networks.
 
-  Both of our approaches - simple voting (SV) and scaled-sum (SS) - combine appropriately filtered networks inferred from correlation-based and direct association methods (see Methods, section "Consensus network and p-value merging").
+  Both of our approaches - simple voting (SV) and scaled-sum (SS) - combine appropriately filtered networks inferred from correlation-based and direct association methods (see Methods).
   We chose the scaled-sum method as the pipeline default since this method takes into account the weights of the associations in the determination of the final consensus.
-  The pipeline allows for any choice of a subset of methods for the consensus calculation. Currently, by default, all direct methods are used, together with \acs{sparcc} and propr.
+  The pipeline enables the selection of any subset of methods for the consensus calculation. Currently, by default, all direct methods are used, together with \acs{sparcc} and propr.
 
@@
   For this purpose, we generated synthetic interaction data using the ``NorTA''~\cite{Kurtz2015} and ``seqtime''~\cite{faustSignaturesEcologicalProcesses2018} methods (see Methods).
-  For each method, an \ac{otu} counts table is generated based on the selected interaction network topology and properties.
-  This counts table is used as the input to the \ac{micone} pipeline to generate predicted associations.
-  The interaction network used to generate the counts table is used as the source of true interactions to calculate the precision (Figure \ref{fig:figure5}) and sensitivity (Figure \ref{fig:figure_s5} and Figure \ref{fig:figure_s6}) for each network inference algorithm.
+  For each method, an \ac{otu} counts table was generated based on the selected parameters and abundance distributions.
+  This counts table was used as the input to the \ac{micone} pipeline to generate predicted associations.
+  The interaction network used to generate the counts table was used as the source of true interactions to calculate the precision (Figure \ref{fig:figure5}) and sensitivity (Figure \ref{fig:figure_s5} and Figure \ref{fig:figure_s6}) for each network inference algorithm.
   As shown in Figure \ref{fig:figure5} the consensus algorithm, especially the scaled-sum method, captures true associations with high precision (through the removal of edges that are either not present in most of the inference methods or whose association strength is low across methods).
@@
   However, if higher precision is required $p>0.5$ can be considered.
-  Therefore, the consensus networks provide the means to obtain a short list of associations that would have a high probability of being present in the real association network.
+  Therefore, the consensus networks provide the means to obtain a short list of associations that would have a high likelihood of being present in the real association network.
 
@@
   Figure \ref{fig:figure6}A, shows the percentage of total variation among the co-occurrence networks due to the different steps of the pipeline.
-  The \ac{ta} step, or more specifically the choice of 16S reference database, contributes the most ($\sim65.4\%$) to the variation in the networks, followed by the \ac{op} step ($\sim26.8\%$).
+  The \ac{ta} step, or more specifically the choice of 16S reference database, contributes the most ($65.4\%$) to the variation in the networks, followed by the \ac{op} step ($26.8\%$).
   This result highlights the importance of the taxonomy assignment step in the 16S data analysis workflow, implying that a change in the reference database will result in drastically different inferred networks.
@@
 
-  The effects of the different steps of the pipeline on the inferred networks can be best visualized through dimensionality reduction.
-  The PCA in Figure \ref{fig:figure6}B shows all the above networks, colored by the databases used in TA (left) and by the NI approach (right).
+  The effects of the different steps of the pipeline on the inferred networks can be visualized through dimensionality reduction.
+  The PCA in Figure \ref{fig:figure6}B shows all the above networks, colored by the tools used in the DC, TA, OP, and NI steps in each subfigure.
   The major effect of the TA step choice, shown before in Figure \ref{fig:figure6}A, is also reflected in the PCA plot, where networks segregate based on the database used (Figure~\ref{fig:figure6}B and Figure~\ref{fig:figure_s1}).
-  It is also evident that some networks, especially those inferred using the direct association network inference methods, are much closer in the PCA plot regardless of the reference database used.
-  Additionally, the variation (on the first two principal axes) between the networks decreases when the low abundance \ac{otu}s are removed from the network.
-  These results suggest that the most important criterion for accurate comparative analysis of co-occurrence networks is the taxonomy reference database followed by the level of filtering of the taxonomy tables.
+  Additionally, the plot also shows that the variation between the networks decreases when the low abundance \ac{otu}s are removed from the network.
+  It is also evident that in the NI step, some networks, especially those inferred using the direct association network inference methods, are much closer in the PCA plot regardless of the reference database used.
+  These results suggest that the most important criterion for accurate comparative analysis of co-occurrence networks is the taxonomy reference database followed by the level of filtering of the taxonomy tables and the network inference algorithm used.
 
@@
 
-  Figure~\ref{fig:figure7} shows the co-occurrence networks inferred for the healthy subjects (control) and subjects with autism specific disorder (ASD) in the fecal microbiome transplant study~\cite{Kang2017} (using the default tools and parameters from Table~\ref{tab:micone_tools}).
+  Figure~\ref{fig:figure7} shows the co-occurrence networks inferred for the healthy subjects (control) and subjects with autism specific disorder (ASD) in the fecal microbiome transplant study~\cite{Kang2017} (constructed using the default tools and parameters from Table~\ref{tab:micone_tools}).
   This figure demonstrates a typical use case of comparative analysis of networks using the \ac{micone} pipeline.
@@
   The analysis of the rewiring of associations in the ASD samples with respect to the control provides a guide for the identification of key genera that could be linked to dysbiosis.
-  We observed 20 unique links in the network for control samples, 12 unique links in the network for ASD subjects, and 7 edges in common between both networks.
-  Although these unique connections do not imply real interactions, they can still serve as potential starting points for literature surveys and further experimental confirmations.
-  For example, the changed connectivity of Prevotella and Porphyromonas in our networks show that these genera could be useful candidates for further downstream verification.
-  Correspondingly, we have found previous studies that indicate that these genera are implicated in ASD~\cite{Kang2017,hoGutMicrobiotaChanges2020} and cognitive impairment~\cite{chiPorphyromonasGingivalisInducedCognitive2021}.
+  We observed 22 unique links in the network for control samples, 12 unique links in the network for ASD subjects, and 7 edges in common between the two networks.
+  Although these unique associations do not imply actual interactions, they can still serve as potential starting points for literature surveys and further experimental exploration of mechanistic processes underlying dysbiosis.
+  For example, \textit{Prevotella} and \textit{Porphyromonas}, genera previously implicated in ASD~\cite{Kang2017,hoGutMicrobiotaChanges2020} and cognitive impairment~\cite{chiPorphyromonasGingivalisInducedCognitive2021} display modified connectivity in our network, suggesting that the observed associations may be relevant for understanding the role of these bacteria in disease.
   Additional visualization and comparison of networks can be performed using the \acl{mind}~\cite{huResourceComparisonIntegration2022}.

### 2022-09-30 sections/acknowledgements.tex
Update on Overleaf.

%!TEX root = ../main.tex

\section*{Acknowledgments}

We are grateful to members of the Segrè lab for helpful discussions and for feedback on the manuscript. This work was partially funded by grants from the National Institutes of Health (National Institute of General Medical Sciences, award R01GM121950; National Institute of Dental and Craniofacial Research, award number R01DE024468; National Institute on Aging, award number UH2AG064704; and National Cancer Institute, grant number R21CA260382), the U.S. Department of Energy, Office of Science, Office of Biological \& Environmental Research through the Microbial Community Analysis and Functional Evaluation in Soils SFA Program (m-CAFEs) under contract number DE-AC02-05CH11231 to Lawrence Berkeley National Laboratory, the National Science Foundation (grants 1457695, NSFOCE-BSF 1635070 and the NSF Center for Chemical Currencies of a Microbial Planet) and the Human Frontiers Science Program (RGP0020/2016 and RGP0060/2021).
DK acknowledges support by the Kilachand Multicellular Design Program graduate fellowship.
KSK was supported by Simons Foundation Grant \#409704, by the Research Corporation for Science Advancement through Cottrell Scholar Award \#24010, by the Scialog grant \#26119, and by the Gordon and Betty Moore Foundation grant \#6790.08.

\section*{Contributions}
Designed the research project: DK, KK, DS, ZH, CDL. Performed analysis: DK, GB. Wrote the first draft of the manuscript: DK. Revised and wrote the final version of the manuscript: DK, DS, KK.

### 2022-09-30 sections/methods.tex
Change: Move tables and fix headers
1. Tables now occupy their own pages at the end (without page numbers)
   we just have to split those pages
2. Updated the preamble such that there are no headers anymore
Update on Overleaf.

--- 2022-09-27 sections/methods.tex
@@
   \vspace{-5mm}
-  This study utilized two kinds of 16S rRNA sequencing datasets: biological datasets and mock datasets.
+  This study utilized two types of 16S rRNA sequencing datasets: biological datasets and mock/synthetic datasets.
   Biological datasets are collections of sequencing reads obtained from naturally occurring microbial community samples.
-  The current analysis used stool samples from a fecal microbiome transplant study of autism~\cite{Kang2017} as biological datasets.
-  This dataset was chosen because the sequences were easily accessible on Qiita~\cite{qiita} and optimally pre-processed according to the \ac{emp}~\cite{Thompson2017} protocol and can be used directly as input to the \ac{micone} pipeline.
+  The current analysis used stool samples from a fecal microbiome transplant study of autism~\cite{Kang2017} as the biological dataset.
+  This dataset was chosen because the sequences were easily accessible on Qiita~\cite{qiita} and optimally pre-processed according to the \ac{emp}~\cite{Thompson2017} protocol, allowing them to be used directly as input to the \ac{micone} pipeline.
   The study was composed of multiple sequencing runs.
   The runs that contained paired-end reads (run 2 (10M reads), run 3 (750K reads) and run 4 (16M reads)), were downloaded from Qiita~\cite{qiita} (study ID 10532) and used as input sequences for the \ac{micone} pipeline.
-  Sequences from both control (212 samples including neurotypical and donors) and autistic (126 samples) patients were included in the analyses, but the groups were analyzed independently in the pipeline after the \ac{op} step.
-  All the network analyses in the study, unless explicitly mentioned, are performed indeed on the healthy and autistic samples in the fecal microbiome transplant study.
-  However, all other analyses (preceding the NI step) are performed together on all the samples in the fecal microbiome transplant study.
-  The mock community 16S datasets are experimental sequencing data obtained for artificially assembled collections of species in known proportions.
+  Sequences from both control (212 samples including neurotypical and donors) and autism spectrum disorder (ASD) (126 samples) patients were included in the analyses.
+  All the network analyses in the study, unless explicitly mentioned, were performed on the healthy and ASD samples in the fecal microbiome transplant study.
+  The mock community 16S datasets are experimental sequencing data obtained for artificially assembled collections of DNA of species in known proportions.
   The mock datasets used for this study, obtained from mockrobiota~\cite{Bokulich2016}, are labeled mock4, mock12, and mock16.
@@
   The configuration file along with the run file (\texttt{main.nf}) lists the inputs, output, and the steps to be performed during runtime, along with the parameters to be used (if different from defaults) for the various steps.
+  The default settings of the pipeline are shown in Table~\ref{tab:micone_tools} (with default parameter values shown in Tables S2-S6 at \href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper}).
   Since the entire pipeline run is stored in the form of a text file (the configuration file), subsequent runs are highly reproducible and changes can be easily tracked using version control.
-  It makes use of the nextflow workflow manager~\cite{Tommaso2015} under the hood, making it readily usable on the local machine, cluster, or cloud with minimal configuration change.
+  The pipeline makes use of the nextflow workflow manager~\cite{Tommaso2015} under the hood, making it readily usable on the local machine, cluster, or cloud with minimal configuration change.
   It also allows for automatic parallelization of all possible processes, both within and across samples.
-  The pipeline is designed to be modular: each tool or method is organized into modules which can be easily modified or replaced.
-  This modular architecture simplifies the process of adding new tools (refer to modules section in the \ac{micone} documentation).
-  % TODO: Write a DockerFile for the package
-  % In addition to the Python package, the entire pipeline has been containerized into a Docker~\cite{Merkel1994} image (\hl{dockerhub link}) for easy deployment and setup.
+  The pipeline is designed to be modular: each tool or method is organized into modules that can be easily modified or replaced.
+  This modular architecture simplifies the process of adding new tools (refer to the modules section in the \ac{micone} documentation).
   The main components of the pipeline are detailed in the subsequent sections.
@@
   The demultiplexing process deals with separating the multiplexed sequences into individual samples based on barcodes.
-  The trimming process handles the quality control steps such as trimming adapters and trimming low-quality nucleotide stretches from the sequences.
+  The trimming process handles the quality control steps such as trimming adapters and low-quality nucleotide stretches from the sequences.
   The parameters and tools in this process are fixed and are not available for user customization.
@@
   The list of tools used in this step, along with their modules and references are provided in Table~\ref{tab:micone_tools}.
-  In all UniFrac comparisons, we set a count threshold of 10, such that if the count of the representative sequences in a particular sample is less than the threshold, it is omitted from the analysis.
-  The expected sequences from the mock communities are trimmed to the V4 region before being subject to UniFrac analyses.
+
+  For the UniFrac analysis in Figure~\ref{fig:figure2}, we had set a count threshold of 10, such that if the count of the representative sequences in a particular sample is less than the threshold, it is omitted from the analysis.
+  Additionally, for Figure~\ref{fig:figure2}C and \ref{fig:figure2}D, the expected sequences from the mock communities were trimmed to the V4 region before being subject to UniFrac analyses.
 
@@
   \vspace{-5mm}
-  This module deals with assigning taxonomies to either the representative sequences of the \ac{otu}s or the \ac{esv}s.
-  In order to assign taxonomies to a particular sequence a taxonomy database and a query tool are necessary.
-  The taxonomy database contains the collection of 16S sequences of microorganisms of interest and the query tool allows one to compare a sequence of interest to all the sequences in the database to identify the best matches.
+  This module deals with assigning taxonomies to the representative sequences (\ac{otu}s or \ac{esv}s).
+  In order to assign taxonomies to a particular sequence, a taxonomy database and a query tool are necessary.
+  The taxonomy database contains a collection of 16S sequences of microorganisms and the query tool allows one to compare a sequence of interest to all the sequences in the database to identify the best matches.
   Finally, a consensus method is used to identify the most probable match from the list of best matches.
@@
   During the assignment, a representative sequence might be assigned an "unknown" Genus for one of two reasons: the first is if the taxonomy identifier associated with the sequence in the database did not contain a given Genus; the second, more likely reason, is that the database contains multiple sequences that are very similar to the query (representative) sequence and the consensus algorithm (from \ac{qiime2}) is unable to assign one particular Genus at the required confidence.
-  The assignment between SILVA and the other two databases were originally significantly different ($40\%$ mismatch) even at the Phylum level.
-  However, this was fixed by making minor adjustments to the taxonomic names, such as changing Bacteroidota to Bacteroidetes in the SILVA Phylum assignments.
+  The assignments in SILVA were originally substantially different from the other two databases ($40\%$ mismatch) even at the Phylum level.
+  However, this was corrected via minor adjustments to the taxonomic names, such as changing Bacteroidota to Bacteroidetes in the SILVA Phylum assignments.
   The full list of changes can be found in \texttt{figure4ab\_data.py} in the data and scripts  repository.
@@
   \vspace{-5mm}
-  This module deals with normalization, filtering, forking, grouping and applying transformations to the \ac{otu} or \ac{esv} counts matrix.
+  This module deals with normalization, filtering, forking, grouping, and applying transformations to the \ac{otu} or \ac{esv} counts matrix.
   Normalization of the count matrix involves converting the count matrix of read counts into a count matrix containing relative abundances.
   The module also supports rarefaction, which is a normalization technique used to overcome the bias that might arise due to variable sampling depth in different samples.
+  This is performed either by sub-sampling or by normalization of the matrix to the lowest sampling depth~\cite{Weiss2015}.
   However, due to the potential biases and false positives~\cite{mcmurdieWasteNotWant2014,chaoCoveragebasedRarefactionExtrapolation2012} that might arise during the process, the rarefaction module is disabled by default and can be enabled in the configuration if needed.
-  This is performed either by sub-sampling or by normalization of the matrix to the lowest sampling depth~\cite{Weiss2015}.
-  Although the pipeline supports normalization and rarefaction, these modules are turned off by default and the analysis reported in the paper do not normalize or rareify the counts matrices.
+  Hence, although the pipeline supports normalization and rarefaction, these modules are turned off by default, and the analyses reported in the paper do not normalize or rareify the counts matrices.
   However, most of the network inference methods perform normalization and other transformation operations on the counts matrix as a part of their workflow.
@@
   By default, when the OP module is ``on'', the samples are filtered out if the total reads in a sample are less than 500 and features are filtered out if the relative abundance is less than 1\%, prevalence (percentage of samples containing feature) is less than 5\% and count sum across all the samples is less than 100.
-  When the OP module is ``off'', the filtering is still performed but only to remove those features that are very low in presence.
+  When the OP module is ``off'', the filtering is still performed but threshold parameters are much more relaxed.
   The parameters used are given in Table~\ref{tab:micone_tools}.
@@
   The group operation transforms the \ac{otu} or \ac{esv} count matrix into a taxonomic count matrix at the requested level by adding up counts that map to the same taxonomy and is carried out at the end of the OP step.
-  Finally, transformations are performed in order to correct for and overcome the compositional bias that is inherent in the counts matrix (in the analysis performed in the study these are handled by the network inference algorithm).
+  Finally, transformations are performed in order to correct for and overcome the compositional bias that is inherent in the counts matrix (in the analysis performed in the study these were disabled and directly handled by the network inference algorithm).
   All the modules in this step were implemented using functions from the \href{http://biom-format.org/}{biom-format} Python package~\cite{mcdonaldBiologicalObservationMatrix2012}.
@@
   These collapsed matrices are used as input to the network inference methods to produce association matrices at the appropriate taxonomy level.
-  These associations can be represented as a network, with nodes representing taxonomies of the microorganisms and edges representing the associations between them.
-  A threshold of 0.3 was set for the correlation-based methods (\ac{sparcc}, Propr, Spearman and Pearson) and a threshold of 0.01 was set for the other direct association methods.
-  After filtering out edges based on the threshold, the isolated nodes are removed.
+  These associations can be represented as a network, with nodes representing the taxonomies of the microorganisms and edges representing the associations between them.
 
-  A null model is created by re-sampling and bootstrapping the counts matrix and recalculating the correlations (see next section for details on network analysis and statistics).
-  These bootstrapped association matrices are used to calculate the significance of the inferred correlations by calculating the p-values against this null model~\cite{Watts2018}.
   The pipeline includes 4 methods for pairwise correlation metrics, and 6 methods for direct association metrics (refer to Table~\ref{tab:micone_tools}).
@@
   On the other hand, direct association methods use conditional independence to avoid the detection of correlated but indirectly connected taxonomic entities~\cite{Kurtz2015,Menon2018}.
+  A null model is created by re-sampling and bootstrapping the counts matrix and recalculating the correlations (see next section for details on network analysis and statistics).
+  These bootstrapped association matrices are used to calculate the significance of the inferred correlations by calculating the p-values against this null model~\cite{Watts2018}.
   Brown's p-value merging method~\cite{brown_400_1975} is used for combining p-values from the pairwise correlations methods to obtain a consensus p-value, which can be used to filter for significance.
@@
  This section defines the notation used below to describe the consensus network algorithm used in the \ac{micone} pipeline.
- Note that all networks to be compared have the same number of nodes.
+ Note that all networks to be compared were updated to have the same number of nodes.
 
-  $w$, the number of co-occurrence networks to be integrated into the consensus network (corresponding to the number of the different approaches used to infer co-occurrence networks excluding Spearman and Pearson, $k$)
+  $w$, the number of co-occurrence networks to be integrated into the consensus network (by default, is equal to the total number of network inference methods excluding Spearman and Pearson, 8)
 
@@
   $P^i$, the matrix of p-values for all edges of the $i^{th}$ co-occurrence network.
-  This is a $q \times q$ matrix, where $i \in \{1,\dots,q\}$
+  This is a $q \times q$ matrix, where $i \in \{1,\dots,w\}$
 
@@
   For all correlation-based methods $k \le w$, $1000$ permutations of the original \ac{otu} counts data were generated~\cite{Watts2018}.
-  All the direct association-based methods used in the study have their own regularization methods built-in and hence do not need to undergo bootstrapping.
   The correlations in the permuted \ac{otu} tables are recalculated using the different correlation-based algorithms.
   Finally, the p-value is determined based on how often a more extreme association is observed for randomly permuted data.
+  Note that, all the direct association-based methods used in the study have their own regularization methods built in and hence do not need to undergo bootstrapping.
 
@@
   \vspace{-5mm}
-  The first part of the consensus algorithm workflow is to merge the p-values for the networks generated by the correlation-based methods.
+  The next step in the consensus algorithm workflow is to merge the p-values for the networks generated by the correlation-based methods.
   This step is performed using the Brown's p-value merging method~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016,faustCoNetAppInference2016}.
@@
 
-  \subsubsection*{Consensus method}
+  \subsubsection*{Consensus methods}
   \vspace{-5mm}
@@
   Both the algorithms make use of a user-defined parameter $p$ ($0 \leq p \leq 1$), in order to threshold the edges from the individual methods.
-  The inputs to both the algorithms are the co-occurrence networks (association matrices) $\bar{N}^i$ (flattened version of $N^{i}$) generated by each method $i$ and the threshold parameter $p$.
+  The inputs to both the algorithms are the co-occurrence networks (association matrices) $\bar{N}^i$ (flattened version of $N^{i}$) generated by each method $i$, and the threshold parameter $p$.
   Here, the $\bar{N}^{i}$ each have the same set of nodes $q$ and only differ by the value of the association inferred between every pair.
-  Networks which do not have a particular node, are updated such that the node is added as an isolated component.
+  Networks that do not have a particular node, are updated such that the node is added as an isolated component.
   In this manner, $\bar{N}^{i}_j$ represents edge $j$ in network $i$.
 
-    Note that this method is only used to filter relevant interactions.
-    If a given pair of nodes is inferred to have edges that satisfy the above relationships, all corresponding edges from the $w$ networks will be returned by the algorithm, as a multigraph.
+    Note that the consensus method is only used to filter relevant interactions.
+    If a given pair of nodes is inferred to have edges that satisfy the consensus requirements, all corresponding edges from the $w$ networks will be returned by the algorithm, as a multigraph.
     Based on this approach, \ac{micone} reports as the default output, the consensus network where each edge is annotated with weights (correlations for the correlation-based methods and direct associations for the other methods) from all the methods used in the consensus algorithm.
@@
   For each pair of nodes, we determine the number of network inference methods that report an edge $j$ between them, i.e. $\bar{N}^{i}_{j}, \forall i \in \{1,\dots,w\}$.
-  Each node-pair will have an edge in the final consensus network if the number of reported edges is larger than a user-defined threshold (Equation~\ref{eqn:simple-voting2}).
+  Each node-pair will have an edge in the final consensus network if the number of reported edges is larger than the threshold (Equation~\ref{eqn:simple-voting2}).
 
@@
 
-  This algorithm generates a consensus network based on the sum of all edges reported between a pair of nodes~\cite{bustinceFuzzySetsTheir2008,tsarevApplicationMajorityVoting2018}.
+  This algorithm generates a consensus network based on the sum of all edges (weights of associations) reported between a pair of nodes~\cite{bustinceFuzzySetsTheir2008,tsarevApplicationMajorityVoting2018}.
   Since in generating a consensus network using this method we sum the edges reported by direct association methods with those from correlation-based methods, summing of the edges is preceded by a pre-processing step, in which all networks are re-scaled.
@@
 
-    In this way, it is guaranteed that $\max_{j}(\mid \bar{S}^i_j \mid) = 1$.
+    In this way, it is guaranteed that $\max(\mid \bar{S}^i \mid) = 1$.
 
@@
   \vspace{-5mm}
-  This section defines notation for the network variability analysis performed for Figure~\ref{fig:figure6}.
+  This section defines the notation used for the network variability analysis performed for Figure~\ref{fig:figure6}.
 
@@
 
-  $\bar{N}_i$, the ``flattened'' version of the adjacency matrix $N^i$ into a $Q^2 \times 1$ column vector, where all columns are stacked onto each other into a $Q^2$ long vector.
+  $\bar{N}^i$, the ``flattened'' version of the adjacency matrix $N^i$ into a $Q^2 \times 1$ column vector, where all columns are stacked onto each other into a $Q^2$ long vector.
 
@@
    In order to compare across different networks and calculate the degree of variability induced by the choice of different modules, we organized multiple networks into a single mathematical structure that we could use for linear regression.
-   First, we obtained the co-occurrence network $\bar{N}_i$ for each of the $W$ possible tool and parameter combinations in the workflow.
+   First, we obtained the co-occurrence network $\bar{N}^i$ for each of the $W$ possible tool and parameter combinations in the workflow.
    We then constructed a matrix $\mathbf{\bar{N}}$ whose $i^{th}$ column is the flattened version of the $i^{th}$ network, i.e. the column vector $\bar{N}^i$.
-   Therefore, $\bar{N}^i_j$ is the weight of edge $j$ in network $i$.
+   Therefore, $\bar{N}^i_j$ is the weight of edge $j$ in the network $i$.
    $\bar{N}^i_j$ is assigned a value of 0 if edge $j$ did not exist in network $i$ but was present in one of the other networks.
@@
 
-  To infer the variability contributed due to the different steps in the pipeline we performed a linear regression on each edge in $\mathbf{\bar{N}}$ and a subsequent ANOVA to extract the within group variances.
+  To infer the variability contributed due to the different steps in the pipeline we can perform a linear regression on each edge in $\mathbf{\bar{N}}$ and a subsequent ANOVA to extract the within-group variances.
   One issue with this approach is that the possibility of correlations existing between the edges of the network could lead to inaccurate estimates of the variance if a linear model were used to directly model the relationships between edges and steps in the workflow.
@@
 
-  In particular, we infer parameters $\alpha_i$ such that:
+  In particular, we infer parameters $\alpha_j$ such that:
   \begin{multline}
-  C_j = &\sum_{i=1}^5 \left( \alpha^{DC(i)}_j.\delta^{DC(i)}_j \right) +
-         \sum_{i=1}^2 \left( \alpha^{CC(i)}_j.\delta^{CC(i)}_j \right) +
-         \sum_{i=1}^3 \left( \alpha^{TA(i)}_j.\delta^{TA(i)}_j \right) +
-         \sum_{i=1}^2 \left( \alpha^{OP(i)}_j.\delta^{OP(i)}_j \right) + \\
-         &\sum_{i=1}^{10} \left( \alpha^{NI(i)}_j.\delta^{NI(i)}_j \right) +
+  C_j = &\sum_{i=1}^5 \left( \alpha^{DC(i)}_j \delta^{DC(i)}_j \right) +
+         \sum_{i=1}^2 \left( \alpha^{CC(i)}_j \delta^{CC(i)}_j \right) +
+         \sum_{i=1}^3 \left( \alpha^{TA(i)}_j \delta^{TA(i)}_j \right) +
+         \sum_{i=1}^2 \left( \alpha^{OP(i)}_j \delta^{OP(i)}_j \right) + \\
+         &\sum_{i=1}^{10} \left( \alpha^{NI(i)}_j \delta^{NI(i)}_j \right) +
          \epsilon_j
@@
 
-   where, $\alpha_i$ are the coefficients of the regression, $\epsilon_i$ are the residuals and $\delta_i$ are the indicator variables that correspond to the processes utilized in the pipeline used to create the network $N_i$; for example, $\delta^{DC(1)}_i = 1$ if the DC(1) process was used in the generation of the network $N_i$.
+   where, $\alpha_i$ are the coefficients of the regression, $\epsilon_i$ are the residuals and $\delta_i$ are the indicator variables that correspond to the processes utilized in the pipeline used to create the network $N_i$; for example, $\delta^{DC(1)}_i = 1$ if the DC(1) process was used in the generation of the network $N^i$.
 
@@
      \item $OP(i) \in$ \{Filter(on), Filter(off)\}
-     \item $NI(i) \in$ \{\ac{sparcc}, propr, Spearman, Pearson, \ac{spieceasi}, \ac{cozine}, \ac{harmonies}, \ac{spring}, \ac{mldm}, FlashWeave\}
+     \item $NI(i) \in$ \{\acs{sparcc}, propr, Spearman, Pearson, \acs{spieceasi}, \acs{cozine}, \acs{harmonies}, \acs{spring}, \acs{mldm}, FlashWeave\}
    \end{enumerate}
 
-   The variance contributed by each step of the pipeline is calculated for every component in $\mathbf{C}$ matrix through ANOVA using the Python statsmodels~\cite{seaboldStatsmodelsEconometricStatistical2010} package and is shown in Figure~\ref{fig:figure6}A.
-  The total variance for the network is calculated by adding the variances for each connection.
-  The merged network table $\mathbf{\bar{N}}$ is used as the input to the PCA analysis to generate Figure~\ref{fig:figure6}B.
+   The variance contributed by each step of the pipeline was calculated for every component in $\mathbf{C}$ matrix through ANOVA using the Python statsmodels~\cite{seaboldStatsmodelsEconometricStatistical2010} package and is shown in Figure~\ref{fig:figure6}A.
+  The total variance for the network was calculated by adding the variances for each connection and normalizing with the degrees of freedom.
+  The merged network table $\mathbf{\bar{N}}$ was used as the input to the PCA analysis to generate Figure~\ref{fig:figure6}B.
 
@@
 
-  The first method, hereby referred to as ``seqtime''~\cite{faustSignaturesEcologicalProcesses2018}, used generalized Lotka-Volterra (gLV) equations to model the microbial community dynamics and utilized the Klemm–Eguı́luz algorithm to generate a clique-based interaction network~\cite{Rottjers2018}.
+  The first method, ``seqtime''~\cite{faustSignaturesEcologicalProcesses2018}, used generalized Lotka-Volterra (gLV) equations to model the microbial community dynamics and utilized the Klemm–Eguı́luz algorithm to generate a clique-based interaction network~\cite{Rottjers2018}.
   We used the \href{https://github.com/hallucigenia-sparsa/seqtime}{seqtime} R package to simulate communities with number of species ($N$) varying from 10 to 150 (10, 25, 50, 100, 150 and 200).
@@
 
-  The second method, hereby referred to as ``NorTA'', used the Normal to Anything (NorTA) approach coupled with a given interaction network topology to generate the abundance distribution of the microbial community~\cite{Kurtz2015}.
+  The second method, ``NorTA'', used the Normal to Anything (NorTA) approach coupled with a given interaction network topology to generate the abundance distribution of the microbial community~\cite{Kurtz2015}.
   We used the \href{https://github.com/zdk123/SpiecEasi}{spieceasi} R package~\cite{Kurtz2015} to simulate communities with different network topologies (scale-free, cluster, block, Erdos-Renyi, band and hub) and target abundance distributions (Negative Binomial, Poisson, Zero-Inflated Negative Binomial).
-  The OTU table was generated using the American Gut Project example in the spieceasi package (amgut1.filt) with the default parameter options.
+  The OTU table was generated using the American Gut Project example in the spieceasi package (\texttt{amgut1.filt}) with the default parameter options.
 
@@
   The interaction matrix was used as the source of expected (true) interactions and the associations predicted using \ac{micone} were the source of predicted interactions.
-  Finally, for each dataset we evaluated the precision of the associations predicted by the individual network inference methods as well as the consensus (Figure \ref{fig:figure5}).
+  Finally, for each dataset we evaluated the precision and sensitivity of the associations predicted by the individual network inference methods as well as the consensus (Figures \ref{fig:figure5}, \ref{fig:figure_s5}, and \ref{fig:figure_s6}).
 
@@
   \vspace{-5mm}
-  In order to compare the representative sequences generated by the various methods in the DC step, we employ both the weighted~\cite{Lozupone2007} (Figure~\ref{fig:figure2}A) and unweighted UniFrac method~\cite{Lozupone2005} (Figure~\ref{fig:figure2}B).
+  In order to compare the representative sequences generated by the various methods in the DC step, we employed both the weighted~\cite{Lozupone2007} (Figure~\ref{fig:figure2}A) and unweighted UniFrac method~\cite{Lozupone2005} (Figure~\ref{fig:figure2}B).
   The UniFrac distance metric (unique fraction metric) is a beta-diversity measure that computes the distance between two sets of taxa as the fraction of the branch length of the tree that leads to descendants from either one environment or the other, but not both~\cite{Lozupone2005}.
   The weighted UniFrac distance metric takes into account the abundances of the representative sequences when calculating shared and unshared branch lengths, whereas the unweighted UniFrac distance metric does not and hence gives equal weights to each sequence.
-  In Figure~\ref{fig:figure2} the distances between methods are the average distance calculated between the reference sequence distribution for a pair of methods generated for every corresponding sample in the dataset.
+  In Figure~\ref{fig:figure2} the distances between methods are the distance between the reference sequence distribution for a pair of methods averaged over every sample in the dataset.
   All UniFrac calculations were performed using the \texttt{scikit-bio}~\cite{thescikit-biodevelopmentteamScikitbioBioinformaticsLibrary2022} v0.5.6 Python package.
@@
   \vspace{-5mm}
-  In Figure~\ref{fig:figure3}C we use the Bray-Curtis distance metric to calculate the distance between the predicted (using the taxonomy databases in the TA step) and expected taxonomic distribution.
+  In Figure~\ref{fig:figure3}C, we used the Bray-Curtis distance metric to calculate the distance between the predicted (using the taxonomy databases in the TA step) and expected taxonomic distribution.
   The Bray-Curtis distance is used to quantify the compositional dissimilarity between two different taxonomic distributions defined by vectors $u$ and $v$.
   It is defined as:
-  $$\sum{|u_i-v_i|} / \sum{|u_i+v_i|}$$
+  $$d = \frac{\sum_i{|u_i-v_i|}}{\sum_i{|u_i+v_i|}}$$
   The Bray-Curtis distance calculations were performed using the \texttt{scipy}~\cite{virtanenSciPyFundamentalAlgorithms2020} v1.8.0 Python package.
@@
   \vspace{-5mm}
-  In Figure~\ref{fig:figure5} we calculate the accuracy and robustness of the inferred association networks (using the various network inference algorithms and the consensus methods) against the original interaction network used to create the taxonomic distribution.
-  We use precision and sensitivity to calculate the accuracy of the predictions.
+  In Figure~\ref{fig:figure5} we evaluated the accuracy and robustness of the inferred association networks (using the various network inference algorithms and the consensus methods) against the original interaction network used to create the taxonomic distribution.
+  We used the following formulations of precision and sensitivity to calculate the accuracy of the predictions:
 
@@
   Sensitivity = $\frac{TP}{FN + TP}$
+
+  where, TP - true positives, FP - false positives and FN - false negatives
 

### 2022-09-30 sections/supplementary.tex
Update on Overleaf.

@@
     The data containing the 16S sequencing reads (V4 region) was downloaded from Qiita~\cite{qiita} (study ID: 10532).
-    Only runs 2, 3, and 4 were used for the subsequent analysis as these runs consisted of paired-end sequencing data and run 1 contained single-end data.
-    The sample metadata was updated to contain only BMI, sex, height, weight and experimental group.
-    This was necessary as two of the network inference algorithms (\ac{mldm} and FlashWeave) required information about environmental heterogeneity.
+    Only runs 2, 3, and 4 were used for the subsequent analysis as these runs consisted of paired-end sequencing data, and run 1 contained single-end data.
+    The sample metadata was updated to contain only BMI, sex, height, weight, and experimental group.
+    This was necessary as two of the network inference algorithms (\acs{mldm} and FlashWeave) required information about environmental heterogeneity.
     However, these environmental correlations were not included in the current analyses.
@@
     The configuration files (main.nf and nextflow.config) used to run the \ac{micone} pipeline as well the details of the pipeline execution (dag, report, timeline and trace) are in the "runs/FMT" directory of the data and scripts repository (\href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper})
-    % TODO: Store pipeline results
     The results of the pipeline execution for reproducing the analyses in the manuscript are stored on Zenodo.
@@
     The configuration files (main.nf and nextflow.config) used to run the \ac{micone} pipeline as well the details of the pipeline execution (dag, report, timeline and trace) are in the "runs/mock*" directory of the data and scripts repository (\href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper})
-    % TODO: Store pipeline results
     The results of the pipeline execution for reproducing the analyses in the manuscript are stored on Zenodo.
 
-% TODO: Update this section and link
     \subsubsection*{Interpretation of Unifrac results in the DC step}
-  In order to verify this claim, for each of these methods, we selected the top 1000 representative sequences for each method and recalculated the distance metrics (Figure \ref{fig:figure_s3}).
-  We observe that both the weighted and unweighted UniFrac distances are increased, implying that the top representative sequences generated by the different methods are not as similar to each other.
+  In Figure \ref{fig:figure_s3}, we observe that both the weighted and unweighted UniFrac distances are increased for the top 1000 representative sequences, implying that the top representative sequences generated by the different methods are not as similar to each other.
   Therefore, since the weighted UniFrac distances are lower than the unweighted distances, we conclude that the representative sequences in the middle range of the abundance distribution are those that must be the most similar between the methods.
 
-    Open-reference and de novo clustering methods perform the best under the weighted UniFrac metric and the worst (marginally) under the unweighted UniFrac metric.
+    Open-reference and de novo clustering methods perform the best under the weighted UniFrac metric and the worst (marginally) under the unweighted UniFrac metric (Figure~\ref{fig:figure2}C and \ref{fig:figure2}D).
     This result can be attributed to the large number of low abundance representative sequences that are generated by these methods.
@@
     The configuration files (main.nf and nextflow.config) used to run the \ac{micone} pipeline as well the details of the pipeline execution (dag, report, timeline and trace) are in the "runs/norta" and "runs/seqtime" directories of the data and scripts repository (\href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper})
-    % TODO: Store pipeline results
     The results of the pipeline execution for reproducing the analyses in the manuscript are stored on Zenodo.
@@
     \item The average shortest path length describes the average of all the shortest paths in the graph.
-No number is reported if the graph is not connected, therefore, the results indicate that none of the networks that make use of \ac{harmonies}, \ac{cozine}, \ac{spring}, \ac{spieceasi} and Pearson are connected.
+No number is reported if the graph is not connected, therefore, the results indicate that none of the networks that make use of \acs{harmonies}, \acs{cozine}, \acs{spring}, \acs{spieceasi} and Pearson are connected.
     \item The average clustering is the average clustering coefficient of the graph.
 The closer the value is to 1.0, the more densely connected is the graph.
-We can observe that the networks that use correlation-based methods have the highest values while the direct-association based methods have the lowest.
-    \item The number of connected components is the highest for the direct-association based methods and the lowest for the correlation-based methods.
+We can observe that the networks that use correlation-based methods have the highest values while the direct association based methods have the lowest.
+    \item The number of connected components is the highest for the direct association based methods and the lowest for the correlation-based methods.
 In the case of propr, all the networks have only one giant component.
@@
 Positive values imply that there are more edges between vertices of the same type than we would expect by chance, and negative implies that there are less.
-The networks inferred by \ac{mldm} report very few edges, and skew the average modularity scores.
-This could also be an artifact of incomplete converge of the \ac{mldm} algorithm for some combinations.
+The networks inferred by \acs{mldm} report very few edges, and skew the average modularity scores.
+This could also be an artifact of incomplete convergence of the \acs{mldm} algorithm for some combinations.
     \item Node connectivity refers to the minimum number of nodes that must be removed from the graph to make it disconnected.
@@
     \item Degree assortativity coefficient measures the similarity of connections in the graph with respect to the node degree.
-Again we observe that the direct-association based methods have negative degree assortativity, meaning that there are many hubs in these networks.
+Again we observe that the direct association based methods have a negative degree of assortativity, meaning that there are many hubs in these networks.
 The correlation-based methods have positive values implying that in these networks nodes with similar degrees attach to one another.
   \end{enumerate}
+  A weight threshold of 0.1 and a p-value threshold of 0.05 were applied to each network before the analysis.
   All the metrics were calculated using the \texttt{networkx} Python package~\cite{hagbergExploringNetworkStructure2008}.
@@
 
-  We first calculate $E[\Psi]$ and $Var[\Psi]$ under the null hypothesis that the data are drawn from a multivariate Gaussian with some covariance matrix.
-  We then use these values to parametrize a $\chi^2$ distribution from which p-value corresponding to $\frac{\psi}{c}$ can be calculated.
-  Furthermore, Brown showed that $\mathrm{E}[\Psi]$ and $\mathrm{Var}[\Psi]$ can be calculated via a numerical approximation:
+  We can calculate $E[\Psi]$ and $Var[\Psi]$ under the null hypothesis that the data are drawn from a multivariate Gaussian with some covariance matrix.
+  We then use these values to parametrize a $\chi^2$ distribution from which the p-value corresponding to $\frac{\psi}{c}$ can be calculated.
+  Furthermore, Brown showed that $\mathrm{E}[\Psi]$ and $\mathrm{Var}[\Psi]$ can be calculated via the following numerical approximation:
   \begin{equation*}
@@
 
-  Using $E[\Psi]$ and $Var[\Psi]$ one then fits a $\chi^2$ distribution with the parameters $c$ and $f$.
+  Using $E[\Psi]$ and $Var[\Psi]$ we then fit a $\chi^2$ distribution with the parameters $c$ and $f$.
   Note that since, in general, $f$ will not be an integer, this should be understood as a Gamma distribution with a shape parameter $f$, as mentioned by Brown~\cite{brown_400_1975}.
-  Using this, one calculates the test $\psi$ and computes the p-value from the CDF of the $\chi^2$ distribution, given in Equation~\ref{eqn:suppl_pvalue-combined}.
+  Using this, we calculate the test $\psi$ and compute the p-value from the CDF of the $\chi^2$ distribution, given in Equation~\ref{eqn:suppl_pvalue-combined}.
   Therefore, the final combined p-value~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016} is then given by:
@@
   Note that we do not use Pearson and Spearman methods in the p-value merging step and these algorithms are only used for demonstration and comparison.
-  The combined p-values are used to threshold for significance the correlation-based networks during the consensus network step.
+  The combined p-values are used to threshold for significance in the correlation-based networks during the consensus network step.
 
@@
     The default format \ac{micone} uses for storing the network files is the JSON (JavaScript Object Notation) format.
-    The custom JSON schema we have designed is able to store all network related information pertaining to nodes, links and the metadata related to the links and datasets.
-    Additionally, \ac{micone} also supports exporting of networks into a variety of other formats such as edge lists, .gml and cytoscape formats.
-    Since, we make use of \texttt{networkx}~\cite{hagbergExploringNetworkStructure2008} for the export functionality, networks can be exported to all formats supported by the package.
-    However, not all the corresponding metadata will be exported appropriately, since most formats do not support this additional metadata.
+    The custom JSON schema we have designed is able to store all network-related information pertaining to nodes, links, and the metadata related to the links and datasets.
+    Additionally, \ac{micone} also supports exporting of networks into a variety of other formats such as edge lists, GML, and Cytoscape formats.
+    Since we make use of \texttt{networkx}~\cite{hagbergExploringNetworkStructure2008} for the export functionality, networks can be exported to all formats supported by the package.
+    However, not all the corresponding metadata will be exported appropriately, as most formats do not support this additional metadata.
     The details of the format and information about importing/exporting it and other network formats can be found in the \ac{micone} documentation.
@@
 
-  Additionally, it is worth pointing out some additional more specific conclusions stemming from the individual steps of our analysis.
-  The different denoising/clustering methods differ mostly in their identification of sequences that are in low abundances.
-  Hence, they do not have much of an impact on the inferred co-occurrence networks when the sequences of low abundance are removed (Figure~\ref{fig:figure_s1}).
+  It is worth pointing out some additional, more specific, conclusions stemming from the individual steps of our analysis.
+  The different denoising/clustering methods differ mostly in their identification of sequences that are in low abundance.
+  Hence, they do not have much of an impact on the inferred co-occurrence networks when the sequences of low abundance are removed (Figure~\ref{fig:figure1}).
   Comparison of inferred and expected reference sequences and their abundances in mock community datasets has allowed us to identify \ac{dada2} as the method which best recapitulates the expected sequence composition.
@@
   Some tools which were not as widely used (e.g.
-dbOTU3~\cite{Olesen2017}) as well as older popular methods like mothur~\cite{Schloss2009} have not been included in the study, but could be added into the pipelines in future updated analyses.
+dbOTU3~\cite{Olesen2017}) as well as older popular methods like mothur~\cite{Schloss2009} have not been included in the study but could be added into the pipelines in future updated analyses.
 
@@
   Additionally, a very large number of nodes also decreases the accuracy of the network inference algorithms for the same sample size~\cite{peschelNetCoMiNetworkConstruction2020} and increases the computational complexity~\cite{tackmannRapidInferenceDirect2019}.
-  We observe that filtering out taxa that are present in low abundances in all samples increases the proportion of taxa in common between taxonomy tables generated using different reference databases (Figure~\ref{fig:figure_s5}), providing another reason for filtering.
-  We also observe that the reduction in the number of taxa leads to better agreement in the networks inferred through different methods (Figure~\ref{fig:figure_s1}).
+  We observe that filtering out taxa that are present in low abundances in all samples increases the proportion of taxa in common between taxonomy tables generated using different reference databases (Figure~\ref{fig:figure_s4}), providing another reason for filtering.
+  We also observe that the reduction in the number of taxa leads to a better agreement in the networks inferred through different methods (Figure~\ref{fig:figure1}).
   Moreover, filtering is necessary in order to increase the power in tests of significance when the number of taxa is much greater than the number of samples.

### 2022-09-30 sections/tables_figures.tex
Update on Overleaf.

@@
       The tools highlighted in gray are the defaults for the pipeline that are recommended based on the benchmarks with the mock and synthetic datasets.
-      The consensus algorithm in the \ac{ni} step incorporates all the modules (bootstrapping, direct association and correlation-based) in that step to generate the consensus network.
+      The consensus algorithm in the Network Inference (\ac{ni}) step incorporates all the modules (bootstrapping, direct association, and correlation-based) to generate the consensus network.
     }
@@
       \textbf{The workflow of the \ac{micone} pipeline}.
-      The steps of the workflow can be broken down into five major groups: \textbf{(SP)} \textbf{S}equence \textbf{P}rocessig, \textbf{(DC)} \textbf{D}enoising and \textbf{C}lustering, \textbf{(TA)} \textbf{T}axonomy \textbf{A}ssignment, \textbf{(OP)} \textbf{O}TU and ESV \textbf{P}rocessing, and \textbf{(NI)} \textbf{N}etwork \textbf{I}nference.
-      Each step incorporates several processes (blue boxes), each of which in turn have several alternate algorithms for the same task (indicated by the text to the right of the blue boxes).
+      The steps of the workflow can be broken down into five major groups: \textbf{(SP)} \textbf{S}equence \textbf{P}rocessing, \textbf{(DC)} \textbf{D}enoising and \textbf{C}lustering, \textbf{(TA)} \textbf{T}axonomy \textbf{A}ssignment, \textbf{(OP)} \textbf{O}TU and ESV \textbf{P}rocessing, and \textbf{(NI)} \textbf{N}etwork \textbf{I}nference.
+      Each step incorporates several processes (blue boxes), each of which in turn has several alternative algorithms for the same task (indicated by the text to the right of the blue boxes).
       Each arrow describes the data that is being passed from one step to another.
-      The input to the pipeline are 16S sequencing reads, and the final output is the consensus network generated from the inferred co-occurrence networks.
-      For details on each process and the outputs, see Methods.
+      The inputs to the pipeline are 16S rRNA sequencing reads, and the final output is the consensus network generated from the inferred co-occurrence networks.
+      For details on each process and the different outputs, see Methods.
     }
@@
       The number of \ac{otu}s or \ac{esv}s generated by the respective methods are provided in the parenthesis next to their names.
-      The data used for the analysis in \textbf{(A, B)} were samples from the fecal microbiome transplant (FMT) dataset~\cite{Kang2017}, containing both healthy subjects and subjects with autism spectrum disorder (ASD).
+      The data used for the analysis in (A, B) were the samples from the fecal microbiome transplant (FMT) dataset~\cite{Kang2017}, containing both healthy subjects and subjects with autism spectrum disorder (ASD).
       \textbf{(C, D)} The distributions of the average weighted and unweighted UniFrac distance between the predicted sequence profile and the expected sequence profile in the mock datasets.
-      The distributions of the average weighted UniFrac distance show that de novo (DN) and open reference (OR) were the best-performing methods in most of the datasets, while they are worst performing methods under the unweighted UniFrac metric.
-      The good performance of dada2 (D2) under both distance metrics combined with its approach of identifying \ac{esv}s using de novo methods, prompts us to propose it as the default method for the DC step.
-      The data used for the analysis in \textbf{(C, D)} were the mock4, mock12 and mock16 datasets from mockrobiota~\cite{Bokulich2016}.
+      The average weighted UniFrac distances show that de novo (DN) and open reference (OR) were the best-performing methods in most of the datasets, while they are the worst-performing methods under the unweighted UniFrac metric.
+      The good performance of dada2 (D2) under both distance metrics combined with its approach of identifying \ac{esv}s using de novo methods, prompts us to use it as the default method for the DC step.
+      The data used for the analysis in (C, D) were the mock4, mock12, and mock16 datasets from mockrobiota~\cite{Bokulich2016}.
     }
@@
       A significant portion of the representative sequences are assigned to an ``unknown'' Genus in two of three databases (\ac{gg} and \ac{ncbi}).
-      The number of assigned genera for each database are displayed at the top of each column.
+      The number of assigned genera for each database is displayed at the top of each column.
       \textbf{(B)} The number of representative sequences assigned to the same taxonomic label when using different reference databases (for the top 100 sequences).
       The mismatches are fewer at higher taxonomic levels, but, even at the Order level there exists greater than 51\% of mismatches, demonstrating the poor agreement in taxonomic labels assigned by the different databases.
-      The data used for the analysis in \textbf{(A, B)} were samples (healthy and ASD) from the FMT dataset.
-      \textbf{(C)} The Bray-Curtis dissimilarity between the predicted taxonomy profile and expected taxonomy profile in the mock datasets shows that there is no singular best choice of database for every dataset, as all the databases have similar performance.
+      The data used for the analysis in (A, B) were samples (healthy and ASD) from the FMT dataset.
+      \textbf{(C)} The Bray-Curtis dissimilarity between the predicted taxonomy profile and expected taxonomy profile in the mock datasets shows that there is no singular best choice of database for every dataset, as all the databases show similar performances.
       The \ac{gg} database and the Naive Bayes classifier are chosen as the defaults for the TA step of \ac{micone} due to their popularity.
-      The datasets used for the analysis in \textbf{(C)} were the mock datasets from mockrobiota.
+      The datasets used for the analysis in (C) were the mock datasets from mockrobiota.
     }
@@
       \textbf{Networks generated using different network inference methods show notable differences in terms of edge-density and connectivity}.
-      \textbf{(A)} The nine different networks (excluding \ac{mldm}) generated by the different network inference methods.
-      The nodes for each network are arranged in the same positions in a circular layout and the differences in the connections can be directly visualized and compared.
+      \textbf{(A)} The nine different networks (excluding \acs{mldm}) generated by the different network inference methods.
+      The nodes for each network (representing taxa) are arranged in the same positions in a circular layout and the differences in the connections can be directly visualized and compared.
       The green links are positive associations and the orange links represent negative associations.
-      The networks are dissimilar and vary widely in terms of connectivity with the correlation-based methods, in general, producing networks with higher edge-densities.
+      The networks look dissimilar and vary widely in terms of connectivity, and it is notable that the correlation-based methods generally produce networks with higher edge-densities.
       A threshold of 0.3 was set for the correlation-based methods (sparcc, propr, spearman and pearson) and a threshold of 0.01 was set for the direct association methods (flashweave, spieceasi, cozine, harmonies, and spring).
       \textbf{(B)} The node overlap Upset plot indicates that all the networks have a large proportion of common nodes involved in connections (33 out of 68).
-      Whereas, \textbf{(C)} the edge overlap Upset plot shows that a very small fraction of these connections are actually shared (8 out of 202).
+      Conversely \textbf{(C)}, the edge overlap Upset plot shows that a very small fraction of these connections are actually shared (8 out of 202).
       The data used in this analysis were the healthy stool samples from the FMT dataset.
-      \ac{mldm} is not shown in the comparisons because the algorithm failed to converge for the particular network combination used here (default setting of the \ac{micone} pipeline).
+      \acs{mldm} is not shown in the comparisons because the algorithm failed to converge for the particular network combination used here (default setting of the \ac{micone} pipeline).
     }
@@
       \textbf{The associations generated by the scaled-sum consensus method show high precision in benchmarks using synthetic datasets}.
-      Each point on the box plot shows the precision of the co-occurrence network generated by the individual and consensus methods using the ``NorTA'' and ``seqtime'' methods (refer to Methods).
+    The different points on the box plot show the precision of co-occurrence networks generated through individual network inference methods and through consensus network construction approaches.
+Precision is estimated based on the comparisons with two sets of synthetic benchmark datasets (``NorTA'' and ``seqtime'', see Methods).
       The independent algorithms chosen for the comparison are the two best-performing correlation-based (propr, sparcc) and direct association based (spieceasi, flashweave) methods.
-      The different consensus based methods used are the scaled-sum (SS) and simple voting (SV) method.
+      For consensus network inference, we used the scaled-sum (SS) and simple voting (SV) methods.
       A weight threshold of 0.1 and a p-value threshold of 0.05 was applied to each network before the calculation of precision.
-      The overall best precision was consistently obtained by the scaled-sum method for $p \geq 0.333$ on both benchmark datasets and among all the individual network inference methods, spieceasi shows the best average precision.
-      The simple voting method when using presence of edges in all inferred networks as a requirement ($p = 1.000$), also outperforms spieceasi on average precision.
-      Therefore, we recommend the scaled-sum consensus method with $p = 0.333$ as the default tool for network inference, since this option provides a good balance of precision and sensitivity.
+      The overall best precision was consistently obtained by the scaled-sum consensus method for $p \geq 0.333$ on both benchmark datasets.
+Among all the individual network inference methods, spieceasi shows the best average precision.
+      The simple voting method, when using the presence of edges in all inferred networks as a requirement ($p = 1.000$), also outperforms spieceasi on average precision.
+      Therefore, we set the scaled-sum consensus method with $p = 0.333$ as the default tool for consensus network inference, since this option provides a good balance of precision and sensitivity (see also Figure~\ref{fig:figure_s5} and \ref{fig:figure_s6}).
     }
@@
       \textbf{The choice of reference database has the largest impact on network variance}.
-      \textbf{(A)} The percentage of variance in the networks contributed by the \acf{dc}, CC (chimera checking), \acf{ta}, \acf{op} and \acf{ni} steps of the pipeline calculated using ANOVA on the linear model, see the Methods section for details.
-      The data used for the analysis were the healthy stool samples from the FMT dataset.
-      A weight threshold of 0.1 and a p-value threshold of 0.05 was applied to each network before the analysis.
+      \textbf{(A)} The percentage of variance in the networks (from the  FMT dataset) contributed by the \acf{dc}, CC (chimera checking), \acf{ta}, \acf{op} and \acf{ni} steps of the pipeline calculated using ANOVA on a linear model (see Methods).
+      A weight threshold of 0.1 and a p-value threshold of 0.05 were applied to each network before the analysis.
       The taxonomy database contributes most to the variance between the networks (65.4\%) followed by the filtering of the counts matrix (26.8\%) in the OP step.
     The variation due to the NI, DC and CC steps are much smaller in comparison (6.553\%, 0.648\%, and 0.003\% respectively).
-      The fraction labeled as the residual is an artifact that arises when multiple steps are changed at the same time, however, this can be ignored as this fraction is negligible.
+      The negligible fraction labeled as the residual is an artifact that arises when multiple steps are changed at the same time.
       \textbf{(B)} All the inferred networks generated from various combinations of tools are shown as points on a PCA plot.
       Each point on the PCA plot represents a network inferred using different combinations of tools and parameters that are available in the \ac{micone} pipeline.
-      The color of the points corresponds to the tools used at each step of the pipeline (DC, TA, OP and NI).
-      The points on the PCA plot can be grouped based on the TA step, but the extent of this grouping decreasing when the filtering is turned on in the OP step, confirming that the variability in the networks decreased upon filtering out the taxonomic entities at low abundance.
+      The color of the points corresponds to the tools used at each step of the pipeline (DC, TA, OP, and NI).
+      The points on the PCA plot can be grouped based on the TA step, but the extent of this separation decreases when the filtering is turned on in the OP step, confirming that the variability in the networks decreased upon filtering out the taxonomic entities at low abundance.
       Some algorithms, especially the direct association methods, at the NI step can also be seen to generate networks that are less variable compared to the others.
-      The \ac{dc} step does not seem to have any correlation with the similarity of networks on the PCA plot.
+      The \ac{dc} step does not seem to have any correlation with the variation in the networks on the PCA plot.
     }
@@
       The networks for the control (left) and ASD (right) samples were generated using the default tools and parameters recommended by the \ac{micone} pipeline as described in Table~\ref{tab:micone_tools}.
-      There are 22 unique links in the network for control samples, 12 unique links in the network for ASD subjects and 7 edges in common between both networks.
+      There are 22 unique links in the network for control samples, 12 unique links in the network for ASD subjects, and 7 edges in common between both networks.
       The changes in these connections can serve as potential starting points for further experimental validations or literature surveys.
@@
       \textbf{Table of global network metrics for networks inferred from all possible combinations of tools}.
-      In each row one tool in a particular step is kept constant, and the metric is calculated for every possible combination of tools for the other steps of the pipeline.
+      In each row, one tool in a particular step is kept constant, and the metric is calculated for every possible combination of tools for the other steps of the pipeline.
       Therefore, each row shows the grouped average metric for each tool in every step of the pipeline.
-      The different network inference methods show the most variation in the global network metrics.
+      The network inference methods show the most variation in the global network metrics compared to tools in other steps of the pipeline.
     }
@@
           There is an increase in both weighted and unweighted UniFrac distances compared to when all the representative sequences are considered.
-          This shows that the 1000 most abundant representative sequences generated by the methods are not similar to each other.
+          This shows that the 1000 most abundant representative sequences generated by the DC methods are not as similar to each other.
           And since the weighted UniFrac is much smaller than the unweighted UniFrac distance, we can conclude that those reference sequences that are present in the middle of the abundance distribution (considering all sequences) are dissimilar.
@@
         \caption{
-          \textbf{The weighted and unweighted UniFrac distances between representative sequences generated using remove bimera and uchime for each denoising method are low.}
-          With the exception of de novo and open reference under the unweighted UniFrac metric, all the other methods have low dissimilarity, implying that the two chimera checking methods, uchime and remove bimera, produce similar outputs.
+          \textbf{The weighted and unweighted UniFrac distances are small for the representative sequences generated using remove bimera and uchime for each denoising method}.
+          With the exception of de novo and open reference under the unweighted UniFrac metric, all the other methods have high similarity, implying that the two chimera checking methods, uchime and remove bimera, produce similar outputs.
           This is especially true for the \ac{dada2} and Deblur methods which are the recommended denoising methods in the \ac{micone} pipeline.
@@
           The comparison made here is similar to Figure~\ref{fig:figure3}B, but instead of the top 100 taxonomic entities (by abundance), all the assignments from one database are matched with those from the other two databases.
-          Higher percentage of mismatches implies that the assigned taxonomies in the more abundant sequences (top 100) match more consistently.
+          The higher percentage of mismatches implies that the assigned taxonomies in the more abundant sequences (top 100) match more consistently.
         }
@@
         \textbf{The precision and sensitivity of the inferred networks on the ``NorTA'' synthetic interaction data}.
-        The different consensus based methods used are: scaled-sum (SS) and simple voting (SV) method.
+        The different consensus methods used are scaled-sum (SS) and simple voting (SV) methods.
         Pearson and Spearman methods are not used in the calculation of the consensus.
-        Among all the independent network inference methods, \ac{spieceasi} has the best average precision (0.944), but the overall best precision was consistently obtained by the scaled-sum method (0.956, 0.985 and 1.000).
-        The simple voting method when using presence of edges in all inferred networks as a requirement ($p = 1.000$), also outperforms \ac{spieceasi} on average precision (0.969).
-        Although \ac{spieceasi} has a higher sensitivity, if the goal of network inference is to obtain the list of associations that have a high probability of existing in the real microbial community, then the consensus methods perform better.
+        Among all the independent network inference methods, \acs{spieceasi} has the best average precision (0.944), but the overall best precision was consistently obtained by the scaled-sum method (0.956, 0.985, and 1.000).
+        The simple voting method when using the presence of edges in all inferred networks as a requirement ($p = 1.000$), also outperforms \acs{spieceasi} on average precision (0.969).
+        Although \acs{spieceasi} has a higher sensitivity, if the goal of network inference is to obtain the list of associations that have a high probability of existing in the real microbial community, then the consensus methods perform better.
       }
@@
         \textbf{The precision and sensitivity of the inferred networks on the ``seqtime'' synthetic interaction data.}
-        The different consensus based methods used are: scaled-sum (SS) and simple voting (SV) method.
+        The different consensus methods used are scaled-sum (SS) and simple voting (SV) methods.
         Pearson and Spearman methods are not used in the calculation of the consensus.
-        Among all the independent network inference methods, \ac{spieceasi} has the best average precision (0.624), but the overall best precision was consistently obtained by the scaled-sum method (0.688, 0.820 and 1.000).
-        The simple voting method when using presence of edges in all inferred networks as a requirement ($p = 1.000$), also outperforms \ac{spieceasi} on average precision (0.692).
+        Among all the independent network inference methods, \acs{spieceasi} has the best average precision (0.624), but the overall best precision was consistently obtained by the scaled-sum method (0.688, 0.820, and 1.000).
+        The simple voting method when using the presence of edges in all inferred networks as a requirement ($p = 1.000$), also outperforms \acs{spieceasi} on average precision (0.692).
         These results show that the scaled-sum method is not only much better suited for inferring robust and accurate interactions from count data generated from network topologies (NorTA), but it is also capable of accurately extracting real associations from Lotka-Volterra simulations.
@@
       The edges colored green are positive associations and those in red are negative associations.
-      We observe that changing the TA and OP steps leads to the creation of the most number of unique edges.
+      We observe that changing the TA and OP steps leads to the creation of the most unique edges.
       \textbf{(B)} The dot plot showing the fraction of nodes (left) and edges (right) in common between the default network and the networks generated by changing one step of the default pipeline.
       The low value of the common fraction for TA and OP steps shows that these steps induce the biggest changes in nodes and edges.
-      The NI step is not shown in this analysis because the consensus method uses edges from the individual network inference methods and a comparison would be biased.
+      The NI step is not shown in this analysis because the consensus methods use edges from the individual network inference methods and a comparison would be biased.
     }

### 2022-09-30 sections/results.tex
Update on Overleaf.

@@
   For example, \textit{Prevotella} and \textit{Porphyromonas}, genera previously implicated in ASD~\cite{Kang2017,hoGutMicrobiotaChanges2020} and cognitive impairment~\cite{chiPorphyromonasGingivalisInducedCognitive2021} display modified connectivity in our network, suggesting that the observed associations may be relevant for understanding the role of these bacteria in disease.
-  Additional visualization and comparison of networks can be performed using the \acl{mind}~\cite{huResourceComparisonIntegration2022}.
+  Additional visualization and comparison of networks can be performed using the \acf{mind}~\cite{huResourceComparisonIntegration2022}.
 
   Figure~\ref{fig:figure_s7} shows a sensitivity analysis in which we compared the default network against networks generated by altering one of the steps of the pipeline relative to the default.
-  This result, both visually (Figure~\ref{fig:figure_s7}A), and quantitatively (Figure~\ref{fig:figure_s7}B)  suggests that the most significant changes occur when the \ac{op} or \ac{ta} steps are changed from the default value.
+  This result, both visually (Figure~\ref{fig:figure_s7} A), and quantitatively (Figure~\ref{fig:figure_s7} B)  suggests that the most significant changes occur when the \ac{op} or \ac{ta} steps are changed from the default value.

### 2022-09-30 sections/introduction.tex
Change: Update references

@@
  For example, multiple studies have shown the importance of specific microbial associations in the healthy microbiome \cite{Lloyd-Price2016,Wu2016,HumanMicrobiomeProjectConsortium2012} and their role in dysbiosis \cite{Wang2017,Gilbert2016,Belizario2015}.
- In the context of terrestrial biogeochemistry, co-occurrence networks were shown to help understand microbiome assembly \cite{Fierer2017}, and  the response of microbial communities to environmental perturbations \cite{Jiao2019}.
+ In the context of terrestrial biogeochemistry, co-occurrence networks were shown to help understand microbiome assembly \cite{fiererEmbracingUnknownDisentangling2017}, and  the response of microbial communities to environmental perturbations \cite{Jiao2019}.
 
@@
 Popular methods for different phases of the analysis of 16S data include tools for: (i) quality checking and trimming the sequencing reads; (ii) denoising and clustering the trimmed reads \cite{Caporaso2010,Callahan2016,Amir2017}; (iii) assigning taxonomy to the denoised reads \cite{bokulichOptimizingTaxonomicClassification2018}; (iv) processing and transforming the taxonomy count matrices \cite{Weiss2015}; and (v) inferring the co-occurrence network \cite{Watts2018,Kurtz2015,tackmannRapidInferenceDirect2019}.
-Different specific algorithms are often aggregated into popular online platforms (like MG-RAST~\cite{Keegan2016}, Qiita~\cite{qiita}) and software packages (such as \ac{qiime2}~\cite{bolyenReproducibleInteractiveScalable2019}).
+Different specific algorithms are often aggregated into popular online platforms (like MG-RAST~\cite{keeganMGRASTMetagenomicsService2016}, Qiita~\cite{gonzalezQiitaRapidWebenabled2018}) and software packages (such as \ac{qiime2}~\cite{bolyenReproducibleInteractiveScalable2019}).
 The different methods and tools can lead to vastly different inferences of community compositions and co-occurrence networks \cite{Golob2017,Weiss2016}, making it difficult to reliably compare networks across different publications and studies.

### 2022-09-30 sections/results.tex
Change: Update references

@@
 
-  To quantify the differences between the networks, we analyzed the distribution of common nodes and edges (Figure~\ref{fig:figure4} B and \ref{fig:figure4}C) using UpSet plots~\cite{Lex}.
+  To quantify the differences between the networks, we analyzed the distribution of common nodes and edges (Figure~\ref{fig:figure4} B and \ref{fig:figure4}C) using UpSet plots~\cite{lexUpSetVisualizationIntersecting2014}.
   The node intersection analysis shows that the networks have $33$ out of $68$ total unique nodes in common and that no network possesses a unique node.

### 2022-09-30 sections/discussion.tex
Update on Overleaf.

@@
   Through \ac{micone}, in addition to transparently revealing the dependence of co-occurrence networks on tool and parameter choices (see Discussion in Supplementary Text for details on the DC, TA and OP steps), we have taken advantage of our spectrum of computational options and the availability of mock and synthetic datasets, to suggest a default standard setting.
+  \todo[noline]{R1.3}
   Additionally, we have developed a consensus approach, that can reliably generate networks that are fairly robust across multiple tool choices.
@@
   The network analyses in this study are primarily at the Genus level, wherein the lowest resolution of a node is a Genus and if an entity cannot be resolved to the Genus level, the next lowest taxonomic level is used (for example, Family).
+  \todo[noline]{R2.1c}
   As a consequence, two entities belonging to the same lineage where one entity is resolved to the Genus level and another is resolved to the Family level are treated as two different nodes in the network.
@@
   Although direct comparisons between co-occurrence networks and directly measured interactions are difficult to interpret and highly debated~\cite{hiranoDifficultyInferringMicrobial2019,gobernaCautionaryNotesUse2022}.
+  \todo[noline]{R2.2}
   Further, benchmarking of co-occurrence networks could also be pursued through the use of literature-based interactions~\cite{lima-mendezDeterminantsCommunityStructure2015a} or biological benchmark interaction data~\cite{sungGlobalMetabolicInteraction2017a}.
@@
   Although in the current analysis, we have only used default parameter values recommended by the tool creators, the \ac{micone} pipeline could be used in the future to explore any combinations of parameters and to optimize these values for improved network inference.
+  \todo[noline]{R1.3}
   Overall, there likely is no ``best method'' for the various steps of 16S data analysis, and hence, \ac{micone} is intended to help researchers to identify the methods and algorithms that are most suitable for their datasets in an easy-to-use and reproducible manner.

### 2022-09-30 sections/methods.tex
Update on Overleaf.

@@
   This study utilized two types of 16S rRNA sequencing datasets: biological datasets and mock/synthetic datasets.
+  \todo[noline]{R2.4}
   Biological datasets are collections of sequencing reads obtained from naturally occurring microbial community samples.
@@
   This modular architecture simplifies the process of adding new tools (refer to the modules section in the \ac{micone} documentation).
+  \todo[noline]{R2.11}
   The main components of the pipeline are detailed in the subsequent sections.
@@
   The module also supports rarefaction, which is a normalization technique used to overcome the bias that might arise due to variable sampling depth in different samples.
+  \todo[noline]{R2.5}
   This is performed either by sub-sampling or by normalization of the matrix to the lowest sampling depth~\cite{Weiss2015}.
@@
   The group operation transforms the \ac{otu} or \ac{esv} count matrix into a taxonomic count matrix at the requested level by adding up counts that map to the same taxonomy and is carried out at the end of the OP step.
+  \todo[noline]{R2.1b}
   Finally, transformations are performed in order to correct for and overcome the compositional bias that is inherent in the counts matrix (in the analysis performed in the study these were disabled and directly handled by the network inference algorithm).
@@
   The consensus methods combine networks inferred from both correlation-based and direct association methods.
+  \todo[noline]{R1.1}
+  \todo[noline]{R1.4}
+  \todo[noline]{R2.2}
+  \todo[noline]{R2.3}
   First, for the correlation-based methods, we calculate p-values using null models and then merge the p-values using Brown's p-value merging method~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016,faustCoNetAppInference2016} (see Methods).
@@
   The next step in the consensus algorithm workflow is to merge the p-values for the networks generated by the correlation-based methods.
+  \todo[noline]{R1.1}
   This step is performed using the Brown's p-value merging method~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016,faustCoNetAppInference2016}.
@@
   This section defines the notation used for the network variability analysis performed for Figure~\ref{fig:figure6}.
+  \todo[noline]{R1.2}
+  \todo[noline]{R1.8}
+  \todo[noline]{R2.1b,d}
 
@@
    Note that row $j$ of $\mathbf{\bar{N}}$, $\bar{N}_j$ is the vector that encodes the values of edge $j$ across all the networks.
+  \todo[noline]{R1.5}
+  \todo[noline]{R2.1b}
 
@@
    The variance contributed by each step of the pipeline was calculated for every component in $\mathbf{C}$ matrix through ANOVA using the Python statsmodels~\cite{seaboldStatsmodelsEconometricStatistical2010} package and is shown in Figure~\ref{fig:figure6}A.
+  \todo[noline]{R1.6}
   The total variance for the network was calculated by adding the variances for each connection and normalizing with the degrees of freedom.
   The merged network table $\mathbf{\bar{N}}$ was used as the input to the PCA analysis to generate Figure~\ref{fig:figure6}B.
+  \todo[noline]{R1.7}
 
@@
   We generated synthetic interaction data using two methodologies previously used for benchmarking network inference methods.
+  \todo[noline]{R1.4}
+  \todo[noline]{R2.2}
+  \todo[noline]{R2.4}
 

### 2022-09-30 sections/results.tex
Update on Overleaf.

@@
   Our analysis involved two types of data: The first type consisted of 16S sequencing data from samples of human stool microbiomes from a fecal microbiome transplant (FMT) study of autism~\cite{Kang2017}.
+  \todo[noline]{R2.4}
   The second type was a collection of datasets synthetically or artificially created for the specific goal of evaluating computational analysis tools.
@@
   This is because the open-reference and de novo clustering methods return a much larger number of \ac{otu}s compared to the other pipelines and would affect the accuracy of the network inference step if stringent filtering is not performed.
+  \todo[noline]{R2.6}
   Overall, since \ac{dada2} as compared to Deblur, displays better performance on all the mock datasets on the weighted UniFrac metric, we set this tool as the default for the DC step of the pipeline.
+  \todo[noline]{R2.10}
   However, if comparison across studies that have sequenced different 16S regions is required, closed-reference and open-reference might be a better option.
@@
   Since no database performs better than others against mock datasets, the choice of which database to use could be driven by other reasons (see Supplementary discussion).
+  \todo[noline]{R2.2a}
   One reason to choose a particular database could be the frequency of updates and the potential for future growth.
@@
   The TA step results in a taxonomic counts table that is used as input to the subsequent steps of the pipeline.
+  \todo[noline]{R2.1b}
   Note that the count tables at different levels can be obtained through aggregation; for example, Genus count tables were obtained by summing up the counts of the lower taxonomy levels (Species and \ac{otu}) that map to the same higher taxonomy level entity.
@@
   Filtered (see \ac{op} step in Methods) genus-level counts table obtained using the default settings in the previous steps were used as input for the different network inference algorithms (Figure~\ref{fig:figure4}).
+  \todo[noline]{R2.1b}
   Even from a visual inspection (Figure~\ref{fig:figure4}A), one can see that the different networks differ vastly in their edge-density and connectivity, with common edges often displaying inverted signs.
@@
  Inspired by previous approaches~\cite{bustinceFuzzySetsTheir2008,tsarevApplicationMajorityVoting2018}, we developed two methods that take into consideration the evidence offered by each network inference algorithm and generate a consensus network that contains the common edges among the inferred networks.
+  \todo[noline]{R1.1}
+  \todo[noline]{R1.4}
 
@@
   The analysis of the rewiring of associations in the ASD samples with respect to the control provides a guide for the identification of key genera that could be linked to dysbiosis.
+  \todo[noline]{R2.14}
   We observed 22 unique links in the network for control samples, 12 unique links in the network for ASD subjects, and 7 edges in common between the two networks.

### 2022-09-30 sections/supplementary.tex
Change: Update references
Update on Overleaf.

@@
     The main biological dataset used in this study was the collection of 16S rRNA sequencing reads from stool samples (healthy and autistic individuals) for a fecal microbiome transplant study~\cite{Kang2017}.
-    The data containing the 16S sequencing reads (V4 region) was downloaded from Qiita~\cite{qiita} (study ID: 10532).
+    The data containing the 16S sequencing reads (V4 region) was downloaded from Qiita~\cite{gonzalezQiitaRapidWebenabled2018} (study ID: 10532).
     Only runs 2, 3, and 4 were used for the subsequent analysis as these runs consisted of paired-end sequencing data, and run 1 contained single-end data.
@@
   In Table~\ref{tab:network_metrics} we show various global network metrics calculated for each tool in the pipeline.
+  \todo[noline]{R2.1a}
+  \todo[noline]{R2.7}
   All the networks that make use of a particular tool are grouped together, and the following average metrics are calculated for each group:
@@
 
-  Fisher~\cite{fisher_224a_1948} proposed that for $k$ independent p-values, each generated by $k$ different methods and denoted by $\bar{P}^i$ (notations are same as used in the "Consensus network and p-value merging" subsection of the Methods), the following will hold true for the statistic $\Psi$:
+  Fisher~\cite{fisher_224a_1948} proposed that for $k$ independent p-values, each generated by $k$ different methods and denoted by $\bar{P}^i$ (notations are same as used in the "Consensus network and p-value merging" subsection of the Methods).
+  \todo[noline]{R1.1}
+  The following will hold true for the statistic $\Psi$:
   \begin{equation*}
@@
 
-  % TODO: Should we mention the MIND database here?
   \subsection*{The JSON network format and network exports}
 
-    The default format \ac{micone} uses for storing the network files is the JSON (JavaScript Object Notation) format.
+    The default format \ac{micone} uses for storing the network files is the JSON (JavaScript Object Notation) format which is supported by the \acf{mind}~\cite{huResourceComparisonIntegration2022}.
     The custom JSON schema we have designed is able to store all network-related information pertaining to nodes, links, and the metadata related to the links and datasets.
@@
   For the chimera checking module, we suggest using the remove bimera method since it was developed in conjunction with \ac{dada2} and its performance does not significantly differ from uchime-denovo.
-  For the current work we have decided to focus on the tools most widely used at the time of the analysis.
+  For the current work, we have decided to focus on the tools most widely used at the time of the analysis.
   Some tools which were not as widely used (e.g.
@@
   The choice of taxonomy database was found to be the most important factor in the inference of microbial co-occurrence networks, contributing $65.4\%$ of the total variance.
+  \todo[noline]{R2.2a}
   The frequent changes in the taxonomy nomenclature coupled with the frequency of updates to the various 16S reference databases create inherent differences \cite{Balvociute2017} in taxonomy hierarchies in these databases.
   Our analysis revealed that no particular reference database performs better than the others across the different mock dataset benchmarks.
+  \todo[noline]{R2.12}
   The default reference database in the pipeline is the \ac{gg} reference database along with the ``Naive Bayes'' classifier as the query tool.
@@
   Additionally, a particular database might be more appropriate than the rest based on specific requirements.
-  For example, in order to generate a dataset that is compatible with the \ac{mind} platform~\cite{huResourceComparisonIntegration2022} \ac{ncbi} is the most appropriate choice as it guarantees compatibility of taxonomic hierarchy and therefore comparability with other datasets.
+  For example, to generate a dataset that is compatible with the \ac{mind} platform~\cite{huResourceComparisonIntegration2022} \ac{ncbi} is the most appropriate choice as it guarantees compatibility of taxonomic hierarchy and therefore comparability with other datasets.
   Furthermore, we also enable users to use custom databases~\cite{Ritari2015,iiRESCRIPtReproducibleSequence2021} with the BLAST and Naive Bayes classifiers that are incorporated into the pipeline (from \ac{qiime2}).
-  We suggest that that choice of the database should be made based on possible reported or inferred biases in the representation of given biomes in a specific databases~\cite{Balvociute2017,iiRESCRIPtReproducibleSequence2021}, as choosing taxon-specific databases have also been observed to compromise classification~\cite{rmarcelinoUseTaxonspecificReference2020}.
+  We suggest that the choice of the database should be made based on possible reported or inferred biases in the representation of given biomes in a specific databases~\cite{Balvociute2017,iiRESCRIPtReproducibleSequence2021}, as choosing taxon-specific databases have also been observed to compromise classification~\cite{rmarcelinoUseTaxonspecificReference2020}.
 
@@
 
+\newpage
   \pagenumbering{goggle}

### 2022-09-30 sections/methods.tex
Update on Overleaf.

@@
   \todo[noline]{R1.4}
+  First, for the correlation-based methods, we calculate p-values using null models and then merge the p-values using Brown's p-value merging method~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016,faustCoNetAppInference2016}.
   \todo[noline]{R2.2}
   \todo[noline]{R2.3}
-  First, for the correlation-based methods, we calculate p-values using null models and then merge the p-values using Brown's p-value merging method~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016,faustCoNetAppInference2016} (see Methods).
   Second, we filter all the inferred networks based on an association strength threshold of 0.1 and a p-value cutoff of 0.05.

### 2022-10-14 manuscript.tex
Add: Combined manuscript file and Text S1 caption

%--------------------------------------------------------%
% Journal Article Manuscript Template
%--------------------------------------------------------%

%--------------------------------------------------------%
% DOCUMENT CLASS
%--------------------------------------------------------%

  % Change "letterpaper" to "a4" if you use a4 paper size
  \documentclass[letterpaper,12pt]{article}

%--------------------------------------------------------%
% TITLE SECTION
%--------------------------------------------------------%

  %Abstract
  \usepackage{abstract} % Allows abstract customization
  % Set the "Abstract" text to bold
  \renewcommand{\abstractnamefont}{\normalfont\bfseries}
  % Set the abstract itself to small italic text
  \renewcommand{\abstracttextfont}{\normalfont\small\itshape}

  %Title
  \usepackage{titlesec} % Allows customization of titles

  %Authors
  \usepackage{authblk} % For multiple authors

  %Date
  \usepackage{datetime} % allows for including today's date
  % These two lines creates a new date format ``Month day(th), year''
  \newdateformat{usvardate}{
  \monthname[\THEMONTH] \ordinal{DAY}, \THEYEAR}

%--------------------------------------------------------%
% HEADERS & FOOTERS
%--------------------------------------------------------%

  %Footnotes
  % \usepackage[bottom]{footmisc} % Makes footnotes stick to bottom of the page

  %Headers from page 2 on
  % \usepackage{fancyhdr}
  \pagestyle{plain}
  % \fancyheadoffset{0cm}
%   \setlength{\headheight}{15pt}

%--------------------------------------------------------%
% MACROS
%--------------------------------------------------------%

  % Define keywords macro command
  \providecommand{\keywords}[1]{\textbf{\textit{Keywords---}} #1}

%--------------------------------------------------------%
% MATH SUPPORT
%--------------------------------------------------------%

  % The amssymb package provides various useful mathematical symbols
  \usepackage{amssymb}
  % The amsthm package provides extended theorem environments
  \usepackage{amsthm}
  % The newtxmath package provides additional math symbol support
  % in Times New Roman symbols, etc.
  \usepackage{newtxmath}
  \usepackage{mathtools}
  \usepackage{blkarray, bigstrut}

%--------------------------------------------------------%
% FONTS
%--------------------------------------------------------%

  \usepackage{microtype} % Slightly tweak font spacing for aesthetics
  \usepackage[utf8]{inputenc}
  \usepackage{newtxtext} % Makes default font Adobe Times New Roman

%--------------------------------------------------------%
% LINES
%--------------------------------------------------------%

  % Spacing
  \usepackage{setspace} % See \doublespacing command at the top of content.tex
  % Numbering
  \usepackage{lineno} 	% See \linenumbers at the top of content.tex
  \usepackage[table,x11names]{xcolor}
  % Lists
  \usepackage{enumitem}
  \setlist{nosep}
  \setlist[itemize]{leftmargin=*}

%--------------------------------------------------------%
% MARGINS
%--------------------------------------------------------%

  %NOTE: All spaces in this template are in inches, because it is
  % formatted for letterpaper (8.5 x 11 inch) paper. If you use a4
  % paper, choose different sizes in millimeters or centimeters.
  \usepackage[top=1.5in, bottom=1.5in, left=1in, right=1in]{geometry}

%--------------------------------------------------------%
% COMMENTS
%--------------------------------------------------------%

  \usepackage[colorinlistoftodos]{todonotes} % allows margin comments
  % See examples in content.tex, and here for manual:
  % http://www.ctan.org/pkg/todonotes
  \usepackage{soul} % allows for highlighting

%--------------------------------------------------------%
% ACRONYMS
%--------------------------------------------------------%

  \usepackage[nohyperlinks,nolist]{acronym} % Managing acronyms

%--------------------------------------------------------%
% GRAPHICS
%--------------------------------------------------------%

  \usepackage{graphicx,caption} % More advanced figure inclusion
  \graphicspath{{figures/}} % Set the default folder for images
  \usepackage{float} % For specifying table/figure locations, i.e. [ht!]

  % The printlen command allows the user to print the exact text width or height.
  % This is useful, when trying to create graphics (outside of LaTeX, of course)
  % with the optimal dimensions. See here for usage: http://www.ctan.org/pkg/printlen
  \usepackage{printlen}

  \usepackage[section]{placeins} % Used to ensure that figures do not go into the next section

%--------------------------------------------------------%
% TABLES
%--------------------------------------------------------%

  \usepackage{longtable} % For long tables that span multiple pages
  \newcommand{\sym}[1]{\rlap{#1}}% For symbols like *** in tables
  \usepackage{tabularx} % Allows advanced table features
  \newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
  \newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
  \newcolumntype{R}[1]{>{\raggedleft\arraybackslash}p{#1}}
  \usepackage{relsize} % Allows precise adjustment of font size,
  %useful for fitting tables to page width
  \usepackage{multirow}
  %for horizontal tables
  \usepackage{lscape}

%--------------------------------------------------------%
% REFERENCES
%--------------------------------------------------------%

  \usepackage{hyperref} % For hyperlinks in the PDF
  \usepackage{csquotes}
  \usepackage[style=nature,url=false,backend=biber,sorting=none]{biblatex}
  \bibliography{references/references.bib}

% Header from Page Three on: Edit below for left and right headers
% \lhead{}
% \rhead{}

%--------------------------------------------------------%
% BEGIN DOCUMENT
%--------------------------------------------------------%

\begin{document}

% COVER PAGE

\begin{titlepage}

  \newcommand{\HRule}{\rule{\linewidth}{0.5mm}} % Defines a new command for the horizontal lines, change thickness here

  \center % Center everything on the page

% HEADING SECTION

  %\textsc{\LARGE University Name}\\[1.5cm] % Name of your university/college
  % \textsc{\Large Manuscript Submission}\\[0.5cm] % Major heading such as course name
  % \textsc{\large The Journal of Blah Blah}\\[0.5cm] % Minor heading such as course title

% TITLE SECTION

  \vspace*{\fill}
  {\huge Inferring microbial co-occurrence networks from amplicon data: a systematic evaluation}\\[0.4cm]
  % {\huge  2. Investigating the best practices for inference of microbial co-occurrence networks from 16S data}\\[0.4cm] % Title of your document
  % {\huge  3. Attempting to find the best practice pipeline for inferring co-occurrence networks from 16S data}\\[0.4cm] % Title of your document
  % {\huge  4. Deciphering the complexities in co-occurrence network inference from 16S data}\\[0.4cm] % Title of your document

% AUTHOR SECTION

  \vspace{1.5 cm}
  Dileep Kishore\textsuperscript{a,b},
  Gabriel Birzu\textsuperscript{c,f},
  Zhenjun Hu\textsuperscript{a},
  Charles DeLisi\textsuperscript{a,c},
  Kirill S. Korolev\textsuperscript{a,b,c}\#,\\
  Daniel Segr\`{e}\textsuperscript{a,b,d,e}\#\\
  \vspace{1cm}
  \textsuperscript{a}Bioinformatics Program, Boston University, Boston, Massachusetts, USA\\
  \textsuperscript{b}Biological Design Center, Boston University, Boston, Massachusetts, USA\\
  \textsuperscript{c}Department of Physics, Boston University, Boston, Massachusetts, USA\\
  \textsuperscript{d}Department of Biomedical Engineering, Boston University, Boston, Massachusetts, USA\\
  \textsuperscript{e}Department of Biology, Boston University, Boston, Massachusetts, USA\\
  \textsuperscript{f}Department of Applied Physics, Stanford University, Stanford, California, USA\\
  \textsuperscript{$\#$}Correspondence should be sent to \href{mailto:korolev@bu.edu}{korolev@bu.edu} or \href{mailto:dsegre@bu.edu}{dsegre@bu.edu}\\

  \vspace{2cm}

% % DATE SECTION

  % \vspace{1.5 cm}
  % {\large Submitted: \today}\\[3cm] % Date, change the \today to a set date if you want to be precise

  \vspace*{\fill} % Fill the rest of the page with whitespace

\end{titlepage}

%-----------------------------------------------------------------

\newpage

\thispagestyle{empty} % Removes header on page two.

% NOTE: Comment out the lines below to remove line numbers
  % Running line numbers:
  \linenumbers
  \setlength\linenumbersep{15pt}
  \renewcommand\linenumberfont{\normalfont\footnotesize\sffamily\color{gray}}
  %\pagewiselinenumbers % Same, but that reset on every page:
  \modulolinenumbers[1] % Number only every line. Change for fewer.

%--------------------------------------------------------%
%   CONTENT
%--------------------------------------------------------%

% ABSTRACT
\begin{abstract}
  {
    \noindent
    Microbes commonly organize into communities consisting of hundreds of species involved in complex interactions with each other.
    16S ribosomal RNA (16S rRNA) amplicon profiling provides snapshots that reveal the phylogenies and abundance profiles of these microbial communities.
    These snapshots, when collected from multiple samples, can reveal the co-occurrence of microbes, providing a glimpse into the network of associations in these communities.
    However, the inference of networks from 16S data involves numerous steps, each requiring specific tools and parameter choices.
    Moreover, the extent to which these steps affect the final network is still unclear.
    In this study, we perform a meticulous analysis of each step of a pipeline that can convert 16S sequencing data into a network of microbial associations.
    Through this process, we map how different choices of algorithms and parameters affect the co-occurrence network and identify the steps that contribute substantially to the variance.
    We further determine the tools and parameters that generate robust co-occurrence networks and develop consensus network algorithms based on benchmarks with mock and synthetic datasets.
    The Microbial Co-occurrence Network Explorer or \acs{micone} (available at \href{https://github.com/segrelab/MiCoNE}{https://github.com/segrelab/MiCoNE}), follows these default tools and parameters and can help explore the outcome of these combinations of choices on the inferred networks.
    We envisage that this pipeline could be used for integrating multiple datasets, and for generating comparative analyses and consensus networks that can guide our understanding of microbial community assembly in different biomes.
  }
\end{abstract}

% Insert keywords here
\keywords{Microbiome, 16S rRNA, Interaction, Denoising, Taxonomy, Network Inference, Correlations, QIIME2, Co-occurrence, Networks, Consensus algorithm, Pipeline, nextflow}

% \doublespacing

\section*{Importance}
  Mapping the interrelationships between different species in a microbial community is important for understanding and controlling their structure and function.
  The surge in the high-throughput sequencing of microbial communities has led to the creation of thousands of datasets containing information about microbial abundances.
  These abundances can be transformed into co-occurrence networks, providing a glimpse into the associations within microbiomes.
  However, processing these datasets to obtain co-occurrence information relies on several complex steps, each of which involves numerous choices of tools and corresponding parameters.
  These multiple options pose questions about the robustness and uniqueness of the inferred networks.
  In this study, we address this workflow and provide a systematic analysis of how these choices of tools affect the final network, and guidelines on appropriate tool selection for a particular dataset.
  We also develop a consensus network algorithm that helps generate more robust co-occurrence networks based on benchmark synthetic datasets.

\doublespacing

% INTRODUCTION
\section*{Introduction}

Microbial communities are ubiquitous and play an important role in marine and terrestrial environments, urban ecosystems, and human health \cite{lima-mendezDeterminantsCommunityStructure2015a,Thompson2017,royo-llonchCompendium530Metagenomeassembled2021,tedersooFungalBiogeographyGlobal2014,dankoGlobalMetagenomicMap2021,mclellanMicrobiomeUrbanWaters2015,HumanMicrobiomeProjectConsortium2012}.
  These microbial communities, or microbiomes, often comprise several hundreds of different microbial strains interacting with each other and their environment, often through complex metabolic and signaling relationships~\cite{zelezniakMetabolicDependenciesDrive2015,Ghoul2016,coyteUnderstandingCompetitionCooperation2019,DSouza2018}.
  Understanding how these interconnections shape community structure and function is a fundamental challenge in microbial ecology, and has applications in the study of microbial ecosystems across different biomes.
  With the advancement in DNA sequencing technologies~\cite{huNextgenerationSequencingTechnologies2021,buermansNextGenerationSequencing2014,Narihiro2017},  more information can be extracted from these microbial community samples than ever before.
  In particular, high-throughput sequencing, including metagenomic sequencing and sequencing of 16S rRNA gene amplicons (hereafter referred to as 16S data) of microbial communities, can help detect, identify and quantify a large portion of the constitutive microorganisms of a microbiome \cite{ju16SRRNAGene2015,Jovel2016,quinceShotgunMetagenomicsSampling2017,sedlarBioinformaticsStrategiesTaxonomy2017}.
  These advances have led to large-scale data collection efforts involving terrestrial~\cite{Thompson2017,gilbertMeetingReportTerabase2010,tedersooFungalBiogeographyGlobal2014}, marine~\cite{lima-mendezDeterminantsCommunityStructure2015a,royo-llonchCompendium530Metagenomeassembled2021} and human-associated microbiota~\cite{HumanMicrobiomeProjectConsortium2012,proctorIntegrativeHumanMicrobiome2019,Lloyd-Price2016}.

 This wealth of information has the potential to help us understand how communities assemble and operate.
 In particular, a powerful tool for translating microbiome composition data into knowledge is the construction of association (co-occurrence or correlation) networks, in which  microbial taxa are represented by nodes, and frequent co-occurrences (or negative correlations) across datasets are encoded as edges between nodes.
 While the relationship between directly measured interactions~\cite{lubbeExometabolomicAnalysisCrossFeeding2017,Jian2020,Hsu2019} and statistically inferred co-occurrence is still poorly understood \cite{Zuniga2017,Rottjers2018}, a significant amount of effort has gone into estimating correlations from large microbiome sequence datasets~\cite{faustMicrobialCooccurrenceRelationships2012,leeCrosskingdomCooccurrenceNetworks2022,faustMicrobialInteractionsNetworks2012a,maEarthMicrobialCooccurrence2020a}.

 The importance of these networks is two-fold: first, they can serve as maps that help identify hubs of keystone species \cite{Menon2018,Rottjers2018}, and the community response to environmental perturbations or underlying host conditions \cite{Gilbert2016}; second, they can serve as a bridge towards building mechanistic models of microbial communities, greatly enhancing our capacity to understand and control them.
 For example, multiple studies have shown the importance of specific microbial associations in the healthy microbiome \cite{Lloyd-Price2016,Wu2016,HumanMicrobiomeProjectConsortium2012} and their role in dysbiosis \cite{Wang2017,Gilbert2016,Belizario2015}.
 In the context of terrestrial biogeochemistry, co-occurrence networks were shown to help understand microbiome assembly \cite{fiererEmbracingUnknownDisentangling2017}, and  the response of microbial communities to environmental perturbations \cite{Jiao2019}.

One of the most frequently used avenues for inferring co-occurrence networks is the parsing and analysis of 16S sequencing data \cite{Rottjers2018,Friedman2012}.
Numerous software tools and pipelines have been developed to analyze 16S sequencing data, with a strong emphasis on the known limitations of this method, including resolution, sequencing depth, compositional nature, sequencing errors, and copy number variations \cite{Bharti2019,pollockMadnessMicrobiomeAttempting2018}.
Popular methods for different phases of the analysis of 16S data include tools for: (i) quality checking and trimming the sequencing reads; (ii) denoising and clustering the trimmed reads \cite{Caporaso2010,Callahan2016,Amir2017}; (iii) assigning taxonomy to the denoised reads \cite{bokulichOptimizingTaxonomicClassification2018}; (iv) processing and transforming the taxonomy count matrices \cite{Weiss2015}; and (v) inferring the co-occurrence network \cite{Watts2018,Kurtz2015,tackmannRapidInferenceDirect2019}.
Different specific algorithms are often aggregated into popular online platforms (like MG-RAST~\cite{keeganMGRASTMetagenomicsService2016}, Qiita~\cite{gonzalezQiitaRapidWebenabled2018}) and software packages (such as \ac{qiime2}~\cite{bolyenReproducibleInteractiveScalable2019}).
The different methods and tools can lead to vastly different inferences of community compositions and co-occurrence networks \cite{Golob2017,Weiss2016}, making it difficult to reliably compare networks across different publications and studies.
 This difference is partially due to the focus of existing platforms on \ac{otu} or \ac{esv} generation and not on the effects of upstream statistical methods on the inferred co-occurrence networks.
 Furthermore, no organized framework currently exists that can systematically analyze and compare each step in the pipeline for processing amplicons into co-occurrence networks.

 In this study, we present a standardized 16S data analysis pipeline called \ac{micone} that produces robust and reproducible co-occurrence networks from 16S sequence data of microbial communities, and enable users to interactively explore how the network would change upon using different alternative tools and parameters at each step.
 Our pipeline is coupled to an online integrative tool for the organization, visualization, and analysis of inter-microbial networks called \ac{mind}~\cite{huResourceComparisonIntegration2022}, which is available at \href{http://microbialnet.org/}{http://microbialnet.org/}.
Through a systematic comparative analysis, we determine which steps of the \ac{micone} pipeline have the largest influence on the final network, and which choice seems to have the optimal agreement with the tested mock and synthetic datasets.
These steps together with our default settings ensure better reproducibility and easier comparison of co-occurrence networks across datasets.
We expect that our tool will also be useful for benchmarking future alternative methods, and for ensuring a transparent evaluation of the possible biases introduced by the use of specific tools.

% RESULTS
\section*{Results}

  \subsection*{\acl{micone} (\acs{micone})}

  We developed \ac{micone}, a flexible and modular pipeline for the inference of co-occurrence networks from 16S data.
  \ac{micone} incorporates various popular, publicly available tools as well as custom Python modules for 16S data analysis and network inference (Methods).
  The different steps that are a part of the \ac{micone} co-occurrence network inference workflow (Figure~\ref{fig:figure1}) can be grouped into five major modules; (i) \ac{sp}; (ii) \ac{dc}; (iii) \ac{ta}; (iv) \ac{op}; and (v) \ac{ni}.
  Each process in the pipeline is implemented through multiple tools (see Methods and Figure~\ref{fig:figure1}).
  The effects of changing any intermediate step of the pipeline can be evaluated in terms of the final network outcome, as well as on any of the intermediate metrics and data outputs.
  The choice of tools and parameters is encoded in a configuration file (with parameters as shown in Tables S2-S6 at \href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper}).
  Through a systematic analysis of tool combinations at each step of the pipeline, we estimated how much the final co-occurrence network depends on the possible choices at each step.

  Our analysis involved two types of data: The first type consisted of 16S sequencing data from samples of human stool microbiomes from a fecal microbiome transplant (FMT) study of autism~\cite{Kang2017}.
  \todo[noline]{R2.4}
  The second type was a collection of datasets synthetically or artificially created for the specific goal of evaluating computational analysis tools.
  In particular, in order to benchmark each step in \ac{micone}, we used both mock data (labeled mock4, mock12, and mock16) from mockrobiota~\cite{Bokulich2016} and synthetic networks generated using the NorTA~\cite{Kurtz2015} and seqtime~\cite{Rottjers2018} approaches (See Methods).

  \FloatBarrier

  \subsection*{DC: Denoising and clustering methods differ in their identification of sequences that are low in abundance}

  The \ac{dc} step is commonly carried out to generate representative sequences (in the form of the \acs{otu}/\acs{esv} tables) from the demultiplexed and trimmed 16S sequencing data.
  In order to compare the count tables generated by different tools, we processed the 16S sequencing reads (from the FMT study~\cite{Kang2017}) using 5 different methods: open-reference clustering, closed-reference clustering, de novo clustering, \ac{dada2}~\cite{Callahan2016} and Deblur~\cite{Amir2017}.
  The first three methods are from the vsearch plugin from \ac{qiime2}~\cite{bolyenReproducibleInteractiveScalable2019}.
  The closed and open reference methods in this analysis use the \acl{gg}~\cite{DeSantis2006} database for reference sequence alignment.

  A comparison of the different methods was carried out by calculating the mean UniFrac distances across all samples (Figure~\ref{fig:figure2}).
  The analysis was performed using both the weighted UniFrac~\cite{Lozupone2007} (Figure~\ref{fig:figure2}A) distance metric, which takes into account the counts of the representative sequences, and the unweighted UniFrac~\cite{Lozupone2005} (Figure~\ref{fig:figure2}B) distance metric, which gives equal weights to each sequence.

  The first main message emerging from this analysis is that the representative sequences generated by the different methods, with the exception of Deblur, are similar to each other when weighted by their abundance (Figure~\ref{fig:figure2}A).
  A second message is that the different methods differ mainly in the assignment of sequences of lower abundance.
  This can be inferred from the unweighted comparison (Figure~\ref{fig:figure2}B) which shows an increase in dissimilarity between each pair of methods (see additional details in Supplementary and Figure \ref{fig:figure_s2}).

  These comparisons only elucidate the similarity between a pair of methods.
  To determine which tool most accurately recapitulates the reference sequences in the samples, we applied the same pipeline step to process the mock datasets (mock4, mock12, and mock16) and compared the predicted representative sequences with the true sequences and their distribution.
  The results (Figure~\ref{fig:figure2}C and \ref{fig:figure2}D) show that the predicted sequence distributions are overall different from the expected ones.
  The variation across datasets indicates that the datasets themselves play a big role in method performance.
  We note that there is no method that outperforms the rest in all datasets (see Supplementary for an extended discussion).
  Based on their slightly better performance on the mock datasets, their de novo error correcting nature and previous independent evaluation~\cite{Nearing2018}, \ac{dada2} and Deblur appear to be the most reliable.
  This is because the open-reference and de novo clustering methods return a much larger number of \ac{otu}s compared to the other pipelines and would affect the accuracy of the network inference step if stringent filtering is not performed.
  \todo[noline]{R2.6}
  Overall, since \ac{dada2} as compared to Deblur, displays better performance on all the mock datasets on the weighted UniFrac metric, we set this tool as the default for the DC step of the pipeline.
  \todo[noline]{R2.10}
  However, if comparison across studies that have sequenced different 16S regions is required, closed-reference and open-reference might be a better option.

  After the denoising, the sequences are subject to Chimera Checking (CC).
  The \ac{micone} pipeline supports two different chimera checking methods, ``uchime-denovo"~\cite{bolyenReproducibleInteractiveScalable2019}, and ``remove bimera"~\cite{Callahan2016}.
  We did not notice any notable difference between the two methods (Figure~\ref{fig:figure_s3}), implying that they identify and remove mostly the same set of sequences as chimeras.
  Since the remove bimera method was originally developed in conjunction with dada2 we use this method as the default.
  The DC step thus results in a reduced set of unique sequences, which will be referred to as representative sequences in the subsequent steps.

  \FloatBarrier

  \subsection*{TA: Taxonomy databases vary widely in taxonomy assignments beyond Order level}

  Taxonomy databases are used to assign taxonomic identities to the representative sequences obtained after the DC step.
  The three 16S taxonomic reference databases used in this study are SILVA~\cite{Quast2012}, \ac{gg}~\cite{DeSantis2006} and \ac{ncbi} RefSeq~\cite{Sayers2009} (Methods).
  These databases vary substantially in terms of taxonomy hierarchies, including species names and phylogenetic relationships~\cite{Balvociute2017}.
  Assignment using a particular database also requires a query tool.
  We used the ``Naive Bayes'' classifier from \ac{qiime2} for the \ac{gg} and SILVA databases and the ``BLAST'' tool (included as a \ac{qiime2} plugin) for the \ac{ncbi} database.
  These tools have been well quantified and optimized~\cite{bokulichOptimizingTaxonomicClassification2018}, hence, we made use of the default parameters in our analyses.

  The representative sequences obtained using the default settings of the DC step were used for taxonomic assignment using the three reference databases.
  Figure~\ref{fig:figure3}A depicts a flow diagram that shows how the top 50 representative sequences (sorted by abundance) are assigned a Genus according to the three databases.
  The different databases lead to assignments that qualitatively display similar distributions. However, the assigned Genus compositions also display clear differences, as does the percentage of unassigned representative sequences (pink).
  Some of the differences in Genus composition have a clear explanation, for example, abundant Genera like Bacteroides and Escherichia are assigned to different representative sequences.
  The large percentage of unassigned sequences is due to the large fraction of the representative sequences assigned to an "unknown" Genus during the assignment process (Methods).

  After the assignment, we performed a pairwise comparison of the similarity between the top 100 assignments (by abundance) from different databases at every taxonomic level (Figure~\ref{fig:figure3}B).
  The comparisons of the assignments below the Order level (Family, Genus, and Species) show less than $45\%$ similarity between any pair of databases.
  This implies that the taxonomy assignments from each reference database are fairly unique.
  The comparison of all assigned genera (Figure~\ref{fig:figure_s4}), instead of just the top 100, contains a higher percentage of mismatches.
  This suggests that, comparatively, the most abundant sequences are more consistently matched to the same taxonomies, at least for the dataset tested in the current analysis.

  To obtain an absolute measure of the accuracy of the taxonomic assignments, we used the representative sequences from the DC step for mock datasets as the query sequences and the expected taxonomic composition as the standard to compare against.
  We used the Bray-Curtis distance metric~\cite{virtanenSciPyFundamentalAlgorithms2020} to calculate the distance between the predicted and expected taxonomic distribution (Figure~\ref{fig:figure3}C).
  We find that none of the databases perform better than the others in absolute terms and that the dissimilarity with the expected composition is high ($>0.5$ for Family and Genus and $>0.9$ for Species), indicating that all the databases have some limitations when trying to recapture the expected taxonomic composition.

  Since no database performs better than others against mock datasets, the choice of which database to use could be driven by other reasons (see Supplementary discussion).
  \todo[noline]{R2.2a}
  One reason to choose a particular database could be the frequency of updates and the potential for future growth.
  Both \ac{gg}, due to its frequent use in the literature~\cite{Balvociute2017}, and \ac{ncbi}, due to its regular revision and maintenance, could be good choices for taxonomy assignment.
  In our default pipeline, we choose \ac{gg} as the default method.

  The TA step results in a taxonomic counts table that is used as input to the subsequent steps of the pipeline.
  \todo[noline]{R2.1b}
  Note that the count tables at different levels can be obtained through aggregation; for example, Genus count tables were obtained by summing up the counts of the lower taxonomy levels (Species and \ac{otu}) that map to the same higher taxonomy level entity.

  \FloatBarrier

  \subsection*{NI: Different network inference methods drastically affect edge-density and connectivity}

   The ten network inference methods we used in this step fall into two groups: the first set of methods (Pearson, Spearman, \acs{sparcc}~\cite{Friedman2012,Watts2018}, and propr~\cite{quinnProprRpackageIdentifying2017}) infer pairwise correlations while the second set (\acs{spieceasi}~\cite{Kurtz2015}, FlashWeave~\cite{tackmannRapidInferenceDirect2019}, \acs{cozine}~\cite{haCompositionalZeroinflatedNetwork2020a}, \acs{harmonies}~\cite{jiangHARMONIESHybridApproach2020}, \acs{spring}~\cite{yoonMicrobialNetworksSPRING2019}, and \acs{mldm}~\cite{Yang2017}) infer direct associations.
   Note that while Pearson and Spearman methods are included in the pipeline for completeness, they tend to generate a large number of spurious edges as they are not intended for compositional datasets.
   Thus, they are not included in subsequent quantitative analyses.

  Filtered (see \ac{op} step in Methods) genus-level counts table obtained using the default settings in the previous steps were used as input for the different network inference algorithms (Figure~\ref{fig:figure4}).
  \todo[noline]{R2.1b}
  Even from a visual inspection (Figure~\ref{fig:figure4}A), one can see that the different networks differ vastly in their edge-density and connectivity, with common edges often displaying inverted signs.

  To quantify the differences between the networks, we analyzed the distribution of common nodes and edges (Figure~\ref{fig:figure4} B and \ref{fig:figure4}C) using UpSet plots~\cite{lexUpSetVisualizationIntersecting2014}.
  The node intersection analysis shows that the networks have $33$ out of $68$ total unique nodes in common and that no network possesses a unique node.
  Edge intersections in contrast show that only $8$ edges (out of $202$ total unique edges) are in common between all the methods and each network has many unique edges.
  These results showed a substantial rewiring of connections in different inferred networks and prompted us to identify associations robust across methods, through consensus algorithms.

  \FloatBarrier

  \subsection*{NI: The scaled-sum consensus method shows high precision on benchmark datasets}

 Inspired by previous approaches~\cite{bustinceFuzzySetsTheir2008,tsarevApplicationMajorityVoting2018}, we developed two methods that take into consideration the evidence offered by each network inference algorithm and generate a consensus network that contains the common edges among the inferred networks.
  \todo[noline]{R1.1}
  \todo[noline]{R1.4}

  Both of our approaches - simple voting (SV) and scaled-sum (SS) - combine appropriately filtered networks inferred from correlation-based and direct association methods (see Methods).
  We chose the scaled-sum method as the pipeline default since this method takes into account the weights of the associations in the determination of the final consensus.
  The pipeline enables the selection of any subset of methods for the consensus calculation. Currently, by default, all direct methods are used, together with \acs{sparcc} and propr.

  Similar to what was done for the previous steps of the pipeline, and in analogy with previous estimations of network inference accuracy~\cite{Kurtz2015,Weiss2016}, we evaluated the network inference algorithms and the final consensus network using synthetic interaction data.
  For this purpose, we generated synthetic interaction data using the ``NorTA''~\cite{Kurtz2015} and ``seqtime''~\cite{faustSignaturesEcologicalProcesses2018} methods (see Methods).
  For each method, an \ac{otu} counts table was generated based on the selected parameters and abundance distributions.
  This counts table was used as the input to the \ac{micone} pipeline to generate predicted associations.
  The interaction network used to generate the counts table was used as the source of true interactions to calculate the precision (Figure \ref{fig:figure5}) and sensitivity (Figure \ref{fig:figure_s5} and Figure \ref{fig:figure_s6}) for each network inference algorithm.
  As shown in Figure \ref{fig:figure5} the consensus algorithm, especially the scaled-sum method, captures true associations with high precision (through the removal of edges that are either not present in most of the inference methods or whose association strength is low across methods).
  Overall, the scaled-sum method for $p=1.000$ performs the best (precision = $1.000$ for both NorTA and seqtime).
  The scaled-sum method for $p=0.333$ (default option in the pipeline) shows a high precision ($0.956$ with NorTA; $0.688$ with seqtime), without displaying significant reduction in sensitivity (Figure~\ref{fig:figure_s5} and Figure~\ref{fig:figure_s6}).
  However, if higher precision is required $p>0.5$ can be considered.
  Therefore, the consensus networks provide the means to obtain a short list of associations that would have a high likelihood of being present in the real association network.

  \FloatBarrier

  \subsection*{Impact of different pipeline steps on co-occurrence networks}

  In order to analyze the effect of different processing methods on the inferred co-occurrence networks (before consensus estimation), we generated networks using all possible combinations of methods and quantified the variability due to each choice (Figure \ref{fig:figure6}A).
 This was achieved by building a linear model of the edges of the network as a function of the various steps in the pipeline workflow (see Methods).
  Figure \ref{fig:figure6}A, shows the percentage of total variation among the co-occurrence networks due to the different steps of the pipeline.
  The \ac{ta} step, or more specifically the choice of 16S reference database, contributes the most ($65.4\%$) to the variation in the networks, followed by the \ac{op} step ($26.8\%$).
  This result highlights the importance of the taxonomy assignment step in the 16S data analysis workflow, implying that a change in the reference database will result in drastically different inferred networks.
  This is likely due to the differential assignment of representative sequences to taxonomic entities (Figure~\ref{fig:figure3} and Figure~\ref{fig:figure_s4}), which drastically alter the nodes and hence the underlying network topology.

  The effects of the different steps of the pipeline on the inferred networks can be visualized through dimensionality reduction.
  The PCA in Figure \ref{fig:figure6}B shows all the above networks, colored by the tools used in the DC, TA, OP, and NI steps in each subfigure.
  The major effect of the TA step choice, shown before in Figure \ref{fig:figure6}A, is also reflected in the PCA plot, where networks segregate based on the database used (Figure~\ref{fig:figure6}B and Figure~\ref{fig:figure_s1}).
  Additionally, the plot also shows that the variation between the networks decreases when the low abundance \ac{otu}s are removed from the network.
  It is also evident that in the NI step, some networks, especially those inferred using the direct association network inference methods, are much closer in the PCA plot regardless of the reference database used.
  These results suggest that the most important criterion for accurate comparative analysis of co-occurrence networks is the taxonomy reference database followed by the level of filtering of the taxonomy tables and the network inference algorithm used.

  \FloatBarrier

  \subsection*{The default pipeline}

  The systematic analyses in the previous sections illustrate that the choice of tools and parameters can have a big impact on the final consensus co-occurrence network.
  However, the mock communities and synthetic data provide an opportunity to select combinations of tools that yield the most accurate and robust results.
  As highlighted in the above sections for individual steps, we propose a set of tools and parameters as the defaults for the pipeline (Table~\ref{tab:micone_tools}).

  Figure~\ref{fig:figure7} shows the co-occurrence networks inferred for the healthy subjects (control) and subjects with autism specific disorder (ASD) in the fecal microbiome transplant study~\cite{Kang2017} (constructed using the default tools and parameters from Table~\ref{tab:micone_tools}).
  This figure demonstrates a typical use case of comparative analysis of networks using the \ac{micone} pipeline.
  As a consequence of using the consensus network algorithm, the final co-occurrence networks are sparse and can be visually compared and examined.

  The analysis of the rewiring of associations in the ASD samples with respect to the control provides a guide for the identification of key genera that could be linked to dysbiosis.
  \todo[noline]{R2.14}
  We observed 22 unique links in the network for control samples, 12 unique links in the network for ASD subjects, and 7 edges in common between the two networks.
  Although these unique associations do not imply actual interactions, they can still serve as potential starting points for literature surveys and further experimental exploration of mechanistic processes underlying dysbiosis.
  For example, \textit{Prevotella} and \textit{Porphyromonas}, genera previously implicated in ASD~\cite{Kang2017,hoGutMicrobiotaChanges2020} and cognitive impairment~\cite{chiPorphyromonasGingivalisInducedCognitive2021} display modified connectivity in our network, suggesting that the observed associations may be relevant for understanding the role of these bacteria in disease.
  Additional visualization and comparison of networks can be performed using the \acf{mind}~\cite{huResourceComparisonIntegration2022}.

  Figure~\ref{fig:figure_s7} shows a sensitivity analysis in which we compared the default network against networks generated by altering one of the steps of the pipeline relative to the default.
  This result, both visually (Figure~\ref{fig:figure_s7} A), and quantitatively (Figure~\ref{fig:figure_s7} B)  suggests that the most significant changes occur when the \ac{op} or \ac{ta} steps are changed from the default value.

% DISCUSSION
\section*{Discussion}

  \subsection*{Why \ac{micone}?}

  A myriad of tools and methods have been developed for different parts of the workflow for inference of co-occurrence networks from 16S rRNA data.
  Our analyses have shown that networks generated using different combinations of tools and approaches can be substantially different from each other, highlighting the need for a clear evaluation of the source of variability and for tools that provide the most robust and accurate results.
  Our newly developed software, \ac{micone}, is a customizable pipeline for the inference of co-occurrence networks from 16S rRNA data that enables users to compare networks generated by multiple possible combinations of tools and parameters.
  Importantly, in addition to revisiting the test cases presented in this work, users will be able to explore the effect of various tool combinations on their own datasets of interest.
  The \ac{micone} pipeline has been built in a modular fashion; its plug-and-play architecture enables users to add new tools and steps, either using existing packages that have not been examined in the present work or those developed in the future.
  The \ac{micone} Python package provides functions and methods to perform a detailed analysis of the count matrices and the co-occurrence networks.
  The inferred networks are exported to a custom JSON format (see Supplementary) by default, but can also be exported to Cytoscape~\cite{shannonCytoscapeSoftwareEnvironment2003}, GML~\cite{himsoltGMLPortableGraph2010}, and many other popular formats via the Python package.

  While several tools/workflows such as \ac{qiime2}~\cite{bolyenReproducibleInteractiveScalable2019} and NetCoMi~\cite{peschelNetCoMiNetworkConstruction2020} can be used to generate co-occurrence networks from 16S sequencing data, no single tool exist that integrates the complete process of inferring microbial interaction networks from 16S sequencing reads.
  \ac{micone} is unique as it offers this functionality packaged in a workflow that can be run locally, on the compute cluster, or in the cloud.

  \subsection*{The default pipeline and recommended tools}

  Through \ac{micone}, in addition to transparently revealing the dependence of co-occurrence networks on tool and parameter choices (see Discussion in Supplementary Text for details on the DC, TA and OP steps), we have taken advantage of our spectrum of computational options and the availability of mock and synthetic datasets, to suggest a default standard setting.
  \todo[noline]{R1.3}
  Additionally, we have developed a consensus approach, that can reliably generate networks that are fairly robust across multiple tool choices.
  An important caveat related to these results is that due to the lack of a universal standard for microbial interaction data, our conclusions are based on the specific datasets used in our analysis.
  While our analysis is based on several mock and synthetic datasets that cover a diverse range of abundance distributions and network topologies, datasets that have drastically different distributions may require a re-assessment of the best settings.

  The networks generated by different network inference methods show considerable differences in edge-density and connectivity, partially due to the underlying assumptions regarding sparsity, distribution, and compositionality.
  To address this issue, we have developed two consensus algorithms (simple voting and scaled-sum method) that generate networks whose links have evidence based on multiple inference algorithms.

 We find that the scaled-sum method performs the best on synthetic datasets, and is therefore chosen as the default for the \ac{ni} step of the pipeline.
 Notably, the consensus network displays a higher precision and returns a concise list of robust associations which represent a valuable set for experimental validation follow-up.

  \subsection*{Future directions}

  Future work building upon our current results could enhance the network inference process in multiple ways.
  The current analyses make use of one fecal microbiome transplant dataset with healthy and ASD samples, three mock community datasets, and several datasets generated by two synthetic interaction methods.
  Incorporating datasets from a broad spectrum of biomes with varying microbial distributions into \ac{micone} will likely increase the robustness and generalizability of the results from these analyses.

  The network analyses in this study are primarily at the Genus level, wherein the lowest resolution of a node is a Genus and if an entity cannot be resolved to the Genus level, the next lowest taxonomic level is used (for example, Family).
  \todo[noline]{R2.1c}
  As a consequence, two entities belonging to the same lineage where one entity is resolved to the Genus level and another is resolved to the Family level are treated as two different nodes in the network.
  Thus, the development of a metric of overlap to compare nodes with shared lineages within and across networks could enable more biologically and phylogenetically relevant comparisons.

  Although direct comparisons between co-occurrence networks and directly measured interactions are difficult to interpret and highly debated~\cite{hiranoDifficultyInferringMicrobial2019,gobernaCautionaryNotesUse2022}.
  \todo[noline]{R2.2}
  Further, benchmarking of co-occurrence networks could also be pursued through the use of literature-based interactions~\cite{lima-mendezDeterminantsCommunityStructure2015a} or biological benchmark interaction data~\cite{sungGlobalMetabolicInteraction2017a}.
  Additionally, \ac{micone} could be extended to enable the processing of metagenomics sequencing data, facilitating the analysis of a much larger and diverse range of datasets and domains of life.

  Although in the current analysis, we have only used default parameter values recommended by the tool creators, the \ac{micone} pipeline could be used in the future to explore any combinations of parameters and to optimize these values for improved network inference.
  \todo[noline]{R1.3}
  Overall, there likely is no ``best method'' for the various steps of 16S data analysis, and hence, \ac{micone} is intended to help researchers to identify the methods and algorithms that are most suitable for their datasets in an easy-to-use and reproducible manner.

  We envision that \ac{micone}, and its underlying tools and databases, will be increasingly useful for building large comparative analyses across studies.
  It enables rapid, configurable, and reproducible inference of microbial networks and furthers the formulation of hypotheses about the role of these interactions on community composition and stability.
  These comparative analyses will require coupled network analysis and visualization tools (such as \ac{mind}~\cite{huResourceComparisonIntegration2022}) and need systematic access to datasets, shared in accordance with FAIR standards~\cite{pachecoFAIRRepresentationsMicrobial2022}.

% METHODS
\section*{Materials and Methods}

  \subsection*{16S rRNA sequencing datasets}
  \vspace{-5mm}
  This study utilized two types of 16S rRNA sequencing datasets: biological datasets and mock/synthetic datasets.
  \todo[noline]{R2.4}
  Biological datasets are collections of sequencing reads obtained from naturally occurring microbial community samples.
  The current analysis used stool samples from a fecal microbiome transplant study of autism~\cite{Kang2017} as the biological dataset.
  This dataset was chosen because the sequences were easily accessible on Qiita~\cite{qiita} and optimally pre-processed according to the \ac{emp}~\cite{Thompson2017} protocol, allowing them to be used directly as input to the \ac{micone} pipeline.
  The study was composed of multiple sequencing runs.
  The runs that contained paired-end reads (run 2 (10M reads), run 3 (750K reads) and run 4 (16M reads)), were downloaded from Qiita~\cite{qiita} (study ID 10532) and used as input sequences for the \ac{micone} pipeline.
  Sequences from both control (212 samples including neurotypical and donors) and autism spectrum disorder (ASD) (126 samples) patients were included in the analyses.
  All the network analyses in the study, unless explicitly mentioned, were performed on the healthy and ASD samples in the fecal microbiome transplant study.
  The mock community 16S datasets are experimental sequencing data obtained for artificially assembled collections of DNA of species in known proportions.
  The mock datasets used for this study, obtained from mockrobiota~\cite{Bokulich2016}, are labeled mock4, mock12, and mock16.
  The mock4 community is composed of 21 bacterial strains.
  Two replicate samples from mock4 contain all species in equal abundances, and two additional replicate samples contain the same species in unequal abundances.
  The mock12 community is composed of 27 bacterial strains that include closely related taxa with some pairs having only one to two nucleotide differences from one another.
  The mock16 community is composed of 49 bacteria and 10 Archaea, all represented in equal amounts.

  \subsection*{\ac{micone}}
  \vspace{-5mm}
  The flowchart describing the workflow of \ac{micone} (\acl{micone}), our complete 16S data-analysis pipeline, is shown in Figure \ref{fig:figure1}.
  The pipeline integrates many publicly available tools as well as custom R or Python modules and scripts to extract co-occurrence associations from 16S sequence data.
  Each of these tools corresponds to a distinct module that recapitulates the relevant analyses.
  All such individual modules are available as part of the \ac{micone} package.
  The inputs to the pipeline by default are raw untrimmed 16S rRNA sequence reads, but the software can be alternatively configured to use trimmed sequences, \ac{otu} tables and other types of intermediate data (see documentation).
  The configuration and modular nature of the \ac{micone} package enables users to start and end the pipeline at any point in the workflow, and to run parts of the pipeline in isolation.
  The pipeline supports both paired-end and single-end reads, and additionally supports independently processing reads from multiple runs and merging the OTU tables in the DC step.
  The final output of the pipeline is the inferred network of co-occurrence relationships among the microbes present in the samples.

  The \ac{micone} pipeline provides both a Python API together with a command-line interface and only uses a single configuration file (\texttt{nextflow.config}) to encode the configuration parameters.
  The \ac{micone} Python API provides several \ac{otu} table and network-related functions and methods, enabling detailed comparison of counts tables and inferred networks if desired.
  Exploring the effects of these combinations of methods on the resultant networks is difficult and inconvenient since different tools differ in their input and output formats and require interconversions between the various formats.
  The pipeline facilitates this comparative exploration by providing a variety of modules for interconversion between various formats, and by allowing for easy incorporation of new tools as modules.
  It also contains helper functions that can help in parsing taxonomies and communicate with the \ac{ncbi} taxonomy database to query taxonomy by name or taxonomic IDs.
  The configuration file along with the run file (\texttt{main.nf}) lists the inputs, output, and the steps to be performed during runtime, along with the parameters to be used (if different from defaults) for the various steps.
  The default settings of the pipeline are shown in Table~\ref{tab:micone_tools} (with default parameter values shown in Tables S2-S6 at \href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper}).
  Since the entire pipeline run is stored in the form of a text file (the configuration file), subsequent runs are highly reproducible and changes can be easily tracked using version control.
  The pipeline makes use of the nextflow workflow manager~\cite{Tommaso2015} under the hood, making it readily usable on the local machine, cluster, or cloud with minimal configuration change.
  It also allows for automatic parallelization of all possible processes, both within and across samples.
  The pipeline is designed to be modular: each tool or method is organized into modules that can be easily modified or replaced.
  This modular architecture simplifies the process of adding new tools (refer to the modules section in the \ac{micone} documentation).
  \todo[noline]{R2.11}
  The main components of the pipeline are detailed in the subsequent sections.

  \subsection*{Sequence Processing (SP)}
  \vspace{-5mm}
  This module deals with processing the raw multiplexed 16S sequence data into demultiplexed, quality-controlled, trimmed sequences.
  It consists of the demultiplexing and trimming processes.
  The demultiplexing process deals with separating the multiplexed sequences into individual samples based on barcodes.
  The trimming process handles the quality control steps such as trimming adapters and low-quality nucleotide stretches from the sequences.
  The parameters and tools in this process are fixed and are not available for user customization.
  The various tools used for the processes were adapted from \ac{qiime2} v2021.8.0~\cite{bolyenReproducibleInteractiveScalable2019}.
  The list of tools used in this step, along with their modules and references are provided in Table~\ref{tab:micone_tools}.

  \subsection*{Denoising and Clustering (DC)}
  \vspace{-5mm}
  This module deals with processing the quality-controlled, trimmed 16S sequence data into \ac{otu} or \ac{esv} count tables.
  It consists of the following processes: denoising (or clustering) and chimera checking.
  The denoise/cluster process handles the conversion of the demultiplexed, trimmed sequences into \ac{otu} or \ac{esv} count tables (some methods, like closed reference and open reference clustering, make use of a taxonomy reference database for clustering).
  The chimera checking process handles the removal of chimeric sequences created during the \ac{pcr} step.
  The output of this module is a matrix of counts, that describes the number of reads of a particular \ac{otu} or \ac{esv} (rows of the matrix) present in each sample (columns of the matrix).
  The options currently available in the pipeline for denoising and clustering are: open reference clustering, closed reference clustering and de novo clustering methods from the vsearch plugin of \ac{qiime2} v2021.8.0~\cite{bolyenReproducibleInteractiveScalable2019} and denoising methods from \ac{dada2} v1.14~\cite{Callahan2016} (from the \ac{dada2} R package) and Deblur v1.1.0~\cite{Amir2017} (from the deblur plugin of \ac{qiime2}).
  The quality filtering and chimera checking tools are derived from those used in \ac{qiime2} v2021.8.0 (uchime-denovo method) and \ac{dada2} (remove bimera method).
  The list of tools used in this step, along with their modules and references are provided in Table~\ref{tab:micone_tools}.

  For the UniFrac analysis in Figure~\ref{fig:figure2}, we had set a count threshold of 10, such that if the count of the representative sequences in a particular sample is less than the threshold, it is omitted from the analysis.
  Additionally, for Figure~\ref{fig:figure2}C and \ref{fig:figure2}D, the expected sequences from the mock communities were trimmed to the V4 region before being subject to UniFrac analyses.

  \subsection*{Taxonomy Assignment (TA)}
  \vspace{-5mm}
  This module deals with assigning taxonomies to the representative sequences (\ac{otu}s or \ac{esv}s).
  In order to assign taxonomies to a particular sequence, a taxonomy database and a query tool are necessary.
  The taxonomy database contains a collection of 16S sequences of microorganisms and the query tool allows one to compare a sequence of interest to all the sequences in the database to identify the best matches.
  Finally, a consensus method is used to identify the most probable match from the list of best matches.
  The pipeline incorporates \ac{gg} 13\_8~\cite{DeSantis2006} (99\% identity), SILVA 138~\cite{Quast2012} (99\% identity) and the \ac{ncbi} (16S RefSeq as of Oct 2021)~\cite{Sayers2009} databases for taxonomy assignment.
  SILVA and \ac{gg} are two popular 16S databases used for taxonomy identification and the \ac{ncbi} RefSeq nucleotide database contains 16S rRNA sequences as a part of two BioProjects - 33175 and 33317.
  The three databases vastly differ in terms of their last update status - \ac{gg} was last updated on May 2013, SILVA was last updated on August 2020 at the time of writing and \ac{ncbi} is updated regularly as new sequences are curated.
  These databases were downloaded and built using the RESCRIPt \ac{qiime2} plugin~\cite{iiRESCRIPtReproducibleSequence2021}.
  The Naive Bayes classifier and the \ac{ncbi} blast used as the query tools in this study were from the \ac{qiime2} package and the parameters used were the defaults of the package.
  The consensus algorithm used is the default method used by the classifiers in \ac{qiime2}.
  During the assignment, a representative sequence might be assigned an "unknown" Genus for one of two reasons: the first is if the taxonomy identifier associated with the sequence in the database did not contain a given Genus; the second, more likely reason, is that the database contains multiple sequences that are very similar to the query (representative) sequence and the consensus algorithm (from \ac{qiime2}) is unable to assign one particular Genus at the required confidence.
  The assignments in SILVA were originally substantially different from the other two databases ($40\%$ mismatch) even at the Phylum level.
  However, this was corrected via minor adjustments to the taxonomic names, such as changing Bacteroidota to Bacteroidetes in the SILVA Phylum assignments.
  The full list of changes can be found in \texttt{figure4ab\_data.py} in the data and scripts  repository.
  The list of tools used in this step, along with their modules and references are provided in Table~\ref{tab:micone_tools}.

  \subsection*{OTU and ESV Processing (OP)}
  \vspace{-5mm}
  This module deals with normalization, filtering, forking, grouping, and applying transformations to the \ac{otu} or \ac{esv} counts matrix.
  Normalization of the count matrix involves converting the count matrix of read counts into a count matrix containing relative abundances.
  The module also supports rarefaction, which is a normalization technique used to overcome the bias that might arise due to variable sampling depth in different samples.
  \todo[noline]{R2.5}
  This is performed either by sub-sampling or by normalization of the matrix to the lowest sampling depth~\cite{Weiss2015}.
  However, due to the potential biases and false positives~\cite{mcmurdieWasteNotWant2014,chaoCoveragebasedRarefactionExtrapolation2012} that might arise during the process, the rarefaction module is disabled by default and can be enabled in the configuration if needed.
  Hence, although the pipeline supports normalization and rarefaction, these modules are turned off by default, and the analyses reported in the paper do not normalize or rareify the counts matrices.
  However, most of the network inference methods perform normalization and other transformation operations on the counts matrix as a part of their workflow.
  Filtering, is performed to remove samples or features (\ac{otu}s or \ac{esv}s) from the counts matrix that are sparse.
  By default, when the OP module is ``on'', the samples are filtered out if the total reads in a sample are less than 500 and features are filtered out if the relative abundance is less than 1\%, prevalence (percentage of samples containing feature) is less than 5\% and count sum across all the samples is less than 100.
  When the OP module is ``off'', the filtering is still performed but threshold parameters are much more relaxed.
  The parameters used are given in Table~\ref{tab:micone_tools}.
  The forking operation splits the count matrix into multiple matrices based on sample metadata column, this is useful for example to compare case vs. control.
  The group operation transforms the \ac{otu} or \ac{esv} count matrix into a taxonomic count matrix at the requested level by adding up counts that map to the same taxonomy and is carried out at the end of the OP step.
  \todo[noline]{R2.1b}
  Finally, transformations are performed in order to correct for and overcome the compositional bias that is inherent in the counts matrix (in the analysis performed in the study these were disabled and directly handled by the network inference algorithm).
  All the modules in this step were implemented using functions from the \href{http://biom-format.org/}{biom-format} Python package~\cite{mcdonaldBiologicalObservationMatrix2012}.

  \subsection*{Network Inference (NI)}
  \vspace{-5mm}
  This module deals with the inference of co-occurrence associations from the processed \ac{otu} or \ac{esv} counts matrix.
  The input count matrices are collapsed to the Genus level (or any other required taxonomy level) using the group module at the OP step.
  These collapsed matrices are used as input to the network inference methods to produce association matrices at the appropriate taxonomy level.
  These associations can be represented as a network, with nodes representing the taxonomies of the microorganisms and edges representing the associations between them.

  The pipeline includes 4 methods for pairwise correlation metrics, and 6 methods for direct association metrics (refer to Table~\ref{tab:micone_tools}).
  Pairwise correlation methods involve the calculation of the correlation coefficient between each pair of nodes (taxonomic entity like Genera) leading to the inclusion of spurious indirect connections.
  On the other hand, direct association methods use conditional independence to avoid the detection of correlated but indirectly connected taxonomic entities~\cite{Kurtz2015,Menon2018}.
  A null model is created by re-sampling and bootstrapping the counts matrix and recalculating the correlations (see next section for details on network analysis and statistics).
  These bootstrapped association matrices are used to calculate the significance of the inferred correlations by calculating the p-values against this null model~\cite{Watts2018}.
  Brown's p-value merging method~\cite{brown_400_1975} is used for combining p-values from the pairwise correlations methods to obtain a consensus p-value, which can be used to filter for significance.
  The bootstrapping and p-value calculations are only performed on the correlations-based methods.
  In the final module of this step, the consensus algorithms are used to create the final consensus network using associations from all the network inference methods (except Pearson and Spearman, by default).
  The outputs of this step are co-occurrence association networks encoded in the JSON format (refer to Supplementary section) and which can also be exportable to a variety of network formats.
  The list of tools used in this step, along with their modules and references are provided in Table~\ref{tab:micone_tools}.

  \subsection*{Consensus network and p-value merging}
  The consensus methods combine networks inferred from both correlation-based and direct association methods.
  \todo[noline]{R1.1}
  \todo[noline]{R1.4}
  First, for the correlation-based methods, we calculate p-values using null models and then merge the p-values using Brown's p-value merging method~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016,faustCoNetAppInference2016}.
  \todo[noline]{R2.2}
  \todo[noline]{R2.3}
  Second, we filter all the inferred networks based on an association strength threshold of 0.1 and a p-value cutoff of 0.05.
  Finally, we apply the consensus algorithms we have developed on these filtered networks.
  These steps are elaborated on in the subsequent sections.

 \subsubsection*{Notation}
  \vspace{-5mm}
 This section defines the notation used below to describe the consensus network algorithm used in the \ac{micone} pipeline.
 Note that all networks to be compared were updated to have the same number of nodes.

  $w$, the number of co-occurrence networks to be integrated into the consensus network (by default, is equal to the total number of network inference methods excluding Spearman and Pearson, 8)

  $q$, the number of unique nodes across all $w$ co-occurrence networks

  $N^i$, the matrix of edge weights for the $i^{th}$ co-occurrence network.
  This is a $q \times q$ matrix, where $i \in \{1,\dots,w\}$.
  $N^i_{a,b}$ represents edge $(a,b)$ in network $i$

  $P^i$, the matrix of p-values for all edges of the $i^{th}$ co-occurrence network.
  This is a $q \times q$ matrix, where $i \in \{1,\dots,w\}$

  $\bar{N}^i$, the ``flattened'' version of the adjacency matrix $N^i$ into a $q^2 \times 1$ column vector, where all columns are stacked onto each other into a $q^2$ long vector.
  Element $\bar{N}^i_j$ corresponds to the $j^{th}$ edge in the $i^{th}$ network.

  $\bar{P}^i$, the ``flattened'' version of the adjacency matrix $P^i$ into a $q^2 \times 1$ column vector, where all columns are stacked onto each other into a $q^2$ long vector.

  \subsubsection*{Bootstrapping and p-value calculation}
  \vspace{-5mm}
  For all correlation-based methods $k \le w$, $1000$ permutations of the original \ac{otu} counts data were generated~\cite{Watts2018}.
  The correlations in the permuted \ac{otu} tables are recalculated using the different correlation-based algorithms.
  Finally, the p-value is determined based on how often a more extreme association is observed for randomly permuted data.
  Note that, all the direct association-based methods used in the study have their own regularization methods built in and hence do not need to undergo bootstrapping.

  \subsubsection*{p-value merging}
  \vspace{-5mm}
  The next step in the consensus algorithm workflow is to merge the p-values for the networks generated by the correlation-based methods.
  \todo[noline]{R1.1}
  This step is performed using the Brown's p-value merging method~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016,faustCoNetAppInference2016}.

  As described in more detail in the Supplementary section and in the original reference~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016}, the final combined p-value is given by:
  \begin{equation}
    \begin{aligned}
        & \hat{P}_j = 1.0 - \Phi_{2f}\left( \psi / c \right) \\
        \text{where},~ &\psi = -2 \sum_{i=1}^k \log(\bar{P}^i_j) ~~~\text{and}~~~ \Phi_{2f} = \mathrm{CDF}\left( \chi^2_{2f} \right)
    \end{aligned}
    \label{eqn:pvalue-combined}
  \end{equation}
  where, $\hat{P}_j$ is the combined p-value for the edge $j$, $f$ is the number of degrees of freedom, and $c$ is a scale factor.

  Note that we do not use Pearson and Spearman methods in the p-value merging step to determine the consensus network.
  These methods are only used for demonstration and comparison.
  The combined p-values are used to threshold for significance right before the consensus algorithm is applied to the inferred networks.

  \subsubsection*{Consensus methods}
  \vspace{-5mm}
  The consensus algorithm was designed to increase the precision (number of true positives) at the end of the network inference step.
  For this purpose, we developed two simple algorithms that combine the edges reported by the different network inference tools.
  Both the algorithms make use of a user-defined parameter $p$ ($0 \leq p \leq 1$), in order to threshold the edges from the individual methods.
  The inputs to both the algorithms are the co-occurrence networks (association matrices) $\bar{N}^i$ (flattened version of $N^{i}$) generated by each method $i$, and the threshold parameter $p$.
  Here, the $\bar{N}^{i}$ each have the same set of nodes $q$ and only differ by the value of the association inferred between every pair.
  Networks that do not have a particular node, are updated such that the node is added as an isolated component.
  In this manner, $\bar{N}^{i}_j$ represents edge $j$ in network $i$.

    Note that the consensus method is only used to filter relevant interactions.
    If a given pair of nodes is inferred to have edges that satisfy the consensus requirements, all corresponding edges from the $w$ networks will be returned by the algorithm, as a multigraph.
    Based on this approach, \ac{micone} reports as the default output, the consensus network where each edge is annotated with weights (correlations for the correlation-based methods and direct associations for the other methods) from all the methods used in the consensus algorithm.

  \paragraph*{Algorithm 1 - Simple voting:}
  The simple voting method performs a voting-based consensus to determine whether an edge will exist between a given node-pair in the final consensus network~\cite{bustinceFuzzySetsTheir2008,tsarevApplicationMajorityVoting2018}.
  For each pair of nodes, we determine the number of network inference methods that report an edge $j$ between them, i.e. $\bar{N}^{i}_{j}, \forall i \in \{1,\dots,w\}$.
  Each node-pair will have an edge in the final consensus network if the number of reported edges is larger than the threshold (Equation~\ref{eqn:simple-voting2}).

 The number of reported edges is computed as follows:

 For each edge $j$, we obtain $M_j$ which represents the number of networks in which edge $j$ is reported.
 Formally, $M_j$ is calculated as the following function:
  \begin{equation}
      M_j = f(g(\bar{N}^{i=1}_j), \dots, g(\bar{N}^{i=w}_j)) \\
    \label{eq:simple-voting}
  \end{equation}

where, $g$ and $f$ are defined as follows:

  \begin{equation*}
    g(x) =
    \begin{cases}
       & 0, \text{ if } x=0, \\
       & -1, \text{ if } x<0, \\
       & 1, \text{ if } x>0
    \end{cases}
  \end{equation*}
  and
  \begin{equation*}
    f(x_1,\dots,x_w) = max \left( \#(i \mid x_i=-1),\#(i \mid x_i=1)
       \right)
  \end{equation*}
  where, $\#$ refers to the cardinality of the set.

 The edge $j$ is selected to be present in the final consensus network if the number of networks in which $j$ appears is greater than a threshold, i.e:
  \begin{equation}
     M_j \geq \lfloor p \times w \rfloor
    \label{eqn:simple-voting2}
  \end{equation}
  where, $p$ is the user-defined threshold parameter.

  The simple voting method returns the union of the networks when $0 \leq p \leq \frac{1}{w}$ and will return the intersection when $\frac{(w - 1)}{w} \leq p \leq 1$.
  In general, if $\frac{(n - 1)}{w} \leq p \leq \frac{n}{w}$, this algorithm will report an edge in the consensus network when at least $n$ network inference methods report this edge.

  \paragraph*{Algorithm 2 - Scaled-sum method:}

  This algorithm generates a consensus network based on the sum of all edges (weights of associations) reported between a pair of nodes~\cite{bustinceFuzzySetsTheir2008,tsarevApplicationMajorityVoting2018}.
  Since in generating a consensus network using this method we sum the edges reported by direct association methods with those from correlation-based methods, summing of the edges is preceded by a pre-processing step, in which all networks are re-scaled.

  First, the network generated by each network inference method ($\bar{N}^i$) is re-scaled into a normalized version ($\bar{S}^i$), as follows:
  \begin{equation}
    \bar{S}^{i} = \frac{\bar{N}^{i}}{\max(\mid \bar{N}^{i} \mid)},~~\forall i \in {1, \dots, w}
    \label{eqn:scaled-sum-rescaling}
  \end{equation}

    In this way, it is guaranteed that $\max(\mid \bar{S}^i \mid) = 1$.

  Next, for each edge $j$, we sum the weights of all reported edges from the different networks.
  \begin{equation}
    s_j = \sum_{i=1}^{w} \bar{S}^i_j
    \label{eqn:scaled-sum}
  \end{equation}

  An edge $j$ will be included in the consensus network if $s_j$ passes a threshold.
  \begin{equation}
    \mid s_j \mid > (w - 1) \times p
    \label{eq:scaled-sum2}
  \end{equation}

  The advantage of this method over the simple voting method is that it also takes into account the strength of the association reported for that particular node in the inferred networks.

  \subsection*{Network variability}

  \subsubsection*{Notation}
  \vspace{-5mm}
  This section defines the notation used for the network variability analysis performed for Figure~\ref{fig:figure6}.
  \todo[noline]{R1.2}
  \todo[noline]{R1.8}
  \todo[noline]{R2.1b,d}

  $W$, the number of co-occurrence networks generated from all possible combinations of tools and parameters in the workflow.
  Note that this is different from $w$, which counted only the different network inference modules.

  $Q$, the number of unique nodes across all $W$ networks.

  $N^i$, the edge weights of the $i^{th}$ co-occurrence network represented as a $Q \times Q$ adjacency matrix, where $i \in {1, \dots, W}$.
  $N^i_{a,b}$ represents the edge $(a,b)$ in network $i$

  $\bar{N}^i$, the ``flattened'' version of the adjacency matrix $N^i$ into a $Q^2 \times 1$ column vector, where all columns are stacked onto each other into a $Q^2$ long vector.

  \subsubsection*{Principal Component Analysis and variability calculation}
  \vspace{-5mm}
   In order to compare across different networks and calculate the degree of variability induced by the choice of different modules, we organized multiple networks into a single mathematical structure that we could use for linear regression.
   First, we obtained the co-occurrence network $\bar{N}^i$ for each of the $W$ possible tool and parameter combinations in the workflow.
   We then constructed a matrix $\mathbf{\bar{N}}$ whose $i^{th}$ column is the flattened version of the $i^{th}$ network, i.e. the column vector $\bar{N}^i$.
   Therefore, $\bar{N}^i_j$ is the weight of edge $j$ in the network $i$.
   $\bar{N}^i_j$ is assigned a value of 0 if edge $j$ did not exist in network $i$ but was present in one of the other networks.
   Note that row $j$ of $\mathbf{\bar{N}}$, $\bar{N}_j$ is the vector that encodes the values of edge $j$ across all the networks.
  \todo[noline]{R1.5}
  \todo[noline]{R2.1b}

  \begin{equation*}
   \mathbf{\bar{N}} =
      \begin{bmatrix}
       \bar{N}^1_1 & \bar{N}^2_1 & \cdots  & \bar{N}^{W}_1  \\
       \bar{N}^1_2 & \bar{N}^2_2 & \cdots  & \bar{N}^{W}_2 \\
       \vdots & \vdots & \vdots  & \vdots \\
       \bar{N}^1_{Q^2} & \bar{N}^2_{Q^2} & \cdots  & \bar{N}^{W}_{Q^2}
      \end{bmatrix}
  \end{equation*}

  To infer the variability contributed due to the different steps in the pipeline we can perform a linear regression on each edge in $\mathbf{\bar{N}}$ and a subsequent ANOVA to extract the within-group variances.
  One issue with this approach is that the possibility of correlations existing between the edges of the network could lead to inaccurate estimates of the variance if a linear model were used to directly model the relationships between edges and steps in the workflow.
  In order to remedy this issue, we performed a PCA (Principal Component Analysis) on the matrix $\mathbf{\bar{N}}$ to obtain the $\mathbf{C}$ matrix ($W \times c$) of components for each network, such that we reduce the dimensions from the $Q^2$ dimensional edge space to a $c$ dimensional component space.

  We then use linear regression to express each component $C_j$ (where $j \in 1:c$) as a linear function of categorical variables that describe the possible options in each of the steps of the pipeline.

  In particular, we infer parameters $\alpha_j$ such that:
  \begin{multline}
  C_j = &\sum_{i=1}^5 \left( \alpha^{DC(i)}_j \delta^{DC(i)}_j \right) +
         \sum_{i=1}^2 \left( \alpha^{CC(i)}_j \delta^{CC(i)}_j \right) +
         \sum_{i=1}^3 \left( \alpha^{TA(i)}_j \delta^{TA(i)}_j \right) +
         \sum_{i=1}^2 \left( \alpha^{OP(i)}_j \delta^{OP(i)}_j \right) + \\
         &\sum_{i=1}^{10} \left( \alpha^{NI(i)}_j \delta^{NI(i)}_j \right) +
         \epsilon_j
  \end{multline}

   where, $\alpha_i$ are the coefficients of the regression, $\epsilon_i$ are the residuals and $\delta_i$ are the indicator variables that correspond to the processes utilized in the pipeline used to create the network $N_i$; for example, $\delta^{DC(1)}_i = 1$ if the DC(1) process was used in the generation of the network $N^i$.

   Here,
   \begin{enumerate}
     \item $DC(i) \in$ \{CR, OR, DN, D2, DB\}
     \item $CC(i) \in$ \{remove bimera, uchime-denovo\}
     \item $TA(i) \in$ \{NaiveBayes(GG), NaiveBayes(SILVA), BLAST(NCBI)\}
     \item $OP(i) \in$ \{Filter(on), Filter(off)\}
     \item $NI(i) \in$ \{\acs{sparcc}, propr, Spearman, Pearson, \acs{spieceasi}, \acs{cozine}, \acs{harmonies}, \acs{spring}, \acs{mldm}, FlashWeave\}
   \end{enumerate}

   The variance contributed by each step of the pipeline was calculated for every component in $\mathbf{C}$ matrix through ANOVA using the Python statsmodels~\cite{seaboldStatsmodelsEconometricStatistical2010} package and is shown in Figure~\ref{fig:figure6}A.
  \todo[noline]{R1.6}
  The total variance for the network was calculated by adding the variances for each connection and normalizing with the degrees of freedom.
  The merged network table $\mathbf{\bar{N}}$ was used as the input to the PCA analysis to generate Figure~\ref{fig:figure6}B.
  \todo[noline]{R1.7}

  \subsection*{Synthetic interaction data}
  \vspace{-5mm}
  We generated synthetic interaction data using two methodologies previously used for benchmarking network inference methods.
  \todo[noline]{R1.4}
  \todo[noline]{R2.2}
  \todo[noline]{R2.4}

  The first method, ``seqtime''~\cite{faustSignaturesEcologicalProcesses2018}, used generalized Lotka-Volterra (gLV) equations to model the microbial community dynamics and utilized the Klemm–Eguı́luz algorithm to generate a clique-based interaction network~\cite{Rottjers2018}.
  We used the \href{https://github.com/hallucigenia-sparsa/seqtime}{seqtime} R package to simulate communities with number of species ($N$) varying from 10 to 150 (10, 25, 50, 100, 150 and 200).
  The initial species concentrations were randomly sampled from a Poisson distribution and the simulation was rerun to generate a number of samples ($S$) varying from 50 to 500 (50, 100, 200, 500) for different communities.
  The abundance values of the species in the community at the end of the simulation time were used to create the OTU table.

  The second method, ``NorTA'', used the Normal to Anything (NorTA) approach coupled with a given interaction network topology to generate the abundance distribution of the microbial community~\cite{Kurtz2015}.
  We used the \href{https://github.com/zdk123/SpiecEasi}{spieceasi} R package~\cite{Kurtz2015} to simulate communities with different network topologies (scale-free, cluster, block, Erdos-Renyi, band and hub) and target abundance distributions (Negative Binomial, Poisson, Zero-Inflated Negative Binomial).
  The OTU table was generated using the American Gut Project example in the spieceasi package (\texttt{amgut1.filt}) with the default parameter options.

  For each method, we generated the OTU table depicting the abundances of species and used this as input to generate association networks using \ac{micone} pipeline.
  The interaction matrix was used as the source of expected (true) interactions and the associations predicted using \ac{micone} were the source of predicted interactions.
  Finally, for each dataset we evaluated the precision and sensitivity of the associations predicted by the individual network inference methods as well as the consensus (Figures \ref{fig:figure5}, \ref{fig:figure_s5}, and \ref{fig:figure_s6}).

  \subsection*{Statistical analyses}

  \subsubsection*{DC step}
  \vspace{-5mm}
  In order to compare the representative sequences generated by the various methods in the DC step, we employed both the weighted~\cite{Lozupone2007} (Figure~\ref{fig:figure2}A) and unweighted UniFrac method~\cite{Lozupone2005} (Figure~\ref{fig:figure2}B).
  The UniFrac distance metric (unique fraction metric) is a beta-diversity measure that computes the distance between two sets of taxa as the fraction of the branch length of the tree that leads to descendants from either one environment or the other, but not both~\cite{Lozupone2005}.
  The weighted UniFrac distance metric takes into account the abundances of the representative sequences when calculating shared and unshared branch lengths, whereas the unweighted UniFrac distance metric does not and hence gives equal weights to each sequence.
  In Figure~\ref{fig:figure2} the distances between methods are the distance between the reference sequence distribution for a pair of methods averaged over every sample in the dataset.
  All UniFrac calculations were performed using the \texttt{scikit-bio}~\cite{thescikit-biodevelopmentteamScikitbioBioinformaticsLibrary2022} v0.5.6 Python package.

  \subsubsection*{TA step}
  \vspace{-5mm}
  In Figure~\ref{fig:figure3}C, we used the Bray-Curtis distance metric to calculate the distance between the predicted (using the taxonomy databases in the TA step) and expected taxonomic distribution.
  The Bray-Curtis distance is used to quantify the compositional dissimilarity between two different taxonomic distributions defined by vectors $u$ and $v$.
  It is defined as:
  $$d = \frac{\sum_i{|u_i-v_i|}}{\sum_i{|u_i+v_i|}}$$
  The Bray-Curtis distance calculations were performed using the \texttt{scipy}~\cite{virtanenSciPyFundamentalAlgorithms2020} v1.8.0 Python package.

  \subsubsection*{NI step}
  \vspace{-5mm}
  In Figure~\ref{fig:figure5} we evaluated the accuracy and robustness of the inferred association networks (using the various network inference algorithms and the consensus methods) against the original interaction network used to create the taxonomic distribution.
  We used the following formulations of precision and sensitivity to calculate the accuracy of the predictions:

  Precision = $\frac{TP}{TP + FP}$

  Sensitivity = $\frac{TP}{FN + TP}$

  where, TP - true positives, FP - false positives and FN - false negatives

  \subsection*{Code and Data Availability}
  Pipeline: \href{https://github.com/segrelab/MiCoNE}{https://github.com/segrelab/MiCoNE} \\
  Documentation: \href{https://micone.readthedocs.io}{https://micone.readthedocs.io} \\
  Data and scripts: \href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper} \\
  Synthetic data and scripts: \href{https://github.com/segrelab/MiCoNE-synthetic-data}{https://github.com/segrelab/MiCoNE-synthetic-data}

% ACKNOWLEDGMENTS
\section*{Acknowledgments}

We are grateful to members of the Segrè lab for helpful discussions and for feedback on the manuscript. This work was partially funded by grants from the National Institutes of Health (National Institute of General Medical Sciences, award R01GM121950; National Institute of Dental and Craniofacial Research, award number R01DE024468; National Institute on Aging, award number UH2AG064704; and National Cancer Institute, grant number R21CA260382), the U.S. Department of Energy, Office of Science, Office of Biological \& Environmental Research through the Microbial Community Analysis and Functional Evaluation in Soils SFA Program (m-CAFEs) under contract number DE-AC02-05CH11231 to Lawrence Berkeley National Laboratory, the National Science Foundation (grants 1457695, NSFOCE-BSF 1635070 and the NSF Center for Chemical Currencies of a Microbial Planet) and the Human Frontiers Science Program (RGP0020/2016 and RGP0060/2021).
DK acknowledges support by the Kilachand Multicellular Design Program graduate fellowship.
KSK was supported by Simons Foundation Grant \#409704, by the Research Corporation for Science Advancement through Cottrell Scholar Award \#24010, by the Scialog grant \#26119, and by the Gordon and Betty Moore Foundation grant \#6790.08.

\section*{Contributions}
Designed the research project: DK, KK, DS, ZH, CDL. Performed analysis: DK, GB. Wrote the first draft of the manuscript: DK. Revised and wrote the final version of the manuscript: DK, DS, KK.

%--------------------------------------------------------%
%   REFERENCE LIST
%--------------------------------------------------------%
\newpage
\singlespacing
\printbibliography

%--------------------------------------------------------%
%   APPENDIX
%--------------------------------------------------------%

\begin{acronym}[XXXXXXXX]
    \acro{ngs}[NGS]{next-generation sequencing}
    \acro{mind}[MIND]{Microbial Interaction Network Database}
    \acro{micone}[MiCoNE]{Microbial Co-occurrence Network Explorer}
    \acro{otu}[OTU]{Operational Taxanomic Unit}
    \acro{esv}[ESV]{Exact Sequence Variant}
    \acro{pcr}[PCR]{Polymerase Chain Reaction}
    \acro{emp}[EMP]{Earth Microbiome Project}
    \acro{gg}[GG]{Greengenes}
    \acro{ncbi}[NCBI]{National Center for Biotechnology Information}
    \acro{comets}[COMETS]{Computation of Microbial Ecosystems in Time and Space}
    \acro{dada2}[DADA2]{Divisive Amplicon Denoising Algorithm 2}
    \acro{qiime1}[QIIME1]{Quantitative Insights Into Microbial Ecology 1}
    \acro{qiime2}[QIIME2]{Quantitative Insights Into Microbial Ecology 2}
    \acro{sparcc}[SparCC]{Sparse Correlations for Compositional data}
    \acro{spieceasi}[SpiecEasi]{Sparse InversE Covariance estimation for Ecological Association and Statistical Inference}
    \acro{mldm}[mLDM]{metagenomic Lognormal-Dirichlet-Multinomial}
    \acro{cozine}[COZINE]{COmpositional Zero-Inflated Network Estimation}
    \acro{spring}[SPRING]{Semi-Parametric Rank-based approach for INference in Graphical model}
    \acro{harmonies}[HARMONIES]{Hybrid Approach for Microbiome Networks Inference via Exploiting Sparsity}
    \acro{kegg}[KEGG]{Kyoto Encyclopedia of Genes and Genomes}
    \acro{fba}[FBA]{Flux Balance Analysis}
    \acro{sp}[SP]{Sequence Processing}
    \acro{dc}[DC]{Denoising and Clustering}
    \acro{ta}[TA]{Taxonomy Assignment}
    \acro{op}[OP]{OTU Processing}
    \acro{ni}[NI]{Network Inference}
\end{acronym}

\newpage
\section*{Tables and Figures}

% Tables

  \begin{table}[H]
    \centering
    \caption{
      \textbf{Tools used in the \ac{micone} pipeline}.
      The tools highlighted in gray are the defaults for the pipeline that are recommended based on the benchmarks with the mock and synthetic datasets.
      The consensus algorithm in the Network Inference (\ac{ni}) step incorporates all the modules (bootstrapping, direct association, and correlation-based) to generate the consensus network.
    }
    \label{tab:micone_tools}
  \end{table}

% Figures

  % \FloatBarrier
  % \newpage
  % \begin{figure}[H]
  %   \centering
  %   \includegraphics[width=0.74\linewidth]{figure1.pdf}
  % \end{figure}
  \begin{figure}[H]
    \centering
    \caption{
      \textbf{The workflow of the \ac{micone} pipeline}.
      The steps of the workflow can be broken down into five major groups: \textbf{(SP)} \textbf{S}equence \textbf{P}rocessing, \textbf{(DC)} \textbf{D}enoising and \textbf{C}lustering, \textbf{(TA)} \textbf{T}axonomy \textbf{A}ssignment, \textbf{(OP)} \textbf{O}TU and ESV \textbf{P}rocessing, and \textbf{(NI)} \textbf{N}etwork \textbf{I}nference.
      Each step incorporates several processes (blue boxes), each of which in turn has several alternative algorithms for the same task (indicated by the text to the right of the blue boxes).
      Each arrow describes the data that is being passed from one step to another.
      The inputs to the pipeline are 16S rRNA sequencing reads, and the final output is the consensus network generated from the inferred co-occurrence networks.
      For details on each process and the different outputs, see Methods.
    }
    \label{fig:figure1}
  \end{figure}

  % \FloatBarrier
  % \newpage
  % \begin{figure}[H]
  %   \centering
  %   \includegraphics[width=\textwidth]{figure2.pdf}
  % \end{figure}
  \begin{figure}[H]
    \centering
    \caption{
      \textbf{The representative sequences generated by the different denoising and clustering methods differ in their identification of sequences that are low in abundance}.
      \textbf{(A)} The average weighted UniFrac distance between the representative sequences shows that the representative sequences and their compositions are fairly identical between the methods (with the exception of Deblur (DB) due to the low ESV count).
      \textbf{(B)} The relatively larger average unweighted UniFrac distance indicates that methods differ in their identification of sequences that are lower in abundance.
      The number of \ac{otu}s or \ac{esv}s generated by the respective methods are provided in the parenthesis next to their names.
      The data used for the analysis in (A, B) were the samples from the fecal microbiome transplant (FMT) dataset~\cite{Kang2017}, containing both healthy subjects and subjects with autism spectrum disorder (ASD).
      \textbf{(C, D)} The distributions of the average weighted and unweighted UniFrac distance between the predicted sequence profile and the expected sequence profile in the mock datasets.
      The average weighted UniFrac distances show that de novo (DN) and open reference (OR) were the best-performing methods in most of the datasets, while they are the worst-performing methods under the unweighted UniFrac metric.
      The good performance of dada2 (D2) under both distance metrics combined with its approach of identifying \ac{esv}s using de novo methods, prompts us to use it as the default method for the DC step.
      The data used for the analysis in (C, D) were the mock4, mock12, and mock16 datasets from mockrobiota~\cite{Bokulich2016}.
    }
    \label{fig:figure2}
  \end{figure}

  % \FloatBarrier
  % \newpage
  % \begin{figure}[H]
  %   \centering
  %   \includegraphics[width=\textwidth]{figure3.pdf}
  % \end{figure}
  \begin{figure}[H]
    \centering
    \caption{
      \textbf{Taxonomic reference databases vary in terms of their taxonomy assignments below the Order level}.
      \textbf{(A)} The taxonomic assignments of the top 50 representative sequences using the three different reference databases.
      This result illustrates how the same sequences are assigned to different genera under different databases.
      A significant portion of the representative sequences are assigned to an ``unknown'' Genus in two of three databases (\ac{gg} and \ac{ncbi}).
      The number of assigned genera for each database is displayed at the top of each column.
      \textbf{(B)} The number of representative sequences assigned to the same taxonomic label when using different reference databases (for the top 100 sequences).
      The mismatches are fewer at higher taxonomic levels, but, even at the Order level there exists greater than 51\% of mismatches, demonstrating the poor agreement in taxonomic labels assigned by the different databases.
      The data used for the analysis in (A, B) were samples (healthy and ASD) from the FMT dataset.
      \textbf{(C)} The Bray-Curtis dissimilarity between the predicted taxonomy profile and expected taxonomy profile in the mock datasets shows that there is no singular best choice of database for every dataset, as all the databases show similar performances.
      The \ac{gg} database and the Naive Bayes classifier are chosen as the defaults for the TA step of \ac{micone} due to their popularity.
      The datasets used for the analysis in (C) were the mock datasets from mockrobiota.
    }
    \label{fig:figure3}
  \end{figure}

  % \FloatBarrier
  % \newpage
  % \begin{figure}[H]
  %   \centering
  %   \includegraphics[width=\textwidth]{figure4.pdf}
  % \end{figure}
  \begin{figure}[H]
    \centering
    \caption{
      \textbf{Networks generated using different network inference methods show notable differences in terms of edge-density and connectivity}.
      \textbf{(A)} The nine different networks (excluding \acs{mldm}) generated by the different network inference methods.
      The nodes for each network (representing taxa) are arranged in the same positions in a circular layout and the differences in the connections can be directly visualized and compared.
      The green links are positive associations and the orange links represent negative associations.
      The networks look dissimilar and vary widely in terms of connectivity, and it is notable that the correlation-based methods generally produce networks with higher edge-densities.
      A threshold of 0.3 was set for the correlation-based methods (sparcc, propr, spearman and pearson) and a threshold of 0.01 was set for the direct association methods (flashweave, spieceasi, cozine, harmonies, and spring).
      \textbf{(B)} The node overlap Upset plot indicates that all the networks have a large proportion of common nodes involved in connections (33 out of 68).
      Conversely \textbf{(C)}, the edge overlap Upset plot shows that a very small fraction of these connections are actually shared (8 out of 202).
      The data used in this analysis were the healthy stool samples from the FMT dataset.
      \acs{mldm} is not shown in the comparisons because the algorithm failed to converge for the particular network combination used here (default setting of the \ac{micone} pipeline).
    }
    \label{fig:figure4}
  \end{figure}

  % \FloatBarrier
  % \newpage
  % \begin{figure}[H]
  %   \centering
  %   \includegraphics[width=1.0\linewidth]{figure5.pdf}
  % \end{figure}
  \begin{figure}[H]
    \centering
    \caption{
      \textbf{The associations generated by the scaled-sum consensus method show high precision in benchmarks using synthetic datasets}.
    The different points on the box plot show the precision of co-occurrence networks generated through individual network inference methods and through consensus network construction approaches. Precision is estimated based on the comparisons with two sets of synthetic benchmark datasets (``NorTA'' and ``seqtime'', see Methods).
      The independent algorithms chosen for the comparison are the two best-performing correlation-based (propr, sparcc) and direct association based (spieceasi, flashweave) methods.
      For consensus network inference, we used the scaled-sum (SS) and simple voting (SV) methods.
      A weight threshold of 0.1 and a p-value threshold of 0.05 was applied to each network before the calculation of precision.
      The overall best precision was consistently obtained by the scaled-sum consensus method for $p \geq 0.333$ on both benchmark datasets. Among all the individual network inference methods, spieceasi shows the best average precision.
      The simple voting method, when using the presence of edges in all inferred networks as a requirement ($p = 1.000$), also outperforms spieceasi on average precision.
      Therefore, we set the scaled-sum consensus method with $p = 0.333$ as the default tool for consensus network inference, since this option provides a good balance of precision and sensitivity (see also Figure~\ref{fig:figure_s5} and \ref{fig:figure_s6}).
    }
    \label{fig:figure5}
  \end{figure}

  % \FloatBarrier
  % \newpage
  % \begin{figure}[H]
  %   \centering
  %   \includegraphics[width=\textwidth]{figure6.pdf}
  % \end{figure}
  \begin{figure}[H]
    \centering
      \caption{
      \textbf{The choice of reference database has the largest impact on network variance}.
      \textbf{(A)} The percentage of variance in the networks (from the  FMT dataset) contributed by the \acf{dc}, CC (chimera checking), \acf{ta}, \acf{op} and \acf{ni} steps of the pipeline calculated using ANOVA on a linear model (see Methods).
      A weight threshold of 0.1 and a p-value threshold of 0.05 were applied to each network before the analysis.
      The taxonomy database contributes most to the variance between the networks (65.4\%) followed by the filtering of the counts matrix (26.8\%) in the OP step.
    The variation due to the NI, DC and CC steps are much smaller in comparison (6.553\%, 0.648\%, and 0.003\% respectively).
      The negligible fraction labeled as the residual is an artifact that arises when multiple steps are changed at the same time.
      \textbf{(B)} All the inferred networks generated from various combinations of tools are shown as points on a PCA plot.
      Each point on the PCA plot represents a network inferred using different combinations of tools and parameters that are available in the \ac{micone} pipeline.
      The color of the points corresponds to the tools used at each step of the pipeline (DC, TA, OP, and NI).
      The points on the PCA plot can be grouped based on the TA step, but the extent of this separation decreases when the filtering is turned on in the OP step, confirming that the variability in the networks decreased upon filtering out the taxonomic entities at low abundance.
      Some algorithms, especially the direct association methods, at the NI step can also be seen to generate networks that are less variable compared to the others.
      The \ac{dc} step does not seem to have any correlation with the variation in the networks on the PCA plot.
    }
    \label{fig:figure6}
  \end{figure}

  % \FloatBarrier
  % \newpage
  % \begin{figure}[H]
  %   \centering
  %   \includegraphics[width=1.0\linewidth]{figure7.pdf}
  % \end{figure}
  \begin{figure}[H]
    \centering
    \caption{
      \textbf{Comparison of networks generated from the control and ASD samples of the FMT dataset using the \ac{micone} pipeline}.
      The networks for the control (left) and ASD (right) samples were generated using the default tools and parameters recommended by the \ac{micone} pipeline as described in Table~\ref{tab:micone_tools}.
      There are 22 unique links in the network for control samples, 12 unique links in the network for ASD subjects, and 7 edges in common between both networks.
      The changes in these connections can serve as potential starting points for further experimental validations or literature surveys.
    }
    \label{fig:figure7}
  \end{figure}

  \FloatBarrier
  \newpage
  \subsection*{Supplementary Tables and Figures}

  \renewcommand{\thefigure}{S\arabic{figure}}
  \setcounter{figure}{0}

  \renewcommand{\thetable}{S\arabic{table}}
  \setcounter{table}{0}

  \begin{table}[H]
    \centering
    \caption{
      \textbf{Table of global network metrics for networks inferred from all possible combinations of tools}.
      In each row, one tool in a particular step is kept constant, and the metric is calculated for every possible combination of tools for the other steps of the pipeline.
      Therefore, each row shows the grouped average metric for each tool in every step of the pipeline.
      The network inference methods show the most variation in the global network metrics compared to tools in other steps of the pipeline.
    }
    \label{tab:network_metrics}
  \end{table}

    % \begin{figure}[H]
    %   \centering
    %   \includegraphics[width=1.0\linewidth]{figure_s1.pdf}
    % \end{figure}
    \begin{figure}[H]
      \centering
        \caption{
          \textbf{The t-SNE plot of all the inferred networks clusters the networks based on the taxonomy reference database used}.
          Each point on the t-SNE plot represents a network inferred using different combinations of tools and parameters that are available in the \ac{micone} pipeline.
          The points are colored by the tools and parameters used in \ac{dc} step (A), \ac{ta} step (B), \ac{op} step (C) and \ac{ni} step (D).
          The separation of the points based on taxonomy reference database shows that the points cluster based on reference database in high-dimensional space.
        }
      \label{fig:figure_s1}
    \end{figure}
    % \FloatBarrier
    % \newpage

    % \begin{figure}[H]
    %   \centering
    %   \includegraphics[width=1.0\linewidth]{figure_s2.pdf}
    % \end{figure}
    \begin{figure}[H]
      \centering
        \caption{
          \textbf{The UniFrac distance between the 1000 most abundant representative sequences is higher than that when all sequences are considered}.
          Each value is the average UniFrac distance between the reference sequences generated by the various methods in the \ac{dc} step (similar to Figure~\ref{fig:figure2}).
          There is an increase in both weighted and unweighted UniFrac distances compared to when all the representative sequences are considered.
          This shows that the 1000 most abundant representative sequences generated by the DC methods are not as similar to each other.
          And since the weighted UniFrac is much smaller than the unweighted UniFrac distance, we can conclude that those reference sequences that are present in the middle of the abundance distribution (considering all sequences) are dissimilar.
        }
      \label{fig:figure_s2}
    \end{figure}
    % \FloatBarrier
    % \newpage

    % \begin{figure}[H]
    %   \centering
    %   \includegraphics[width=1.0\linewidth]{figure_s3.pdf}
    % \end{figure}
    \begin{figure}[H]
      \centering
        \caption{
          \textbf{The weighted and unweighted UniFrac distances are small for the representative sequences generated using remove bimera and uchime for each denoising method}.
          With the exception of de novo and open reference under the unweighted UniFrac metric, all the other methods have high similarity, implying that the two chimera checking methods, uchime and remove bimera, produce similar outputs.
          This is especially true for the \ac{dada2} and Deblur methods which are the recommended denoising methods in the \ac{micone} pipeline.
          Therefore, remove bimera is recommended as the default chimera method if one is using \ac{dada2} and uchime-denovo when one is using Deblur, since these methods were developed for these respective algorithms (\ac{qiime2} uses uchime-denovo in the Deblur workflow).
        }
      \label{fig:figure_s3}
    \end{figure}
    % \FloatBarrier
    % \newpage

    % \begin{figure}[H]
    %   \centering
    %   \includegraphics[width=1.0\linewidth]{figure_s4.pdf}
    % \end{figure}
    \begin{figure}[H]
      \centering
        \caption{
          \textbf{The pairwise comparison of assignments generated using different databases for all representative sequences has a higher proportion of mismatches}.
          The comparison made here is similar to Figure~\ref{fig:figure3}B, but instead of the top 100 taxonomic entities (by abundance), all the assignments from one database are matched with those from the other two databases.
          The higher percentage of mismatches implies that the assigned taxonomies in the more abundant sequences (top 100) match more consistently.
        }
      \label{fig:figure_s4}
    \end{figure}
    % \FloatBarrier
    % \newpage

  % \begin{figure}[H]
  %   \centering
  %   \includegraphics[width=1.0\linewidth]{figure_s5.pdf}
  % \end{figure}
  \begin{figure}[H]
    \centering
      \caption{
        \textbf{The precision and sensitivity of the inferred networks on the ``NorTA'' synthetic interaction data}.
        The different consensus methods used are scaled-sum (SS) and simple voting (SV) methods.
        Pearson and Spearman methods are not used in the calculation of the consensus.
        Among all the independent network inference methods, \acs{spieceasi} has the best average precision (0.944), but the overall best precision was consistently obtained by the scaled-sum method (0.956, 0.985, and 1.000).
        The simple voting method when using the presence of edges in all inferred networks as a requirement ($p = 1.000$), also outperforms \acs{spieceasi} on average precision (0.969).
        Although \acs{spieceasi} has a higher sensitivity, if the goal of network inference is to obtain the list of associations that have a high probability of existing in the real microbial community, then the consensus methods perform better.
      }
    \label{fig:figure_s5}
  \end{figure}
  % \FloatBarrier
  % \newpage

  % \begin{figure}[H]
  %   \centering
  %   \includegraphics[width=1.0\linewidth]{figure_s6.pdf}
  % \end{figure}
  \begin{figure}[H]
    \centering
      \caption{
        \textbf{The precision and sensitivity of the inferred networks on the ``seqtime'' synthetic interaction data.}
        The different consensus methods used are scaled-sum (SS) and simple voting (SV) methods.
        Pearson and Spearman methods are not used in the calculation of the consensus.
        Among all the independent network inference methods, \acs{spieceasi} has the best average precision (0.624), but the overall best precision was consistently obtained by the scaled-sum method (0.688, 0.820, and 1.000).
        The simple voting method when using the presence of edges in all inferred networks as a requirement ($p = 1.000$), also outperforms \acs{spieceasi} on average precision (0.692).
        These results show that the scaled-sum method is not only much better suited for inferring robust and accurate interactions from count data generated from network topologies (NorTA), but it is also capable of accurately extracting real associations from Lotka-Volterra simulations.
      }
    \label{fig:figure_s6}
  \end{figure}
  % \FloatBarrier
  % \newpage

  % \begin{figure}[H]
  %   \centering
  %   \includegraphics[width=1.0\linewidth]{figure_s7.pdf}
  % \end{figure}
  \begin{figure}[H]
    \centering
    \caption{
      \textbf{Sensitivity analysis of the default settings of the \ac{micone} pipeline}.
      \textbf{(A)} The network constructed using the default pipeline parameters (DC=\ac{dada2}, CC=remove bimera, TA=\ac{gg}, OP=Filter(on), NI=scaled-sum consensus) is compared against networks generated when one of the steps uses a different tool.
      The layout is created by fixing the positions of all the nodes from all networks and then drawing only the relevant edges, making the connections directly comparable.
      The edges colored green are positive associations and those in red are negative associations.
      We observe that changing the TA and OP steps leads to the creation of the most number of unique edges.
      \textbf{(B)} The dot plot showing the fraction of nodes (left) and edges (right) in common between the default network and the networks generated by changing one step of the default pipeline.
      The low value of the common fraction for TA and OP steps shows that these steps induce the biggest changes in nodes and edges.
      The NI step is not shown in this analysis because the consensus methods use edges from the individual network inference methods and a comparison would be biased.
    }
    \label{fig:figure_s7}
  \end{figure}
  % \FloatBarrier
  % \newpage

  \renewcommand{\figurename}{Text}
  \renewcommand{\thefigure}{S\arabic{figure}}
  \begin{figure}[H]
    \centering
    \caption{Supplemental text describing data processing, synthetic interaction data generation, network metrics, p-value merging and supplementary discussion sections.}
    \label{txt:text_s1}
  \end{figure}

% \input{sections/supplementary.tex}

%--------------------------------------------------------%
%   END DOCUMENT
%--------------------------------------------------------%

\end{document}

### 2022-10-14 sections/tables_figures.tex
Update on Overleaf.
Add: Combined manuscript file and Text S1 caption

@@
       The edges colored green are positive associations and those in red are negative associations.
-      We observe that changing the TA and OP steps leads to the creation of the most unique edges.
+      We observe that changing the TA and OP steps leads to the creation of the most number of unique edges.
       \textbf{(B)} The dot plot showing the fraction of nodes (left) and edges (right) in common between the default network and the networks generated by changing one step of the default pipeline.
@@
   % \newpage
+
+  \renewcommand{\figurename}{Text}
+  \renewcommand{\thefigure}{S\arabic{figure}}
+  \begin{figure}[H]
+    \centering
+    \caption{Supplemental text describing data processing, synthetic interaction data generation, network metrics, p-value merging and supplementary discussion sections.}
+    \label{txt:text_s1}
+  \end{figure}

### 2022-10-14 manuscript.tex
Change: Comment out annotations

@@
   Our analysis involved two types of data: The first type consisted of 16S sequencing data from samples of human stool microbiomes from a fecal microbiome transplant (FMT) study of autism~\cite{Kang2017}.
-  \todo[noline]{R2.4}
+  %\todo[noline]{R2.4}
   The second type was a collection of datasets synthetically or artificially created for the specific goal of evaluating computational analysis tools.
@@
   This is because the open-reference and de novo clustering methods return a much larger number of \ac{otu}s compared to the other pipelines and would affect the accuracy of the network inference step if stringent filtering is not performed.
-  \todo[noline]{R2.6}
+  %\todo[noline]{R2.6}
   Overall, since \ac{dada2} as compared to Deblur, displays better performance on all the mock datasets on the weighted UniFrac metric, we set this tool as the default for the DC step of the pipeline.
-  \todo[noline]{R2.10}
+  %\todo[noline]{R2.10}
   However, if comparison across studies that have sequenced different 16S regions is required, closed-reference and open-reference might be a better option.
@@
   Since no database performs better than others against mock datasets, the choice of which database to use could be driven by other reasons (see Supplementary discussion).
-  \todo[noline]{R2.2a}
+  %\todo[noline]{R2.2a}
   One reason to choose a particular database could be the frequency of updates and the potential for future growth.
@@
   The TA step results in a taxonomic counts table that is used as input to the subsequent steps of the pipeline.
-  \todo[noline]{R2.1b}
+  %\todo[noline]{R2.1b}
   Note that the count tables at different levels can be obtained through aggregation; for example, Genus count tables were obtained by summing up the counts of the lower taxonomy levels (Species and \ac{otu}) that map to the same higher taxonomy level entity.
@@
   Filtered (see \ac{op} step in Methods) genus-level counts table obtained using the default settings in the previous steps were used as input for the different network inference algorithms (Figure~\ref{fig:figure4}).
-  \todo[noline]{R2.1b}
+  %\todo[noline]{R2.1b}
   Even from a visual inspection (Figure~\ref{fig:figure4}A), one can see that the different networks differ vastly in their edge-density and connectivity, with common edges often displaying inverted signs.
@@
  Inspired by previous approaches~\cite{bustinceFuzzySetsTheir2008,tsarevApplicationMajorityVoting2018}, we developed two methods that take into consideration the evidence offered by each network inference algorithm and generate a consensus network that contains the common edges among the inferred networks.
-  \todo[noline]{R1.1}
-  \todo[noline]{R1.4}
+  %\todo[noline]{R1.1}
+  %\todo[noline]{R1.4}
 
@@
   The analysis of the rewiring of associations in the ASD samples with respect to the control provides a guide for the identification of key genera that could be linked to dysbiosis.
-  \todo[noline]{R2.14}
+  %\todo[noline]{R2.14}
   We observed 22 unique links in the network for control samples, 12 unique links in the network for ASD subjects, and 7 edges in common between the two networks.
@@
   Through \ac{micone}, in addition to transparently revealing the dependence of co-occurrence networks on tool and parameter choices (see Discussion in Supplementary Text for details on the DC, TA and OP steps), we have taken advantage of our spectrum of computational options and the availability of mock and synthetic datasets, to suggest a default standard setting.
-  \todo[noline]{R1.3}
+  %\todo[noline]{R1.3}
   Additionally, we have developed a consensus approach, that can reliably generate networks that are fairly robust across multiple tool choices.
@@
   The network analyses in this study are primarily at the Genus level, wherein the lowest resolution of a node is a Genus and if an entity cannot be resolved to the Genus level, the next lowest taxonomic level is used (for example, Family).
-  \todo[noline]{R2.1c}
+  %\todo[noline]{R2.1c}
   As a consequence, two entities belonging to the same lineage where one entity is resolved to the Genus level and another is resolved to the Family level are treated as two different nodes in the network.
@@
   Although direct comparisons between co-occurrence networks and directly measured interactions are difficult to interpret and highly debated~\cite{hiranoDifficultyInferringMicrobial2019,gobernaCautionaryNotesUse2022}.
-  \todo[noline]{R2.2}
+  %\todo[noline]{R2.2}
   Further, benchmarking of co-occurrence networks could also be pursued through the use of literature-based interactions~\cite{lima-mendezDeterminantsCommunityStructure2015a} or biological benchmark interaction data~\cite{sungGlobalMetabolicInteraction2017a}.
@@
   Although in the current analysis, we have only used default parameter values recommended by the tool creators, the \ac{micone} pipeline could be used in the future to explore any combinations of parameters and to optimize these values for improved network inference.
-  \todo[noline]{R1.3}
+  %\todo[noline]{R1.3}
   Overall, there likely is no ``best method'' for the various steps of 16S data analysis, and hence, \ac{micone} is intended to help researchers to identify the methods and algorithms that are most suitable for their datasets in an easy-to-use and reproducible manner.
@@
   This study utilized two types of 16S rRNA sequencing datasets: biological datasets and mock/synthetic datasets.
-  \todo[noline]{R2.4}
+  %\todo[noline]{R2.4}
   Biological datasets are collections of sequencing reads obtained from naturally occurring microbial community samples.
@@
   This modular architecture simplifies the process of adding new tools (refer to the modules section in the \ac{micone} documentation).
-  \todo[noline]{R2.11}
+  %\todo[noline]{R2.11}
   The main components of the pipeline are detailed in the subsequent sections.
@@
   The module also supports rarefaction, which is a normalization technique used to overcome the bias that might arise due to variable sampling depth in different samples.
-  \todo[noline]{R2.5}
+  %\todo[noline]{R2.5}
   This is performed either by sub-sampling or by normalization of the matrix to the lowest sampling depth~\cite{Weiss2015}.
@@
   The group operation transforms the \ac{otu} or \ac{esv} count matrix into a taxonomic count matrix at the requested level by adding up counts that map to the same taxonomy and is carried out at the end of the OP step.
-  \todo[noline]{R2.1b}
+  %\todo[noline]{R2.1b}
   Finally, transformations are performed in order to correct for and overcome the compositional bias that is inherent in the counts matrix (in the analysis performed in the study these were disabled and directly handled by the network inference algorithm).
@@
   The consensus methods combine networks inferred from both correlation-based and direct association methods.
-  \todo[noline]{R1.1}
-  \todo[noline]{R1.4}
+  %\todo[noline]{R1.1}
+  %\todo[noline]{R1.4}
   First, for the correlation-based methods, we calculate p-values using null models and then merge the p-values using Brown's p-value merging method~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016,faustCoNetAppInference2016}.
-  \todo[noline]{R2.2}
-  \todo[noline]{R2.3}
+  %\todo[noline]{R2.2}
+  %\todo[noline]{R2.3}
   Second, we filter all the inferred networks based on an association strength threshold of 0.1 and a p-value cutoff of 0.05.
@@
   The next step in the consensus algorithm workflow is to merge the p-values for the networks generated by the correlation-based methods.
-  \todo[noline]{R1.1}
+  %\todo[noline]{R1.1}
   This step is performed using the Brown's p-value merging method~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016,faustCoNetAppInference2016}.
@@
   This section defines the notation used for the network variability analysis performed for Figure~\ref{fig:figure6}.
-  \todo[noline]{R1.2}
-  \todo[noline]{R1.8}
-  \todo[noline]{R2.1b,d}
+  %\todo[noline]{R1.2}
+  %\todo[noline]{R1.8}
+  %\todo[noline]{R2.1b,d}
 
@@
    Note that row $j$ of $\mathbf{\bar{N}}$, $\bar{N}_j$ is the vector that encodes the values of edge $j$ across all the networks.
-  \todo[noline]{R1.5}
-  \todo[noline]{R2.1b}
+  %\todo[noline]{R1.5}
+  %\todo[noline]{R2.1b}
 
@@
    The variance contributed by each step of the pipeline was calculated for every component in $\mathbf{C}$ matrix through ANOVA using the Python statsmodels~\cite{seaboldStatsmodelsEconometricStatistical2010} package and is shown in Figure~\ref{fig:figure6}A.
-  \todo[noline]{R1.6}
+  %\todo[noline]{R1.6}
   The total variance for the network was calculated by adding the variances for each connection and normalizing with the degrees of freedom.
   The merged network table $\mathbf{\bar{N}}$ was used as the input to the PCA analysis to generate Figure~\ref{fig:figure6}B.
-  \todo[noline]{R1.7}
+  %\todo[noline]{R1.7}
 
@@
   We generated synthetic interaction data using two methodologies previously used for benchmarking network inference methods.
-  \todo[noline]{R1.4}
-  \todo[noline]{R2.2}
-  \todo[noline]{R2.4}
+  %\todo[noline]{R1.4}
+  %\todo[noline]{R2.2}
+  %\todo[noline]{R2.4}
 

### 2022-10-14 sections/methods.tex
Change: Comment out annotations

--- 2022-09-30 sections/methods.tex
@@
   This study utilized two types of 16S rRNA sequencing datasets: biological datasets and mock/synthetic datasets.
+  %\todo[noline]{R2.4}
   Biological datasets are collections of sequencing reads obtained from naturally occurring microbial community samples.
@@
   This modular architecture simplifies the process of adding new tools (refer to the modules section in the \ac{micone} documentation).
+  %\todo[noline]{R2.11}
   The main components of the pipeline are detailed in the subsequent sections.
@@
   The module also supports rarefaction, which is a normalization technique used to overcome the bias that might arise due to variable sampling depth in different samples.
+  %\todo[noline]{R2.5}
   This is performed either by sub-sampling or by normalization of the matrix to the lowest sampling depth~\cite{Weiss2015}.
@@
   The group operation transforms the \ac{otu} or \ac{esv} count matrix into a taxonomic count matrix at the requested level by adding up counts that map to the same taxonomy and is carried out at the end of the OP step.
+  %\todo[noline]{R2.1b}
   Finally, transformations are performed in order to correct for and overcome the compositional bias that is inherent in the counts matrix (in the analysis performed in the study these were disabled and directly handled by the network inference algorithm).
@@
   The consensus methods combine networks inferred from both correlation-based and direct association methods.
-  First, for the correlation-based methods, we calculate p-values using null models and then merge the p-values using Brown's p-value merging method~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016,faustCoNetAppInference2016} (see Methods).
+  %\todo[noline]{R1.1}
+  %\todo[noline]{R1.4}
+  First, for the correlation-based methods, we calculate p-values using null models and then merge the p-values using Brown's p-value merging method~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016,faustCoNetAppInference2016}.
+  %\todo[noline]{R2.2}
+  %\todo[noline]{R2.3}
   Second, we filter all the inferred networks based on an association strength threshold of 0.1 and a p-value cutoff of 0.05.
@@
   The next step in the consensus algorithm workflow is to merge the p-values for the networks generated by the correlation-based methods.
+  %\todo[noline]{R1.1}
   This step is performed using the Brown's p-value merging method~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016,faustCoNetAppInference2016}.
@@
   This section defines the notation used for the network variability analysis performed for Figure~\ref{fig:figure6}.
+  %\todo[noline]{R1.2}
+  %\todo[noline]{R1.8}
+  %\todo[noline]{R2.1b,d}
 
@@
    Note that row $j$ of $\mathbf{\bar{N}}$, $\bar{N}_j$ is the vector that encodes the values of edge $j$ across all the networks.
+  %\todo[noline]{R1.5}
+  %\todo[noline]{R2.1b}
 
@@
    The variance contributed by each step of the pipeline was calculated for every component in $\mathbf{C}$ matrix through ANOVA using the Python statsmodels~\cite{seaboldStatsmodelsEconometricStatistical2010} package and is shown in Figure~\ref{fig:figure6}A.
+  %\todo[noline]{R1.6}
   The total variance for the network was calculated by adding the variances for each connection and normalizing with the degrees of freedom.
   The merged network table $\mathbf{\bar{N}}$ was used as the input to the PCA analysis to generate Figure~\ref{fig:figure6}B.
+  %\todo[noline]{R1.7}
 
@@
   We generated synthetic interaction data using two methodologies previously used for benchmarking network inference methods.
+  %\todo[noline]{R1.4}
+  %\todo[noline]{R2.2}
+  %\todo[noline]{R2.4}
 

### 2022-10-14 sections/supplementary.tex
Change: Comment out annotations

@@
   In Table~\ref{tab:network_metrics} we show various global network metrics calculated for each tool in the pipeline.
-  \todo[noline]{R2.1a}
-  \todo[noline]{R2.7}
+  %\todo[noline]{R2.1a}
+  %\todo[noline]{R2.7}
   All the networks that make use of a particular tool are grouped together, and the following average metrics are calculated for each group:
@@
   Fisher~\cite{fisher_224a_1948} proposed that for $k$ independent p-values, each generated by $k$ different methods and denoted by $\bar{P}^i$ (notations are same as used in the "Consensus network and p-value merging" subsection of the Methods).
-  \todo[noline]{R1.1}
+  %\todo[noline]{R1.1}
   The following will hold true for the statistic $\Psi$:
@@
   The choice of taxonomy database was found to be the most important factor in the inference of microbial co-occurrence networks, contributing $65.4\%$ of the total variance.
-  \todo[noline]{R2.2a}
+  %\todo[noline]{R2.2a}
   The frequent changes in the taxonomy nomenclature coupled with the frequency of updates to the various 16S reference databases create inherent differences \cite{Balvociute2017} in taxonomy hierarchies in these databases.
   Our analysis revealed that no particular reference database performs better than the others across the different mock dataset benchmarks.
-  \todo[noline]{R2.12}
+  %\todo[noline]{R2.12}
   The default reference database in the pipeline is the \ac{gg} reference database along with the ``Naive Bayes'' classifier as the query tool.

### 2022-12-01 final_submission/manuscript.tex
Add: Draft for final submission
Change: Bootstrap -> Permutation in the manuscript

--- 2022-10-14 manuscript.tex
@@
   \usepackage[style=nature,url=false,backend=biber,sorting=none]{biblatex}
-  \bibliography{references/references.bib}
+  \bibliography{references.bib}
 
@@
   On the other hand, direct association methods use conditional independence to avoid the detection of correlated but indirectly connected taxonomic entities~\cite{Kurtz2015,Menon2018}.
-  A null model is created by re-sampling and bootstrapping the counts matrix and recalculating the correlations (see next section for details on network analysis and statistics).
-  These bootstrapped association matrices are used to calculate the significance of the inferred correlations by calculating the p-values against this null model~\cite{Watts2018}.
+  A null model is created by re-sampling and permuting the counts matrix and recalculating the correlations (see next section for details on network analysis and statistics).
+  These permuted association matrices are used to calculate the significance of the inferred correlations by calculating the p-values against this null model~\cite{Watts2018}.
   Brown's p-value merging method~\cite{brown_400_1975} is used for combining p-values from the pairwise correlations methods to obtain a consensus p-value, which can be used to filter for significance.
-  The bootstrapping and p-value calculations are only performed on the correlations-based methods.
+  The permutations and p-value calculations are only performed on the correlations-based methods.
   In the final module of this step, the consensus algorithms are used to create the final consensus network using associations from all the network inference methods (except Pearson and Spearman, by default).
@@
 
-  \subsubsection*{Bootstrapping and p-value calculation}
+  \subsubsection*{Permutations and p-value calculation}
   \vspace{-5mm}
@@
   Finally, the p-value is determined based on how often a more extreme association is observed for randomly permuted data.
-  Note that, all the direct association-based methods used in the study have their own regularization methods built in and hence do not need to undergo bootstrapping.
+  Note that, all the direct association-based methods used in the study have their own regularization methods built in and hence do not need to undergo this procedure.
 
@@
       The tools highlighted in gray are the defaults for the pipeline that are recommended based on the benchmarks with the mock and synthetic datasets.
-      The consensus algorithm in the Network Inference (\ac{ni}) step incorporates all the modules (bootstrapping, direct association, and correlation-based) to generate the consensus network.
+      The consensus algorithm in the Network Inference (\ac{ni}) step incorporates all the modules (permutations, direct association, and correlation-based) to generate the consensus network.
     }
@@
 
+  \setcounter{figure}{0}
   \renewcommand{\figurename}{Text}

### 2022-12-01 final_submission/manuscript.tex
Change: Address R2 comments + change p -> \theta

@@
   We note that there is no method that outperforms the rest in all datasets (see Supplementary for an extended discussion).
-  Based on their slightly better performance on the mock datasets, their de novo error correcting nature and previous independent evaluation~\cite{Nearing2018}, \ac{dada2} and Deblur appear to be the most reliable.
+  Based on being among the top performers on the mock datasets, their de novo error correcting nature and previous independent evaluation~\cite{Nearing2018}, \ac{dada2} and Deblur appear to be the most reliable.
   This is because the open-reference and de novo clustering methods return a much larger number of \ac{otu}s compared to the other pipelines and would affect the accuracy of the network inference step if stringent filtering is not performed.
@@
   We chose the scaled-sum method as the pipeline default since this method takes into account the weights of the associations in the determination of the final consensus.
-  The pipeline enables the selection of any subset of methods for the consensus calculation. Currently, by default, all direct methods are used, together with \acs{sparcc} and propr.
+  The pipeline enables the selection of any subset of methods for the consensus calculation.
+Currently, by default, all direct association methods (\acs{spieceasi}, \acs{cozine}, \acs{harmonies}, \acs{spring}, \acs{mldm}, and FlashWeave) are used, together with \acs{sparcc} and propr.
 
@@
   As shown in Figure \ref{fig:figure5} the consensus algorithm, especially the scaled-sum method, captures true associations with high precision (through the removal of edges that are either not present in most of the inference methods or whose association strength is low across methods).
-  Overall, the scaled-sum method for $p=1.000$ performs the best (precision = $1.000$ for both NorTA and seqtime).
-  The scaled-sum method for $p=0.333$ (default option in the pipeline) shows a high precision ($0.956$ with NorTA; $0.688$ with seqtime), without displaying significant reduction in sensitivity (Figure~\ref{fig:figure_s5} and Figure~\ref{fig:figure_s6}).
-  However, if higher precision is required $p>0.5$ can be considered.
+  Overall, the scaled-sum method for $\theta=1.000$ performs the best (precision = $1.000$ for both NorTA and seqtime).
+  The scaled-sum method for $\theta=0.333$ (default option in the pipeline) shows a high precision ($0.956$ with NorTA; $0.688$ with seqtime), without displaying significant reduction in sensitivity (Figure~\ref{fig:figure_s5} and Figure~\ref{fig:figure_s6}).
+  However, if higher precision is required $\theta>0.5$ can be considered.
   Therefore, the consensus networks provide the means to obtain a short list of associations that would have a high likelihood of being present in the real association network.
@@
   For this purpose, we developed two simple algorithms that combine the edges reported by the different network inference tools.
-  Both the algorithms make use of a user-defined parameter $p$ ($0 \leq p \leq 1$), in order to threshold the edges from the individual methods.
-  The inputs to both the algorithms are the co-occurrence networks (association matrices) $\bar{N}^i$ (flattened version of $N^{i}$) generated by each method $i$, and the threshold parameter $p$.
+  Both the algorithms make use of a user-defined parameter $\theta$ ($0 \leq \theta \leq 1$), in order to threshold the edges from the individual methods.
+  The inputs to both the algorithms are the co-occurrence networks (association matrices) $\bar{N}^i$ (flattened version of $N^{i}$) generated by each method $i$, and the threshold parameter $\theta$.
   Here, the $\bar{N}^{i}$ each have the same set of nodes $q$ and only differ by the value of the association inferred between every pair.
@@
   \begin{equation}
-     M_j \geq \lfloor p \times w \rfloor
+     M_j \geq \lfloor \theta \times w \rfloor
     \label{eqn:simple-voting2}
   \end{equation}
-  where, $p$ is the user-defined threshold parameter.
-
-  The simple voting method returns the union of the networks when $0 \leq p \leq \frac{1}{w}$ and will return the intersection when $\frac{(w - 1)}{w} \leq p \leq 1$.
-  In general, if $\frac{(n - 1)}{w} \leq p \leq \frac{n}{w}$, this algorithm will report an edge in the consensus network when at least $n$ network inference methods report this edge.
+  where, $\theta$ is the user-defined threshold parameter.
+
+  The simple voting method returns the union of the networks when $0 \leq \theta \leq \frac{1}{w}$ and will return the intersection when $\frac{(w - 1)}{w} \leq \theta \leq 1$.
+  In general, if $\frac{(n - 1)}{w} \leq \theta \leq \frac{n}{w}$, this algorithm will report an edge in the consensus network when at least $n$ network inference methods report this edge.
 
@@
   \begin{equation}
-    \mid s_j \mid > (w - 1) \times p
+    \mid s_j \mid > (w - 1) \times \theta
     \label{eq:scaled-sum2}
@@
       A weight threshold of 0.1 and a p-value threshold of 0.05 was applied to each network before the calculation of precision.
-      The overall best precision was consistently obtained by the scaled-sum consensus method for $p \geq 0.333$ on both benchmark datasets.
+      The overall best precision was consistently obtained by the scaled-sum consensus method for $\theta \geq 0.333$ on both benchmark datasets.
 Among all the individual network inference methods, spieceasi shows the best average precision.
-      The simple voting method, when using the presence of edges in all inferred networks as a requirement ($p = 1.000$), also outperforms spieceasi on average precision.
-      Therefore, we set the scaled-sum consensus method with $p = 0.333$ as the default tool for consensus network inference, since this option provides a good balance of precision and sensitivity (see also Figure~\ref{fig:figure_s5} and \ref{fig:figure_s6}).
+      The simple voting method, when using the presence of edges in all inferred networks as a requirement ($\theta = 1.000$), also outperforms spieceasi on average precision.
+      Therefore, we set the scaled-sum consensus method with $\theta = 0.333$ as the default tool for consensus network inference, since this option provides a good balance of precision and sensitivity (see also Figure~\ref{fig:figure_s5} and \ref{fig:figure_s6}).
     }
@@
         Among all the independent network inference methods, \acs{spieceasi} has the best average precision (0.944), but the overall best precision was consistently obtained by the scaled-sum method (0.956, 0.985, and 1.000).
-        The simple voting method when using the presence of edges in all inferred networks as a requirement ($p = 1.000$), also outperforms \acs{spieceasi} on average precision (0.969).
+        The simple voting method when using the presence of edges in all inferred networks as a requirement ($\theta = 1.000$), also outperforms \acs{spieceasi} on average precision (0.969).
         Although \acs{spieceasi} has a higher sensitivity, if the goal of network inference is to obtain the list of associations that have a high probability of existing in the real microbial community, then the consensus methods perform better.
@@
         Among all the independent network inference methods, \acs{spieceasi} has the best average precision (0.624), but the overall best precision was consistently obtained by the scaled-sum method (0.688, 0.820, and 1.000).
-        The simple voting method when using the presence of edges in all inferred networks as a requirement ($p = 1.000$), also outperforms \acs{spieceasi} on average precision (0.692).
+        The simple voting method when using the presence of edges in all inferred networks as a requirement ($\theta = 1.000$), also outperforms \acs{spieceasi} on average precision (0.692).
         These results show that the scaled-sum method is not only much better suited for inferring robust and accurate interactions from count data generated from network topologies (NorTA), but it is also capable of accurately extracting real associations from Lotka-Volterra simulations.
@@
 
-% \input{sections/supplementary.tex}
+\newpage
+\section*{Supplementary Text}
+
+  \subsection*{Processing the FMT data}
+
+    \subsubsection*{Data download and pre-processing}
+    The main biological dataset used in this study was the collection of 16S rRNA sequencing reads from stool samples (healthy and autistic individuals) for a fecal microbiome transplant study~\cite{Kang2017}.
+    The data containing the 16S sequencing reads (V4 region) was downloaded from Qiita~\cite{gonzalezQiitaRapidWebenabled2018} (study ID: 10532).
+    Only runs 2, 3, and 4 were used for the subsequent analysis as these runs consisted of paired-end sequencing data, and run 1 contained single-end data.
+    The sample metadata was updated to contain only BMI, sex, height, weight, and experimental group.
+    This was necessary as two of the network inference algorithms (\acs{mldm} and FlashWeave) required information about environmental heterogeneity.
+    However, these environmental correlations were not included in the current analyses.
+
+    \subsubsection*{Processing using the \ac{micone} pipeline}
+    The data was then processed using the \ac{micone} pipeline starting at the \ac{sp} step and ending at the \ac{ni} step with the consensus algorithm.
+    The configuration files (main.nf and nextflow.config) used to run the \ac{micone} pipeline as well the details of the pipeline execution (dag, report, timeline and trace) are in the "runs/FMT" directory of the data and scripts repository (\href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper})
+    The results of the pipeline execution for reproducing the analyses in the manuscript are stored on Zenodo.
+
+  \subsection*{Processing the mock data}
+
+    \subsubsection*{Data download and pre-processing}
+    The mock datasets, mock4, mock12 and mock16 used for this study, were obtained from mockrobiota~\cite{Bokulich2016}.
+    Mock 4 is a mock community composed of 21 bacterial strains represented in equal abundances in two replicate samples, and the same strains represented in uneven abundances in two other replicate samples.
+    Mock 12 is composed of 27 bacterial strains containing closely related taxa, the members of which were chosen in part for their well-separated 16S rRNA gene sequences.
+Some pairs of strains differ by as little as one nucleotide, but all the strains are distinguishable over the sequenced region of the 16S rRNA gene.
+    Mock 16 is a mock community composed of even amounts of purified genomic DNA from 49 bacteria and 10 archaea.
+    The datasets did not require any preprocessing and could be directly used as input to the pipeline
+
+    \subsubsection*{Processing using the \ac{micone} pipeline}
+    The data was processed using the \ac{micone} pipeline starting at the \ac{sp} step and ending at the \ac{op} step with the filtered taxonomic tables as the final output.
+    The configuration files (main.nf and nextflow.config) used to run the \ac{micone} pipeline as well the details of the pipeline execution (dag, report, timeline and trace) are in the "runs/mock*" directory of the data and scripts repository (\href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper})
+    The results of the pipeline execution for reproducing the analyses in the manuscript are stored on Zenodo.
+
+    \subsubsection*{Interpretation of Unifrac results in the DC step}
+  In Figure \ref{fig:figure_s3}, we observe that both the weighted and unweighted UniFrac distances are increased for the top 1000 representative sequences, implying that the top representative sequences generated by the different methods are not as similar to each other.
+  Therefore, since the weighted UniFrac distances are lower than the unweighted distances, we conclude that the representative sequences in the middle range of the abundance distribution are those that must be the most similar between the methods.
+
+    Open-reference and de novo clustering methods perform the best under the weighted UniFrac metric and the worst (marginally) under the unweighted UniFrac metric (Figure~\ref{fig:figure2}C and \ref{fig:figure2}D).
+    This result can be attributed to the large number of low abundance representative sequences that are generated by these methods.
+    Deblur performs poorly under weighted Unifrac and although its performance on the mock4 dataset is the best under unweighted UniFrac, its performance on the other datasets is average.
+    The Deblur method returns a very small number of representative sequences (2388) and this could account for the reason for the high dissimilarity with the other methods as well as irregular performance on the mock data.
+
+  \subsection*{Synthetic interaction data}
+
+    \subsubsection*{Data generation}
+    The synthetic interaction data for the study were generated using two methods.
+    The first method, ``seqtime''~\cite{faustSignaturesEcologicalProcesses2018} utilized generalized Lotka-Volterra (gLV) equations to model the microbial community dynamics and made use of the Klemm–Eguı́luz algorithm to generate clique-based interaction networks~\cite{Rottjers2018}.
+    We used the seqtime R package to simulate communities with different numbers of species and samples (see Methods for details).
+    The second method, ``NorTA'' used the Normal to Anything (NorTA) approach coupled with a given interaction network topology to generate the abundance distribution of the microbial community~\cite{Kurtz2015}.
+    We used the spieceasi R package to simulate communities with different abundance distributions and network topologies (see Methods for details).
+    The scripts to generate these datasets can be found in the synthetic data and scripts repository (\href{https://github.com/segrelab/MiCoNE-synthetic-data}{https://github.com/segrelab/MiCoNE-synthetic-data})
+
+    \subsubsection*{Processing using the \ac{micone} pipeline}
+    The data was processed using the \ac{micone} pipeline using only the \ac{ni} step with the consensus networks as the final output.
+    The configuration files (main.nf and nextflow.config) used to run the \ac{micone} pipeline as well the details of the pipeline execution (dag, report, timeline and trace) are in the "runs/norta" and "runs/seqtime" directories of the data and scripts repository (\href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper})
+    The results of the pipeline execution for reproducing the analyses in the manuscript are stored on Zenodo.
+
+  \subsection*{Network metrics}
+
+  In Table~\ref{tab:network_metrics} we show various global network metrics calculated for each tool in the pipeline.
+  %\todo[noline]{R2.1a}
+  %\todo[noline]{R2.7}
+  All the networks that make use of a particular tool are grouped together, and the following average metrics are calculated for each group:
+  \begin{enumerate}
+    \item The average shortest path length describes the average of all the shortest paths in the graph.
+No number is reported if the graph is not connected, therefore, the results indicate that none of the networks that make use of \acs{harmonies}, \acs{cozine}, \acs{spring}, \acs{spieceasi} and Pearson are connected.
+    \item The average clustering is the average clustering coefficient of the graph.
+The closer the value is to 1.0, the more densely connected is the graph.
+We can observe that the networks that use correlation-based methods have the highest values while the direct association based methods have the lowest.
+    \item The number of connected components is the highest for the direct association based methods and the lowest for the correlation-based methods.
+In the case of propr, all the networks have only one giant component.
+    \item The modularity metric is the modularity over all partitions in a graph calculated using a label propagation algorithm~\cite{cordascoCommunityDetectionSemisynchronous2010}.
+Positive values imply that there are more edges between vertices of the same type than we would expect by chance, and negative implies that there are less.
+The networks inferred by \acs{mldm} report very few edges, and skew the average modularity scores.
+This could also be an artifact of incomplete convergence of the \acs{mldm} algorithm for some combinations.
+    \item Node connectivity refers to the minimum number of nodes that must be removed from the graph to make it disconnected.
+We observe that only the networks generated using propr have a high value since most of these networks are connected.
+    \item Degree assortativity coefficient measures the similarity of connections in the graph with respect to the node degree.
+Again we observe that the direct association based methods have a negative degree of assortativity, meaning that there are many hubs in these networks.
+The correlation-based methods have positive values implying that in these networks nodes with similar degrees attach to one another.
+  \end{enumerate}
+  A weight threshold of 0.1 and a p-value threshold of 0.05 were applied to each network before the analysis.
+  All the metrics were calculated using the \texttt{networkx} Python package~\cite{hagbergExploringNetworkStructure2008}.
+
+  \subsection*{p-value merging}
+
+  Fisher~\cite{fisher_224a_1948} proposed that for $k$ independent p-values, each generated by $k$ different methods and denoted by $\bar{P}^i$ (notations are same as used in the "Consensus network and p-value merging" subsection of the Methods).
+  %\todo[noline]{R1.1}
+  The following will hold true for the statistic $\Psi$:
+  \begin{equation*}
+    \begin{aligned}
+      \Psi &= \sum_{i=1}^k -2 \log \left( \bar{P}^i \right) \\
+        \Psi &\sim \chi^2_{2k}
+    \end{aligned}
+  \end{equation*}
+
+  Brown~\cite{brown_400_1975} extended Fisher's method to dependent p-values by using a re-scaled $\chi^2$ distribution:
+  \begin{equation*}
+    \Psi \sim c \chi^2_{2f}
+  \end{equation*}
+  where, $f$ is the degrees of freedom and $c$ is the scale factor and are given by:
+  \begin{equation*}
+    f = \frac{\mathrm{E}[\Psi]^2}{\mathrm{Var}[\Psi]} ~~~\text{and}~~~ c = \frac{\mathrm{Var}[\Psi]}{2\mathrm{E}[\Psi]} = \frac{k}{f}
+  \end{equation*}
+
+  We can calculate $E[\Psi]$ and $Var[\Psi]$ under the null hypothesis that the data are drawn from a multivariate Gaussian with some covariance matrix.
+  We then use these values to parametrize a $\chi^2$ distribution from which the p-value corresponding to $\frac{\psi}{c}$ can be calculated.
+  Furthermore, Brown showed that $\mathrm{E}[\Psi]$ and $\mathrm{Var}[\Psi]$ can be calculated via the following numerical approximation:
+  \begin{equation*}
+    \mathrm{E}[\Psi] = 2k ~~~\text{and}~~~ \mathrm{Var}[\Psi] = 4k + 2\sum_{i<j} \mathrm{Cov}\left( -2\log(\bar{P}^i), -2\log(\bar{P}^j) \right)
+  \end{equation*}
+
+  The above formulation was improved by Kost and McDermott~\cite{kost_combining_2002} by further fitting a third-order polynomial to approximate the covariance
+  \begin{equation}
+    \mathrm{Cov}\left( -2\log(\bar{P}^i), -2\log(\bar{P}^j) \right) \approx 3.263 \rho_{ij} + 0.710 \rho_{ij}^2 + 0.027 \rho_{ij}^3
+    \label{eqn:suppl_covariance-pvalues}
+  \end{equation}
+  where, $\rho_{ij}$ is the correlation between method $i$ and method $j$
+
+  Using $E[\Psi]$ and $Var[\Psi]$ we then fit a $\chi^2$ distribution with the parameters $c$ and $f$.
+  Note that since, in general, $f$ will not be an integer, this should be understood as a Gamma distribution with a shape parameter $f$, as mentioned by Brown~\cite{brown_400_1975}.
+  Using this, we calculate the test $\psi$ and compute the p-value from the CDF of the $\chi^2$ distribution, given in Equation~\ref{eqn:suppl_pvalue-combined}.
+  Therefore, the final combined p-value~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016} is then given by:
+  \begin{equation}
+    \begin{aligned*}
+        & \hat{P}_j = 1.0 - \Phi_{2f}\left( \psi / c \right) \\
+        \text{where},~ &\psi = -2 \sum_{i=1}^k \log(\bar{P}^i_j) ~~~\text{and}~~~ \Phi_{2f} = \mathrm{CDF}\left( \chi^2_{2f} \right)
+    \end{aligned*}
+    \label{eqn:suppl_pvalue-combined}
+  \end{equation}
+
+  The p-value merging and consensus method in \ac{micone} (see Methods) uses Equation~\ref{eqn:suppl_covariance-pvalues} to estimate the covariance of the p-values and Equation~\ref{eqn:suppl_pvalue-combined} to merge the p-values (obtained from bootstrapping) from the different correlation methods.
+  Note that we do not use Pearson and Spearman methods in the p-value merging step and these algorithms are only used for demonstration and comparison.
+  The combined p-values are used to threshold for significance in the correlation-based networks during the consensus network step.
+
+  \subsection*{The JSON network format and network exports}
+
+    The default format \ac{micone} uses for storing the network files is the JSON (JavaScript Object Notation) format which is supported by the \acf{mind}~\cite{huResourceComparisonIntegration2022}.
+    The custom JSON schema we have designed is able to store all network-related information pertaining to nodes, links, and the metadata related to the links and datasets.
+    Additionally, \ac{micone} also supports exporting of networks into a variety of other formats such as edge lists, GML, and Cytoscape formats.
+    Since we make use of \texttt{networkx}~\cite{hagbergExploringNetworkStructure2008} for the export functionality, networks can be exported to all formats supported by the package.
+    However, not all the corresponding metadata will be exported appropriately, as most formats do not support this additional metadata.
+    The details of the format and information about importing/exporting it and other network formats can be found in the \ac{micone} documentation.
+
+  \subsection*{Supplementary discussion}
+
+  It is worth pointing out some additional, more specific, conclusions stemming from the individual steps of our analysis.
+  The different denoising/clustering methods differ mostly in their identification of sequences that are in low abundance.
+  Hence, they do not have much of an impact on the inferred co-occurrence networks when the sequences of low abundance are removed (Figure~\ref{fig:figure1}).
+  Comparison of inferred and expected reference sequences and their abundances in mock community datasets has allowed us to identify \ac{dada2} as the method which best recapitulates the expected sequence composition.
+  For the chimera checking module, we suggest using the remove bimera method since it was developed in conjunction with \ac{dada2} and its performance does not significantly differ from uchime-denovo.
+  For the current work, we have decided to focus on the tools most widely used at the time of the analysis.
+  Some tools which were not as widely used (e.g.
+dbOTU3~\cite{Olesen2017}) as well as older popular methods like mothur~\cite{Schloss2009} have not been included in the study but could be added into the pipelines in future updated analyses.
+
+  The choice of taxonomy database was found to be the most important factor in the inference of microbial co-occurrence networks, contributing $65.4\%$ of the total variance.
+  %\todo[noline]{R2.2a}
+  The frequent changes in the taxonomy nomenclature coupled with the frequency of updates to the various 16S reference databases create inherent differences \cite{Balvociute2017} in taxonomy hierarchies in these databases.
+  Our analysis revealed that no particular reference database performs better than the others across the different mock dataset benchmarks.
+  %\todo[noline]{R2.12}
+  The default reference database in the pipeline is the \ac{gg} reference database along with the ``Naive Bayes'' classifier as the query tool.
+  The reason for our choice stems from the popularity of the \ac{gg} database~\cite{parkEvaluation16SRRNA2018} in taxonomic studies, which would enable easy comparison across datasets.
+  However, for newer studies we recommend using SILVA database because of its size and taxonomic comprehensiveness~\cite{iiRESCRIPtReproducibleSequence2021} and since \ac{gg} has not been updated since 2013.
+  Additionally, a particular database might be more appropriate than the rest based on specific requirements.
+  For example, to generate a dataset that is compatible with the \ac{mind} platform~\cite{huResourceComparisonIntegration2022} \ac{ncbi} is the most appropriate choice as it guarantees compatibility of taxonomic hierarchy and therefore comparability with other datasets.
+  Furthermore, we also enable users to use custom databases~\cite{Ritari2015,iiRESCRIPtReproducibleSequence2021} with the BLAST and Naive Bayes classifiers that are incorporated into the pipeline (from \ac{qiime2}).
+  We suggest that the choice of the database should be made based on possible reported or inferred biases in the representation of given biomes in a specific databases~\cite{Balvociute2017,iiRESCRIPtReproducibleSequence2021}, as choosing taxon-specific databases have also been observed to compromise classification~\cite{rmarcelinoUseTaxonspecificReference2020}.
+
+  The \ac{op} step of the pipeline is second in its contribution to total network variance.
+  This can be attributed to the large number of nodes that are added to the final networks when the filtering is turned off.
+  Additionally, a very large number of nodes also decreases the accuracy of the network inference algorithms for the same sample size~\cite{peschelNetCoMiNetworkConstruction2020} and increases the computational complexity~\cite{tackmannRapidInferenceDirect2019}.
+  We observe that filtering out taxa that are present in low abundances in all samples increases the proportion of taxa in common between taxonomy tables generated using different reference databases (Figure~\ref{fig:figure_s4}), providing another reason for filtering.
+  We also observe that the reduction in the number of taxa leads to a better agreement in the networks inferred through different methods (Figure~\ref{fig:figure6} and Figure~\ref{fig:figures1}).
+  Moreover, filtering is necessary in order to increase the power in tests of significance when the number of taxa is much greater than the number of samples.
+
+\newpage
+  \pagenumbering{goggle}
+  \thispagestyle{empty}
+  \input{sections/table1.tex}
+  \thispagestyle{empty}
+  \input{sections/table_s1.tex}
 

### 2022-12-01 final_submission/manuscript.tex
Change: Address R3 comments

@@
  This wealth of information has the potential to help us understand how communities assemble and operate.
- In particular, a powerful tool for translating microbiome composition data into knowledge is the construction of association (co-occurrence or correlation) networks, in which  microbial taxa are represented by nodes, and frequent co-occurrences (or negative correlations) across datasets are encoded as edges between nodes.
- While the relationship between directly measured interactions~\cite{lubbeExometabolomicAnalysisCrossFeeding2017,Jian2020,Hsu2019} and statistically inferred co-occurrence is still poorly understood \cite{Zuniga2017,Rottjers2018}, a significant amount of effort has gone into estimating correlations from large microbiome sequence datasets~\cite{faustMicrobialCooccurrenceRelationships2012,leeCrosskingdomCooccurrenceNetworks2022,faustMicrobialInteractionsNetworks2012a,maEarthMicrobialCooccurrence2020a}.
+ In particular, a powerful tool for translating microbiome composition data into knowledge is the construction of association (co-occurrence) networks, in which  microbial taxa are represented by nodes, and frequent co-occurrences (or negative co-occurrences) across datasets are encoded as edges between nodes.
+ While the relationship between directly measured interactions~\cite{lubbeExometabolomicAnalysisCrossFeeding2017,Jian2020,Hsu2019} and statistically inferred co-occurrence is still poorly understood \cite{Zuniga2017,Rottjers2018}, a significant amount of effort has gone into estimating co-occurrences from large microbiome sequence datasets~\cite{faustMicrobialCooccurrenceRelationships2012,leeCrosskingdomCooccurrenceNetworks2022,faustMicrobialInteractionsNetworks2012a,maEarthMicrobialCooccurrence2020a}.
 
@@
 Numerous software tools and pipelines have been developed to analyze 16S sequencing data, with a strong emphasis on the known limitations of this method, including resolution, sequencing depth, compositional nature, sequencing errors, and copy number variations \cite{Bharti2019,pollockMadnessMicrobiomeAttempting2018}.
-Popular methods for different phases of the analysis of 16S data include tools for: (i) quality checking and trimming the sequencing reads; (ii) denoising and clustering the trimmed reads \cite{Caporaso2010,Callahan2016,Amir2017}; (iii) assigning taxonomy to the denoised reads \cite{bokulichOptimizingTaxonomicClassification2018}; (iv) processing and transforming the taxonomy count matrices \cite{Weiss2015}; and (v) inferring the co-occurrence network \cite{Watts2018,Kurtz2015,tackmannRapidInferenceDirect2019}.
+Popular methods for different phases of the analysis of 16S data include tools for: (i) quality checking and trimming the sequencing reads; (ii) denoising and clustering the trimmed reads \cite{Caporaso2010,Callahan2016,Amir2017}; (iii) assigning taxonomy to the denoised reads \cite{bokulichOptimizingTaxonomicClassification2018}; (iv) processing and transforming the taxonomy count matrices \cite{weissNormalizationMicrobialDifferential2017}; and (v) inferring the co-occurrence network \cite{Watts2018,Kurtz2015,tackmannRapidInferenceDirect2019}.
 Different specific algorithms are often aggregated into popular online platforms (like MG-RAST~\cite{keeganMGRASTMetagenomicsService2016}, Qiita~\cite{gonzalezQiitaRapidWebenabled2018}) and software packages (such as \ac{qiime2}~\cite{bolyenReproducibleInteractiveScalable2019}).
@@
    The ten network inference methods we used in this step fall into two groups: the first set of methods (Pearson, Spearman, \acs{sparcc}~\cite{Friedman2012,Watts2018}, and propr~\cite{quinnProprRpackageIdentifying2017}) infer pairwise correlations while the second set (\acs{spieceasi}~\cite{Kurtz2015}, FlashWeave~\cite{tackmannRapidInferenceDirect2019}, \acs{cozine}~\cite{haCompositionalZeroinflatedNetwork2020a}, \acs{harmonies}~\cite{jiangHARMONIESHybridApproach2020}, \acs{spring}~\cite{yoonMicrobialNetworksSPRING2019}, and \acs{mldm}~\cite{Yang2017}) infer direct associations.
+   In general, when we refer to co-occurrences we include the associations from both correlation based methods and direct association based methods.
+   Whereas, correlations exclusively refer to those inferred by the correlation methods.
    Note that while Pearson and Spearman methods are included in the pipeline for completeness, they tend to generate a large number of spurious edges as they are not intended for compositional datasets.
@@
   %\todo[noline]{R2.5}
-  This is performed either by sub-sampling or by normalization of the matrix to the lowest sampling depth~\cite{Weiss2015}.
+  This is performed either by sub-sampling of the matrix to a specified rarefaction depth~\cite{weissNormalizationMicrobialDifferential2017} in order to obtain samples with equal library sizes.
   However, due to the potential biases and false positives~\cite{mcmurdieWasteNotWant2014,chaoCoveragebasedRarefactionExtrapolation2012} that might arise during the process, the rarefaction module is disabled by default and can be enabled in the configuration if needed.
-  Hence, although the pipeline supports normalization and rarefaction, these modules are turned off by default, and the analyses reported in the paper do not normalize or rareify the counts matrices.
-  However, most of the network inference methods perform normalization and other transformation operations on the counts matrix as a part of their workflow.
+  Hence, although the pipeline supports rarefaction, it is turned off by default.
+  In addition to rarefaction, the \ac{micone} pipeline also supports total sum scaling (TSS) and the centered log-ratio (clr) transformation (from the speiceasi R package).
+  However, since most of the network inference methods perform normalization and other transformation operations on the counts matrix as a part of their workflow, the analyses reported in the paper do not explicitly normalize the counts matrices.
   Filtering, is performed to remove samples or features (\ac{otu}s or \ac{esv}s) from the counts matrix that are sparse.

### 2023-02-01 final_submission/manuscript.tex
Change: Update Figure 5, text + discussion

@@
   This counts table was used as the input to the \ac{micone} pipeline to generate predicted associations.
-  The interaction network used to generate the counts table was used as the source of true interactions to calculate the precision (Figure \ref{fig:figure5}) and sensitivity (Figure \ref{fig:figure_s5} and Figure \ref{fig:figure_s6}) for each network inference algorithm.
+  The interaction network used to generate the counts table was used as the source of true interactions to calculate the precision and sensitivity (Figure \ref{fig:figure5}) of the consensus algorithms and the four best-performing individual network inference methods (propr, FlashWeave, \acs{sparcc}, and \acs{spieceasi}).
   As shown in Figure \ref{fig:figure5} the consensus algorithm, especially the scaled-sum method, captures true associations with high precision (through the removal of edges that are either not present in most of the inference methods or whose association strength is low across methods).
+  Although this increase in precision is followed by a drop in sensitivity (as parameter $\theta$ increases), this approach is advantageous as the consensus networks provide the means to obtain a short list of associations that would have a high likelihood of being present in the real association network.
   Overall, the scaled-sum method for $\theta=1.000$ performs the best (precision = $1.000$ for both NorTA and seqtime).
+  Additionally, Figure \ref{fig:figure_s5} and Figure \ref{fig:figure_s6} show precision and sensitivity values of all network inference and consensus algorithms for each interaction network in the synthetic datasets.
   The scaled-sum method for $\theta=0.333$ (default option in the pipeline) shows a high precision ($0.956$ with NorTA; $0.688$ with seqtime), without displaying significant reduction in sensitivity (Figure~\ref{fig:figure_s5} and Figure~\ref{fig:figure_s6}).
   However, if higher precision is required $\theta>0.5$ can be considered.
-  Therefore, the consensus networks provide the means to obtain a short list of associations that would have a high likelihood of being present in the real association network.
 
@@
   While our analysis is based on several mock and synthetic datasets that cover a diverse range of abundance distributions and network topologies, datasets that have drastically different distributions may require a re-assessment of the best settings.
+  However, the \ac{micone} pipeline provides a platform for easy evaluation of accuracy, variance and other properities at each step of the workflow for any other dataset of interest.
 
@@
   \vspace{-5mm}
-  In Figure~\ref{fig:figure5} we evaluated the accuracy and robustness of the inferred association networks (using the various network inference algorithms and the consensus methods) against the original interaction network used to create the taxonomic distribution.
+  In Figure~\ref{fig:figure5} we evaluated the precision and sensitivity of the inferred association networks (using the various network inference algorithms and the consensus methods) against the original interaction network used to create the taxonomic distribution.
   We used the following formulations of precision and sensitivity to calculate the accuracy of the predictions:
@@
       \textbf{The associations generated by the scaled-sum consensus method show high precision in benchmarks using synthetic datasets}.
-    The different points on the box plot show the precision of co-occurrence networks generated through individual network inference methods and through consensus network construction approaches.
-Precision is estimated based on the comparisons with two sets of synthetic benchmark datasets (``NorTA'' and ``seqtime'', see Methods).
+    The different points on the box plot show the precision (red) and sensitivity (blue) of co-occurrence networks generated through individual network inference methods and through consensus network construction approaches.
+Precision and sensitivity are estimated based on the comparisons with two sets of synthetic benchmark datasets (``NorTA'' and ``seqtime'', see Methods).
       The independent algorithms chosen for the comparison are the two best-performing correlation-based (propr, sparcc) and direct association based (spieceasi, flashweave) methods.
       For consensus network inference, we used the scaled-sum (SS) and simple voting (SV) methods.
-      A weight threshold of 0.1 and a p-value threshold of 0.05 was applied to each network before the calculation of precision.
+      A weight threshold of 0.1 and a p-value threshold of 0.05 was applied to each network before the calculation of precision and sensitivity.
+      The purpose behind the construction of the consensus algorithms is to capture true associations in the data through the removal of assocations that have a lower probability of being present in the networks inferred by different inference algorithms.
+      Therefore, an increase in precision is followed by a decrease in sensitivity.
       The overall best precision was consistently obtained by the scaled-sum consensus method for $\theta \geq 0.333$ on both benchmark datasets.
@@
       Therefore, we set the scaled-sum consensus method with $\theta = 0.333$ as the default tool for consensus network inference, since this option provides a good balance of precision and sensitivity (see also Figure~\ref{fig:figure_s5} and \ref{fig:figure_s6}).
+      The correlation-based methods (propr, and \acs{sparcc}) and the simple-voting consensus method return networks with higher sensitivities.
     }

### 2023-03-08 final_submission/manuscript.tex
Update on Overleaf.

@@
   This counts table was used as the input to the \ac{micone} pipeline to generate predicted associations.
-  The interaction network used to generate the counts table was used as the source of true interactions to calculate the precision and sensitivity (Figure \ref{fig:figure5}) of the consensus algorithms and the four best-performing individual network inference methods (propr, FlashWeave, \acs{sparcc}, and \acs{spieceasi}).
+  The interaction network used to generate the counts table was used as the source of true interactions to calculate the precision and sensitivity (Figure \ref{fig:figure5}) of the consensus algorithms.
+  These values are also compared with the precision and sensitivity of four individual network inference methods (two correlation-based, i.e.
+propr, FlashWeave, and two direct association methods, i.e.
+\acs{sparcc}, and \acs{spieceasi}), which were selected based on their high precision (see Figure ~\ref{fig:figure_s5} and Figure ~\ref{fig:figure_s6}).
   As shown in Figure \ref{fig:figure5} the consensus algorithm, especially the scaled-sum method, captures true associations with high precision (through the removal of edges that are either not present in most of the inference methods or whose association strength is low across methods).
-  Although this increase in precision is followed by a drop in sensitivity (as parameter $\theta$ increases), this approach is advantageous as the consensus networks provide the means to obtain a short list of associations that would have a high likelihood of being present in the real association network.
+  Although this increase in precision is associated with a drop in sensitivity (as the consensus parameter $\theta$ increases), the consensus networks provide valuable and practically helpful results, in the form of a short list of high-confidence associations.
   Overall, the scaled-sum method for $\theta=1.000$ performs the best (precision = $1.000$ for both NorTA and seqtime).
-  Additionally, Figure \ref{fig:figure_s5} and Figure \ref{fig:figure_s6} show precision and sensitivity values of all network inference and consensus algorithms for each interaction network in the synthetic datasets.
-  The scaled-sum method for $\theta=0.333$ (default option in the pipeline) shows a high precision ($0.956$ with NorTA; $0.688$ with seqtime), without displaying significant reduction in sensitivity (Figure~\ref{fig:figure_s5} and Figure~\ref{fig:figure_s6}).
+  Additionally, Figure \ref{fig:figure_s5} and Figure \ref{fig:figure_s6} show the precision and sensitivity values of all network inference and consensus algorithms for each interaction network in the synthetic datasets.
+  The scaled-sum method for $\theta=0.333$ (default option in the pipeline) shows a high precision ($0.956$ with NorTA; $0.688$ with seqtime), without displaying a significant reduction in sensitivity (Figure~\ref{fig:figure_s5} and Figure~\ref{fig:figure_s6}).
   However, if higher precision is required $\theta>0.5$ can be considered.
@@
 
-  Through \ac{micone}, in addition to transparently revealing the dependence of co-occurrence networks on tool and parameter choices (see Discussion in Supplementary Text for details on the DC, TA and OP steps), we have taken advantage of our spectrum of computational options and the availability of mock and synthetic datasets, to suggest a default standard setting.
+  Through \ac{micone}, in addition to transparently revealing the dependence of co-occurrence networks on tool and parameter choices (see Discussion in Supplementary Text for details on the DC, TA, and OP steps), we have taken advantage of our spectrum of computational options and the availability of mock and synthetic datasets, to suggest a default standard setting that streamlines comparisons across datasets.
   %\todo[noline]{R1.3}
-  Additionally, we have developed a consensus approach, that can reliably generate networks that are fairly robust across multiple tool choices.
-  An important caveat related to these results is that due to the lack of a universal standard for microbial interaction data, our conclusions are based on the specific datasets used in our analysis.
-  While our analysis is based on several mock and synthetic datasets that cover a diverse range of abundance distributions and network topologies, datasets that have drastically different distributions may require a re-assessment of the best settings.
-  However, the \ac{micone} pipeline provides a platform for easy evaluation of accuracy, variance and other properities at each step of the workflow for any other dataset of interest.
+  Additionally, we have developed a consensus approach that can reliably generate fairly robust networks across multiple tool choices.
+  Even if, in the current analysis, we have shown the relevance of our approach to two very different types of microbiome datasets (human and vole stool), it is important to remember that there is no universal standard for microbial interaction data, and our conclusions are based on the specific datasets used in our analysis.
+  While our analysis is based on several mock and synthetic datasets that cover a diverse range of abundance distributions and network topologies, datasets with drastically different distributions may require a re-assessment of the best settings.
+  However, the \ac{micone} pipeline provides a platform for easy evaluation of accuracy, variance, and other properties at each workflow step for any other dataset of interest.
 
@@
  We find that the scaled-sum method performs the best on synthetic datasets, and is therefore chosen as the default for the \ac{ni} step of the pipeline.
- Notably, the consensus network displays a higher precision and returns a concise list of robust associations which represent a valuable set for experimental validation follow-up.
+ Notably, the consensus network displays a higher precision and returns a concise list of robust associations representing a valuable set for experimental validation follow-up.
 
@@
   %\todo[noline]{R2.1c}
-  As a consequence, two entities belonging to the same lineage where one entity is resolved to the Genus level and another is resolved to the Family level are treated as two different nodes in the network.
-  Thus, the development of a metric of overlap to compare nodes with shared lineages within and across networks could enable more biologically and phylogenetically relevant comparisons.
-
-  Although direct comparisons between co-occurrence networks and directly measured interactions are difficult to interpret and highly debated~\cite{hiranoDifficultyInferringMicrobial2019,gobernaCautionaryNotesUse2022}.
+  Consequently, two entities belonging to the same lineage where one entity is resolved to the Genus level and another is resolved to the Family level are treated as two different nodes in the network.
+  Thus, developing an overlap metric to compare nodes with shared lineages within and across networks could enable more biologically and phylogenetically relevant comparisons.
+
+  In thinking about the possible biological interpretations of the co-occurrence network computed by us and others, it is important to remember that there is no solid basis for assuming that these networks carry information about physical or metabolic interactions (see for example, REFS Pankaj , also FAIR ).
+  Comparing co-occurrence networks and directly measured interactions remains a major unresolved challenge~\cite{hiranoDifficultyInferringMicrobial2019,gobernaCautionaryNotesUse2022}, which needs to be investigated further.
+  Understanding this connection will enable the utilization of co-occurrence networks, for systems in which direct interaction measurements cannot be taken.
   %\todo[noline]{R2.2}
@@
       \textbf{The associations generated by the scaled-sum consensus method show high precision in benchmarks using synthetic datasets}.
-    The different points on the box plot show the precision (red) and sensitivity (blue) of co-occurrence networks generated through individual network inference methods and through consensus network construction approaches.
+      The different points on the box plot show the precision (A, C) and sensitivity (B, D) of co-occurrence networks generated through individual network inference methods and consensus network construction approaches.
       Precision and sensitivity are estimated based on the comparisons with two sets of synthetic benchmark datasets (``NorTA'' and ``seqtime'', see Methods).
-      The independent algorithms chosen for the comparison are the two best-performing correlation-based (propr, sparcc) and direct association based (spieceasi, flashweave) methods.
+      The independent algorithms chosen for the comparison are the two best-performing correlation-based (propr, sparcc) and direct-association-based (spieceasi, flashweave) methods.
       For consensus network inference, we used the scaled-sum (SS) and simple voting (SV) methods.
       A weight threshold of 0.1 and a p-value threshold of 0.05 was applied to each network before the calculation of precision and sensitivity.
-      The purpose behind the construction of the consensus algorithms is to capture true associations in the data through the removal of assocations that have a lower probability of being present in the networks inferred by different inference algorithms.
+      The purpose behind the construction of the consensus algorithms is to capture true associations in the data through the removal of associations that have a lower probability of being present in the networks inferred by different inference algorithms.
       Therefore, an increase in precision is followed by a decrease in sensitivity.
-      The overall best precision was consistently obtained by the scaled-sum consensus method for $\theta \geq 0.333$ on both benchmark datasets.
+      The scaled-sum consensus method consistently obtained the overall best precision for $\theta \geq 0.333$ on both benchmark datasets.
 Among all the individual network inference methods, spieceasi shows the best average precision.
-      The simple voting method, when using the presence of edges in all inferred networks as a requirement ($\theta = 1.000$), also outperforms spieceasi on average precision.
+      When using the presence of edges in all inferred networks as a requirement ($\theta = 1.000$), the simple voting method also outperforms spieceasi on average precision.
       Therefore, we set the scaled-sum consensus method with $\theta = 0.333$ as the default tool for consensus network inference, since this option provides a good balance of precision and sensitivity (see also Figure~\ref{fig:figure_s5} and \ref{fig:figure_s6}).
@@
   %\todo[noline]{R2.12}
-  The default reference database in the pipeline is the \ac{gg} reference database along with the ``Naive Bayes'' classifier as the query tool.
+  The default reference database in the pipeline is currently chosen to be the \ac{gg} reference database along with the ``Naive Bayes'' classifier as the query tool.
   The reason for our choice stems from the popularity of the \ac{gg} database~\cite{parkEvaluation16SRRNA2018} in taxonomic studies, which would enable easy comparison across datasets.
-  However, for newer studies we recommend using SILVA database because of its size and taxonomic comprehensiveness~\cite{iiRESCRIPtReproducibleSequence2021} and since \ac{gg} has not been updated since 2013.
+  However, we recommend using the SILVA database for newer studies due to its size and taxonomic comprehensiveness~\cite{iiRESCRIPtReproducibleSequence2021} and since \ac{gg} has not been updated since 2013.
   Additionally, a particular database might be more appropriate than the rest based on specific requirements.
-  For example, to generate a dataset that is compatible with the \ac{mind} platform~\cite{huResourceComparisonIntegration2022} \ac{ncbi} is the most appropriate choice as it guarantees compatibility of taxonomic hierarchy and therefore comparability with other datasets.
+  For example, to generate a dataset that is compatible with the \ac{mind} platform~\cite{huResourceComparisonIntegration2022}, \ac{ncbi}, is the most appropriate choice as it guarantees compatibility of taxonomic hierarchy and therefore comparability with other datasets.
+  A detailed study of other taxonomy mapping approaches[REFS IDTAXA] and integrative approaches combining different databases might offer orthogonal information and better matches.
   Furthermore, we also enable users to use custom databases~\cite{Ritari2015,iiRESCRIPtReproducibleSequence2021} with the BLAST and Naive Bayes classifiers that are incorporated into the pipeline (from \ac{qiime2}).

### 2023-03-10 final_submission/manuscript.tex
Add: Files for final submission

@@
  In particular, a powerful tool for translating microbiome composition data into knowledge is the construction of association (co-occurrence) networks, in which  microbial taxa are represented by nodes, and frequent co-occurrences (or negative co-occurrences) across datasets are encoded as edges between nodes.
+  %\todo[noline]{R3.7}
  While the relationship between directly measured interactions~\cite{lubbeExometabolomicAnalysisCrossFeeding2017,Jian2020,Hsu2019} and statistically inferred co-occurrence is still poorly understood \cite{Zuniga2017,Rottjers2018}, a significant amount of effort has gone into estimating co-occurrences from large microbiome sequence datasets~\cite{faustMicrobialCooccurrenceRelationships2012,leeCrosskingdomCooccurrenceNetworks2022,faustMicrobialInteractionsNetworks2012a,maEarthMicrobialCooccurrence2020a}.
@@
   Our analysis involved two types of data: The first type consisted of 16S sequencing data from samples of human stool microbiomes from a fecal microbiome transplant (FMT) study of autism~\cite{Kang2017}.
-  %\todo[noline]{R2.4}
+  %%\todo[noline]{R2.4}
   The second type was a collection of datasets synthetically or artificially created for the specific goal of evaluating computational analysis tools.
@@
   Based on being among the top performers on the mock datasets, their de novo error correcting nature and previous independent evaluation~\cite{Nearing2018}, \ac{dada2} and Deblur appear to be the most reliable.
+  %\todo[noline]{R2.4}
   This is because the open-reference and de novo clustering methods return a much larger number of \ac{otu}s compared to the other pipelines and would affect the accuracy of the network inference step if stringent filtering is not performed.
-  %\todo[noline]{R2.6}
+  %%\todo[noline]{R2.6}
   Overall, since \ac{dada2} as compared to Deblur, displays better performance on all the mock datasets on the weighted UniFrac metric, we set this tool as the default for the DC step of the pipeline.
-  %\todo[noline]{R2.10}
+  %%\todo[noline]{R2.10}
   However, if comparison across studies that have sequenced different 16S regions is required, closed-reference and open-reference might be a better option.
@@
   Since no database performs better than others against mock datasets, the choice of which database to use could be driven by other reasons (see Supplementary discussion).
-  %\todo[noline]{R2.2a}
+  %%\todo[noline]{R2.2a}
   One reason to choose a particular database could be the frequency of updates and the potential for future growth.
@@
   The TA step results in a taxonomic counts table that is used as input to the subsequent steps of the pipeline.
-  %\todo[noline]{R2.1b}
+  %%\todo[noline]{R2.1b}
   Note that the count tables at different levels can be obtained through aggregation; for example, Genus count tables were obtained by summing up the counts of the lower taxonomy levels (Species and \ac{otu}) that map to the same higher taxonomy level entity.
@@
    In general, when we refer to co-occurrences we include the associations from both correlation based methods and direct association based methods.
-   Whereas, correlations exclusively refer to those inferred by the correlation methods.
+  %\todo[noline]{R3.7}
+   On the other hand, when mentioning correlations, we will exclusively refer to edges inferred by the correlation methods.
    Note that while Pearson and Spearman methods are included in the pipeline for completeness, they tend to generate a large number of spurious edges as they are not intended for compositional datasets.
@@
   Filtered (see \ac{op} step in Methods) genus-level counts table obtained using the default settings in the previous steps were used as input for the different network inference algorithms (Figure~\ref{fig:figure4}).
-  %\todo[noline]{R2.1b}
+  %%\todo[noline]{R2.1b}
   Even from a visual inspection (Figure~\ref{fig:figure4}A), one can see that the different networks differ vastly in their edge-density and connectivity, with common edges often displaying inverted signs.
@@
  Inspired by previous approaches~\cite{bustinceFuzzySetsTheir2008,tsarevApplicationMajorityVoting2018}, we developed two methods that take into consideration the evidence offered by each network inference algorithm and generate a consensus network that contains the common edges among the inferred networks.
-  %\todo[noline]{R1.1}
-  %\todo[noline]{R1.4}
+  %%\todo[noline]{R1.1}
+  %%\todo[noline]{R1.4}
 
@@
 Currently, by default, all direct association methods (\acs{spieceasi}, \acs{cozine}, \acs{harmonies}, \acs{spring}, \acs{mldm}, and FlashWeave) are used, together with \acs{sparcc} and propr.
+  %\todo[noline]{R3.2}
 
@@
   As shown in Figure \ref{fig:figure5} the consensus algorithm, especially the scaled-sum method, captures true associations with high precision (through the removal of edges that are either not present in most of the inference methods or whose association strength is low across methods).
+  %\todo[noline]{R1}
+  %\todo[noline]{R2.1}
   Although this increase in precision is associated with a drop in sensitivity (as the consensus parameter $\theta$ increases), the consensus networks provide valuable and practically helpful results, in the form of a short list of high-confidence associations.
   Overall, the scaled-sum method for $\theta=1.000$ performs the best (precision = $1.000$ for both NorTA and seqtime).
-  Additionally, Figure \ref{fig:figure_s5} and Figure \ref{fig:figure_s6} show the precision and sensitivity values of all network inference and consensus algorithms for each interaction network in the synthetic datasets.
-  The scaled-sum method for $\theta=0.333$ (default option in the pipeline) shows a high precision ($0.956$ with NorTA; $0.688$ with seqtime), without displaying a significant reduction in sensitivity (Figure~\ref{fig:figure_s5} and Figure~\ref{fig:figure_s6}).
+  Figure \ref{fig:figure_s5} and Figure \ref{fig:figure_s6} show the precision and sensitivity values of all network inference and consensus algorithms for each interaction network in the synthetic datasets.
+  The scaled-sum method for $\theta=0.333$ (default option in the pipeline) shows a high precision ($0.956$ with NorTA; $0.688$ with seqtime), without displaying a significant reduction in sensitivity (Figure~\ref{fig:figure5}, Figure~\ref{fig:figure_s5} and Figure~\ref{fig:figure_s6}).
   However, if higher precision is required $\theta>0.5$ can be considered.
@@
   It is also evident that in the NI step, some networks, especially those inferred using the direct association network inference methods, are much closer in the PCA plot regardless of the reference database used.
+  The network variance analysis performed on the supplementary dataset~\cite{shafferStandardizedMultiomicsEarth2022} (stool samples from radiation-exposed bank vole) shown in Figure~\ref{fig:figure_s8} support these observations, implying that these findings are fairly consistent across different datasets.
   These results suggest that the most important criterion for accurate comparative analysis of co-occurrence networks is the taxonomy reference database followed by the level of filtering of the taxonomy tables and the network inference algorithm used.
@@
   The analysis of the rewiring of associations in the ASD samples with respect to the control provides a guide for the identification of key genera that could be linked to dysbiosis.
-  %\todo[noline]{R2.14}
+  %%\todo[noline]{R2.14}
   We observed 22 unique links in the network for control samples, 12 unique links in the network for ASD subjects, and 7 edges in common between the two networks.
@@
   Through \ac{micone}, in addition to transparently revealing the dependence of co-occurrence networks on tool and parameter choices (see Discussion in Supplementary Text for details on the DC, TA, and OP steps), we have taken advantage of our spectrum of computational options and the availability of mock and synthetic datasets, to suggest a default standard setting that streamlines comparisons across datasets.
-  %\todo[noline]{R1.3}
+  %%\todo[noline]{R1.3}
   Additionally, we have developed a consensus approach that can reliably generate fairly robust networks across multiple tool choices.
@@
   The network analyses in this study are primarily at the Genus level, wherein the lowest resolution of a node is a Genus and if an entity cannot be resolved to the Genus level, the next lowest taxonomic level is used (for example, Family).
-  %\todo[noline]{R2.1c}
+  %%\todo[noline]{R2.1c}
   Consequently, two entities belonging to the same lineage where one entity is resolved to the Genus level and another is resolved to the Family level are treated as two different nodes in the network.
@@
 
-  In thinking about the possible biological interpretations of the co-occurrence network computed by us and others, it is important to remember that there is no solid basis for assuming that these networks carry information about physical or metabolic interactions (see for example, REFS Pankaj , also FAIR ).
+  In thinking about the possible biological interpretations of the co-occurrence network computed by us and others, it is important to remember that there is no solid basis for assuming that these networks carry information about physical or metabolic interactions~\cite{fisherIdentifyingKeystoneSpecies2014,pachecoFAIRRepresentationsMicrobial2022}.
   Comparing co-occurrence networks and directly measured interactions remains a major unresolved challenge~\cite{hiranoDifficultyInferringMicrobial2019,gobernaCautionaryNotesUse2022}, which needs to be investigated further.
-  Understanding this connection will enable the utilization of co-occurrence networks, for systems in which direct interaction measurements cannot be taken.
-  %\todo[noline]{R2.2}
+  Understanding this connection will be benficial for predicting interactions in systems where direct interaction measurements cannot be taken.
+  %%\todo[noline]{R2.2}
   Further, benchmarking of co-occurrence networks could also be pursued through the use of literature-based interactions~\cite{lima-mendezDeterminantsCommunityStructure2015a} or biological benchmark interaction data~\cite{sungGlobalMetabolicInteraction2017a}.
@@
   Although in the current analysis, we have only used default parameter values recommended by the tool creators, the \ac{micone} pipeline could be used in the future to explore any combinations of parameters and to optimize these values for improved network inference.
-  %\todo[noline]{R1.3}
+  %%\todo[noline]{R1.3}
   Overall, there likely is no ``best method'' for the various steps of 16S data analysis, and hence, \ac{micone} is intended to help researchers to identify the methods and algorithms that are most suitable for their datasets in an easy-to-use and reproducible manner.
@@
   This study utilized two types of 16S rRNA sequencing datasets: biological datasets and mock/synthetic datasets.
-  %\todo[noline]{R2.4}
+  %%\todo[noline]{R2.4}
   Biological datasets are collections of sequencing reads obtained from naturally occurring microbial community samples.
@@
   The mock16 community is composed of 49 bacteria and 10 Archaea, all represented in equal amounts.
+  In addition to these datasets, we have utilized a dataset containing stool samples from radiation-exposed bank vole~\cite{shafferStandardizedMultiomicsEarth2022} (Qiita study ID 13114) for supplementary analyses.
 
@@
   This modular architecture simplifies the process of adding new tools (refer to the modules section in the \ac{micone} documentation).
-  %\todo[noline]{R2.11}
+  %%\todo[noline]{R2.11}
   The main components of the pipeline are detailed in the subsequent sections.
@@
   The module also supports rarefaction, which is a normalization technique used to overcome the bias that might arise due to variable sampling depth in different samples.
-  %\todo[noline]{R2.5}
+  %%\todo[noline]{R2.5}
   This is performed either by sub-sampling of the matrix to a specified rarefaction depth~\cite{weissNormalizationMicrobialDifferential2017} in order to obtain samples with equal library sizes.
+  %\todo[noline]{R3.5}
   However, due to the potential biases and false positives~\cite{mcmurdieWasteNotWant2014,chaoCoveragebasedRarefactionExtrapolation2012} that might arise during the process, the rarefaction module is disabled by default and can be enabled in the configuration if needed.
@@
   In addition to rarefaction, the \ac{micone} pipeline also supports total sum scaling (TSS) and the centered log-ratio (clr) transformation (from the speiceasi R package).
+  %\todo[noline]{R3.4}
   However, since most of the network inference methods perform normalization and other transformation operations on the counts matrix as a part of their workflow, the analyses reported in the paper do not explicitly normalize the counts matrices.
@@
   The group operation transforms the \ac{otu} or \ac{esv} count matrix into a taxonomic count matrix at the requested level by adding up counts that map to the same taxonomy and is carried out at the end of the OP step.
-  %\todo[noline]{R2.1b}
+  %%\todo[noline]{R2.1b}
   Finally, transformations are performed in order to correct for and overcome the compositional bias that is inherent in the counts matrix (in the analysis performed in the study these were disabled and directly handled by the network inference algorithm).
@@
   The consensus methods combine networks inferred from both correlation-based and direct association methods.
-  %\todo[noline]{R1.1}
-  %\todo[noline]{R1.4}
+  %%\todo[noline]{R1.1}
+  %%\todo[noline]{R1.4}
   First, for the correlation-based methods, we calculate p-values using null models and then merge the p-values using Brown's p-value merging method~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016,faustCoNetAppInference2016}.
-  %\todo[noline]{R2.2}
-  %\todo[noline]{R2.3}
+  %%\todo[noline]{R2.2}
+  %%\todo[noline]{R2.3}
   Second, we filter all the inferred networks based on an association strength threshold of 0.1 and a p-value cutoff of 0.05.
@@
   The next step in the consensus algorithm workflow is to merge the p-values for the networks generated by the correlation-based methods.
-  %\todo[noline]{R1.1}
+  %%\todo[noline]{R1.1}
   This step is performed using the Brown's p-value merging method~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016,faustCoNetAppInference2016}.
@@
   This section defines the notation used for the network variability analysis performed for Figure~\ref{fig:figure6}.
-  %\todo[noline]{R1.2}
-  %\todo[noline]{R1.8}
-  %\todo[noline]{R2.1b,d}
+  %%\todo[noline]{R1.2}
+  %%\todo[noline]{R1.8}
+  %%\todo[noline]{R2.1b,d}
 
@@
    Note that row $j$ of $\mathbf{\bar{N}}$, $\bar{N}_j$ is the vector that encodes the values of edge $j$ across all the networks.
-  %\todo[noline]{R1.5}
-  %\todo[noline]{R2.1b}
+  %%\todo[noline]{R1.5}
+  %%\todo[noline]{R2.1b}
 
@@
   To infer the variability contributed due to the different steps in the pipeline we can perform a linear regression on each edge in $\mathbf{\bar{N}}$ and a subsequent ANOVA to extract the within-group variances.
-  One issue with this approach is that the possibility of correlations existing between the edges of the network could lead to inaccurate estimates of the variance if a linear model were used to directly model the relationships between edges and steps in the workflow.
-  In order to remedy this issue, we performed a PCA (Principal Component Analysis) on the matrix $\mathbf{\bar{N}}$ to obtain the $\mathbf{C}$ matrix ($W \times c$) of components for each network, such that we reduce the dimensions from the $Q^2$ dimensional edge space to a $c$ dimensional component space.
+  A major issue with this approach is that the possibility of correlations existing between the edges of the network could lead to inaccurate estimates of the variance if a linear model were used to directly model the relationships between edges and steps in the workflow.
+  %\todo[noline]{R2.5}
+  Therefore, in order to remedy this issue, we performed a PCA (Principal Component Analysis) on the matrix $\mathbf{\bar{N}}$ to obtain the $\mathbf{C}$ matrix ($W \times c$) of components for each network, such that we reduce the dimensions from the $Q^2$ dimensional edge space to a $c$ dimensional component space.
 
@@
    The variance contributed by each step of the pipeline was calculated for every component in $\mathbf{C}$ matrix through ANOVA using the Python statsmodels~\cite{seaboldStatsmodelsEconometricStatistical2010} package and is shown in Figure~\ref{fig:figure6}A.
-  %\todo[noline]{R1.6}
+  %%\todo[noline]{R1.6}
   The total variance for the network was calculated by adding the variances for each connection and normalizing with the degrees of freedom.
   The merged network table $\mathbf{\bar{N}}$ was used as the input to the PCA analysis to generate Figure~\ref{fig:figure6}B.
-  %\todo[noline]{R1.7}
+  %%\todo[noline]{R1.7}
 
@@
   We generated synthetic interaction data using two methodologies previously used for benchmarking network inference methods.
-  %\todo[noline]{R1.4}
-  %\todo[noline]{R2.2}
-  %\todo[noline]{R2.4}
+  %%\todo[noline]{R1.4}
+  %%\todo[noline]{R2.2}
+  %%\todo[noline]{R2.4}
 
@@
 
+  % \begin{figure}[H]
+  %   \centering
+  %   \includegraphics[width=1.0\linewidth]{figure_s8.pdf}
+  % \end{figure}
+  \begin{figure}[H]
+    \centering
+    \caption{
+      \textbf{Network variance analysis of the stool samples from radiation-exposed bank vole (EMP dataset)}.
+      \textbf{(A)} The percentage of variance in the networks contributed by the \acf{dc}, CC (chimera checking), \acf{ta}, \acf{op} and \acf{ni} steps of the pipeline calculated using ANOVA on a linear model (see Methods).
+      A weight threshold of 0.1 and a p-value threshold of 0.05 were applied to each network before the analysis.
+      The taxonomy database contributes most to the variance between the networks (57.96\%) followed by the filtering of the counts matrix (33.48\%) in the OP step.
+    The variation due to the NI, DC and CC steps are much smaller in comparison (6.69\%, 1.30\%, and 0.00\% respectively).
+      The negligible fraction labeled as the residual is an artifact that arises when multiple steps are changed at the same time.
+      \textbf{(B)} All the inferred networks generated from various combinations of tools are shown as points on a PCA plot.
+      Each point on the PCA plot represents a network inferred using different combinations of tools and parameters that are available in the \ac{micone} pipeline.
+      The color of the points corresponds to the tools used at each step of the pipeline (DC, TA, OP, and NI).
+      The points on the PCA plot can be largely grouped based on the TA step, and the extent of this separation decreases when the filtering is turned on in the OP step.
+      However, unlike Figure~\ref{fig:figure6} B, here we observe that the networks that utilize \ac{gg} and \ac{ncbi} databases have the variation along similar axes in the 2 dimensional PCA plot.
+      This could imply that the taxonomy assignments returned by these two databases for this dataset, might be more similar to each other compared to those in the FMT dataset.
+    }
+    \label{fig:figure_s8}
+  \end{figure}
+  % \FloatBarrier
+  % \newpage
+
   \setcounter{figure}{0}
@@
     \centering
-    \caption{Supplemental text describing data processing, synthetic interaction data generation, network metrics, p-value merging and supplementary discussion sections.}
+    \caption{Supplemental text describing data processing, synthetic interaction data generation, network metrics, p-value merging, additional network variance analysis and supplementary discussion sections.}
     \label{txt:text_s1}
@@
 
-\newpage
-\section*{Supplementary Text}
-
-  \subsection*{Processing the FMT data}
-
-    \subsubsection*{Data download and pre-processing}
-    The main biological dataset used in this study was the collection of 16S rRNA sequencing reads from stool samples (healthy and autistic individuals) for a fecal microbiome transplant study~\cite{Kang2017}.
-    The data containing the 16S sequencing reads (V4 region) was downloaded from Qiita~\cite{gonzalezQiitaRapidWebenabled2018} (study ID: 10532).
-    Only runs 2, 3, and 4 were used for the subsequent analysis as these runs consisted of paired-end sequencing data, and run 1 contained single-end data.
-    The sample metadata was updated to contain only BMI, sex, height, weight, and experimental group.
-    This was necessary as two of the network inference algorithms (\acs{mldm} and FlashWeave) required information about environmental heterogeneity.
-    However, these environmental correlations were not included in the current analyses.
-
-    \subsubsection*{Processing using the \ac{micone} pipeline}
-    The data was then processed using the \ac{micone} pipeline starting at the \ac{sp} step and ending at the \ac{ni} step with the consensus algorithm.
-    The configuration files (main.nf and nextflow.config) used to run the \ac{micone} pipeline as well the details of the pipeline execution (dag, report, timeline and trace) are in the "runs/FMT" directory of the data and scripts repository (\href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper})
-    The results of the pipeline execution for reproducing the analyses in the manuscript are stored on Zenodo.
-
-  \subsection*{Processing the mock data}
-
-    \subsubsection*{Data download and pre-processing}
-    The mock datasets, mock4, mock12 and mock16 used for this study, were obtained from mockrobiota~\cite{Bokulich2016}.
-    Mock 4 is a mock community composed of 21 bacterial strains represented in equal abundances in two replicate samples, and the same strains represented in uneven abundances in two other replicate samples.
-    Mock 12 is composed of 27 bacterial strains containing closely related taxa, the members of which were chosen in part for their well-separated 16S rRNA gene sequences.
-Some pairs of strains differ by as little as one nucleotide, but all the strains are distinguishable over the sequenced region of the 16S rRNA gene.
-    Mock 16 is a mock community composed of even amounts of purified genomic DNA from 49 bacteria and 10 archaea.
-    The datasets did not require any preprocessing and could be directly used as input to the pipeline
-
-    \subsubsection*{Processing using the \ac{micone} pipeline}
-    The data was processed using the \ac{micone} pipeline starting at the \ac{sp} step and ending at the \ac{op} step with the filtered taxonomic tables as the final output.
-    The configuration files (main.nf and nextflow.config) used to run the \ac{micone} pipeline as well the details of the pipeline execution (dag, report, timeline and trace) are in the "runs/mock*" directory of the data and scripts repository (\href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper})
-    The results of the pipeline execution for reproducing the analyses in the manuscript are stored on Zenodo.
-
-    \subsubsection*{Interpretation of Unifrac results in the DC step}
-  In Figure \ref{fig:figure_s3}, we observe that both the weighted and unweighted UniFrac distances are increased for the top 1000 representative sequences, implying that the top representative sequences generated by the different methods are not as similar to each other.
-  Therefore, since the weighted UniFrac distances are lower than the unweighted distances, we conclude that the representative sequences in the middle range of the abundance distribution are those that must be the most similar between the methods.
-
-    Open-reference and de novo clustering methods perform the best under the weighted UniFrac metric and the worst (marginally) under the unweighted UniFrac metric (Figure~\ref{fig:figure2}C and \ref{fig:figure2}D).
-    This result can be attributed to the large number of low abundance representative sequences that are generated by these methods.
-    Deblur performs poorly under weighted Unifrac and although its performance on the mock4 dataset is the best under unweighted UniFrac, its performance on the other datasets is average.
-    The Deblur method returns a very small number of representative sequences (2388) and this could account for the reason for the high dissimilarity with the other methods as well as irregular performance on the mock data.
-
-  \subsection*{Synthetic interaction data}
-
-    \subsubsection*{Data generation}
-    The synthetic interaction data for the study were generated using two methods.
-    The first method, ``seqtime''~\cite{faustSignaturesEcologicalProcesses2018} utilized generalized Lotka-Volterra (gLV) equations to model the microbial community dynamics and made use of the Klemm–Eguı́luz algorithm to generate clique-based interaction networks~\cite{Rottjers2018}.
-    We used the seqtime R package to simulate communities with different numbers of species and samples (see Methods for details).
-    The second method, ``NorTA'' used the Normal to Anything (NorTA) approach coupled with a given interaction network topology to generate the abundance distribution of the microbial community~\cite{Kurtz2015}.
-    We used the spieceasi R package to simulate communities with different abundance distributions and network topologies (see Methods for details).
-    The scripts to generate these datasets can be found in the synthetic data and scripts repository (\href{https://github.com/segrelab/MiCoNE-synthetic-data}{https://github.com/segrelab/MiCoNE-synthetic-data})
-
-    \subsubsection*{Processing using the \ac{micone} pipeline}
-    The data was processed using the \ac{micone} pipeline using only the \ac{ni} step with the consensus networks as the final output.
-    The configuration files (main.nf and nextflow.config) used to run the \ac{micone} pipeline as well the details of the pipeline execution (dag, report, timeline and trace) are in the "runs/norta" and "runs/seqtime" directories of the data and scripts repository (\href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper})
-    The results of the pipeline execution for reproducing the analyses in the manuscript are stored on Zenodo.
-
-  \subsection*{Network metrics}
-
-  In Table~\ref{tab:network_metrics} we show various global network metrics calculated for each tool in the pipeline.
-  %\todo[noline]{R2.1a}
-  %\todo[noline]{R2.7}
-  All the networks that make use of a particular tool are grouped together, and the following average metrics are calculated for each group:
-  \begin{enumerate}
-    \item The average shortest path length describes the average of all the shortest paths in the graph.
-No number is reported if the graph is not connected, therefore, the results indicate that none of the networks that make use of \acs{harmonies}, \acs{cozine}, \acs{spring}, \acs{spieceasi} and Pearson are connected.
-    \item The average clustering is the average clustering coefficient of the graph.
-The closer the value is to 1.0, the more densely connected is the graph.
-We can observe that the networks that use correlation-based methods have the highest values while the direct association based methods have the lowest.
-    \item The number of connected components is the highest for the direct association based methods and the lowest for the correlation-based methods.
-In the case of propr, all the networks have only one giant component.
-    \item The modularity metric is the modularity over all partitions in a graph calculated using a label propagation algorithm~\cite{cordascoCommunityDetectionSemisynchronous2010}.
-Positive values imply that there are more edges between vertices of the same type than we would expect by chance, and negative implies that there are less.
-The networks inferred by \acs{mldm} report very few edges, and skew the average modularity scores.
-This could also be an artifact of incomplete convergence of the \acs{mldm} algorithm for some combinations.
-    \item Node connectivity refers to the minimum number of nodes that must be removed from the graph to make it disconnected.
-We observe that only the networks generated using propr have a high value since most of these networks are connected.
-    \item Degree assortativity coefficient measures the similarity of connections in the graph with respect to the node degree.
-Again we observe that the direct association based methods have a negative degree of assortativity, meaning that there are many hubs in these networks.
-The correlation-based methods have positive values implying that in these networks nodes with similar degrees attach to one another.
-  \end{enumerate}
-  A weight threshold of 0.1 and a p-value threshold of 0.05 were applied to each network before the analysis.
-  All the metrics were calculated using the \texttt{networkx} Python package~\cite{hagbergExploringNetworkStructure2008}.
-
-  \subsection*{p-value merging}
-
-  Fisher~\cite{fisher_224a_1948} proposed that for $k$ independent p-values, each generated by $k$ different methods and denoted by $\bar{P}^i$ (notations are same as used in the "Consensus network and p-value merging" subsection of the Methods).
-  %\todo[noline]{R1.1}
-  The following will hold true for the statistic $\Psi$:
-  \begin{equation*}
-    \begin{aligned}
-      \Psi &= \sum_{i=1}^k -2 \log \left( \bar{P}^i \right) \\
-        \Psi &\sim \chi^2_{2k}
-    \end{aligned}
-  \end{equation*}
-
-  Brown~\cite{brown_400_1975} extended Fisher's method to dependent p-values by using a re-scaled $\chi^2$ distribution:
-  \begin{equation*}
-    \Psi \sim c \chi^2_{2f}
-  \end{equation*}
-  where, $f$ is the degrees of freedom and $c$ is the scale factor and are given by:
-  \begin{equation*}
-    f = \frac{\mathrm{E}[\Psi]^2}{\mathrm{Var}[\Psi]} ~~~\text{and}~~~ c = \frac{\mathrm{Var}[\Psi]}{2\mathrm{E}[\Psi]} = \frac{k}{f}
-  \end{equation*}
-
-  We can calculate $E[\Psi]$ and $Var[\Psi]$ under the null hypothesis that the data are drawn from a multivariate Gaussian with some covariance matrix.
-  We then use these values to parametrize a $\chi^2$ distribution from which the p-value corresponding to $\frac{\psi}{c}$ can be calculated.
-  Furthermore, Brown showed that $\mathrm{E}[\Psi]$ and $\mathrm{Var}[\Psi]$ can be calculated via the following numerical approximation:
-  \begin{equation*}
-    \mathrm{E}[\Psi] = 2k ~~~\text{and}~~~ \mathrm{Var}[\Psi] = 4k + 2\sum_{i<j} \mathrm{Cov}\left( -2\log(\bar{P}^i), -2\log(\bar{P}^j) \right)
-  \end{equation*}
-
-  The above formulation was improved by Kost and McDermott~\cite{kost_combining_2002} by further fitting a third-order polynomial to approximate the covariance
-  \begin{equation}
-    \mathrm{Cov}\left( -2\log(\bar{P}^i), -2\log(\bar{P}^j) \right) \approx 3.263 \rho_{ij} + 0.710 \rho_{ij}^2 + 0.027 \rho_{ij}^3
-    \label{eqn:suppl_covariance-pvalues}
-  \end{equation}
-  where, $\rho_{ij}$ is the correlation between method $i$ and method $j$
-
-  Using $E[\Psi]$ and $Var[\Psi]$ we then fit a $\chi^2$ distribution with the parameters $c$ and $f$.
-  Note that since, in general, $f$ will not be an integer, this should be understood as a Gamma distribution with a shape parameter $f$, as mentioned by Brown~\cite{brown_400_1975}.
-  Using this, we calculate the test $\psi$ and compute the p-value from the CDF of the $\chi^2$ distribution, given in Equation~\ref{eqn:suppl_pvalue-combined}.
-  Therefore, the final combined p-value~\cite{Poole_Gibbs_Shmulevich_Bernard_Knijnenburg_2016} is then given by:
-  \begin{equation}
-    \begin{aligned*}
-        & \hat{P}_j = 1.0 - \Phi_{2f}\left( \psi / c \right) \\
-        \text{where},~ &\psi = -2 \sum_{i=1}^k \log(\bar{P}^i_j) ~~~\text{and}~~~ \Phi_{2f} = \mathrm{CDF}\left( \chi^2_{2f} \right)
-    \end{aligned*}
-    \label{eqn:suppl_pvalue-combined}
-  \end{equation}
-
-  The p-value merging and consensus method in \ac{micone} (see Methods) uses Equation~\ref{eqn:suppl_covariance-pvalues} to estimate the covariance of the p-values and Equation~\ref{eqn:suppl_pvalue-combined} to merge the p-values (obtained from bootstrapping) from the different correlation methods.
-  Note that we do not use Pearson and Spearman methods in the p-value merging step and these algorithms are only used for demonstration and comparison.
-  The combined p-values are used to threshold for significance in the correlation-based networks during the consensus network step.
-
-  \subsection*{The JSON network format and network exports}
-
-    The default format \ac{micone} uses for storing the network files is the JSON (JavaScript Object Notation) format which is supported by the \acf{mind}~\cite{huResourceComparisonIntegration2022}.
-    The custom JSON schema we have designed is able to store all network-related information pertaining to nodes, links, and the metadata related to the links and datasets.
-    Additionally, \ac{micone} also supports exporting of networks into a variety of other formats such as edge lists, GML, and Cytoscape formats.
-    Since we make use of \texttt{networkx}~\cite{hagbergExploringNetworkStructure2008} for the export functionality, networks can be exported to all formats supported by the package.
-    However, not all the corresponding metadata will be exported appropriately, as most formats do not support this additional metadata.
-    The details of the format and information about importing/exporting it and other network formats can be found in the \ac{micone} documentation.
-
-  \subsection*{Supplementary discussion}
-
-  It is worth pointing out some additional, more specific, conclusions stemming from the individual steps of our analysis.
-  The different denoising/clustering methods differ mostly in their identification of sequences that are in low abundance.
-  Hence, they do not have much of an impact on the inferred co-occurrence networks when the sequences of low abundance are removed (Figure~\ref{fig:figure1}).
-  Comparison of inferred and expected reference sequences and their abundances in mock community datasets has allowed us to identify \ac{dada2} as the method which best recapitulates the expected sequence composition.
-  For the chimera checking module, we suggest using the remove bimera method since it was developed in conjunction with \ac{dada2} and its performance does not significantly differ from uchime-denovo.
-  For the current work, we have decided to focus on the tools most widely used at the time of the analysis.
-  Some tools which were not as widely used (e.g.
-dbOTU3~\cite{Olesen2017}) as well as older popular methods like mothur~\cite{Schloss2009} have not been included in the study but could be added into the pipelines in future updated analyses.
-
-  The choice of taxonomy database was found to be the most important factor in the inference of microbial co-occurrence networks, contributing $65.4\%$ of the total variance.
-  %\todo[noline]{R2.2a}
-  The frequent changes in the taxonomy nomenclature coupled with the frequency of updates to the various 16S reference databases create inherent differences \cite{Balvociute2017} in taxonomy hierarchies in these databases.
-  Our analysis revealed that no particular reference database performs better than the others across the different mock dataset benchmarks.
-  %\todo[noline]{R2.12}
-  The default reference database in the pipeline is currently chosen to be the \ac{gg} reference database along with the ``Naive Bayes'' classifier as the query tool.
-  The reason for our choice stems from the popularity of the \ac{gg} database~\cite{parkEvaluation16SRRNA2018} in taxonomic studies, which would enable easy comparison across datasets.
-  However, we recommend using the SILVA database for newer studies due to its size and taxonomic comprehensiveness~\cite{iiRESCRIPtReproducibleSequence2021} and since \ac{gg} has not been updated since 2013.
-  Additionally, a particular database might be more appropriate than the rest based on specific requirements.
-  For example, to generate a dataset that is compatible with the \ac{mind} platform~\cite{huResourceComparisonIntegration2022}, \ac{ncbi}, is the most appropriate choice as it guarantees compatibility of taxonomic hierarchy and therefore comparability with other datasets.
-  A detailed study of other taxonomy mapping approaches[REFS IDTAXA] and integrative approaches combining different databases might offer orthogonal information and better matches.
-  Furthermore, we also enable users to use custom databases~\cite{Ritari2015,iiRESCRIPtReproducibleSequence2021} with the BLAST and Naive Bayes classifiers that are incorporated into the pipeline (from \ac{qiime2}).
-  We suggest that the choice of the database should be made based on possible reported or inferred biases in the representation of given biomes in a specific databases~\cite{Balvociute2017,iiRESCRIPtReproducibleSequence2021}, as choosing taxon-specific databases have also been observed to compromise classification~\cite{rmarcelinoUseTaxonspecificReference2020}.
-
-  The \ac{op} step of the pipeline is second in its contribution to total network variance.
-  This can be attributed to the large number of nodes that are added to the final networks when the filtering is turned off.
-  Additionally, a very large number of nodes also decreases the accuracy of the network inference algorithms for the same sample size~\cite{peschelNetCoMiNetworkConstruction2020} and increases the computational complexity~\cite{tackmannRapidInferenceDirect2019}.
-  We observe that filtering out taxa that are present in low abundances in all samples increases the proportion of taxa in common between taxonomy tables generated using different reference databases (Figure~\ref{fig:figure_s4}), providing another reason for filtering.
-  We also observe that the reduction in the number of taxa leads to a better agreement in the networks inferred through different methods (Figure~\ref{fig:figure6} and Figure~\ref{fig:figures1}).
-  Moreover, filtering is necessary in order to increase the power in tests of significance when the number of taxa is much greater than the number of samples.
-
-\newpage
-  \pagenumbering{goggle}
-  \thispagestyle{empty}
-  \input{sections/table1.tex}
-  \thispagestyle{empty}
-  \input{sections/table_s1.tex}
-
 %--------------------------------------------------------%

### 2023-03-10 final_submission/manuscript_full.tex
Add: Files for final submission

--- 2023-03-08 final_submission/manuscript.tex
@@
   In Table~\ref{tab:network_metrics} we show various global network metrics calculated for each tool in the pipeline.
-  %\todo[noline]{R2.1a}
-  %\todo[noline]{R2.7}
+  %%\todo[noline]{R2.1a}
+  %%\todo[noline]{R2.7}
   All the networks that make use of a particular tool are grouped together, and the following average metrics are calculated for each group:
@@
   Fisher~\cite{fisher_224a_1948} proposed that for $k$ independent p-values, each generated by $k$ different methods and denoted by $\bar{P}^i$ (notations are same as used in the "Consensus network and p-value merging" subsection of the Methods).
-  %\todo[noline]{R1.1}
+  %%\todo[noline]{R1.1}
   The following will hold true for the statistic $\Psi$:
@@
 
+  \subsection*{Network variance analysis of stool samples from radiation-exposed bank vole (EMP dataset)}
+
+  In order to verify the consistency of the network variance analysis, we processed an additional 16S rRNA dataset through the \ac{micone} pipeline.
+  %\todo[noline]{R3.1}
+  Specifically, we used stool samples from radiation-exposed bank vole that were a part of the Earth Microbiome Project (EMP)~\cite{shafferStandardizedMultiomicsEarth2022}.
+  We chose a dataset from the EMP because the \ac{micone} pipeline inherently supports the EMP amplicon protocol with minimal pre-processing requirements.
+  The data containing the 16S sequencing reads (V4 region) was downloaded from Qiita~\cite{gonzalezQiitaRapidWebenabled2018} (study ID: 13114).
+  For the analysis, the paired-end sequencing data extracted from stools samples of radiation-exposed bank vole (named ``mousseau'') were chosen from run ``EMP500\_6-9''.
+  The data was then processed using the \ac{micone} pipeline starting at the \ac{sp} step and ending at the \ac{ni} step with the consensus algorithm.
+  The configuration files (main.nf and nextflow.config) used to run the \ac{micone} pipeline as well the details of the pipeline execution (dag, report, timeline and trace) are in the "runs/EMP" directory of the data and scripts repository (\href{https://github.com/segrelab/MiCoNE-pipeline-paper}{https://github.com/segrelab/MiCoNE-pipeline-paper})
+
+  Similar to the previous network variance analysis performed in Figure~\ref{fig:figure6}, we analyzed the effect of different processing methods on the inferred co-occurrence networks generated using all possible combinations of methods.
+  Figure \ref{fig:figure_s8}A, shows the percentage of total variation among the co-occurrence networks due to the different steps of the pipeline.
+  The \ac{ta} step, or more specifically the choice of 16S reference database, contributes the most ($57.96\%$) to the variation in the networks, followed by the \ac{op} step ($33.48\%$).
+  These results are very similar to those from the network variance analysis of the FMT dataset (Figure~\ref{fig:figure6}A).
+  This implies that the effects of various workflow steps on the inferred networks are fairly consistent across different datasets.
+
+  The effects of the different steps of the pipeline on the inferred networks can be visualized through dimensionality reduction.
+  The PCA in Figure \ref{fig:figure_s8}B shows all possible inferred networks, colored by the tools used in the DC, TA, OP, and NI steps in each subfigure.
+  We observe that the networks largely segregate based on the database used (TA step), and the extent of this separation decreases when the filtering is turned on in the OP step.
+  However, unlike the case of the FMT dataset (Figure~\ref{fig:figure6}B), here we observe that the networks that utilize \ac{gg} and \ac{ncbi} databases have the variation along similar axes in the 2 dimensional PCA plot.
+  This could imply that the taxonomy assignments returned by these two databases for this dataset, might be more similar to each other compared to those in the FMT dataset.
+  The variation in the networks due to the other workflow steps is consistent with that observed in the FMT data analysis (Figure~\ref{fig:figure6}B).
+
   \subsection*{Supplementary discussion}
@@
   The choice of taxonomy database was found to be the most important factor in the inference of microbial co-occurrence networks, contributing $65.4\%$ of the total variance.
-  %\todo[noline]{R2.2a}
+  %%\todo[noline]{R2.2a}
   The frequent changes in the taxonomy nomenclature coupled with the frequency of updates to the various 16S reference databases create inherent differences \cite{Balvociute2017} in taxonomy hierarchies in these databases.
   Our analysis revealed that no particular reference database performs better than the others across the different mock dataset benchmarks.
-  %\todo[noline]{R2.12}
+  %%\todo[noline]{R2.12}
   The default reference database in the pipeline is currently chosen to be the \ac{gg} reference database along with the ``Naive Bayes'' classifier as the query tool.
   The reason for our choice stems from the popularity of the \ac{gg} database~\cite{parkEvaluation16SRRNA2018} in taxonomic studies, which would enable easy comparison across datasets.
+  %\todo[noline]{R2.6}
   However, we recommend using the SILVA database for newer studies due to its size and taxonomic comprehensiveness~\cite{iiRESCRIPtReproducibleSequence2021} and since \ac{gg} has not been updated since 2013.
@@
   For example, to generate a dataset that is compatible with the \ac{mind} platform~\cite{huResourceComparisonIntegration2022}, \ac{ncbi}, is the most appropriate choice as it guarantees compatibility of taxonomic hierarchy and therefore comparability with other datasets.
-  A detailed study of other taxonomy mapping approaches[REFS IDTAXA] and integrative approaches combining different databases might offer orthogonal information and better matches.
+  A detailed study of other taxonomy mapping approaches~\cite{muraliIDTAXANovelApproach2018,matiasrodriguesMAPseqHighlyEfficient2017} and integrative approaches combining different databases~\cite{gwakDataDrivenModelingSpeciesLevel2020} might offer orthogonal information and better matches.
+  %\todo[noline]{R2.5a}
   Furthermore, we also enable users to use custom databases~\cite{Ritari2015,iiRESCRIPtReproducibleSequence2021} with the BLAST and Naive Bayes classifiers that are incorporated into the pipeline (from \ac{qiime2}).
@@
   We observe that filtering out taxa that are present in low abundances in all samples increases the proportion of taxa in common between taxonomy tables generated using different reference databases (Figure~\ref{fig:figure_s4}), providing another reason for filtering.
-  We also observe that the reduction in the number of taxa leads to a better agreement in the networks inferred through different methods (Figure~\ref{fig:figure6} and Figure~\ref{fig:figures1}).
+  We also observe that the reduction in the number of taxa leads to a better agreement in the networks inferred through different methods (Figure~\ref{fig:figure6} and Figure~\ref{fig:figure_s1}).
+  %\todo[noline]{R2.3}
   Moreover, filtering is necessary in order to increase the power in tests of significance when the number of taxa is much greater than the number of samples.
 
-\newpage
-  \pagenumbering{goggle}
-  \thispagestyle{empty}
-  \input{sections/table1.tex}
-  \thispagestyle{empty}
-  \input{sections/table_s1.tex}
+% \newpage
+%   \pagenumbering{goggle}
+%   \thispagestyle{empty}
+%   \input{table1.tex}
+%   \thispagestyle{empty}
+%   \input{table_s1.tex}
 
